### 2020-06-25 paper.Rmd
added first files

---
title             : "An improved k-means criterion for optimizing between-group similarity"
shorttitle        : "K-means anticlustering extended"

author:
  - name          : "Martin Papenberg"
    affiliation   : "1"
    corresponding : yes
    address       : "Heinrich Heine Universität Düsseldorf,  Institut für Experimentelle Psychologie, Universitätsstraße 1, 40225 Düsseldorf, Germany"
    email         : "martin.papenberg@hhu.de"

affiliation:
  - id            : "1"
    institution   : "Heinrich Heine University Düsseldorf"

authornote: |

  Martin Papenberg, Department of Experimental Psychology, Heinrich 
  Heine University Düsseldorf.

abstract: |
  This is the abstract.

keywords          : "K-Means, Anticlustering, Variance"
wordcount         : "X"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

This is my new paper.

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2020-06-25 improved_objective.R

@@ -1,0 +1,104 @@
+
+library(anticlust)
+
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
+
+improved_exchange_ <- function(nrep, x, K, objective, obj_function, 
+                               categories, preclustering) {
+  x <- as.matrix(x)
+  N <- nrow(x)
+  clusters <- anticlustering(
+    x, K, objective = objective, categories = categories
+  )
+  new_obj <- obj_function(x, clusters)
+  old_obj <- -1
+  while (new_obj > old_obj) {
+    clusters <- anticlustering(
+      x, K, objective = objective
+    )
+    old_obj <- new_obj
+    new_obj <- obj_function(x, clusters)
+  }
+  clusters
+}
+
+improved_exchange <- function(x, K, objective, obj_function, 
+                              categories = NULL, preclustering = FALSE, 
+                              nrep = 10) {
+  candidate_solutions <- lapply(
+    1:nrep, 
+    improved_exchange_,
+    x = x, 
+    K = K, 
+    objective = objective, 
+    obj_function = obj_function,
+    categories = categories, 
+    preclustering = preclustering
+  )
+  obj_function_ <- function(clusters, x) {
+    obj_function(x, clusters)
+  }
+  objs <- lapply(
+    candidate_solutions,
+    obj_function_,
+    x = x
+  )
+  candidate_solutions[[which.max(objs)]]
+}
+
+
+N <- 200
+M <- 2
+K <- 3
+features <- matrix(rnorm(N * M), ncol = M)
+features <- schaper2019[, 3:6]
+
+nrep <- 50
+preclustering <- TRUE
+
+ac_kmeans <- improved_exchange(
+  scale(cbind(features)),
+  K = K,
+  objective = "variance",
+  obj_function = variance_objective, 
+  preclustering = preclustering,
+  nrep = nrep
+)
+
+ac_distance <- improved_exchange(
+  scale(cbind(features)),
+  K = K,
+  objective = "distance",
+  obj_function = variance_objective, 
+  preclustering = preclustering,
+  nrep = nrep
+)
+
+ac_kmeans_var_only <- improved_exchange(
+  squared_from_mean(features),
+  K = K,
+  objective = "variance",
+  obj_function = variance_objective,
+  preclustering = preclustering,
+  nrep = nrep
+)
+
[…]

### 2020-06-25 testing.R

@@ -1,0 +1,94 @@
+
+options(scipen = 999)
+
+library(anticlust)
+
+# Determine the maximum range in a vector
+range_diff <- function(x) {
+  diff(range(x))
+}
+
+compare_objectives <- function(i, N = 100, K = 2, preclustering = FALSE) {
+  cat("Simulation run", i, "\n")
+  data <- rnorm(N)
+  data_var <- abs(data - mean(data))
+  ac <- anticlustering(
+    data, 
+    K = K,
+    objective = "variance"
+  )
+  var_only <- anticlustering(
+    data_var, 
+    K = K,
+    objective = "variance"
+  )
+  ac_var <- anticlustering(
+    cbind(data, data_var), 
+    K = K,
+    objective = "variance"
+  )
+  ac_var_pre <- anticlustering(
+    cbind(data, data_var), 
+    K = K,
+    objective = "variance",
+    preclustering = TRUE
+  )
+  ac_dist <- anticlustering(
+    data, 
+    K = K,
+    objective = "distance",
+  )
+  ac_dist_pre <- anticlustering(
+    data, 
+    K = K,
+    objective = "distance",
+    preclustering = TRUE
+  )
+  
+  random <- sample(ac_dist)
+  
+  list(
+    DIFF_MEANS = c(
+      kmeans = range_diff(tapply(data, ac, mean)),
+      kmeansVAR = range_diff(tapply(data, ac_var, mean)),
+      kmeansVAR_PRE = range_diff(tapply(data, ac_var_pre, mean)),
+      kmeansVARONLY = range_diff(tapply(data, var_only, mean)),
+      MDGP = range_diff(tapply(data, ac_dist, mean)),
+      MDGP_PRE = range_diff(tapply(data, ac_dist_pre, mean)),
+      RANDOM = range_diff(tapply(data, random, mean))
+    ),
+    DIFF_VARS = c(
+      kmeans = range_diff(tapply(data, ac, var)),
+      kmeansVAR = range_diff(tapply(data, ac_var, var)),
+      kmeansVAR_PRE = range_diff(tapply(data, ac_var_pre, var)),
+      kmeansVARONLY = range_diff(tapply(data, var_only, var)),
+      MDGP = range_diff(tapply(data, ac_dist, var)),
+      MDGP_PRE = range_diff(tapply(data, ac_dist_pre, var)),
+      RANDOM = range_diff(tapply(data, random, var))
+    )
+  )
+}
+
[…]

### 2020-06-25 improved_objective.R
only scale data where features are on different scales

@@ -49,17 +49,17 @@ improved_exchange <- function(x, K, objective, obj_function,
 }
 
 
-N <- 200
+N <- 500
 M <- 2
-K <- 3
-features <- matrix(rnorm(N * M), ncol = M)
-features <- schaper2019[, 3:6]
+K <- 10 
+features <- matrix(rchisq(N * M, 2), ncol = M)
+#features <- scale(schaper2019[, 3:6])
 
-nrep <- 50
-preclustering <- TRUE
+nrep <- 20
+preclustering <- FALSE
 
 ac_kmeans <- improved_exchange(
-  scale(cbind(features)),
+  features,
   K = K,
   objective = "variance",
   obj_function = variance_objective, 
@@ -68,7 +68,7 @@ ac_kmeans <- improved_exchange(
 )
 
 ac_distance <- improved_exchange(
-  scale(cbind(features)),
+  features,
   K = K,
   objective = "distance",
   obj_function = variance_objective, 
@@ -86,7 +86,7 @@ ac_kmeans_var_only <- improved_exchange(
 )
 
 ac_kmeans_var <- improved_exchange(
-  scale(cbind(features, squared_from_mean(features))),
+  cbind(features, squared_from_mean(features)),
   K = K,
   objective = "variance",
   obj_function = variance_objective,
@@ -100,5 +100,8 @@ cat("K-Means + Squared Diff\n")
 print(mean_sd_tab(features, ac_kmeans_var, return_diff = TRUE))
 cat("Anticluster Editing \n")
 print(mean_sd_tab(features, ac_distance, return_diff = TRUE))
-cat("Only Squared Diff\n")
-print(mean_sd_tab(features, ac_kmeans_var_only, return_diff = TRUE))
+#cat("Only Squared Diff\n")
+#print(mean_sd_tab(features, ac_kmeans_var_only, return_diff = TRUE))
+
+
+

### 2020-06-25 paper.Rmd
added some initial text /code to paper

@@
 library("papaja")
+options(digits = 2)
 ```
 
-This is my new paper.
+- It is often important to divide a set of elements into similar 
+subgroups
+- This task can be automized and open source solutions exist 
+- Anticlustering = opposite of clustering, does that
+- Objectives based on pairwise distances exist, but k-means also exists 
+- k-means = optimizes similarity wrt to mean, but not other distribution
+parameters, in particular not wrt to the variance
+- here, I present an improved objective function for k-means 
+anticlustering, which optimizes similarity wrt to means and variance / 
+standard deviation at the same time
+
+## A Motivating Example
+
+```{r}
+
+# set.seed(26178) # uncomment later
+
+library(anticlust)
+
+N <- 14
+K <- 2
+partitions <- generate_partitions(N, K)
+length(partitions) # number of possible partitions
+features <- rnorm(N)
+
+## Create an objective function that takes the partition
+## as first argument (then, we can use sapply to compute
+## the objective for each partition)
+diff_mean <- function(clusters, features) {
+  abs(diff(tapply(features, clusters, mean)))
+}
+
+diff_vars <- function(clusters, features) {
+  abs(diff(tapply(features, clusters, var)))
+}
+
+all_objectives_mean <- sapply(
+  partitions,
+  FUN = diff_mean,
+  features = features
+)
+
+all_objectives_var <- sapply(
+  partitions,
+  FUN = diff_vars,
+  features = features
+)
+
+df <- data.frame(
+  diff_mean = all_objectives_mean,
+  diff_var  = all_objectives_var
+)
+
+plot(df, pch = 4, col = "darkgrey", las = 1)
+# Illustrate best partition wrt simililarity in means
+points(
+  all_objectives_mean[which.min(all_objectives_mean)],
+  all_objectives_var[which.min(all_objectives_mean)],
+  cex = 2, col = "#DF536B", pch = 16
+)
+
+df$sum <- rowSums(df)
+
+points(
+  all_objectives_mean[which.min(df$sum)],
+  all_objectives_var[which.min(df$sum)],
+  cex = 2, col = "#2297E6", pch = 16
+)
+
+points(
+  all_objectives_mean[which.min(df$diff_var)],
+  all_objectives_var[which.min(df$diff_var)],
+  cex = 2, col = "#61D04F", pch = 16
+)
+
+```
+
+The plot illustrates all possible ways to partition a data set of $N = 
+14$ into two parts. Red point: maximum similarity in means; green point 
+= maximum similarity in variance. Blue point: Almost no difference to 
+red point wrt similarity in means, but variance is much more similar. 
+This blue partition is certainly preferable to the red one when maximal 
+overall similarity is required; the red one just optimizes similarity in 
+means (which is what maximizing the k-means objective does).
+
+Unfortunately, improving algorithmic procedures will not help to find 
+better partitions -- the objective function has to change. I will show 
+an objective function that captures both similarity in means and 
+variance.
+
+## K-Means criterion
+
+The popular k-means clustering method aims at minimizing the 
+within-group variance, i.e., the sum of the squared Euclidean distances 
+between each data point and its cluster center:
+
+More specifically:
+
+$$
+\mathit{Var}_{\mathrm{within}} = \sum\limits_{j=1}^{K} 
+\sum\limits_{x_i \in c_j} \vert \vert x_i - \mu_j \vert \vert^2
+$$
+
+The squared distance between an element $x$ and the average of all 
+elements $\overline{\mathrm{\boldsymbol{x}}}$ is:
+
+$$
+\vert \vert x - \overline{\mathrm{\boldsymbol{x}}} \vert \vert^2 = 
+  \sum\limits_{m = 1}^{M}(x_{m} - \overline{\mathrm{\boldsymbol{x_m}}})^2
+$$
+
+Now define new data points representing the squared distances between 
+data points and the grand mean:
+
+$$ 
+y_i = (x_i - \mathrm{\boldsymbol{x}})^2, \forall i \in N
+$$
+
+where $\mathrm{\boldsymbol{x}}$ is a vector of length $M$ containing 
+the average for each feature.
+
+Now:
+
+$$
+\sum\limits_{m = 1}^{M}(x_{m} - \overline{\mathrm{\boldsymbol{x_m}}})^2 + 
+  \sum\limits_{m = 1}^{M}(y_{m} - \overline{\mathrm{\boldsymbol{y_m}}})^2 =
+  \sum\limits_{m = 1}^{M}(x_{m} - \overline{\mathrm{\boldsymbol{x_m}}})^2 + 
+    (y_{m} - \overline{\mathrm{\boldsymbol{y_m}}})^2
+$$
+
+If we define $x_{m + 1} = y_1$, $x_{m + 2} = y_2$, ..., $x_{2M} = y_M$.
+this is the same as:
+
+$$
+\sum\limits_{m = 1}^{2M}(x_{m} - \overline{\mathrm{\boldsymbol{x_m}}})^2 
+$$
+
+Thus, we redefine the data points and include an additional $M$ features 
+for each element. Then, we maximize $\mathit{Var}_{\mathrm{within}}$, 
+which corresponds to a tradeoff between minimizing differences in mean 
+features between clusters, and the average squared distance to the grand 
+mean, i.e., the within-cluster variance. That is: We try to maximize the 
+within-cluster variance, but to a certain degree we also strive to 
+obtain similar within-cluster variance in each cluster.
+
+Important to note: We solve the same optimization problem, we just 
+change the input a bit (we double the number of "columns").
+
+**Example table!!**
+
+# Stuff
+
+New objective seems to work best when N is high relative to K; 
+optimizing two objectives at the same time is hard when N is low. Easier 
+when only optimizing one objective (i.e., the mean, as done by normal
+k-means anticlustering).
+
+# Test equivalence of objectives
+
+Check that I can actually just append the squared difference data as 
+additional columns to the original data frame:
+
+```{r}
+
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
+
+N <- 500
+M <- 20
+data <- matrix(rnorm(N * M), ncol = M)
+var_data <- squared_from_mean(data)
+
+K <- 10
+clusters <- sample(rep_len(1:K, N))
+
+obj1 <- variance_objective(data, clusters) + variance_objective(var_data, clusters)
+obj2 <- variance_objective(cbind(data, var_data), clusters)
+
+tolerance <- 1e-10
+
+obj1 - obj2 < tolerance
+
+obj1
+obj2
+
+```
 

### 2020-07-06 paper.Rmd
work on k-means formalization

@@
 partitions <- generate_partitions(N, K)
-length(partitions) # number of possible partitions
 features <- rnorm(N)
@@
 The plot illustrates all possible ways to partition a data set of $N = 
-14$ into two parts. Red point: maximum similarity in means; green point 
+14$ into two parts (there are `r length(partitions)` possible 
+partitions). Red point: maximum similarity in means; green point 
 = maximum similarity in variance. Blue point: Almost no difference to 
@@
 
-The popular k-means clustering method aims at minimizing the 
-within-group variance, i.e., the sum of the squared Euclidean distances 
-between each data point and its cluster center:
+In my treatment of k-means clustering, I adopt the notation Steinley 
+(2006) provided in his thorough synthesis of the k-means literature. In 
+clustering---or anticlustering---, an item pool $X$ = $\{x_1, \ldots, 
+x_N\}$ has to be partitioned into $K$ groups $\{C_1, \ldots, C_K\}$. We 
+will refer to the groups $C_j$, $j = 1, \ldots, K$, as anticlusters, 
+clusters, or groups, and to the collection of anticlusters $C$ as the 
+partitioning of the data set. We assume that each element $x_i$ ($i = 1, 
+\ldots, N$), is a vector of length $P$, and each entry describes one of 
+its numeric measurements. Generally hence, $X$ can be interpreted as an 
+$N \times M$ matrix where each row represents an data point $x_i$ (e.g., 
+representing a stimulus or a person) and each column is a numeric 
+attribute (e.g., the number of syllables in a word, or the 
+attractiveness rating of a face).
 
-More specifically:
+K-Means anticlustering aims to maximize the error sum of squares (SSE), 
+also known as the within-group variance, which is the sum of the 
+squared Euclidean distances between each data point and its cluster 
+centroid. The cluster centroid is given by averaging the values on each 
+variable over the elements within that cluster. The centroid of the 
+*j*th variable in the *k*th cluster, consisting of $n_k$ elements, is 
+given as:
 
 $$
-\mathit{Var}_{\mathrm{within}} = \sum\limits_{j=1}^{K} 
-\sum\limits_{x_i \in c_j} \vert \vert x_i - \mu_j \vert \vert^2
+\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits {i \in C_k} x_{ij}
 $$
 
-The squared distance between an element $x$ and the average of all 
-elements $\overline{\mathrm{\boldsymbol{x}}}$ is:
+The k-means objective function, the error sum of squares, is then given 
+as:
+
+$$
+\mathit{SSE} = 
+  \sum\limits_{j=1}^{P} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (x_{ij} - \overline{x}_j ^{(k)})^2
+$$
+
+---
+
+**This is old, adjust to Steinley (2006):**
+
+The squared distance between an element $x = (x_1, \ldots, x_M)$ and the 
+average of all elements $\overline{\mathrm{\boldsymbol{x}}}$ is:
 

### 2020-07-06 improved_objective.R
implemented improved exchange method into anticlust

@@ -5,93 +5,45 @@ squared_from_mean <- function(data) {
   apply(data, 2, function(x) (x - mean(x))^2)
 }
 
-improved_exchange_ <- function(nrep, x, K, objective, obj_function, 
-                               categories, preclustering) {
-  x <- as.matrix(x)
-  N <- nrow(x)
-  clusters <- anticlustering(
-    x, K, objective = objective, categories = categories
-  )
-  new_obj <- obj_function(x, clusters)
-  old_obj <- -1
-  while (new_obj > old_obj) {
-    clusters <- anticlustering(
-      x, K, objective = objective
-    )
-    old_obj <- new_obj
-    new_obj <- obj_function(x, clusters)
-  }
-  clusters
-}
 
-improved_exchange <- function(x, K, objective, obj_function, 
-                              categories = NULL, preclustering = FALSE, 
-                              nrep = 10) {
-  candidate_solutions <- lapply(
-    1:nrep, 
-    improved_exchange_,
-    x = x, 
-    K = K, 
-    objective = objective, 
-    obj_function = obj_function,
-    categories = categories, 
-    preclustering = preclustering
-  )
-  obj_function_ <- function(clusters, x) {
-    obj_function(x, clusters)
-  }
-  objs <- lapply(
-    candidate_solutions,
-    obj_function_,
-    x = x
-  )
-  candidate_solutions[[which.max(objs)]]
-}
+N <- 60
+M <- 1
+K <- 3
+features <- matrix(rnorm(N * M), ncol = M)
+#features <- schaper2019[, 3:6]
 
+nrep <- 30
 
-N <- 500
-M <- 2
-K <- 10 
-features <- matrix(rchisq(N * M, 2), ncol = M)
-#features <- scale(schaper2019[, 3:6])
-
-nrep <- 20
-preclustering <- FALSE
-
-ac_kmeans <- improved_exchange(
+ac_kmeans <- anticlustering(
   features,
   K = K,
   objective = "variance",
-  obj_function = variance_objective, 
-  preclustering = preclustering,
-  nrep = nrep
+  method = "local-maximum",
+  repetitions = nrep
 )
 
-ac_distance <- improved_exchange(
+ac_distance <- anticlustering(
   features,
   K = K,
   objective = "distance",
-  obj_function = variance_objective, 
-  preclustering = preclustering,
-  nrep = nrep
+  method = "local-maximum",
[…]

### 2020-07-06 example-oasis.R
added example using oasis data

@@ -1,0 +1,24 @@
+
+library(anticlust)
+library(ggplot2)
+
+oasis <- read.csv("https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv")
+
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
+
+features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
+extended <- cbind(features, squared_from_mean(features))
+
+oasis$ac <- anticlustering(
+  scale(extended),
+  K = 10,
+  categories = oasis$Category, 
+  objective = "variance",
+  method = "local-maximum"
+)
+
+mean_sd_tab(features, oasis$ac)
+
+table(ac, oasis$Category)

### 2020-07-06 paper.Rmd
minor stuff paper

@@
 
-In my treatment of k-means clustering, I adopt the notation Steinley 
-(2006) provided in his thorough synthesis of the k-means literature. In 
-clustering---or anticlustering---, an item pool $X$ = $\{x_1, \ldots, 
-x_N\}$ has to be partitioned into $K$ groups $\{C_1, \ldots, C_K\}$. We 
-will refer to the groups $C_j$, $j = 1, \ldots, K$, as anticlusters, 
-clusters, or groups, and to the collection of anticlusters $C$ as the 
-partitioning of the data set. We assume that each element $x_i$ ($i = 1, 
-\ldots, N$), is a vector of length $P$, and each entry describes one of 
-its numeric measurements. Generally hence, $X$ can be interpreted as an 
-$N \times M$ matrix where each row represents an data point $x_i$ (e.g., 
-representing a stimulus or a person) and each column is a numeric 
-attribute (e.g., the number of syllables in a word, or the 
-attractiveness rating of a face).
+In this treatment of k-means, I adopt the notation Steinley (2006) 
+provided in his synthesis of the k-means literature. In clustering---or 
+anticlustering---, an item pool $X$ = $\{x_1, \ldots, x_N\}$ has to be 
+partitioned into $K$ groups $\{C_1, \ldots, C_K\}$. We assume that each 
+element $x_i$ ($i = 1, \ldots, N$), is a vector of length $P$, and each 
+entry describes one of its numeric measurements. Generally, $X$ 
+can is interpreted as an $N \times M$ matrix where each row represents 
+an data point $x_i$ (e.g., representing a stimulus or a person) and each 
+column is a numeric attribute (e.g., the number of syllables in a word, 
+or the attractiveness rating of a face).
 
 K-Means anticlustering aims to maximize the error sum of squares (SSE), 
-also known as the within-group variance, which is the sum of the 
-squared Euclidean distances between each data point and its cluster 
-centroid. The cluster centroid is given by averaging the values on each 
-variable over the elements within that cluster. The centroid of the 
-*j*th variable in the *k*th cluster, consisting of $n_k$ elements, is 
-given as:
+the within-group variance, which is the sum of the squared Euclidean 
+distances between each data point and its cluster centroid. The cluster 
+centroid is given by averaging the values on each variable over the 
+elements within that cluster. The centroid of the *j*th variable in the 
+*k*th cluster, consisting of $n_k$ elements, is given as:
 
@@
   (x_{ij} - \overline{x}_j ^{(k)})^2
+$$
+
+In anticlustering applications, we usually assume that all cluster are 
+of equal size. Hence
+
+$$
+n_k = \frac{N}{K}, \forall k \in \{1, \ldots, K\}
 $$
@@
 
+- Two examples: $K = 2$, $N = 20$ vs. $N = 100$, using classical
+k-means: only mean; using k-variance: only variance; combination:
+both, but works better for larger $N$
+
+# Weighted k-means
+
+Using variable weighting (see Steinley, 2006, p. 13), we can 
+adjust the relative importance of minimizing differences in means
+and variances. Standard in `anticlust`: Equal weight, this is 
+achieved by first creating the new variables, representing 
+the squared differences from the mean, and then standardizing
+the entire feature matrix. If $N$ is large enough (relative to $K$),
+there is usually no need for weighting.
+
+# Simulation
+
+- show that k-variance anticlustering outperforms other objectives
+
+# Examples
+
+- use some examples (e.g., OASIS data set)
+
 # Test equivalence of objectives

### 2020-07-07 solution-structure.R
illustrate exchange algorithm for different objectives

@@ -1,0 +1,110 @@
+
+# Investigate why k-variance is worse than k-means-variance for 
+# optimizing between-group similarity in variance
+
+library(anticlust)
+
+N <- 16
+K <- 2
+all_partitions <- generate_partitions(N, K)
+features <- as.matrix(rnorm(N))
+
+## Create an objective function that takes the partition
+## as first argument (then, we can use sapply to compute
+## the objective for each partition)
+diff_mean <- function(clusters, features) {
+  abs(diff(range(tapply(features, clusters, mean))))
+}
+
+diff_vars <- function(clusters, features) {
+  abs(diff(range(tapply(features, clusters, var))))
+}
+
+all_objectives_mean <- sapply(
+  all_partitions,
+  FUN = diff_mean,
+  features = features
+)
+
+all_objectives_var <- sapply(
+  all_partitions,
+  FUN = diff_vars,
+  features = features
+)
+
+df <- data.frame(
+  diff_mean = all_objectives_mean,
+  diff_var  = all_objectives_var
+)
+
+# exchange method for finding all objectives
+
+local_maximum_exchange <- function(features, init) {
+  best_partition <- init
+  found_partitions <- list(best_partition)
+  new_obj <- variance_objective(features, best_partition)
+  old_obj <- -Inf
+  while (new_obj > old_obj) {
+    partitions <- anticlustering(
+      features,
+      K = best_partition, 
+      objective = variance_objective
+    )
+    best_partition <- partitions[[length(partitions)]]
+    old_obj <- new_obj
+    new_obj <- variance_objective(features, best_partition)
+    found_partitions <- c(found_partitions, partitions)
+  }
+  found_partitions[!duplicated(found_partitions)]
+} 
+
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
+
+# Initialize a partition
+init <- sample(rep_len(1:K, N))
+init <- anticlust:::order_cluster_vector(init)
+
+# Do local search for three objectives
+partitions_means <- local_maximum_exchange(
+  features, 
+  init
+)
+partitions_variance <- local_maximum_exchange(
+  squared_from_mean(features), 
+  init
+)
+partitions_means_variance <- local_maximum_exchange(
+  cbind(features, squared_from_mean(features)), 
[…]

### 2020-07-07 paper.Rmd
misc minor

@@
 
+bibliography      : ["lit.bib"]
+
 keywords          : "K-Means, Anticlustering, Variance"
@@
 
-- It is often important to divide a set of elements into similar 
-subgroups
-- This task can be automized and open source solutions exist 
-- Anticlustering = opposite of clustering, does that
-- Objectives based on pairwise distances exist, but k-means also exists 
-- k-means = optimizes similarity wrt to mean, but not other distribution
-parameters, in particular not wrt to the variance
+Oftentimes researchers have to divide a set of elements different, but 
+equivalent groups. For example, this problem arises when 
+splitting a test into parts of equal difficulty [@gierl2017], when 
+assigning students to work groups [@baker2002], or when assigning 
+stimuli to experimental conditions [@lahl2006]. Solving such problems 
+manually is a tedious and time-consuming task and the similarity of 
+manually established groups is usually improvable. Fortunately, these 
+problems can be formalized mathematically [e.g., @brusco2019; 
+@spath1986] and by now, accessible open source solutions for tackling 
+this automatically exist [@papenberg2020].
+
+Forming similar groups is also known as the *anticlustering* problem. 
+Anticlustering reverses the logic of its better known twin, cluster 
+analysis [@rokach2005]. In cluster analysis, groups are sought that are 
+internally homogenous, whereas there is heterogeneity between groups 
+[@steinley2006]. Anticlustering instead strives for high between-group 
+similarity high within-group heterogeneity. Usually, cluster analysis 
+aims to uncover "true" latent structure contained in a data set, whereas 
+anticlustering imposes an "unnatural" division of the data.
+
+One well-known criterion for anticlustering reverses the k-means 
+clustering method by maximizing the within-group variation, or 
+*variance* [@jain2010; @steinley2006; @spath1986; @valev1998; 
+@valev1983]. Späth and Valev independently coined the term 
+anticlustering specifically to denote the reversal of classical 
+k-means.
+
+--- k-means
+
+--- dissimilarity data vs. rectangular data
+
+Although forming groups with high between-group similarity is often 
+equivalent to establishing high within-group diversity, researchers 
+usually only wish to obtain one of these goals. In experimental 
+psychology, sets of stimuli should be as similar as possible; within-set 
+heterogeneity is only an unimportant side-product. In this paper, I 
+address such problems that require high between-group similarity. I 
+present a novel extension to the k-means objective function, 
+within-group sum-of-squares that improves its capacity to form similar 
+groups [@steinley2006], but is no longer aims to maximize heterogeneity.
+
+Ironically, even though maximizing the k-means objective maximizes 
+similarity with regard to the means of the
+
 - here, I present an improved objective function for k-means 
@@
 
-plot(df, pch = 4, col = "darkgrey", las = 1)
+plot(df, pch = 4, col = rainbow(length(partitions)), las = 1)
 # Illustrate best partition wrt simililarity in means
@@
 
+# Combined approach better than k-variance alone
+
+- Often, maximizing k-means-variance yields more similar variances 
+than k-variance alone. Why is that? I guess: structure in the solution 
+space, local maxima are real. When also considering means, you more 
+often escape bad local maxima?
+
 # Test equivalence of objectives

### 2020-07-07 example-oasis.R

@@ -16,9 +16,10 @@ oasis$ac <- anticlustering(
   K = 10,
   categories = oasis$Category, 
   objective = "variance",
-  method = "local-maximum"
+  method = "local-maximum",
+  preclustering = FALSE
 )
 
 mean_sd_tab(features, oasis$ac)
 
-table(ac, oasis$Category)
+table(oasis$ac, oasis$Category)

### 2020-07-08 paper.Rmd
some work on intro etc

--- 2020-07-06 paper.Rmd
@@
 ---
-title             : "An improved k-means criterion for optimizing between-group similarity"
+title             : "An extension of k-means anticlustering to optimize between-group similarity"
 shorttitle        : "K-means anticlustering extended"
@@
 abstract: |
-  This is the abstract.
+  Anticlustering is an automated method to divide a data set into groups
+  in such a way that high between-group similarity and high within-group 
+  heterogeneity is obtained. Anticlustering is accomplished by maximizing 
+  instead of minimizing a clustering objective.
+
+bibliography      : ["lit.bib"]
 
@@
 library("papaja")
+library("anticlust")
+library("knitr")
 options(digits = 2)
+knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE) 
 ```
 
-- It is often important to divide a set of elements into similar 
-subgroups
-- This task can be automized and open source solutions exist 
-- Anticlustering = opposite of clustering, does that
-- Objectives based on pairwise distances exist, but k-means also exists 
-- k-means = optimizes similarity wrt to mean, but not other distribution
-parameters, in particular not wrt to the variance
-- here, I present an improved objective function for k-means 
-anticlustering, which optimizes similarity wrt to means and variance / 
-standard deviation at the same time
+Anticlustering is a method to partition a set of elements into different 
+groups with the focus of obtaining high between-group similarity and 
+within-group heterogeneity. Anticlustering reverses the logic of its 
+better known twin, cluster analysis that seeks homogeneity within 
+clusters and heterogeneity between clusters [@rokach2005; 
+@steinley2006]. The method of anticlustering is useful in many 
+tasks surrounding the work of researchers in psychology, for example, 
+when splitting a test into parts of equal difficulty [@gierl2017], when 
+assigning students to work groups [@baker2002], or when assigning 
+stimuli to experimental conditions [@lahl2006]. Solving such problems 
+"by hand" is a tedious and time-consuming task and the quality of 
+manually established groups is usually improvable. Fortunately, these 
+problems can be formalized mathematically via several anticlustering 
+objective functions [e.g., @brusco2019; @spath1986; @baker2002; 
+@fernandez2013], and accessible open source software solutions exist 
+[@papenberg2020].
+
+Several anticlustering objectives are available to assess between-group 
+similarity or within-group heterogeneity. In many cases, both objectives 
+are directly related, that is, maximizing heterogeneity within groups 
+simultaneously maximizes between-group similarity. Several 
+anticlustering objectives are based pairwise dissimilarity ratings; the 
+most prominent criterion of this kind is the *diversity*, which is the 
+total sum of dissimilarities between any elements within the same group. 
+Note that maximizing the diversity is equivalent to minimizing 
+the sum of dissimilarities between elements in different clusters, 
+directly illustrating the correspondence between within-group 
+heterogeneity and between-group similarity. Another important criterion 
+based on pairwise dissimilarities is the dispersion, which is the 
+minimum dissimilarity between any two items in the same group 
+[@fernandez2013; @brusco2019]. Maximizing the dispersion increases the 
+within-group heterogeneity by ensuring that any two elements as 
+dissimilar as possible from each other. Note that maximizing the 
+dispersion no longer simultaneously maximize between-group similarity. 
+In this case, these two potential aims of anticlustering diverge.
+
+Oftentimes, the available data does not represent pairwise 
+similarity, but rather numeric, categorical or binary attributes 
+[@dry2009]. That is, the data is a two-way table where rows represent 
+elements and columns represent variables describing the elements. For 
+example, numeric ratings may describe the attributes of word stimuli for 
+psycholinguistic experiments. Such data may include Norms for imagery, 
+concreteness and orthographic variables [@friendly1982]. In other cases, 
+attributes are binary and may represent the presence or absence of 
+a feature [@tversky1977]. Attribute data can be converted to pairwise 
+dissimilarities for example by computing an appropriate dissimilarity 
+measure, such as the pairwise Euclidean or squared Euclidean distance 
+across a set of numeric attributes [@brusco2019]. Then, anticlustering 
+criteria based on pairwise dissimilarity can be applied to partition the 
+elements.
+
+A different approach to anticlustering directly works with feature based 
+data and does not require a transformation into pairwise 
+dissimilarities: k-means anticlustering. Späth [-@spath1986] and Valev 
+[-@valev1998; -@valev1983] independently recognized that by optimizing 
+the reversed objective function of the popular k-means algorithm 
+[@jain2010; @steinley2006], a data set is partitioned into similar 
+groups. They originally---and independently from each other---coined the 
+term anticlustering for this purpose. K-means anticlustering maximizes the 
+sum of squared Euclidean distances from each element to the center of 
+the cluster to which it belongs [cf. @aloise2009np]. As @spath1986 
+noted, k-means anticlustering thereby directly minimizes differences 
+with regard to the means of the numeric variables between groups. Hence, 
+groups only become similar with regard to the mean of the distribution 
+of the numeric attributes, which may perform poorly as a measure of 
+overall similarity; the spread of the data is not targeted and may 
+differ between sets. @papenberg2020 showed that maximizing the 
+diversity (based on the pairwise Euclidean distances) is better suited 
+to minimize difference with regard to both the mean and the variance of 
+the data; it was shown that even an entirely random split usually 
+resulted in more similar standard deviations as k-means solely focuses 
+on the mean. When requiring that groups are overall similar to each 
+other, this focus may be detrimental as similar means can be obtained 
+even when the underlying distributions are very different 
+[@anscombe1973].
+
+In this paper, I present a simple extension of the k-means objective 
+function that ensures when maximizing the criterion, not only the means
+but also the variances become similar between groups. Hence, this paper 
+is more concerned with the goal of forming similar groups, rather than 
+maximizing within-group heterogeneity.
 
@@
 
-# set.seed(26178) # uncomment later
-
-library(anticlust)
+set.seed(4318) # 4318
 
@@
 partitions <- generate_partitions(N, K)
-features <- rnorm(N)
+features <- round(rnorm(N, 15, 3), 1)
 
@@
 
-The plot illustrates all possible ways to partition a data set of $N = 
-14$ into two parts (there are `r length(partitions)` possible 
-partitions). Red point: maximum similarity in means; green point 
+An example will show that aiming not only for similarity in attribute 
+means is not only useful, but also possible. I created 14 data points 
+from a normal distribution ($M = 15$, $SD = 3$, see Table 1). The data 
+points in Figure 1 illustrate all `r length(partitions)`  ways to 
+partition a data set of $N = 14$ into two parts (the Appendix contains 
+an explanation how to compute the number of partitions and how to 
+generate them). Red point: maximum similarity in means; green point 
 = maximum similarity in variance. Blue point: Almost no difference to 
@@
 means (which is what maximizing the k-means objective does).
+
+```{r datatables}
+
+best_partition_mean <- partitions[[which.min(df$diff_mean)]]
+best_partition_var  <- partitions[[which.min(df$diff_var)]]
+best_partition_combined <- partitions[[which.min(df$sum)]]
+
+data <- data.frame(
+  x = features,
+  best_mean = best_partition_mean,
+  best_var = best_partition_var,
+  best_combined = best_partition_combined
+
+)
+
+descriptives <- cbind(
+  mean_sd_tab(as.matrix(features), best_partition_mean),
+  mean_sd_tab(as.matrix(features), best_partition_var),
+  mean_sd_tab(as.matrix(features), best_partition_combined)
+)
+
+descriptives <- paste("M =", gsub("[(]", "(SD = ", descriptives))
+dim(descriptives) <- c(2, 3)
+colnames(descriptives) <- colnames(data[,-1])
+apa_table(data, caption = "Illustrative data set")
+apa_table(descriptives)
+
+```
 

### 2020-07-10 solution-structure-large-data.R
added visualization of exchange method for larger data

@@ -1,0 +1,94 @@
+
+# Investigate why k-variance is worse than k-means-variance for 
+# optimizing between-group similarity in variance
+
+source("exchange_method.R")
+
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
+
+library(anticlust)
+library(tidyverse)
+
+N <- 200
+M <- 4
+K <- 4
+features <- matrix(rnorm(N * M), ncol = M)
+dist_features <- squared_from_mean(features)
+combined_features <- cbind(features, dist_features)
+
+# Initialize a partition
+init <- sample(rep_len(1:K, N))
+init <- order_cluster_vector(init)
+
+# Do local search for three objectives
+partitions_means <- local_maximum_exchange(
+  features, 
+  init,
+  anticlust::variance_objective
+)
+partitions_variance <- local_maximum_exchange(
+  dist_features, 
+  init,
+  anticlust::variance_objective
+)
+partitions_means_variance <- local_maximum_exchange(
+  combined_features, 
+  init,
+  anticlust::variance_objective
+)
+
+var_objective <- function(clusters, data) {
+  anticlust::variance_objective(data, clusters)
+}
+
+partitions <- list(
+  kMeans = partitions_means, 
+  kVariance = partitions_variance, 
+  kExtended = partitions_means_variance
+)
+all_features <- list(
+  kMeans = features, 
+  kVariance = dist_features
+)
+
+
+results_list <- list()
+# iterate over data and objectives
+for (i in names(partitions)) {
+  iteration <- 1:length(partitions[[i]])
+  for (j in names(all_features)) {
+    objectives <- sapply(partitions[[i]], var_objective, data = all_features[[j]])
+    results_list[[paste0(i, "-", j)]] <- data.frame(
+      objectives, method = i, objective = j, iteration = iteration
+    )
+  }
+}
+results <- do.call(rbind, results_list)
+rownames(results) <- NULL
+
+results <- pivot_wider(results, names_from = objective, values_from = objectives)
+
+
+# Plot data
+ggplot(results, aes(x = kMeans, y = kVariance, group = method)) +
+  geom_path(aes(col = method)) + 
+  geom_point(aes(col = method)) + 
+  geom_point(aes(x = kMeans[1], y = kVariance[1]))
+
+options(pillar.sigfig = 6)
+
+results %>% 
+  group_by(method) %>% 
+  summarize(
[…]

### 2020-07-10 Application-1-Papenberg-Klau-2020.R
added application from last paper, added new objective

@@ -1,0 +1,82 @@
+
+# Author: Martin Papenberg
+# Year: 2019-2020
+
+# This script reproduces Application I (Stimulus assignment) in 
+# "Using anticlustering to partition data sets into equivalent parts" 
+# (Papenberg & Klau, 2020).
+
+# The code can simply be executed from top to bottom. 
+
+## 1. Load - and, if required, install - package `anticlust`
+
+if (!requireNamespace("anticlust")) {
+  if (!requireNamespace("remotes")) {
+    install.packages("remotes")
+  }
+  remotes::install_github("m-Py/anticlust")
+}
+
+library(anticlust)
+
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
+
+## 2. Set random seed for reproducibility
+set.seed(1234) 
+
+## 3. Load data (is included in the package)
+data(schaper2019)
+
+# Look at data:
+head(schaper2019)
+# For more information on the data set, umcomment and execute the
+# following line:
+# ?schaper2019
+
+## 4. Anticlustering
+
+# Select the four features that should be similar across 3 lists:
+features <- schaper2019[, 3:6]
+
+# Anticluster editing
+ac_ace <- anticlustering(
+  features, 
+  K = 3, # 3 stimulus sets
+  categories = schaper2019$room, # balances category across lists
+  objective = "diversity", 
+  method = "local-maximum",
+  repetitions = 10
+)
+
+# K-Means Anticlustering
+ac_kmeans <- anticlustering(
+  features, 
+  K = 3, 
+  categories = schaper2019$room,
+  objective = "variance", 
+  method = "local-maximum",
+  repetitions = 10
+)
+
+# K-Means-Extended
+ac_kmeans_extended <- anticlustering(
+  scale(cbind(features, squared_from_mean(features))), 
+  K = 3, 
+  categories = schaper2019$room,
+  objective = "variance", 
+  method = "local-maximum",
+  repetitions = 10
+)
+
+## 5. Evaluate results
+
+# Compare feature means and standard deviations
+# Anticluster editing
+mean_sd_tab(features, ac_ace)
+
+# k-means anticlustering
+mean_sd_tab(features, ac_kmeans)
+
+mean_sd_tab(features, ac_kmeans_extended)

### 2020-07-10 pareto_front.R
visualizing paretor front for k-means / variance

@@ -1,0 +1,65 @@
+
+# Investigate why k-variance is worse than k-means-variance for 
+# optimizing between-group similarity in variance
+
+library(anticlust)
+
+N <- 14
+K <- 2
+M <- 2
+all_partitions <- generate_partitions(N, K)
+features <- matrix(rnorm(N * M), ncol = M)
+
+obj_fun <- function(partitions, features) {
+  variance_objective(features, partitions)
+}
+
+all_objectives_mean <- sapply(
+  all_partitions,
+  FUN = obj_fun,
+  features = features
+)
+
+all_objectives_var <- sapply(
+  all_partitions,
+  FUN = obj_fun,
+  features = squared_from_mean(features)
+)
+
+df <- data.frame(
+  kMeans = all_objectives_mean,
+  kVariance  = all_objectives_var
+)
+
+# exchange method for finding all objectives
+
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
+
+# Compute Pareto efficient set!
+
+plot(df, col = "darkgrey", las = 1, pch = 4, cex = 0.8)
+
+# For a data frame (rows = partitions, columns = objectives), determine which 
+# elements are part of the Pareto efficient set. Output is a logical vector,
+# illustrating for each row if it is part of the Pareto efficient set.
+pareto_set <- function(df) {
+  is_dominated <- rep(FALSE, nrow(df)) 
+  for (i in 1:nrow(df)) {
+    for (j in 1:nrow(df)) {
+      j_dominates_i <- all(df[i, ] <= df[j, ]) && any(df[i, ] < df[j, ])
+      if (j_dominates_i) {
+        is_dominated[i] <- TRUE
+        break
+      }
+    }
+  }
+  !is_dominated
+}
+
+efficient_set <- pareto_set(df)
+plot(df, col = "darkgrey", las = 1, pch = 4, cex = 0.8)
+points(df[efficient_set, ], col = "red", cex = 1.2, pch = 19)
+
+

### 2020-07-10 paper.Rmd
some notes on writing

@@
 In this paper, I present a simple extension of the k-means objective 
-function that ensures when maximizing the criterion, not only the means
-but also the variances become similar between groups. Hence, this paper 
-is more concerned with the goal of forming similar groups, rather than 
-maximizing within-group heterogeneity.
+function to maximize simililarity with regard to means and variances 
+at the same time.
 
@@
 
-## K-Means criterion
+## K-Means objective
 
@@
 
----
+## K-variances
 
@@
 
-# Stuff
+## Bicriterion optimization model
 
-New objective seems to work best when N is high relative to K; 
+- formalize the weighted-sum objective and use the weighted-sum method.
+works well on this specific problem [although in general there are 
+caveats to the weighted-sum method, see @brusco2019]. is fast, good 
+results. in fact, the objectives seem to work well together (optimizing 
+k-means extended often yields more similar variances than just 
+k-variances!!!)
+
+- In this paper: Formalization of the objective function, recognition
+that this formulation does not change the basic structure of the k-means
+objective, because it reduces to augmentation of the data input
+
+- not interested in the Pareto efficient set! Because: Overall 
+similarity should be maximized, which requires that both criteria are 
+satisfied to a strong degree (not interested in just maximizing 
+similarity wrt variance; this may be interesting in other applications). 
+Instead: A tradeoff is possible, and in many cases, simultaneous 
+optimization will meet both criteria almost as well as optimizing just
+either criterion. To justify this approach: show the pareto front, which 
+is not a diagonal / circle [see @brusco2012] but a RIGHT TRIANGLE! (this 
+is probably because this is *anti*clustering!)
+
+- Evaluation will consist of a comparison of different objectives based 
+on the same (simple) exchange algorithm
+
+- Future research: Application of more sophisticated bicriterion 
+optimization algorithms are interesting
+
+- How to select weights? (Oftentimes: for equal weights, the results 
+are good)
+
+- Standardization of the extended data matrix to ensure that the two
+criterion functions do no differ markedly in scale [@brusco2019]
+
+- Using variable weighting (see Steinley, 2006, p. 13), we can 
+adjust the relative importance of minimizing differences in means
+and variances. Standard in `anticlust`: Equal weight, this is 
+achieved by first creating the new variables, representing 
+the squared differences from the mean, and then standardizing
+the entire feature matrix. If $N$ is large enough (relative to $K$),
+there is usually no need for weighting.
+
+# Evaluation
+
+- Do simulations to show that k-variance anticlustering outperforms 
+other objectives with regard to similarity in means / variances
+
+- New objective seems to work best when N is high relative to K; 
 optimizing two objectives at the same time is hard when N is low. Easier 
@@
 
-# Weighted k-means
-
-Using variable weighting (see Steinley, 2006, p. 13), we can 
-adjust the relative importance of minimizing differences in means
-and variances. Standard in `anticlust`: Equal weight, this is 
-achieved by first creating the new variables, representing 
-the squared differences from the mean, and then standardizing
-the entire feature matrix. If $N$ is large enough (relative to $K$),
-there is usually no need for weighting.
-
-# Simulation
-
-- show that k-variance anticlustering outperforms other objectives
+- Often, maximizing k-means-variance yields more similar variances 
+than k-variance alone. Why is that? I guess: structure in the solution 
+space, local maxima are real. When also considering means, you more 
+often escape bad local maxima?
 
@@
 - use some examples (e.g., OASIS data set)
-
-# Combined approach better than k-variance alone
-
-- Often, maximizing k-means-variance yields more similar variances 
-than k-variance alone. Why is that? I guess: structure in the solution 
-space, local maxima are real. When also considering means, you more 
-often escape bad local maxima?
 

### 2020-07-14 K-Means-Vs-Difference.R
extracted some often used functions

@@ -9,36 +9,12 @@
 
 library(anticlust)
 source("exchange_method.R")
+source("misc-functions.R")
 
 N <- 12
 K <- 3
 all_partitions <- generate_partitions(N, K)
 features <- as.matrix(rnorm(N))
-
-## Create an objective function that takes the partition
-## as first argument (then, we can use sapply to compute
-## the objective for each partition)
-
-sum_diff_vars <- function(clusters, features) {
-  features <- as.matrix(features) 
-  sum(apply(features, 2, diff_vars, features = features, clusters = clusters))
-}
-
-diff_vars <- function(clusters, features) {
-  abs(diff(range(tapply(features, clusters, var))))
-}
-
-diff_mean <- function(clusters, features) {
-  abs(diff(range(tapply(features, clusters, mean))))
-}
-
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-obj_fun <- function(clusters, features) {
-  variance_objective(features, clusters)
-}
 
 all_objectives_var <- sapply(
   all_partitions,
@@ -54,19 +30,19 @@ all_objectives_mean <- sapply(
 
 all_objectives_kVariances <- sapply(
   all_partitions,
-  FUN = obj_fun,
+  FUN = var_objective,
   features = squared_from_mean(features)
 )
 
-all_objectives_kMeans<- sapply(
+all_objectives_kMeans <- sapply(
   all_partitions,
-  FUN = obj_fun,
+  FUN = var_objective,
   features = features
 )
 
 all_objectives_kExtended <- sapply(
   all_partitions,
-  FUN = obj_fun,
+  FUN = var_objective,
   features = cbind(squared_from_mean(features), features)
 )
 
@@ -85,7 +61,6 @@ plot(
   pch = 4, 
   cex = 0.8
 )
-
 
 # Initialize a partition
 init <- sample(rep_len(1:K, N))
@@ -142,28 +117,10 @@ points(plot_df [match(list(init), all_partitions), , drop = FALSE], cex = 2, pch
 legend("topright", legend = c("k-means", "k-variance", "k-extended"),
        pch = c(19, 18, 17), col = c("#28E2E5", "#2297E6", "#CD0BBC"))
 
-nth <- function(x) {
-  x[[length(x)]]
-}
-
-# Check out similarity of item means (rounded on two decimals)
-fun_by_group <- function(features, clusters, fun) {
[…]

### 2020-07-14 improved_objective.R

@@ -1,14 +1,9 @@
 
 library(anticlust)
 
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-
 N <- 150
-M <- 1
-K <- 2
+M <- 2
+K <- 3
 features <- matrix(rnorm(N * M), ncol = M)
 #features <- schaper2019[, 3:6]
 
@@ -64,7 +59,7 @@ variance_objective(features, ac_kmeans_var) / variance_objective(features, ac_km
 # WHY IS K-MEANS EXTENDED BETTER AT MINIMIZING MAXIMUM DIFFERENCE IN 
 # VARIANCE THAN K-VARIANCE ALONE? NOT SURE YET. (and this is only the 
 # case for K > 2)
-sum(group_diff_min_max(features, ac_kmeans, var))
-sum(group_diff_min_max(features, ac_kmeans_var, var))
-sum(group_diff_min_max(features, ac_kmeans_var_only, var))
+sum(group_diff_min_max(ac_kmeans, features, var))
+sum(group_diff_min_max(ac_kmeans_var, features, var))
+sum(group_diff_min_max(ac_kmeans_var_only, features, var))
 

### 2020-07-14 misc-functions.R

@@ -1,0 +1,45 @@
+
+# Compute features for k-means-extended criterion
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
+
+# Get the last element of a list / vector
+last <- function(x) {
+  x[[length(x)]]
+}
+
+# Difference between minimum and maximum value per feature of a data frame,
+# by a group variable
+group_diff_min_max <- function(clusters, features, fun) {
+  apply(fun_by_group(features, clusters, fun), 1, diff_min_max)
+}
+
+# Function that computes a function for each feature of a data frame,
+# by a grouping variable
+fun_by_group <- function(features, clusters, fun) {
+  data.frame(lapply(by(features, clusters, function(x) apply(x, 2, fun)), c))
+}
+
+# Difference between minimum and maximum in a vector
+diff_min_max <- function(x) {
+  diff(range(x))
+}
+
+# short-cut function to determine difference in minimum and maximum
+# variance, across groups, for one feature
+diff_vars <- function(clusters, features) {
+  abs(diff(range(tapply(features, clusters, var))))
+}
+
+# short-cut function to determine difference in minimum and maximum
+# means, across groups, for one feature
+diff_mean <- function(clusters, features) {
+  abs(diff(range(tapply(features, clusters, mean))))
+}
+
+# Variance objective function with changed order of the arguments 
+# (can be called by sapply() etc.)
+var_objective <- function(clusters, features) {
+  anticlust::variance_objective(features, clusters)
+}

### 2020-07-14 solution-structure-large-data.R

@@ -3,15 +3,13 @@
 # optimizing between-group similarity in variance
 
 source("exchange_method.R")
+source("misc-functions.R")
 
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
 
 library(anticlust)
 library(tidyverse)
 
-N <- 200
+N <- 100
 M <- 4
 K <- 4
 features <- matrix(rnorm(N * M), ncol = M)
@@ -39,10 +37,6 @@ partitions_means_variance <- local_maximum_exchange(
   anticlust::variance_objective
 )
 
-var_objective <- function(clusters, data) {
-  anticlust::variance_objective(data, clusters)
-}
-
 partitions <- list(
   kMeans = partitions_means, 
   kVariance = partitions_variance, 
@@ -52,7 +46,6 @@ all_features <- list(
   kMeans = features, 
   kVariance = dist_features
 )
-
 
 results_list <- list()
 # iterate over data and objectives
@@ -89,6 +82,6 @@ best_mean <- partitions_means[[length(partitions_means)]]
 best_var <- partitions_variance[[length(partitions_variance)]]
 best_mean_var <- partitions_means_variance[[length(partitions_means_variance)]]
 
-mean_sd_tab(features, best_mean, return_diff = TRUE)
-mean_sd_tab(features, best_var, return_diff = TRUE)
-mean_sd_tab(features, best_mean_var, return_diff = TRUE)
+sum(group_diff_min_max(best_mean, features, var))
+sum(group_diff_min_max(best_var, features, var))
+sum(group_diff_min_max(best_mean_var, features, var))

### 2020-07-14 paper.Rmd
some notes on validation

@@
 
-- Do simulations to show that k-variance anticlustering outperforms 
-other objectives with regard to similarity in means / variances
-
-- New objective seems to work best when N is high relative to K; 
-optimizing two objectives at the same time is hard when N is low. Easier 
-when only optimizing one objective (i.e., the mean, as done by normal
-k-means anticlustering).
-
-- Two examples: $K = 2$, $N = 20$ vs. $N = 100$, using classical
+- Simulation 1: Show that k-means extended obtains 99.999% of the 
+k-means and k-variance objective, on average. 
+  + New objective seems to work best when N is high relative to K; optimizing 
+two objectives at the same time is hard when N is low. Easier when only 
+optimizing one objective (i.e., the mean, as done by normal k-means 
+anticlustering). What I can say: When N / K > X, the objective is 
+usually 99.99% of both objectives.
+  + Two examples: $K = 2$, $N = 20$ vs. $N = 100$, using classical
 k-means: only mean; using k-variance: only variance; combination:
 both, but works better for larger $N$
+
+- Simulation 2: show that k-variance anticlustering outperforms 
+other objectives with regard to similarity in means / variances 
+(in particular: anticluster editing)
 
@@
 space, local maxima are real. When also considering means, you more 
-often escape bad local maxima?
+often escape bad local maxima? The finding is: k-means-extended better 
+minimizes the maximum difference between the variance across groups 
+than k-variance (for K > 2), which in the eye of the researcher, may 
+be a more intuitive criterion.
 
@@
 
-- use some examples (e.g., OASIS data set)
-
-# Test equivalence of objectives
-
-Check that I can actually just append the squared difference data as 
-additional columns to the original data frame:
-
-```{r}
-
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-N <- 500
-M <- 20
-data <- matrix(rnorm(N * M), ncol = M)
-var_data <- squared_from_mean(data)
-
-K <- 10
-clusters <- sample(rep_len(1:K, N))
-
-obj1 <- variance_objective(data, clusters) + variance_objective(var_data, clusters)
-obj2 <- variance_objective(cbind(data, var_data), clusters)
-
-tolerance <- 1e-10
-
-obj1 - obj2 < tolerance
-
-obj1
-obj2
-
-```
+- use some examples (e.g., OASIS data set, 900 items allocated to 10 
+sets, same mean and variance across the 10 sets)
 

### 2020-07-14 equivalent-objectives.R
script to verify equivalence of objectives

@@ -1,0 +1,28 @@
+
+# Test equivalence of objectives
+
+# Check that I can actually just append the squared difference data as 
+# additional columns to the original data frame:
+
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
+
+
+N <- 500
+M <- 20
+data <- matrix(rnorm(N * M), ncol = M)
+var_data <- squared_from_mean(data)
+
+K <- 10
+clusters <- sample(rep_len(1:K, N))
+
+obj1 <- variance_objective(data, clusters) + variance_objective(var_data, clusters)
+obj2 <- variance_objective(cbind(data, var_data), clusters)
+
+tolerance <- 1e-10
+
+obj1 - obj2 < tolerance
+
+obj1
+obj2

### 2020-07-14 K-Means-Vs-Difference.R
misc changes to R scripts

@@ -11,21 +11,24 @@ library(anticlust)
 source("exchange_method.R")
 source("misc-functions.R")
 
-N <- 12
+N <- 15
+M <- 3
 K <- 3
+features <- matrix(rnorm(N * M), ncol = M)
 all_partitions <- generate_partitions(N, K)
-features <- as.matrix(rnorm(N))
 
 all_objectives_var <- sapply(
   all_partitions,
-  FUN = diff_vars,
-  features = features
+  FUN = sum_group_diff_min_max,
+  features = features,
+  fun = var
 )
 
 all_objectives_mean <- sapply(
   all_partitions,
-  FUN = diff_mean,
-  features = features
+  FUN = sum_group_diff_min_max,
+  features = features,
+  fun = mean
 )
 
 all_objectives_kVariances <- sapply(
@@ -54,13 +57,8 @@ df <- data.frame(
   kExt = all_objectives_kExtended
 )
 
-plot(
-  df, 
-  col = "darkgrey", 
-  las = 1, 
-  pch = 4, 
-  cex = 0.8
-)
+plot(df, col = "darkgrey", las = 1, pch = 4, cex = 0.8)
+
 
 # Initialize a partition
 init <- sample(rep_len(1:K, N))
@@ -112,7 +110,7 @@ points(
   type = "b",
   lwd = 2
 )
-points(plot_df [match(list(init), all_partitions), , drop = FALSE], cex = 2, pch = 19)
+points(plot_df[match(list(init), all_partitions), , drop = FALSE], cex = 3, pch = 15)
 
 legend("topright", legend = c("k-means", "k-variance", "k-extended"),
        pch = c(19, 18, 17), col = c("#28E2E5", "#2297E6", "#CD0BBC"))
@@ -121,6 +119,10 @@ legend("topright", legend = c("k-means", "k-variance", "k-extended"),
 # with regard to similarity of the groups!
 fun <- var
 
-group_diff_min_max(last(partitions_means), features, fun)
-group_diff_min_max(last(partitions_means_variance), features, fun)
-group_diff_min_max(last(partitions_variance), features, fun)
+variance_objective(squared_from_mean(features), last(partitions_means_variance))
+variance_objective(squared_from_mean(features), last(partitions_variance))
+
+sum_group_diff_min_max(last(partitions_means_variance), features, fun)
+sum_group_diff_min_max(last(partitions_variance), features, fun)
+
+round(cor(plot_df), 2)

### 2020-07-14 misc-functions.R

@@ -9,6 +9,13 @@ last <- function(x) {
   x[[length(x)]]
 }
 
+# Sum of difference between minimum and maximum value per feature of a 
+# data frame, by a group variable (i.e., total deviation in variances,
+# measured by difference between min/max variance, across several features)
+sum_group_diff_min_max <- function(clusters, features, fun) {
+  sum(group_diff_min_max(clusters, features, fun))
+}
+
 # Difference between minimum and maximum value per feature of a data frame,
 # by a group variable
 group_diff_min_max <- function(clusters, features, fun) {

### 2020-07-14 pareto_front.R

@@ -3,6 +3,7 @@
 # optimizing between-group similarity in variance
 
 library(anticlust)
+source("misc-functions.R")
 
 N <- 14
 K <- 2
@@ -12,6 +13,10 @@ features <- matrix(rnorm(N * M), ncol = M)
 
 obj_fun <- function(partitions, features) {
   variance_objective(features, partitions)
+}
+
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
 }
 
 all_objectives_mean <- sapply(
@@ -30,12 +35,6 @@ df <- data.frame(
   kMeans = all_objectives_mean,
   kVariance  = all_objectives_var
 )
-
-# exchange method for finding all objectives
-
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
 
 # Compute Pareto efficient set!
 

### 2020-07-14 K-Means-Vs-Difference.R
more exploration of the solution space

@@ -57,7 +57,7 @@ df <- data.frame(
   kExt = all_objectives_kExtended
 )
 
-plot(df, col = "darkgrey", las = 1, pch = 4, cex = 0.8)
+#plot(df, col = "darkgrey", las = 1, pch = 4, cex = 0.8)
 
 
 # Initialize a partition

### 2020-07-14 diff-min-max-mean-variance.R

@@ -1,0 +1,126 @@
+
+# Investigate why k-variance is worse than k-means-variance for 
+# optimizing between-group similarity in variance
+
+# Reason: I validated the results using difference between minimum and
+# maximum *standard deviation* across groups. For K > 2, this is not 
+# equivalent to minimum and maximum *variance* across groups (which
+# is however minimized using the k-variance criterion!)
+
+library(anticlust)
+source("exchange_method.R")
+source("misc-functions.R")
+
+N <- 100
+M <- 4
+K <- 4
+features <- matrix(rnorm(N * M), ncol = M)
+dist_features <- squared_from_mean(features)
+combined_features <- cbind(features, dist_features)
+
+# Initialize a partition
+init <- sample(rep_len(1:K, N))
+init <- order_cluster_vector(init)
+
+# Do local search for three objectives
+partitions_means <- local_maximum_exchange(
+  features, 
+  init
+)
+partitions_variance <- local_maximum_exchange(
+  dist_features, 
+  init
+)
+partitions_means_variance <- local_maximum_exchange(
+  combined_features, 
+  init
+)
+
+objectives <- list()
+objectives[["diff_mean"]] <- list()
+objectives[["diff_var"]] <- list()
+
+funs <- list(diff_mean = mean, diff_var = var)
+
+for (i in c("diff_mean", "diff_var")) {
+  objectives[[i]][["kMeans"]] <- sapply(
+    partitions_means,
+    sum_group_diff_min_max,
+    features = features,
+    fun = funs[[i]]
+  )
+  objectives[[i]][["kVar"]] <- sapply(
+    partitions_variance,
+    sum_group_diff_min_max,
+    features = features,
+    fun = funs[[i]]
+  )
+  objectives[[i]][["kExt"]] <- sapply(
+    partitions_means_variance,
+    sum_group_diff_min_max,
+    features = features,
+    fun = funs[[i]]
+  )
+}
+
+
+
+sample_point <- function(clusters, features) {
+  cl <- sample(clusters)
+  c(
+    mean = sum_group_diff_min_max(cl, features, fun = mean),
+    var = sum_group_diff_min_max(cl, features, fun = var)
+  )
+}
+
[…]

### 2020-07-14 solution-structure-large-data.R

@@ -23,18 +23,15 @@ init <- order_cluster_vector(init)
 # Do local search for three objectives
 partitions_means <- local_maximum_exchange(
   features, 
-  init,
-  anticlust::variance_objective
+  init
 )
 partitions_variance <- local_maximum_exchange(
   dist_features, 
-  init,
-  anticlust::variance_objective
+  init
 )
 partitions_means_variance <- local_maximum_exchange(
   combined_features, 
-  init,
-  anticlust::variance_objective
+  init
 )
 
 partitions <- list(

### 2020-07-14 exchange_method.R
using speed optimized local update exchange method

@@ -1,6 +1,6 @@
 
 # Call the exchange method below until a local maximum is reached
-local_maximum_exchange <- function(features, init, obj_function) {
+local_maximum_exchange <- function(features, init) {
   best_partition <- init
   found_partitions <- list(best_partition)
   new_obj <- variance_objective(features, best_partition)
@@ -8,64 +8,79 @@ local_maximum_exchange <- function(features, init, obj_function) {
   while (new_obj > old_obj) {
     partitions <- exchange_method(
       features,
-      clusters = best_partition, 
-      obj_function = obj_function
+      clusters = best_partition
     )
     best_partition <- partitions[[length(partitions)]]
     old_obj <- new_obj
-    new_obj <- obj_function(features, best_partition)
+    new_obj <- anticlust::variance_objective(features, best_partition)
     found_partitions <- c(found_partitions, partitions)
   }
   found_partitions[!duplicated(found_partitions)]
 } 
 
 
-#' Solve anticlustering using the modified exchange method
+#' Solve k-means anticlustering using the modified exchange method
 #'
-#' @param data the data -- a N x N dissimilarity matrix or a N x M
-#'     table of item features
-#' @param init An initial assignment of clusters to groups (integer vector)
-#' @param obj_function the objective function. Takes as first argument
-#'     a cluster assignment and as second argument the data set `data`
-#'     (`data` is matrix, no `data.frame`).
+#' @param data the data -- an N x M table of item features
+#' @param clusters An initial cluster assignment
 #'
 #' @return The anticluster assignment
 #'
 #' @noRd
 #'
-#'
 
-exchange_method <- function(data, clusters, obj_function) {
-
-  all_partitions <- list(order_cluster_vector(clusters))
-  u <- 2
+exchange_method <- function(data, clusters) {
   N <- nrow(data)
-  best_total <- obj_function(data, clusters)
+  best_total <- variance_objective_(clusters, data)
+  centers <- cluster_centers(data, clusters)
+  distances <- dist_from_centers(data, centers, squared = TRUE)
+  all_partitions <- list(clusters)
+  u <- 2
+  
+  ## frequencies of each cluster are required for updating cluster centers:
+  tab <- c(table(clusters))
   for (i in 1:N) {
-    # exchange partners to not have the same cluster
-    feasible_partners <- (clusters != clusters[i])
-    exchange_partners <- (1:N)[feasible_partners]
-
+    # cluster of current item
[…]

### 2020-07-14 diff-min-max-mean-variance.R
minor

@@ -11,7 +11,7 @@ library(anticlust)
 source("exchange_method.R")
 source("misc-functions.R")
 
-N <- 100
+N <- 200
 M <- 4
 K <- 4
 features <- matrix(rnorm(N * M), ncol = M)
@@ -73,7 +73,7 @@ sample_point <- function(clusters, features) {
   )
 }
 
-pts <- data.frame(t(replicate(10000, sample_point(init, features))))
+pts <- data.frame(t(replicate(1000, sample_point(init, features))))
 plot(
   pts, 
   col = "darkgrey", 
@@ -122,5 +122,8 @@ points(objectives[["diff_mean"]][[c(1, 1)]],
 legend("topright", legend = c("k-means", "k-variance", "k-extended"),
        pch = c(19, 18, 17), col = c("#28E2E5", "#2297E6", "#CD0BBC"))
 
+sum_group_diff_min_max(last(partitions_means), features, var)
+sum_group_diff_min_max(last(partitions_variance), features, var)
+sum_group_diff_min_max(last(partitions_means_variance), features, var)
 
 # randomly sample points to illustrate the solution space

### 2020-07-14 solution-structure-large-data.R

@@ -49,7 +49,7 @@ results_list <- list()
 for (i in names(partitions)) {
   iteration <- 1:length(partitions[[i]])
   for (j in names(all_features)) {
-    objectives <- sapply(partitions[[i]], var_objective, data = all_features[[j]])
+    objectives <- sapply(partitions[[i]], var_objective, features = all_features[[j]])
     results_list[[paste0(i, "-", j)]] <- data.frame(
       objectives, method = i, objective = j, iteration = iteration
     )
@@ -69,16 +69,15 @@ ggplot(results, aes(x = kMeans, y = kVariance, group = method)) +
 
 options(pillar.sigfig = 6)
 
+# K-Variance better optimizes k-variance criterion ....
 results %>% 
   group_by(method) %>% 
   summarize(
     kVarianceObj = max(kVariance),
     kMeansObj = max(kMeans))
 
-best_mean <- partitions_means[[length(partitions_means)]]
-best_var <- partitions_variance[[length(partitions_variance)]]
-best_mean_var <- partitions_means_variance[[length(partitions_means_variance)]]
-
-sum(group_diff_min_max(best_mean, features, var))
-sum(group_diff_min_max(best_var, features, var))
-sum(group_diff_min_max(best_mean_var, features, var))
+# ... but is worse at minimizing the difference in range in variance,
+# compared to k-extended  
+sum(group_diff_min_max(last(partitions_means), features, var))
+sum(group_diff_min_max(last(partitions_variance), features, var))
+sum(group_diff_min_max(last(partitions_means_variance), features, var))

### 2020-07-15 render.sh
simple render script

@@ -1,0 +1,3 @@
+#!/bin/sh
+
+R -e 'rmarkdown::render("paper.Rmd")'

### 2020-07-15 paper.Rmd
some work on introduction

@@
 ---
-title             : "An extension of k-means anticlustering to optimize between-group similarity"
+title             : "A bicriterion extension of k-means anticlustering to optimize between-group similarity"
 shorttitle        : "K-means anticlustering extended"
@@
 documentclass     : "apa6"
-classoption       : "man"
+classoption       : "doc"
 output            : papaja::apa6_pdf
@@
 
-Oftentimes, the available data does not represent pairwise 
-similarity, but rather numeric, categorical or binary attributes 
-[@dry2009]. That is, the data is a two-way table where rows represent 
-elements and columns represent variables describing the elements. For 
-example, numeric ratings may describe the attributes of word stimuli for 
-psycholinguistic experiments. Such data may include Norms for imagery, 
-concreteness and orthographic variables [@friendly1982]. In other cases, 
-attributes are binary and may represent the presence or absence of 
-a feature [@tversky1977]. Attribute data can be converted to pairwise 
-dissimilarities for example by computing an appropriate dissimilarity 
-measure, such as the pairwise Euclidean or squared Euclidean distance 
-across a set of numeric attributes [@brusco2019]. Then, anticlustering 
-criteria based on pairwise dissimilarity can be applied to partition the 
-elements.
+Oftentimes, the available data does not represent pairwise similarity, 
+but rather numeric, categorical or binary attributes [@dry2009]. That 
+is, the data is a two-way two-mode [@arabie1992] table where rows 
+represent the entities of study, and columns represent variables 
+describing the elements. For example, numeric ratings may describe the 
+attributes of word stimuli for psycholinguistic experiments. Such data 
+may include Norms for imagery, concreteness and orthographic variables 
+[@friendly1982]. In other cases, attributes are binary and may represent 
+the presence or absence of a feature [@tversky1977]. Such data can be 
+converted to pairwise dissimilarities for example by computing an 
+appropriate dissimilarity measure, such as the pairwise Euclidean or 
+squared Euclidean distance across a set of numeric attributes 
+[@brusco2019]. Then, anticlustering criteria based on pairwise 
+dissimilarity can be applied to partition the elements.
 
@@
 data and does not require a transformation into pairwise 
-dissimilarities: k-means anticlustering. Späth [-@spath1986] and Valev 
-[-@valev1998; -@valev1983] independently recognized that by optimizing 
-the reversed objective function of the popular k-means algorithm 
-[@jain2010; @steinley2006], a data set is partitioned into similar 
-groups. They originally---and independently from each other---coined the 
-term anticlustering for this purpose. K-means anticlustering maximizes the 
-sum of squared Euclidean distances from each element to the center of 
-the cluster to which it belongs [cf. @aloise2009np]. As @spath1986 
-noted, k-means anticlustering thereby directly minimizes differences 
-with regard to the means of the numeric variables between groups. Hence, 
-groups only become similar with regard to the mean of the distribution 
-of the numeric attributes, which may perform poorly as a measure of 
-overall similarity; the spread of the data is not targeted and may 
-differ between sets. @papenberg2020 showed that maximizing the 
-diversity (based on the pairwise Euclidean distances) is better suited 
-to minimize difference with regard to both the mean and the variance of 
-the data; it was shown that even an entirely random split usually 
-resulted in more similar standard deviations as k-means solely focuses 
-on the mean. When requiring that groups are overall similar to each 
-other, this focus may be detrimental as similar means can be obtained 
-even when the underlying distributions are very different 
-[@anscombe1973].
+dissimilarities. Instead, the sum of the squared Euclidean distances 
+between elements and the cluster they are assigned to is minimized 
+[brusco2006branch], most prominently using the k-means heuristic 
+[@jain2010; @steinley2006]. K-means clustering maximizes between-cluster 
+separation and within-cluster homogeneity simultaneously 
+[@aloise2009np]. Späth [-@spath1986] and Valev [-@valev1998; 
+-@valev1983] independently recognized that by optimizing the reversed 
+k-means objective that is, maximizing the sum of distances between 
+elements and cluster centroids -- a data set is partitioned into similar 
+groups, thereby reversing the logic of classical clustering. They 
+independently from each other coined the term anticlustering for this 
+purpose. K-means anticlustering maximizes the sum of squared Euclidean 
+distances from each element to the center of the cluster to which it 
+belongs. As @spath1986 noted, k-means anticlustering thereby directly 
+minimizes differences with regard to the means of the numeric variables 
+between groups. Hence, groups only become similar with regard to the 
+mean of the distribution of the numeric attributes, which may perform 
+poorly as a measure of overall similarity; the spread of the data is not 
+targeted and may differ between sets. @papenberg2020 showed that 
+maximizing the diversity (based on the pairwise Euclidean distances) is 
+better suited to minimize difference with regard to both the mean and 
+the variance of the data; it was shown that even an entirely random 
+split usually resulted in more similar standard deviations as k-means 
+solely focuses on the mean. When requiring that groups are overall 
+similar to each other, this focus may be detrimental as similar means 
+can be obtained even when the underlying distributions are very 
+different [@anscombe1973].
 
 In this paper, I present a simple extension of the k-means objective 
-function to maximize simililarity with regard to means and variances 
-at the same time.
+function to maximize simililarity with regard to means and variances at 
+the same time. The aim is to establish groups that are overall similar 
+to each other. It is shown that this is achieved by formulating a 
+bicriterion optimization objective for k-means anticlustering. 
+Interestingly, under specific---and important---circumstances, this 
+reformulation reduces to an augmentation of the original data matrix 
+while the optimization objective remains unchanged. This insight has 
+important consequences as it allows to adapt standard optimization 
+schemes for single criterion anticlustering. I show that the extended 
+k-means criterion outperforms other anticlustering criteria when the 
+goal is the minimize difference with regard to the mean and variance of 
+the different groups. In most cases, both objectives can be optimized
+satisfactorily at the same time without any loss on either; when the 
+data set is large enough (usually, when $N > XX$), the bicriterion 
+optimization leads to > 99.99% objective performance as compared to the 
+single optimization of either k-means or k-variance.
 
@@
 
-An example will show that aiming not only for similarity in attribute 
-means is not only useful, but also possible. I created 14 data points 
-from a normal distribution ($M = 15$, $SD = 3$, see Table 1). The data 
-points in Figure 1 illustrate all `r length(partitions)`  ways to 
-partition a data set of $N = 14$ into two parts (the Appendix contains 
-an explanation how to compute the number of partitions and how to 
-generate them). Red point: maximum similarity in means; green point 
-= maximum similarity in variance. Blue point: Almost no difference to 
-red point wrt similarity in means, but variance is much more similar. 
-This blue partition is certainly preferable to the red one when maximal 
-overall similarity is required; the red one just optimizes similarity in 
-means (which is what maximizing the k-means objective does).
+An motivating example illustrates that aiming to optimize similarity of 
+means of standard deviations in a data set is not only desirable, but 
+also expected to be feasible. To this end, I created 14 data points from 
+a normal distribution ($M = 15$, $SD = 3$, see Table 1). The data points 
+in Figure 1 illustrate all `r length(partitions)`  ways to partition a 
+data set of $N = 14$ into two parts. Red point: maximum similarity in 
+means; green point = maximum similarity in variance. Blue point: Almost 
+no difference to red point wrt similarity in means, but variance is much 
+more similar. This blue partition is certainly preferable to the red one 
+when maximal overall similarity is required; the red one just optimizes 
+similarity in means (which is what maximizing the k-means objective 
+does).
 

### 2020-07-16 paper.Rmd
work on paper

@@
 title             : "A bicriterion extension of k-means anticlustering to optimize between-group similarity"
-shorttitle        : "K-means anticlustering extended"
+shorttitle        : "Bicriterion K-means anticlustering"
 
@@
 abstract: |
-  Anticlustering is an automated method to divide a data set into groups
+    Anticlustering is an automated method to partition data sets into groups 
     in such a way that high between-group similarity and high within-group 
-  heterogeneity is obtained. Anticlustering is accomplished by maximizing 
-  instead of minimizing a clustering objective.
+    heterogeneity is obtained. Anticlustering has important applications in 
+    psychological research, for example to partition stimulus sets into 
+    equivalent sets, or to divide a group of students into heterogeneous 
+    classes. K-means anticlustering is maximizes the error sum of squares, 
+    i.e., the sum of squared Euclidean distances from each element to the 
+    center of the group to which it belongs. Thereby, k-means anticlustering 
+    directly maximizes the similarity of the mean values of numeric 
+    attributes across groups, but ignores the spread of the data. 
+    This paper presents a straight-forward extension of 
+    the k-means objective function that allows to simultaneously maximize 
+    similarity with regard to the mean as well as the variance across 
+    groups. It is shown that when both criteria receive equal weights, the 
+    bicriterion optimization problem reduces to an augmentation of the input 
+    data, leaving the original k-means criterion unchanged. Tests on 
+    simulated and real data show that similarity with regard to means and 
+    variance criteria can usually be achieved to an astonishing degree with 
+    almost no cost to either criterion. It is therefore concluded that the 
+    extended model should be preferred to classical k-means anticlustering 
+    for optimizing between-group similarity. An exception is given for 
+    binary data that does not profit from considering equality in 
+    variances.
 
@@
 
-keywords          : "K-Means, Anticlustering, Variance"
+keywords          : "Anticlustering, variance, k-means, k-variance, bicriterion optimization"
 wordcount         : "X"
@@
 knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE)
+
+source("misc-functions.R")
+
 ```
@@
 between elements and the cluster they are assigned to is minimized 
-[brusco2006branch], most prominently using the k-means heuristic 
-[@jain2010; @steinley2006]. K-means clustering maximizes between-cluster 
-separation and within-cluster homogeneity simultaneously 
-[@aloise2009np]. Späth [-@spath1986] and Valev [-@valev1998; 
+[@brusco2006branch], most prominently using the k-means heuristic 
+[@jain2010; @steinley2006]. Minimizing the k-means criterion 
+simultaneously maximizes between-cluster separation and within-cluster 
+homogeneity [@aloise2009np]. Späth [-@spath1986] and Valev [-@valev1998; 
 -@valev1983] independently recognized that by optimizing the reversed 
@@
 
-## A Motivating Example
+# Problem formalization
 
-```{r}
+For the purpose of problem formalization, I adopt the notation Steinley 
+(2006) provided in his thorough synthesis of a half-century literature 
+on k-means. K-means clustering is used to partition two-way, two-mode 
+data (that is, $N$ elements each having $P$ attributes) into $K$ groups 
+$(C_1., \ldots, C_K)$ where $C_k$ is the set of $n_k$ objects in group 
+$k$. The partitioning is subject to the requirements that each element 
+is assigned to exactly one group, formalized by the two constraints
 
-set.seed(4318) # 4318
+$$
+\bigcup\limits_{k = 1}^{K} C_j = X
+$$
+$$
+C_j \cap C_k = \emptyset, \; \forall j, k \in \{1, \ldots, K\}, \; j \ne k
+$$
 
-N <- 14
-K <- 2
-partitions <- generate_partitions(N, K)
-features <- round(rnorm(N, 15, 3), 1)
+In anticlustering applications, we usually impose restrictions on 
+the group sizes $n_k$, with the most common restriction being that all 
+groups have equal size:
 
-## Create an objective function that takes the partition
-## as first argument (then, we can use sapply to compute
-## the objective for each partition)
-diff_mean <- function(clusters, features) {
-  abs(diff(tapply(features, clusters, mean)))
-}
+$$
+n_k = \frac{N}{K}, \forall k \in \{1, \ldots, K\}
+$$
 
-diff_vars <- function(clusters, features) {
-  abs(diff(tapply(features, clusters, var)))
-}
+In this paper, I will discuss only such applications where groups are of 
+equal size.
 
-all_objectives_mean <- sapply(
-  partitions,
-  FUN = diff_mean,
-  features = features
-)
-
-all_objectives_var <- sapply(
-  partitions,
-  FUN = diff_vars,
-  features = features
-)
-
-df <- data.frame(
-  diff_mean = all_objectives_mean,
-  diff_var  = all_objectives_var
-)
-
-plot(df, pch = 4, col = "darkgrey", las = 1)
-# Illustrate best partition wrt simililarity in means
-points(
-  all_objectives_mean[which.min(all_objectives_mean)],
-  all_objectives_var[which.min(all_objectives_mean)],
-  cex = 2, col = "#DF536B", pch = 16
-)
-
-df$sum <- rowSums(df)
-
-points(
-  all_objectives_mean[which.min(df$sum)],
-  all_objectives_var[which.min(df$sum)],
-  cex = 2, col = "#2297E6", pch = 16
-)
-
-points(
-  all_objectives_mean[which.min(df$diff_var)],
-  all_objectives_var[which.min(df$diff_var)],
-  cex = 2, col = "#61D04F", pch = 16
-)
-
-```
-
-An motivating example illustrates that aiming to optimize similarity of 
-means of standard deviations in a data set is not only desirable, but 
-also expected to be feasible. To this end, I created 14 data points from 
-a normal distribution ($M = 15$, $SD = 3$, see Table 1). The data points 
-in Figure 1 illustrate all `r length(partitions)`  ways to partition a 
-data set of $N = 14$ into two parts. Red point: maximum similarity in 
-means; green point = maximum similarity in variance. Blue point: Almost 
-no difference to red point wrt similarity in means, but variance is much 
-more similar. This blue partition is certainly preferable to the red one 
-when maximal overall similarity is required; the red one just optimizes 
-similarity in means (which is what maximizing the k-means objective 
-does).
-
-```{r datatables}
-
-best_partition_mean <- partitions[[which.min(df$diff_mean)]]
-best_partition_var  <- partitions[[which.min(df$diff_var)]]
-best_partition_combined <- partitions[[which.min(df$sum)]]
-
-data <- data.frame(
-  x = features,
-  best_mean = best_partition_mean,
-  best_var = best_partition_var,
-  best_combined = best_partition_combined
-
-)
-
-descriptives <- cbind(
-  mean_sd_tab(as.matrix(features), best_partition_mean),
-  mean_sd_tab(as.matrix(features), best_partition_var),
-  mean_sd_tab(as.matrix(features), best_partition_combined)
-)
-
-descriptives <- paste("M =", gsub("[(]", "(SD = ", descriptives))
-dim(descriptives) <- c(2, 3)
-colnames(descriptives) <- colnames(data[,-1])
-apa_table(data, caption = "Illustrative data set")
-apa_table(descriptives)
-
-```
-
-Unfortunately, improving algorithmic procedures will not help to find 
-better partitions -- the objective function has to change. I will show 
-an objective function that captures both similarity in means and 
-variance.
-
-## K-Means objective
-
-In this treatment of k-means, I adopt the notation Steinley (2006) 
-provided in his synthesis of the k-means literature. In clustering---or 
-anticlustering---, an item pool $X$ = $\{x_1, \ldots, x_N\}$ has to be 
-partitioned into $K$ groups $\{C_1, \ldots, C_K\}$. We assume that each 
-element $x_i$ ($i = 1, \ldots, N$), is a vector of length $P$, and each 
-entry describes one of its numeric measurements. Generally, $X$ 
-can is interpreted as an $N \times M$ matrix where each row represents 
-an data point $x_i$ (e.g., representing a stimulus or a person) and each 
-column is a numeric attribute (e.g., the number of syllables in a word, 
-or the attractiveness rating of a face).
-
-K-Means anticlustering aims to maximize the error sum of squares (SSE), 
-the within-group variance, which is the sum of the squared Euclidean 
-distances between each data point and its cluster centroid. The cluster 
+For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, 
+k-means clustering aims to minimize the error sum of squares (SSE)---the 
+within-group variance---which is the sum of the squared Euclidean 
+distances between each data point and its cluster centroid. Conversely, 
+k-means anticlustering aims to maximize the same objective. The cluster 
 centroid is given by averaging the values on each variable over the 
 elements within that cluster. The centroid of the *j*th variable in the 
-*k*th cluster, consisting of $n_k$ elements, is given as:
+*k*th cluster is given as:
 
 $$
-\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits {i \in C_k} x_{ij}
+\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} x_{ij}
 $$
 
-The k-means objective function, the error sum of squares, is then given 
-as:
+The k-means objective function is then given as
 
@@
 
-In anticlustering applications, we usually assume that all cluster are 
-of equal size. Hence
+Note there is a direct connection between $SSE$ and the cluster 
+centroids [@spath1986]. If the complete centroid vector for cluster 
+$C_k$ is given by
 
 $$
-n_k = \frac{N}{K}, \forall k \in \{1, \ldots, K\}
+\overline{\mathbf{x}}^{(k)} = 
+  (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})
 $$
 
-## K-variances
+and the $SSE$ is minimal, the centroid values 
+$\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as 
+possible from the overall data centroid $\overline{\mathbf{x}} = 
+(\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, 
+\overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the 
+overall mean value on the $p$th attribute:
+
+$$
+\overline{\mathbf{x}}_p = \frac{1}{N} \sum\limits_{i=1}^{N} x_{ip}
+$$
+
+If the $SSE$ is maximal---which is the goal of k-means 
+anticlustering---the cluster centroids are as close as possible to the 
+overall centroid $\overline{\mathbf{x}}$ and therefore to each other 
+[@spath1986]. Thus, k-means anticlustering directly optimizes the 
+similarity of the mean values of the $P$ attributes across the $K$ 
+clusters.
+
+## K-Variances
 
@@
 
-Important to note: We solve the same optimization problem, we just 
-change the input a bit (we double the number of "columns").
+## Criterion Tradeoff
 
-**Example table!!**
+```{r}
+
+# Generate data and all partitions
+
+set.seed(2207) # 2207
+
+N <- 14
+K <- 2
+partitions <- generate_partitions(N, K)
+features <- rnorm(N, 15, 3)
+
+# Compute objectives for each partition
+all_objectives_mean <- sapply(
+  partitions,
+  FUN = diff_mean,
+  features = features
+)
+
+all_objectives_var <- sapply(
+  partitions,
+  FUN = diff_vars,
+  features = features
+)
+
+df <- data.frame(
+  diff_mean = all_objectives_mean,
+  diff_var  = all_objectives_var
+)
+
+```
+
+A motivating example illustrates that aiming to simultaneously optimize 
+similarity of means and variances at the same time is not only 
+desirable, but may also be feasible. I created 14 data points from a 
+univariate normal distribution ($M = 15$, $SD = 3$) and generated all `r 
+length(partitions)` possible ways to partition the 14 data points into 
+$K = 2$ equal-sized groups. I computed the difference in the two means 
+and two variances for each partition. This simple difference metric was 
+chosen for purposes of interpretability; in fact, there is a a 1-to-1 
+correspondence between these differences and $SSE$ and 
+$SSE_{\mathrm{VAR}}$, respectively (Figure 1). Figure 2 illustrates 
+similarity in means plotted against similarity in variances across all 
+partitions. Ideally, a partition is chosen that minimizes the 
+difference with regard to means and variances (see the triangular 
+symbol). However, optimal k-means anticlustering would create two groups 
+that are not particularly similar with regard to their variance and 
+chose the partition that is represented by the squared symbol. The 
+diamond symbol represents that partition that where the two groups have 
+the least difference in variances. The black symbols illustrate the 
+other members of the optimal pareto set.
+
+```{r, fig.cap = "Illustrates the 1-to-1 relationship between SSE and difference in means (Panel A) and variance (Panel B).
+Each data point illustrates a partition, i.e., a way to divide a data set ($N = 14$) into two groups.", fig.height = 3.5}
+
+# For each partition: Compute SSE objective, and SSE(VAR)
+# SSE
+all_objectives_kMeans <- sapply(
+  partitions,
+  FUN = var_objective,
+  features = features
+)
+
+# SSE(VAR)
+all_objectives_kVar <- sapply(
+  partitions,
+  FUN = var_objective,
+  features = squared_from_mean(as.matrix(features))
+)
+
+# Plot relationship between objectives
+par(mfrow = c(1, 2))
+plot(
+  all_objectives_mean, 
+  all_objectives_kMeans, 
+  xlab = "mean(f1) - mean(f2)",
+  ylab = "SSE",
+  pch = 4, 
+  cex = .5,
+  col = "darkgrey"
+)
+legend("topright", legend = "A", cex = 1.3, bty = "n")
+plot(
+  all_objectives_var, 
+  all_objectives_kVar, 
+  xlab = "var(f1) - var(f2)",
+  ylab = "SSE_var",
+  pch = 4, 
+  cex = .5,
+  col = "darkgrey"
+)
+legend("topright", legend = "B", cex = 1.3, bty = "n")
+```
+
+```{r, fig.width = 4.5, fig.height = 4.5, fig.cap = "All partitions. Points = pareto efficient set."}
+
+# Plot SSE vs SSE(VAR)
+plot(
+  df, pch = 4, col = "darkgrey", 
+  las = 1, cex = .5,
+  xlab = "mean(f1) - mean(f2)",
+  ylab = "var(f1) - var(f2)"
+)
+
+# Illustrate Pareto set
+pareto_pts <- pareto_set(df, sense = "min")
+points(df[pareto_pts, ], pch = 19)
+
+# Illustrate best partition wrt SSE (difference in means)
+points(
+  all_objectives_mean[which.min(all_objectives_mean)],
+  all_objectives_var[which.min(all_objectives_mean)],
+  cex = 1.2, pch = 22, bg = "white", lwd = 1.7
+)
+
+df$sum <- rowSums(df)
+
+# Illustrate best partition wrt SSE(VAR) (i.e., difference in variance)
+points(
+  all_objectives_mean[which.min(df$sum)],
+  all_objectives_var[which.min(df$sum)],
+  cex = 1.2, pch = 24, bg = "white", lwd = 1.7
+)
+
+# Illustrate best partition wrt simililarity combined similarity / variance
+points(
+  all_objectives_mean[which.min(df$diff_var)],
+  all_objectives_var[which.min(df$diff_var)],
+  cex = 1.2, pch = 23, bg = "white", lwd = 1.7
+)
+```
 
@@
 
-- formalize the weighted-sum objective and use the weighted-sum method.
-works well on this specific problem [although in general there are 
-caveats to the weighted-sum method, see @brusco2019]. is fast, good 
-results. in fact, the objectives seem to work well together (optimizing 
-k-means extended often yields more similar variances than just 
-k-variances!!!)
+```{r echo = FALSE}
+# From Wikipedia on Multi-objective optimization:
+
+# Without additional subjective preference information, all Pareto optimal 
+# solutions are considered equally good. Researchers study multi-objective 
+# optimization problems from different viewpoints and, thus, there exist 
+# different solution philosophies and goals when setting and solving them. 
+# The goal may be to find a representative set of Pareto optimal 
+# solutions, and/or quantify the trade-offs in satisfying the different 
+# objectives, and/or finding a single solution that satisfies the 
+# subjective preferences of a human decision maker (DM).
+
+# The most preferred results can be found using different philosophies. 
+# Multi-objective optimization methods can be divided into four 
+# classes. In so-called no preference methods, no DM is expected to be 
+# available, but a neutral compromise solution is identified without 
+# preference information. The other classes are so-called a priori, a 
+# posteriori and interactive methods and they all involve preference 
+# information from the DM in different ways.
+
+# Solving a multi-objective optimization problem is sometimes understood 
+# as approximating or computing all or a representative set of Pareto 
+# optimal solutions.
+
+# Scalarizing a multi-objective optimization problem is an a priori 
+# method, which means formulating a single-objective optimization problem 
+# such that optimal solutions to the single-objective optimization problem 
+# are Pareto optimal solutions to the multi-objective optimization 
+# problem. In addition, it is often required that every Pareto optimal 
+# solution can be reached with some parameters of the scalarization.
+
+```
+
+- pragmatic solution [easy and practical approach, @marler2010; 
+@naidu2014]: formalize the weighted-sum objective and use the 
+weighted-sum approach. Then check out whether the simplest version is 
+sufficient in this case (but see the problems of weighted sum approach 
+in several contributions by Brusco et al.). Refer to @marler2010 to 
+argue for weighted sum for the current problem
+
+- it should be noted that both mathematical objectives capture different 
+aspects of an underlying substantive objective, that is to maximize 
+similarity of data distributions between different groups
+
+- Weighted sum approach works well on this specific problem [although in 
+general there are caveats to the weighted-sum method, see @brusco2019]. 
+is fast, good results. in fact, the objectives seem to work well 
+together
+
+- weighted sum approach generates a single solution from the 
+(approximated) pareto set
 
@@
 
-- Future research: Application of more sophisticated bicriterion 
-optimization algorithms are interesting
+### Weight selection
 
@@
 
+- Preference to weight [e.g., @marler2010]. However, the relationship 
+between weights and outcome is not always straightforward and the 
+results may disappoint. On the one hand, vastly different weights need 
+not lead to different solutions; and even equal weights can be 
+sufficient if the solution space offers partitions that maximize both 
+objectives at the same time. "the solution may not preserve one’s
+initial preferences no matter how the weights are set"; "even if one 
+determines acceptable values for the weights a priori, the final 
+solution may not accurately reflect initial preferences" [@marler2010, 
+p. 860].
+
 - Standardization of the extended data matrix to ensure that the two
-criterion functions do no differ markedly in scale [@brusco2019]
+criterion functions "do not differ markedly in scale" [@brusco2019]. 
+See Figure 1: the objectives do differ in scale, because k-variance 
+is based on a quadratic distance and therefore has larger values.
 
@@
 
+# Discussion
+
+- Focus on presentation of new objective function, the bicriterion 
+optimization process might be improved [e.g., by using a direct 
+algorithm as @brusco2019 did]
+
+- Future research: Application of more sophisticated bicriterion 
+optimization algorithms are interesting
+
+- new criterion: optimizes between-group similarity, no correspondence 
+to within-group heterogeneity [see @brusco2019 for a exceptional method 
+maximizing within-group heterogeneity]
+
+- Similarly: the reversed objective no longer has a useful clustering 
+interpretation: Maximizing differences in variances does not lead to 
+homogeneous and/or well separated clusters
+
 \newpage

### 2020-07-16 misc-functions.R
added pareto set finder to misc-functions

@@ -50,3 +50,32 @@ diff_mean <- function(clusters, features) {
 var_objective <- function(clusters, features) {
   anticlust::variance_objective(features, clusters)
 }
+
+
+# For a data frame (rows = partitions, columns = objectives), determine which 
+# elements are part of the Pareto efficient set. Output is a logical vector,
+# illustrating for each row if it is part of the Pareto efficient set.
+# `sense` can be "max" or "min", depending on whether high objective values
+# are good or low objective values. 
+pareto_set <- function(df, sense = "max") {
+  if (sense == "min") {
+    df <- -df
+  }
+  df <- data.frame(df)
+  ncrits <- ncol(df)
+  df$ID <- 1:nrow(df)
+  df <- df[order(rowSums(df[, 1:ncrits]), decreasing = TRUE), ]
+  
+  is_dominated <- rep(FALSE, nrow(df)) 
+  for (i in nrow(df):1) {
+    for (j in 1:nrow(df)) {
+      j_dominates_i <- all(df[i, 1:ncrits] <= df[j, 1:ncrits]) && 
+                       any(df[i, 1:ncrits] < df[j, 1:ncrits])
+      if (j_dominates_i) {
+        is_dominated[df$ID[i]] <- TRUE
+        break
+      }
+    }
+  }
+  !is_dominated
+}

### 2020-07-16 paper.Rmd
added footnote

@@
 correspondence between these differences and $SSE$ and 
-$SSE_{\mathrm{VAR}}$, respectively (Figure 1). Figure 2 illustrates 
-similarity in means plotted against similarity in variances across all 
-partitions. Ideally, a partition is chosen that minimizes the 
-difference with regard to means and variances (see the triangular 
+$SSE_{\mathrm{VAR}}$, respectively (Figure 1).[^onlyonegroup] Figure 2 
+illustrates similarity in means plotted against similarity in variances 
+across all partitions. Ideally, a partition is chosen that minimizes 
+the difference with regard to means and variances (see the triangular 
 symbol). However, optimal k-means anticlustering would create two groups 
@@
 other members of the optimal pareto set.
+
+[^onlyonegroup]: Note that this such a correspondence no longer holds 
+when there are more than two groups or there is more than one 
+numeric attribute (i.e., if $K > 2$ or $P > 1$).
 

### 2020-07-16 paper.Rmd
using one plot to illustrate the criteria in the example instead of two

@@
 ```{r, fig.cap = "Illustrates the 1-to-1 relationship between SSE and difference in means (Panel A) and variance (Panel B).
-Each data point illustrates a partition, i.e., a way to divide a data set ($N = 14$) into two groups.", fig.height = 3.5}
+Each data point illustrates a partition, i.e., a way to divide a data set ($N = 14$) into two groups.
+In Panel C, the relationship between the two criteria across partitions is plotted.
+The highlighted symbols indicate membership of the pareto set.
+The partition represented by the diamond minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the triangular minimizes the sum of the differences in means and variances.
+", fig.height = 7, fig.width = 6}
 
@@
 # Plot relationship between objectives
-par(mfrow = c(1, 2))
+layout(matrix(c(1,2,3,3), 2, 2, byrow = TRUE), heights = c(1, 1.4))
 plot(
@@
 legend("topright", legend = "B", cex = 1.3, bty = "n")
-```
 
-```{r, fig.width = 4.5, fig.height = 4.5, fig.cap = "All partitions. Points = pareto efficient set."}
+par(mar = c(7, 10, 0.4, 8) + 0.1)
 
@@
 )
+legend("topright", legend = "C", cex = 1.3, bty = "n")
 

### 2020-10-06 paper.Rmd
storing data to save time during rendering

@@
 source("misc-functions.R")
+set.seed(2207) # 2207
 
@@
 For the purpose of problem formalization, I adopt the notation Steinley 
-(2006) provided in his thorough synthesis of a half-century literature 
+(2006) provided in his comprehensive synthesis of the literature 
 on k-means. K-means clustering is used to partition two-way, two-mode 
@@
 
+# First rendering is slow because the data needs to be generated
+path <- "partition_objectives.csv"
+
+if (!file.exists(path)) {
   # Generate data and all partitions
-
-set.seed(2207) # 2207
-
   N <- 14
@@
 
+  # For each partition: Compute SSE objective, and SSE(VAR)
+  # SSE
+  all_objectives_kMeans <- sapply(
+    partitions,
+    FUN = var_objective,
+    features = features
+  )
+
+  # SSE(VAR)
+  all_objectives_kVar <- sapply(
+    partitions,
+    FUN = var_objective,
+    features = squared_from_mean(as.matrix(features))
+  )
+
   df <- data.frame(
     diff_mean = all_objectives_mean,
-  diff_var  = all_objectives_var
+    diff_var  = all_objectives_var,
+    kmeans = all_objectives_kMeans,
+    kvar = all_objectives_kVar
   )
+
+  write.table(
+    df, path, row.names = FALSE, sep = ","
+  )
+} else {
+  df <- read.csv(path)
+}
 
@@
 univariate normal distribution ($M = 15$, $SD = 3$) and generated all `r 
-length(partitions)` possible ways to partition the 14 data points into 
+n_partitions(14, 2)` possible ways to partition the 14 data points into 
 $K = 2$ equal-sized groups. I computed the difference in the two means 
@@
 
-# For each partition: Compute SSE objective, and SSE(VAR)
-# SSE
-all_objectives_kMeans <- sapply(
-  partitions,
-  FUN = var_objective,
-  features = features
-)
-
-# SSE(VAR)
-all_objectives_kVar <- sapply(
-  partitions,
-  FUN = var_objective,
-  features = squared_from_mean(as.matrix(features))
-)
-
 # Plot relationship between objectives
@@
 plot(
-  all_objectives_mean, 
-  all_objectives_kMeans, 
+  df[, c("diff_mean", "kmeans")],
   xlab = "mean(f1) - mean(f2)",
@@
 plot(
-  all_objectives_var, 
-  all_objectives_kVar, 
+  df[, c("diff_var", "kvar")],
   xlab = "var(f1) - var(f2)",
@@
 plot(
-  df, pch = 4, col = "darkgrey", 
+  df[, c("kmeans", "kvar")], 
+  pch = 4, col = "darkgrey", 
   las = 1, cex = .5,
-  xlab = "mean(f1) - mean(f2)",
-  ylab = "var(f1) - var(f2)"
+  xlab = "kmeans",
+  ylab = "kvar"
 )
-legend("topright", legend = "C", cex = 1.3, bty = "n")
+legend("bottomleft", legend = "C", cex = 1.3, bty = "n")
 
 # Illustrate Pareto set
-pareto_pts <- pareto_set(df, sense = "min")
-points(df[pareto_pts, ], pch = 19)
+#pareto_pts <- pareto_set(df[, c("kmeans", "kvar")])
+#points(df[pareto_pts, c("kmeans", "kvar")], pch = 19)
 
-# Illustrate best partition wrt SSE (difference in means)
+# Illustrate best partition wrt difference in means
 points(
-  all_objectives_mean[which.min(all_objectives_mean)],
-  all_objectives_var[which.min(all_objectives_mean)],
+  df$kmeans[which.max(df$kmeans)],
+  df$kvar[which.max(df$kmeans)],
   cex = 1.2, pch = 22, bg = "white", lwd = 1.7
@@
 
-df$sum <- rowSums(df)
-
-# Illustrate best partition wrt SSE(VAR) (i.e., difference in variance)
+# Illustrate best partition wrt difference in variance
 points(
-  all_objectives_mean[which.min(df$sum)],
-  all_objectives_var[which.min(df$sum)],
+  df$kmeans[which.max(df$kvar)],
+  df$kvar[which.max(df$kvar)],
   cex = 1.2, pch = 24, bg = "white", lwd = 1.7
@@
 
+df$kplus <- rowSums(df[, c("kmeans", "kvar")])
+
 # Illustrate best partition wrt simililarity combined similarity / variance
 points(
-  all_objectives_mean[which.min(df$diff_var)],
-  all_objectives_var[which.min(df$diff_var)],
+  df$kmeans[which.max(df$kplus)],
+  df$kvar[which.max(df$kplus)],
   cex = 1.2, pch = 23, bg = "white", lwd = 1.7

### 2020-10-06 scripts/K-Means-Vs-Difference.R
moved test scripts to new directory

@@ -9,7 +9,7 @@
 
 library(anticlust)
 source("exchange_method.R")
-source("misc-functions.R")
+source("../misc-functions.R")
 
 N <- 15
 M <- 3

### 2020-10-06 scripts/diff-min-max-mean-variance.R

@@ -9,7 +9,7 @@
 
 library(anticlust)
 source("exchange_method.R")
-source("misc-functions.R")
+source("../misc-functions.R")
 
 N <- 200
 M <- 4

### 2020-10-06 scripts/pareto_front.R

@@ -3,7 +3,7 @@
 # optimizing between-group similarity in variance
 
 library(anticlust)
-source("misc-functions.R")
+source("../misc-functions.R")
 
 N <- 12
 K <- 3

### 2020-10-06 scripts/solution-structure-large-data.R

@@ -3,7 +3,7 @@
 # optimizing between-group similarity in variance
 
 source("exchange_method.R")
-source("misc-functions.R")
+source("../misc-functions.R")
 
 
 library(anticlust)

### 2020-10-06 paper.Rmd
minor

@@
 
-For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, 
-k-means clustering aims to minimize the error sum of squares (SSE)---the 
-within-group variance---which is the sum of the squared Euclidean 
-distances between each data point and its cluster centroid. Conversely, 
-k-means anticlustering aims to maximize the same objective. The cluster 
-centroid is given by averaging the values on each variable over the 
-elements within that cluster. The centroid of the *j*th variable in the 
-*k*th cluster is given as:
+For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means 
+clustering aims to minimize the error sum of squares (SSE), also known as the 
+within-group variance. The SSE is the sum of the squared Euclidean distances 
+between each data point and its cluster centroid. The cluster centroid is given 
+by averaging the values on each variable over the elements within that cluster, 
+where the centroid $\overline{x}_j ^{(k)}$ of the *j*th variable in the *k*th cluster is
 
@@
 
-The k-means objective function is then given as
+The SSE is then given as:
 
@@
 
-If the $SSE$ is maximal---which is the goal of k-means 
-anticlustering---the cluster centroids are as close as possible to the 
+If the $SSE$ is maximal, the cluster centroids are as close as possible to the 
 overall centroid $\overline{\mathbf{x}}$ and therefore to each other 
 [@spath1986]. Thus, k-means anticlustering directly optimizes the 
-similarity of the mean values of the $P$ attributes across the $K$ 
-clusters.
+similarity of the mean values of the attributes across the different clusters.
 
@@
 
-**This is old, adjust to Steinley (2006):**
+K-means anticlustering, i.e., maximizing SSE minimizes differences between the 
+means of the input variables. I present a rather simple extension of the SSE 
+criterion to simultaneously minimize differences with regard to the variances 
+of the input variables. The variance is defined as the sum of the squared 
+distances between all
 
@@
 )
-legend("topright", legend = "A", cex = 1.3, bty = "n")
+legend("bottomleft", legend = "A", cex = 1.3, bty = "n")
 plot(
@@
 )
-legend("topright", legend = "B", cex = 1.3, bty = "n")
+legend("bottomleft", legend = "B", cex = 1.3, bty = "n")
 

### 2020-10-07 paper.Rmd
working on problem formalization

@@
 ---
-title             : "A bicriterion extension of k-means anticlustering to optimize between-group similarity"
+title             : "A bicriterion extension of k-means anticlustering"
 shorttitle        : "Bicriterion K-means anticlustering"
@@
 For the purpose of problem formalization, I adopt the notation Steinley 
-(2006) provided in his comprehensive synthesis of the literature 
-on k-means. K-means clustering is used to partition two-way, two-mode 
-data (that is, $N$ elements each having $P$ attributes) into $K$ groups 
-$(C_1., \ldots, C_K)$ where $C_k$ is the set of $n_k$ objects in group 
+(2006) provided in his comprehensive synthesis of half a century of research on 
+k-means. K-means clustering is used to partition two-way, two-mode 
+data. That is, $N$ elements each having $P$ attributes into $K$ groups 
+$(C_1., \ldots, C_K)$, $C_k$ being the set of $n_k$ objects in group 
 $k$. The partitioning is subject to the requirements that each element 
@@
 
-In this paper, I will discuss only such applications where groups are of 
-equal size.
-
 For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means 
 clustering aims to minimize the error sum of squares (SSE), also known as the 
-within-group variance. The SSE is the sum of the squared Euclidean distances 
-between each data point and its cluster centroid. The cluster centroid is given 
-by averaging the values on each variable over the elements within that cluster, 
-where the centroid $\overline{x}_j ^{(k)}$ of the *j*th variable in the *k*th cluster is
+within-group variance [@spath1986]. The SSE is the sum of the squared Euclidean 
+distances between each data point and its cluster centroid; the *k*'th centroid 
+is a vector of length $P$, where each entry is the mean value of one of the $P$ 
+attributes across the observations belonging the *k*'th cluster ($k = 1, \ldots, 
+K$). The centroid of the *k*'th cluster has the notation 
+$\overline{\mathbf{x}}^{(k)}$ with
+$$
+\overline{\mathbf{x}}^{(k)} = 
+  (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})
+$$
+
+The mean $\overline{x}_j ^{(k)}$ of the *j*th variable on the *k*th cluster is 
+computed as:
 
@@
 
-The SSE is then given as:
+Based on the original data and the cluster centroids, the SSE---i.e., the k-means 
+objective function---is given as:
 
@@
 
-Note there is a direct connection between $SSE$ and the cluster 
-centroids [@spath1986]. If the complete centroid vector for cluster 
-$C_k$ is given by
-
-$$
-\overline{\mathbf{x}}^{(k)} = 
-  (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})
-$$
-
-and the $SSE$ is minimal, the centroid values 
+Note there is a direct connection between the $SSE$ and the location of the 
+cluster centroids [@spath1986]. If the $SSE$ is minimal, the centroid values 
 $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as 
@@
 [@spath1986]. Thus, k-means anticlustering directly optimizes the 
-similarity of the mean values of the attributes across the different clusters.
+similarity of the mean values of the $P$ attributes across the $K$ clusters.
 
-## K-Variances
+## Bi-criterion extension
 
-K-means anticlustering, i.e., maximizing SSE minimizes differences between the 
-means of the input variables. I present a rather simple extension of the SSE 
-criterion to simultaneously minimize differences with regard to the variances 
-of the input variables. The variance is defined as the sum of the squared 
-distances between all
+K-means anticlustering---i.e., maximizing the SSE---minimizes differences 
+between the means of the input variables. In the following, a rather 
+straight-forward extension of the SSE criterion is formulated that can be used 
+to minimize differences with regard to the variances of the input variables at 
+the same time.
 
-The squared distance between an element $x = (x_1, \ldots, x_M)$ and the 
-average of all elements $\overline{\mathrm{\boldsymbol{x}}}$ is:
+For the sake of simplicity, we first consider a set of $N$ one-dimensional 
+observations $(x_1, \ldots, x_N)$. The extension to the general case of $P$ 
+attributes will be straight forward, because the formulation of the new 
+criterion effectively reduces to the generation of new variables, leaving the 
+k-means criterion SSE per sé unchanged.
+
+The sample variance of this set of 
+observations is computed as the mean of the squared distances between 
+observations and their mean $\mathbf{x}$:
 
 $$
-\vert \vert x - \overline{\mathrm{\boldsymbol{x}}} \vert \vert^2 = 
-  \sum\limits_{m = 1}^{M}(x_{m} - \overline{\mathrm{\boldsymbol{x_m}}})^2
+\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^2
 $$
 
-Now define new data points representing the squared distances between 
-data points and the grand mean:
+We next define a new variable $y = (y_1, \ldots, y_N)$, representing the squared 
+deviation of each observation from the mean:
 
 $$
-y_i = (x_i - \mathrm{\boldsymbol{x}})^2, \forall i \in N
+y_i = (x_i - \overline{\mathbf{x}})^2
 $$
 
-where $\mathrm{\boldsymbol{x}}$ is a vector of length $M$ containing 
-the average for each feature.
-
-Now:
+The sample variance of $x$ is consequently the mean of these new values:
 
 $$
-\sum\limits_{m = 1}^{M}(x_{m} - \overline{\mathrm{\boldsymbol{x_m}}})^2 + 
-  \sum\limits_{m = 1}^{M}(y_{m} - \overline{\mathrm{\boldsymbol{y_m}}})^2 =
-  \sum\limits_{m = 1}^{M}(x_{m} - \overline{\mathrm{\boldsymbol{x_m}}})^2 + 
-    (y_{m} - \overline{\mathrm{\boldsymbol{y_m}}})^2
+\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} y_i
 $$
 
-If we define $x_{m + 1} = y_1$, $x_{m + 2} = y_2$, ..., $x_{2M} = y_M$.
-this is the same as:
+Note at this point that the sample variance is basically defined as a mean 
+value. We can use this insight and employ k-means anticlustering to equalize 
+this *mean* across groups. For a given $K$, groups can be created that are 
+similar with regard to the means of $y$---i.e., the variance of $x$---by 
+maximizing the SSE, in this one-dimensional case given as:
 
 $$
-\sum\limits_{m = 1}^{2M}(x_{m} - \overline{\mathrm{\boldsymbol{x_m}}})^2 
+\mathit{SSE} = 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (y_{i} - \overline{y} ^{(k)})^2
 $$
 
-Thus, we redefine the data points and include an additional $M$ features 
-for each element. Then, we maximize $\mathit{Var}_{\mathrm{within}}$, 
-which corresponds to a tradeoff between minimizing differences in mean 
-features between clusters, and the average squared distance to the grand 
-mean, i.e., the within-cluster variance. That is: We try to maximize the 
-within-cluster variance, but to a certain degree we also strive to 
-obtain similar within-cluster variance in each cluster.
+## Weighted-sum approach
+
+$$
+\mathit{SSE_{comb}} = 
+  w_1
+  \sum\limits_{k=1}^{K} 
+    \sum\limits_{i \in C_k}
+    (x_{i} - \overline{x} ^{(k)})^2 + 
+  w_2
+  \sum\limits_{k=1}^{K} 
+    \sum\limits_{i \in C_k}
+    (y_{i} - \overline{y} ^{(k)})^2
+$$
+
+The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of 
+equalizing means and variances, respectively.
+
+## Unweighted case
+
+Some simplifications can be arranged when ignoring the weights on the two 
+criteria.
+
+$$
+\mathit{SSE_{comb}} = 
+  \sum\limits_{k=1}^{K} 
+    \sum\limits_{i \in C_k}
+    (x_{i} - \overline{x} ^{(k)})^2 + 
+  \sum\limits_{k=1}^{K} 
+    \sum\limits_{i \in C_k}
+    (y_{i} - \overline{y} ^{(k)})^2 \\
+  = \sum\limits_{k=1}^{K} 
+     \sum\limits_{i \in C_k}
+     ((x_{i} - \overline{x} ^{(k)})^2 + 
+     (y_{i} - \overline{y} ^{(k)})^2)
+$$
+
+When defining $z = (x, y)$, the SSE reduces to the standard case in (X):
+
+$$
+\mathit{SSE} = 
+  \sum\limits_{j=1}^{2} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (z_{ij} - \overline{z}_j ^{(k)})^2
+$$
+
+This logic can be extended to the general case of $P$ attributes; for each 
+attribute, a new variable is created as described above, and the entirety of 
+all $2P$ variables is then used as input to standard k-means anticlustering.
 
@@
 
+- extending on covariances?
+
 \newpage

### 2020-10-07 paper.Rmd
more work on formalization

@@
 $$
+
 $$
@@
 within-group variance [@spath1986]. The SSE is the sum of the squared Euclidean 
-distances between each data point and its cluster centroid; the *k*'th centroid 
+distances between each data point and its cluster centroid. The *k*'th centroid 
 is a vector of length $P$, where each entry is the mean value of one of the $P$ 
-attributes across the observations belonging the *k*'th cluster ($k = 1, \ldots, 
-K$). The centroid of the *k*'th cluster has the notation 
-$\overline{\mathbf{x}}^{(k)}$ with
-$$
-\overline{\mathbf{x}}^{(k)} = 
-  (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})
-$$
-
-The mean $\overline{x}_j ^{(k)}$ of the *j*th variable on the *k*th cluster is 
+attributes, computed across the observations belonging the *k*'th cluster ($k = 
+1, \ldots, K$). The centroid of the *k*'th cluster $\overline{\mathbf{x}}^{(k)} 
+= (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is 
 computed as:
@@
 
-Based on the original data and the cluster centroids, the SSE---i.e., the k-means 
-objective function---is given as:
+Based on the data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, 
+a partitioning of the data into $K$ clusters, the cluster centroids, the 
+SSE---i.e., the k-means objective function---is then given as:
 
@@
 between the means of the input variables. In the following, a rather 
-straight-forward extension of the SSE criterion is formulated that can be used 
-to minimize differences with regard to the variances of the input variables at 
-the same time.
+straight-forward bicriterion extension of the SSE criterion is formulated that 
+can be used to minimize differences with regard to the variances of the input 
+variables at the same time.
 
-For the sake of simplicity, we first consider a set of $N$ one-dimensional 
-observations $(x_1, \ldots, x_N)$. The extension to the general case of $P$ 
-attributes will be straight forward, because the formulation of the new 
-criterion effectively reduces to the generation of new variables, leaving the 
-k-means criterion SSE per sé unchanged.
+For the sake of simplicity, we first consider a set of one-dimensional 
+observations $(x_1, \ldots, x_N)$ to motivate the new criterion. The extension 
+to the general case of $P$ attributes will be straight forward, because the 
+formulation effectively reduces to the generation of new variables, whereas the 
+original k-means criterion remains unchanged.
 
-The sample variance of this set of 
-observations is computed as the mean of the squared distances between 
-observations and their mean $\mathbf{x}$:
+The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is 
+computed as the mean of the squared distances between observations and their 
+mean $\mathbf{x}$:
 
@@
 
-Note at this point that the sample variance is basically defined as a mean 
-value. We can use this insight and employ k-means anticlustering to equalize 
-this *mean* across groups. For a given $K$, groups can be created that are 
-similar with regard to the means of $y$---i.e., the variance of $x$---by 
-maximizing the SSE, in this one-dimensional case given as:
+We can use the insight that the sample variance is basically defined as an 
+average, implying that k-means anticlustering can be employed to equalize this 
+average across different groups. This is the central insight leading to the new 
+criterion. In the one-dimensional case as discussed here, the SSE that 
+represents discrepancies with regard to the variance, is given as:
 
 $$
-\mathit{SSE} = 
+\mathit{SSE_{Var}} = 
   \sum\limits_{k=1}^{K} 
@@
 
-## Weighted-sum approach
+This criterion, reflecting discrepancies in variance across groups, is extended 
+to the general case of $P$ attributes as
 
 $$
-\mathit{SSE_{comb}} = 
-  w_1
+\mathit{SSE_{Var}} = 
+  \sum\limits_{j=1}^{P} 
   \sum\limits_{k=1}^{K} 
   \sum\limits_{i \in C_k}
-    (x_{i} - \overline{x} ^{(k)})^2 + 
-  w_2
+  (y_{ij} - \overline{y}_j ^{(k)})^2
+$$
+
+with
+
+$$
+\overline{y}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} y_{ij}
+$$
+
+and
+
+$$
+y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2.
+$$
+
+## Bicriterion formulation
+
+It will be rarely of interest to maximize $SSE_{Var}$ (equalizing variances) 
+without also considering the standard $SSE$ (i.e., equalizing means). A 
+combination of $SSE_{Var}$ and $SSE$, however, can be employed to simultaneously 
+optimize both objectives. That is, formulating a criterion incorporating both 
+discrepancies in means as well as variances to minimize difference with regard 
+to either. The most simple approach to such a bicriterion optimization is the 
+weighted sum approach. That is,
+
+Combining (X) and (Y), we can obtain a combined criterion formulation as a 
+weighted sum criterion.
+
+$$
+\mathit{SSE_{bicriterion}} = 
+  w_1
+  \sum\limits_{j=1}^{P} 
   \sum\limits_{k=1}^{K} 
   \sum\limits_{i \in C_k}
-    (y_{i} - \overline{y} ^{(k)})^2
+  (x_{ij} - \overline{x}_j ^{(k)})^2 + 
+  w_2
+  \sum\limits_{j=1}^{P} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (y_{ij} - \overline{y}_j ^{(k)})^2
 $$
@@
 $$
-\mathit{SSE_{comb}} = 
+\mathit{SSE_{bicriterion}} = 
   \sum\limits_{k=1}^{K} 
@@
 
-When defining $z = (x, y)$, the SSE reduces to the standard case in (X):
+When defining $z_i = (z_{i1}, z_{i2})$ with $z_{i1} = x_i$ and $z_{i2} = y_i$, 
+the bicriterion SSE is computed as
 
 $$
-\mathit{SSE} = 
+\mathit{SSE_{bicriterion}} =
   \sum\limits_{j=1}^{2} 

### 2021-03-04 paper.Rmd
finally working on paper again

--- 2020-10-07 paper.Rmd
@@
 ---
-title             : "A bicriterion extension of k-means anticlustering"
-shorttitle        : "Bicriterion K-means anticlustering"
+title             : "k-plus: a bicriterion extension for k-means anticlustering"
+shorttitle        : "k-plus anticlustering"
 
@@
 abstract: |
-    Anticlustering is an automated method to partition data sets into groups 
+
+    Anticlustering is a method to partition data sets into groups 
     in such a way that high between-group similarity and high within-group 
-    heterogeneity is obtained. Anticlustering has important applications in 
-    psychological research, for example to partition stimulus sets into 
-    equivalent sets, or to divide a group of students into heterogeneous 
-    classes. K-means anticlustering is maximizes the error sum of squares, 
-    i.e., the sum of squared Euclidean distances from each element to the 
-    center of the group to which it belongs. Thereby, k-means anticlustering 
-    directly maximizes the similarity of the mean values of numeric 
-    attributes across groups, but ignores the spread of the data. 
-    This paper presents a straight-forward extension of 
-    the k-means objective function that allows to simultaneously maximize 
-    similarity with regard to the mean as well as the variance across 
-    groups. It is shown that when both criteria receive equal weights, the 
-    bicriterion optimization problem reduces to an augmentation of the input 
-    data, leaving the original k-means criterion unchanged. Tests on 
-    simulated and real data show that similarity with regard to means and 
-    variance criteria can usually be achieved to an astonishing degree with 
-    almost no cost to either criterion. It is therefore concluded that the 
-    extended model should be preferred to classical k-means anticlustering 
-    for optimizing between-group similarity. An exception is given for 
-    binary data that does not profit from considering equality in 
-    variances.
+    heterogeneity is obtained. In psychological research, anticlustering is for example 
+    used to to divide stimulus sets into equivalent sets, to split tests into parallel 
+    halfs, or to divide a group of students into heterogeneous classes. Different 
+    objective functions can be optimized in anticlustering applications. K-means 
+    anticlustering maximizes the error sum of squares, i.e., the sum of squared 
+    Euclidean distances from each element to the center of the group to which it 
+    belongs. K-means anticlustering thereby directly maximizes the similarity of 
+    the mean values of numeric attributes across groups, but it ignores the spread of 
+    the data. This paper presents k-plus anticlustering, a straight-forward 
+    extension of the k-means objective function that allows to simultaneously 
+    maximize similarity with regard to the mean as well as the variance across 
+    groups. It is shown that the bicriterion optimization problem reduces to an 
+    augmentation of the input data, leaving the original k-means criterion 
+    unchanged. Tests on simulated and real data show that similarity with regard to 
+    means and variance criteria can usually be achieved to an astonishing degree 
+    with almost no cost to either criterion. It is therefore concluded that the 
+    extended model should be preferred to classical k-means anticlustering for 
+    optimizing between-group similarity. An exception is given for binary data that 
+    does not profit from considering equality in variances.
 
@@
 
-keywords          : "Anticlustering, variance, k-means, k-variance, bicriterion optimization"
+keywords          : "Anticlustering, k-means, k-plus, variance, bicriterion optimization"
 wordcount         : "X"
@@
 
-A different approach to anticlustering directly works with feature based 
-data and does not require a transformation into pairwise 
-dissimilarities. Instead, the sum of the squared Euclidean distances 
-between elements and the cluster they are assigned to is minimized 
-[@brusco2006branch], most prominently using the k-means heuristic 
-[@jain2010; @steinley2006]. Minimizing the k-means criterion 
+A different approach to anticlustering directly works with feature based data 
+and does not require a transformation into pairwise dissimilarities. Instead, 
+the sum of the squared Euclidean distances between elements and the cluster they 
+are assigned to is minimized [@brusco2006branch], most prominently using the 
+k-means heuristic [@jain2010; @steinley2006]. Minimizing the k-means criterion 
 simultaneously maximizes between-cluster separation and within-cluster 
 homogeneity [@aloise2009np]. Späth [-@spath1986] and Valev [-@valev1998; 
--@valev1983] independently recognized that by optimizing the reversed 
-k-means objective that is, maximizing the sum of distances between 
-elements and cluster centroids -- a data set is partitioned into similar 
-groups, thereby reversing the logic of classical clustering. They 
-independently from each other coined the term anticlustering for this 
-purpose. K-means anticlustering maximizes the sum of squared Euclidean 
-distances from each element to the center of the cluster to which it 
-belongs. As @spath1986 noted, k-means anticlustering thereby directly 
-minimizes differences with regard to the means of the numeric variables 
-between groups. Hence, groups only become similar with regard to the 
-mean of the distribution of the numeric attributes, which may perform 
-poorly as a measure of overall similarity; the spread of the data is not 
-targeted and may differ between sets. @papenberg2020 showed that 
-maximizing the diversity (based on the pairwise Euclidean distances) is 
-better suited to minimize difference with regard to both the mean and 
-the variance of the data; it was shown that even an entirely random 
-split usually resulted in more similar standard deviations as k-means 
-solely focuses on the mean. When requiring that groups are overall 
-similar to each other, this focus may be detrimental as similar means 
-can be obtained even when the underlying distributions are very 
-different [@anscombe1973].
+-@valev1983] independently recognized that by optimizing the reversed k-means 
+objective that is, maximizing the sum of distances between elements and cluster 
+centroids -- a data set is partitioned into similar groups, thereby reversing 
+the logic of classical clustering. They coined the term anticlustering for this 
+purpose. K-means anticlustering maximizes the sum of squared Euclidean distances 
+from each element to the center of the cluster to which it belongs. As 
+@spath1986 noted, k-means anticlustering thereby directly minimizes differences 
+with regard to the means of the numeric variables between groups. Hence, groups 
+only become similar with regard to the mean of the distribution of the numeric 
+attributes, which may perform poorly as a measure of overall similarity; the 
+spread of the data is not targeted and may differ between sets. @papenberg2020 
+showed that maximizing the diversity (based on the pairwise Euclidean distances) 
+is better suited to minimize difference with regard to both the mean and the 
+variance of the data; it was shown that even an entirely random split usually 
+resulted in more similar standard deviations as k-means solely focuses on the 
+mean. When requiring that groups are overall similar to each other, this focus 
+may be misguided as similar means can be obtained even when the underlying 
+distributions are different [@anscombe1973].
 
-In this paper, I present a simple extension of the k-means objective 
-function to maximize simililarity with regard to means and variances at 
-the same time. The aim is to establish groups that are overall similar 
-to each other. It is shown that this is achieved by formulating a 
-bicriterion optimization objective for k-means anticlustering. 
-Interestingly, under specific---and important---circumstances, this 
-reformulation reduces to an augmentation of the original data matrix 
-while the optimization objective remains unchanged. This insight has 
-important consequences as it allows to adapt standard optimization 
-schemes for single criterion anticlustering. I show that the extended 
-k-means criterion outperforms other anticlustering criteria when the 
-goal is the minimize difference with regard to the mean and variance of 
-the different groups. In most cases, both objectives can be optimized
-satisfactorily at the same time without any loss on either; when the 
-data set is large enough (usually, when $N > XX$), the bicriterion 
-optimization leads to > 99.99% objective performance as compared to the 
-single optimization of either k-means or k-variance.
+In this paper, I present k-plus anticlustering, a simple extension of the 
+k-means objective function to maximize similarity with regard to means and 
+variances at the same time. The aim is to establish groups that are overall 
+similar to each other. It is shown that this is achieved by formulating a 
+bicriterion optimization objective for k-means anticlustering. Interestingly, 
+this reformulation reduces to an augmentation of the original data matrix while 
+the optimization objective remains unchanged. This insight has important 
+consequences as it allows to adapt standard optimization schemes for single 
+criterion anticlustering. I show that the k-plus criterion outperforms other 
+anticlustering criteria when the goal is the minimize difference with regard to 
+the mean and variance of the different groups. In most cases, both objectives 
+can be optimized satisfactorily at the same time without any loss on either; 
+when the data set is large enough (usually, when $N > XX$), the bicriterion 
+optimization leads to > 99.99% objective performance as compared to the single 
+optimization of either k-means or k-variance.
 
@@
 
-For the purpose of problem formalization, I adopt the notation Steinley 
-(2006) provided in his comprehensive synthesis of half a century of research on 
-k-means. K-means clustering is used to partition two-way, two-mode 
-data. That is, $N$ elements each having $P$ attributes into $K$ groups 
-$(C_1., \ldots, C_K)$, $C_k$ being the set of $n_k$ objects in group 
-$k$. The partitioning is subject to the requirements that each element 
-is assigned to exactly one group, formalized by the two constraints
+For the purpose of problem formalization, I adopt the notation Steinley (2006) 
+provided in his comprehensive synthesis of half a century of research on 
+k-means. K-means clustering is used to partition two-way, two-mode data into 
+groups. That is, $N$ elements each having $P$ attributes are assigned to $K$ 
+groups $(C_1., \ldots, C_K)$, $C_k$ being the set of $n_k$ objects in group $k$. 
+The partitioning is subject to the requirements that each element is assigned to 
+exactly one group, formalized by the two constraints
 
@@
 
-In anticlustering applications, we usually impose restrictions on 
-the group sizes $n_k$, with the most common restriction being that all 
-groups have equal size:
+In anticlustering applications, we  impose an additional constraint on the group 
+sizes $n_k$, where the most common one requires that all groups have equal size:
 
@@
 For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means 
-clustering aims to minimize the error sum of squares (SSE), also known as the 
-within-group variance [@spath1986]. The SSE is the sum of the squared Euclidean 
-distances between each data point and its cluster centroid; the *k*'th centroid 
-is a vector of length $P$, where each entry is the mean value of one of the $P$ 
-attributes across the observations belonging the *k*'th cluster ($k = 1, \ldots, 
-K$). The centroid of the *k*'th cluster has the notation 
-$\overline{\mathbf{x}}^{(k)}$ with
-$$
-\overline{\mathbf{x}}^{(k)} = 
-  (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})
-$$
-
-The mean $\overline{x}_j ^{(k)}$ of the *j*th variable on the *k*th cluster is 
-computed as:
-
-$$
-\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} x_{ij}
-$$
-
-Based on the original data and the cluster centroids, the SSE---i.e., the k-means 
-objective function---is given as:
+clustering aims to minimize the error sum of squares (SSE), which is the sum of 
+the squared Euclidean distances between each data point and its cluster 
+centroid:
 
@@
 
+The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, 
+\overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, 
+where each entry is the mean value of one of the $P$ attributes, computed across 
+all observations belonging the $k$th cluster $C_k$ ($k = 1, \ldots, K$):
+
+$$
+\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} x_{ij}
+$$
+
 Note there is a direct connection between the $SSE$ and the location of the 
 cluster centroids [@spath1986]. If the $SSE$ is minimal, the centroid values 
-$\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as 
-possible from the overall data centroid $\overline{\mathbf{x}} = 
+$\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible 
+from the overall data centroid $\overline{\mathbf{x}} = 
 (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, 
-\overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the 
-overall mean value on the $p$th attribute:
+\overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the overall mean 
+value on the $p$th attribute:
 
@@
 [@spath1986]. Thus, k-means anticlustering directly optimizes the 
-similarity of the mean values of the $P$ attributes across the $K$ clusters.
+similarity of the mean attribute values across clusters.
 
-## Bi-criterion extension
+## A criterion to quantify differences in variances
 
 K-means anticlustering---i.e., maximizing the SSE---minimizes differences 
-between the means of the input variables. In the following, a rather 
-straight-forward extension of the SSE criterion is formulated that can be used 
-to minimize differences with regard to the variances of the input variables at 
-the same time.
+between the means of the input variables. I now present a rather 
+straight-forward adaption of the SSE to capture differences in variance between 
+groups. To motivate the new criterion, we first consider a set of 
+one-dimensional observations $(x_1, \ldots, x_N)$. The extension to the general 
+case of $P$ attributes will be straight forward
 
-For the sake of simplicity, we first consider a set of $N$ one-dimensional 
-observations $(x_1, \ldots, x_N)$. The extension to the general case of $P$ 
-attributes will be straight forward, because the formulation of the new 
-criterion effectively reduces to the generation of new variables, leaving the 
-k-means criterion SSE per sé unchanged.
-
-The sample variance of this set of 
-observations is computed as the mean of the squared distances between 
-observations and their mean $\mathbf{x}$:
+The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is 
+computed as the mean of the squared distances between observations and their 
+mean $\overline{\mathbf{x}}$:
 
@@
 
-Note at this point that the sample variance is basically defined as a mean 
-value. We can use this insight and employ k-means anticlustering to equalize 
-this *mean* across groups. For a given $K$, groups can be created that are 
-similar with regard to the means of $y$---i.e., the variance of $x$---by 
-maximizing the SSE, in this one-dimensional case given as:
+Since the sample variance is basically defined as an average, k-means 
+anticlustering can be employed to equalize the variance across groups. That is, 
+we can formulate an adaption of the SSE that quantifies the degree to which 
+variances differ between groups:
 
 $$
-\mathit{SSE} = 
+\mathit{SSE_{Var}} = 
   \sum\limits_{k=1}^{K} 
@@
 
-## Weighted-sum approach
+Maximizing $\mathit{SSE_{Var}}$ will equalize the variance between groups on one 
+attribute. The criterion is extended to the general case of $P$ attributes as
 
 $$
-\mathit{SSE_{comb}} = 
-  w_1
+\mathit{SSE_{Var}} = 
+  \sum\limits_{j=1}^{P} 
   \sum\limits_{k=1}^{K} 
   \sum\limits_{i \in C_k}
-    (x_{i} - \overline{x} ^{(k)})^2 + 
-  w_2
+  (y_{ij} - \overline{y}_j ^{(k)})^2
+$$
+
+where
+
+$$
+\overline{y}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} y_{ij}
+$$
+
+and
+
+$$
+y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2.
+$$
+
+Note that $\mathit{SSE_{Var}}$ has the same form as the standard $\mathit{SSE}$. 
+However, the input data reflect the squared difference between each data point 
+and the average values instead of the raw data points.
+
+## A bicriterion model: k-plus anticlustering
+
+In most applications, it will not be of interest to only maximize $SSE_{Var}$ 
+(i.e., equalizing variances) without also considering the standard $SSE$ (i.e., 
+equalizing means). A combination of $SSE_{Var}$ and $SSE$, however, can be 
+employed to simultaneously optimize both objectives. The most simple approach to 
+such a bicriterion optimization is the weighted sum approach [@naidu2014]. That 
+is, both criteria are computed independently from each other and then combined 
+into a single criterion through a weighted sum:
+
+$$
+\mathit{SSE_{kplus}} = w_1 
+  \sum\limits_{j=1}^{P} 
   \sum\limits_{k=1}^{K} 
   \sum\limits_{i \in C_k}
-    (y_{i} - \overline{y} ^{(k)})^2
+  (y_{ij} - \overline{x}_j ^{(k)})^2 + 
+  w_2 \sum\limits_{j=1}^{P} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (x_{ij} - \overline{y}_j ^{(k)})^2
 $$
@@
 The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of 
-equalizing means and variances, respectively.
+equalizing means and variances, respectively. Often, they are chosen such that 
+$w_1 + w_2 = 1$ [@brusco2019].
 
-## Unweighted case
-
-Some simplifications can be arranged when ignoring the weights on the two 
-criteria.
+Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$ 
+(i.e., or setting them to the same value). When defining $z_i = (x_{i1}, \ldots, 
+x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and 
+$y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ is computed as
 
 $$
-\mathit{SSE_{comb}} = 
-  \sum\limits_{k=1}^{K} 
-    \sum\limits_{i \in C_k}
-    (x_{i} - \overline{x} ^{(k)})^2 + 
-  \sum\limits_{k=1}^{K} 
-    \sum\limits_{i \in C_k}
-    (y_{i} - \overline{y} ^{(k)})^2 \\
-  = \sum\limits_{k=1}^{K} 
-     \sum\limits_{i \in C_k}
-     ((x_{i} - \overline{x} ^{(k)})^2 + 
-     (y_{i} - \overline{y} ^{(k)})^2)
-$$
-
-When defining $z = (x, y)$, the SSE reduces to the standard case in (X):
-
-$$
-\mathit{SSE} = 
-  \sum\limits_{j=1}^{2} 
+\mathit{SSE_{kplus}} =
+  \sum\limits_{j=1}^{2 P} 
   \sum\limits_{k=1}^{K} 
@@
 
-This logic can be extended to the general case of $P$ attributes; for each 
-attribute, a new variable is created as described above, and the entirety of 
-all $2P$ variables is then used as input to standard k-means anticlustering.
+This formula again has the same form as SSE (and $SSE_{Var}$), but is computed 
+on different data. In this case, both the original data $x_{ij}$ as well as the 
+squared distances $y_{ij}$ are used as attributes that enter the objective 
+function. That is, computing the k-plus criterion is done by first computing the 
+values $y_{ij}$, appending these to the original data matrix, and then compute 
+SSE on the augmented data matrix.
+
+Note that a minor problem arises when ignoring the weights during optimizing the 
+k-plus objective. Since their computation involves squaring, the squared 
+distances $y_{ij}$ have much larger values, on average, than the raw values 
+$x_{ij}$. Therefore, the $SSE_{Var}$ criterion naturally receives greater weight 
+when computing the k-plus criterion than the standard $SSE$ criterion. This 
+imbalance is unwarranted because optimizing similarity with regard to variances 
+is not more important than optimizing similarity with regard to means. A simple 
+solution for this issue is to standardize all values $x_{ij}$ and $y_{ij}$ to 
+the same scale before computing $SSE_{kplus}$, e.g. via $z$-standardization.
 
@@
 
-A motivating example illustrates that aiming to simultaneously optimize 
-similarity of means and variances at the same time is not only 
-desirable, but may also be feasible. I created 14 data points from a 
-univariate normal distribution ($M = 15$, $SD = 3$) and generated all `r 
-n_partitions(14, 2)` possible ways to partition the 14 data points into 
-$K = 2$ equal-sized groups. I computed the difference in the two means 
-and two variances for each partition. This simple difference metric was 
-chosen for purposes of interpretability; in fact, there is a a 1-to-1 
-correspondence between these differences and $SSE$ and 
-$SSE_{\mathrm{VAR}}$, respectively (Figure 1).[^onlyonegroup] Figure 2 
-illustrates similarity in means plotted against similarity in variances 
-across all partitions. Ideally, a partition is chosen that minimizes 
-the difference with regard to means and variances (see the triangular 
-symbol). However, optimal k-means anticlustering would create two groups 
-that are not particularly similar with regard to their variance and 
-chose the partition that is represented by the squared symbol. The 
-diamond symbol represents that partition that where the two groups have 
-the least difference in variances. The black symbols illustrate the 
-other members of the optimal pareto set.
+Optimizing a bicriterion objective such as $SSE_{kplus}$ may not succeed as one 
+may hope. The ideal outcome would be to find a partition where both the means 
+and the variances are very similar between groups. However, research on 
+multi-objective optimization generally shows that cases are rare where a single 
+solution satisfies all criteria to a satisfactory degree. Usually, a tradeoff is 
+inevitable. It is an open question whether k-plus anticlustering is a expection 
+from this rule. Next, a motivating example illustrates that in anticlustering, 
+aiming to simultaneously optimize similarity of means and variances is not only 
+desirable, but may actually be feasible. This notion is later confirmed when 
+applying k-plus anticlustering in simulation studies and examples using real 
+data.
+
+For this initial example, I created 14 data points from a univariate normal 
+distribution ($M = 15$, $SD = 3$) and generated all `r n_partitions(14, 2)` 
+possible ways to partition the 14 data points into $K = 2$ equal-sized groups. 
+For each partition, the criteria SSE and $SSE_{Var}$ were computed. Figure 1 
+shows the scatterplot of these criteria across partitions. Note that for both 
+criteria, a larger value is better, i.e., indicates higher similarity with 
+regard to means and variances, respectively. Ideally, to obtain overall 
+between-group similarity, a partition is chosen that minimizes the difference 
+with regard to both means and variances (see the triangular symbol). However, 
+optimal k-means anticlustering would create two groups that are not particularly 
+similar with regard to their variance. The partition that is represented by the 
+squared symbol has the overall best highest SSE, but is outperformed by many 
+other partitions with regard to $SSE_{Var}$. The diamond symbol represents 
+that partition that where the two groups have the least difference in 
+variances.
+
+Next: use the partition that optimizes $SSE_{kplus}$!
 
@@
 
-```{r, fig.cap = "Illustrates the 1-to-1 relationship between SSE and difference in means (Panel A) and variance (Panel B).
-Each data point illustrates a partition, i.e., a way to divide a data set ($N = 14$) into two groups.
-In Panel C, the relationship between the two criteria across partitions is plotted.
-The highlighted symbols indicate membership of the pareto set.
-The partition represented by the diamond minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the triangular minimizes the sum of the differences in means and variances.
-", fig.height = 7, fig.width = 6}
-
-# Plot relationship between objectives
-layout(matrix(c(1,2,3,3), 2, 2, byrow = TRUE), heights = c(1, 1.4))
-plot(
-  df[, c("diff_mean", "kmeans")],
-  xlab = "mean(f1) - mean(f2)",
-  ylab = "SSE",
-  pch = 4, 
-  cex = .5,
-  col = "darkgrey"
-)
-legend("bottomleft", legend = "A", cex = 1.3, bty = "n")
-plot(
-  df[, c("diff_var", "kvar")],
-  xlab = "var(f1) - var(f2)",
-  ylab = "SSE_var",
-  pch = 4, 
-  cex = .5,
-  col = "darkgrey"
-)
-legend("bottomleft", legend = "B", cex = 1.3, bty = "n")
-
-par(mar = c(7, 10, 0.4, 8) + 0.1)
+```{r, fig.cap = "The relationship between the two criteria SSE and $SSE_{Var}$ across partitions is shown.
+The partition represented by the diamond minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the triangular is a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
 
@@
   las = 1, cex = .5,
-  xlab = "kmeans",
-  ylab = "kvar"
+  xlab = "SSE",
+  ylab = "SSE_Var"
 )
-legend("bottomleft", legend = "C", cex = 1.3, bty = "n")
 

### 2021-03-05 paper.Rmd
added bullet point for arguments
working on initial example with plot

@@
 source("misc-functions.R")
-set.seed(2207) # 2207
 
@@
 is better suited to minimize difference with regard to both the mean and the 
-variance of the data; it was shown that even an entirely random split usually 
+variance of the data; it was# Variance objective function with changed order of the arguments 
+# (can be called by sapply() etc.)
+var_objective <- function(clusters, features) {
+  anticlust::variance_objective(features, clusters)
+}
+ shown that even an entirely random split usually 
 resulted in more similar standard deviations as k-means solely focuses on the 
@@
 distances $y_{ij}$ have much larger values, on average, than the raw values 
-$x_{ij}$. Therefore, the $SSE_{Var}$ criterion naturally receives greater weight 
-when computing the k-plus criterion than the standard $SSE$ criterion. This 
-imbalance is unwarranted because optimizing similarity with regard to variances 
-is not more important than optimizing similarity with regard to means. A simple 
-solution for this issue is to standardize all values $x_{ij}$ and $y_{ij}$ to 
-the same scale before computing $SSE_{kplus}$, e.g. via $z$-standardization.
+$x_{ij}$ (also see the axes in Figure 1). Therefore, the $SSE_{Var}$ criterion 
+naturally receives greater weight when computing the k-plus criterion than the 
+standard $SSE$ criterion. This imbalance is unwarranted because optimizing 
+similarity with regard to variances is not more important than optimizing 
+similarity with regard to means. A simple solution for this issue is to 
+standardize all values $x_{ij}$ and $y_{ij}$ to the same scale before computing 
+$SSE_{kplus}$, e.g. via $z$-standardization.
 
@@
 if (!file.exists(path)) {
+  set.seed(2207) # 2207
   # Generate data and all partitions
@@
   # Compute objectives for each partition
-  all_objectives_mean <- sapply(
-    partitions,
-    FUN = diff_mean,
-    features = features
-  )
-
-  all_objectives_var <- sapply(
-    partitions,
-    FUN = diff_vars,
-    features = features
-  )
-
   # For each partition: Compute SSE objective, and SSE(VAR)
-  # SSE
+  # 1. SSE
   all_objectives_kMeans <- sapply(
@@
 
-  # SSE(VAR)
+  squared_distances_from_mean <- squared_from_mean(as.matrix(features))
+
+  # 2. SSE(VAR)
   all_objectives_kVar <- sapply(
@@
     FUN = var_objective,
-    features = squared_from_mean(as.matrix(features))
+    features = squared_distances_from_mean
+  )
+
+  # 3. SSE(KPLUS)
+  # Use standardization to compute k-plus criterion!
+  all_objectives_kPlus <- sapply(
+    partitions,
+    FUN = var_objective,
+    features = scale(cbind(features, squared_distances_from_mean))
   )
@@
   df <- data.frame(
-    diff_mean = all_objectives_mean,
-    diff_var  = all_objectives_var,
     kmeans = all_objectives_kMeans,
-    kvar = all_objectives_kVar
+    kvar = all_objectives_kVar,
+    kplus = all_objectives_kPlus
   )
@@
 
-Optimizing a bicriterion objective such as $SSE_{kplus}$ may not succeed as one 
-may hope. The ideal outcome would be to find a partition where both the means 
-and the variances are very similar between groups. However, research on 
-multi-objective optimization generally shows that cases are rare where a single 
-solution satisfies all criteria to a satisfactory degree. Usually, a tradeoff is 
-inevitable. It is an open question whether k-plus anticlustering is a expection 
-from this rule. Next, a motivating example illustrates that in anticlustering, 
-aiming to simultaneously optimize similarity of means and variances is not only 
-desirable, but may actually be feasible. This notion is later confirmed when 
-applying k-plus anticlustering in simulation studies and examples using real 
-data.
-
-For this initial example, I created 14 data points from a univariate normal 
-distribution ($M = 15$, $SD = 3$) and generated all `r n_partitions(14, 2)` 
-possible ways to partition the 14 data points into $K = 2$ equal-sized groups. 
-For each partition, the criteria SSE and $SSE_{Var}$ were computed. Figure 1 
-shows the scatterplot of these criteria across partitions. Note that for both 
-criteria, a larger value is better, i.e., indicates higher similarity with 
-regard to means and variances, respectively. Ideally, to obtain overall 
-between-group similarity, a partition is chosen that minimizes the difference 
-with regard to both means and variances (see the triangular symbol). However, 
-optimal k-means anticlustering would create two groups that are not particularly 
-similar with regard to their variance. The partition that is represented by the 
-squared symbol has the overall best highest SSE, but is outperformed by many 
-other partitions with regard to $SSE_{Var}$. The diamond symbol represents 
-that partition that where the two groups have the least difference in 
-variances.
-
-Next: use the partition that optimizes $SSE_{kplus}$!
-
-[^onlyonegroup]: Note that this such a correspondence no longer holds 
-when there are more than two groups or there is more than one 
-numeric attribute (i.e., if $K > 2$ or $P > 1$).
+When a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be 
+to find a partition where both the means and the variances are very similar 
+between groups, i.e., a which partition where both SSE and $SSE_{Var}$ are close 
+to the optimum value. However, research on multi-objective optimization 
+generally shows that cases are rare where a single solution satisfies all 
+criteria to a satisfactory degree. Usually, a tradeoff is necessary. For 
+anticlustering, however, a motivating example illustrates that aiming to 
+simultaneously optimize similarity of means and variances is not only desirable, 
+but may actually be feasible.
 
@@
 
-# Illustrate Pareto set
-#pareto_pts <- pareto_set(df[, c("kmeans", "kvar")])
-#points(df[pareto_pts, c("kmeans", "kvar")], pch = 19)
-
-# Illustrate best partition wrt difference in means
+# Illustrate best partition wrt difference in means (i.e., SSE)
 points(
@@
 
-# Illustrate best partition wrt difference in variance
+# Illustrate best partition wrt difference in variance (i.e., SSE(VAR))
 points(
@@
 
-df$kplus <- rowSums(df[, c("kmeans", "kvar")])
-
-# Illustrate best partition wrt simililarity combined similarity / variance
+# Illustrate best partition wrt k-plus criterion
 points(
@@
 ```
+
+For this initial example, I created 14 data points from a univariate normal 
+distribution ($M = 15$, $SD = 3$) and generated all `r n_partitions(14, 2)` 
+possible ways to partition the 14 data points into $K = 2$ equal-sized groups. 
+Figure 1 shows the scatterplot of the criteria SSE and $SSE_{Var}$ across 
+partitions. Note that for both criteria, a larger value is better, i.e., 
+indicates higher similarity with regard to means and variances, respectively. 
+Therefore, to obtain overall between-group similarity, a partition should be 
+chosen that minimizes the difference with regard to both means and variances, 
+i.e., a partition that is depicted in the upper right corner of the scatterplot. 
+However, optimal k-means anticlustering would create two groups that are not 
+particularly similar with regard to their variance. The partition that is 
+represented by the squared symbol has the overall best highest SSE, but is 
+outperformed by many other partitions with regard to $SSE_{Var}$. Similarly, the 
+partition that maximizes $SSE_{Var}$ (the triangular symbol in Figure 1) is not 
+well-suited to optimize SSE at the same time. The diamond symbol represents the 
+partition that maximizes $SSE_{kplus}$, i.e., sum of SSE and $SSE_{Var}$. This 
+partition is located exactly in the right upper corner and yields `r 
+df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum 
+SSE, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global 
+maximum value of $SSE_{Var}$. Hence, this initial example shows that maximizing 
+the combined k-plus criterion may well be suited to optimize both similarity 
+with regard to means as well as variances between groups. Later, I will show 
+that this result also holds for larger data sets, more groups, and more 
+features.
+
+```{r}
+
+```
+
+[^onlyonegroup]: Note that this such a correspondence no longer holds 
+when there are more than two groups or there is more than one 
+numeric attribute (i.e., if $K > 2$ or $P > 1$).
 

### 2021-06-29 paper.Rmd
working on the paper again, yay Martin

@@
 
-  Martin Papenberg, Department of Experimental Psychology, Heinrich 
-  Heine University Düsseldorf.
+  Martin Papenberg, Department of Experimental Psychology, Heinrich Heine University Düsseldorf.
 
@@
 
-    Anticlustering is a method to partition data sets into groups 
-    in such a way that high between-group similarity and high within-group 
-    heterogeneity is obtained. In psychological research, anticlustering is for example 
-    used to to divide stimulus sets into equivalent sets, to split tests into parallel 
-    halfs, or to divide a group of students into heterogeneous classes. Different 
-    objective functions can be optimized in anticlustering applications. K-means 
-    anticlustering maximizes the error sum of squares, i.e., the sum of squared 
-    Euclidean distances from each element to the center of the group to which it 
-    belongs. K-means anticlustering thereby directly maximizes the similarity of 
-    the mean values of numeric attributes across groups, but it ignores the spread of 
-    the data. This paper presents k-plus anticlustering, a straight-forward 
-    extension of the k-means objective function that allows to simultaneously 
-    maximize similarity with regard to the mean as well as the variance across 
-    groups. It is shown that the bicriterion optimization problem reduces to an 
-    augmentation of the input data, leaving the original k-means criterion 
-    unchanged. Tests on simulated and real data show that similarity with regard to 
-    means and variance criteria can usually be achieved to an astonishing degree 
-    with almost no cost to either criterion. It is therefore concluded that the 
-    extended model should be preferred to classical k-means anticlustering for 
-    optimizing between-group similarity. An exception is given for binary data that 
-    does not profit from considering equality in variances.
+    Anticlustering refers to a collection of methods that partition data sets into disjunct groups.
+The partitioning process may pursue two goals: (a) high between-group similarity and (b) high within-group heterogeneity.
+Many anticlustering objectives simultaneously pursue both goals while other methods focus on a particular aspect.
+Recently, Brusco, Cradit and Steinley (2020) presented a highly effective bicriterion optimization algorithm that focuses primarily on maximzing within-group heterogeneity.
+The present paper introduces a novel bicriterion objective for optimizing between-group similarity.
+*K-plus anticlustering* is a straightforward extension of k-means anticlustering, which maximizes the sum of squared Euclidean distances between each data point and the center of the cluster to which it belongs.
+Thereby, k-means anticlustering directly minimizes the difference between the mean values of numeric attributes across clusters, but neglects optimizing similarity with regard to the variance of the data sets.
+K-plus anticlustering extends the k-means objective by adding a term that aims to simultaneously optimize similarity with regard to the variance.
+Surprisingly, it is shown that the combined objective (i.e., the bicriterion) can be reduced to an augmentation of the input data, leaving the original k-means criterion unchanged.
+Tests on simulated and real data show that k-plus anticlustering achieves high similarity both with regard to mean attribute values as well as the variance, usually with no noticable cost to either criterion.
+It is therefore concluded that the extended model should be preferred to classical k-means anticlustering when optimizing between-group similarity.
 
@@
 
-Anticlustering is a method to partition a set of elements into different 
-groups with the focus of obtaining high between-group similarity and 
-within-group heterogeneity. Anticlustering reverses the logic of its 
-better known twin, cluster analysis that seeks homogeneity within 
-clusters and heterogeneity between clusters [@rokach2005; 
-@steinley2006]. The method of anticlustering is useful in many 
-tasks surrounding the work of researchers in psychology, for example, 
-when splitting a test into parts of equal difficulty [@gierl2017], when 
-assigning students to work groups [@baker2002], or when assigning 
-stimuli to experimental conditions [@lahl2006]. Solving such problems 
-"by hand" is a tedious and time-consuming task and the quality of 
-manually established groups is usually improvable. Fortunately, these 
-problems can be formalized mathematically via several anticlustering 
-objective functions [e.g., @brusco2019; @spath1986; @baker2002; 
-@fernandez2013], and accessible open source software solutions exist 
-[@papenberg2020].
+Anticlustering is a collection of methods to partition elements into disjunct groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
+Oftentimes, both goals coincide: several anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990].
+Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
+Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012 for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements exhaustively into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].
 
-Several anticlustering objectives are available to assess between-group 
-similarity or within-group heterogeneity. In many cases, both objectives 
-are directly related, that is, maximizing heterogeneity within groups 
-simultaneously maximizes between-group similarity. Several 
-anticlustering objectives are based pairwise dissimilarity ratings; the 
-most prominent criterion of this kind is the *diversity*, which is the 
-total sum of dissimilarities between any elements within the same group. 
-Note that maximizing the diversity is equivalent to minimizing 
-the sum of dissimilarities between elements in different clusters, 
-directly illustrating the correspondence between within-group 
-heterogeneity and between-group similarity. Another important criterion 
-based on pairwise dissimilarities is the dispersion, which is the 
-minimum dissimilarity between any two items in the same group 
-[@fernandez2013; @brusco2019]. Maximizing the dispersion increases the 
-within-group heterogeneity by ensuring that any two elements as 
-dissimilar as possible from each other. Note that maximizing the 
-dispersion no longer simultaneously maximize between-group similarity. 
-In this case, these two potential aims of anticlustering diverge.
+Anticlustering is useful in many tasks surrounding research work in psychology.
+For example, anticlustering can be used to split a test into parts of equal difficulty [@gierl2017], to assign students to work groups [@baker2002] or when assigning stimuli to experimental conditions [@lahl2006].
+For an overview of anticlustering applications in psychology and other areas, see @brusco2019 and @papenberg2020.
+Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual group assignment is usually suboptimal.
+Fortunately, these problems can be formalized as mathematical optimization problems [e.g., @brusco2019; @spath1986; @baker2002; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
+
+---
+
+**TODO; hier weitermachen**
+
+Several anticlustering objectives are based pairwise dissimilarity ratings; the most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
+Another important criterion based on pairwise dissimilarities is the dispersion, which is the minimum dissimilarity between any two items in the same group [@fernandez2013; @brusco2019].
+Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements as dissimilar as possible from each other.
 
@@
 is better suited to minimize difference with regard to both the mean and the 
-variance of the data; it was# Variance objective function with changed order of the arguments 
-# (can be called by sapply() etc.)
-var_objective <- function(clusters, features) {
-  anticlust::variance_objective(features, clusters)
-}
- shown that even an entirely random split usually 
+variance of the data; it was shown that even an entirely random split usually
 resulted in more similar standard deviations as k-means solely focuses on the 
@@
 one-dimensional observations $(x_1, \ldots, x_N)$. The extension to the general 
-case of $P$ attributes will be straight forward
+case of $P$ attributes will be straight forward.
 
@@
 
-Note that a minor problem arises when ignoring the weights during optimizing the 
+Note that a problem may arise when ignoring the weights during optimizing the
 k-plus objective. Since their computation involves squaring, the squared 
@@
 ```{r, fig.cap = "The relationship between the two criteria SSE and $SSE_{Var}$ across partitions is shown.
-The partition represented by the diamond minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the triangular is a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond is a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
 
@@
 
-For this initial example, I created 14 data points from a univariate normal 
-distribution ($M = 15$, $SD = 3$) and generated all `r n_partitions(14, 2)` 
-possible ways to partition the 14 data points into $K = 2$ equal-sized groups. 
-Figure 1 shows the scatterplot of the criteria SSE and $SSE_{Var}$ across 
-partitions. Note that for both criteria, a larger value is better, i.e., 
-indicates higher similarity with regard to means and variances, respectively. 
-Therefore, to obtain overall between-group similarity, a partition should be 
-chosen that minimizes the difference with regard to both means and variances, 
-i.e., a partition that is depicted in the upper right corner of the scatterplot. 
-However, optimal k-means anticlustering would create two groups that are not 
-particularly similar with regard to their variance. The partition that is 
-represented by the squared symbol has the overall best highest SSE, but is 
-outperformed by many other partitions with regard to $SSE_{Var}$. Similarly, the 
-partition that maximizes $SSE_{Var}$ (the triangular symbol in Figure 1) is not 
-well-suited to optimize SSE at the same time. The diamond symbol represents the 
-partition that maximizes $SSE_{kplus}$, i.e., sum of SSE and $SSE_{Var}$. This 
-partition is located exactly in the right upper corner and yields `r 
-df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum 
-SSE, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global 
-maximum value of $SSE_{Var}$. Hence, this initial example shows that maximizing 
-the combined k-plus criterion may well be suited to optimize both similarity 
-with regard to means as well as variances between groups. Later, I will show 
-that this result also holds for larger data sets, more groups, and more 
-features.
+For this motivating example, I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
+Figure 1 shows the scatterplot of the criteria $SSE$ and $SSE_{Var}$ across partitions.
+Note that for both criteria, a larger value is better, i.e., indicates higher similarity with regard to means and variances, respectively.
+Therefore, to obtain overall between-group similarity, a partition should be chosen that is depicted in the upper right corner of the scatterplot.
+However, optimal k-means anticlustering would create two groups that are not particularly similar with regard to their variance.
+The partition that is represented by the squared symbol has the overall best highest $SSE$, but is outperformed by many other partitions with regard to $SSE_{Var}$.
+Similarly, the partition that maximizes $SSE_{Var}$ (the triangular symbol in Figure 1) is not well-suited to optimize SSE at the same time.
+The diamond symbol represents the partition that maximizes $SSE_{kplus}$, i.e., sum of SSE and $SSE_{Var}$.
+This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum SSE, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $SSE_{Var}$.
+
+This initial example shows that maximizing the combined k-plus criterion may well be suited to optimize both similarity with regard to means as well as variances between groups.
+Later, I will show that this result also holds for real data sets, using more groups and a larger number of numeric features.
 

### 2021-06-30 paper.Rmd
Abstract

@@
 ---
-title             : "k-plus: a bicriterion extension for k-means anticlustering"
+title             : "k-plus: a bicriterion extension of k-means anticlustering"
 shorttitle        : "k-plus anticlustering"
@@
 
+#  If the study extends or replicates previous research, cite the relevant work with an author–date citation.
+
 abstract: |
 
-    Anticlustering refers to a collection of methods that partition data sets into disjunct groups.
-The partitioning process may pursue two goals: (a) high between-group similarity and (b) high within-group heterogeneity.
+    Anticlustering refers to a collection of methods that partition data sets into disjunct groups with two primary goals: (a) high between-group similarity and (b) high within-group heterogeneity.
 Many anticlustering objectives simultaneously pursue both goals while other methods focus on a particular aspect.
-Recently, Brusco, Cradit and Steinley (2020) presented a highly effective bicriterion optimization algorithm that focuses primarily on maximzing within-group heterogeneity.
-The present paper introduces a novel bicriterion objective for optimizing between-group similarity.
-*K-plus anticlustering* is a straightforward extension of k-means anticlustering, which maximizes the sum of squared Euclidean distances between each data point and the center of the cluster to which it belongs.
-Thereby, k-means anticlustering directly minimizes the difference between the mean values of numeric attributes across clusters, but neglects optimizing similarity with regard to the variance of the data sets.
-K-plus anticlustering extends the k-means objective by adding a term that aims to simultaneously optimize similarity with regard to the variance.
+The present paper introduces a novel bicriterion objective aiming at optimizing between-group similarity.
+*K-plus anticlustering* is an extension of k-means anticlustering, which minimizes the difference between mean values of numeric attributes across clusters.
+The k-plus objective extends k-means by including a term that aims to optimize similarity with regard to the variance of data across groups, which is ignored by k-means.
 Surprisingly, it is shown that the combined objective (i.e., the bicriterion) can be reduced to an augmentation of the input data, leaving the original k-means criterion unchanged.
-Tests on simulated and real data show that k-plus anticlustering achieves high similarity both with regard to mean attribute values as well as the variance, usually with no noticable cost to either criterion.
-It is therefore concluded that the extended model should be preferred to classical k-means anticlustering when optimizing between-group similarity.
+Tests on simulated and real data show that k-plus anticlustering usually achieves high similarity both with regard to mean attribute values as well as the variance with no noticable sacrifice to either criterion.
+The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity.
 
@@
 Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
-Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012 for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements exhaustively into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].
+Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012 for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].
 
@@
 
+- show that k-plus is strongly preferred over diversity when the groups are of unequal size
+
 # Discussion

### 2021-06-30 paper.Rmd
save; introduction is being worked on

@@
 For an overview of anticlustering applications in psychology and other areas, see @brusco2019 and @papenberg2020.
-Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual group assignment is usually suboptimal.
+Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual partitioning is usually suboptimal.
 Fortunately, these problems can be formalized as mathematical optimization problems [e.g., @brusco2019; @spath1986; @baker2002; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
 
----
-
-**TODO; hier weitermachen**
-
-Several anticlustering objectives are based pairwise dissimilarity ratings; the most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
+Several anticlustering objectives exist that differ in the assumed input data and their mathematical definition of cluster similarity.
+Some anticlustering objectives are based pairwise dissimilarity ratings; the most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
 Another important criterion based on pairwise dissimilarities is the dispersion, which is the minimum dissimilarity between any two items in the same group [@fernandez2013; @brusco2019].
@@
 
-Oftentimes, the available data does not represent pairwise similarity, 
-but rather numeric, categorical or binary attributes [@dry2009]. That 
-is, the data is a two-way two-mode [@arabie1992] table where rows 
-represent the entities of study, and columns represent variables 
-describing the elements. For example, numeric ratings may describe the 
-attributes of word stimuli for psycholinguistic experiments. Such data 
-may include Norms for imagery, concreteness and orthographic variables 
-[@friendly1982]. In other cases, attributes are binary and may represent 
-the presence or absence of a feature [@tversky1977]. Such data can be 
-converted to pairwise dissimilarities for example by computing an 
-appropriate dissimilarity measure, such as the pairwise Euclidean or 
-squared Euclidean distance across a set of numeric attributes 
-[@brusco2019]. Then, anticlustering criteria based on pairwise 
-dissimilarity can be applied to partition the elements.
+Oftentimes, the available data does not represent pairwise similarity, but rather numeric, categorical or binary attributes [@dry2009].
+That is, the data is a two-way two-mode [@arabie1992] table where rows represent the entities of study, and columns represent variables describing the elements.
+For example, numeric ratings may describe the attributes of word stimuli for psycholinguistic experiments.
+Such data may include Norms for imagery, concreteness and orthographic variables [@friendly1982].
+In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977].
+Such data can be converted to pairwise dissimilarities for example by computing an appropriate dissimilarity measure, such as the pairwise Euclidean or squared Euclidean distance across a set of numeric attributes [@brusco2019].
+Then, anticlustering criteria based on pairwise dissimilarity can be applied to partition the elements.
 
-A different approach to anticlustering directly works with feature based data 
-and does not require a transformation into pairwise dissimilarities. Instead, 
-the sum of the squared Euclidean distances between elements and the cluster they 
-are assigned to is minimized [@brusco2006branch], most prominently using the 
-k-means heuristic [@jain2010; @steinley2006]. Minimizing the k-means criterion 
-simultaneously maximizes between-cluster separation and within-cluster 
-homogeneity [@aloise2009np]. Späth [-@spath1986] and Valev [-@valev1998; 
--@valev1983] independently recognized that by optimizing the reversed k-means 
-objective that is, maximizing the sum of distances between elements and cluster 
-centroids -- a data set is partitioned into similar groups, thereby reversing 
-the logic of classical clustering. They coined the term anticlustering for this 
-purpose. K-means anticlustering maximizes the sum of squared Euclidean distances 
-from each element to the center of the cluster to which it belongs. As 
-@spath1986 noted, k-means anticlustering thereby directly minimizes differences 
-with regard to the means of the numeric variables between groups. Hence, groups 
-only become similar with regard to the mean of the distribution of the numeric 
-attributes, which may perform poorly as a measure of overall similarity; the 
-spread of the data is not targeted and may differ between sets. @papenberg2020 
-showed that maximizing the diversity (based on the pairwise Euclidean distances) 
-is better suited to minimize difference with regard to both the mean and the 
-variance of the data; it was shown that even an entirely random split usually
-resulted in more similar standard deviations as k-means solely focuses on the 
-mean. When requiring that groups are overall similar to each other, this focus 
-may be misguided as similar means can be obtained even when the underlying 
-distributions are different [@anscombe1973].
+A different approach to anticlustering directly works with feature based data and does not require a transformation into pairwise dissimilarities.
+Instead, the sum of the squared Euclidean distances between elements and the cluster they are assigned to is minimized [@brusco2006branch], most prominently using the k-means heuristic [@jain2010; @steinley2006].
+Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np].
+Späth [-@spath1986] and Valev [-@valev1998; -@valev1983] independently recognized that by optimizing the reversed k-means objective that is, maximizing the sum of distances between elements and cluster centroids -- a data set is partitioned into similar groups, thereby reversing the logic of classical clustering.
+They coined the term anticlustering for this purpose.
+K-means anticlustering maximizes the sum of squared Euclidean distances from each element to the center of the cluster to which it belongs.
+As @spath1986 noted, k-means anticlustering thereby directly minimizes differences with regard to the means of the numeric variables between groups.
+Hence, groups only become similar with regard to the mean of the distribution of the numeric attributes, which may perform poorly as a measure of overall similarity; the spread of the data is not targeted and may differ between sets.
+@papenberg2020 showed that maximizing the diversity (based on the pairwise Euclidean distances) is better suited to minimize difference with regard to both the mean and the variance of the data; it was shown that even an entirely random split usually resulted in more similar standard deviations as k-means solely focuses on the mean.
+When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [@anscombe1973].
 
-In this paper, I present k-plus anticlustering, a simple extension of the 
-k-means objective function to maximize similarity with regard to means and 
-variances at the same time. The aim is to establish groups that are overall 
-similar to each other. It is shown that this is achieved by formulating a 
-bicriterion optimization objective for k-means anticlustering. Interestingly, 
-this reformulation reduces to an augmentation of the original data matrix while 
-the optimization objective remains unchanged. This insight has important 
-consequences as it allows to adapt standard optimization schemes for single 
-criterion anticlustering. I show that the k-plus criterion outperforms other 
-anticlustering criteria when the goal is the minimize difference with regard to 
-the mean and variance of the different groups. In most cases, both objectives 
-can be optimized satisfactorily at the same time without any loss on either; 
-when the data set is large enough (usually, when $N > XX$), the bicriterion 
-optimization leads to > 99.99% objective performance as compared to the single 
-optimization of either k-means or k-variance.
+In this paper, I present k-plus anticlustering, a simple extension of the k-means objective function to maximize similarity with regard to means and variances at the same time.
+The aim is to establish groups that are overall similar to each other.
+It is shown that this is achieved by formulating a bicriterion optimization objective for k-means anticlustering.
+Interestingly, this reformulation reduces to an augmentation of the original data matrix while the optimization objective remains unchanged.
+This insight has important consequences as it allows to adapt standard optimization schemes for single criterion anticlustering.
+I show that the k-plus criterion outperforms other anticlustering criteria when the goal is the minimize difference with regard to the mean and variance of the different groups.
+In most cases, both objectives can be optimized satisfactorily at the same time without any loss on either; when the data set is large enough (usually, when $N > XX$), the bicriterion optimization leads to > 99.99% objective performance as compared to the single optimization of either k-means or k-variance.
 

### 2021-06-30 paper.Rmd
working on paper, added application (OASIS)

@@
     Anticlustering refers to a collection of methods that partition data sets into disjunct groups with two primary goals: (a) high between-group similarity and (b) high within-group heterogeneity.
-Many anticlustering objectives simultaneously pursue both goals while other methods focus on a particular aspect.
+Some anticlustering objectives simultaneously pursue both goals while other methods focus on a particular aspect.
 The present paper introduces a novel bicriterion objective aiming at optimizing between-group similarity.
@@
 Surprisingly, it is shown that the combined objective (i.e., the bicriterion) can be reduced to an augmentation of the input data, leaving the original k-means criterion unchanged.
-Tests on simulated and real data show that k-plus anticlustering usually achieves high similarity both with regard to mean attribute values as well as the variance with no noticable sacrifice to either criterion.
+Tests on simulated and real data show that k-plus anticlustering usually achieves high similarity both with regard to mean attribute values as well as the variance, with no noticable sacrifice to either criterion.
 The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity.
@@
 Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
-Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012 for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].
+Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].
 
@@
 For example, anticlustering can be used to split a test into parts of equal difficulty [@gierl2017], to assign students to work groups [@baker2002] or when assigning stimuli to experimental conditions [@lahl2006].
-For an overview of anticlustering applications in psychology and other areas, see @brusco2019 and @papenberg2020.
 Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual partitioning is usually suboptimal.
 Fortunately, these problems can be formalized as mathematical optimization problems [e.g., @brusco2019; @spath1986; @baker2002; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
+For a general overview of anticlustering approaches and applications in Psychology, see @brusco2019 and @papenberg2020.
 
-Several anticlustering objectives exist that differ in the assumed input data and their mathematical definition of cluster similarity.
-Some anticlustering objectives are based pairwise dissimilarity ratings; the most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
-Another important criterion based on pairwise dissimilarities is the dispersion, which is the minimum dissimilarity between any two items in the same group [@fernandez2013; @brusco2019].
-Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements as dissimilar as possible from each other.
+Anticlustering objectives differ in the assumed input data and their mathematical definition of cluster similarity.
+Some anticlustering objectives employ pairwise dissimilarity ratings to compute a total measure of cluster (dis)similarity.
+The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
+Maximizing the diversity simultaneously optimizes within-group heterogeneity as well as between-group similarity.
+Another important criterion based on pairwise dissimilarities is the *dispersion*, i.e., the minimum dissimilarity between any two elements within the same group [@fernandez2013; @brusco2019].
+Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements as dissimilar from each other as possible.
 
-Oftentimes, the available data does not represent pairwise similarity, but rather numeric, categorical or binary attributes [@dry2009].
+Oftentimes, the data available to researchers does not represent pairwise similarity, but numeric, categorical or binary attributes [@dry2009].
 That is, the data is a two-way two-mode [@arabie1992] table where rows represent the entities of study, and columns represent variables describing the elements.
 For example, numeric ratings may describe the attributes of word stimuli for psycholinguistic experiments.
-Such data may include Norms for imagery, concreteness and orthographic variables [@friendly1982].
+Such data may include norms for imagery, concreteness and orthographic variables [@friendly1982].
 In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977].
-Such data can be converted to pairwise dissimilarities for example by computing an appropriate dissimilarity measure, such as the pairwise Euclidean or squared Euclidean distance across a set of numeric attributes [@brusco2019].
-Then, anticlustering criteria based on pairwise dissimilarity can be applied to partition the elements.
+In the context of anticlustering, One possible approach for handling two-way two-mode data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across a set of numeric attributes [@brusco2019].
+Then, anticlustering criteria based on pairwise dissimilarities such as the dispersion or diversity can be optimized to partition the elements into groups.
 
-A different approach to anticlustering directly works with feature based data and does not require a transformation into pairwise dissimilarities.
-Instead, the sum of the squared Euclidean distances between elements and the cluster they are assigned to is minimized [@brusco2006branch], most prominently using the k-means heuristic [@jain2010; @steinley2006].
+A different approach to anticlustering directly works with the two-way two-mode feature data using the k-means criterion.
+K-means is probably the best-known clustering method and its reversal has been recognized as a useful anticlustering tool by Späth [-@spath1986] and Valev [-@valev1998; -@valev1983].
+In k-means clustering, the sum of the squared Euclidean distances between elements and the center of the cluster they are assigned to is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006].
 Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np].
-Späth [-@spath1986] and Valev [-@valev1998; -@valev1983] independently recognized that by optimizing the reversed k-means objective that is, maximizing the sum of distances between elements and cluster centroids -- a data set is partitioned into similar groups, thereby reversing the logic of classical clustering.
-They coined the term anticlustering for this purpose.
-K-means anticlustering maximizes the sum of squared Euclidean distances from each element to the center of the cluster to which it belongs.
-As @spath1986 noted, k-means anticlustering thereby directly minimizes differences with regard to the means of the numeric variables between groups.
-Hence, groups only become similar with regard to the mean of the distribution of the numeric attributes, which may perform poorly as a measure of overall similarity; the spread of the data is not targeted and may differ between sets.
-@papenberg2020 showed that maximizing the diversity (based on the pairwise Euclidean distances) is better suited to minimize difference with regard to both the mean and the variance of the data; it was shown that even an entirely random split usually resulted in more similar standard deviations as k-means solely focuses on the mean.
+For the anticlustering application, the k-means criterion is maximized instead, thereby achieving similarity between groups.
+Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters.
+Hence, groups are similar with regard to the mean of the distribution of the numeric attributes wheras other parameters---such as the variance---are ignored.
 When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [@anscombe1973].
+@papenberg2020 showed that maximizing the diversity (based on the pairwise Euclidean distances) is better suited to minimize difference with regard to both the mean and the variance of the data; it was shown that even an entirely random split usually resulted in more similar variances because k-means solely focuses on the mean similarity.
 
-In this paper, I present k-plus anticlustering, a simple extension of the k-means objective function to maximize similarity with regard to means and variances at the same time.
-The aim is to establish groups that are overall similar to each other.
-It is shown that this is achieved by formulating a bicriterion optimization objective for k-means anticlustering.
-Interestingly, this reformulation reduces to an augmentation of the original data matrix while the optimization objective remains unchanged.
-This insight has important consequences as it allows to adapt standard optimization schemes for single criterion anticlustering.
+To optimize overall group similarity---and hence to overcome the limitations of k-means anticlustering---I present k-plus anticlustering in this paper.
+The k-plus criterion is an extension of the k-means objective that includes a term aiming at optimizing similarity with regard to the variance of data across groups.
+Interestingly, this bicriterion reformulation can be understood as an augmentation of the original data matrix, leaving the original k-means objective unchanged.
 I show that the k-plus criterion outperforms other anticlustering criteria when the goal is the minimize difference with regard to the mean and variance of the different groups.
-In most cases, both objectives can be optimized satisfactorily at the same time without any loss on either; when the data set is large enough (usually, when $N > XX$), the bicriterion optimization leads to > 99.99% objective performance as compared to the single optimization of either k-means or k-variance.
+In most cases, both objectives can be optimized satisfactorily at the same time without any loss on either.
+In several examples I illustrate how readers can use k-plus anticlustering easily using the free and open source software anticlust, which is an extension to the statistical programming language R [@papenberg2020].
 
@@
 
-For the purpose of problem formalization, I adopt the notation Steinley (2006) 
-provided in his comprehensive synthesis of half a century of research on 
-k-means. K-means clustering is used to partition two-way, two-mode data into 
-groups. That is, $N$ elements each having $P$ attributes are assigned to $K$ 
-groups $(C_1., \ldots, C_K)$, $C_k$ being the set of $n_k$ objects in group $k$. 
-The partitioning is subject to the requirements that each element is assigned to 
-exactly one group, formalized by the two constraints
+For the purpose of problem formalization, I adopt the notation @steinley2006 provided in his comprehensive synthesis of half a century of research on k-means.
+K-means clustering (and anticlustering) is used to partition two-way, two-mode data into groups.
+That is, $N$ elements each having $P$ attributes are assigned to $K$ groups $(C_1., \ldots, C_K)$, $C_k$ being the set of $n_k$ objects in group $k$.
+The partitioning is subject to the requirement that each element is assigned to exactly one group, formalized by the following two constraints:
 
@@
 
-In anticlustering applications, we  impose an additional constraint on the group 
-sizes $n_k$, where the most common one requires that all groups have equal size:
+In anticlustering applications, we impose an additional constraint on the group sizes $n_k$, where the most common restriction requires that all groups have equal size:
 
@@
 
-For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means 
-clustering aims to minimize the error sum of squares (SSE), which is the sum of 
-the squared Euclidean distances between each data point and its cluster 
-centroid:
+For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means anticlustering aims to maximize the error sum of squares (SSE), which is the sum of the squared Euclidean distances between each data point and its cluster centroid:
 
@@
 
-The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, 
-\overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, 
-where each entry is the mean value of one of the $P$ attributes, computed across 
-all observations belonging the $k$th cluster $C_k$ ($k = 1, \ldots, K$):
+The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all observations belonging the $k$th cluster $C_k$ ($k = 1, \ldots, K$):
 
@@
 
-Note there is a direct connection between the $SSE$ and the location of the 
-cluster centroids [@spath1986]. If the $SSE$ is minimal, the centroid values 
-$\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible 
-from the overall data centroid $\overline{\mathbf{x}} = 
-(\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, 
-\overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the overall mean 
-value on the $p$th attribute:
+Note there is a direct connection between the $SSE$ and the location of the cluster centroids [@spath1986].
+If the $SSE$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the overall mean value on the $p$th attribute:
 
@@
 
-If the $SSE$ is maximal, the cluster centroids are as close as possible to the 
-overall centroid $\overline{\mathbf{x}}$ and therefore to each other 
-[@spath1986]. Thus, k-means anticlustering directly optimizes the 
-similarity of the mean attribute values across clusters.
+If the $SSE$ is maximal---which is the goal of k-means *anticlustering*---, the cluster centroids are as close as possible to the overall centroid $\overline{\mathbf{x}}$ and therefore to each other [@spath1986].
+Thus, k-means anticlustering directly optimizes the similarity of the mean attribute values across clusters.
 
-## A criterion to quantify differences in variances
+## Quantifying differences in variance
 
-K-means anticlustering---i.e., maximizing the SSE---minimizes differences 
-between the means of the input variables. I now present a rather 
-straight-forward adaption of the SSE to capture differences in variance between 
-groups. To motivate the new criterion, we first consider a set of 
-one-dimensional observations $(x_1, \ldots, x_N)$. The extension to the general 
-case of $P$ attributes will be straight forward.
+In the following, I present a rather straight-forward adaption of the $SSE$ to quantify how the variance of numeric data differs between groups.
+To motivate the new criterion, we first consider a set of one-dimensional observations $(x_1, \ldots, x_N)$.
+The extension to the general case of $P$ attributes will be straight-forward.
 
-The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is 
-computed as the mean of the squared distances between observations and their 
-mean $\overline{\mathbf{x}}$:
+The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
 
@@
 
-We next define a new variable $y = (y_1, \ldots, y_N)$, representing the squared 
-deviation of each observation from the mean:
+When defining a new variable $y = (y_1, \ldots, y_N)$ as the squared deviation of each observation from the mean $\overline{\mathbf{x}}$, the sample variance of $x$ can be reformulated as the mean of $y$:
 
@@
 
-The sample variance of $x$ is consequently the mean of these new values:
-
 $$
@@
 
-Since the sample variance is basically defined as an average, k-means 
-anticlustering can be employed to equalize the variance across groups. That is, 
-we can formulate an adaption of the SSE that quantifies the degree to which 
-variances differ between groups:
+Because the sample variance is basically defined as an average, k-means anticlustering can be employed to equalize the variance across groups.
+That is, we can formulate an adaption of the $SSE$ to quantify the degree to which the variance differs between groups.
+First, however, we generalize the idea to the case of $P$ attributes.
+For each attribute, we have to compute the squared deviation between all values and the mean of the attribute.
+We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$, where
 
 $$
-\mathit{SSE_{Var}} = 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (y_{i} - \overline{y} ^{(k)})^2
+y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2.
 $$
 
-Maximizing $\mathit{SSE_{Var}}$ will equalize the variance between groups on one 
-attribute. The criterion is extended to the general case of $P$ attributes as
+To reflect how strongly the variances of all attributes vary across the $K$ clusters, we now define the criterion $\mathit{SSE_{Var}}$:
 
@@
 
-where
+Analogously to the standard $SSE$, the centroid values $\overline{y}_j ^{(k)}$ are given as:
 
@@
 
-and
+Note that $\mathit{SSE_{Var}}$ has the same form as the standard $\mathit{SSE}$.
+However, the input data reflects the squared difference between each data point and the attribute mean instead of the raw data points.
 
-$$
-y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2.
-$$
+## A bicriterion model: K-plus anticlustering
 
-Note that $\mathit{SSE_{Var}}$ has the same form as the standard $\mathit{SSE}$. 
-However, the input data reflect the squared difference between each data point 
-and the average values instead of the raw data points.
-
-## A bicriterion model: k-plus anticlustering
-
-In most applications, it will not be of interest to only maximize $SSE_{Var}$ 
-(i.e., equalizing variances) without also considering the standard $SSE$ (i.e., 
-equalizing means). A combination of $SSE_{Var}$ and $SSE$, however, can be 
-employed to simultaneously optimize both objectives. The most simple approach to 
-such a bicriterion optimization is the weighted sum approach [@naidu2014]. That 
-is, both criteria are computed independently from each other and then combined 
-into a single criterion through a weighted sum:
+It is unlikely that an exclusive optimization of $SSE_{Var}$ (i.e., equalizing variances) is of interest without also considering the $SSE$ (i.e., equalizing means).
+A combination of $SSE_{Var}$ and $SSE$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall cluster similarity.
+The most simple approach for such a bicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
+That is, both criteria are computed independently from each other and then combined into a single criterion through a weighted sum:
 
@@
 
-Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$ 
-(i.e., or setting them to the same value). When defining $z_i = (x_{i1}, \ldots, 
-x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and 
-$y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ is computed as
+Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
+When defining $z_i = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and $y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ is then computed as
 
@@
 
-This formula again has the same form as SSE (and $SSE_{Var}$), but is computed 
-on different data. In this case, both the original data $x_{ij}$ as well as the 
-squared distances $y_{ij}$ are used as attributes that enter the objective 
-function. That is, computing the k-plus criterion is done by first computing the 
-values $y_{ij}$, appending these to the original data matrix, and then compute 
-SSE on the augmented data matrix.
+This formula again has the same form as $SSE$ (and $SSE_{Var}$), but is computed on different data.
+In this case, both the original data $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function.
+That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending these to the original data matrix, and then computing the $SSE$ on the augmented data matrix.
+Hence, k-plus anticlustering reduces to an augmentation of the data matrix, leaving the original k-means criterion unchanged.
 
-Note that a problem may arise when ignoring the weights during optimizing the
-k-plus objective. Since their computation involves squaring, the squared 
-distances $y_{ij}$ have much larger values, on average, than the raw values 
-$x_{ij}$ (also see the axes in Figure 1). Therefore, the $SSE_{Var}$ criterion 
-naturally receives greater weight when computing the k-plus criterion than the 
-standard $SSE$ criterion. This imbalance is unwarranted because optimizing 
-similarity with regard to variances is not more important than optimizing 
-similarity with regard to means. A simple solution for this issue is to 
-standardize all values $x_{ij}$ and $y_{ij}$ to the same scale before computing 
-$SSE_{kplus}$, e.g. via $z$-standardization.
+Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective.
+Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ have much larger values, on average, than the raw values $x_{ij}$ (also see Figure 1).
+Therefore, the $SSE_{Var}$ criterion naturally receives greater weight when computing $\mathit{SSE_{kplus}}$ than the standard $SSE$ criterion.
+This imbalance is most likely unwarranted because optimizing similarity with regard to variances is not more important than optimizing similarity with regard to means.
+A solution is to standardize all values $x_{ij}$ and $y_{ij}$ to the same scale before computing $SSE_{kplus}$, e.g.
+via $z$-standardization.
 
-## Criterion Tradeoff
+## Criterion tradeoff
 
@@
 
-When a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be 
-to find a partition where both the means and the variances are very similar 
-between groups, i.e., a which partition where both SSE and $SSE_{Var}$ are close 
-to the optimum value. However, research on multi-objective optimization 
-generally shows that cases are rare where a single solution satisfies all 
-criteria to a satisfactory degree. Usually, a tradeoff is necessary. For 
-anticlustering, however, a motivating example illustrates that aiming to 
-simultaneously optimize similarity of means and variances is not only desirable, 
-but may actually be feasible.
+When optimizing a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $SSE$ and $SSE_{Var}$ are close to the global optimum value.
+Generally, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree [@ferligoj1992direct].
+Usually, a tradeoff is necessary.
+For anticlustering, a motivating example illustrates that aiming to simultaneously optimize similarity of means and variances is not only desirable, but may actually be feasible.
 
-```{r, fig.cap = "The relationship between the two criteria SSE and $SSE_{Var}$ across partitions is shown.
-The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond is a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all partitions (for $N = 14$ and $K = 2$).
+The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$ and represents a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
 
@@
 Similarly, the partition that maximizes $SSE_{Var}$ (the triangular symbol in Figure 1) is not well-suited to optimize SSE at the same time.
-The diamond symbol represents the partition that maximizes $SSE_{kplus}$, i.e., sum of SSE and $SSE_{Var}$.
-This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum SSE, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $SSE_{Var}$.
+The diamond symbol represents the partition that maximizes the unweighted $SSE_{kplus}$, i.e., the sum of $SSE$ and $SSE_{Var}$.
+This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $SSE$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $SSE_{Var}$.
 
@@
 ```
-
-[^onlyonegroup]: Note that this such a correspondence no longer holds 
-when there are more than two groups or there is more than one 
-numeric attribute (i.e., if $K > 2$ or $P > 1$).
-
-## Bicriterion optimization model
-
-```{r echo = FALSE}
-# From Wikipedia on Multi-objective optimization:
-
-# Without additional subjective preference information, all Pareto optimal 
-# solutions are considered equally good. Researchers study multi-objective 
-# optimization problems from different viewpoints and, thus, there exist 
-# different solution philosophies and goals when setting and solving them. 
-# The goal may be to find a representative set of Pareto optimal 
-# solutions, and/or quantify the trade-offs in satisfying the different 
-# objectives, and/or finding a single solution that satisfies the 
-# subjective preferences of a human decision maker (DM).
-
-# The most preferred results can be found using different philosophies. 
-# Multi-objective optimization methods can be divided into four 
-# classes. In so-called no preference methods, no DM is expected to be 
-# available, but a neutral compromise solution is identified without 
-# preference information. The other classes are so-called a priori, a 
-# posteriori and interactive methods and they all involve preference 
-# information from the DM in different ways.
-
-# Solving a multi-objective optimization problem is sometimes understood 
-# as approximating or computing all or a representative set of Pareto 
-# optimal solutions.
-
-# Scalarizing a multi-objective optimization problem is an a priori 
-# method, which means formulating a single-objective optimization problem 
-# such that optimal solutions to the single-objective optimization problem 
-# are Pareto optimal solutions to the multi-objective optimization 
-# problem. In addition, it is often required that every Pareto optimal 
-# solution can be reached with some parameters of the scalarization.
-
-```
-
-- pragmatic solution [easy and practical approach, @marler2010; 
-@naidu2014]: formalize the weighted-sum objective and use the 
-weighted-sum approach. Then check out whether the simplest version is 
-sufficient in this case (but see the problems of weighted sum approach 
-in several contributions by Brusco et al.). Refer to @marler2010 to 
-argue for weighted sum for the current problem
-
-- it should be noted that both mathematical objectives capture different 
-aspects of an underlying substantive objective, that is to maximize 
-similarity of data distributions between different groups
-
-- Weighted sum approach works well on this specific problem [although in 
-general there are caveats to the weighted-sum method, see @brusco2019]. 
-is fast, good results. in fact, the objectives seem to work well 
-together
-
-- weighted sum approach generates a single solution from the 
-(approximated) pareto set
-
-- In this paper: Formalization of the objective function, recognition
-that this formulation does not change the basic structure of the k-means
-objective, because it reduces to augmentation of the data input
-
-- not interested in the Pareto efficient set! Because: Overall 
-similarity should be maximized, which requires that both criteria are 
-satisfied to a strong degree (not interested in just maximizing 
-similarity wrt variance; this may be interesting in other applications). 
-Instead: A tradeoff is possible, and in many cases, simultaneous 
-optimization will meet both criteria almost as well as optimizing just
-either criterion. To justify this approach: show the pareto front, which 
-is not a diagonal / circle [see @brusco2012] but a RIGHT TRIANGLE! (this 
-is probably because this is *anti*clustering!)
-
-- Evaluation will consist of a comparison of different objectives based 
-on the same (simple) exchange algorithm
-
-### Weight selection
-
-- How to select weights? (Oftentimes: for equal weights, the results 
-are good)
-
-- Preference to weight [e.g., @marler2010]. However, the relationship 
-between weights and outcome is not always straightforward and the 
-results may disappoint. On the one hand, vastly different weights need 
-not lead to different solutions; and even equal weights can be 
-sufficient if the solution space offers partitions that maximize both 
-objectives at the same time. "the solution may not preserve one’s
-initial preferences no matter how the weights are set"; "even if one 
-determines acceptable values for the weights a priori, the final 
-solution may not accurately reflect initial preferences" [@marler2010, 
-p. 860].
-
-- Standardization of the extended data matrix to ensure that the two
-criterion functions "do not differ markedly in scale" [@brusco2019]. 
-See Figure 1: the objectives do differ in scale, because k-variance 
-is based on a quadratic distance and therefore has larger values.
-
-- Using variable weighting (see Steinley, 2006, p. 13), we can 
-adjust the relative importance of minimizing differences in means
-and variances. Standard in `anticlust`: Equal weight, this is 
-achieved by first creating the new variables, representing 
-the squared differences from the mean, and then standardizing
-the entire feature matrix. If $N$ is large enough (relative to $K$),
-there is usually no need for weighting.
 
@@
 
-# Examples
+# Applying k-plus anticlusting using the R package anticlust
 
-- use some examples (e.g., OASIS data set, 900 items allocated to 10 
-sets, same mean and variance across the 10 sets)
+In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package anticlust [@papenberg2020].
+The results presented here used anticlust version 0.5.6.
+The package can be installed in the R environment using the `install.packages()` command:
 
-- show that k-plus is strongly preferred over diversity when the groups are of unequal size
+```R
+install.packages("anticlust")
+```
+
+In this application, I use norming data for the OASIS image data set that is freely available online [@kurdi2017introducing].
+@kurdi2017introducing assembled 900 open-access color images and collected ratings on two affective dimensions: arousal and valence.
+@brielmann2019intense collected an additional rating dimension by measuring how the same 900 images were rated with regard to their beauty.
+In my application, I use all three features, currently available from Github.[^githubbeauty]
+
+[^githubbeauty]: https://github.com/aenneb/OASIS-beauty
+
+I used k-plus anticlustering to divide the data set of 900 images into 9 groups of 100 images each; this process takes about 3-5 seconds on a contemporary personal computer, depending on how many iterations are needed to find the local maximum.
+The task was accomplished using the following R code:
+
+```R
+library(anticlust) # load the package anticlust
+
+anticlustering(
+  features,
+  standardize = TRUE,
+  K = 9,
+  objective = "kplus",
+  method = "local-maximum"
+)
+```
+
+Here, the variable `features` has to be a data table containing the image data: 900 rows representing images; 3 columns representing the features beauty, arousal, and valence.
+The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the `features` variable (**TODO**).
+The arguments `objective = "kplus"` and `method = "local-maximum"` ensure that the k-plus criterion is optimized using the local maximum search method described in the previous section (**TODO**).
+The argument `K` describes the number of groups.
+The argument `standardize = TRUE` ensures that all features are standardized before the optimization starts; in particular, this option enforces that equalizing means and variances receives the same weight during the optimization process.
+
+Table 1 shows the results of the anticlustering application by listing the descriptive statistics (means and standard deviations) for each of the 9 groups and for each of the 3 features.
+Additionaly, Table 1 shows the descriptive statistics for an application of k-means anticlustering and anticluster editing (which maximizes the diversity criterion).
+It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations.
+K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations.
+Anticluster editing was also well-suited to match both means and standard deviations, but was slightly outperformed by k-plus.
+The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to the means and standard deviations.
+
+```{r}
+
+path <- "results_oasis_same_size.csv"
+
+if (!file.exists(path)) {
+  oasis <- read.csv(
+    "https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv"
+  )
+
+  features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
+
+  K <- 9
+  mean_sd_tabs <- list()
+  for (criterion in c("kplus", "variance", "diversity")) {
+    oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
+      features,
+      standardize = TRUE,
+      K = K,
+      objective = criterion,
+      method = "local-maximum"
+    )
+    mean_sd_tabs[[criterion]] <- mean_sd_tab(
+      features,
+      oasis[[paste(criterion, "group", sep = "_")]]
+    )
+    mean_sd_tabs[[criterion]] <- cbind(
+      1:K,
+      nrow(features) / K,
+      mean_sd_tabs[[criterion]]
+    )
+  }
+
+  # append all tables to the same table:
+  oasis_tab <- do.call(rbind, mean_sd_tabs)
+  rownames(oasis_tab) <- NULL
+  colnames(oasis_tab) <- c("Group", "N", "Beauty", "Valence", "Arousal")
+
+  write.table(
+    oasis_tab, path, row.names = FALSE, sep = ";"
+  )
+
+} else {
+  oasis_tab <- read.csv(path, sep = ";")
+}
+
+apa_table(
+  oasis_tab,
+  caption = "Descriptive statistics for OASIS features by group and anticlustering method.",
+  note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
+  escape = FALSE,
+  align = c("r"),
+  stub_indents = list("\\textbf{K-plus anticlustering}" = 1:9,
+                      "\\textbf{K-means anticlustering}" = 10:18,
+                      "\\textbf{Anticluster editing}" = 19:27)
+)
+```
+
+Table 2 illustrates another advantage of k-plus anticlustering over anticluster editing.
+When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups.
+K-plus anticlustering strives for equal means and variances in all groups, regardless of group size.
+
+Unequal group sizes can be requested in anticlust by passing the group sizes as a vector to the `K` argument:
+
+```R
+anticlustering(
+  features,
+  standardize = TRUE,
+  K = c(100, 100, 100, 600),
+  objective = "kplus",
+  method = "local-maximum"
+)
+```
+
+```{r}
+
+path <- "results_oasis_unequal_size.csv"
+
+if (!file.exists(path)) {
+  oasis <- read.csv(
+    "https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv"
+  )
+
+  features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
+
+  K <- c(100, 100, 100, 600)
+  mean_sd_tabs <- list()
+  for (criterion in c("kplus", "variance", "diversity")) {
+    oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
+      features,
+      standardize = TRUE,
+      K = K,
+      objective = criterion,
+      method = "local-maximum"
+    )
+    mean_sd_tabs[[criterion]] <- mean_sd_tab(
+      features,
+      oasis[[paste(criterion, "group", sep = "_")]]
+    )
+    mean_sd_tabs[[criterion]] <- cbind(1:length(K), K, mean_sd_tabs[[criterion]])
+  }
+
+  # append all tables to the same table:
+  oasis_tab <- do.call(rbind, mean_sd_tabs)
+  rownames(oasis_tab) <- NULL
+  colnames(oasis_tab) <- c("Group", "N", "Beauty", "Valence", "Arousal")
+
+  write.table(
+    oasis_tab, path, row.names = FALSE, sep = ";"
+  )
+
+} else {
+  oasis_tab <- read.csv(path, sep = ";")
+}
+
+apa_table(
+  oasis_tab,
+  caption = "Performance of the diversity and k-plus objectives for unequal group sizes.",
+  note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
+  escape = FALSE,
+  align = c("r"),
+  stub_indents = list("\\textbf{K-plus anticlustering}" = 1:4,
+                      "\\textbf{K-means anticlustering}" = 5:8,
+                      "\\textbf{Anticluster editing}" = 9:12)
+)
+```
 

### 2021-06-30 potpurri.Rmd
working on paper, added application (OASIS)

## Bicriterion optimization model

```{r echo = FALSE}
# From Wikipedia on Multi-objective optimization:

# Without additional subjective preference information, all Pareto optimal
# solutions are considered equally good. Researchers study multi-objective
# optimization problems from different viewpoints and, thus, there exist
# different solution philosophies and goals when setting and solving them.
# The goal may be to find a representative set of Pareto optimal
# solutions, and/or quantify the trade-offs in satisfying the different
# objectives, and/or finding a single solution that satisfies the
# subjective preferences of a human decision maker (DM).

# The most preferred results can be found using different philosophies.
# Multi-objective optimization methods can be divided into four
# classes. In so-called no preference methods, no DM is expected to be
# available, but a neutral compromise solution is identified without
# preference information. The other classes are so-called a priori, a
# posteriori and interactive methods and they all involve preference
# information from the DM in different ways.

# Solving a multi-objective optimization problem is sometimes understood
# as approximating or computing all or a representative set of Pareto
# optimal solutions.

# Scalarizing a multi-objective optimization problem is an a priori
# method, which means formulating a single-objective optimization problem
# such that optimal solutions to the single-objective optimization problem
# are Pareto optimal solutions to the multi-objective optimization
# problem. In addition, it is often required that every Pareto optimal
# solution can be reached with some parameters of the scalarization.

```

- pragmatic solution [easy and practical approach, @marler2010;
@naidu2014]: formalize the weighted-sum objective and use the
weighted-sum approach. Then check out whether the simplest version is
sufficient in this case (but see the problems of weighted sum approach
in several contributions by Brusco et al.). Refer to @marler2010 to
argue for weighted sum for the current problem

- it should be noted that both mathematical objectives capture different
aspects of an underlying substantive objective, that is to maximize
similarity of data distributions between different groups

- Weighted sum approach works well on this specific problem [although in
general there are caveats to the weighted-sum method, see @brusco2019].
is fast, good results. in fact, the objectives seem to work well
together

- weighted sum approach generates a single solution from the
(approximated) pareto set

- In this paper: Formalization of the objective function, recognition
that this formulation does not change the basic structure of the k-means
objective, because it reduces to augmentation of the data input

- not interested in the Pareto efficient set! Because: Overall
similarity should be maximized, which requires that both criteria are
satisfied to a strong degree (not interested in just maximizing
similarity wrt variance; this may be interesting in other applications).
Instead: A tradeoff is possible, and in many cases, simultaneous
optimization will meet both criteria almost as well as optimizing just
either criterion. To justify this approach: show the pareto front, which
is not a diagonal / circle [see @brusco2012] but a RIGHT TRIANGLE! (this
is probably because this is *anti*clustering!)

- Evaluation will consist of a comparison of different objectives based
on the same (simple) exchange algorithm

### Weight selection

- How to select weights? (Oftentimes: for equal weights, the results
are good)

- Preference to weight [e.g., @marler2010]. However, the relationship
between weights and outcome is not always straightforward and the
results may disappoint. On the one hand, vastly different weights need
not lead to different solutions; and even equal weights can be
sufficient if the solution space offers partitions that maximize both
objectives at the same time. "the solution may not preserve one’s
initial preferences no matter how the weights are set"; "even if one
determines acceptable values for the weights a priori, the final
solution may not accurately reflect initial preferences" [@marler2010,
p. 860].

- Standardization of the extended data matrix to ensure that the two
criterion functions "do not differ markedly in scale" [@brusco2019].
See Figure 1: the objectives do differ in scale, because k-variance
is based on a quadratic distance and therefore has larger values.

- Using variable weighting (see Steinley, 2006, p. 13), we can
adjust the relative importance of minimizing differences in means
and variances. Standard in `anticlust`: Equal weight, this is
achieved by first creating the new variables, representing
the squared differences from the mean, and then standardizing
the entire feature matrix. If $N$ is large enough (relative to $K$),
there is usually no need for weighting.

### 2021-07-23 paper.Rmd
simulation and example applications

@@
 classoption       : "doc"
-output            : papaja::apa6_pdf
+output            : 
+  papaja::apa6_pdf: 
+    number_sections: true
 ---
@@
 Maximizing the diversity simultaneously optimizes within-group heterogeneity as well as between-group similarity.
+Because the diversity is the reversal of the clustering editing objective [@zahn1964; @shamir2004vf], @papenberg2020 used the term *anticluster editing* to refer to anticlustering approaches that optimize the diversity.
 Another important criterion based on pairwise dissimilarities is the *dispersion*, i.e., the minimum dissimilarity between any two elements within the same group [@fernandez2013; @brusco2019].
@@
 When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [@anscombe1973].
-@papenberg2020 showed that maximizing the diversity (based on the pairwise Euclidean distances) is better suited to minimize difference with regard to both the mean and the variance of the data; it was shown that even an entirely random split usually resulted in more similar variances because k-means solely focuses on the mean similarity.
+@papenberg2020, showed that maximizing the diversity (based on the pairwise Euclidean distances) is better suited to minimize difference with regard to both the mean and the variance of the data; it was shown that even an entirely random split usually resulted in more similar variances because k-means solely focuses on the mean similarity.
 
@@
 
-## A bicriterion model: K-plus anticlustering
+## A bicriterion model: K-plus anticlustering {#kplus}
 
@@
 Similarly, the partition that maximizes $SSE_{Var}$ (the triangular symbol in Figure 1) is not well-suited to optimize SSE at the same time.
-The diamond symbol represents the partition that maximizes the unweighted $SSE_{kplus}$, i.e., the sum of $SSE$ and $SSE_{Var}$.
+The diamond symbol represents the partition that maximizes (the unweighted) $SSE_{kplus}$, i.e., the sum of $SSE$ and $SSE_{Var}$.
 This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $SSE$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $SSE_{Var}$.
@@
 This initial example shows that maximizing the combined k-plus criterion may well be suited to optimize both similarity with regard to means as well as variances between groups.
-Later, I will show that this result also holds for real data sets, using more groups and a larger number of numeric features.
+In the next section I present a simulation study that more systematically investigates the performance of the k-plus criterion.
 
-```{r}
+# Simulation study
 
-```
+To systematically evaluate the new k-plus criterion, I compared three anticlustering approaches in a simulation study: (a) k-means anticlustering, optimizing the $SSE$; (b) k-plus anticlustering, optimizing the $SSE_{kplus}$; (c) anticluster editing, optimizing the diversity criterion.
+All methods employed an a priori standardization of the data sets.
+In the case of k-plus anticlustering, this means that the augmented input data (i.e., the original data and the squared distance data) were subjected to a standardization, which is recommended.
 
-# Evaluation
+The simulation used the statistical programming environment R [@R-base].
+The R package `anticlust` [Version 0.5.6; @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
+The R packages `dplyr` [Version 1.0.7; @R-dplyr], `ggplot2` [Version 3.3.5; @R-ggplot2], `papaja` [Version 0.1.0.9997; @R-papaja] and `tidyr` [Version 1.1.3; @R-tidyr] were used for data processing and result presentation.
+The simulation study is fully reproducible via code and data that is accessible from the accompanying OSF repository (**link TODO**).
 
-- Simulation 1: Show that k-means extended obtains 99.999% of the 
-k-means and k-variance objective, on average. 
-  + New objective seems to work best when N is high relative to K; optimizing 
-two objectives at the same time is hard when N is low. Easier when only 
-optimizing one objective (i.e., the mean, as done by normal k-means 
-anticlustering). What I can say: When N / K > X, the objective is 
-usually 99.99% of both objectives.
-  + Two examples: $K = 2$, $N = 20$ vs. $N = 100$, using classical
-k-means: only mean; using k-variance: only variance; combination:
-both, but works better for larger $N$
+## Conditions
 
-- Simulation 2: show that k-variance anticlustering outperforms 
-other objectives with regard to similarity in means / variances 
-(in particular: anticluster editing)
+First, I generated 10,000 data sets which were subsequently subjected to the three anticlustering methods.
+The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features $M$ varied between 1 and 10; (c) the distribution of features was randomly determined to be either (i) uniform in [0, 1], (ii) a standard normal distribution with *M* = 0 and *SD* = 1, or (iii) a wider normal distribution with *M* = 0 and *SD* = 2.
+For a given data set, all $M$ features were independently sampled from the given distribution and no correlations were assumed between features.
+All three anticlustering methods were applied to each of the 10,000 data sets using $K = 2$, $K = 3$ and $K = 4$; groups sizes were always equal.
 
-- Often, maximizing k-means-variance yields more similar variances 
-than k-variance alone. Why is that? I guess: structure in the solution 
-space, local maxima are real. When also considering means, you more 
-often escape bad local maxima? The finding is: k-means-extended better 
-minimizes the maximum difference between the variance across groups 
-than k-variance (for K > 2), which in the eye of the researcher, may 
-be a more intuitive criterion.
+## Optimization algorithm
 
-# Applying k-plus anticlusting using the R package anticlust
+To optimize the three anticlustering criteria, a local-maximum search was applied that systematically swaps elements between groups to gradually improve the objective function [Method LCW in @weitz1998].
+Building on an initial random assignment of elements to groups, elements are swapped between groups in such a way that each swap improves the objective criterion by the largest amount that is possible.
+That is, for each element, all possible swaps are simulated and the resulting objective values are stored.
+After each exchange has been tested, the one exchange is realized that improves the objective the most.
+No swap is realized if no swap improves the criterion.
+The exchange procedure is repeated for each element.
+Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements.
+The search stops as soon as an iteration through the entire set no longer yields any improvement, i.e., as soon as a local maximum has been found.
+To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
+This simulation employed five repetitions of the local maximum search.
 
-In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package anticlust [@papenberg2020].
-The results presented here used anticlust version 0.5.6.
+## Evaluation criteria
+
+Two primary performance criteria were assessed.
+First, for k-plus anticlustering, I checked how well the identified solution approximated the ordinary k-means criterion.
+Specifically, I computed the $SSE$ for the solution that was returned by k-means anticlustering---that directly optimizes $SSE$---and for the solution that was returned by k-plus anticlustering that only indirectly optimizes $SSE$ through $SSE_{kplus}$.
+If optimizing $SSE_{kplus}$ simultaneously optimizes the standard $SSE$ to a satisfactory degree, we may conclude that simultaneously striving for similar means and variances is feasible by optimizing the k-plus criterion.
+
+Second, to compare the three anticlustering criteria with regard to their ability to create similar groups, I computed the means and standard deviations and assessed the degree to which discrepancies occurred between groups.
+To this end, I then computed the difference between the minimum and maximum value of the means and standard deviations between groups, for each of the $M$ features.
+The mean of the differences across all features quantified the total dissimilarity in means ($\Delta M$) and standard deviations ($\Delta \mathit{SD}$).
+Therefore, lower values of $\Delta M$ and $\Delta \mathit{SD}$ indicated that groups were more similar with regard means or standard deviations, respectively.
+
+## Results
+
+- k-plus approximates k-means very well, especially for large $N$ (for larger $K$, $N$ schould increase as well)
+- k-plus is the best method to optimize similarity with regard to means and standard deviations at the same time
+
+# K-plus anticlustering using the R package anticlust
+
+In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020].
+The results presented here used `anticlust` version 0.5.6.
 The package can be installed in the R environment using the `install.packages()` command:
@@
 ```
+
+## Example Application I
 
@@
 The argument `K` describes the number of groups.
-The argument `standardize = TRUE` ensures that all features are standardized before the optimization starts; in particular, this option enforces that equalizing means and variances receives the same weight during the optimization process.
+The argument `standardize = TRUE` ensures that all features are standardized before the optimization starts; in particular, this option enforces that equalizing means and variances receives the same weight during the optimization process (see [Section 1.2](#kplus)).
 
@@
   features,
+  K = c(100, 100, 100, 600),
   standardize = TRUE,
-  K = c(100, 100, 100, 600),
   objective = "kplus",
@@
 
+## Example Application II: Small data set (N = 96)
+
+The simulation study indicated that k-plus may show decreased performance for smaller group sizes.
+Therefore,    this second example application illustrates that k-plus anticlustering may also yield satisfactory results when group are smaller than in the OASIS example.
+I employ a norming data set of 96 word stimuli that was contributed to the `anticlust` package by Dr.
+Schaper [@schaper2019metacognitive; @schaper2019metamory; also see @papenberg2020 for a description of the norming data].
+After loading the `anticlust` package, the data set can be accessed as follows:
+
+```{r, echo = TRUE}
+data(schaper2019)
+```
+
+Table 3 illustrates the results that are obtained when using k-plus anticlustering to divide the stimulus set into 6 groups of size `r 96 / 6` each.
+Even though the means and standard deviations are no longer perfectly matched between groups, arguably, they are still similar enough for any practical purpose, especially when considering inevitable measurement error in norming studies.
+The following code was used to achieve the results reported in Table 3:
+
+```{r, echo = TRUE}
+features <- schaper2019[, 3:6]
+anticlusters <- anticlustering(
+  features,
+  K = 6,
+  objective = "kplus",
+  method = "local-maximum",
+  repetitions = 5,
+  standardize = TRUE
+)
+```
+
+Note that standardization is particularly useful for anticlustering when the score range differs strongly between features, as is the case for the Schaper data set.
+
+```{r}
+
+tab <- mean_sd_tab(features, anticlusters)
+tab <- cbind(1:6, 96 / 6, tab)
+colnames(tab) <- c("Group", "N", "Typicality", "Atypicality", "Syllables", "Frequency")
+
+apa_table(
+  tab,
+  caption = "Descriptive statistics by group for the Schaper et al. (2019) data set, after applying k-plus anticlustering.",
+  note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
+  escape = FALSE,
+  align = c("r")
+)
+
+```
+
 # Discussion
 
-- Focus on presentation of new objective function, the bicriterion 
-optimization process might be improved [e.g., by using a direct 
-algorithm as @brusco2019 did]
+- k-plus anticlustering is useful and should in the default case be preferred to k-means anticlustering.
+For optimizing similarity in means and variances it also outperforms anticluster editing, which, to be fair, does not directly strive to do so.
+Unequal group sizes give another advantage for k-plus anticlustering in comparison to anticluster editing.
+For small samples sizes (or large $K$), k-plus anticlustering may return slightly less-than-optimal results in the eye of the user, but in this case it is no problem to return to k-means and only focus on the means.
+There is not "anticlustering hacking"; users may simply choose the results they like best.
+For now, I strongly recommend using k-plus anticlustering for creating groups with high between group similarity.
 
-- Future research: Application of more sophisticated bicriterion 
-optimization algorithms are interesting
+## Limitations and outlook
 
-- new criterion: optimizes between-group similarity, no correspondence 
-to within-group heterogeneity [see @brusco2019 for a exceptional method 
-maximizing within-group heterogeneity]
+- This paper focused on the presentation of new objective function, the bicriterion optimization process might be improved [e.g., by using a direct algorithm as @brusco2019 did].
+From the algorithmic point of view, this paper did not strive to apply the latest cutting edge methodology.
+For most practical purposes, however, I assume that optimizing the k-plus criterion will yield very satisfactory results for researchers.
+I have already worked with researchers planning their experiments who are generally amazed by the groupd that are returned by k-plus anticlustering.
 
-- Similarly: the reversed objective no longer has a useful clustering 
-interpretation: Maximizing differences in variances does not lead to 
-homogeneous and/or well separated clusters
+## Future research
 
-- extending on covariances?
+- Future research: Application of more sophisticated bicriterion optimization algorithms are interesting
+
+- extending k-means anticlustering on maximizing similarity with regard to covariances? Mahalanobis distance?
+
+# General thoughts on anticlustering
+
+- new criterion: optimizes between-group similarity, no correspondence to within-group heterogeneity [see @brusco2019 for a exceptional method maximizing within-group heterogeneity]
+
+- Similarly: the reversed objective no longer has a useful clustering interpretation: Maximizing differences in variances does not lead to homogeneous and/or well separated clusters
 

### 2021-10-05 paper.Rmd
some work on paper

@@
 
-    Anticlustering refers to a collection of methods that partition data sets into disjunct groups with two primary goals: (a) high between-group similarity and (b) high within-group heterogeneity.
-Some anticlustering objectives simultaneously pursue both goals while other methods focus on a particular aspect.
-The present paper introduces a novel bicriterion objective aiming at optimizing between-group similarity.
-*K-plus anticlustering* is an extension of k-means anticlustering, which minimizes the difference between mean values of numeric attributes across clusters.
-The k-plus objective extends k-means by including a term that aims to optimize similarity with regard to the variance of data across groups, which is ignored by k-means.
-Surprisingly, it is shown that the combined objective (i.e., the bicriterion) can be reduced to an augmentation of the input data, leaving the original k-means criterion unchanged.
-Tests on simulated and real data show that k-plus anticlustering usually achieves high similarity both with regard to mean attribute values as well as the variance, with no noticable sacrifice to either criterion.
+    Anticlustering refers to a collection of methods that partition data sets into disjunct while addressing two objectives: (a) high between-group similarity and (b) high within-group heterogeneity.
+For some anticlustering formulations, these two objectives coincide: high between-group similarity is achieved when within-group heterogeneity is maximized.
+Other anticlustering methods focus on one of the two aspects.
+The present paper focuces on maximizing between-group similarity by introducing a novel bicriterion objective.
+*K-plus anticlustering* is an extension of k-means anticlustering, which minimizes the difference between the mean of numeric attributes across clusters.
+The k-plus objective extends k-means by adding a term that aims to optimize similarity with regard to the variance of the attributes data across clusters.
+It is shown that the combined bicriterion objective can be reduced to an augmentation of the input data, leaving the original k-means criterion unchanged.
+A computer simulation and practical examples on real data show that k-plus anticlustering usually achieves high similarity both with regard to mean attribute values as well as the variance, with no noticable sacrifice to either criterion.
 The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity.
+K-plus anticlustering is easily accessible to researchers and practitioners: the method has been implemented as part of the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
 
@@
 
-Anticlustering objectives differ in the assumed input data and their mathematical definition of cluster similarity.
-Some anticlustering objectives employ pairwise dissimilarity ratings to compute a total measure of cluster (dis)similarity.
-The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
-Maximizing the diversity simultaneously optimizes within-group heterogeneity as well as between-group similarity.
+Anticlustering methods can be distinguished by the input data (pairwise dissimilarity vs.
+attributes values) and the objective function that is used to optimize group between-group similarity or within-group heterogeneity.
+The most prominent criterion on the basis of pairwise dissimilarity ratings is the *diversity*, i.e., the total sum of dissimilarities between any elements within the same group.
+Maximizing the diversity simultaneously optimizes within-group heterogeneity and between-group similarity.
 Because the diversity is the reversal of the clustering editing objective [@zahn1964; @shamir2004vf], @papenberg2020 used the term *anticluster editing* to refer to anticlustering approaches that optimize the diversity.
@@
 
-Oftentimes, the data available to researchers does not represent pairwise similarity, but numeric, categorical or binary attributes [@dry2009].
+Oftentimes, the data available to researchers does not represent pairwise similarity, but numeric, categorical or binary attribute values (or *features*; @dry2009).
 That is, the data is a two-way two-mode [@arabie1992] table where rows represent the entities of study, and columns represent variables describing the elements.
@@
 Such data may include norms for imagery, concreteness and orthographic variables [@friendly1982].
-In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977].
-In the context of anticlustering, One possible approach for handling two-way two-mode data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across a set of numeric attributes [@brusco2019].
-Then, anticlustering criteria based on pairwise dissimilarities such as the dispersion or diversity can be optimized to partition the elements into groups.
+In other cases, attributes are binary and may represent the presence or absence of a facial feature [@tversky1977].
+In the context of anticlustering, one possible approach for handling two-way two-mode data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across a set of numeric attributes [@brusco2019].
+Subsequently, an anticlustering criterion such as the diversity can be optimized to partition the elements into similar groups.
 
-A different approach to anticlustering directly works with the two-way two-mode feature data using the k-means criterion.
-K-means is probably the best-known clustering method and its reversal has been recognized as a useful anticlustering tool by Späth [-@spath1986] and Valev [-@valev1998; -@valev1983].
+A different approach to anticlustering directly works with the two-way two-mode feature data by maximizing the k-means criterion.
+K-means is probably the best-known clustering method.
+Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful anticlustering tool by Späth [-@spath1986] and Valev [-@valev1998; -@valev1983].
 In k-means clustering, the sum of the squared Euclidean distances between elements and the center of the cluster they are assigned to is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006].
@@
 Hence, groups are similar with regard to the mean of the distribution of the numeric attributes wheras other parameters---such as the variance---are ignored.
-When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [@anscombe1973].
-@papenberg2020, showed that maximizing the diversity (based on the pairwise Euclidean distances) is better suited to minimize difference with regard to both the mean and the variance of the data; it was shown that even an entirely random split usually resulted in more similar variances because k-means solely focuses on the mean similarity.
+When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [e.g., @anscombe1973].
+@papenberg2020, showed that maximizing the diversity (based on the Euclidean distance) is better suited to minimize difference with regard to both the mean and the variance of the data; it was shown that even an entirely random split usually resulted in more similar variances because k-means solely focuses on the mean similarity.
 
@@
 Interestingly, this bicriterion reformulation can be understood as an augmentation of the original data matrix, leaving the original k-means objective unchanged.
-I show that the k-plus criterion outperforms other anticlustering criteria when the goal is the minimize difference with regard to the mean and variance of the different groups.
+I show that the k-plus criterion outperforms other anticlustering criteria when aiming at minimizing differences with regard to the mean and the variance of the different groups.
 In most cases, both objectives can be optimized satisfactorily at the same time without any loss on either.
-In several examples I illustrate how readers can use k-plus anticlustering easily using the free and open source software anticlust, which is an extension to the statistical programming language R [@papenberg2020].
+Moreover, in two pratical examples I illustrate how readers can easily apply k-plus anticlustering using the free and open source software `anticlust`, which is an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
 
@@
 That is, $N$ elements each having $P$ attributes are assigned to $K$ groups $(C_1., \ldots, C_K)$, $C_k$ being the set of $n_k$ objects in group $k$.
-The partitioning is subject to the requirement that each element is assigned to exactly one group, formalized by the following two constraints:
+The partitioning is subject to the requirement that each element is assigned to exactly one group, formalized by the following two constraints [@papenberg2020]:
 
@@
 
-For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means anticlustering aims to maximize the error sum of squares (SSE), which is the sum of the squared Euclidean distances between each data point and its cluster centroid:
+For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means anticlustering aims to maximize the error sum of squares ($\mathit{SSE}$), which is the sum of the squared Euclidean distances between each data point and its cluster centroid:
 
@@
 
-Note there is a direct connection between the $SSE$ and the location of the cluster centroids [@spath1986].
-If the $SSE$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the overall mean value on the $p$th attribute:
+Note there is a direct connection between the $\mathit{SSE}$ and the location of the cluster centroids [@spath1986].
+If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the overall mean value on the $p$th attribute:
 
@@
 
-If the $SSE$ is maximal---which is the goal of k-means *anticlustering*---, the cluster centroids are as close as possible to the overall centroid $\overline{\mathbf{x}}$ and therefore to each other [@spath1986].
+If the $\mathit{SSE}$ is maximal---which is the goal of k-means *anticlustering*---, the cluster centroids are as close as possible to the overall centroid $\overline{\mathbf{x}}$ and therefore to each other [@spath1986].
 Thus, k-means anticlustering directly optimizes the similarity of the mean attribute values across clusters.
@@
 
-In the following, I present a rather straight-forward adaption of the $SSE$ to quantify how the variance of numeric data differs between groups.
+In the following, I present a variation of the $\mathit{SSE}$ that quantifies how the variance of numeric data differs between groups.
 To motivate the new criterion, we first consider a set of one-dimensional observations $(x_1, \ldots, x_N)$.
@@
 Because the sample variance is basically defined as an average, k-means anticlustering can be employed to equalize the variance across groups.
-That is, we can formulate an adaption of the $SSE$ to quantify the degree to which the variance differs between groups.
+That is, we can formulate an adaption of the $\mathit{SSE}$ to quantify the degree to which the variance differs between groups.
 First, however, we generalize the idea to the case of $P$ attributes.
@@
 
-Analogously to the standard $SSE$, the centroid values $\overline{y}_j ^{(k)}$ are given as:
+Analogously to the standard $\mathit{SSE}$, the centroid values $\overline{y}_j ^{(k)}$ are given as:
 
@@
 
-It is unlikely that an exclusive optimization of $SSE_{Var}$ (i.e., equalizing variances) is of interest without also considering the $SSE$ (i.e., equalizing means).
-A combination of $SSE_{Var}$ and $SSE$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall cluster similarity.
+It is unlikely that an exclusive optimization of $SSE_{Var}$ (i.e., equalizing variances) is of interest without also considering the $\mathit{SSE}$ (i.e., equalizing means).
+A combination of $SSE_{Var}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall cluster similarity.
 The most simple approach for such a bicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
@@
 
-This formula again has the same form as $SSE$ (and $SSE_{Var}$), but is computed on different data.
+This formula again has the same form as $\mathit{SSE}$ (and $SSE_{Var}$), but is computed on different data.
 In this case, both the original data $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function.
-That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending these to the original data matrix, and then computing the $SSE$ on the augmented data matrix.
+That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending these to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data matrix.
 Hence, k-plus anticlustering reduces to an augmentation of the data matrix, leaving the original k-means criterion unchanged.
@@
 Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ have much larger values, on average, than the raw values $x_{ij}$ (also see Figure 1).
-Therefore, the $SSE_{Var}$ criterion naturally receives greater weight when computing $\mathit{SSE_{kplus}}$ than the standard $SSE$ criterion.
+Therefore, the $SSE_{Var}$ criterion naturally receives greater weight when computing $\mathit{SSE_{kplus}}$ than the standard $\mathit{SSE}$ criterion.
 This imbalance is most likely unwarranted because optimizing similarity with regard to variances is not more important than optimizing similarity with regard to means.
@@
 
-When optimizing a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $SSE$ and $SSE_{Var}$ are close to the global optimum value.
+When optimizing a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $\mathit{SSE}$ and $SSE_{Var}$ are close to the global optimum value.
 Generally, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree [@ferligoj1992direct].
@@
 For this motivating example, I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
-Figure 1 shows the scatterplot of the criteria $SSE$ and $SSE_{Var}$ across partitions.
+Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $SSE_{Var}$ across partitions.
 Note that for both criteria, a larger value is better, i.e., indicates higher similarity with regard to means and variances, respectively.
@@
 However, optimal k-means anticlustering would create two groups that are not particularly similar with regard to their variance.
-The partition that is represented by the squared symbol has the overall best highest $SSE$, but is outperformed by many other partitions with regard to $SSE_{Var}$.
+The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $SSE_{Var}$.
 Similarly, the partition that maximizes $SSE_{Var}$ (the triangular symbol in Figure 1) is not well-suited to optimize SSE at the same time.
-The diamond symbol represents the partition that maximizes (the unweighted) $SSE_{kplus}$, i.e., the sum of $SSE$ and $SSE_{Var}$.
-This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $SSE$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $SSE_{Var}$.
+The diamond symbol represents the partition that maximizes (the unweighted) $SSE_{kplus}$, i.e., the sum of $\mathit{SSE}$ and $SSE_{Var}$.
+This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $SSE_{Var}$.
 
@@
 
-To systematically evaluate the new k-plus criterion, I compared three anticlustering approaches in a simulation study: (a) k-means anticlustering, optimizing the $SSE$; (b) k-plus anticlustering, optimizing the $SSE_{kplus}$; (c) anticluster editing, optimizing the diversity criterion.
+To systematically evaluate the new k-plus criterion, I compared three anticlustering approaches in a simulation study: (a) k-means anticlustering, optimizing the $\mathit{SSE}$; (b) k-plus anticlustering, optimizing the $SSE_{kplus}$; (c) anticluster editing, optimizing the diversity criterion.
 All methods employed an a priori standardization of the data sets.
@@
 First, for k-plus anticlustering, I checked how well the identified solution approximated the ordinary k-means criterion.
-Specifically, I computed the $SSE$ for the solution that was returned by k-means anticlustering---that directly optimizes $SSE$---and for the solution that was returned by k-plus anticlustering that only indirectly optimizes $SSE$ through $SSE_{kplus}$.
-If optimizing $SSE_{kplus}$ simultaneously optimizes the standard $SSE$ to a satisfactory degree, we may conclude that simultaneously striving for similar means and variances is feasible by optimizing the k-plus criterion.
+Specifically, I computed the $\mathit{SSE}$ for the solution that was returned by k-means anticlustering---that directly optimizes $\mathit{SSE}$---and for the solution that was returned by k-plus anticlustering that only indirectly optimizes $\mathit{SSE}$ through $SSE_{kplus}$.
+If optimizing $SSE_{kplus}$ simultaneously optimizes the standard $\mathit{SSE}$ to a satisfactory degree, we may conclude that simultaneously striving for similar means and variances is feasible by optimizing the k-plus criterion.
 
@@
 Anticluster editing was also well-suited to match both means and standard deviations, but was slightly outperformed by k-plus.
-The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to the means and standard deviations.
+The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to the means and standard deviations, but instead optimizes the within-group heterogeneity.
 

### 2021-10-05 paper.Rmd
minor stuff

@@
 
-#  If the study extends or replicates previous research, cite the relevant work with an author–date citation.
-
 abstract: |
 
-    Anticlustering refers to a collection of methods that partition data sets into disjunct while addressing two objectives: (a) high between-group similarity and (b) high within-group heterogeneity.
+    Anticlustering refers to a collection of methods that partition data sets into disjunct groups while addressing two objectives: (a) high between-group similarity and (b) high within-group heterogeneity.
 For some anticlustering formulations, these two objectives coincide: high between-group similarity is achieved when within-group heterogeneity is maximized.
 Other anticlustering methods focus on one of the two aspects.
-The present paper focuces on maximizing between-group similarity by introducing a novel bicriterion objective.
+The present paper focuces on maximizing between-group similarity by introducing a novel bicriterion objective for anticlustering.
 *K-plus anticlustering* is an extension of k-means anticlustering, which minimizes the difference between the mean of numeric attributes across clusters.
@@
 
-Table 2 illustrates another advantage of k-plus anticlustering over anticluster editing.
-When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups.
+Table 2 illustrates an advantage of k-plus anticlustering over anticluster editing: When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups.
 K-plus anticlustering strives for equal means and variances in all groups, regardless of group size.
-
-Unequal group sizes can be requested in anticlust by passing the group sizes as a vector to the `K` argument:
+Unequal group sizes can be requested in anticlust by passing the different group sizes to the `K` argument:
 
@@
 
-Table 3 illustrates the results that are obtained when using k-plus anticlustering to divide the stimulus set into 6 groups of size `r 96 / 6` each.
-Even though the means and standard deviations are no longer perfectly matched between groups, arguably, they are still similar enough for any practical purpose, especially when considering inevitable measurement error in norming studies.
+Table 3 illustrates the results that are obtained when using k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
+Even though the means and standard deviations are no longer perfectly matched between groups, arguably, they are still similar enough for any practical purpose---especially when considering inevitable measurement error in norming studies.
 The following code was used to achieve the results reported in Table 3:

### 2021-10-06 paper.Rmd
some stuff

@@
 
-    Anticlustering refers to a collection of methods that partition data sets into disjunct groups while addressing two objectives: (a) high between-group similarity and (b) high within-group heterogeneity.
-For some anticlustering formulations, these two objectives coincide: high between-group similarity is achieved when within-group heterogeneity is maximized.
-Other anticlustering methods focus on one of the two aspects.
+    Anticlustering refers to a collection of methods that partition data sets into disjunct groups while aiming for two objectives: (a) high between-group similarity and (b) high within-group heterogeneity.
+For some anticlustering formulations, these two objectives coincide, i.e., high between-group similarity is achieved when within-group heterogeneity is maximized.
+Other anticlustering criteria focus on one of the two aspects.
 The present paper focuces on maximizing between-group similarity by introducing a novel bicriterion objective for anticlustering.
@@
 The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity.
-K-plus anticlustering is easily accessible to researchers and practitioners: the method has been implemented as part of the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
+K-plus anticlustering is easily accessible to researchers and practitioners: the method has been implemented as part of the open source R package `anticlust` and is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
 
@@
 Anticlustering is useful in many tasks surrounding research work in psychology.
-For example, anticlustering can be used to split a test into parts of equal difficulty [@gierl2017], to assign students to work groups [@baker2002] or when assigning stimuli to experimental conditions [@lahl2006].
-Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual partitioning is usually suboptimal.
+For example, anticlustering can be used to split a test into parts of equal difficulty [@gierl2017], to assign students to work groups [@baker2002] or when assigning stimuli to different---but parallel---experimental conditions [@lahl2006].
+For a more comprehensive overview of anticlustering applications in Psychology, see @brusco2019 and @papenberg2020.
+Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual partitioning is usually subpar.
 Fortunately, these problems can be formalized as mathematical optimization problems [e.g., @brusco2019; @spath1986; @baker2002; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
-For a general overview of anticlustering approaches and applications in Psychology, see @brusco2019 and @papenberg2020.
 
@@
 attributes values) and the objective function that is used to optimize group between-group similarity or within-group heterogeneity.
+Some anticlustering objectives quantify group similarity on the basis of pairwise dissimilarity rating such as the Euclidean distance.
 The most prominent criterion on the basis of pairwise dissimilarity ratings is the *diversity*, i.e., the total sum of dissimilarities between any elements within the same group.
@@
 Another important criterion based on pairwise dissimilarities is the *dispersion*, i.e., the minimum dissimilarity between any two elements within the same group [@fernandez2013; @brusco2019].
-Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements as dissimilar from each other as possible.
+Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
 
-Oftentimes, the data available to researchers does not represent pairwise similarity, but numeric, categorical or binary attribute values (or *features*; @dry2009).
-That is, the data is a two-way two-mode [@arabie1992] table where rows represent the entities of study, and columns represent variables describing the elements.
-For example, numeric ratings may describe the attributes of word stimuli for psycholinguistic experiments.
-Such data may include norms for imagery, concreteness and orthographic variables [@friendly1982].
-In other cases, attributes are binary and may represent the presence or absence of a facial feature [@tversky1977].
-In the context of anticlustering, one possible approach for handling two-way two-mode data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across a set of numeric attributes [@brusco2019].
-Subsequently, an anticlustering criterion such as the diversity can be optimized to partition the elements into similar groups.
-
-A different approach to anticlustering directly works with the two-way two-mode feature data by maximizing the k-means criterion.
+Oftentimes, researchers work with attribute values [or *features*; @dry2009] instead of dissimilarity ratings.
+This use case is for example common when selecting stimuli for an experiment on the basis of norming data [e.g., @kurdi2017introducing].
+For word stimuli in psycholinguistic experiments, norming data may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982].
+In other cases, attribute values are binary and may represent the presence or absence of a feature [@tversky1977].
+In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019].
+Subsequently, an anticlustering criterion such as the diversity can be optimized to partition the elements into groups.
+A different approach to anticlustering directly works with the attribute data by maximizing the k-means criterion.
 K-means is probably the best-known clustering method.
@@
 To optimize overall group similarity---and hence to overcome the limitations of k-means anticlustering---I present k-plus anticlustering in this paper.
-The k-plus criterion is an extension of the k-means objective that includes a term aiming at optimizing similarity with regard to the variance of data across groups.
+The k-plus criterion is an extension of the k-means objective that includes a term that quantifies similarity with regard to the variance of data across groups.
 Interestingly, this bicriterion reformulation can be understood as an augmentation of the original data matrix, leaving the original k-means objective unchanged.
-I show that the k-plus criterion outperforms other anticlustering criteria when aiming at minimizing differences with regard to the mean and the variance of the different groups.
+I show that the k-plus criterion outperforms other anticlustering criteria when the goal is to minimize differences with regard to the mean and the variance of groups.
 In most cases, both objectives can be optimized satisfactorily at the same time without any loss on either.
-Moreover, in two pratical examples I illustrate how readers can easily apply k-plus anticlustering using the free and open source software `anticlust`, which is an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
+Moreover, in two pratical examples I illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
 
@@
 
-It is unlikely that an exclusive optimization of $SSE_{Var}$ (i.e., equalizing variances) is of interest without also considering the $\mathit{SSE}$ (i.e., equalizing means).
-A combination of $SSE_{Var}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall cluster similarity.
+It is unlikely that an exclusive optimization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is of interest without also considering the $\mathit{SSE}$ (i.e., equalizing means).
+A combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall cluster similarity.
 The most simple approach for such a bicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
@@
 
-The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of 
-equalizing means and variances, respectively. Often, they are chosen such that 
-$w_1 + w_2 = 1$ [@brusco2019].
-
+The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
 Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
@@
 
-This formula again has the same form as $\mathit{SSE}$ (and $SSE_{Var}$), but is computed on different data.
+This formula again has the same form as $\mathit{SSE}$ (and $\mathit{SSE_{Var}}$), but is computed on different data.
 In this case, both the original data $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function.
@@
 Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ have much larger values, on average, than the raw values $x_{ij}$ (also see Figure 1).
-Therefore, the $SSE_{Var}$ criterion naturally receives greater weight when computing $\mathit{SSE_{kplus}}$ than the standard $\mathit{SSE}$ criterion.
+Therefore, the $\mathit{SSE_{Var}}$ criterion naturally receives greater weight when computing $\mathit{SSE_{kplus}}$ than the standard $\mathit{SSE}$ criterion.
 This imbalance is most likely unwarranted because optimizing similarity with regard to variances is not more important than optimizing similarity with regard to means.
@@
 
-When optimizing a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $\mathit{SSE}$ and $SSE_{Var}$ are close to the global optimum value.
+When optimizing a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are close to the global optimum value.
 Generally, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree [@ferligoj1992direct].
 Usually, a tradeoff is necessary.
-For anticlustering, a motivating example illustrates that aiming to simultaneously optimize similarity of means and variances is not only desirable, but may actually be feasible.
+However, a motivating example illustrates that aiming to simultaneously optimize similarity of means and variances is not only desirable, but may actually be feasible.
 
@@
 
-For this motivating example, I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
-Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $SSE_{Var}$ across partitions.
+For this example I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
+Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
 Note that for both criteria, a larger value is better, i.e., indicates higher similarity with regard to means and variances, respectively.
@@
 However, optimal k-means anticlustering would create two groups that are not particularly similar with regard to their variance.
-The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $SSE_{Var}$.
-Similarly, the partition that maximizes $SSE_{Var}$ (the triangular symbol in Figure 1) is not well-suited to optimize SSE at the same time.
-The diamond symbol represents the partition that maximizes (the unweighted) $SSE_{kplus}$, i.e., the sum of $\mathit{SSE}$ and $SSE_{Var}$.
-This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $SSE_{Var}$.
+The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$.
+Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (the triangular symbol in Figure 1) is not well-suited to optimize SSE at the same time.
+The diamond symbol represents the partition that maximizes (the unweighted) $SSE_{kplus}$, i.e., the sum of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$.
+This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
 

### 2021-10-08 paper.Rmd
working on results plot

@@
 options(digits = 2)
-knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE)
+knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE, dev = "cairo_pdf")
 
@@
 To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
-This simulation employed five repetitions of the local maximum search.
+The simulation employed five repetitions of the local maximum search.
 
@@
 ## Results
+
+```{r, fig.cap = "Displays how well maximizing the k-plus criterion approximates the k-means criterion that is found via direct maximization of the k-means criterion.", fig.height = 5, fig.width = 5}
+
+# Load required packages
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+
+## Analyze data for K = 2 and K = 3 and K = 4
+simulation_results <- list()
+for (K in 2:4) {
+  filename <- paste0("./Simulation_Study/results-K", K, "-objectives-raw.csv")
+  df <- read.csv(filename, sep = ";", stringsAsFactors = FALSE)
+  df$K <- K
+  simulation_results[[paste0("K-", K)]] <- df
+}
+
+df <- do.call(rbind, simulation_results)
+rownames(df) <- NULL
+
+# Make long format
+ldf <- pivot_longer(
+  df,
+  cols = paste0(c("kvar", "kmeans", "means", "sd"), "_obj"),
+  names_to = "Objective",
+  names_pattern = "(.*)_obj"
+)
+
+# Check how well k-plus approximates k-means
+rel <- ldf %>% 
+  group_by(method, Objective, N, K) %>% 
+  summarise(Mean = mean(value)) %>% 
+  filter(Objective == "kmeans", method %in% c("kplus", "k-means-exchange")) %>% 
+  pivot_wider(names_from = method, values_from = Mean) %>% 
+  group_by(N, K) %>% 
+  summarise(rel = kplus / `k-means-exchange`)
+
+rel %>% 
+  ggplot(aes(x = N, y = rel, colour = factor(K), linetype = factor(K))) + 
+  geom_line(size = 1) +
+  geom_point(aes(colour = factor(K), shape = factor(K)), size = 3) +
+  scale_x_continuous(breaks = seq(36, 300, 24)) + 
+  theme_apa() +
+  theme(
+    legend.position = "top",
+    legend.title = element_text()
+  ) +
+  scale_colour_grey(name = "K") +
+  guides(
+    linetype = guide_legend(title = "K"),
+    shape = guide_legend(title = "K")
+  ) + 
+  labs(y = "k-plus approximation of k-means criterion")
+
+# approximation at N > 100
+# mean(filter(rel, N > 100)$rel) * 100
+
+```
+
+First, I investigated how well maximizing the k-plus criterion approximates the ordinary k-means criterion at the same time.
+Figure 2 illustrates the relative values of $\mathit{SSE}$ that were found when optimizing $\mathit{SSE_{kplus}}$ as compared to the direct optimization of $\mathit{SSE}$.
+Generally, optimizing $\mathit{SSE_{kplus}}$ yields an adequate approximation of $\mathit{SSE}$.
+However, the quality of this approximation depends on the group sizes of the partition problem: For larger $N$, k-plus anticlustering very closely follows k-means anticlustering's performance with regard to optimizing $\mathit{SSE}$.
+For $N > 100$, $\mathit{SSE}$ was approximated to more than 99.99% when conducting k-plus anticlustering as compared to k-means anticlustering.
+The minimum approximation for N = 24 and K = 4 was ...
 

### 2021-10-11 paper.Rmd
some results stuff

@@
 library("knitr")
+library("dplyr")
+library("ggplot2")
+library("tidyr")
 options(digits = 2)
@@
 
-# Load required packages
-library(dplyr)
-library(ggplot2)
-library(tidyr)
-
 ## Analyze data for K = 2 and K = 3 and K = 4
@@
 # approximation at N > 100
-# mean(filter(rel, N > 100)$rel) * 100
+approx_n100 <- mean(filter(rel, N > 100)$rel) 
+approx_n24_k4 <- mean(filter(rel, N == 24, K == 4)$rel)
+
+# cut decimals after n decimals
+cut_decimals <- function(x, n) {
+  x <- as.character(x)
+  x1 <- strsplit(x, "\\.")[[1]][1]
+  x2 <- strsplit(x, "\\.")[[1]][2]
+  x2 <- strsplit(x2, "")[[1]]
+  as.numeric(paste0(x1, ".", paste(x2[1:n], collapse = "")))
+}
 
@@
 First, I investigated how well maximizing the k-plus criterion approximates the ordinary k-means criterion at the same time.
-Figure 2 illustrates the relative values of $\mathit{SSE}$ that were found when optimizing $\mathit{SSE_{kplus}}$ as compared to the direct optimization of $\mathit{SSE}$.
-Generally, optimizing $\mathit{SSE_{kplus}}$ yields an adequate approximation of $\mathit{SSE}$.
-However, the quality of this approximation depends on the group sizes of the partition problem: For larger $N$, k-plus anticlustering very closely follows k-means anticlustering's performance with regard to optimizing $\mathit{SSE}$.
-For $N > 100$, $\mathit{SSE}$ was approximated to more than 99.99% when conducting k-plus anticlustering as compared to k-means anticlustering.
-The minimum approximation for N = 24 and K = 4 was ...
+Figure 2 illustrates the relative values of $\mathit{SSE}$ that were found when optimizing $\mathit{SSE_{kplus}}$---as compared to the direct optimization of $\mathit{SSE}$.
+Generally, optimizing $\mathit{SSE_{kplus}}$ yielded an adequate approximation of $\mathit{SSE}$.
+In particular, for large group sizes, there is virtually no sacrifice in $\mathit{SSE}$ when  $\mathit{SSE_{kplus}}$ is optimized: For $N > 100$, $\mathit{SSE}$ was approximated to more than `r cut_decimals(approx_n100, 4) * 100`% when conducting k-plus anticlustering as compared to k-means anticlustering.
+In the simulation, the worst approximation was found when the group sizes were smallest: for $N = 24$ and $K = 4$ was approximated to $\mathit{SSE}$ `r approx_n24_k4 * 100`%.
+Thus, the quality of the approximation depended on the group size: smaller groups led to worse approximations, while in general there was very little in $\mathit{SSE}$, expecially for larger groups.
+
+In a next step, I compared k-means anticlustering, k-plus anticlustering and anticluster editing in their ability to equalize means and standard deviations between groups.
+
+```{r, fig.cap = "Awesome caption", fig.height = 6.5, fig.width = 5}
+
+ldf %>% 
+  group_by(method, Objective, N, K) %>% 
+  summarise(Mean = mean(value)) %>% 
+  filter(Objective %in% c("means", "sd"), method != "random") %>% 
+  ggplot(aes(x = N, y = Mean, colour = method)) + 
+  geom_line(aes(colour = method), size = .7) + 
+  geom_point(aes(colour = method, shape = method)) +
+  facet_grid(cols = vars(Objective), rows = vars(K), scales = "free") + 
+  theme_apa() +
+  theme(
+    legend.position = "top",
+    legend.title = element_text()
+  ) + 
+  guides(
+    linetype = guide_legend(title = "Method"),
+    shape = guide_legend(title = "Method")
+  ) +
+  theme(
+    legend.position = "top",
+    legend.title = element_text()
+  ) +
+  scale_colour_grey(name = "Method")
+
+```
 

### 2021-10-22 paper.Rmd
working on results for simulation, using Table

@@
 library("tidyr")
+library("prmisc") # remotes::install_github("m-Py/prmisc")
 options(digits = 2)
@@
 In the simulation, the worst approximation was found when the group sizes were smallest: for $N = 24$ and $K = 4$ was approximated to $\mathit{SSE}$ `r approx_n24_k4 * 100`%.
-Thus, the quality of the approximation depended on the group size: smaller groups led to worse approximations, while in general there was very little in $\mathit{SSE}$, expecially for larger groups.
+Thus, the quality of the approximation depended on the group size: smaller groups led to worse approximations, while in general there was very little cost with regard $\mathit{SSE}$ when optimizing $\mathit{SSE_{kplus}}$, expecially for larger groups.
 
 In a next step, I compared k-means anticlustering, k-plus anticlustering and anticluster editing in their ability to equalize means and standard deviations between groups.
+The results are shown in Table 1.
 
-```{r, fig.cap = "Awesome caption", fig.height = 6.5, fig.width = 5}
+```{r, fig.cap = "Awesome caption"}
 
-ldf %>% 
-  group_by(method, Objective, N, K) %>% 
+# Recode N into categories so the results can be displayed in a more informative way
+ldf <- within(ldf, {
+  N_category <- ifelse(N <= 200, "N <= 200", "N > 200")
+  N_category <- ifelse(N <= 100, "N <= 100", N_category)
+  N_category <- ifelse(N <= 50,  "N <= 50", N_category)
+})
+
+ldf$N_category <- ordered(
+  ldf$N_category, 
+  levels = c("N <= 50", "N <= 100", "N <= 200", "N > 200")
+)
+
+# for sorting by method, make it ordered factor:
+ldf$method <- ordered(
+  ldf$method, 
+  levels = c("k-means-exchange", "kplus", "ace-exchange"),
+  labels = c("k-means", "k-plus", "anticluster editing")
+)
+
+results_table <- ldf %>% 
+  group_by(method, Objective, N_category) %>% 
   summarise(Mean = mean(value)) %>% 
   filter(Objective %in% c("means", "sd"), method != "random") %>% 
-  ggplot(aes(x = N, y = Mean, colour = method)) + 
-  geom_line(aes(colour = method), size = .7) + 
-  geom_point(aes(colour = method, shape = method)) +
-  facet_grid(cols = vars(Objective), rows = vars(K), scales = "free") + 
-  theme_apa() +
-  theme(
-    legend.position = "top",
-    legend.title = element_text()
-  ) + 
-  guides(
-    linetype = guide_legend(title = "Method"),
-    shape = guide_legend(title = "Method")
-  ) +
-  theme(
-    legend.position = "top",
-    legend.title = element_text()
-  ) +
-  scale_colour_grey(name = "Method")
+  pivot_wider(names_from = Objective, values_from = Mean) %>%
+  arrange(N_category, method)
+
+# enforce two decimals
+results_table$means <- force_decimals(results_table$means, 2)
+results_table$sd <- force_decimals(results_table$sd, 2)
+
+# Generate table
+footnote <- "$\\Delta M$ and $\\Delta \\mathit{SD}$ quantify how strongly means and standard deviations differed between anticlusters.
+Table cells contain the average objectives across simulation runs, split by $N$."
+
+results_table$N_category <- NULL
+
+colnames <- c("", "$\\Delta M$", "$\\Delta \\mathit{SD}$")
+colnames(results_table) <- colnames
+
+apa_table(
+  results_table, 
+  caption = "Results of the simulation study.", 
+  note = footnote, 
+  escape = FALSE,
+  stub_indents = list("\\textbf{$\\textit{N} \\leq 50$}" = 1:3,
+                      "\\textbf{$50 <\\textit{N} \\leq 100$}" = 4:6,
+                      "\\textbf{$100 <\\textit{N} \\leq 200$}" = 7:9,
+                      "\\textbf{$\\textit{N} > 200$}" = 10:12),
+  span_text_columns = TRUE
+)
 
@@
 
-- k-plus approximates k-means very well, especially for large $N$ (for larger $K$, $N$ schould increase as well)
 - k-plus is the best method to optimize similarity with regard to means and standard deviations at the same time
@@
 The simulation study indicated that k-plus may show decreased performance for smaller group sizes.
-Therefore,    this second example application illustrates that k-plus anticlustering may also yield satisfactory results when group are smaller than in the OASIS example.
-I employ a norming data set of 96 word stimuli that was contributed to the `anticlust` package by Dr.
+Therefore, a second application illustrates that k-plus anticlustering may yield satisfactory results in practice, when group are smaller.
+I make use of a norming data set of 96 word stimuli that was contributed to the `anticlust` package by Dr.
 Schaper [@schaper2019metacognitive; @schaper2019metamory; also see @papenberg2020 for a description of the norming data].

### 2021-10-22 paper.Rmd
'finished' results of simulation study

@@
 To systematically evaluate the new k-plus criterion, I compared three anticlustering approaches in a simulation study: (a) k-means anticlustering, optimizing the $\mathit{SSE}$; (b) k-plus anticlustering, optimizing the $SSE_{kplus}$; (c) anticluster editing, optimizing the diversity criterion.
-All methods employed an a priori standardization of the data sets.
-In the case of k-plus anticlustering, this means that the augmented input data (i.e., the original data and the squared distance data) were subjected to a standardization, which is recommended.
+All methods used a standardization of the data sets as a preprocessing step.
+In the case of k-plus anticlustering, this means that the augmented input data (i.e., the data table that contains the original data and the squared distance data) were subjected to a standardization, which is recommended.
 
@@
 
-```{r, fig.cap = "Displays how well maximizing the k-plus criterion approximates the k-means criterion that is found via direct maximization of the k-means criterion.", fig.height = 5, fig.width = 5}
+```{r, fig.cap = "Displays how well maximizing the k-plus criterion approximates the k-means criterion that is found via direct maximization of the k-means criterion.", fig.height = 4.5, fig.width = 4.5}
 
@@
 
-First, I investigated how well maximizing the k-plus criterion approximates the ordinary k-means criterion at the same time.
+First, I investigated how well maximizing the k-plus criterion ($\mathit{SSE_{kplus}}$) approximates the ordinary k-means criterion ($\mathit{SSE}$) at the same time.
 Figure 2 illustrates the relative values of $\mathit{SSE}$ that were found when optimizing $\mathit{SSE_{kplus}}$---as compared to the direct optimization of $\mathit{SSE}$.
@@
 In a next step, I compared k-means anticlustering, k-plus anticlustering and anticluster editing in their ability to equalize means and standard deviations between groups.
-The results are shown in Table 1.
+To clarify the general pattern of results, Table 1 illustrates the outcome variables  $\Delta M$ and $\Delta \mathit{SD}$ after splitting the sample size into categories and averaging across simulation runs.
+Note that the results have also been collapsed across the different group sizes ($K = 2, 3, 4$) because the relative performance of the anticlustering methods did not vary by $K$.
+Only a main effect was observed: across all methods, more groups led to higher discrepancies between means and standard deviations between groups.
+Table 1 shows that k-means best equalizes means between groups; its advantage is most pronounced for smaller samples sizes.
+For larger sample sizes, both k-plus anticlustering and anticluster editing perform similarly well with regard to $\Delta M$.
+Replicating @papenberg2020, it is shown that anticluster editing better minimizes differences with regard to standard deviations between groups than k-means anticlustering.
+However, noteworthy, k-plus anticlustering strongly outperforms both competitors in this matter; $\Delta \mathit{SD}$ is consistently lowest for the new k-plus method.
+In conclusion, k-plus anticlustering is well suited to minimize differences with regard to mean values across groups---especially when groups are larger---and at the same time is able to minimize differences in standard deviations between groups.
 
@@
 
-- k-plus is the best method to optimize similarity with regard to means and standard deviations at the same time
-
 # K-plus anticlustering using the R package anticlust
 
-In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020].
+In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] and show that k-plus anticlustering yields excellent results when being applied to real data sets.
 The results presented here used `anticlust` version 0.5.6.
@@
 
-Table 1 shows the results of the anticlustering application by listing the descriptive statistics (means and standard deviations) for each of the 9 groups and for each of the 3 features.
-Additionaly, Table 1 shows the descriptive statistics for an application of k-means anticlustering and anticluster editing (which maximizes the diversity criterion).
+Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means and standard deviations) for each of the 9 groups and for each of the 3 features.
+Additionaly, Table 2 shows the descriptive statistics for an application of k-means anticlustering and anticluster editing (which maximizes the diversity criterion).
 It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations.
@@
 
-Table 2 illustrates an advantage of k-plus anticlustering over anticluster editing: When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups.
+Table 3 illustrates an advantage of k-plus anticlustering over anticluster editing: When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups.
 K-plus anticlustering strives for equal means and variances in all groups, regardless of group size.
@@
 
-Table 3 illustrates the results that are obtained when using k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
+Table 4 illustrates the results that are obtained when using k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
 Even though the means and standard deviations are no longer perfectly matched between groups, arguably, they are still similar enough for any practical purpose---especially when considering inevitable measurement error in norming studies.
-The following code was used to achieve the results reported in Table 3:
+The following code was used to achieve the results reported in Table 4:
 

### 2021-10-22 paper.Rmd
minor work

@@
 The k-plus objective extends k-means by adding a term that aims to optimize similarity with regard to the variance of the attributes data across clusters.
-It is shown that the combined bicriterion objective can be reduced to an augmentation of the input data, leaving the original k-means criterion unchanged.
-A computer simulation and practical examples on real data show that k-plus anticlustering usually achieves high similarity both with regard to mean attribute values as well as the variance, with no noticable sacrifice to either criterion.
+It is shown that the combined bicriterion objective can be reduced to an augmentation of the input data, leaving the original k-means criterion unchanged: For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the original values.
+A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity both with regard to mean attribute values as well as the variance, usually with no noticable sacrifice to either criterion.
 The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity.
@@
 
-First, I generated 10,000 data sets which were subsequently subjected to the three anticlustering methods.
+First, I generated 10,000 data sets which were subsequently processed by to the three competing anticlustering methods.
 The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features $M$ varied between 1 and 10; (c) the distribution of features was randomly determined to be either (i) uniform in [0, 1], (ii) a standard normal distribution with *M* = 0 and *SD* = 1, or (iii) a wider normal distribution with *M* = 0 and *SD* = 2.
 For a given data set, all $M$ features were independently sampled from the given distribution and no correlations were assumed between features.
-All three anticlustering methods were applied to each of the 10,000 data sets using $K = 2$, $K = 3$ and $K = 4$; groups sizes were always equal.
+All three anticlustering methods were applied to each of the 10,000 data sets using $K = 2$, $K = 3$ and $K = 4$.
+Groups sizes were always equal.
 
@@
 To clarify the general pattern of results, Table 1 illustrates the outcome variables  $\Delta M$ and $\Delta \mathit{SD}$ after splitting the sample size into categories and averaging across simulation runs.
-Note that the results have also been collapsed across the different group sizes ($K = 2, 3, 4$) because the relative performance of the anticlustering methods did not vary by $K$.
-Only a main effect was observed: across all methods, more groups led to higher discrepancies between means and standard deviations between groups.
-Table 1 shows that k-means best equalizes means between groups; its advantage is most pronounced for smaller samples sizes.
-For larger sample sizes, both k-plus anticlustering and anticluster editing perform similarly well with regard to $\Delta M$.
+Note that the results have also been collapsed across the different group sizes ($K = 2, 3, 4$) because the relative performance of the anticlustering methods did not vary by $K$.[^kmaineffect] Table 1 shows that k-means was best suited to equalize mean values between groups; this advantage was most pronounced for smaller samples sizes.
+For larger sample sizes, k-plus anticlustering and anticluster editing performed similarly well with regard to $\Delta M$.
 Replicating @papenberg2020, it is shown that anticluster editing better minimizes differences with regard to standard deviations between groups than k-means anticlustering.
-However, noteworthy, k-plus anticlustering strongly outperforms both competitors in this matter; $\Delta \mathit{SD}$ is consistently lowest for the new k-plus method.
+Noteworthy, however, k-plus anticlustering strongly outperformed both competitors in this matter; $\Delta \mathit{SD}$ was consistently lowest for the new k-plus method.
 In conclusion, k-plus anticlustering is well suited to minimize differences with regard to mean values across groups---especially when groups are larger---and at the same time is able to minimize differences in standard deviations between groups.
+
+[^kmaineffect]: Only a main effect was observed: across all anticlustering methods, more groups led to higher discrepancies between means and standard deviations between groups.
 
@@
 
-In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] and show that k-plus anticlustering yields excellent results when being applied to real data sets.
+In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020].
+It is shown that k-plus anticlustering yields excellent results when being applied to real data sets.
 The results presented here used `anticlust` version 0.5.6.
@@
 
-In this application, I use norming data for the OASIS image data set that is freely available online [@kurdi2017introducing].
+In the first example, I use norming data for the OASIS image data set that is freely available online [@kurdi2017introducing].
 @kurdi2017introducing assembled 900 open-access color images and collected ratings on two affective dimensions: arousal and valence.
@@
 The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the `features` variable (**TODO**).
-The arguments `objective = "kplus"` and `method = "local-maximum"` ensure that the k-plus criterion is optimized using the local maximum search method described in the previous section (**TODO**).
+The arguments `objective = "kplus"` and `method = "local-maximum"` ensure that the k-plus criterion is optimized using the local maximum search method described in the previous section.
 The argument `K` describes the number of groups.
@@
 Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means and standard deviations) for each of the 9 groups and for each of the 3 features.
-Additionaly, Table 2 shows the descriptive statistics for an application of k-means anticlustering and anticluster editing (which maximizes the diversity criterion).
+Additionaly, Table 2 shows the descriptive statistics for an application of k-means anticlustering and anticluster editing.
 It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations.
@@
 Anticluster editing was also well-suited to match both means and standard deviations, but was slightly outperformed by k-plus.
-The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to the means and standard deviations, but instead optimizes the within-group heterogeneity.
+The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to the means and standard deviations, but instead maximizes within-group heterogeneity.
 
@@
 
-Table 3 illustrates an advantage of k-plus anticlustering over anticluster editing: When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups.
+Table 3 illustrates another advantage of k-plus anticlustering over anticluster editing: When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups.
 K-plus anticlustering strives for equal means and variances in all groups, regardless of group size.
@@
 
-## Example Application II: Small data set (N = 96)
+## Example Application II: Small data set ($N = 96$)
 
 The simulation study indicated that k-plus may show decreased performance for smaller group sizes.
-Therefore, a second application illustrates that k-plus anticlustering may yield satisfactory results in practice, when group are smaller.
+Therefore, a second application illustrates that k-plus anticlustering yields satisfactory results in practice even when group sizes are smaller.
 I make use of a norming data set of 96 word stimuli that was contributed to the `anticlust` package by Dr.
@@
 Table 4 illustrates the results that are obtained when using k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
-Even though the means and standard deviations are no longer perfectly matched between groups, arguably, they are still similar enough for any practical purpose---especially when considering inevitable measurement error in norming studies.
+Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are still similar enough for any practical purpose---especially when considering inevitable measurement error in norming studies.
 The following code was used to achieve the results reported in Table 4:
@@
 
-Note that standardization is particularly useful for anticlustering when the score range differs strongly between features, as is the case for the Schaper data set.
-
 ```{r}
@@
   tab,
-  caption = "Descriptive statistics by group for the Schaper et al. (2019) data set, after applying k-plus anticlustering.",
+  caption = "Descriptive statistics by group for the Schaper et al. (2019a; 2019b) data set, after applying k-plus anticlustering.",
   note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
@@
 
-- k-plus anticlustering is useful and should in the default case be preferred to k-means anticlustering.
+In this paper, I presented k-plus anticlustering, a method to divide a data set into similar subgroups.
+The k-plus criterion extends the classical k-means criterion by incorporating a term that represents how groups differ with regard to their variances.
+The combined bicriterion objective could be reduced to an augmentation of the input data; for each input variable, a new variable can be computed as the squared deviation between the original values and the overall mean of the original variable.
+A simulation study and examples on norming data indicate that optimizing the new k-plus objective is well-suited for minimizing differences with regard to means and variances between groups.
+
+K-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups.
+For small samples sizes (or large $K$), k-plus anticlustering may return slightly less-than-optimal results in the eye of the user, but in this case it is no problem to return to k-means and only focus on the means.
+There is no anticluster hacking; as opposed to some statistical workflows where "mindlessly" trying different approaches until one approach yiels satisfactory results is bad practice, such a trial-and-error approach is unproblematic for anticlustering.
+Users may simply choose the results they like best.
+As a starter, I strongly recommend trying out k-plus anticlustering when creating groups that aim for high between-group similarity.
+
 For optimizing similarity in means and variances it also outperforms anticluster editing, which, to be fair, does not directly strive to do so.
 Unequal group sizes give another advantage for k-plus anticlustering in comparison to anticluster editing.
-For small samples sizes (or large $K$), k-plus anticlustering may return slightly less-than-optimal results in the eye of the user, but in this case it is no problem to return to k-means and only focus on the means.
-There is not "anticlustering hacking"; users may simply choose the results they like best.
-For now, I strongly recommend using k-plus anticlustering for creating groups with high between group similarity.
 

### 2021-10-22 paper.Rmd
some discussion, abstract

@@
 For some anticlustering formulations, these two objectives coincide, i.e., high between-group similarity is achieved when within-group heterogeneity is maximized.
-Other anticlustering criteria focus on one of the two aspects.
-The present paper focuces on maximizing between-group similarity by introducing a novel bicriterion objective for anticlustering.
-*K-plus anticlustering* is an extension of k-means anticlustering, which minimizes the difference between the mean of numeric attributes across clusters.
-The k-plus objective extends k-means by adding a term that aims to optimize similarity with regard to the variance of the attributes data across clusters.
-It is shown that the combined bicriterion objective can be reduced to an augmentation of the input data, leaving the original k-means criterion unchanged: For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the original values.
-A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity both with regard to mean attribute values as well as the variance, usually with no noticable sacrifice to either criterion.
+The present paper focuces on maximizing between-group similarity by introducing the k-plus objective, a novel bicriterion for anticlustering.
+K-plus anticlustering is an extension of classical k-means anticlustering.
+While k-means anticlustering focuses on minimizing between-group-differences with regard to the means of attribute values, k-plus anticlustering includes a term to optimize similarity with regard to the variance at the same time.
+The k-plus bicriterion can be reduced to an augmentation of the input data: For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the original variable.
+Thus, k-plus leaves the original k-means criterion unchanged and only differs in the input data.
+A computer simulation and practical examples on real norming data show that k-plus anticlustering achieves high similarity both with regard to mean attribute values as well as the variance, usually with no noticable sacrifice to either criterion.
 The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity.
-K-plus anticlustering is easily accessible to researchers and practitioners: the method has been implemented as part of the open source R package `anticlust` and is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
+K-plus anticlustering is easily accessible as it has been implemented as part of the open source R package `anticlust` and is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
 
@@
 
-In this paper, I presented k-plus anticlustering, a method to divide a data set into similar subgroups.
-The k-plus criterion extends the classical k-means criterion by incorporating a term that represents how groups differ with regard to their variances.
-The combined bicriterion objective could be reduced to an augmentation of the input data; for each input variable, a new variable can be computed as the squared deviation between the original values and the overall mean of the original variable.
-A simulation study and examples on norming data indicate that optimizing the new k-plus objective is well-suited for minimizing differences with regard to means and variances between groups.
+In this paper, I presented the k-plus criterion for anticlustering.
+Anticlustering methods are used to divide sets of elements into subgroups while striving for high between-group similarity or high within-group heterogeneity.
+The k-plus criterion represents between-group similarity as the discrepancy between the mean and variance of attribute values.
+These descriptive statistics are usually "eyeballed" by applicants of anticlustering methods when they try to assess the quality of of a partitioning; hence, a criterion that directly optimizes these measures has been needed and is now provided.
+The k-plus criterion extends the classical k-means criterion---which represents how similar groups are with regard to their means---by incorporating a term that represents how groups differ with regard to their variances.
+The combined bicriterion objective could be reduced to an augmentation of the input data; for each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the variable.
+Thus, the original k-means criterion remains basically unchanged.
+A simulation study and examples on norming data indicated that optimizing the new k-plus objective is well-suited for minimizing differences with regard to means and variances between groups.
 
-K-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups.
-For small samples sizes (or large $K$), k-plus anticlustering may return slightly less-than-optimal results in the eye of the user, but in this case it is no problem to return to k-means and only focus on the means.
-There is no anticluster hacking; as opposed to some statistical workflows where "mindlessly" trying different approaches until one approach yiels satisfactory results is bad practice, such a trial-and-error approach is unproblematic for anticlustering.
+This work indicates that k-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups, given that similarity in means and variances can usually be achieved succesfully at the same time.
+However, for small group sizes (small $N$ and / or large $K$) k-plus anticlustering may return slightly less-than-optimal results in the eye of the user.
+In this case, it is not a problem to return to k-means that only focuses on the means.
+As opposed to some statistical workflows where "mindlessly" trying different approaches until one approach yiels satisfactory results is bad practice, such a trial-and-error approach is unproblematic for anticlustering.
+There is no anticluster hacking.
 Users may simply choose the results they like best.
-As a starter, I strongly recommend trying out k-plus anticlustering when creating groups that aim for high between-group similarity.
+As a default, it is recommended to apply out k-plus anticlustering when creating groups that aim for high between-group similarity.
 

### 2021-10-22 paper.Rmd
typo

@@
 To clarify the general pattern of results, Table 1 illustrates the outcome variables  $\Delta M$ and $\Delta \mathit{SD}$ after splitting the sample size into categories and averaging across simulation runs.
-Note that the results have also been collapsed across the different group sizes ($K = 2, 3, 4$) because the relative performance of the anticlustering methods did not vary by $K$.[^kmaineffect] Table 1 shows that k-means was best suited to equalize mean values between groups; this advantage was most pronounced for smaller samples sizes.
-For larger sample sizes, k-plus anticlustering and anticluster editing performed similarly well with regard to $\Delta M$.
+Note that the results have also been collapsed across the different group sizes ($K = 2, 3, 4$) because the relative performance of the anticlustering methods did not vary by $K$.[^kmaineffect] Table 1 shows that k-means was best suited to equalize mean values between groups; this advantage was most pronounced for smaller group sizes.
+For larger group sizes, k-plus anticlustering and anticluster editing performed similarly well with regard to $\Delta M$.
 Replicating @papenberg2020, it is shown that anticluster editing better minimizes differences with regard to standard deviations between groups than k-means anticlustering.

### 2021-11-23 paper.Rmd
some unordered stuff

@@
 
-Anticlustering methods can be distinguished by the input data (pairwise dissimilarity vs.
-attributes values) and the objective function that is used to optimize group between-group similarity or within-group heterogeneity.
+Anticlustering methods can be distinguished by the input data and the objective function that is used to optimize group between-group similarity or within-group heterogeneity.
 Some anticlustering objectives quantify group similarity on the basis of pairwise dissimilarity rating such as the Euclidean distance.
-The most prominent criterion on the basis of pairwise dissimilarity ratings is the *diversity*, i.e., the total sum of dissimilarities between any elements within the same group.
-Maximizing the diversity simultaneously optimizes within-group heterogeneity and between-group similarity.
+The most prominent criterion on the basis of pairwise dissimilarity ratings is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
+Maximizing the diversity optimizes within-group heterogeneity, which simultaneously leads to similar groups---if all groups have the same size.
+Anticlustering objective functions usually represent the reversal of a clustering objective, only the direction of the optimization is changed.
 Because the diversity is the reversal of the clustering editing objective [@zahn1964; @shamir2004vf], @papenberg2020 used the term *anticluster editing* to refer to anticlustering approaches that optimize the diversity.
@@
 When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [e.g., @anscombe1973].
-@papenberg2020, showed that maximizing the diversity (based on the Euclidean distance) is better suited to minimize difference with regard to both the mean and the variance of the data; it was shown that even an entirely random split usually resulted in more similar variances because k-means solely focuses on the mean similarity.
+@papenberg2020, showed that maximizing the diversity (based on the Euclidean distance) is better suited to minimize difference with regard to both the mean and the variance of the data; even an entirely random split usually resulted in more similar variances because k-means solely focuses on means.
 
@@
 
-```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all partitions (for $N = 14$ and $K = 2$).
+```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
 The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$ and represents a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
@@
 To systematically evaluate the new k-plus criterion, I compared three anticlustering approaches in a simulation study: (a) k-means anticlustering, optimizing the $\mathit{SSE}$; (b) k-plus anticlustering, optimizing the $SSE_{kplus}$; (c) anticluster editing, optimizing the diversity criterion.
-All methods used a standardization of the data sets as a preprocessing step.
+All methods used a $z$-standardization of the data sets as a preprocessing step.
 In the case of k-plus anticlustering, this means that the augmented input data (i.e., the data table that contains the original data and the squared distance data) were subjected to a standardization, which is recommended.
@@
 In this paper, I presented the k-plus criterion for anticlustering.
-Anticlustering methods are used to divide sets of elements into subgroups while striving for high between-group similarity or high within-group heterogeneity.
-The k-plus criterion represents between-group similarity as the discrepancy between the mean and variance of attribute values.
+K-plus anticlustering focuses on constructing groups that are highly similar to each other, operationalized by similar means as well as variances.
 These descriptive statistics are usually "eyeballed" by applicants of anticlustering methods when they try to assess the quality of of a partitioning; hence, a criterion that directly optimizes these measures has been needed and is now provided.
-The k-plus criterion extends the classical k-means criterion---which represents how similar groups are with regard to their means---by incorporating a term that represents how groups differ with regard to their variances.
-The combined bicriterion objective could be reduced to an augmentation of the input data; for each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the variable.
-Thus, the original k-means criterion remains basically unchanged.
-A simulation study and examples on norming data indicated that optimizing the new k-plus objective is well-suited for minimizing differences with regard to means and variances between groups.
+The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means.
+Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data.
+For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the variable.
+A simulation study and examples on real norming data indicate that minimizing differences with regard to means and variances between groups is well possible using the k-plus criterion.
 

### 2022-02-24 paper.Rmd
extending kplus logic to higher order moments and covariances

@@
 ---
-title             : "k-plus: a bicriterion extension of k-means anticlustering"
+title             : "k-plus Anticlustering: An Improved k-means Criterion for Maximizing Between-Group Similarity"
 shorttitle        : "k-plus anticlustering"
@@
 
-    Anticlustering refers to a collection of methods that partition data sets into disjunct groups while aiming for two objectives: (a) high between-group similarity and (b) high within-group heterogeneity.
-For some anticlustering formulations, these two objectives coincide, i.e., high between-group similarity is achieved when within-group heterogeneity is maximized.
-The present paper focuces on maximizing between-group similarity by introducing the k-plus objective, a novel bicriterion for anticlustering.
-K-plus anticlustering is an extension of classical k-means anticlustering.
-While k-means anticlustering focuses on minimizing between-group-differences with regard to the means of attribute values, k-plus anticlustering includes a term to optimize similarity with regard to the variance at the same time.
-The k-plus bicriterion can be reduced to an augmentation of the input data: For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the original variable.
-Thus, k-plus leaves the original k-means criterion unchanged and only differs in the input data.
-A computer simulation and practical examples on real norming data show that k-plus anticlustering achieves high similarity both with regard to mean attribute values as well as the variance, usually with no noticable sacrifice to either criterion.
+    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity and high within-group heterogeneity.
+Anticlustering is usually approached by optimizing a reversed clustering objective such as the k-means criterion.
+Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity.
+While k-means anticlustering minimizes between-group-differences only with regard to mean attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, covariance structure, and higher order moments.
+To minimize differences in variance, each variable is duplicated by computing the squared deviation between the original values and the variable's mean.
+Because k-means anticlustering minimizes differences with regard to variables' means, appending the duplicated variables to the original data set---and then applying standard k-means anticlustering---suffices to equalize the variance between groups.
+Following a similar logic, differences in skewness, kurtosis and covariances can be minimized.
+Thus, while k-plus is a separate criterion for anticlustering, it can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
+A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity with regard to mean attribute values as well as the variance and higher order moments, usually with no noticable sacrifice to either criterion.
 The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity.
-K-plus anticlustering is easily accessible as it has been implemented as part of the open source R package `anticlust` and is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
+Example are given on how k-plus anticlustering can be applied using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
 
@@
 
-keywords          : "Anticlustering, k-means, k-plus, variance, bicriterion optimization"
+keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, higher order moments"
 wordcount         : "X"
@@
 To optimize overall group similarity---and hence to overcome the limitations of k-means anticlustering---I present k-plus anticlustering in this paper.
-The k-plus criterion is an extension of the k-means objective that includes a term that quantifies similarity with regard to the variance of data across groups.
-Interestingly, this bicriterion reformulation can be understood as an augmentation of the original data matrix, leaving the original k-means objective unchanged.
+The k-plus criterion is an extension of the k-means objective that includes a term that quantifies similarity with regard to the variance, covariances, and higher order moments of data across groups.
+Interestingly, this extended criterion can be understood as an augmentation of the original data matrix, leaving the original k-means objective unchanged.
 I show that the k-plus criterion outperforms other anticlustering criteria when the goal is to minimize differences with regard to the mean and the variance of groups.
@@
 In the next section I present a simulation study that more systematically investigates the performance of the k-plus criterion.
+
+## Skewness and kurtosis
+
+## Covariance structure
 

### 2022-02-24 paper.Rmd
removed unnecessary abbreviation

@@
 First, I generated 10,000 data sets which were subsequently processed by to the three competing anticlustering methods.
-The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features $M$ varied between 1 and 10; (c) the distribution of features was randomly determined to be either (i) uniform in [0, 1], (ii) a standard normal distribution with *M* = 0 and *SD* = 1, or (iii) a wider normal distribution with *M* = 0 and *SD* = 2.
+The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features varied between 1 and 10; (c) the distribution of features was randomly determined to be either (i) uniform in [0, 1], (ii) a standard normal distribution with *M* = 0 and *SD* = 1, or (iii) a wider normal distribution with *M* = 0 and *SD* = 2.
 For a given data set, all $M$ features were independently sampled from the given distribution and no correlations were assumed between features.

### 2022-02-25 scripts/Example-4-OASIS-Skew_Kurtosis.R
general function for any higher order moments

@@ -5,26 +5,17 @@ library(DescTools) # for computing skew and kurtosis
 oasis <- read.csv("https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv")
 
 # function to compute features for variance
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-# function to compute features for skewness
-third_moment_features <- function(data) {
-  apply(data, 2, function(x) ((x - mean(x))/sd(x))^3 )
-}
-
-# function to compute features for skewness
-fourth_moment_features <- function(data) {
-  apply(data, 2, function(x) ((x - mean(x))/sd(x))^4 )
+# moment = 2 -> variance; 3 = skew; 4 = kurtosis
+moment_features <- function(data, moment = 2) {
+  apply(data, 2, function(x) (x - mean(x))^moment)
 }
 
 features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
-extended <- cbind(features, squared_from_mean(features))
+extended <- cbind(features, moment_features(features))
 extended_skew_kurtosis <- cbind(
   extended, 
-  third_moment_features(features),
-  fourth_moment_features(features)
+  moment_features(features, 3),
+  moment_features(features, 4)
 )
 
 # standard k-plus anticlustering
@@ -51,4 +42,3 @@ by(features, anticlusters_cov, function(x) round(apply(x, 2, Kurt), 2))
 mean_sd_tab(features, anticlusters_skew_kurtosis)
 mean_sd_tab(features, anticlusters_cov)
 
-table(oasis$ac, oasis$Category)

### 2022-02-25 paper.Rmd
minor changes in abstract

@@
     Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity and high within-group heterogeneity.
-Anticlustering is usually approached by optimizing a reversed clustering objective such as the k-means criterion.
+Anticlustering is usually approached by maximzing instead of minimizing a clustering objective such as the k-means criterion.
 Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity.
@@
 To minimize differences in variance, each variable is duplicated by computing the squared deviation between the original values and the variable's mean.
-Because k-means anticlustering minimizes differences with regard to variables' means, appending the duplicated variables to the original data set---and then applying standard k-means anticlustering---suffices to equalize the variance between groups.
+Because the variance is defined as the expected value of the squared deviations from the mean, appending these duplicated variables to the data set---and then applying standard k-means anticlustering---suffices to equalize the variance between groups.
 Following a similar logic, differences in skewness, kurtosis and covariances can be minimized.

### 2022-02-28 paper.Rmd
started writing about skew kurtosis etc.

@@
 
-## Skewness and kurtosis
+## Higher order moments, skewness and kurtosis
+
+The outlined logic to equalize variance in different groups is easily generalized to higher order moments such as skewness and kurtosis, i.e., the third and fourth distribution moments, respectively.
+Generally, the *j*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ is computed as [@joanes1998]
+
+$$
+\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^j
+$$
+
+where $j = 2$ gives the variance, $j = 3$ the skewness, and $j = 4$ the kurtosis.
+
+The approach outlined in sections 1.1 and 1.2 illustrated how the definition of the variance can be used to add variables to the data set to minimize differences with regard to the variance between groups.
+Changing the power of the deviation between mean and data points, we can use the same logic to equalize any desired sample moment between groups.
+I expect this to be interesting in particular for the skewness and kurtosis to minimize differences with regard to distribution asymmetry (skewness) and the propensity to include outliers (kurtosis).
 
 ## Covariance structure
+
+Because the covariance is also defined as an expected value, the k-plus criterion can be extended to reflect differences in covariances between data sets.
+In this case, an additional variable has to be computed for each pair of variables, i.e., $\binom{N}{2}$ new variables, using the definition of the sample covariance:
+
+## The k-plus family of anticlustering objectives
+
+Some clarifications with respect to nomenclature are needed.
+The k-plus criterion is not a single objective function for anticlustering, but instead is a familiy of objectives that use the standard k-means criterion $SSE$ in order to minimize differences between groups with regard to additional characterizations of a data set; that is, not just the mean, but also the variance, covariances, or higher order moments such as skewness and kurtosis.
+Which of these objectives are pursued depends on a researchers needs.
+In the most basic---and arguably most important---case, k-plus can be used to not just minimize differences with regard to the mean of variables but also their variance.
+Measures of location (mean) and spread (variance / standard deviation) are arguably most important to researchers when they "eyeball" their data and they are routinely reported as characterizations of a data set.
+Hence, by default, the k-plus anticlustering objective implies that means and variances should be equal between groups.
+
+While less routinely reported (and less well understood), skewness are kurtosis are also important descriptive characterizations of a data set [@decarlo1997; @westfall2014].
+To optimize overall similarity between groups, it may thus be desirable to also minimize differences with regard to these distribution moments.
+In principle, any higher order moments can also be included as part of the k-plus objective.
+Note, however, that the inclusion each additional moment requires appending as many features to the data set as there were to begin with.
+
+K-plus anticlustering can also be used to obtain similar covariance structures among groups.
+This approach may be of interest in cross validation applications where prediction accuracy depends on the correlations among features and criteria [@zeng2000; @papenberg2020].
+Note that each additional objective that is pursued with k-plus anticlustering comes with a cost.
+It becomes more difficult to satisfactorily fulfil many criteria at the same time.
+As the simulation study below will show, including many k-plus objectives at the same time becomes more appropriate when the data set becomes larger.
+In the end, it is a tradeoff between fulfilling few objectives to a precise degree or fulfilling many objectives to a less compelling degree.
 

### 2022-02-28 example_tradeoff.Rmd
added example to potpurri

--- 2022-02-24 paper.Rmd
@@
----
-title             : "k-plus Anticlustering: An Improved k-means Criterion for Maximizing Between-Group Similarity"
-shorttitle        : "k-plus anticlustering"
-
-author:
-  - name          : "Martin Papenberg"
-    affiliation   : "1"
-    corresponding : yes
-    address       : "Heinrich Heine Universität Düsseldorf,  Institut für Experimentelle Psychologie, Universitätsstraße 1, 40225 Düsseldorf, Germany"
-    email         : "martin.papenberg@hhu.de"
-
-affiliation:
-  - id            : "1"
-    institution   : "Heinrich Heine University Düsseldorf"
-
-authornote: |
-
-  Martin Papenberg, Department of Experimental Psychology, Heinrich Heine University Düsseldorf.
-
-abstract: |
-
-    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity and high within-group heterogeneity.
-Anticlustering is usually approached by optimizing a reversed clustering objective such as the k-means criterion.
-Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity.
-While k-means anticlustering minimizes between-group-differences only with regard to mean attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, covariance structure, and higher order moments.
-To minimize differences in variance, each variable is duplicated by computing the squared deviation between the original values and the variable's mean.
-Because k-means anticlustering minimizes differences with regard to variables' means, appending the duplicated variables to the original data set---and then applying standard k-means anticlustering---suffices to equalize the variance between groups.
-Following a similar logic, differences in skewness, kurtosis and covariances can be minimized.
-Thus, while k-plus is a separate criterion for anticlustering, it can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
-A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity with regard to mean attribute values as well as the variance and higher order moments, usually with no noticable sacrifice to either criterion.
-The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity.
-Example are given on how k-plus anticlustering can be applied using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
-
-bibliography      : ["lit.bib"]
-
-keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, higher order moments"
-wordcount         : "X"
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : no
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "doc"
-output            : 
-  papaja::apa6_pdf: 
-    number_sections: true
----
-
-```{r setup, include = FALSE}
-library("papaja")
-library("anticlust")
-library("knitr")
-library("dplyr")
-library("ggplot2")
-library("tidyr")
-library("prmisc") # remotes::install_github("m-Py/prmisc")
-options(digits = 2)
-knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE, dev = "cairo_pdf")
-
-source("misc-functions.R")
-
-```
-
-Anticlustering is a collection of methods to partition elements into disjunct groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
-Oftentimes, both goals coincide: several anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990].
-Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
-Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].
-
-Anticlustering is useful in many tasks surrounding research work in psychology.
-For example, anticlustering can be used to split a test into parts of equal difficulty [@gierl2017], to assign students to work groups [@baker2002] or when assigning stimuli to different---but parallel---experimental conditions [@lahl2006].
-For a more comprehensive overview of anticlustering applications in Psychology, see @brusco2019 and @papenberg2020.
-Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual partitioning is usually subpar.
-Fortunately, these problems can be formalized as mathematical optimization problems [e.g., @brusco2019; @spath1986; @baker2002; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
-
-Anticlustering methods can be distinguished by the input data and the objective function that is used to optimize group between-group similarity or within-group heterogeneity.
-Some anticlustering objectives quantify group similarity on the basis of pairwise dissimilarity rating such as the Euclidean distance.
-The most prominent criterion on the basis of pairwise dissimilarity ratings is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
-Maximizing the diversity optimizes within-group heterogeneity, which simultaneously leads to similar groups---if all groups have the same size.
-Anticlustering objective functions usually represent the reversal of a clustering objective, only the direction of the optimization is changed.
-Because the diversity is the reversal of the clustering editing objective [@zahn1964; @shamir2004vf], @papenberg2020 used the term *anticluster editing* to refer to anticlustering approaches that optimize the diversity.
-Another important criterion based on pairwise dissimilarities is the *dispersion*, i.e., the minimum dissimilarity between any two elements within the same group [@fernandez2013; @brusco2019].
-Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
-
-Oftentimes, researchers work with attribute values [or *features*; @dry2009] instead of dissimilarity ratings.
-This use case is for example common when selecting stimuli for an experiment on the basis of norming data [e.g., @kurdi2017introducing].
-For word stimuli in psycholinguistic experiments, norming data may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982].
-In other cases, attribute values are binary and may represent the presence or absence of a feature [@tversky1977].
-In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019].
-Subsequently, an anticlustering criterion such as the diversity can be optimized to partition the elements into groups.
-A different approach to anticlustering directly works with the attribute data by maximizing the k-means criterion.
-K-means is probably the best-known clustering method.
-Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful anticlustering tool by Späth [-@spath1986] and Valev [-@valev1998; -@valev1983].
-In k-means clustering, the sum of the squared Euclidean distances between elements and the center of the cluster they are assigned to is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006].
-Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np].
-For the anticlustering application, the k-means criterion is maximized instead, thereby achieving similarity between groups.
-Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters.
-Hence, groups are similar with regard to the mean of the distribution of the numeric attributes wheras other parameters---such as the variance---are ignored.
-When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [e.g., @anscombe1973].
-@papenberg2020, showed that maximizing the diversity (based on the Euclidean distance) is better suited to minimize difference with regard to both the mean and the variance of the data; even an entirely random split usually resulted in more similar variances because k-means solely focuses on means.
-
-To optimize overall group similarity---and hence to overcome the limitations of k-means anticlustering---I present k-plus anticlustering in this paper.
-The k-plus criterion is an extension of the k-means objective that includes a term that quantifies similarity with regard to the variance, covariances, and higher order moments of data across groups.
-Interestingly, this extended criterion can be understood as an augmentation of the original data matrix, leaving the original k-means objective unchanged.
-I show that the k-plus criterion outperforms other anticlustering criteria when the goal is to minimize differences with regard to the mean and the variance of groups.
-In most cases, both objectives can be optimized satisfactorily at the same time without any loss on either.
-Moreover, in two pratical examples I illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
-
-# Problem formalization
-
-For the purpose of problem formalization, I adopt the notation @steinley2006 provided in his comprehensive synthesis of half a century of research on k-means.
-K-means clustering (and anticlustering) is used to partition two-way, two-mode data into groups.
-That is, $N$ elements each having $P$ attributes are assigned to $K$ groups $(C_1., \ldots, C_K)$, $C_k$ being the set of $n_k$ objects in group $k$.
-The partitioning is subject to the requirement that each element is assigned to exactly one group, formalized by the following two constraints [@papenberg2020]:
-
-$$
-\bigcup\limits_{k = 1}^{K} C_j = X
-$$
-
-$$
-C_j \cap C_k = \emptyset, \; \forall j, k \in \{1, \ldots, K\}, \; j \ne k
-$$
-
-In anticlustering applications, we impose an additional constraint on the group sizes $n_k$, where the most common restriction requires that all groups have equal size:
-
-$$
-n_k = \frac{N}{K}, \forall k \in \{1, \ldots, K\}
-$$
-
-For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means anticlustering aims to maximize the error sum of squares ($\mathit{SSE}$), which is the sum of the squared Euclidean distances between each data point and its cluster centroid:
-
-$$
-\mathit{SSE} = 
-  \sum\limits_{j=1}^{P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (x_{ij} - \overline{x}_j ^{(k)})^2
-$$
-
-The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all observations belonging the $k$th cluster $C_k$ ($k = 1, \ldots, K$):
-
-$$
-\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} x_{ij}
-$$
-
-Note there is a direct connection between the $\mathit{SSE}$ and the location of the cluster centroids [@spath1986].
-If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the overall mean value on the $p$th attribute:
-
-$$
-\overline{\mathbf{x}}_p = \frac{1}{N} \sum\limits_{i=1}^{N} x_{ip}
-$$
-
-If the $\mathit{SSE}$ is maximal---which is the goal of k-means *anticlustering*---, the cluster centroids are as close as possible to the overall centroid $\overline{\mathbf{x}}$ and therefore to each other [@spath1986].
-Thus, k-means anticlustering directly optimizes the similarity of the mean attribute values across clusters.
-
-## Quantifying differences in variance
-
-In the following, I present a variation of the $\mathit{SSE}$ that quantifies how the variance of numeric data differs between groups.
-To motivate the new criterion, we first consider a set of one-dimensional observations $(x_1, \ldots, x_N)$.
-The extension to the general case of $P$ attributes will be straight-forward.
-
-The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
-
-$$
-\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^2
-$$
-
-When defining a new variable $y = (y_1, \ldots, y_N)$ as the squared deviation of each observation from the mean $\overline{\mathbf{x}}$, the sample variance of $x$ can be reformulated as the mean of $y$:
-
-$$
-y_i = (x_i - \overline{\mathbf{x}})^2
-$$
-
-$$
-\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} y_i
-$$
-
-Because the sample variance is basically defined as an average, k-means anticlustering can be employed to equalize the variance across groups.
-That is, we can formulate an adaption of the $\mathit{SSE}$ to quantify the degree to which the variance differs between groups.
-First, however, we generalize the idea to the case of $P$ attributes.
-For each attribute, we have to compute the squared deviation between all values and the mean of the attribute.
-We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$, where
-
-$$
-y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2.
-$$
-
-To reflect how strongly the variances of all attributes vary across the $K$ clusters, we now define the criterion $\mathit{SSE_{Var}}$:
-
-$$
-\mathit{SSE_{Var}} = 
-  \sum\limits_{j=1}^{P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (y_{ij} - \overline{y}_j ^{(k)})^2
-$$
-
-Analogously to the standard $\mathit{SSE}$, the centroid values $\overline{y}_j ^{(k)}$ are given as:
-
-$$
-\overline{y}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} y_{ij}
-$$
-
-Note that $\mathit{SSE_{Var}}$ has the same form as the standard $\mathit{SSE}$.
-However, the input data reflects the squared difference between each data point and the attribute mean instead of the raw data points.
-
-## A bicriterion model: K-plus anticlustering {#kplus}
-
-It is unlikely that an exclusive optimization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is of interest without also considering the $\mathit{SSE}$ (i.e., equalizing means).
-A combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall cluster similarity.
-The most simple approach for such a bicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
-That is, both criteria are computed independently from each other and then combined into a single criterion through a weighted sum:
-
-$$
-\mathit{SSE_{kplus}} = w_1 
-  \sum\limits_{j=1}^{P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (y_{ij} - \overline{x}_j ^{(k)})^2 + 
-  w_2 \sum\limits_{j=1}^{P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (x_{ij} - \overline{y}_j ^{(k)})^2
-$$
-
-The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
-Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
-When defining $z_i = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and $y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ is then computed as
-
-$$
-\mathit{SSE_{kplus}} =
-  \sum\limits_{j=1}^{2 P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (z_{ij} - \overline{z}_j ^{(k)})^2
-$$
-
-This formula again has the same form as $\mathit{SSE}$ (and $\mathit{SSE_{Var}}$), but is computed on different data.
-In this case, both the original data $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function.
-That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending these to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data matrix.
-Hence, k-plus anticlustering reduces to an augmentation of the data matrix, leaving the original k-means criterion unchanged.
-
-Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective.
-Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ have much larger values, on average, than the raw values $x_{ij}$ (also see Figure 1).
-Therefore, the $\mathit{SSE_{Var}}$ criterion naturally receives greater weight when computing $\mathit{SSE_{kplus}}$ than the standard $\mathit{SSE}$ criterion.
-This imbalance is most likely unwarranted because optimizing similarity with regard to variances is not more important than optimizing similarity with regard to means.
-A solution is to standardize all values $x_{ij}$ and $y_{ij}$ to the same scale before computing $SSE_{kplus}$, e.g.
-via $z$-standardization.
-
 ## Criterion tradeoff
@@
 In the next section I present a simulation study that more systematically investigates the performance of the k-plus criterion.
-
-## Skewness and kurtosis
-
-## Covariance structure
-
-# Simulation study
-
-To systematically evaluate the new k-plus criterion, I compared three anticlustering approaches in a simulation study: (a) k-means anticlustering, optimizing the $\mathit{SSE}$; (b) k-plus anticlustering, optimizing the $SSE_{kplus}$; (c) anticluster editing, optimizing the diversity criterion.
-All methods used a $z$-standardization of the data sets as a preprocessing step.
-In the case of k-plus anticlustering, this means that the augmented input data (i.e., the data table that contains the original data and the squared distance data) were subjected to a standardization, which is recommended.
-
-The simulation used the statistical programming environment R [@R-base].
-The R package `anticlust` [Version 0.5.6; @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
-The R packages `dplyr` [Version 1.0.7; @R-dplyr], `ggplot2` [Version 3.3.5; @R-ggplot2], `papaja` [Version 0.1.0.9997; @R-papaja] and `tidyr` [Version 1.1.3; @R-tidyr] were used for data processing and result presentation.
-The simulation study is fully reproducible via code and data that is accessible from the accompanying OSF repository (**link TODO**).
-
-## Conditions
-
-First, I generated 10,000 data sets which were subsequently processed by to the three competing anticlustering methods.
-The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features varied between 1 and 10; (c) the distribution of features was randomly determined to be either (i) uniform in [0, 1], (ii) a standard normal distribution with *M* = 0 and *SD* = 1, or (iii) a wider normal distribution with *M* = 0 and *SD* = 2.
-For a given data set, all $M$ features were independently sampled from the given distribution and no correlations were assumed between features.
-All three anticlustering methods were applied to each of the 10,000 data sets using $K = 2$, $K = 3$ and $K = 4$.
-Groups sizes were always equal.
-
-## Optimization algorithm
-
-To optimize the three anticlustering criteria, a local-maximum search was applied that systematically swaps elements between groups to gradually improve the objective function [Method LCW in @weitz1998].
-Building on an initial random assignment of elements to groups, elements are swapped between groups in such a way that each swap improves the objective criterion by the largest amount that is possible.
-That is, for each element, all possible swaps are simulated and the resulting objective values are stored.
-After each exchange has been tested, the one exchange is realized that improves the objective the most.
-No swap is realized if no swap improves the criterion.
-The exchange procedure is repeated for each element.
-Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements.
-The search stops as soon as an iteration through the entire set no longer yields any improvement, i.e., as soon as a local maximum has been found.
-To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
-The simulation employed five repetitions of the local maximum search.
-
-## Evaluation criteria
-
-Two primary performance criteria were assessed.
-First, for k-plus anticlustering, I checked how well the identified solution approximated the ordinary k-means criterion.
-Specifically, I computed the $\mathit{SSE}$ for the solution that was returned by k-means anticlustering---that directly optimizes $\mathit{SSE}$---and for the solution that was returned by k-plus anticlustering that only indirectly optimizes $\mathit{SSE}$ through $SSE_{kplus}$.
-If optimizing $SSE_{kplus}$ simultaneously optimizes the standard $\mathit{SSE}$ to a satisfactory degree, we may conclude that simultaneously striving for similar means and variances is feasible by optimizing the k-plus criterion.
-
-Second, to compare the three anticlustering criteria with regard to their ability to create similar groups, I computed the means and standard deviations and assessed the degree to which discrepancies occurred between groups.
-To this end, I then computed the difference between the minimum and maximum value of the means and standard deviations between groups, for each of the $M$ features.
-The mean of the differences across all features quantified the total dissimilarity in means ($\Delta M$) and standard deviations ($\Delta \mathit{SD}$).
-Therefore, lower values of $\Delta M$ and $\Delta \mathit{SD}$ indicated that groups were more similar with regard means or standard deviations, respectively.
-
-## Results
-
-```{r, fig.cap = "Displays how well maximizing the k-plus criterion approximates the k-means criterion that is found via direct maximization of the k-means criterion.", fig.height = 4.5, fig.width = 4.5}
-
-## Analyze data for K = 2 and K = 3 and K = 4
-simulation_results <- list()
-for (K in 2:4) {
-  filename <- paste0("./Simulation_Study/results-K", K, "-objectives-raw.csv")
-  df <- read.csv(filename, sep = ";", stringsAsFactors = FALSE)
-  df$K <- K
-  simulation_results[[paste0("K-", K)]] <- df
-}
-
-df <- do.call(rbind, simulation_results)
-rownames(df) <- NULL
-
-# Make long format
-ldf <- pivot_longer(
-  df,
-  cols = paste0(c("kvar", "kmeans", "means", "sd"), "_obj"),
-  names_to = "Objective",
-  names_pattern = "(.*)_obj"
-)
-
-# Check how well k-plus approximates k-means
-rel <- ldf %>% 
-  group_by(method, Objective, N, K) %>% 
-  summarise(Mean = mean(value)) %>% 
-  filter(Objective == "kmeans", method %in% c("kplus", "k-means-exchange")) %>% 
-  pivot_wider(names_from = method, values_from = Mean) %>% 
-  group_by(N, K) %>% 
-  summarise(rel = kplus / `k-means-exchange`)
-
-rel %>% 
-  ggplot(aes(x = N, y = rel, colour = factor(K), linetype = factor(K))) + 
-  geom_line(size = 1) +
-  geom_point(aes(colour = factor(K), shape = factor(K)), size = 3) +
-  scale_x_continuous(breaks = seq(36, 300, 24)) + 
-  theme_apa() +
-  theme(
-    legend.position = "top",
-    legend.title = element_text()
-  ) +
-  scale_colour_grey(name = "K") +
-  guides(
-    linetype = guide_legend(title = "K"),
-    shape = guide_legend(title = "K")
-  ) + 
-  labs(y = "k-plus approximation of k-means criterion")
-
-# approximation at N > 100
-approx_n100 <- mean(filter(rel, N > 100)$rel) 
-approx_n24_k4 <- mean(filter(rel, N == 24, K == 4)$rel)
-
-# cut decimals after n decimals
-cut_decimals <- function(x, n) {
-  x <- as.character(x)
-  x1 <- strsplit(x, "\\.")[[1]][1]
-  x2 <- strsplit(x, "\\.")[[1]][2]
-  x2 <- strsplit(x2, "")[[1]]
-  as.numeric(paste0(x1, ".", paste(x2[1:n], collapse = "")))
-}
-
-```
-
-First, I investigated how well maximizing the k-plus criterion ($\mathit{SSE_{kplus}}$) approximates the ordinary k-means criterion ($\mathit{SSE}$) at the same time.
-Figure 2 illustrates the relative values of $\mathit{SSE}$ that were found when optimizing $\mathit{SSE_{kplus}}$---as compared to the direct optimization of $\mathit{SSE}$.
-Generally, optimizing $\mathit{SSE_{kplus}}$ yielded an adequate approximation of $\mathit{SSE}$.
-In particular, for large group sizes, there is virtually no sacrifice in $\mathit{SSE}$ when  $\mathit{SSE_{kplus}}$ is optimized: For $N > 100$, $\mathit{SSE}$ was approximated to more than `r cut_decimals(approx_n100, 4) * 100`% when conducting k-plus anticlustering as compared to k-means anticlustering.
-In the simulation, the worst approximation was found when the group sizes were smallest: for $N = 24$ and $K = 4$ was approximated to $\mathit{SSE}$ `r approx_n24_k4 * 100`%.
-Thus, the quality of the approximation depended on the group size: smaller groups led to worse approximations, while in general there was very little cost with regard $\mathit{SSE}$ when optimizing $\mathit{SSE_{kplus}}$, expecially for larger groups.
-
-In a next step, I compared k-means anticlustering, k-plus anticlustering and anticluster editing in their ability to equalize means and standard deviations between groups.
-To clarify the general pattern of results, Table 1 illustrates the outcome variables  $\Delta M$ and $\Delta \mathit{SD}$ after splitting the sample size into categories and averaging across simulation runs.
-Note that the results have also been collapsed across the different group sizes ($K = 2, 3, 4$) because the relative performance of the anticlustering methods did not vary by $K$.[^kmaineffect] Table 1 shows that k-means was best suited to equalize mean values between groups; this advantage was most pronounced for smaller group sizes.
-For larger group sizes, k-plus anticlustering and anticluster editing performed similarly well with regard to $\Delta M$.
-Replicating @papenberg2020, it is shown that anticluster editing better minimizes differences with regard to standard deviations between groups than k-means anticlustering.
-Noteworthy, however, k-plus anticlustering strongly outperformed both competitors in this matter; $\Delta \mathit{SD}$ was consistently lowest for the new k-plus method.
-In conclusion, k-plus anticlustering is well suited to minimize differences with regard to mean values across groups---especially when groups are larger---and at the same time is able to minimize differences in standard deviations between groups.
-
-[^kmaineffect]: Only a main effect was observed: across all anticlustering methods, more groups led to higher discrepancies between means and standard deviations between groups.
-
-```{r, fig.cap = "Awesome caption"}
-
-# Recode N into categories so the results can be displayed in a more informative way
-ldf <- within(ldf, {
-  N_category <- ifelse(N <= 200, "N <= 200", "N > 200")
-  N_category <- ifelse(N <= 100, "N <= 100", N_category)
-  N_category <- ifelse(N <= 50,  "N <= 50", N_category)
-})
-
-ldf$N_category <- ordered(
-  ldf$N_category, 
-  levels = c("N <= 50", "N <= 100", "N <= 200", "N > 200")
-)
-
-# for sorting by method, make it ordered factor:
-ldf$method <- ordered(
-  ldf$method, 
-  levels = c("k-means-exchange", "kplus", "ace-exchange"),
-  labels = c("k-means", "k-plus", "anticluster editing")
-)
-
-results_table <- ldf %>% 
-  group_by(method, Objective, N_category) %>% 
-  summarise(Mean = mean(value)) %>% 
-  filter(Objective %in% c("means", "sd"), method != "random") %>% 
-  pivot_wider(names_from = Objective, values_from = Mean) %>%
-  arrange(N_category, method)
-
-# enforce two decimals
-results_table$means <- force_decimals(results_table$means, 2)
-results_table$sd <- force_decimals(results_table$sd, 2)
-
-# Generate table
-footnote <- "$\\Delta M$ and $\\Delta \\mathit{SD}$ quantify how strongly means and standard deviations differed between anticlusters.
-Table cells contain the average objectives across simulation runs, split by $N$."
-
-results_table$N_category <- NULL
-
-colnames <- c("", "$\\Delta M$", "$\\Delta \\mathit{SD}$")
-colnames(results_table) <- colnames
-
-apa_table(
-  results_table, 
-  caption = "Results of the simulation study.", 
-  note = footnote, 
-  escape = FALSE,
-  stub_indents = list("\\textbf{$\\textit{N} \\leq 50$}" = 1:3,
-                      "\\textbf{$50 <\\textit{N} \\leq 100$}" = 4:6,
-                      "\\textbf{$100 <\\textit{N} \\leq 200$}" = 7:9,
-                      "\\textbf{$\\textit{N} > 200$}" = 10:12),
-  span_text_columns = TRUE
-)
-
-```
-
-# K-plus anticlustering using the R package anticlust
-
-In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020].
-It is shown that k-plus anticlustering yields excellent results when being applied to real data sets.
-The results presented here used `anticlust` version 0.5.6.
-The package can be installed in the R environment using the `install.packages()` command:
-
-```R
-install.packages("anticlust")
-```
-
-## Example Application I
-
-In the first example, I use norming data for the OASIS image data set that is freely available online [@kurdi2017introducing].
-@kurdi2017introducing assembled 900 open-access color images and collected ratings on two affective dimensions: arousal and valence.
-@brielmann2019intense collected an additional rating dimension by measuring how the same 900 images were rated with regard to their beauty.
-In my application, I use all three features, currently available from Github.[^githubbeauty]
-
-[^githubbeauty]: https://github.com/aenneb/OASIS-beauty
-
-I used k-plus anticlustering to divide the data set of 900 images into 9 groups of 100 images each; this process takes about 3-5 seconds on a contemporary personal computer, depending on how many iterations are needed to find the local maximum.
-The task was accomplished using the following R code:
-
-```R
-library(anticlust) # load the package anticlust
-
-anticlustering(
-  features,
-  standardize = TRUE,
-  K = 9,
-  objective = "kplus",
-  method = "local-maximum"
-)
-```
-
-Here, the variable `features` has to be a data table containing the image data: 900 rows representing images; 3 columns representing the features beauty, arousal, and valence.
-The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the `features` variable (**TODO**).
-The arguments `objective = "kplus"` and `method = "local-maximum"` ensure that the k-plus criterion is optimized using the local maximum search method described in the previous section.
-The argument `K` describes the number of groups.
-The argument `standardize = TRUE` ensures that all features are standardized before the optimization starts; in particular, this option enforces that equalizing means and variances receives the same weight during the optimization process (see [Section 1.2](#kplus)).
-
-Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means and standard deviations) for each of the 9 groups and for each of the 3 features.
-Additionaly, Table 2 shows the descriptive statistics for an application of k-means anticlustering and anticluster editing.
-It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations.
-K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations.
-Anticluster editing was also well-suited to match both means and standard deviations, but was slightly outperformed by k-plus.
-The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to the means and standard deviations, but instead maximizes within-group heterogeneity.
-
-```{r}
-
-path <- "results_oasis_same_size.csv"
-
-if (!file.exists(path)) {
-  oasis <- read.csv(
-    "https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv"
-  )
-
-  features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
-
-  K <- 9
-  mean_sd_tabs <- list()
-  for (criterion in c("kplus", "variance", "diversity")) {
-    oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
-      features,
-      standardize = TRUE,
-      K = K,
-      objective = criterion,
-      method = "local-maximum"
-    )
-    mean_sd_tabs[[criterion]] <- mean_sd_tab(
-      features,
-      oasis[[paste(criterion, "group", sep = "_")]]
-    )
-    mean_sd_tabs[[criterion]] <- cbind(
-      1:K,
-      nrow(features) / K,
-      mean_sd_tabs[[criterion]]
-    )
-  }
-
-  # append all tables to the same table:
-  oasis_tab <- do.call(rbind, mean_sd_tabs)
-  rownames(oasis_tab) <- NULL
-  colnames(oasis_tab) <- c("Group", "N", "Beauty", "Valence", "Arousal")
-
-  write.table(
-    oasis_tab, path, row.names = FALSE, sep = ";"
-  )
-
-} else {
-  oasis_tab <- read.csv(path, sep = ";")
-}
-
-apa_table(
-  oasis_tab,
-  caption = "Descriptive statistics for OASIS features by group and anticlustering method.",
-  note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
-  escape = FALSE,
-  align = c("r"),
-  stub_indents = list("\\textbf{K-plus anticlustering}" = 1:9,
-                      "\\textbf{K-means anticlustering}" = 10:18,
-                      "\\textbf{Anticluster editing}" = 19:27)
-)
-```
-
-Table 3 illustrates another advantage of k-plus anticlustering over anticluster editing: When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups.
-K-plus anticlustering strives for equal means and variances in all groups, regardless of group size.
-Unequal group sizes can be requested in anticlust by passing the different group sizes to the `K` argument:
-
-```R
-anticlustering(
-  features,
-  K = c(100, 100, 100, 600),
-  standardize = TRUE,
-  objective = "kplus",
-  method = "local-maximum"
-)
-```
-
-```{r}
-
-path <- "results_oasis_unequal_size.csv"
-
-if (!file.exists(path)) {
-  oasis <- read.csv(
-    "https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv"
-  )
-
-  features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
-
-  K <- c(100, 100, 100, 600)
-  mean_sd_tabs <- list()
-  for (criterion in c("kplus", "variance", "diversity")) {
-    oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
-      features,
-      standardize = TRUE,
-      K = K,
-      objective = criterion,
-      method = "local-maximum"
-    )
-    mean_sd_tabs[[criterion]] <- mean_sd_tab(
-      features,
-      oasis[[paste(criterion, "group", sep = "_")]]
-    )
-    mean_sd_tabs[[criterion]] <- cbind(1:length(K), K, mean_sd_tabs[[criterion]])
-  }
-
-  # append all tables to the same table:
-  oasis_tab <- do.call(rbind, mean_sd_tabs)
-  rownames(oasis_tab) <- NULL
-  colnames(oasis_tab) <- c("Group", "N", "Beauty", "Valence", "Arousal")
-
-  write.table(
-    oasis_tab, path, row.names = FALSE, sep = ";"
-  )
-
-} else {
-  oasis_tab <- read.csv(path, sep = ";")
-}
-
-apa_table(
-  oasis_tab,
-  caption = "Performance of the diversity and k-plus objectives for unequal group sizes.",
-  note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
-  escape = FALSE,
-  align = c("r"),
-  stub_indents = list("\\textbf{K-plus anticlustering}" = 1:4,
-                      "\\textbf{K-means anticlustering}" = 5:8,
-                      "\\textbf{Anticluster editing}" = 9:12)
-)
-```
-
-## Example Application II: Small data set ($N = 96$)
-
-The simulation study indicated that k-plus may show decreased performance for smaller group sizes.
-Therefore, a second application illustrates that k-plus anticlustering yields satisfactory results in practice even when group sizes are smaller.
-I make use of a norming data set of 96 word stimuli that was contributed to the `anticlust` package by Dr.
-Schaper [@schaper2019metacognitive; @schaper2019metamory; also see @papenberg2020 for a description of the norming data].
-After loading the `anticlust` package, the data set can be accessed as follows:
-
-```{r, echo = TRUE}
-data(schaper2019)
-```
-
-Table 4 illustrates the results that are obtained when using k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
-Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are still similar enough for any practical purpose---especially when considering inevitable measurement error in norming studies.
-The following code was used to achieve the results reported in Table 4:
-
-```{r, echo = TRUE}
-features <- schaper2019[, 3:6]
-anticlusters <- anticlustering(
-  features,
-  K = 6,
-  objective = "kplus",
-  method = "local-maximum",
-  repetitions = 5,
-  standardize = TRUE
-)
-```
-
-```{r}
-
-tab <- mean_sd_tab(features, anticlusters)
-tab <- cbind(1:6, 96 / 6, tab)
-colnames(tab) <- c("Group", "N", "Typicality", "Atypicality", "Syllables", "Frequency")
-
-apa_table(
-  tab,
-  caption = "Descriptive statistics by group for the Schaper et al. (2019a; 2019b) data set, after applying k-plus anticlustering.",
-  note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
-  escape = FALSE,
-  align = c("r")
-)
-
-```
-
-# Discussion
-
-In this paper, I presented the k-plus criterion for anticlustering.
-K-plus anticlustering focuses on constructing groups that are highly similar to each other, operationalized by similar means as well as variances.
-These descriptive statistics are usually "eyeballed" by applicants of anticlustering methods when they try to assess the quality of of a partitioning; hence, a criterion that directly optimizes these measures has been needed and is now provided.
-The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means.
-Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data.
-For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the variable.
-A simulation study and examples on real norming data indicate that minimizing differences with regard to means and variances between groups is well possible using the k-plus criterion.
-
-This work indicates that k-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups, given that similarity in means and variances can usually be achieved succesfully at the same time.
-However, for small group sizes (small $N$ and / or large $K$) k-plus anticlustering may return slightly less-than-optimal results in the eye of the user.
-In this case, it is not a problem to return to k-means that only focuses on the means.
-As opposed to some statistical workflows where "mindlessly" trying different approaches until one approach yiels satisfactory results is bad practice, such a trial-and-error approach is unproblematic for anticlustering.
-There is no anticluster hacking.
-Users may simply choose the results they like best.
-As a default, it is recommended to apply out k-plus anticlustering when creating groups that aim for high between-group similarity.
-
-For optimizing similarity in means and variances it also outperforms anticluster editing, which, to be fair, does not directly strive to do so.
-Unequal group sizes give another advantage for k-plus anticlustering in comparison to anticluster editing.
-
-## Limitations and outlook
-
-- This paper focused on the presentation of new objective function, the bicriterion optimization process might be improved [e.g., by using a direct algorithm as @brusco2019 did].
-From the algorithmic point of view, this paper did not strive to apply the latest cutting edge methodology.
-For most practical purposes, however, I assume that optimizing the k-plus criterion will yield very satisfactory results for researchers.
-I have already worked with researchers planning their experiments who are generally amazed by the groupd that are returned by k-plus anticlustering.
-
-## Future research
-
-- Future research: Application of more sophisticated bicriterion optimization algorithms are interesting
-
-- extending k-means anticlustering on maximizing similarity with regard to covariances? Mahalanobis distance?
-
-# General thoughts on anticlustering
-
-- new criterion: optimizes between-group similarity, no correspondence to within-group heterogeneity [see @brusco2019 for a exceptional method maximizing within-group heterogeneity]
-
-- Similarly: the reversed objective no longer has a useful clustering interpretation: Maximizing differences in variances does not lead to homogeneous and/or well separated clusters
-
-\newpage
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup

### 2022-02-28 paper.Rmd
added example to potpurri

@@
 via $z$-standardization.
-
-## Criterion tradeoff
-
-```{r}
-
-# First rendering is slow because the data needs to be generated
-path <- "partition_objectives.csv"
-
-if (!file.exists(path)) {
-  set.seed(2207) # 2207
-  # Generate data and all partitions
-  N <- 14
-  K <- 2
-  partitions <- generate_partitions(N, K)
-  features <- rnorm(N, 15, 3)
-
-  # Compute objectives for each partition
-  # For each partition: Compute SSE objective, and SSE(VAR)
-  # 1. SSE
-  all_objectives_kMeans <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = features
-  )
-
-  squared_distances_from_mean <- squared_from_mean(as.matrix(features))
-
-  # 2. SSE(VAR)
-  all_objectives_kVar <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = squared_distances_from_mean
-  )
-
-  # 3. SSE(KPLUS)
-  # Use standardization to compute k-plus criterion!
-  all_objectives_kPlus <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = scale(cbind(features, squared_distances_from_mean))
-  )
-
-  df <- data.frame(
-    kmeans = all_objectives_kMeans,
-    kvar = all_objectives_kVar,
-    kplus = all_objectives_kPlus
-  )
-
-  write.table(
-    df, path, row.names = FALSE, sep = ","
-  )
-} else {
-  df <- read.csv(path)
-}
-
-```
-
-When optimizing a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are close to the global optimum value.
-Generally, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree [@ferligoj1992direct].
-Usually, a tradeoff is necessary.
-However, a motivating example illustrates that aiming to simultaneously optimize similarity of means and variances is not only desirable, but may actually be feasible.
-
-```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
-The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$ and represents a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
-
-# Plot SSE vs SSE(VAR)
-plot(
-  df[, c("kmeans", "kvar")], 
-  pch = 4, col = "darkgrey", 
-  las = 1, cex = .5,
-  xlab = "SSE",
-  ylab = "SSE_Var"
-)
-
-# Illustrate best partition wrt difference in means (i.e., SSE)
-points(
-  df$kmeans[which.max(df$kmeans)],
-  df$kvar[which.max(df$kmeans)],
-  cex = 1.2, pch = 22, bg = "white", lwd = 1.7
-)
-
-# Illustrate best partition wrt difference in variance (i.e., SSE(VAR))
-points(
-  df$kmeans[which.max(df$kvar)],
-  df$kvar[which.max(df$kvar)],
-  cex = 1.2, pch = 24, bg = "white", lwd = 1.7
-)
-
-# Illustrate best partition wrt k-plus criterion
-points(
-  df$kmeans[which.max(df$kplus)],
-  df$kvar[which.max(df$kplus)],
-  cex = 1.2, pch = 23, bg = "white", lwd = 1.7
-)
-```
-
-For this example I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
-Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
-Note that for both criteria, a larger value is better, i.e., indicates higher similarity with regard to means and variances, respectively.
-Therefore, to obtain overall between-group similarity, a partition should be chosen that is depicted in the upper right corner of the scatterplot.
-However, optimal k-means anticlustering would create two groups that are not particularly similar with regard to their variance.
-The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$.
-Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (the triangular symbol in Figure 1) is not well-suited to optimize SSE at the same time.
-The diamond symbol represents the partition that maximizes (the unweighted) $SSE_{kplus}$, i.e., the sum of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$.
-This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
-
-This initial example shows that maximizing the combined k-plus criterion may well be suited to optimize both similarity with regard to means as well as variances between groups.
-In the next section I present a simulation study that more systematically investigates the performance of the k-plus criterion.
 

### 2022-02-28 paper.Rmd
continued

@@
 To motivate the new criterion, we first consider a set of one-dimensional observations $(x_1, \ldots, x_N)$.
-The extension to the general case of $P$ attributes will be straight-forward.
+The extension to the general case of $P$ attributes will be straight forward.
 
-The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
+The sample variance[^biasedestimate] of a set of observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
 
@@
 $$
+
+[^biasedestimate]: this
 
@@
 It is unlikely that an exclusive optimization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is of interest without also considering the $\mathit{SSE}$ (i.e., equalizing means).
-A combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall cluster similarity.
+A combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall between-group similarity.
 The most simple approach for such a bicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
@@
 $$
-\mathit{SSE_{kplus}} = w_1 
-  \sum\limits_{j=1}^{P} 
+w_1 \sum\limits_{j=1}^{P} 
   \sum\limits_{k=1}^{K} 
   \sum\limits_{i \in C_k}
-  (y_{ij} - \overline{x}_j ^{(k)})^2 + 
+  (x_{ij} - \overline{x}_j ^{(k)})^2 + 
 w_2 \sum\limits_{j=1}^{P} 
@@
   \sum\limits_{i \in C_k}
-  (x_{ij} - \overline{y}_j ^{(k)})^2
+  (y_{ij} - \overline{y}_j ^{(k)})^2
 $$
@@
 Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
-When defining $z_i = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and $y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ is then computed as
+When defining $z_i = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and $y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
 
@@
 Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective.
-Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ have much larger values, on average, than the raw values $x_{ij}$ (also see Figure 1).
+Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ have much larger values, on average, than the raw values $x_{ij}$.
 Therefore, the $\mathit{SSE_{Var}}$ criterion naturally receives greater weight when computing $\mathit{SSE_{kplus}}$ than the standard $\mathit{SSE}$ criterion.
@@
 
-## Higher order moments, skewness and kurtosis
+## Skewness, kurtosis, and higher order moments
 
-The outlined logic to equalize variance in different groups is easily generalized to higher order moments such as skewness and kurtosis, i.e., the third and fourth distribution moments, respectively.
-Generally, the *j*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ is computed as [@joanes1998]
+The outlined logic to equalize the variance, which is the second moment of a distribution, can be generalized to higher order moments in a straight forward manner.
+The *j*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ can be computed as
 
@@
 
-where $j = 2$ gives the variance, $j = 3$ the skewness, and $j = 4$ the kurtosis.
+where $j = 2$ gives the variance, $j = 3$ the skewness, and $j = 4$ the kurtosis [@joanes1998].
+Thus, by changing $j$---the power of the deviation between mean and data points---we can use the k-plus logic to equalize any desired sample moment between groups.
+For each moment $j$ and for each variable $x$, this is accomplished by adding a new variable $y^{(j)} = (y^{(j)}_1, \ldots, y^{(j)}_N)$ with $y^{(j)}_i = (x_i - \overline{\mathbf{x}})^j$ to the data set and subsequently applying standard k-means anticlustering.
+I expect this to be particularly interesting for the skewness and kurtosis when aiming to minimize differences with regard to distribution asymmetry (skewness) and the propensity to include outliers (kurtosis).
 
-The approach outlined in sections 1.1 and 1.2 illustrated how the definition of the variance can be used to add variables to the data set to minimize differences with regard to the variance between groups.
-Changing the power of the deviation between mean and data points, we can use the same logic to equalize any desired sample moment between groups.
-I expect this to be interesting in particular for the skewness and kurtosis to minimize differences with regard to distribution asymmetry (skewness) and the propensity to include outliers (kurtosis).
-
-## Covariance structure
+## Covariances
 
@@
 Some clarifications with respect to nomenclature are needed.
-The k-plus criterion is not a single objective function for anticlustering, but instead is a familiy of objectives that use the standard k-means criterion $SSE$ in order to minimize differences between groups with regard to additional characterizations of a data set; that is, not just the mean, but also the variance, covariances, or higher order moments such as skewness and kurtosis.
+The k-plus criterion is not a single objective function for anticlustering, but instead a familiy of objectives that extend the standard k-means objective ($SSE$) to minimize differences between groups with regard to several characterizations of a data set.
+That is: not just the mean, but also the variance, covariances, or higher order moments such as skewness and kurtosis.
 Which of these objectives are pursued depends on a researchers needs.
-In the most basic---and arguably most important---case, k-plus can be used to not just minimize differences with regard to the mean of variables but also their variance.
+In the most basic---and arguably most important---case, k-plus anticlustering can be used to not just minimize differences with regard to the mean of variables but also their variance.
 Measures of location (mean) and spread (variance / standard deviation) are arguably most important to researchers when they "eyeball" their data and they are routinely reported as characterizations of a data set.
@@
 
-While less routinely reported (and less well understood), skewness are kurtosis are also important descriptive characterizations of a data set [@decarlo1997; @westfall2014].
+While less routinely reported, skewness are kurtosis are also important descriptive characterizations of a data set [@decarlo1997; @westfall2014].
 To optimize overall similarity between groups, it may thus be desirable to also minimize differences with regard to these distribution moments.

### 2022-05-19 paper.Rmd
writing

@@
 Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity.
-While k-means anticlustering minimizes between-group-differences only with regard to mean attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, covariance structure, and higher order moments.
+While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, higher order moments, and covariance structure.
 To minimize differences in variance, each variable is duplicated by computing the squared deviation between the original values and the variable's mean.
@@
 
-Anticlustering methods can be distinguished by the input data and the objective function that is used to optimize group between-group similarity or within-group heterogeneity.
+Anticlustering methods can be distinguished by the input data and the objective function that is used to optimize group between-group similarity / within-group heterogeneity.
 Some anticlustering objectives quantify group similarity on the basis of pairwise dissimilarity rating such as the Euclidean distance.
@@
 Anticlustering objective functions usually represent the reversal of a clustering objective, only the direction of the optimization is changed.
-Because the diversity is the reversal of the clustering editing objective [@zahn1964; @shamir2004vf], @papenberg2020 used the term *anticluster editing* to refer to anticlustering approaches that optimize the diversity.
-Another important criterion based on pairwise dissimilarities is the *dispersion*, i.e., the minimum dissimilarity between any two elements within the same group [@fernandez2013; @brusco2019].
+Another important criterion based on pairwise dissimilarities is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013; @brusco2019].
 Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
@@
 K-means is probably the best-known clustering method.
-Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful anticlustering tool by Späth [-@spath1986] and Valev [-@valev1998; -@valev1983].
+Reversing the k-means objective function---using maximization instead of minimization---has been independently recognized as a useful anticlustering tool by Späth [-@spath1986] and Valev [-@valev1998; -@valev1983].
+
 In k-means clustering, the sum of the squared Euclidean distances between elements and the center of the cluster they are assigned to is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006].
@@
 When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [e.g., @anscombe1973].
-@papenberg2020, showed that maximizing the diversity (based on the Euclidean distance) is better suited to minimize difference with regard to both the mean and the variance of the data; even an entirely random split usually resulted in more similar variances because k-means solely focuses on means.
+@papenberg2020 showed that maximizing the diversity (using the Euclidean distance) is better suited to minimize difference with regard to both the mean and the variance of the data; even an entirely random split usually resulted in more similar variances because k-means solely focuses on the features' means.
 
-To optimize overall group similarity---and hence to overcome the limitations of k-means anticlustering---I present k-plus anticlustering in this paper.
-The k-plus criterion is an extension of the k-means objective that includes a term that quantifies similarity with regard to the variance, covariances, and higher order moments of data across groups.
-Interestingly, this extended criterion can be understood as an augmentation of the original data matrix, leaving the original k-means objective unchanged.
-I show that the k-plus criterion outperforms other anticlustering criteria when the goal is to minimize differences with regard to the mean and the variance of groups.
-In most cases, both objectives can be optimized satisfactorily at the same time without any loss on either.
-Moreover, in two pratical examples I illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
+To optimize overall group similarity---and hence to overcome the limitations of k-means anticlustering---this paper presents k-plus anticlustering.
+The k-plus criterion is an extension of the k-means objective that includes one or several terms to quantify similarity with regard to other important characteristics of a data set rather than only the mean.
+Hence, the k-plus criterion is not a single objective function for anticlustering, but instead a familiy of objectives that extend the standard k-means objective to minimize differences between groups with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
+Which of these objectives is pursued depends on a researchers' needs.
+In the most basic---and arguably most important---case, k-plus anticlustering can be used to not just minimize differences with regard to the mean of variables but also their variance.
+Measures of location and spread are arguably most important to researchers when they "eyeball" their data and they are routinely reported as characterizations of a data set.
+Hence, by default, the k-plus objective is refered to when considering the mean and the variance of variables in the anticlustering process.
+
+While less routinely reported, skewness are kurtosis are also important descriptive characterizations of a data set [@decarlo1997; @westfall2014].
+To optimize overall similarity between groups, it may thus be desirable to also minimize differences with regard to these distribution moments.
+In principle, any other higher order moments can also be included as part of the k-plus objective.
+Futhermore, k-plus anticlustering can also be used to obtain similar covariance structures among groups.
+This approach may be of interest in cross validation applications where prediction accuracy depends on the correlations among features and criteria [@zeng2000; @papenberg2020].
+Interestingly, in each case, the k-plus criterion can be reduced to an augmentation of the original data matrix, leaving the original k-means objective unchanged.
+
+After mathematically introducing the k-plus objective, a simulation study is reported showing that k-plus anticlustering is well suited to create groups that have minimal differences according to multiple distribution characteristics.
+Fulfilling multiple criteria at the same time becomes more feasible as the size of the data set increases.
+In two pratical examples I illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
 
@@
 
-## The k-plus family of anticlustering objectives
-
-Some clarifications with respect to nomenclature are needed.
-The k-plus criterion is not a single objective function for anticlustering, but instead a familiy of objectives that extend the standard k-means objective ($SSE$) to minimize differences between groups with regard to several characterizations of a data set.
-That is: not just the mean, but also the variance, covariances, or higher order moments such as skewness and kurtosis.
-Which of these objectives are pursued depends on a researchers needs.
-In the most basic---and arguably most important---case, k-plus anticlustering can be used to not just minimize differences with regard to the mean of variables but also their variance.
-Measures of location (mean) and spread (variance / standard deviation) are arguably most important to researchers when they "eyeball" their data and they are routinely reported as characterizations of a data set.
-Hence, by default, the k-plus anticlustering objective implies that means and variances should be equal between groups.
-
-While less routinely reported, skewness are kurtosis are also important descriptive characterizations of a data set [@decarlo1997; @westfall2014].
-To optimize overall similarity between groups, it may thus be desirable to also minimize differences with regard to these distribution moments.
-In principle, any higher order moments can also be included as part of the k-plus objective.
-Note, however, that the inclusion each additional moment requires appending as many features to the data set as there were to begin with.
-
-K-plus anticlustering can also be used to obtain similar covariance structures among groups.
-This approach may be of interest in cross validation applications where prediction accuracy depends on the correlations among features and criteria [@zeng2000; @papenberg2020].
-Note that each additional objective that is pursued with k-plus anticlustering comes with a cost.
-It becomes more difficult to satisfactorily fulfil many criteria at the same time.
-As the simulation study below will show, including many k-plus objectives at the same time becomes more appropriate when the data set becomes larger.
-In the end, it is a tradeoff between fulfilling few objectives to a precise degree or fulfilling many objectives to a less compelling degree.
+$$
+\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}}) 
+$$
 
@@
 
-To systematically evaluate the new k-plus criterion, I compared three anticlustering approaches in a simulation study: (a) k-means anticlustering, optimizing the $\mathit{SSE}$; (b) k-plus anticlustering, optimizing the $SSE_{kplus}$; (c) anticluster editing, optimizing the diversity criterion.
-All methods used a $z$-standardization of the data sets as a preprocessing step.
-In the case of k-plus anticlustering, this means that the augmented input data (i.e., the data table that contains the original data and the squared distance data) were subjected to a standardization, which is recommended.
+Three k-plus criteria were implemented in a simulation study: (i) standard *k-plus anticlustering*, minimizing differences with regard to the mean and variance; (ii) *k-plus skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) *k-plus correlation*, minimizing differences with regard to the mean, variance and covariance structure.
+The k-plus methods were compared to (a) standard k-means anticlustering, (b) diversity anticlustering using the Euclidean distance, and (c) a simple random allocation of elements to groups.
+All anticlustering methods used a $z$-standardization of the data sets as a preprocessing step.
+In the case of k-plus anticlustering, this means that the augmented input data (i.e., the data table that contains the original data and the features that were added to incorporate the additional optimization criteria) were subjected to a standardization, which is recommended.
 
 The simulation used the statistical programming environment R [@R-base].
-The R package `anticlust` [Version 0.5.6; @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
-The R packages `dplyr` [Version 1.0.7; @R-dplyr], `ggplot2` [Version 3.3.5; @R-ggplot2], `papaja` [Version 0.1.0.9997; @R-papaja] and `tidyr` [Version 1.1.3; @R-tidyr] were used for data processing and result presentation.
+The R package `anticlust` [Version 0.6.1; @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
+The R packages `dplyr` [Version 1.0.7; @R-dplyr], `ggplot2` [Version 3.3.5; @R-ggplot2], `papaja` [Version 0.1.0.9997; @R-papaja] and `tidyr` [Version 1.1.3; @R-tidyr] (**TODO: citation faux**) were used for data processing and result presentation.
 The simulation study is fully reproducible via code and data that is accessible from the accompanying OSF repository (**link TODO**).
-
-## Conditions
-
-First, I generated 10,000 data sets which were subsequently processed by to the three competing anticlustering methods.
-The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features varied between 1 and 10; (c) the distribution of features was randomly determined to be either (i) uniform in [0, 1], (ii) a standard normal distribution with *M* = 0 and *SD* = 1, or (iii) a wider normal distribution with *M* = 0 and *SD* = 2.
-For a given data set, all $M$ features were independently sampled from the given distribution and no correlations were assumed between features.
-All three anticlustering methods were applied to each of the 10,000 data sets using $K = 2$, $K = 3$ and $K = 4$.
-Groups sizes were always equal.
 
@@
 
-To optimize the three anticlustering criteria, a local-maximum search was applied that systematically swaps elements between groups to gradually improve the objective function [Method LCW in @weitz1998].
+The same local maximum search was applied to optimize each of the competing five anticlustering methods [Method LCW in @weitz1998].
 Building on an initial random assignment of elements to groups, elements are swapped between groups in such a way that each swap improves the objective criterion by the largest amount that is possible.
@@
 
+## Conditions
+
+First, I generated 10,000 data sets following a normal distribution.
+The data sets were subsequently processed by the competing methods.
+The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features varied between 2 and 5; (c) the standard deviation of all features was set to 1, 2 or 3 (the mean was always zero); (d) the correlation between all features was r = 0, r = .1, r = .2, r = .3, r = .4 or r = .5.
+All five anticlustering methods were applied to each of the 10,000 data sets using $K = 2$, $K = 3$ and $K = 4$ (i.e., each method was applied 30,000 times).
+Groups sizes were always equal.
+
 ## Evaluation criteria
 
-Two primary performance criteria were assessed.
-First, for k-plus anticlustering, I checked how well the identified solution approximated the ordinary k-means criterion.
-Specifically, I computed the $\mathit{SSE}$ for the solution that was returned by k-means anticlustering---that directly optimizes $\mathit{SSE}$---and for the solution that was returned by k-plus anticlustering that only indirectly optimizes $\mathit{SSE}$ through $SSE_{kplus}$.
-If optimizing $SSE_{kplus}$ simultaneously optimizes the standard $\mathit{SSE}$ to a satisfactory degree, we may conclude that simultaneously striving for similar means and variances is feasible by optimizing the k-plus criterion.
+To compare the five anticlustering criteria with regard to their ability to create similar groups, I investigated the discrepancy in means, standard deviations, skewness, kurtosis, and correlation between features across groups.
 
-Second, to compare the three anticlustering criteria with regard to their ability to create similar groups, I computed the means and standard deviations and assessed the degree to which discrepancies occurred between groups.
-To this end, I then computed the difference between the minimum and maximum value of the means and standard deviations between groups, for each of the $M$ features.
+**todo hier weitermachen**
+
+I then computed the difference between the minimum and maximum value of the means and standard deviations between groups, for each of the $M$ features.
 The mean of the differences across all features quantified the total dissimilarity in means ($\Delta M$) and standard deviations ($\Delta \mathit{SD}$).

### 2022-05-20 paper.Rmd
writing the results section

@@
 
-The sample variance[^biasedestimate] of a set of observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
+The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
 
@@
 $$
-
-[^biasedestimate]: this
 
@@
 
-Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective.
+During the remainder of this paper, whenever the k-plus criterion is optimized, the original data matrix is simply extended and the original k-means criterion $SSE$ is optimized.
+For a multicriterion objective such as the k-plus criterion, this corresponds to a rather basic approach, and future research may investigate more sophisticated optimization schemes for k-plus anticlustering [e.g., see @brusco2012; @brusco2019].
+In this paper, however, I am primarily concerned with investigating the properties of new objective itself rather than extending state of the art multicriterion optimization methods.
+
+Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective, i.e., when just appending the squared-distance-features to the data set.
 Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ have much larger values, on average, than the raw values $x_{ij}$.
@@
 
-Three k-plus criteria were implemented in a simulation study: (i) standard *k-plus anticlustering*, minimizing differences with regard to the mean and variance; (ii) *k-plus skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) *k-plus correlation*, minimizing differences with regard to the mean, variance and covariance structure.
-The k-plus methods were compared to (a) standard k-means anticlustering, (b) diversity anticlustering using the Euclidean distance, and (c) a simple random allocation of elements to groups.
+To evaluate the new k-plus objective function, three criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus skew/kurtosis, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus correlation, minimizing differences with regard to the mean, variance and covariance structure.
+
+The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means; (ii) diversity anticlustering, maximizing the sum of Euclidean distances within groups; (c) a simple random allocation of elements to groups.
 All anticlustering methods used a $z$-standardization of the data sets as a preprocessing step.
-In the case of k-plus anticlustering, this means that the augmented input data (i.e., the data table that contains the original data and the features that were added to incorporate the additional optimization criteria) were subjected to a standardization, which is recommended.
+In the case of k-plus anticlustering, the augmented input data---i.e., the original data as well as the features that were added to incorporate the additional optimization criteria---were subjected to a standardization, which is recommended.
 
@@
 The same local maximum search was applied to optimize each of the competing five anticlustering methods [Method LCW in @weitz1998].
-Building on an initial random assignment of elements to groups, elements are swapped between groups in such a way that each swap improves the objective criterion by the largest amount that is possible.
+Building on an initial random assignment of elements to groups, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
 That is, for each element, all possible swaps are simulated and the resulting objective values are stored.
@@
 No swap is realized if no swap improves the criterion.
-The exchange procedure is repeated for each element.
+This exchange procedure is repeated for each element.
 Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements.
-The search stops as soon as an iteration through the entire set no longer yields any improvement, i.e., as soon as a local maximum has been found.
+The search stops as soon as an iteration through the entire data set no longer yielded an improvement, i.e., as soon as a local maximum has been found.
 To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
@@
 The data sets were subsequently processed by the competing methods.
-The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features varied between 2 and 5; (c) the standard deviation of all features was set to 1, 2 or 3 (the mean was always zero); (d) the correlation between all features was r = 0, r = .1, r = .2, r = .3, r = .4 or r = .5.
-All five anticlustering methods were applied to each of the 10,000 data sets using $K = 2$, $K = 3$ and $K = 4$ (i.e., each method was applied 30,000 times).
+The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features varied between 2 and 5; (c) the standard deviation of all features was set to 1, 2 or 3 (the mean was always zero); (d) the correlation between all features was $r = 0$, $r = .1$, $r = .2$, $r = .3$, $r = .4$ or $r = .5$.
+All anticlustering methods were applied to each of the 10,000 data sets using $K = 2$, $K = 3$ and $K = 4$ (i.e., each method was applied 30,000 times).
 Groups sizes were always equal.
@@
 
-To compare the five anticlustering criteria with regard to their ability to create similar groups, I investigated the discrepancy in means, standard deviations, skewness, kurtosis, and correlation between features across groups.
-
-**todo hier weitermachen**
-
-I then computed the difference between the minimum and maximum value of the means and standard deviations between groups, for each of the $M$ features.
-The mean of the differences across all features quantified the total dissimilarity in means ($\Delta M$) and standard deviations ($\Delta \mathit{SD}$).
-Therefore, lower values of $\Delta M$ and $\Delta \mathit{SD}$ indicated that groups were more similar with regard means or standard deviations, respectively.
+After the five anticlustering methods and the random allocation had been applied to the 10,000 data sets for $K = 2$, $K = 3$, $K = 4$, I compared the competing methods with regard to their ability to create similar groups.
+To this end, I investigated the discrepancy in means, standard deviations, skewness, kurtosis, and correlation between groups---for each data set for each $K$.
+To quantify the discrepancy in group means, the following computation was used: for each of the (2, 3, 4 or 5) features, all (2, 3 or 4) group means were computed.
+For each feature, the difference between the minimum and maximum group mean was used as a measure of discrepancy; the global discrepancy in means ($\Delta M$) was then determined as the average discrepancy across features.
+The same procedure was applied to compute the discrepancy in standard deviations ($\Delta \mathit{SD}$), skewness ($\Delta \mathit{Skew}$) and kurtosis ($\Delta \mathit{Kurtosis}$).
+Quantifying discrepancy in correlations ($\Delta \mathit{Cor}$) followed the same rule, but discrepancies in correlations between groups had to be aggregated across pairs of features instead of single features.
 
@@
 
-```{r, fig.cap = "Displays how well maximizing the k-plus criterion approximates the k-means criterion that is found via direct maximization of the k-means criterion.", fig.height = 4.5, fig.width = 4.5}
+```{r Table1}
+
+# Load required packages
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+source("./Simulation_Study/0-functions-generate-data.R")
 
@@
 
+## add information on simulation conditions
+## TODO: use this and not the one during anticlustering computations
+conditions <- info_from_filename_vectorized(df$file)
+df <- subset(df, select = -c(N, M, SD))
+df <- data.frame(df, conditions)
+
 # Make long format
@@
   df,
-  cols = paste0(c("kvar", "kmeans", "means", "sd"), "_obj"),
+  cols = paste0(c("kvar", "kmeans", "means", "sd", "skew", "kur", "cor"), "_obj"),
   names_to = "Objective",
@@
 
-# Check how well k-plus approximates k-means
-rel <- ldf %>% 
-  group_by(method, Objective, N, K) %>% 
-  summarise(Mean = mean(value)) %>% 
-  filter(Objective == "kmeans", method %in% c("kplus", "k-means-exchange")) %>% 
-  pivot_wider(names_from = method, values_from = Mean) %>% 
-  group_by(N, K) %>% 
-  summarise(rel = kplus / `k-means-exchange`)
+## Global results, aggregated across all simulation variables
+tab <- ldf %>% 
+  group_by(method, Objective) %>% 
+  summarise(Mean = round(mean(value), 3)) %>% 
+  filter(Objective %in% c("means", "sd", "skew", "kur", "cor")) %>% 
+  pivot_wider(names_from = Objective, values_from = Mean) %>% 
+  select(c(means, sd, skew, kur, cor))
 
-rel %>% 
-  ggplot(aes(x = N, y = rel, colour = factor(K), linetype = factor(K))) + 
-  geom_line(size = 1) +
-  geom_point(aes(colour = factor(K), shape = factor(K)), size = 3) +
-  scale_x_continuous(breaks = seq(36, 300, 24)) + 
-  theme_apa() +
-  theme(
-    legend.position = "top",
-    legend.title = element_text()
-  ) +
-  scale_colour_grey(name = "K") +
-  guides(
-    linetype = guide_legend(title = "K"),
-    shape = guide_legend(title = "K")
-  ) + 
-  labs(y = "k-plus approximation of k-means criterion")
+colnames(tab) <- c("",
+  "$\\Delta M$", 
+  "$\\Delta \\mathit{SD}$",
+  "$\\Delta \\mathit{Skew}$",
+  "$\\Delta \\mathit{Kurtosis}$",
+  "$\\Delta \\mathit{Cor}$"
+)
 
-# approximation at N > 100
-approx_n100 <- mean(filter(rel, N > 100)$rel) 
-approx_n24_k4 <- mean(filter(rel, N == 24, K == 4)$rel)
-
-# cut decimals after n decimals
-cut_decimals <- function(x, n) {
-  x <- as.character(x)
-  x1 <- strsplit(x, "\\.")[[1]][1]
-  x2 <- strsplit(x, "\\.")[[1]][2]
-  x2 <- strsplit(x2, "")[[1]]
-  as.numeric(paste0(x1, ".", paste(x2[1:n], collapse = "")))
-}
+apa_table(
+  tab,
+  caption = "Results of the simulation study",
+  note = "The global results of the simulation study aggregated across all conditions.
+Cells contain information about the average between-group discrepancy with regard to means, standard deviations, skewness, kurtosis, and correlations.
+Lower values indicate less discrepancy, i.e., higher between-group similarity.
+Each cell value is the result of averaging across 30,000 data points (10,000 data sets $\\times$ $K = 2, 3$, or 4).",
+  escape = FALSE,
+  align = c("r")
+)
 
@@
 
-First, I investigated how well maximizing the k-plus criterion ($\mathit{SSE_{kplus}}$) approximates the ordinary k-means criterion ($\mathit{SSE}$) at the same time.
-Figure 2 illustrates the relative values of $\mathit{SSE}$ that were found when optimizing $\mathit{SSE_{kplus}}$---as compared to the direct optimization of $\mathit{SSE}$.
-Generally, optimizing $\mathit{SSE_{kplus}}$ yielded an adequate approximation of $\mathit{SSE}$.
-In particular, for large group sizes, there is virtually no sacrifice in $\mathit{SSE}$ when  $\mathit{SSE_{kplus}}$ is optimized: For $N > 100$, $\mathit{SSE}$ was approximated to more than `r cut_decimals(approx_n100, 4) * 100`% when conducting k-plus anticlustering as compared to k-means anticlustering.
-In the simulation, the worst approximation was found when the group sizes were smallest: for $N = 24$ and $K = 4$ was approximated to $\mathit{SSE}$ `r approx_n24_k4 * 100`%.
-Thus, the quality of the approximation depended on the group size: smaller groups led to worse approximations, while in general there was very little cost with regard $\mathit{SSE}$ when optimizing $\mathit{SSE_{kplus}}$, expecially for larger groups.
+```{r Figure1, fig.width = 8, fig.height = 10}
 
-In a next step, I compared k-means anticlustering, k-plus anticlustering and anticluster editing in their ability to equalize means and standard deviations between groups.
-To clarify the general pattern of results, Table 1 illustrates the outcome variables  $\Delta M$ and $\Delta \mathit{SD}$ after splitting the sample size into categories and averaging across simulation runs.
-Note that the results have also been collapsed across the different group sizes ($K = 2, 3, 4$) because the relative performance of the anticlustering methods did not vary by $K$.[^kmaineffect] Table 1 shows that k-means was best suited to equalize mean values between groups; this advantage was most pronounced for smaller group sizes.
-For larger group sizes, k-plus anticlustering and anticluster editing performed similarly well with regard to $\Delta M$.
-Replicating @papenberg2020, it is shown that anticluster editing better minimizes differences with regard to standard deviations between groups than k-means anticlustering.
-Noteworthy, however, k-plus anticlustering strongly outperformed both competitors in this matter; $\Delta \mathit{SD}$ was consistently lowest for the new k-plus method.
-In conclusion, k-plus anticlustering is well suited to minimize differences with regard to mean values across groups---especially when groups are larger---and at the same time is able to minimize differences in standard deviations between groups.
-
-[^kmaineffect]: Only a main effect was observed: across all anticlustering methods, more groups led to higher discrepancies between means and standard deviations between groups.
-
-```{r, fig.cap = "Awesome caption"}
-
-# Recode N into categories so the results can be displayed in a more informative way
-ldf <- within(ldf, {
-  N_category <- ifelse(N <= 200, "N <= 200", "N > 200")
-  N_category <- ifelse(N <= 100, "N <= 100", N_category)
-  N_category <- ifelse(N <= 50,  "N <= 50", N_category)
-})
-
-ldf$N_category <- ordered(
-  ldf$N_category, 
-  levels = c("N <= 50", "N <= 100", "N <= 200", "N > 200")
-)
-
-# for sorting by method, make it ordered factor:
-ldf$method <- ordered(
-  ldf$method, 
-  levels = c("k-means-exchange", "kplus", "ace-exchange"),
-  labels = c("k-means", "k-plus", "anticluster editing")
-)
-
-results_table <- ldf %>% 
-  group_by(method, Objective, N_category) %>% 
+# Plot the results, by N
+ldf %>% 
+  group_by(method, Objective, N) %>% 
   summarise(Mean = mean(value)) %>% 
-  filter(Objective %in% c("means", "sd"), method != "random") %>% 
-  pivot_wider(names_from = Objective, values_from = Mean) %>%
-  arrange(N_category, method)
-
-# enforce two decimals
-results_table$means <- force_decimals(results_table$means, 2)
-results_table$sd <- force_decimals(results_table$sd, 2)
-
-# Generate table
-footnote <- "$\\Delta M$ and $\\Delta \\mathit{SD}$ quantify how strongly means and standard deviations differed between anticlusters.
-Table cells contain the average objectives across simulation runs, split by $N$."
-
-results_table$N_category <- NULL
-
-colnames <- c("", "$\\Delta M$", "$\\Delta \\mathit{SD}$")
-colnames(results_table) <- colnames
-
-apa_table(
-  results_table, 
-  caption = "Results of the simulation study.", 
-  note = footnote, 
-  escape = FALSE,
-  stub_indents = list("\\textbf{$\\textit{N} \\leq 50$}" = 1:3,
-                      "\\textbf{$50 <\\textit{N} \\leq 100$}" = 4:6,
-                      "\\textbf{$100 <\\textit{N} \\leq 200$}" = 7:9,
-                      "\\textbf{$\\textit{N} > 200$}" = 10:12),
-  span_text_columns = TRUE
-)
+  filter(
+    Objective %in% c("means", "sd", "skew", "kur", "cor"),
+    method != "random"
+  ) %>% 
+  mutate(
+    Objective = ordered(
+      Objective, 
+      levels = c("means", "sd", "skew", "kur", "cor"),
+      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
+  ) %>%
+  ggplot(aes(x = N, y = Mean, colour = method)) + 
+  geom_line(aes(linetype = method), size = .85) + 
+  facet_grid(rows = vars(Objective), scales = "free") + 
+  ylab("Mean discrepancy") +
+  theme_bw(base_size = 16)
 
 ```
+
+Table 1 displays the global simulation results aggregated across all conditions.
+Standard k-means is best at minimizing discrepancy with regard to means, but disregards all other distribution characteristics.
+K-plus anticlustering also addresses $\Delta \mathit{M}$ quite well and at the same time tends to minimize discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$).
+With regard to minimizing $\Delta \mathit{SD}$, all three k-plus approaches outperform all other methods---including diversity anticlustering.
+
+As Figure 1 shows, with increasing $N$, all k-means and the k-plus methods methods converge on minimizing $\Delta \mathit{M}$.
+This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time.
+In particular, standard k-plus anticlustering is not much worse than k-means anticlustering when it comes to $\Delta \mathit{M}$, but it offers the immense benefit of also equalizing the spread of the data across groups.
+
+While k-plus anticlustering apparently is capable of fulfilling multiple objectives at the same time, it is apparent that increasing the number of optimization criteria aggravates fulfilling each single criterion.
+For example, k-plus skew/kurtosis and k-plus correlation are worse at minimizing discrepancy with regard to standard deviations than standard k-plus.
+Still, every k-plus method does what it is supposed to do: k-plus skew/kurtosis is best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation is best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$.
+At the same time, these two k-plus objectives maintain a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
+
+It is maybe surprising that k-plus skew/kurtosis does not seem to be much better than diversity anticlustering with regard to skew and---in particular---kurtosis, even though this methods is specialized for these criteria.
+Figure 1 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis is even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
+However, with increasing $N$, it clearly outperforms diversity anticlustering when it comes to $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
+These results show the cost of optimizing several criteria at the same time; increasing $N$ facilitates fulfilling multiple criteria.
+For low $N$, all k-plus methods show the highest discrepancies in skew and kurtosis; here k-plus seems to be preoccupied with fulfilling the other objectives.
+But even for small $N$, all k-plus methods outperform all other methods with regard to discrepancies in standard deviations.
+
+Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to somewhat address all distribution characteristics.
+For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering is clearly outperformed.
+As as side note, apart from k-plus correlation, diversity anticlustering is the only method that equalizes correlation structure among groups.
+As far as I know, this observation has not been made previously, even though the maximum diversity problem is quite well studied in the literature.
 

### 2022-05-20 paper.Rmd
results / discussion

@@
 
-To evaluate the new k-plus objective function, three criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus skew/kurtosis, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus correlation, minimizing differences with regard to the mean, variance and covariance structure.
+To evaluate the new family of k-plus objectives, three specific criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus skew/kurtosis, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus correlation, minimizing differences with regard to the mean, variance and covariance structure.
 
@@
 
-The same local maximum search was applied to optimize each of the competing five anticlustering methods [Method LCW in @weitz1998].
+The same local maximum search was applied to optimize each of the five competing anticlustering methods [Method LCW in @weitz1998].
 Building on an initial random assignment of elements to groups, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
@@
 
-After the five anticlustering methods and the random allocation had been applied to the 10,000 data sets for $K = 2$, $K = 3$, $K = 4$, I compared the competing methods with regard to their ability to create similar groups.
+After the five anticlustering methods and the random allocation procedure had been applied to the 10,000 data sets for $K = 2$, $K = 3$, $K = 4$, I investigated how well the competing methods were able to create similar groups.
 To this end, I investigated the discrepancy in means, standard deviations, skewness, kurtosis, and correlation between groups---for each data set for each $K$.
@@
 
-As Figure 1 shows, with increasing $N$, all k-means and the k-plus methods methods converge on minimizing $\Delta \mathit{M}$.
+As Figure 1 shows, with increasing $N$, k-means and all k-plus methods methods converge on minimizing $\Delta \mathit{M}$.
 This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time.
@@
 
-While k-plus anticlustering apparently is capable of fulfilling multiple objectives at the same time, it is apparent that increasing the number of optimization criteria aggravates fulfilling each single criterion.
+While k-plus anticlustering is capable of fulfilling multiple objectives at the same time, it is apparent that increasing the number of optimization criteria aggravates fulfilling each single criterion.
 For example, k-plus skew/kurtosis and k-plus correlation are worse at minimizing discrepancy with regard to standard deviations than standard k-plus.
@@
 
-It is maybe surprising that k-plus skew/kurtosis does not seem to be much better than diversity anticlustering with regard to skew and---in particular---kurtosis, even though this methods is specialized for these criteria.
+It is maybe surprising that k-plus skew/kurtosis does not seem to be much better than diversity anticlustering with regard to skew and---in particular---kurtosis, even though this method is specifically tailored to these criteria.
 Figure 1 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis is even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
 However, with increasing $N$, it clearly outperforms diversity anticlustering when it comes to $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
-These results show the cost of optimizing several criteria at the same time; increasing $N$ facilitates fulfilling multiple criteria.
-For low $N$, all k-plus methods show the highest discrepancies in skew and kurtosis; here k-plus seems to be preoccupied with fulfilling the other objectives.
-But even for small $N$, all k-plus methods outperform all other methods with regard to discrepancies in standard deviations.
+These results show the cost of optimizing several criteria at the same time; for low $N$, k-plus seems skew/kurtosis to be preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
+Increasing $N$ then facilitates to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
 
 Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to somewhat address all distribution characteristics.
-For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering is clearly outperformed.
+For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering gets outperformed.
 As as side note, apart from k-plus correlation, diversity anticlustering is the only method that equalizes correlation structure among groups.
@@
 
+Moreover, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
+Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes.
+It was of particular interest to compare the k-plus objective with the popular diversity objective.
+@papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups, whereas the diversity maintains an appropriate balance between location and spread.
+Using the k-plus extension, however, the k-means criterion regains a lot of attractivity: when optimizing for similarity in means and variance, k-plus anticlustering clearly outperformed the diversity criterion in the simulation study.
+It should also be noted that the k-plus criterion should almost always be preferred to standard k-means.
+This is because even for low $N$, k-plus equalizes means quite well, and for larger $N$ the practical differences between k-means and k-plus vanish completely.
+A reasonable exception is given when the data set consists of binary variables where mean and standard deviation are directly related.
+
+[^clusterediting]: In their study, they used the term anticluster editing to refer to the maximization of the diversity criterion.
+The diversity criterion is the objective function used in the clustering method cluster editing, where it has to be minimized [@zahn1964; @shamir2004vf].
+
 ## Future research

### 2022-05-20 paper.Rmd
some restructuring
writing

@@
 
-source("misc-functions.R")
-
 ```
 
-Anticlustering is a collection of methods to partition elements into disjunct groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
-Oftentimes, both goals coincide: several anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990].
+Anticlustering is a collection of methods to partition a set of elements into disjunct groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
+Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990].
 Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
@@
 Table 1 displays the global simulation results aggregated across all conditions.
-Standard k-means is best at minimizing discrepancy with regard to means, but disregards all other distribution characteristics.
+Standard k-means performed best at minimizing discrepancy with regard to means, but disregards all other distribution characteristics.
 K-plus anticlustering also addresses $\Delta \mathit{M}$ quite well and at the same time tends to minimize discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$).
 With regard to minimizing $\Delta \mathit{SD}$, all three k-plus approaches outperform all other methods---including diversity anticlustering.
-
 As Figure 1 shows, with increasing $N$, k-means and all k-plus methods methods converge on minimizing $\Delta \mathit{M}$.
 This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time.
-In particular, standard k-plus anticlustering is not much worse than k-means anticlustering when it comes to $\Delta \mathit{M}$, but it offers the immense benefit of also equalizing the spread of the data across groups.
 
-While k-plus anticlustering is capable of fulfilling multiple objectives at the same time, it is apparent that increasing the number of optimization criteria aggravates fulfilling each single criterion.
-For example, k-plus skew/kurtosis and k-plus correlation are worse at minimizing discrepancy with regard to standard deviations than standard k-plus.
+While k-plus anticlustering is capable of addressing multiple objectives simultaneously, it is apparent that increasing the number of optimization criteria aggravates fulfilling each single criterion.
+K-plus skew/kurtosis and k-plus correlation are worse at minimizing discrepancy with regard to standard deviations than standard k-plus, which is slightly worse at minimizing $\Delta \mathit{M}$ than k-means anticlustering.
 Still, every k-plus method does what it is supposed to do: k-plus skew/kurtosis is best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation is best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$.
@@
 
-It is maybe surprising that k-plus skew/kurtosis does not seem to be much better than diversity anticlustering with regard to skew and---in particular---kurtosis, even though this method is specifically tailored to these criteria.
+It is maybe surprising that k-plus skew/kurtosis does not seem to be much better than diversity anticlustering with regard to skew and---in particular---kurtosis, even though the method has been specifically tailored to these criteria.
 Figure 1 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis is even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
-However, with increasing $N$, it clearly outperforms diversity anticlustering when it comes to $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
-These results show the cost of optimizing several criteria at the same time; for low $N$, k-plus seems skew/kurtosis to be preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
+However, with increasing $N$, it clearly outperforms diversity anticlustering.
+These results illustrate the cost of optimizing several criteria at the same time: for low $N$, k-plus seems skew/kurtosis to be preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
 Increasing $N$ then facilitates to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
@@
 As as side note, apart from k-plus correlation, diversity anticlustering is the only method that equalizes correlation structure among groups.
-As far as I know, this observation has not been made previously, even though the maximum diversity problem is quite well studied in the literature.
+As far as I know, this observation has not been made previously, even though the maximum diversity problem has been widely studied.
 
@@
 
-In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020].
+This section demonstrates how researchers and other data analysists can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020].
 It is shown that k-plus anticlustering yields excellent results when being applied to real data sets.
-The results presented here used `anticlust` version 0.5.6.
+The results presented here used `anticlust` version 0.6.1.
 The package can be installed in the R environment using the `install.packages()` command:
@@
 The argument `K` describes the number of groups.
-The argument `standardize = TRUE` ensures that all features are standardized before the optimization starts; in particular, this option enforces that equalizing means and variances receives the same weight during the optimization process (see [Section 1.2](#kplus)).
+The argument `standardize = TRUE` ensures that all features are standardized before the optimization starts; in particular, this option enforces that equalizing means and variances receives the same weight during the optimization process.
 
@@
 
-Table 3 illustrates another advantage of k-plus anticlustering over anticluster editing: When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups.
+Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups.
 K-plus anticlustering strives for equal means and variances in all groups, regardless of group size.
@@
 In this paper, I presented the k-plus criterion for anticlustering.
-K-plus anticlustering focuses on constructing groups that are highly similar to each other, operationalized by similar means as well as variances.
-These descriptive statistics are usually "eyeballed" by applicants of anticlustering methods when they try to assess the quality of of a partitioning; hence, a criterion that directly optimizes these measures has been needed and is now provided.
+K-plus anticlustering focuses on dividing a data set into subgroups in such a way that these groups are similar to each other.
 The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means.
+The k-plus family may be used to equalize the variance as well, any higher order moments such as skewness and kurtosis, as well as covariances.
 Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data.
-For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the variable.
-A simulation study and examples on real norming data indicate that minimizing differences with regard to means and variances between groups is well possible using the k-plus criterion.
+A simulation study and examples on real norming data indicated that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.
 
 This work indicates that k-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups, given that similarity in means and variances can usually be achieved succesfully at the same time.
-However, for small group sizes (small $N$ and / or large $K$) k-plus anticlustering may return slightly less-than-optimal results in the eye of the user.
-In this case, it is not a problem to return to k-means that only focuses on the means.
-As opposed to some statistical workflows where "mindlessly" trying different approaches until one approach yiels satisfactory results is bad practice, such a trial-and-error approach is unproblematic for anticlustering.
-There is no anticluster hacking.
-Users may simply choose the results they like best.
-As a default, it is recommended to apply out k-plus anticlustering when creating groups that aim for high between-group similarity.
-
-For optimizing similarity in means and variances it also outperforms anticluster editing, which, to be fair, does not directly strive to do so.
-Unequal group sizes give another advantage for k-plus anticlustering in comparison to anticluster editing.
+Depending on a user's needs, equalizing higher order moments can also be requested, but it should be noted that fulfilling several criteria at the same time may be difficult to achieve for small data sets.
+In any way, a user can just try out different objectives, i.e., start greedy by requesting that multiple distribution moments are equalized, and if the results for the means and standard deviations are subpar, try out fewer criteria.
+K-plus anticlustering was also shown to outperform diversity anticlustering with regard to the specific objectives that are included in the optimization process; diversity is still a reasonable all-arounder that strives for overall betwee-group similarity.
+Which method should be employed depends on a user's preferences.
+In doubt, several anticlustering methods can be applied sequentially and the results can be compared directly.
+Unlike in statistical methods, there is no problem with trying out different approaches when it comes to anticlustering.
 

### 2022-05-24 Simulation_Study/1-Generate-Data.R
added all data sets and simulation results

@@ -6,7 +6,7 @@
 source("0-functions-generate-data.R")
 
 sample_sizes <- 12 * 2:25 # 24 - 300
-nsets <- 1000
+nsets <- 10000
 
 N  <- sample(sample_sizes, replace = TRUE, size = nsets)
 M  <- sample(2:5, replace = TRUE, size = nsets)

### 2022-05-24 paper.Rmd

@@
 
-Analogously to the standard $\mathit{SSE}$, the centroid values $\overline{y}_j ^{(k)}$ are given as:
+Analogously to the standard $\mathit{SSE}$, the centroid values $\overline{y}_j ^{(k)}$ are given as
 
 $$
-\overline{y}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} y_{ij}
+\overline{y}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} y_{ij}.
 $$
@@
   (z_{ij} - \overline{z}_j ^{(k)})^2
-$$
+$$.
 
@@
 For a multicriterion objective such as the k-plus criterion, this corresponds to a rather basic approach, and future research may investigate more sophisticated optimization schemes for k-plus anticlustering [e.g., see @brusco2012; @brusco2019].
-In this paper, however, I am primarily concerned with investigating the properties of new objective itself rather than extending state of the art multicriterion optimization methods.
+In this paper, however, I am primarily concerned with investigating the properties of new objective itself rather than extending the state of the art on multicriterion optimization methods.
 
@@
 $$
-\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^j
+\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^j.
 $$
@@
 Because the covariance is also defined as an expected value, the k-plus criterion can be extended to reflect differences in covariances between data sets.
-In this case, an additional variable has to be computed for each pair of variables, i.e., $\binom{N}{2}$ new variables, using the definition of the sample covariance:
+The covariance between two variables $x$ and $y$ is defined as
 
 $$
-\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}}) 
+\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}}).
 $$
+
+Thus, to maximize between-groups similarity with regard to covariance structure, it is possible to add an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
+This requires computing $\binom{N}{2}$ new variables and thus turns out be be rather costly: The number of additional variables needed to minimize differences with regard to covariance structure grows quadratically with increasing number of variables; the number of additional variables needed to minimize differences with regard to distribution moments only grows linearly with increasing number of variables.
 
@@
 
-To evaluate the new family of k-plus objectives, three specific criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus skew/kurtosis, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus correlation, minimizing differences with regard to the mean, variance and covariance structure.
+To thorougly evaluate k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.
 
@@
 
-The simulation used the statistical programming environment R [@R-base].
+The simulation was implemented using the statistical programming language R [Version 4.2; @R-base].
 The R package `anticlust` [Version 0.6.1; @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
-The R packages `dplyr` [Version 1.0.7; @R-dplyr], `ggplot2` [Version 3.3.5; @R-ggplot2], `papaja` [Version 0.1.0.9997; @R-papaja] and `tidyr` [Version 1.1.3; @R-tidyr] (**TODO: citation faux**) were used for data processing and result presentation.
+The R package `faux` [Version 1.1.0; @R-faux] was used to generate the multivariate normal data sets that were processed during the simulation.
+The R packages `dplyr` [Version 1.0.7; @R-dplyr], `ggplot2` [Version 3.3.5; @R-ggplot2], `papaja` [Version 0.1.0.9997; @R-papaja], `tidyr` [Version 1.1.3; @R-tidyr] and `DescTools` [Version 0.99.45; @R-desctools] were used for data analysis and result presentation.
 The simulation study is fully reproducible via code and data that is accessible from the accompanying OSF repository (**link TODO**).
@@
 
-The same local maximum search was applied to optimize each of the five competing anticlustering methods [Method LCW in @weitz1998].
+The same local maximum search was applied to optimize each of the five competing anticlustering objectives [Method LCW in @weitz1998].
 Building on an initial random assignment of elements to groups, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
-That is, for each element, all possible swaps are simulated and the resulting objective values are stored.
+That is, for each element, all possible exchanges are simulated and the resulting objective values are stored.
 After each exchange has been tested, the one exchange is realized that improves the objective the most.
-No swap is realized if no swap improves the criterion.
-This exchange procedure is repeated for each element.
-Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements.
-The search stops as soon as an iteration through the entire data set no longer yielded an improvement, i.e., as soon as a local maximum has been found.
+No exchange is realized if no improvement is possible.
+The exchange procedure is repeated for each element.
+Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements, repeating the exchange procedure.
+The procedure stops as soon as an iteration through the entire data set no longer yielded an improvement, i.e., as soon as a local maximum has been found.
 To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
@@
 Lower values indicate less discrepancy, i.e., higher between-group similarity.
-Each cell value is the result of averaging across 30,000 data points (10,000 data sets $\\times$ $K = 2, 3$, or 4).",
+Each cell value is the result of averaging across 30,000 data points (10,000 data sets $\\times$ $K = 2$, $K = 3$, $K = or 4$).",
   escape = FALSE,
@@
 K-plus anticlustering also addresses $\Delta \mathit{M}$ quite well and at the same time tends to minimize discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$).
-With regard to minimizing $\Delta \mathit{SD}$, all three k-plus approaches outperform all other methods---including diversity anticlustering.
-As Figure 1 shows, with increasing $N$, k-means and all k-plus methods methods converge on minimizing $\Delta \mathit{M}$.
+With regard to minimizing $\Delta \mathit{SD}$, all three k-plus approaches outperform all other methods.
+As Figure 1 shows, with increasing $N$, k-means and all k-plus methods converge on minimizing $\Delta \mathit{M}$.
 This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time.
@@
 Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to somewhat address all distribution characteristics.
-For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering gets outperformed.
+For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering is outperformed.
 As as side note, apart from k-plus correlation, diversity anticlustering is the only method that equalizes correlation structure among groups.
@@
 
-This section demonstrates how researchers and other data analysists can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020].
-It is shown that k-plus anticlustering yields excellent results when being applied to real data sets.
-The results presented here used `anticlust` version 0.6.1.
+This section demonstrates how researchers and other data analysists can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020], using an interface function called `kplus_anticlustering()`, which is available from version 0.6.3 onward.
 The package can be installed in the R environment using the `install.packages()` command:
@@
 
-anticlustering(
+kplus_anticlustering(
   features,
-  standardize = TRUE,
   K = 9,
-  objective = "kplus",
-  method = "local-maximum"
+  variance = TRUE,
+  skew = TRUE,
+  kurtosis = FALSE,
+  covariances = FALSE,
+  moments = NULL,
+  method = "local-maximum",
+  standardize = TRUE
 )
@@
 The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the `features` variable (**TODO**).
-The arguments `objective = "kplus"` and `method = "local-maximum"` ensure that the k-plus criterion is optimized using the local maximum search method described in the previous section.
+The `kplus_anticlustering()` function has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion.
+Only the argument `variance` is set to `TRUE` by default; the other arguments have to be "turned on".
+If other higher order moments should be included as part of the optimization, the argument `moments` can be used to specify the desired moments as an integer vector.
+
+The argument `method = "local-maximum"` ensures that the k-plus criterion is optimized using the local maximum search method described in the previous section.
 The argument `K` describes the number of groups.
-The argument `standardize = TRUE` ensures that all features are standardized before the optimization starts; in particular, this option enforces that equalizing means and variances receives the same weight during the optimization process.
+By default, the function ensures that all features are standardized before the optimization starts; in particular, this option enforces that all k-plus criteria receive the same weight during the optimization process (this behaviour can be adjusted using the boolean argument `standardize`).
 

### 2022-05-24 paper.Rmd
prmisc not needed in current manuscript

@@
 library("tidyr")
-library("prmisc") # remotes::install_github("m-Py/prmisc")
 options(digits = 2)

### 2022-05-24 Simulation_Study/4-Aggregate-Results.R
do not use random method in plot

@@ -7,6 +7,8 @@ library(dplyr)
 library(ggplot2)
 library(tidyr)
 
+source("0-functions-generate-data.R")
+
 ## Analyze data for K = 2 and K = 3 and K = 4
 simulation_results <- list()
 for (K in 2:4) {
@@ -40,7 +42,7 @@ ldf <- pivot_longer(
 ## Global results, aggregated across all simulation variables
 ldf %>% 
   group_by(method, Objective) %>% 
-  summarise(Mean = round(mean(value), 3)) %>% 
+  summarise(Mean = round(mean(value), 2)) %>% 
   filter(Objective %in% c("means", "sd", "skew", "kur", "cor")) %>% 
   pivot_wider(names_from = Objective, values_from = Mean) %>% 
   select(c(means, sd, skew, kur, cor))
@@ -50,7 +52,8 @@ ldf %>%
 ldf %>% 
   group_by(method, Objective, N) %>% 
   summarise(Mean = mean(value)) %>% 
-  filter(Objective %in% c("means", "sd", "skew", "kur", "cor")) %>% 
+  filter(Objective %in% c("means", "sd", "skew", "kur", "cor"),
+         method != "random") %>% 
   ggplot(aes(x = N, y = Mean, colour = method)) + 
   geom_line(size = 1) + 
   facet_grid(rows = vars(Objective), scales = "free") + 

### 2022-05-24 paper.Rmd
table and figure caption

@@
 Lower values indicate less discrepancy, i.e., higher between-group similarity.
-Each cell value is the result of averaging across 30,000 data points (10,000 data sets $\\times$ $K = 2$, $K = 3$, $K = or 4$).",
+Each cell value is the result of averaging across 30,000 data points (10,000 data sets $\\times$ $K = 2$, $K = 3$ or $K = 4$).",
   escape = FALSE,
@@
 
-```{r Figure1, fig.width = 8, fig.height = 10}
+```{r Figure1, fig.width = 8, fig.height = 10, fig.cap = "Simulation results by $N$."}
 

### 2022-05-24 Simulation_Study/0-functions-analyze-data.R
rerun simulation because file info was not correctly inferred (i'm too stupid to change it in the csv files without causing damage apparently)

@@ -14,8 +14,6 @@
 compute_objectives <- function(row, K) {
   filename <- paste0("./datasets/", row["file"], ".csv")
   #print(filename)
-  #print(row)
-  #print(row$file)
   data <- read.csv(filename)
   anticlusters <- anticlusters_from_string(row["result"])
   kvar_obj <- variance_objective(squared_from_mean(data), anticlusters)

### 2022-05-24 Simulation_Study/0-functions-generate-data.R

@@ -49,18 +49,3 @@ info_from_filename <- function(file) {
 }
 
 number_from_string <- function(x) as.numeric(gsub("[^0-9]", "", x))
-
-
-info_from_filename_vectorized <- function(file) {
-  info <- lapply(
-    strsplit(file, "_"), 
-    FUN = gsub,
-    pattern = "_", 
-    replacement = ""
-  )
-  N <- sapply(info, function(x) number_from_string(x[1]))
-  M <- sapply(info, function(x) number_from_string(x[2]))
-  SD <- sapply(info, function(x) number_from_string(x[3]))
-  r <- sapply(info, function(x) as.numeric(gsub("r", "", x[4])))
-  data.frame(N, M, SD, r)
-}

### 2022-05-24 Simulation_Study/2-Call-Methods.R

@@ -29,7 +29,7 @@ for (K in 2:4) {
   
   # Uncomment the following line and exchange the value of the size 
   # argument if not all files should be processed at the same time
-  #files <- sample(files, size = 1000)
+  files <- sample(files, size = 100)
   
   ## Apply all methods to the data sets
   message("Starting to work on ", length(files), " data sets")

### 2022-05-24 Simulation_Study/4-Aggregate-Results.R

@@ -7,8 +7,6 @@ library(dplyr)
 library(ggplot2)
 library(tidyr)
 
-source("0-functions-generate-data.R")
-
 ## Analyze data for K = 2 and K = 3 and K = 4
 simulation_results <- list()
 for (K in 2:4) {
@@ -23,13 +21,6 @@ rownames(df) <- NULL
 
 length(unique(df$ID))
 table(table(df$ID))
-
-## add information on simulation conditions
-## TODO: use this and not the one during anticlustering computations
-conditions <- info_from_filename_vectorized(df$file)
-df <- subset(df, select = -c(N, M, SD))
-df <- data.frame(df, conditions)
-
 
 # Make long format
 ldf <- pivot_longer(

### 2022-05-24 Simulation_Study/0-functions-anticlustering-methods.R
change method names for k-plus

@@ -55,9 +55,9 @@ anticluster_data_ <- function(data, K) {
   objectives <- c(
     "diversity",
     "k-means",
-    "kplus",
-    "kplus-skew-kurtosis",
-    "kplus-correlation",
+    "k-plus",
+    "k-plus-skew-kurtosis",
+    "k-plus-correlation",
     "random"
   )
   clusters <- as.list(rep(NA, length(objectives)))
@@ -87,7 +87,7 @@ anticluster_data_ <- function(data, K) {
         repetitions = repetitions,
         standardize = standardize
       )
-    } else if (i == "kplus") {
+    } else if (i == "k-plus") {
       clusters[[i]] <- anticlustering(
         data, 
         K = K,
@@ -96,7 +96,7 @@ anticluster_data_ <- function(data, K) {
         repetitions = repetitions,
         standardize = standardize
       )
-    } else if (i == "kplus-skew-kurtosis") {
+    } else if (i == "k-plus-skew-kurtosis") {
       var_data <- moment_features(data, moment = 2)
       skew_data <- moment_features(data, moment = 3)
       kurtosis_data <- moment_features(data, moment = 4)
@@ -109,7 +109,7 @@ anticluster_data_ <- function(data, K) {
           repetitions = repetitions,
           standardize = FALSE
         )
-    } else if (i == "kplus-correlation") {
+    } else if (i == "k-plus-correlation") {
       var_data <- moment_features(data, moment = 2)
       covar_data <- covariance_features(data)
       data_correlation <- scale(cbind(data, var_data, covar_data))
@@ -123,6 +123,8 @@ anticluster_data_ <- function(data, K) {
       )
     } else if (i == "random") {
       clusters[[i]] <- sample(rep_len(1:K, N))
+    } else {
+      stop("There is a typo in your method names!")
     }
   }
   clusters

### 2022-05-24 Simulation_Study/0-functions-analyze-data.R
no longer compute k-means and k-plus criterion for each data set

@@ -16,8 +16,6 @@ compute_objectives <- function(row, K) {
   #print(filename)
   data <- read.csv(filename)
   anticlusters <- anticlusters_from_string(row["result"])
-  kvar_obj <- variance_objective(squared_from_mean(data), anticlusters)
-  kmeans_obj <- variance_objective(data, anticlusters)
   means_obj <- var_means(anticlusters, data)
   sd_obj <- var_sds(anticlusters, data)
   skew_obj <- var_skew(anticlusters, data)
@@ -27,8 +25,6 @@ compute_objectives <- function(row, K) {
   return(c(
     row["ID"],
     row["method"],
-    kvar_obj = kvar_obj,
-    kmeans_obj = kmeans_obj,
     means_obj = means_obj,
     sd_obj = sd_obj,
     skew_obj = skew_obj,

### 2022-05-24 Simulation_Study/4-Aggregate-Results.R
only keep objectives that are actually used

@@ -25,7 +25,7 @@ table(table(df$ID))
 # Make long format
 ldf <- pivot_longer(
   df,
-  cols = paste0(c("kvar", "kmeans", "means", "sd", "skew", "kur", "cor"), "_obj"),
+  cols = paste0(c("means", "sd", "skew", "kur", "cor"), "_obj"),
   names_to = "Objective",
   names_pattern = "(.*)_obj"
 )
@@ -34,7 +34,6 @@ ldf <- pivot_longer(
 ldf %>% 
   group_by(method, Objective) %>% 
   summarise(Mean = round(mean(value), 2)) %>% 
-  filter(Objective %in% c("means", "sd", "skew", "kur", "cor")) %>% 
   pivot_wider(names_from = Objective, values_from = Mean) %>% 
   select(c(means, sd, skew, kur, cor))
 
@@ -43,8 +42,7 @@ ldf %>%
 ldf %>% 
   group_by(method, Objective, N) %>% 
   summarise(Mean = mean(value)) %>% 
-  filter(Objective %in% c("means", "sd", "skew", "kur", "cor"),
-         method != "random") %>% 
+  filter(method != "random") %>% 
   ggplot(aes(x = N, y = Mean, colour = method)) + 
   geom_line(size = 1) + 
   facet_grid(rows = vars(Objective), scales = "free") + 
@@ -56,7 +54,6 @@ ldf %>%
 ldf %>% 
   group_by(method, Objective, M) %>% 
   summarise(Mean = mean(value)) %>% 
-  filter(Objective %in% c("means", "sd", "skew", "kur", "cor")) %>% 
   ggplot(aes(x = M, y = Mean, colour = method)) + 
   geom_line(size = 1) + 
   facet_grid(rows = vars(Objective), scales = "free") + 
@@ -66,7 +63,6 @@ ldf %>%
 ldf %>% 
   group_by(method, Objective, SD) %>% 
   summarise(Mean = mean(value)) %>% 
-  filter(Objective %in% c("means", "sd", "skew", "kur", "cor")) %>% 
   ggplot(aes(x = SD, y = Mean, colour = method)) + 
   geom_point(size = 3) + 
   geom_line(size = 1) + 
@@ -87,7 +83,6 @@ ldf %>%
 ldf %>% 
   group_by(method, Objective, K) %>% 
   summarise(Mean = mean(value)) %>% 
[…]

### 2022-05-24 paper.Rmd
this is not needed when file info is correctly stored
...same in the paper

@@
 
-## add information on simulation conditions
-## TODO: use this and not the one during anticlustering computations
-conditions <- info_from_filename_vectorized(df$file)
-df <- subset(df, select = -c(N, M, SD))
-df <- data.frame(df, conditions)
-
 # Make long format
@@
   df,
-  cols = paste0(c("kvar", "kmeans", "means", "sd", "skew", "kur", "cor"), "_obj"),
+  cols = paste0(c("means", "sd", "skew", "kur", "cor"), "_obj"),
   names_to = "Objective",
@@
   summarise(Mean = round(mean(value), 3)) %>% 
-  filter(Objective %in% c("means", "sd", "skew", "kur", "cor")) %>% 
   pivot_wider(names_from = Objective, values_from = Mean) %>% 
@@
   summarise(Mean = mean(value)) %>% 
-  filter(
-    Objective %in% c("means", "sd", "skew", "kur", "cor"),
-    method != "random"
-  ) %>% 
+  filter(method != "random") %>% 
   mutate(

### 2022-05-24 paper.Rmd
working on example

@@
 Only the argument `variance` is set to `TRUE` by default; the other arguments have to be "turned on".
-If other higher order moments should be included as part of the optimization, the argument `moments` can be used to specify the desired moments as an integer vector.
+If other higher order moments should be included as part of the optimization, the optional argument `moments` can be used to specify the desired moments as an integer vector.
 
@@
 
-```{r}
+```{r, eval = FALSE}
 
@@
   K <- 9
-  mean_sd_tabs <- list()
   for (criterion in c("kplus", "variance", "diversity")) {
+    if (criterion == "kplus") {
+        oasis[[paste(criterion, "group", sep = "_")]] <- kplus_anticlustering(
+            features,
+            K = K,
+            variance = TRUE,
+            skew = TRUE,
+            kurtosis = FALSE,
+            covariances = FALSE,
+            moments = NULL,
+            method = "local-maximum",
+            standardize = TRUE
+        )
+    } else {
         oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
@@
         )
-    mean_sd_tabs[[criterion]] <- mean_sd_tab(
-      features,
-      oasis[[paste(criterion, "group", sep = "_")]]
-    )
-    mean_sd_tabs[[criterion]] <- cbind(
-      1:K,
-      nrow(features) / K,
-      mean_sd_tabs[[criterion]]
-    )
     }
 
+  }
+
+    descriptive_by_group <- function(features, anticlusters, FUN, name) {
+    df <- data.frame(
+      t(data.frame(lapply(by(features, anticlusters, FUN), c)))
+    )
+    df$group <- 1:nrow(df)
+    df$Descriptive = name
+    df
+  }
+
+  descriptive_by_group_all <- function(features, anticlusters) {
+    means <- descriptive_by_group(features, anticlusters, colMeans, "mean")
+    sds <- descriptive_by_group(
+      features, 
+      anticlusters, 
+      function(x) sapply(x, sd),
+      "sds"
+    )
+    skew <- descriptive_by_group(
+      features, 
+      anticlusters, 
+      function(x) sapply(x, DescTools::Skew),
+      "skew"
+    )
+    rbind(means, sds, skew)
+  }
+
+kplus <- descriptive_by_group_all(features, oasis$kplus_group)
+kplus <- data.frame(kplus, Objective = "kplus")
+
+kmeans <- descriptive_by_group_all(features, oasis$variance_group)
+kmeans <- data.frame(kmeans, Objective = "kmeans")
+diversity <- descriptive_by_group_all(features, oasis$variance_group)
+diversity <- data.frame(diversity, Objective = "diversity")
+
+all_objs <- rbind(kplus, kmeans, diversity) %>% 
+  pivot_wider(names_from = Descriptive, values_from = c(beauty_mean, Valence_mean, Arousal_mean))
+
   # append all tables to the same table:
-  oasis_tab <- do.call(rbind, mean_sd_tabs)
-  rownames(oasis_tab) <- NULL
-  colnames(oasis_tab) <- c("Group", "N", "Beauty", "Valence", "Arousal")
-
-  write.table(
-    oasis_tab, path, row.names = FALSE, sep = ";"
-  )
+  rownames(all_objs) <- NULL
+  colnames(all_objs) <- c("Group", "N", "Beauty", "Valence", "Arousal")
 

### 2022-05-24 Simulation_Study/2-Call-Methods.R
redo simulation, hopefully last time

@@ -29,7 +29,7 @@ for (K in 2:4) {
   
   # Uncomment the following line and exchange the value of the size 
   # argument if not all files should be processed at the same time
-  files <- sample(files, size = 100)
+  #files <- sample(files, size = 200)
   
   ## Apply all methods to the data sets
   message("Starting to work on ", length(files), " data sets")

### 2022-05-24 paper.Rmd
implemented new example 1 (including skew now) in paper

--- 2022-05-24 paper.Rmd
@@
 library("tidyr")
+# for formatting numbers in Rmarkdown, use my package prmisc (https://github.com/m-Py/prmisc)
+library("prmisc") # remotes::install_github("m-Py/prmisc") 
 options(digits = 2)
@@
 
-path <- "results_oasis_same_size.csv"
+example1_file <- "results_oasis_same_size.csv"
+if (!file.exists(example1_file)) {
 
-if (!file.exists(path)) {
     oasis <- read.csv(
@@
     K <- 9
-  mean_sd_tabs <- list()
-  for (criterion in c("kplus", "variance", "diversity")) {
-    oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
+
+    # k-plus variance / skew
+    groups_kplus <- kplus_anticlustering(
         features,
-      standardize = TRUE,
         K = K,
-      objective = criterion,
-      method = "local-maximum"
+        variance = TRUE,
+        skew = TRUE,
+        method = "local-maximum",
+        standardize = TRUE
     )
-    mean_sd_tabs[[criterion]] <- mean_sd_tab(
+
+    # k-means
+    groups_kmeans <- anticlustering(
         features,
-      oasis[[paste(criterion, "group", sep = "_")]]
+        K = K,
+        objective = "variance",
+        method = "local-maximum",
+        standardize = TRUE
     )
-    mean_sd_tabs[[criterion]] <- cbind(
-      1:K,
-      nrow(features) / K,
-      mean_sd_tabs[[criterion]]
+
+    # diversity
+    groups_diversity <- anticlustering(
+        features,
+        K = K,
+        objective = "diversity",
+        method = "local-maximum",
+        standardize = TRUE
     )
+
+    # Functions to compute means / SD / skew per group 
+    descriptives_by_group <- function(features, anticlusters, FUN, name) {
+        df <- data.frame(
+            t(data.frame(lapply(by(features, anticlusters, FUN), c)))
+        )
+        df$group <- 1:nrow(df)
+        df$Descriptive = name
+        df
     }
 
-  # append all tables to the same table:
-  oasis_tab <- do.call(rbind, mean_sd_tabs)
-  rownames(oasis_tab) <- NULL
-  colnames(oasis_tab) <- c("Group", "N", "Beauty", "Valence", "Arousal")
+    all_descriptives_by_group <- function(features, anticlusters) {
+    means <- descriptives_by_group(features, anticlusters, colMeans, "mean")
+    sds <- descriptives_by_group(
+        features, 
+        anticlusters, 
+        function(x) sapply(x, sd),
+        "sds"
+    )
+    skew <- descriptives_by_group(
+        features, 
+        anticlusters, 
+        function(x) sapply(x, DescTools::Skew),
+        "skew"
+    )
+    rbind(means, sds, skew)
+    }
+
+    kplus <- all_descriptives_by_group(features, groups_kplus)
+    kplus <- data.frame(kplus, Objective = "kplus")
+
+    kmeans <- all_descriptives_by_group(features, groups_kmeans)
+    kmeans <- data.frame(kmeans, Objective = "kmeans")
+    diversity <- all_descriptives_by_group(features, groups_diversity)
+    diversity <- data.frame(diversity, Objective = "diversity")
+
+    all_objs <- rbind(kplus, kmeans, diversity) %>% 
+    pivot_wider(names_from = Descriptive, values_from = c(beauty_mean, Valence_mean, Arousal_mean))
 
     write.table(
-    oasis_tab, path, row.names = FALSE, sep = ";"
+        all_objs, example1_file, row.names = FALSE, sep = ";"
     )
@@
 } else {
-  oasis_tab <- read.csv(path, sep = ";")
+  all_objs <- read.csv(example1_file, sep = ";")
 }
 
+table_all_objs <- subset(all_objs, select = -Objective)
+table_all_objs <- as.matrix(table_all_objs)
+colnames(table_all_objs) <- c("Group", rep(c("$M$", "$\\mathit{SD}$", "Skew"), 3))
+table_all_objs <- apply(table_all_objs, 2, force_or_cut)
+#mode(table_all_objs) <- "character"
+
 apa_table(
-  oasis_tab,
+  table_all_objs,
   caption = "Descriptive statistics for OASIS features by group and anticlustering method.",
-  note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
+  note = "The k-plus approach minimized differences with regard to means, variances and skew.",
   escape = FALSE,
-  align = c("r"),
+  align = c("c", rep("r", 9)),
+  col_spanners = list(Beauty = c(2, 4), Valence = c(5, 7), Arousal = c(8, 10)), 
   stub_indents = list("\\textbf{K-plus anticlustering}" = 1:9,
                       "\\textbf{K-means anticlustering}" = 10:18,
-                      "\\textbf{Anticluster editing}" = 19:27)
+                      "\\textbf{Diversity anticlustering}" = 19:27)
 )

### 2022-05-24 Example1.R

@@ -1,0 +1,88 @@
+
+library(tidyr)
+library(dplyr)
+library(anticlust)
+
+oasis <- read.csv(
+  "https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv"
+)
+
+features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
+
+K <- 9
+
+# k-plus variance / skew
+start <- Sys.time()
+groups_kplus <- kplus_anticlustering(
+  features,
+  K = K,
+  variance = TRUE,
+  skew = TRUE,
+  method = "local-maximum",
+  standardize = TRUE
+)
+Sys.time() - start
+
+# k-means
+start <- Sys.time()
+groups_kmeans <- anticlustering(
+  features,
+  K = K,
+  objective = "variance",
+  method = "local-maximum",
+  standardize = TRUE
+)
+Sys.time() - start
+
+# diversity
+start <- Sys.time()
+groups_diversity <- anticlustering(
+  features,
+  K = K,
+  objective = "diversity",
+  method = "local-maximum",
+  standardize = TRUE
+)
+Sys.time() - start
+
+
+# Functions to compute means / SD / skew per group 
+descriptives_by_group <- function(features, anticlusters, FUN, name) {
+  df <- data.frame(
+    t(data.frame(lapply(by(features, anticlusters, FUN), c)))
+  )
+  df$group <- 1:nrow(df)
+  df$Descriptive = name
+  df
+}
+
+
+all_descriptives_by_group <- function(features, anticlusters) {
+  means <- descriptives_by_group(features, anticlusters, colMeans, "mean")
+  sds <- descriptives_by_group(
+    features, 
+    anticlusters, 
+    function(x) sapply(x, sd),
+    "sds"
+  )
+  skew <- descriptives_by_group(
+    features, 
+    anticlusters, 
+    function(x) sapply(x, DescTools::Skew),
+    "skew"
+  )
+  rbind(means, sds, skew)
+}
+
+
+kplus <- all_descriptives_by_group(features, groups_kplus)
+kplus <- data.frame(kplus, Objective = "kplus")
+
+kmeans <- all_descriptives_by_group(features, groups_kmeans)
+kmeans <- data.frame(kmeans, Objective = "kmeans")
+diversity <- all_descriptives_by_group(features, groups_diversity)
+diversity <- data.frame(diversity, Objective = "diversity")
+
[…]

### 2022-05-30 paper.Rmd

@@
 
-To thorougly evaluate k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.
+To thorougly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.
 
-The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means; (ii) diversity anticlustering, maximizing the sum of Euclidean distances within groups; (c) a simple random allocation of elements to groups.
+The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means; (ii) diversity anticlustering, maximizing the sum of Euclidean distances within groups; (iii) a simple random allocation of elements to groups.
 All anticlustering methods used a $z$-standardization of the data sets as a preprocessing step.
@@
 
-The same local maximum search was applied to optimize each of the five competing anticlustering objectives [Method LCW in @weitz1998].
+The same local maximum search [Method LCW in @weitz1998] was applied to optimize each of the five competing anticlustering objectives.
 Building on an initial random assignment of elements to groups, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
@@
 
-I used k-plus anticlustering to divide the data set of 900 images into 9 groups of 100 images each; this process takes about 3-5 seconds on a contemporary personal computer, depending on how many iterations are needed to find the local maximum.
-The task was accomplished using the following R code:
+I used k-plus anticlustering to divide the data set of 900 images into 9 groups of 100 images each. The task was accomplished using the following R code:
 
@@
 
-Here, the variable `features` has to be a data table containing the image data: 900 rows representing images; 3 columns representing the features beauty, arousal, and valence.
-The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the `features` variable (**TODO**).
+The code runs in about 10 seconds on a contemporary personal computer and varies depending on how many iterations are needed to find the local maximum.
+The variable `features` has to be a data table containing the image data: 900 rows representing images; 3 columns representing the features beauty, arousal, and valence.
+The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the variable `features` (**TODO**).
+
 The `kplus_anticlustering()` function has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion.
-Only the argument `variance` is set to `TRUE` by default; the other arguments have to be "turned on".
+Only the argument `variance` is set to `TRUE` by default---corresponding to the understanding that k-plus is an extension to k-means that at least equalizes the variance in addition to the mean.
+The arguments corresponding to skewness, kurtosis and convariance have to be spefically "turned on".
 If other higher order moments should be included as part of the optimization, the optional argument `moments` can be used to specify the desired moments as an integer vector.
-
 The argument `method = "local-maximum"` ensures that the k-plus criterion is optimized using the local maximum search method described in the previous section.
 The argument `K` describes the number of groups.
-By default, the function ensures that all features are standardized before the optimization starts; in particular, this option enforces that all k-plus criteria receive the same weight during the optimization process (this behaviour can be adjusted using the boolean argument `standardize`).
+By default, the function ensures that all features are standardized before the optimization starts; in particular, this option enforces that all k-plus criteria receive the same weight during the optimization process.
+This behaviour can be adjusted using the boolean argument `standardize`.
+The documentation of the `anticlust` package also provides detailed information on how to use the `kplus_anticlustering()` function.
 
-Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means and standard deviations) for each of the 9 groups and for each of the 3 features.
-Additionaly, Table 2 shows the descriptive statistics for an application of k-means anticlustering and anticluster editing.
+Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features.
+Table 2 also entails the results that were obtained when applying k-means and diversity anticlustering with the same specifications.
 It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations.
-K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations.
-Anticluster editing was also well-suited to match both means and standard deviations, but was slightly outperformed by k-plus.
-The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to the means and standard deviations, but instead maximizes within-group heterogeneity.
+The skewness is also very evenly matched.
+K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations and especially skewness.
+Diversity anticlustering was well-suited to match means, standard deviations and skew, but was slightly outperformed by k-plus.
+The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to these criteria, but instead maximizes within-group heterogeneity.
+Generally, all methods fulfilled their goals pretty well because the data set they were applied to was rather large.
 
@@
 K-plus anticlustering strives for equal means and variances in all groups, regardless of group size.
-Unequal group sizes can be requested in anticlust by passing the different group sizes to the `K` argument:
+Unequal group sizes can be requested in anticlust by passing the different group sizes to the argument `K`.
+The following code was used to equalize means and variances between four groups:
 
 ```R
-anticlustering(
+kplus_anticlustering(
   features,
   K = c(100, 100, 100, 600),
-  standardize = TRUE,
-  objective = "kplus",
   method = "local-maximum"
@@
   for (criterion in c("kplus", "variance", "diversity")) {
+    if (criterion == "kplus") {
+        oasis[[paste(criterion, "group", sep = "_")]] <-  kplus_anticlustering(
+            features,
+            K = c(100, 100, 100, 600),
+            method = "local-maximum"
+        )
+    } else {
         oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
@@
         )
+    }
     mean_sd_tabs[[criterion]] <- mean_sd_tab(

### 2022-05-30 paper.Rmd
some work on example etc

@@
 In the first example, I use norming data for the OASIS image data set that is freely available online [@kurdi2017introducing].
-@kurdi2017introducing assembled 900 open-access color images and collected ratings on two affective dimensions: arousal and valence.
+Kurdi et al.
+assembled 900 open-access color images and collected ratings on two affective dimensions: arousal and valence.
 @brielmann2019intense collected an additional rating dimension by measuring how the same 900 images were rated with regard to their beauty.
-In my application, I use all three features, currently available from Github.[^githubbeauty]
+In my application, I use all three features; the corresponding data is available from Github.[^githubbeauty]
 
@@
 
-The code runs in about 10 seconds on a contemporary personal computer and varies depending on how many iterations are needed to find the local maximum.
-The variable `features` has to be a data table containing the image data: 900 rows representing images; 3 columns representing the features beauty, arousal, and valence.
-The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the variable `features` (**TODO**).
+The code runs in about 10 seconds on a contemporary personal computer; the run time may vary depending on how many iterations are needed to find the local maximum.
+The variable `features` is a data table containing the image data: 900 rows representing images and 3 columns representing the features beauty, arousal, and valence.
+The accompanying OSF repository contains code to reproduce the example (**TODO**).
 
 The `kplus_anticlustering()` function has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion.
-Only the argument `variance` is set to `TRUE` by default---corresponding to the understanding that k-plus is an extension to k-means that at least equalizes the variance in addition to the mean.
-The arguments corresponding to skewness, kurtosis and convariance have to be spefically "turned on".
+Only the argument `variance` is set to `TRUE` by default---corresponding to the understanding that k-plus is an extension to k-means that at least equalizes variances in addition to means.
+The arguments corresponding to skewness, kurtosis and convariance have to be spefically "turned on", i.e., by setting them to `TRUE` in the function call of `kplus_anticlustering()`.
 If other higher order moments should be included as part of the optimization, the optional argument `moments` can be used to specify the desired moments as an integer vector.
@@
 Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features.
-Table 2 also entails the results that were obtained when applying k-means and diversity anticlustering with the same specifications.
-It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations.
-The skewness is also very evenly matched.
+Table 2 also includes results of k-means anticlustering and diversity anticlustering.
+It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations, and the skewness was also very evenly matched.
 K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations and especially skewness.
 Diversity anticlustering was well-suited to match means, standard deviations and skew, but was slightly outperformed by k-plus.
-The latter result is not surprising because---unlike k-plus anticlustering---anticluster editing does not directly maximize similarity with regard to these criteria, but instead maximizes within-group heterogeneity.
-Generally, all methods fulfilled their goals pretty well because the data set they were applied to was rather large.
+The latter result is not surprising because---unlike k-plus anticlustering---diversity anticlustering does not directly maximize similarity with regard to these criteria, but instead maximizes within-group heterogeneity.
 
@@
   caption = "Descriptive statistics for OASIS features by group and anticlustering method.",
-  note = "The k-plus approach minimized differences with regard to means, variances and skew.",
+  note = "The k-plus criterion that was optimized included a term to minimize differences with regard to means, variances, and skewness.",
   escape = FALSE,
@@
 Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups.
-K-plus anticlustering strives for equal means and variances in all groups, regardless of group size.
+This is due the diversity criterion being a measure of within-group heterogeneity; maximizing within-group heterogeneity only incidentally maximizes between-group similarity.
+However, because the diversity criterion is a sum rather than an average, the correspondence between within-group heterogeneity and between-group similarity no longer holds unequal-sized groups.
+K-plus anticlustering strives for between-group similarity regardless of group size.
+
 Unequal group sizes can be requested in anticlust by passing the different group sizes to the argument `K`.
-The following code was used to equalize means and variances between four groups:
+The following code was used to equalize means and variances between four groups using the k-plus approach:
 
@@
                       "\\textbf{K-means anticlustering}" = 5:8,
-                      "\\textbf{Anticluster editing}" = 9:12)
+                      "\\textbf{Diversity anticlustering}" = 9:12)
 )
 ```
+
+This function call relied on the default k-plus settings, ensuring that variances are equalized -- but no higher order moments or covariances.
 
@@
 
-Table 4 illustrates the results that are obtained when using k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
+Table 4 illustrates the results that are obtained when using standard k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
 Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are still similar enough for any practical purpose---especially when considering inevitable measurement error in norming studies.
-The following code was used to achieve the results reported in Table 4:
 
-```{r, echo = TRUE}
+```{r, echo = FALSE}
 features <- schaper2019[, 3:6]
@@
   tab,
-  caption = "Descriptive statistics by group for the Schaper et al. (2019a; 2019b) data set, after applying k-plus anticlustering.",
+  caption = "Descriptive statistics by group for the Schaper et al. (2019a; 2019b) data set, after applying default k-plus anticlustering.",
   note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",

### 2022-05-30 paper.Rmd
discussion

@@
 The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means.
-The k-plus family may be used to equalize the variance as well, any higher order moments such as skewness and kurtosis, as well as covariances.
+The k-plus criterion however may be used to also equalize the variance between groups, and any higher order moments such as skewness and kurtosis, as even covariances.
 Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data.
-A simulation study and examples on real norming data indicated that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.
 
-This work indicates that k-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups, given that similarity in means and variances can usually be achieved succesfully at the same time.
-Depending on a user's needs, equalizing higher order moments can also be requested, but it should be noted that fulfilling several criteria at the same time may be difficult to achieve for small data sets.
-In any way, a user can just try out different objectives, i.e., start greedy by requesting that multiple distribution moments are equalized, and if the results for the means and standard deviations are subpar, try out fewer criteria.
+A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.
+In particular, this work indicates that k-plus anticlustering should be a replacement of the classical k-means approach when striving for between-group similarity.
+K-means anticlustering disregards any distribution characteristics other than the mean; k-plus anticlustering ensures between-group similarity with regard to multiple criteria while still taking into account differences in means.
+For large data sets, there is virtually no reason to use k-means anticlustering because the practical differences between k-means and k-plus regarding similarity in means vanish completely.
+A reasonable exception is given when the input data consists of binary variables because in this case the mean and the standard deviation are directly related.
+
+The results showed that fulfilling several criteria at the same time may be difficult to achieve for small data sets.
+Still, users can try out different objectives and see which specification fulfils their needs.
+A reasonable approach it to start greedy, i.e., requesting multiple distribution moments to be equalized, and only if the results are subpar, go for fewer criteria.
+Unlike in statistical analysis, there is no problem with "just trying out" different approaches when it comes to anticlustering.
+
 K-plus anticlustering was also shown to outperform diversity anticlustering with regard to the specific objectives that are included in the optimization process; diversity is still a reasonable all-arounder that strives for overall betwee-group similarity.
@@
 In doubt, several anticlustering methods can be applied sequentially and the results can be compared directly.
-Unlike in statistical methods, there is no problem with trying out different approaches when it comes to anticlustering.
 
@@
 
-- This paper focused on the presentation of new objective function, the bicriterion optimization process might be improved [e.g., by using a direct algorithm as @brusco2019 did].
-From the algorithmic point of view, this paper did not strive to apply the latest cutting edge methodology.
-For most practical purposes, however, I assume that optimizing the k-plus criterion will yield very satisfactory results for researchers.
-I have already worked with researchers planning their experiments who are generally amazed by the groupd that are returned by k-plus anticlustering.
+This paper focused on the presentation of new objective function but the did not properly consider the multicriterion nature of the k-plus objective.
+Instead, a simple (unweighted) sum approach was applied to optimize the k-plus objective, independent of the number of criteria that were incorporated.
+Other approaches, such as direct algorithms, may yield improved results and better possibilities to select partitionings according to differently weighted criteria.
+@brusco2019 showed how direct algorithms can be used to optimize multiple critera in anticlustering applications.
+Their study focused on an approach to find the pareto set of solutions that simultaneously optimized diversity and dispersion, thereby maximizing within-group heterogeneity with regard to both criteria.[^inanticlust]
 
-Moreover, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
+[^inanticlust]: The `anticlust` package now also includes an implementation of the bicriterion anticlustering algorithm by @brusco2019.
+
+Focusing on the objetive function only and not on specialized algorithmic optimization approaches followed two reasonings.
+First, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
 Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes.
@@
 Using the k-plus extension, however, the k-means criterion regains a lot of attractivity: when optimizing for similarity in means and variance, k-plus anticlustering clearly outperformed the diversity criterion in the simulation study.
-It should also be noted that the k-plus criterion should almost always be preferred to standard k-means.
-This is because even for low $N$, k-plus equalizes means quite well, and for larger $N$ the practical differences between k-means and k-plus vanish completely.
-A reasonable exception is given when the data set consists of binary variables where mean and standard deviation are directly related.
+Second, the applications show that the simple sum approach using a general local maximum search yields very satisfying results in practice.
+It seems that the space of solutions in anticlustering settings includes partitionings that fulfil multiple objectives and these partitionings can be found using a local maximum search.
 
@@
 
-## Future research
-
-- Future research: Application of more sophisticated bicriterion optimization algorithms are interesting
-
-- extending k-means anticlustering on maximizing similarity with regard to covariances? Mahalanobis distance?
-
-# General thoughts on anticlustering
+## Closing thoughts on the nature of anticlustering
 
@@
 - Similarly: the reversed objective no longer has a useful clustering interpretation: Maximizing differences in variances does not lead to homogeneous and/or well separated clusters
+
+## Conclusion
+
+This paper introduced k-plus anticlustering for maximizing between-group similarity, improving the traditional k-means approach.
+K-plus anticlustering can be adopted whenever a data set has to be partitioned into equivalent parts.
+The method is accessible and easily usable via the R package `anticlust`.
 

### 2022-05-30 paper.Rmd
more discussion

@@
 
-In this paper, I presented the k-plus criterion for anticlustering.
-K-plus anticlustering focuses on dividing a data set into subgroups in such a way that these groups are similar to each other.
+In this paper, I presented the k-plus criterion for anticlustering, which can be used to divided a data set into subgroups in such a way that the groups are similar to each other.
 The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means.
-The k-plus criterion however may be used to also equalize the variance between groups, and any higher order moments such as skewness and kurtosis, as even covariances.
-Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data.
+The k-plus criterion also equalizes the variance between groups, and if needed any higher order moments such as skewness and kurtosis, or even the covariances.
+Thus, k-plus represents a family of anticlustering criteria, which can be combined by users according to their needs.
+Interestingly, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
 
 A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.
-In particular, this work indicates that k-plus anticlustering should be a replacement of the classical k-means approach when striving for between-group similarity.
-K-means anticlustering disregards any distribution characteristics other than the mean; k-plus anticlustering ensures between-group similarity with regard to multiple criteria while still taking into account differences in means.
-For large data sets, there is virtually no reason to use k-means anticlustering because the practical differences between k-means and k-plus regarding similarity in means vanish completely.
-A reasonable exception is given when the input data consists of binary variables because in this case the mean and the standard deviation are directly related.
+In particular, this work indicates that k-plus anticlustering should be a replacement of the classical k-means approach when striving for between-group similarity.[^binarydata]  K-means anticlustering disregards any distribution characteristics other than the mean, while k-plus anticlustering ensures between-group similarity with regard to multiple criteria---and still minimizes differences in means quite well.
+The results showed that fulfilling several criteria at the same time may be aggregated in small data sets.
+Still, users can try out different objectives and see which specification fulfils their needs.
+A reasonable approach it to start greedy by requesting that multiple distribution moments should equalized.
+Only if the results of the greedy setup are subpar, users can reduce their requirements and opt for fewer criteria.
+Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.
 
-The results showed that fulfilling several criteria at the same time may be difficult to achieve for small data sets.
-Still, users can try out different objectives and see which specification fulfils their needs.
-A reasonable approach it to start greedy, i.e., requesting multiple distribution moments to be equalized, and only if the results are subpar, go for fewer criteria.
-Unlike in statistical analysis, there is no problem with "just trying out" different approaches when it comes to anticlustering.
-
-K-plus anticlustering was also shown to outperform diversity anticlustering with regard to the specific objectives that are included in the optimization process; diversity is still a reasonable all-arounder that strives for overall betwee-group similarity.
-Which method should be employed depends on a user's preferences.
-In doubt, several anticlustering methods can be applied sequentially and the results can be compared directly.
+[^binarydata]: A reasonable exception is given when the input data consists of binary variables because in this case the mean and the standard deviation are directly related.
 
@@
 
-This paper focused on the presentation of new objective function but the did not properly consider the multicriterion nature of the k-plus objective.
-Instead, a simple (unweighted) sum approach was applied to optimize the k-plus objective, independent of the number of criteria that were incorporated.
-Other approaches, such as direct algorithms, may yield improved results and better possibilities to select partitionings according to differently weighted criteria.
+This paper focused on the presentation of new objective function but the did not really consider the multicriterion nature of the k-plus objective.
+Instead, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
+Other approaches, such as direct algorithms, may yield improved results and better opportunities to select among partitionings according to multiple differently weighted k-plus criteria.
 @brusco2019 showed how direct algorithms can be used to optimize multiple critera in anticlustering applications.
-Their study focused on an approach to find the pareto set of solutions that simultaneously optimized diversity and dispersion, thereby maximizing within-group heterogeneity with regard to both criteria.[^inanticlust]
+Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion -- convincingly arguing that maximizing within-group heterogeneity should incorporate both of these criteria.[^inanticlust] Future research should also investigate how k-plus anticlustering may profit from more sophisticated algorithmic treatment.
 
@@
 
-Focusing on the objetive function only and not on specialized algorithmic optimization approaches followed two reasonings.
-First, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
-Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes.
+Focusing on the objective function only, and not on specialized algorithms, followed two reasonings.
+First, the applications showed that the local maximum search---maximizing an unweighted k-plus sum---yields satisfying results in practice that probably suffice most users.
+Second, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
+Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes tailored to each individual criterion.
 It was of particular interest to compare the k-plus objective with the popular diversity objective.
 @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups, whereas the diversity maintains an appropriate balance between location and spread.
-Using the k-plus extension, however, the k-means criterion regains a lot of attractivity: when optimizing for similarity in means and variance, k-plus anticlustering clearly outperformed the diversity criterion in the simulation study.
-Second, the applications show that the simple sum approach using a general local maximum search yields very satisfying results in practice.
-It seems that the space of solutions in anticlustering settings includes partitionings that fulfil multiple objectives and these partitionings can be found using a local maximum search.
+Using the k-plus extension, however, the k-means criterion regained attractivity: when opting for similarity in means and variances, k-plus anticlustering clearly outperformed the diversity criterion.
 

### 2022-05-30 paper.Rmd
typo
words

@@
 Still, users can try out different objectives and see which specification fulfils their needs.
-A reasonable approach it to start greedy by requesting that multiple distribution moments should equalized.
-Only if the results of the greedy setup are subpar, users can reduce their requirements and opt for fewer criteria.
+A reasonable approach is to start greedy by requesting that multiple distribution moments should equalized.
+If the results of the greedy setup are subpar, users can reduce their requirements by opting for fewer criteria.
 Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.
@@
 Other approaches, such as direct algorithms, may yield improved results and better opportunities to select among partitionings according to multiple differently weighted k-plus criteria.
-@brusco2019 showed how direct algorithms can be used to optimize multiple critera in anticlustering applications.
-Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion -- convincingly arguing that maximizing within-group heterogeneity should incorporate both of these criteria.[^inanticlust] Future research should also investigate how k-plus anticlustering may profit from more sophisticated algorithmic treatment.
+@brusco2019 showed am example of how direct algorithms can be used to optimize multiple critera in anticlustering applications.
+Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion -- convincingly arguing that applications requiring high within-group heterogeneity should incorporate both of these criteria.[^inanticlust] Future research should also investigate how k-plus anticlustering may profit from more sophisticated algorithmic treatment.
 

### 2022-05-31 paper.Rmd
shortened abstract

@@
 Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity.
-While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, higher order moments, and covariance structure.
-To minimize differences in variance, each variable is duplicated by computing the squared deviation between the original values and the variable's mean.
-Because the variance is defined as the expected value of the squared deviations from the mean, appending these duplicated variables to the data set---and then applying standard k-means anticlustering---suffices to equalize the variance between groups.
-Following a similar logic, differences in skewness, kurtosis and covariances can be minimized.
-Thus, while k-plus is a separate criterion for anticlustering, it can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
-A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity with regard to mean attribute values as well as the variance and higher order moments, usually with no noticable sacrifice to either criterion.
-The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity.
+While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, higher order moments like skewness and kurtosis, and even covariance structure.
+It is shown that all k-plus variations can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
+A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity with regard to mutliple distribution characteristics with little sacrifice differences in means.
+The k-plus extension is therefore preferred over classical k-means anticlustering when optimizing between-group similarity.
 Example are given on how k-plus anticlustering can be applied using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

### 2022-05-31 paper.Rmd
work on abstract

@@
 
-    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity and high within-group heterogeneity.
-Anticlustering is usually approached by maximzing instead of minimizing a clustering objective such as the k-means criterion.
-Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity.
-While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, higher order moments like skewness and kurtosis, and even covariance structure.
+    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity or high within-group heterogeneity.
+Anticlustering is usually approached by maximzing instead of minimizing a clustering objective such as the sum of squared errors used in k-means clustering.
+Introducing k-plus, this paper presents an extension of the k-means criterion to maximize between-group similarity.
+While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize differences with regard to multiple distribution characteristics: the variance, higher order moments such as skewness and kurtosis, and even to preserve covariance structure.
+Thus, k-plus corresponds to a family of new anticlustering criteria that can be combined in arbitrary ways.
 It is shown that all k-plus variations can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
-A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity with regard to mutliple distribution characteristics with little sacrifice differences in means.
+A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple criteria with little sacrifice to each single criterion.
 The k-plus extension is therefore preferred over classical k-means anticlustering when optimizing between-group similarity.
-Example are given on how k-plus anticlustering can be applied using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
+Example are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
 
@@
 
-Anticlustering is a collection of methods to partition a set of elements into disjunct groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
+Anticlustering is a collection of methods to partition a set of elements into disjoint groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
 Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990].
 Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
-Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].
+Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjoint groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].
 
@@
 
-## Closing thoughts on the nature of anticlustering
-
-- new criterion: optimizes between-group similarity, no correspondence to within-group heterogeneity [see @brusco2019 for a exceptional method maximizing within-group heterogeneity]
-
-- Similarly: the reversed objective no longer has a useful clustering interpretation: Maximizing differences in variances does not lead to homogeneous and/or well separated clusters
-
 ## Conclusion

### 2022-05-31 paper.Rmd
introduction

@@
 Anticlustering is a collection of methods to partition a set of elements into disjoint groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
-Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990].
-Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
-Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjoint groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].
+Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@feo1990].
+Anticlustering thereby reverses the logic of its better known twin cluster analysis, which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
+Whereas a variety of different approaches exist for cluster analysis---including as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview]---anticlustering is usually formalized as a partitioning process that optimizes an objective function, representing between-group similarity and/or within-group heterogeneity [@papenberg2020; @spath1986; @brusco2019].
 
 Anticlustering is useful in many tasks surrounding research work in psychology.
-For example, anticlustering can be used to split a test into parts of equal difficulty [@gierl2017], to assign students to work groups [@baker2002] or when assigning stimuli to different---but parallel---experimental conditions [@lahl2006].
+For example, anticlustering can be used to split tests into parts of equal difficulty [@gierl2017], to assign students to work groups [@baker2002] or when assigning stimuli to different---but parallel---experimental conditions [@lahl2006].
 For a more comprehensive overview of anticlustering applications in Psychology, see @brusco2019 and @papenberg2020.
@@
 
-Anticlustering methods can be distinguished by the input data and the objective function that is used to optimize group between-group similarity / within-group heterogeneity.
+Anticlustering methods can be distinguished on the basis of the input data and the objective function that is used to quantify between-group similarity or within-group heterogeneity.
 Some anticlustering objectives quantify group similarity on the basis of pairwise dissimilarity rating such as the Euclidean distance.
-The most prominent criterion on the basis of pairwise dissimilarity ratings is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
+The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
 Maximizing the diversity optimizes within-group heterogeneity, which simultaneously leads to similar groups---if all groups have the same size.
-Anticlustering objective functions usually represent the reversal of a clustering objective, only the direction of the optimization is changed.
-Another important criterion based on pairwise dissimilarities is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013; @brusco2019].
+Another important criterion based on pairwise dissimilarities is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
 Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
+@brusco2019 convincingly argued that anticlustering applications requiring high within-group heterogeneity should incorporate both dispersion and diversity, presenting a bicriterion algorithm to find efficient solutions according to both criteria.
 
-Oftentimes, researchers work with attribute values [or *features*; @dry2009] instead of dissimilarity ratings.
+Oftentimes, partitioning applications in psychology use attribute data [or *features*; @dry2009] instead of dissimilarity ratings as input.
 This use case is for example common when selecting stimuli for an experiment on the basis of norming data [e.g., @kurdi2017introducing].
@@
 Subsequently, an anticlustering criterion such as the diversity can be optimized to partition the elements into groups.
-A different approach to anticlustering directly works with the attribute data by maximizing the k-means criterion.
-K-means is probably the best-known clustering method.
-Reversing the k-means objective function---using maximization instead of minimization---has been independently recognized as a useful anticlustering tool by Späth [-@spath1986] and Valev [-@valev1998; -@valev1983].
+A different approach to anticlustering directly works with the attribute data, using an appropriate clustering objective that is computed on the basis of the features directly.
+K-means is probably the best-known clustering method of this kind.
+In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006].
+Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np].
+Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful approach when aiming for high between-group similarity [@spath1986; @valev1998; @valev1983].
+Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters.
+However, other parameters---such as the variance---are ignored.
+@papenberg2020 showed that k-means anticlustering tends to over-optimize similarity with regard to the mean; even an entirely random split usually resulted in more similar variances.
+When requiring that groups are overall similar to each other, k-means' focus may be misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
 
-In k-means clustering, the sum of the squared Euclidean distances between elements and the center of the cluster they are assigned to is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006].
-Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np].
-For the anticlustering application, the k-means criterion is maximized instead, thereby achieving similarity between groups.
-Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters.
-Hence, groups are similar with regard to the mean of the distribution of the numeric attributes wheras other parameters---such as the variance---are ignored.
-When requiring that groups are overall similar to each other, this focus may be misguided as similar means can be obtained even when the underlying distributions are different [e.g., @anscombe1973].
-@papenberg2020 showed that maximizing the diversity (using the Euclidean distance) is better suited to minimize difference with regard to both the mean and the variance of the data; even an entirely random split usually resulted in more similar variances because k-means solely focuses on the features' means.
-
-To optimize overall group similarity---and hence to overcome the limitations of k-means anticlustering---this paper presents k-plus anticlustering.
-The k-plus criterion is an extension of the k-means objective that includes one or several terms to quantify similarity with regard to other important characteristics of a data set rather than only the mean.
-Hence, the k-plus criterion is not a single objective function for anticlustering, but instead a familiy of objectives that extend the standard k-means objective to minimize differences between groups with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
-Which of these objectives is pursued depends on a researchers' needs.
-In the most basic---and arguably most important---case, k-plus anticlustering can be used to not just minimize differences with regard to the mean of variables but also their variance.
-Measures of location and spread are arguably most important to researchers when they "eyeball" their data and they are routinely reported as characterizations of a data set.
-Hence, by default, the k-plus objective is refered to when considering the mean and the variance of variables in the anticlustering process.
-
-While less routinely reported, skewness are kurtosis are also important descriptive characterizations of a data set [@decarlo1997; @westfall2014].
-To optimize overall similarity between groups, it may thus be desirable to also minimize differences with regard to these distribution moments.
-In principle, any other higher order moments can also be included as part of the k-plus objective.
-Futhermore, k-plus anticlustering can also be used to obtain similar covariance structures among groups.
-This approach may be of interest in cross validation applications where prediction accuracy depends on the correlations among features and criteria [@zeng2000; @papenberg2020].
-Interestingly, in each case, the k-plus criterion can be reduced to an augmentation of the original data matrix, leaving the original k-means objective unchanged.
-
-After mathematically introducing the k-plus objective, a simulation study is reported showing that k-plus anticlustering is well suited to create groups that have minimal differences according to multiple distribution characteristics.
-Fulfilling multiple criteria at the same time becomes more feasible as the size of the data set increases.
-In two pratical examples I illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
+To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering.
+The k-plus objective is an extension of the k-means objective that includes one or several terms to quantify between-group similarity with regard to other distribution characteristics and not only the mean.
+Specifically, k-plus offers the possibility to minimize differences with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
+Hence, the k-plus criterion is not a single criterion for anticlustering, but instead a familiy of objectives implementing different requirements with regard to between-group similarity.
+After mathematically introducing the k-plus objective, a simulation study is reported showing that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
+Two pratical are included that illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
 
@@
 Other approaches, such as direct algorithms, may yield improved results and better opportunities to select among partitionings according to multiple differently weighted k-plus criteria.
-@brusco2019 showed am example of how direct algorithms can be used to optimize multiple critera in anticlustering applications.
-Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion -- convincingly arguing that applications requiring high within-group heterogeneity should incorporate both of these criteria.[^inanticlust] Future research should also investigate how k-plus anticlustering may profit from more sophisticated algorithmic treatment.
+@brusco2019 showed am example of how direct algorithms can be used to optimize multiple criteria in anticlustering applications.
+Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion.[^inanticlust] Future research should also investigate how k-plus anticlustering may profit from more sophisticated algorithmic treatment.
 

### 2022-05-31 potpurri.Rmd
introduction

@@
+Which of these objectives is pursued depends on a researchers' needs.
+In the most basic---and arguably most important---case, k-plus anticlustering can be used to not just minimize differences with regard to the mean of variables but also their variance.
+Measures of location and spread are arguably most important to researchers when they "eyeball" their data and they are routinely reported as characterizations of a data set.
+Hence, by default, the k-plus objective is refered to when considering the mean and the variance of variables in the anticlustering process.
+
+While less routinely reported, skewness are kurtosis are also important descriptive characterizations of a data set [@decarlo1997; @westfall2014].
+To optimize overall similarity between groups, it may thus be desirable to also minimize differences with regard to these distribution moments.
+In principle, any other higher order moments can also be included as part of the k-plus objective.
+Futhermore, k-plus anticlustering can also be used to obtain similar covariance structures among groups.
+This approach may be of interest in cross validation applications where prediction accuracy depends on the correlations among features and criteria [@zeng2000; @papenberg2020].
+Interestingly, in each case, the k-plus criterion can be reduced to an augmentation of the original data matrix, leaving the original k-means objective unchanged.
+
 ## Bicriterion optimization model

### 2022-05-31 paper.Rmd
minor editing

@@
 output            : 
-  papaja::apa6_pdf: 
-    number_sections: true
+  papaja::apa6_pdf
 ---
@@
 
-Anticlustering is a collection of methods to partition a set of elements into disjoint groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
+Anticlustering is a collection of methods to partition elements into disjoint groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
 Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@feo1990].
@@
 Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
-When defining $z_i = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and $y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
+This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining $z_i = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and $y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
 
@@
 That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending these to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data matrix.
-Hence, k-plus anticlustering reduces to an augmentation of the data matrix, leaving the original k-means criterion unchanged.
-
-During the remainder of this paper, whenever the k-plus criterion is optimized, the original data matrix is simply extended and the original k-means criterion $SSE$ is optimized.
-For a multicriterion objective such as the k-plus criterion, this corresponds to a rather basic approach, and future research may investigate more sophisticated optimization schemes for k-plus anticlustering [e.g., see @brusco2012; @brusco2019].
-In this paper, however, I am primarily concerned with investigating the properties of new objective itself rather than extending the state of the art on multicriterion optimization methods.
+Hence, using the unweighted sum approach, k-plus anticlustering reduces to an augmentation of the data matrix, leaving the original k-means criterion unchanged.
 
@@
 To thorougly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.
+To implement all k-plus objectives, the simple unweighted sum outlined above approach was employed.
+Thus, when optimizing a k-plus criterion, the original data matrix was extended to incorporate the additional criteria.
+Subsequently, the original k-means criterion $SSE$ was maximized on the basis of the augmented data set.
 

### 2022-05-31 potpurri.Rmd
minor editing

@@
+During the remainder of this paper, whenever the k-plus criterion is optimized, the original data matrix is simply extended and the original k-means criterion $SSE$ is optimized.
+For a multicriterion objective such as the k-plus criterion, this corresponds to a rather basic approach, and future research may investigate more sophisticated optimization schemes for k-plus anticlustering [e.g., see @brusco2012; @brusco2019].
+In this paper, however, I am primarily concerned with investigating the properties of new objective itself rather than extending the state of the art on multicriterion optimization methods.
+
 Which of these objectives is pursued depends on a researchers' needs.

### 2022-06-21 paper.Rmd
minor editing in example and discussion

@@
 This is due the diversity criterion being a measure of within-group heterogeneity; maximizing within-group heterogeneity only incidentally maximizes between-group similarity.
-However, because the diversity criterion is a sum rather than an average, the correspondence between within-group heterogeneity and between-group similarity no longer holds unequal-sized groups.
+Because the diversity criterion is a sum rather than an average, the correspondence between within-group heterogeneity and between-group similarity does not hold for unequal-sized groups.
 K-plus anticlustering strives for between-group similarity regardless of group size.
 
-Unequal group sizes can be requested in anticlust by passing the different group sizes to the argument `K`.
+Unequal group sizes can be requested in `anticlust` by passing the different group sizes to the argument `K`.
 The following code was used to equalize means and variances between four groups using the k-plus approach:
@@
 
-This function call relied on the default k-plus settings, ensuring that variances are equalized -- but no higher order moments or covariances.
+This function call relied on the default k-plus settings ensuring similarity with regard to means and variances equalized---but with regard to higher order moments or covariances.
 
@@
 In this paper, I presented the k-plus criterion for anticlustering, which can be used to divided a data set into subgroups in such a way that the groups are similar to each other.
+The k-plus criterion represents between-group similarity as discrepancies in distribution moments: mean, variances, skewness, kurtosis, any higher order moments, and covariances.
+Depending on the application, a different selection of these objectives can be included in the anticlustering process.
 The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means.
-The k-plus criterion also equalizes the variance between groups, and if needed any higher order moments such as skewness and kurtosis, or even the covariances.
-Thus, k-plus represents a family of anticlustering criteria, which can be combined by users according to their needs.
-Interestingly, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
+Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
 
 A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.
-In particular, this work indicates that k-plus anticlustering should be a replacement of the classical k-means approach when striving for between-group similarity.[^binarydata]  K-means anticlustering disregards any distribution characteristics other than the mean, while k-plus anticlustering ensures between-group similarity with regard to multiple criteria---and still minimizes differences in means quite well.
-The results showed that fulfilling several criteria at the same time may be aggregated in small data sets.
-Still, users can try out different objectives and see which specification fulfils their needs.
+Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, this work indicates that k-plus anticlustering should be a default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be aggregated, especially in small data sets.
+In practice, users can try out different objectives and see which specification fulfils their needs.
 A reasonable approach is to start greedy by requesting that multiple distribution moments should equalized.
-If the results of the greedy setup are subpar, users can reduce their requirements by opting for fewer criteria.
+If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria.
 Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.
+Using the R package `anticlust`, application of k-plus anticlustering is easy and free.
 
-[^binarydata]: A reasonable exception is given when the input data consists of binary variables because in this case the mean and the standard deviation are directly related.
+[^binarydata]: A reasonable exception is given when the input data consists of binary variables because for binary data the mean and the variance are directly related.
 
@@
 
-This paper focused on the presentation of new objective function but the did not really consider the multicriterion nature of the k-plus objective.
+This paper focused on the presentation of new objective function but did not really consider the multicriterion nature of the k-plus criterion.
 Instead, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
 Other approaches, such as direct algorithms, may yield improved results and better opportunities to select among partitionings according to multiple differently weighted k-plus criteria.
-@brusco2019 showed am example of how direct algorithms can be used to optimize multiple criteria in anticlustering applications.
-Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion.[^inanticlust] Future research should also investigate how k-plus anticlustering may profit from more sophisticated algorithmic treatment.
+For example, @brusco2019 showed how direct algorithms can be used to optimize multiple criteria in anticlustering applications.
+Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion in applications that require high within-group heterogeneity.[^inanticlust] Future research should investigate how k-plus anticlustering may profit from similar---more sophisticated---algorithmic treatment.
 
@@
 
-Focusing on the objective function only, and not on specialized algorithms, followed two reasonings.
-First, the applications showed that the local maximum search---maximizing an unweighted k-plus sum---yields satisfying results in practice that probably suffice most users.
-Second, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
-Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes tailored to each individual criterion.
-It was of particular interest to compare the k-plus objective with the popular diversity objective.
-@papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups, whereas the diversity maintains an appropriate balance between location and spread.
-Using the k-plus extension, however, the k-means criterion regained attractivity: when opting for similarity in means and variances, k-plus anticlustering clearly outperformed the diversity criterion.
+Focusing only on the objective function and not on optimization algorithms followed two reasonings.
+First, the applications showed that the local maximum search---maximizing the unweighted k-plus sum---yields satisfying results in practice that probably suffice most applications.
+Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
+In the simulation study, to obtain a fair assessment, it was important that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes tailored to each individual criterion.
+It was of particular interest to compare the k-plus objective with the common diversity objective.
+@papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups.
+In contrast, maximizing the diversity leads to an appropriate balance between location and spread.
+Using the k-plus extension, however, the k-means criterion has regained attractivity: when opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.
 
 [^clusterediting]: In their study, they used the term anticluster editing to refer to the maximization of the diversity criterion.
-The diversity criterion is the objective function used in the clustering method cluster editing, where it has to be minimized [@zahn1964; @shamir2004vf].
+This is because the diversity criterion is minimized as part of the cluster editing method [@zahn1964; @shamir2004vf].
 
@@
 
-This paper introduced k-plus anticlustering for maximizing between-group similarity, improving the traditional k-means approach.
-K-plus anticlustering can be adopted whenever a data set has to be partitioned into equivalent parts.
+This paper introduced k-plus anticlustering for maximizing between-group similarity, improving the classical k-means approach.
+K-plus anticlustering can be adopted whenever a group of elements has to be partitioned into equivalent parts.
 The method is accessible and easily usable via the R package `anticlust`.

### 2022-06-22 paper.Rmd
editing abstract and introduction

@@
 
-    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity or high within-group heterogeneity.
-Anticlustering is usually approached by maximzing instead of minimizing a clustering objective such as the sum of squared errors used in k-means clustering.
-Introducing k-plus, this paper presents an extension of the k-means criterion to maximize between-group similarity.
-While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize differences with regard to multiple distribution characteristics: the variance, higher order moments such as skewness and kurtosis, and even to preserve covariance structure.
-Thus, k-plus corresponds to a family of new anticlustering criteria that can be combined in arbitrary ways.
+    Anticlustering is a collection of methods to partition data sets into disjoint groups while aiming for high between-group similarity or high within-group heterogeneity.
+Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximzing instead of minimizing a clustering objective function.
+Introducing k-plus, this paper presents an extension of the k-means clustering objective to maximize between-group similarity in anticlustering applications.
+While k-means anticlustering minimizes between-group differences only with regard to mean attribute values, k-plus anticlustering can be used to minimize differences with regard to several additional distribution characteristics, including the variance, skewness and kurtosis, and even covariance structure.
 It is shown that all k-plus variations can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
-A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple criteria with little sacrifice to each single criterion.
-The k-plus extension is therefore preferred over classical k-means anticlustering when optimizing between-group similarity.
+A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple criteria.
+In particular, optimizing similarity with regard to variances as well as means comes with almost no cost to either criterion; the k-plus extension is therefore generally preferred over classical k-means anticlustering.
+Higher order moments and covariance structure can also succesfully be equalized, while fulfilling multiple criteria at the same time is improved when the sample size increases.
 Example are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
@@
 Anticlustering thereby reverses the logic of its better known twin cluster analysis, which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
-Whereas a variety of different approaches exist for cluster analysis---including as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview]---anticlustering is usually formalized as a partitioning process that optimizes an objective function, representing between-group similarity and/or within-group heterogeneity [@papenberg2020; @spath1986; @brusco2019].
+Whereas a variety of different approaches exist for cluster analysis---including hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview]---anticlustering is usually formalized as a partitioning process that optimizes an objective function that represents between-group similarity and/or within-group heterogeneity [@papenberg2020; @spath1986; @brusco2019].
 
-Anticlustering is useful in many tasks surrounding research work in psychology.
+Anticlustering is useful in many tasks surrounding research work in psychology [@steinley2006].
 For example, anticlustering can be used to split tests into parts of equal difficulty [@gierl2017], to assign students to work groups [@baker2002] or when assigning stimuli to different---but parallel---experimental conditions [@lahl2006].
@@
 Anticlustering methods can be distinguished on the basis of the input data and the objective function that is used to quantify between-group similarity or within-group heterogeneity.
-Some anticlustering objectives quantify group similarity on the basis of pairwise dissimilarity rating such as the Euclidean distance.
-The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group.
-Maximizing the diversity optimizes within-group heterogeneity, which simultaneously leads to similar groups---if all groups have the same size.
-Another important criterion based on pairwise dissimilarities is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
+Some anticlustering objectives are based on pairwise dissimilarity ratings such as the Euclidean distance.
+The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group [@brusco2019].
+By summing dissimilarities within-groups, the diversity is primarily a measure of within-group heterogeneity.
+For equal-sized groups, however, maximizing the diversity simultaneously minimizes the average dissimilarity between elements in different groups, thereby maximizing between-group similarity.
+
+Another important anticlustering criterion that is based on pairwise dissimilarities is the *dispersion*, i.e., the minimum dissimilarity between any two elements within the same group [@fernandez2013].
 Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
-@brusco2019 convincingly argued that anticlustering applications requiring high within-group heterogeneity should incorporate both dispersion and diversity, presenting a bicriterion algorithm to find efficient solutions according to both criteria.
+@brusco2019 convincingly argued that anticlustering applications requiring high within-group heterogeneity should incorporate both dispersion and diversity, presenting a bicriterion algorithm to approximate the Pareto efficient set of solutions according to both criteria.
+
+[^averagesum]:
 
@@
 In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019].
-Subsequently, an anticlustering criterion such as the diversity can be optimized to partition the elements into groups.
-A different approach to anticlustering directly works with the attribute data, using an appropriate clustering objective that is computed on the basis of the features directly.
+Subsequently, an anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the elements.
+A different anticlusting approach is based on clustering objectives that can directly be computed on the attribute values.
 K-means is probably the best-known clustering method of this kind.
@@
 To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering.
-The k-plus objective is an extension of the k-means objective that includes one or several terms to quantify between-group similarity with regard to other distribution characteristics and not only the mean.
+The k-plus objective is an extension of the k-means objective that includes one or several terms to quantify between-group similarity with regard to other distribution characteristics instead of only the mean.
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
-Hence, the k-plus criterion is not a single criterion for anticlustering, but instead a familiy of objectives implementing different requirements with regard to between-group similarity.
-After mathematically introducing the k-plus objective, a simulation study is reported showing that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
+Hence, the k-plus criterion is not a single criterion for anticlustering, but a familiy of objectives implementing different requirements with regard to between-group similarity.
+After formally introducing the k-plus objective, a simulation study is reported showing that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
 Two pratical are included that illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
@@
 
-This section demonstrates how researchers and other data analysists can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020], using an interface function called `kplus_anticlustering()`, which is available from version 0.6.3 onward.
+This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020], using an interface function called `kplus_anticlustering()`, which is available from version 0.6.3 onward.
 The package can be installed in the R environment using the `install.packages()` command:

### 2022-06-22 paper.Rmd
remove unused footnote
some minor editing

@@
 
-[^averagesum]:
-
 Oftentimes, partitioning applications in psychology use attribute data [or *features*; @dry2009] instead of dissimilarity ratings as input.
@@
 In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019].
-Subsequently, an anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the elements.
+Subsequently, an anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the data set.
 A different anticlusting approach is based on clustering objectives that can directly be computed on the attribute values.
-K-means is probably the best-known clustering method of this kind.
+K-means is probably the best-known method of this kind.
 In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006].
@@
 To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering.
-The k-plus objective is an extension of the k-means objective that includes one or several terms to quantify between-group similarity with regard to other distribution characteristics instead of only the mean.
+K-plus is an extension of the k-means objective that includes one or several terms to quantify between-group similarity with regard to other distribution characteristics instead of only the mean.
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
-Hence, the k-plus criterion is not a single criterion for anticlustering, but a familiy of objectives implementing different requirements with regard to between-group similarity.
+The k-plus criterion is not a single criterion for anticlustering, but a familiy of objectives implementing different requirements with regard to between-group similarity.
 After formally introducing the k-plus objective, a simulation study is reported showing that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
@@
 
-In anticlustering applications, we impose an additional constraint on the group sizes $n_k$, where the most common restriction requires that all groups have equal size:
+In anticlustering applications, an additional constraint is usually imposed on the group sizes $n_k$, where the most common restriction requires that all groups have equal size:
 
@@
 Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective, i.e., when just appending the squared-distance-features to the data set.
-Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ have much larger values, on average, than the raw values $x_{ij}$.
+Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ usually have much larger values, on average, than the raw values $x_{ij}$.
 Therefore, the $\mathit{SSE_{Var}}$ criterion naturally receives greater weight when computing $\mathit{SSE_{kplus}}$ than the standard $\mathit{SSE}$ criterion.
@@
 
-Thus, to maximize between-groups similarity with regard to covariance structure, it is possible to add an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
+Thus, to maximize between-group similarity with regard to covariance structure, it is possible to add an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
 This requires computing $\binom{N}{2}$ new variables and thus turns out be be rather costly: The number of additional variables needed to minimize differences with regard to covariance structure grows quadratically with increasing number of variables; the number of additional variables needed to minimize differences with regard to distribution moments only grows linearly with increasing number of variables.
@@
 
-To thorougly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.
+To thoroughly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.
 To implement all k-plus objectives, the simple unweighted sum outlined above approach was employed.
@@
 
-```{r Figure1, fig.width = 8, fig.height = 10, fig.cap = "Simulation results by $N$."}
+```{r Figure1, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by $N$."}
 
 # Plot the results, by N
-ldf %>% 
+ldf <- ldf %>% 
   group_by(method, Objective, N) %>% 
@@
       labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
-  ) %>%
+  )
+
+ldf %>%
   ggplot(aes(x = N, y = Mean, colour = method)) + 
@@
 Table 1 displays the global simulation results aggregated across all conditions.
-Standard k-means performed best at minimizing discrepancy with regard to means, but disregards all other distribution characteristics.
-K-plus anticlustering also addresses $\Delta \mathit{M}$ quite well and at the same time tends to minimize discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$).
-With regard to minimizing $\Delta \mathit{SD}$, all three k-plus approaches outperform all other methods.
+Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but disregarded all other distribution characteristics, where it did not outperform a random assignment.
+K-plus anticlustering also addressed $\Delta \mathit{M}$ quite well and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); all three k-plus approaches outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
+
 As Figure 1 shows, with increasing $N$, k-means and all k-plus methods converge on minimizing $\Delta \mathit{M}$.
-This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time.
+This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time---especially when $N$ increases.
+However, while k-plus anticlustering is capable of addressing multiple objectives simultaneously, it is apparent that increasing the number of optimization criteria aggravated fulfilling each single criterion.
+K-plus skew/kurtosis and k-plus correlation performed worse at minimizing discrepancy with regard to standard deviations than standard k-plus, which itself performed slightly worse than k-means anticlustering.
+Still, every k-plus method did what it was supposed to do: k-plus skew/kurtosis was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation was best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$.
+At the same time, these two k-plus objectives maintained a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
 
-While k-plus anticlustering is capable of addressing multiple objectives simultaneously, it is apparent that increasing the number of optimization criteria aggravates fulfilling each single criterion.
-K-plus skew/kurtosis and k-plus correlation are worse at minimizing discrepancy with regard to standard deviations than standard k-plus, which is slightly worse at minimizing $\Delta \mathit{M}$ than k-means anticlustering.
-Still, every k-plus method does what it is supposed to do: k-plus skew/kurtosis is best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation is best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$.
-At the same time, these two k-plus objectives maintain a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
-
-It is maybe surprising that k-plus skew/kurtosis does not seem to be much better than diversity anticlustering with regard to skew and---in particular---kurtosis, even though the method has been specifically tailored to these criteria.
-Figure 1 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis is even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
-However, with increasing $N$, it clearly outperforms diversity anticlustering.
+It is maybe surprising that k-plus skew/kurtosis does not seem to have performed much better than diversity anticlustering with regard to skewness and---in particular---kurtosis, even though the method has been specifically tailored to these criteria.
+Figure 1 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis was even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
+However, with increasing $N$, it clearly outperformed diversity anticlustering.
 These results illustrate the cost of optimizing several criteria at the same time: for low $N$, k-plus seems skew/kurtosis to be preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
@@
 
-Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to somewhat address all distribution characteristics.
-For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering is outperformed.
-As as side note, apart from k-plus correlation, diversity anticlustering is the only method that equalizes correlation structure among groups.
+Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to somewhat addressed all distribution characteristics.
+For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering was outperformed.
+As as side note, apart from k-plus correlation, diversity anticlustering was the only method that equalizes correlation structure among groups.
 As far as I know, this observation has not been made previously, even though the maximum diversity problem has been widely studied.

### 2022-06-22 Simulation_Study/4-Aggregate-Results.R

@@ -54,6 +54,13 @@ ldf %>%
 ldf %>% 
   group_by(method, Objective, M) %>% 
   summarise(Mean = mean(value)) %>% 
+  filter(method != "random") %>% 
+  mutate(
+    Objective = ordered(
+      Objective, 
+      levels = c("means", "sd", "skew", "kur", "cor"),
+      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
+  ) %>% 
   ggplot(aes(x = M, y = Mean, colour = method)) + 
   geom_line(size = 1) + 
   facet_grid(rows = vars(Objective), scales = "free") + 
@@ -63,6 +70,13 @@ ldf %>%
 ldf %>% 
   group_by(method, Objective, SD) %>% 
   summarise(Mean = mean(value)) %>% 
+  filter(method != "random") %>% 
+  mutate(
+    Objective = ordered(
+      Objective, 
+      levels = c("means", "sd", "skew", "kur", "cor"),
+      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
+  ) %>% 
   ggplot(aes(x = SD, y = Mean, colour = method)) + 
   geom_point(size = 3) + 
   geom_line(size = 1) + 

### 2022-06-22 Simulation_Study/4-Aggregate-Results.R
plots for appendix

@@ -83,20 +83,17 @@ ldf %>%
   facet_grid(rows = vars(Objective), scales = "free") + 
   theme_bw(base_size = 22)
 
-# Plot the results, for delte(Mean) by SD and N
-ldf %>% 
-  group_by(method, Objective, SD, N) %>% 
-  summarise(Mean = mean(value)) %>% 
-  filter(Objective %in% c("means")) %>% 
-  ggplot(aes(x = N, y = Mean, colour = method)) + 
-  geom_line(size = 1) + 
-  facet_grid(cols = vars(SD)) + 
-  theme_bw(base_size = 22)
-
 # Plot the results, by K
 ldf %>% 
   group_by(method, Objective, K) %>% 
   summarise(Mean = mean(value)) %>% 
+  filter(method != "random") %>% 
+  mutate(
+    Objective = ordered(
+      Objective, 
+      levels = c("means", "sd", "skew", "kur", "cor"),
+      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
+  ) %>% 
   ggplot(aes(x = K, y = Mean, colour = method)) + 
   geom_point(size = 3) + 
   geom_line(size = 1) + 
@@ -107,6 +104,13 @@ ldf %>%
 ldf %>% 
   group_by(method, Objective, r) %>% 
   summarise(Mean = mean(value)) %>% 
+  filter(method != "random") %>% 
+  mutate(
+    Objective = ordered(
+      Objective, 
+      levels = c("means", "sd", "skew", "kur", "cor"),
+      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
+  ) %>% 
   ggplot(aes(x = r, y = Mean, colour = method)) + 
   geom_point(size = 3) + 
   geom_line(size = 1) + 

### 2022-06-22 paper.Rmd
fix

@@
 # Plot the results, by N
-ldf <- ldf %>% 
+ldf %>% 
   group_by(method, Objective, N) %>% 
@@
       labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
-  )
-
-ldf %>%
+  ) %>%
   ggplot(aes(x = N, y = Mean, colour = method)) + 

### 2022-06-22 paper.Rmd
minor stuff

@@
 Table 1 displays the global simulation results aggregated across all conditions.
-Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but disregarded all other distribution characteristics, where it did not outperform a random assignment.
-K-plus anticlustering also addressed $\Delta \mathit{M}$ quite well and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); all three k-plus approaches outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
+Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but did not outperform a random assignment with regard to all other distribution characteristics; replicating @papenberg2020, k-means' performance with regard to similarity in standard deviations was even below that of a random assignment.
+K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); all three k-plus approaches outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
 
@@
 This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time---especially when $N$ increases.
-However, while k-plus anticlustering is capable of addressing multiple objectives simultaneously, it is apparent that increasing the number of optimization criteria aggravated fulfilling each single criterion.
+However, while k-plus anticlustering is capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria apparently aggravates fulfilling each single criterion.
 K-plus skew/kurtosis and k-plus correlation performed worse at minimizing discrepancy with regard to standard deviations than standard k-plus, which itself performed slightly worse than k-means anticlustering.

### 2022-11-29 paper.Rmd
worked on discussion

@@
 
-In this paper, I presented the k-plus criterion for anticlustering, which can be used to divided a data set into subgroups in such a way that the groups are similar to each other.
-The k-plus criterion represents between-group similarity as discrepancies in distribution moments: mean, variances, skewness, kurtosis, any higher order moments, and covariances.
-Depending on the application, a different selection of these objectives can be included in the anticlustering process.
+Anticlustering is a partitioning method that aims for high between-group similarity and within-group heterogeneity.
+In this paper, I presented the k-plus criterion for anticlustering, which focuses on establishing groups that are similar to each other.
+The k-plus criterion represents between-group similarity as discrepancies in distribution moments: mean, variances, skewness, kurtosis, higher order moments, and even covariances.
+Depending on the application, a different subset of these objectives can be included in the anticlustering process.
 The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means.
@@
 A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.
-Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, this work indicates that k-plus anticlustering should be a default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be aggregated, especially in small data sets.
+Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, this work indicates that k-plus anticlustering should be a default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time is aggregated, especially in small data sets.
 In practice, users can try out different objectives and see which specification fulfils their needs.
-A reasonable approach is to start greedy by requesting that multiple distribution moments should equalized.
+A reasonable approach is to start "greedy", e.g.
+by specifying that multiple distribution moments should equalized.
 If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria.
 Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.
-Using the R package `anticlust`, application of k-plus anticlustering is easy and free.
+Moreover, using the open source R package `anticlust`, the application of k-plus anticlustering is easily accessible and free.
 
-[^binarydata]: A reasonable exception is given when the input data consists of binary variables because for binary data the mean and the variance are directly related.
+[^binarydata]: A reasonable exception is given when the input data consists of binary variables because for binary data, the mean and the variance are directly dependent.
 
@@
 
-This paper focused on the presentation of new objective function but did not really consider the multicriterion nature of the k-plus criterion.
-Instead, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
-Other approaches, such as direct algorithms, may yield improved results and better opportunities to select among partitionings according to multiple differently weighted k-plus criteria.
-For example, @brusco2019 showed how direct algorithms can be used to optimize multiple criteria in anticlustering applications.
-Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion in applications that require high within-group heterogeneity.[^inanticlust] Future research should investigate how k-plus anticlustering may profit from similar---more sophisticated---algorithmic treatment.
+The present paper focused on a new criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion.
+Instead, a simple (unweighted) sum approach, combined with a generic local maximum neighbourhood search, was used to optimize the k-plus objective.
+Other approaches may yield improved results and opportunities to select among partitionings that are preferred according to the different k-plus criteria.
+@brusco2019 showed how direct algorithms can be used to optimize multiple criteria at the same time in anticlustering applications.
+Their study focused on finding the pareto set of efficient solutions that simultaneously optimize diversity and dispersion, which is essential in applications that require high within-group heterogeneity.
+Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.
 
-[^inanticlust]: The `anticlust` package now also includes an implementation of the bicriterion anticlustering algorithm by @brusco2019.
-
-Focusing only on the objective function and not on optimization algorithms followed two reasonings.
+This paper's focus on the k-plus objective only and not on optimization algorithms followed two reasonings.
 First, the applications showed that the local maximum search---maximizing the unweighted k-plus sum---yields satisfying results in practice that probably suffice most applications.
 Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
-In the simulation study, to obtain a fair assessment, it was important that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes tailored to each individual criterion.
+To obtain a fair assessment In the simulation study, it was therefore important that each anticlustering objective is optimized using the same algorithm, instead of using specialized optimization schemes tailored to each individual criterion.
 It was of particular interest to compare the k-plus objective with the common diversity objective.
-@papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups.
+@papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should be preferred because k-means does not equalize the spread of the distribution between groups.
 In contrast, maximizing the diversity leads to an appropriate balance between location and spread.
-Using the k-plus extension, however, the k-means criterion has regained attractivity: when opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.
+Using the k-plus extension, however, the k-means criterion has regained attractivity: When opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.
 

### 2022-11-29 paper.Rmd
minor

@@
 After formally introducing the k-plus objective, a simulation study is reported showing that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
-Two pratical are included that illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
+Two practical examples are included that illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
 
@@
 As as side note, apart from k-plus correlation, diversity anticlustering was the only method that equalizes correlation structure among groups.
-As far as I know, this observation has not been made previously, even though the maximum diversity problem has been widely studied.
+To the best of my knowledge, this observation has not been made previously, even though the maximum diversity problem has been widely studied.
 
@@
 Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups.
-This is due the diversity criterion being a measure of within-group heterogeneity; maximizing within-group heterogeneity only incidentally maximizes between-group similarity.
+This is due the diversity criterion being a measure of within-group heterogeneity.
 Because the diversity criterion is a sum rather than an average, the correspondence between within-group heterogeneity and between-group similarity does not hold for unequal-sized groups.

### 2022-11-29 paper.Rmd
minor in discussion

@@
 Anticlustering is a partitioning method that aims for high between-group similarity and within-group heterogeneity.
-In this paper, I presented the k-plus criterion for anticlustering, which focuses on establishing groups that are similar to each other.
+In this paper, I presented the k-plus criterion, which focuses on the between-group similarity aspect of anticlustering.
 The k-plus criterion represents between-group similarity as discrepancies in distribution moments: mean, variances, skewness, kurtosis, higher order moments, and even covariances.
@@
 by specifying that multiple distribution moments should equalized.
-If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria.
+If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variances).
 Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.
@@
 The present paper focused on a new criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion.
-Instead, a simple (unweighted) sum approach, combined with a generic local maximum neighbourhood search, was used to optimize the k-plus objective.
-Other approaches may yield improved results and opportunities to select among partitionings that are preferred according to the different k-plus criteria.
+Instead of applying a specialized multicriterion optimization algorithm, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
+More sophisticated approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus criteria.
 @brusco2019 showed how direct algorithms can be used to optimize multiple criteria at the same time in anticlustering applications.
-Their study focused on finding the pareto set of efficient solutions that simultaneously optimize diversity and dispersion, which is essential in applications that require high within-group heterogeneity.
+Their study focused on approximating the pareto set of efficient solutions that optimize diversity and dispersion, which is essential in applications that require high within-group heterogeneity.
 Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.
 
-This paper's focus on the k-plus objective only and not on optimization algorithms followed two reasonings.
+This paper's focus on the k-plus objective only instead of optimization algorithms followed two reasonings.
 First, the applications showed that the local maximum search---maximizing the unweighted k-plus sum---yields satisfying results in practice that probably suffice most applications.

### 2022-12-06 paper.Rmd
discussion

@@
 Anticlustering is a partitioning method that aims for high between-group similarity and within-group heterogeneity.
-In this paper, I presented the k-plus criterion, which focuses on the between-group similarity aspect of anticlustering.
-The k-plus criterion represents between-group similarity as discrepancies in distribution moments: mean, variances, skewness, kurtosis, higher order moments, and even covariances.
-Depending on the application, a different subset of these objectives can be included in the anticlustering process.
-The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means.
-Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
+In this paper, I presented the k-plus criterion, which focuses on maximizing between-group similarity.
+K-plus represents between-group similarity as discrepancies in distribution moments between groups: mean, variances, skewness, kurtosis, higher order moments, and even covariances.
+Depending on the application and the needs of a user, a different subset of these objectives can be included in the anticlustering process.
+The k-plus criterion is an extension of the classical k-means criterion, which only represents how similar groups are with regard to their means.
+Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by merely augmenting the input data.
 
 A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.
-Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, this work indicates that k-plus anticlustering should be a default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time is aggregated, especially in small data sets.
+Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be a default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be aggravated in small data sets.
 In practice, users can try out different objectives and see which specification fulfils their needs.
@@
 Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.
-Moreover, using the open source R package `anticlust`, the application of k-plus anticlustering is easily accessible and free.
+Moreover, using the open source R package `anticlust`, the (repeated) application of k-plus anticlustering is easily accessible and free.
 
@@
 
-The present paper focused on a new criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion.
+The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion.
 Instead of applying a specialized multicriterion optimization algorithm, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
-More sophisticated approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus criteria.
+The same approach was used independent of the number of objectives that were included as part of the anticlustering process.
+More sophisticated algorithmic approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
 @brusco2019 showed how direct algorithms can be used to optimize multiple criteria at the same time in anticlustering applications.
@@
 To obtain a fair assessment In the simulation study, it was therefore important that each anticlustering objective is optimized using the same algorithm, instead of using specialized optimization schemes tailored to each individual criterion.
-It was of particular interest to compare the k-plus objective with the common diversity objective.
+It was of particular interest to compare the k-plus objective with the popular diversity objective.
 @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should be preferred because k-means does not equalize the spread of the distribution between groups.

### 2022-12-07 paper.Rmd
Abstract

@@
 Introducing k-plus, this paper presents an extension of the k-means clustering objective to maximize between-group similarity in anticlustering applications.
-While k-means anticlustering minimizes between-group differences only with regard to mean attribute values, k-plus anticlustering can be used to minimize differences with regard to several additional distribution characteristics, including the variance, skewness and kurtosis, and even covariance structure.
-It is shown that all k-plus variations can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
-A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple criteria.
-In particular, optimizing similarity with regard to variances as well as means comes with almost no cost to either criterion; the k-plus extension is therefore generally preferred over classical k-means anticlustering.
-Higher order moments and covariance structure can also succesfully be equalized, while fulfilling multiple criteria at the same time is improved when the sample size increases.
+The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, higher order moments, and covariances), thereby extending the classical k-means criterion, which only reflects how groups differ in their means.
+It is shown that the k-plus criterion can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
+A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple objectives.
+In particular, optimizing between-group similarity with regard to variances as well as means comes with almost no cost to either criterion; the k-plus extension is therefore generally preferred over classical k-means anticlustering.
+Higher order moments and covariance structure can also succesfully be equalized.
 Example are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
@@
 
-keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, higher order moments"
+keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis"
 wordcount         : "X"

### 2022-12-27 paper.Rmd
round of editing

@@
 
-    Anticlustering is a collection of methods to partition data sets into disjoint groups while aiming for high between-group similarity or high within-group heterogeneity.
+    Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity.
 Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximzing instead of minimizing a clustering objective function.
@@
 
-Anticlustering is a collection of methods to partition elements into disjoint groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
-Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@feo1990].
-Anticlustering thereby reverses the logic of its better known twin cluster analysis, which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006].
-Whereas a variety of different approaches exist for cluster analysis---including hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview]---anticlustering is usually formalized as a partitioning process that optimizes an objective function that represents between-group similarity and/or within-group heterogeneity [@papenberg2020; @spath1986; @brusco2019].
-
-Anticlustering is useful in many tasks surrounding research work in psychology [@steinley2006].
-For example, anticlustering can be used to split tests into parts of equal difficulty [@gierl2017], to assign students to work groups [@baker2002] or when assigning stimuli to different---but parallel---experimental conditions [@lahl2006].
-For a more comprehensive overview of anticlustering applications in Psychology, see @brusco2019 and @papenberg2020.
+Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
+Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and separation between clusters [@rokach2005; @steinley2006].
+Anticlustering has many applications in research psychology [@steinley2006; @brusco2019; @papenberg2020]: Examples include splitting tests into parts of equal difficulty [@gierl2017], assigning students to work groups [@baker2002], and assigning stimuli to different, but parallel, experimental conditions [@lahl2006].
 Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual partitioning is usually subpar.
@@
 
-Anticlustering methods can be distinguished on the basis of the input data and the objective function that is used to quantify between-group similarity or within-group heterogeneity.
-Some anticlustering objectives are based on pairwise dissimilarity ratings such as the Euclidean distance.
-The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group [@brusco2019].
-By summing dissimilarities within-groups, the diversity is primarily a measure of within-group heterogeneity.
-For equal-sized groups, however, maximizing the diversity simultaneously minimizes the average dissimilarity between elements in different groups, thereby maximizing between-group similarity.
-
-Another important anticlustering criterion that is based on pairwise dissimilarities is the *dispersion*, i.e., the minimum dissimilarity between any two elements within the same group [@fernandez2013].
+Anticlustering methods are characterized by (a) an objective function that quantifies between-group similarity and/or within-group heterogeneity and (b) an algorithm that determines how elements are assigned to groups in such a way that the objective function is maximized.
+Several anticlustering objective functions use pairwise dissimilarity ratings such as the Euclidean distance as input.
+The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group [@brusco2019; @gallego2013].
+By summing all pairwise dissimilarities of elements within the same group, the diversity reflects the total degree of within-group heterogeneity.
+High within-group diversity simultaneously ensures high similarity of the different groups.[^equalsizedgroups] Another important anticlustering criterion is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
 Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
@@
 
-Oftentimes, partitioning applications in psychology use attribute data [or *features*; @dry2009] instead of dissimilarity ratings as input.
-This use case is for example common when selecting stimuli for an experiment on the basis of norming data [e.g., @kurdi2017introducing].
+[^equalsizedgroups]: As will be discussed later, the diversity objective only stricly equalizes within-group heterogeneity and between-group similarity when groups are equal-sized.
+
+Oftentimes, partitioning applications in psychology are focused on ensuring between-group similarity rather than within-group heterogeneity (even though both goals oftentimes coincide).
+High similarity between-group similarity is for example desireable when designing experimental conditions that have to incorporate different stimulus sets that should however be as similar as possible according to response-relevant attributes [@lahl2006].
+Such stimulus assignment tasks are usually conducted on the basis of attribute data [or *features*; @dry2009] rather than pairwise dissimilarity ratings, for example when selecting stimuli for an experiment on the basis of norming data [e.g., @kurdi2017introducing].
 For word stimuli in psycholinguistic experiments, norming data may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982].
@@
 In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019].
-Subsequently, an anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the data set.
-A different anticlusting approach is based on clustering objectives that can directly be computed on the attribute values.
-K-means is probably the best-known method of this kind.
+Subsequently, an anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the stimuli.
+A different anticlusting approach employs clustering objectives that can directly be computed on the attribute values.
+K-means is probably the best-known objective of this kind.
 In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006].
@@
 Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters.
-However, other parameters---such as the variance---are ignored.
-@papenberg2020 showed that k-means anticlustering tends to over-optimize similarity with regard to the mean; even an entirely random split usually resulted in more similar variances.
-When requiring that groups are overall similar to each other, k-means' focus may be misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
+However, other parameters---such as the variance---are not targeted.
+@papenberg2020 showed that k-means anticlustering tends to over-optimize similarity with regard to the mean; even an entirely random group assignment usually yielded more similar variances.
+When requiring overall between-group similarity, the negligence of other distribution characteristics than the mean is misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
 
 To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering.
-K-plus is an extension of the k-means objective that includes one or several terms to quantify between-group similarity with regard to other distribution characteristics instead of only the mean.
+K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to other distribution characteristics instead of only the mean.
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
-The k-plus criterion is not a single criterion for anticlustering, but a familiy of objectives implementing different requirements with regard to between-group similarity.
 After formally introducing the k-plus objective, a simulation study is reported showing that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
-Two practical examples are included that illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
+Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
 
@@
 
-Because the sample variance is basically defined as an average, k-means anticlustering can be employed to equalize the variance across groups.
-That is, we can formulate an adaption of the $\mathit{SSE}$ to quantify the degree to which the variance differs between groups.
-First, however, we generalize the idea to the case of $P$ attributes.
-For each attribute, we have to compute the squared deviation between all values and the mean of the attribute.
+Because the sample variance is basically defined as an average, k-means anticlustering can be employed to equalize the variance across groups, which is the central insight motivating the new k-plus criterion.
+To obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, we have to compute the squared deviation between all individual values and the mean of the attribute.
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$, where
@@
 The same procedure was applied to compute the discrepancy in standard deviations ($\Delta \mathit{SD}$), skewness ($\Delta \mathit{Skew}$) and kurtosis ($\Delta \mathit{Kurtosis}$).
-Quantifying discrepancy in correlations ($\Delta \mathit{Cor}$) followed the same rule, but discrepancies in correlations between groups had to be aggregated across pairs of features instead of single features.
+Quantifying discrepancy in correlations ($\Delta \mathit{Cor}$) followed the same rule, but discrepancies in correlations between groups had to be averaged across pairs of features instead of single features.
 
@@
 Table 1 displays the global simulation results aggregated across all conditions.
-Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but did not outperform a random assignment with regard to all other distribution characteristics; replicating @papenberg2020, k-means' performance with regard to similarity in standard deviations was even below that of a random assignment.
+Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but did not outperform a random assignment with regard to all other distribution characteristics.
+Replicating @papenberg2020, k-means' performance with regard to similarity in standard deviations was even below that of a random assignment.
 K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); all three k-plus approaches outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
 
-As Figure 1 shows, with increasing $N$, k-means and all k-plus methods converge on minimizing $\Delta \mathit{M}$.
+As Figure 1 shows, with increasing $N$, k-means and the k-plus methods converge on the same level with regard to minimizing $\Delta \mathit{M}$.
 This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time---especially when $N$ increases.
@@
 For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering was outperformed.
-As as side note, apart from k-plus correlation, diversity anticlustering was the only method that equalizes correlation structure among groups.
+As as side note, apart from k-plus correlation, diversity anticlustering was the only method that equalized correlation structure among groups.
 To the best of my knowledge, this observation has not been made previously, even though the maximum diversity problem has been widely studied.
@@
 Therefore, a second application illustrates that k-plus anticlustering yields satisfactory results in practice even when group sizes are smaller.
-I make use of a norming data set of 96 word stimuli that was contributed to the `anticlust` package by Dr.
+I make use of a norming data set of 96 word stimuli that was contributed to the `anticlust` package by Marie L.
 Schaper [@schaper2019metacognitive; @schaper2019metamory; also see @papenberg2020 for a description of the norming data].
@@
   tab,
-  caption = "Descriptive statistics by group for the Schaper et al. (2019a; 2019b) data set, after applying default k-plus anticlustering.",
+  caption = "Descriptive statistics by group for the Schaper et al. (2019a; 2019b) data set, after applying k-plus anticlustering.",
   note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
@@
 
-The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion.
-Instead of applying a specialized multicriterion optimization algorithm, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
-The same approach was used independent of the number of objectives that were included as part of the anticlustering process.
-More sophisticated algorithmic approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
+The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion: A simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
+More sophisticated multicriterion approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives (means, variances, higher order moments, and covariances).
 @brusco2019 showed how direct algorithms can be used to optimize multiple criteria at the same time in anticlustering applications.
@@
 This paper's focus on the k-plus objective only instead of optimization algorithms followed two reasonings.
-First, the applications showed that the local maximum search---maximizing the unweighted k-plus sum---yields satisfying results in practice that probably suffice most applications.
+First, the applications showed that the local maximum search---maximizing the unweighted k-plus sum---yields satisfying results in practice that probably suffice in most applications.
 Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it.

### 2022-12-27 paper.Rmd
minor

@@
 Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and separation between clusters [@rokach2005; @steinley2006].
-Anticlustering has many applications in research psychology [@steinley2006; @brusco2019; @papenberg2020]: Examples include splitting tests into parts of equal difficulty [@gierl2017], assigning students to work groups [@baker2002], and assigning stimuli to different, but parallel, experimental conditions [@lahl2006].
+Anticlustering has many applications in research psychology [@steinley2006; @brusco2019; @papenberg2020].
+Examples include splitting tests into parts of equal difficulty [@gierl2017], assigning students to work groups [@baker2002], and assigning stimuli to different, but parallel, experimental conditions [@lahl2006].
 Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual partitioning is usually subpar.
-Fortunately, these problems can be formalized as mathematical optimization problems [e.g., @brusco2019; @spath1986; @baker2002; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
+Fortunately, anticlustering problems can be formalized as mathematical optimization problems [e.g., @brusco2019; @spath1986; @baker2002; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
 
-Anticlustering methods are characterized by (a) an objective function that quantifies between-group similarity and/or within-group heterogeneity and (b) an algorithm that determines how elements are assigned to groups in such a way that the objective function is maximized.
-Several anticlustering objective functions use pairwise dissimilarity ratings such as the Euclidean distance as input.
+Anticlustering methods are characterized by (a) an objective function that quantifies between-group similarity and/or within-group heterogeneity and (b) an algorithm that determines how elements are assigned to groups to maximize the objective function Several anticlustering objective functions use pairwise dissimilarity ratings such as the Euclidean distance as input.
 The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group [@brusco2019; @gallego2013].
 By summing all pairwise dissimilarities of elements within the same group, the diversity reflects the total degree of within-group heterogeneity.
-High within-group diversity simultaneously ensures high similarity of the different groups.[^equalsizedgroups] Another important anticlustering criterion is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
+High within-group diversity simultaneously ensures high between-group similarity.[^equalsizedgroups] Another important anticlustering criterion is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
 Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
-@brusco2019 convincingly argued that anticlustering applications requiring high within-group heterogeneity should incorporate both dispersion and diversity, presenting a bicriterion algorithm to approximate the Pareto efficient set of solutions according to both criteria.
+@brusco2019 convincingly argued that anticlustering applications requiring high within-group heterogeneity should incorporate both dispersion and diversity, and presented a bicriterion algorithm to approximate the Pareto efficient set of solutions according to both criteria.
 
-[^equalsizedgroups]: As will be discussed later, the diversity objective only stricly equalizes within-group heterogeneity and between-group similarity when groups are equal-sized.
+[^equalsizedgroups]: As will be shown later, the diversity objective only strictly equalizes within-group heterogeneity and between-group similarity when groups are equal-sized.
 
-Oftentimes, partitioning applications in psychology are focused on ensuring between-group similarity rather than within-group heterogeneity (even though both goals oftentimes coincide).
-High similarity between-group similarity is for example desireable when designing experimental conditions that have to incorporate different stimulus sets that should however be as similar as possible according to response-relevant attributes [@lahl2006].
-Such stimulus assignment tasks are usually conducted on the basis of attribute data [or *features*; @dry2009] rather than pairwise dissimilarity ratings, for example when selecting stimuli for an experiment on the basis of norming data [e.g., @kurdi2017introducing].
-For word stimuli in psycholinguistic experiments, norming data may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982].
-In other cases, attribute values are binary and may represent the presence or absence of a feature [@tversky1977].
+Oftentimes, partitioning applications in psychology are focused on ensuring between-group similarity rather than within-group heterogeneity (even though both goals usually coincide).
+High similarity between-group similarity is for example desirable when preparing experimental conditions that use different stimulus sets that should however be as similar as possible with regard to response-relevant attributes [@lahl2006].
+Rather than relying on pairwise (dis)similarity ratings, stimulus selection tasks are usually conducted on the basis of attribute data [or *features*; @dry2009].
+For stimuli in psycholinguistic experiments, for example, attributes may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982].
+In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977].
 In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019].
-Subsequently, an anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the stimuli.
+Subsequently, a distance-based anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the stimuli.
 A different anticlusting approach employs clustering objectives that can directly be computed on the attribute values.
@@
 Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
-To obtain a fair assessment In the simulation study, it was therefore important that each anticlustering objective is optimized using the same algorithm, instead of using specialized optimization schemes tailored to each individual criterion.
+To obtain a fair assessment In the simulation study, it was therefore important that each anticlustering objective was optimized using the same algorithm.
+Differences in performance cannot be interpreted if the criterion as well as the algorithm differs between methods.
 It was of particular interest to compare the k-plus objective with the popular diversity objective.

### 2022-12-28 appendix.Rmd
added appendix with figures

#

```{r Figure2, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *M*."}
### For the appendix -- other splits for the main results as plot (N is most important)
# Plot the results, by M
ldf %>% 
  group_by(method, Objective, M) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>% 
  ggplot(aes(x = M, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16) + 
  scale_x_continuous(breaks = 2:5)
```

```{r Figure3, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *SD*."}
# Plot the results, by SD
ldf %>% 
  group_by(method, Objective, SD) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>% 
  ggplot(aes(x = SD, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16) + 
  scale_x_continuous(breaks = 1:3)
```

```{r Figure4, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *K*."}
# Plot the results, by K
ldf %>% 
  group_by(method, Objective, K) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>% 
  ggplot(aes(x = K, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16) + 
  scale_x_continuous(breaks = 2:4)
```

```{r Figure5, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *r*."}
# Plot the results, by r
ldf %>% 
  group_by(method, Objective, r) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>% 
  ggplot(aes(x = r, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16) + 
  scale_x_continuous(breaks = 0:5 / 10)
```

### 2022-12-28 paper.Rmd
editing

@@
 Introducing k-plus, this paper presents an extension of the k-means clustering objective to maximize between-group similarity in anticlustering applications.
-The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, higher order moments, and covariances), thereby extending the classical k-means criterion, which only reflects how groups differ in their means.
+The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, higher order moments, and covariances), while the classical k-means criterion only reflects the degree to which groups differ in their means.
 It is shown that the k-plus criterion can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
@@
 documentclass     : "apa6"
-classoption       : "doc"
 output            : 
   papaja::apa6_pdf
+appendix: "appendix.Rmd"
+
+header-includes:
+  - \usepackage{setspace}
 ---
@@
 Anticlustering has many applications in research psychology [@steinley2006; @brusco2019; @papenberg2020].
-Examples include splitting tests into parts of equal difficulty [@gierl2017], assigning students to work groups [@baker2002], and assigning stimuli to different, but parallel, experimental conditions [@lahl2006].
+Examples include splitting tests into parts of equal difficulty [@gierl2017], assigning students to work groups [@baker2002], and assigning stimuli to different, but parallel experimental conditions [@lahl2006].
 Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual partitioning is usually subpar.
-Fortunately, anticlustering problems can be formalized as mathematical optimization problems [e.g., @brusco2019; @spath1986; @baker2002; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
+Fortunately, anticlustering problems can be formalized as mathematical optimization problems [e.g., @baker2002; @brusco2019; @spath1986; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
 
-Anticlustering methods are characterized by (a) an objective function that quantifies between-group similarity and/or within-group heterogeneity and (b) an algorithm that determines how elements are assigned to groups to maximize the objective function Several anticlustering objective functions use pairwise dissimilarity ratings such as the Euclidean distance as input.
-The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between any elements within the same group [@brusco2019; @gallego2013].
-By summing all pairwise dissimilarities of elements within the same group, the diversity reflects the total degree of within-group heterogeneity.
-High within-group diversity simultaneously ensures high between-group similarity.[^equalsizedgroups] Another important anticlustering criterion is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
+Anticlustering methods are characterized by (a) an objective function that quantifies between-group similarity and/or within-group heterogeneity, and (b) an algorithm that determines how elements are assigned to groups to maximize the objective function.
+Several anticlustering objective functions use pairwise dissimilarity ratings such as the Euclidean distance as input.
+The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between elements within the same group [@brusco2019; @gallego2013].
+By summing all pairwise dissimilarities within the same group, the diversity reflects the total degree of within-group heterogeneity.
+For anticlustering, high within-group heterogeneity is desired, which simultaneously ensures high between-group similarity.[^equalsizedgroups] Another important anticlustering criterion is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
 Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
-@brusco2019 convincingly argued that anticlustering applications requiring high within-group heterogeneity should incorporate both dispersion and diversity, and presented a bicriterion algorithm to approximate the Pareto efficient set of solutions according to both criteria.
+@brusco2019 convincingly argued that anticlustering applications striving for within-group heterogeneity should incorporate both dispersion and diversity, and they presented a bicriterion algorithm to approximate the Pareto efficient set of solutions according to both criteria.
 
@@
 
-Oftentimes, partitioning applications in psychology are focused on ensuring between-group similarity rather than within-group heterogeneity (even though both goals usually coincide).
-High similarity between-group similarity is for example desirable when preparing experimental conditions that use different stimulus sets that should however be as similar as possible with regard to response-relevant attributes [@lahl2006].
-Rather than relying on pairwise (dis)similarity ratings, stimulus selection tasks are usually conducted on the basis of attribute data [or *features*; @dry2009].
+Oftentimes, partitioning applications in psychology are focused on ensuring between-group similarity rather than within-group heterogeneity (even though both goals oftentimes coincide).
+High between-group similarity is for example desirable when preparing experimental conditions that use different stimulus sets, that should however be as similar as possible with regard to response-relevant attributes [@lahl2006].
+Rather than relying on pairwise (dis)similarity ratings, stimulus selection tasks are usually conducted on the basis of attribute data.
+That is, individual stimuli are numerically coded on relevant dimensions [@dry2009].
 For stimuli in psycholinguistic experiments, for example, attributes may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982].
@@
 Subsequently, a distance-based anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the stimuli.
-A different anticlusting approach employs clustering objectives that can directly be computed on the attribute values.
+A different anticlusting approach relies on clustering objectives that can directly be computed using the attribute values.
 K-means is probably the best-known objective of this kind.
-In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized [@brusco2006branch], usually using the k-means heuristic [@jain2010; @steinley2006].
+In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized, usually using the k-means heuristic [@jain2010; @steinley2006; @brusco2006branch].
 Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np].
@@
 Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters.
-However, other parameters---such as the variance---are not targeted.
-@papenberg2020 showed that k-means anticlustering tends to over-optimize similarity with regard to the mean; even an entirely random group assignment usually yielded more similar variances.
-When requiring overall between-group similarity, the negligence of other distribution characteristics than the mean is misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
+However, other distribution characteristics---such as the variance---are not targeted.
+@papenberg2020 showed that k-means anticlustering tends to over-optimize similarity with regard to the mean; even an entirely random group assignment usually resulted in more similar variances.
+When requiring overall between-group similarity, the negligence of all distribution characteristics other than the mean is misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
 
 To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering.
-K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to other distribution characteristics instead of only the mean.
+K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to several distribution characteristics instead of only the mean.
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
@@
 For the purpose of problem formalization, I adopt the notation @steinley2006 provided in his comprehensive synthesis of half a century of research on k-means.
-K-means clustering (and anticlustering) is used to partition two-way, two-mode data into groups.
-That is, $N$ elements each having $P$ attributes are assigned to $K$ groups $(C_1., \ldots, C_K)$, $C_k$ being the set of $n_k$ objects in group $k$.
+K-means anticlustering is used to partition $N$ elements each having $P$ attributes into $K$ groups $(C_1, \ldots, C_K)$, where $C_k$ is the set of $n_k$ objects in the $k$'th group.
 The partitioning is subject to the requirement that each element is assigned to exactly one group, formalized by the following two constraints [@papenberg2020]:
@@
 
-The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all observations belonging the $k$th cluster $C_k$ ($k = 1, \ldots, K$):
+The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all observations belonging the $k$'th cluster $C_k$ ($k = 1, \ldots, K$):
 
@@
 To motivate the new criterion, we first consider a set of one-dimensional observations $(x_1, \ldots, x_N)$.
-The extension to the general case of $P$ attributes will be straight forward.
-
 The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
@@
 Because the sample variance is basically defined as an average, k-means anticlustering can be employed to equalize the variance across groups, which is the central insight motivating the new k-plus criterion.
-To obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, we have to compute the squared deviation between all individual values and the mean of the attribute.
+Therefore, to obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, we have to compute the squared deviation between all individual values and the mean of an attribute.
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$, where
@@
 
-It is unlikely that an exclusive optimization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is of interest without also considering the $\mathit{SSE}$ (i.e., equalizing means).
+It is unlikely that an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is of interest without also considering the $\mathit{SSE}$ (i.e., equalizing means).
 A combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall between-group similarity.
 The most simple approach for such a bicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
-That is, both criteria are computed independently from each other and then combined into a single criterion through a weighted sum:
+That is, both criteria are computed independently from each other and are then combined into a single criterion through a weighted sum:
 
@@
 Thus, when optimizing a k-plus criterion, the original data matrix was extended to incorporate the additional criteria.
-Subsequently, the original k-means criterion $SSE$ was maximized on the basis of the augmented data set.
+Subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
 
@@
 
-The simulation was implemented using the statistical programming language R [Version 4.2; @R-base].
-The R package `anticlust` [Version 0.6.1; @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
-The R package `faux` [Version 1.1.0; @R-faux] was used to generate the multivariate normal data sets that were processed during the simulation.
-The R packages `dplyr` [Version 1.0.7; @R-dplyr], `ggplot2` [Version 3.3.5; @R-ggplot2], `papaja` [Version 0.1.0.9997; @R-papaja], `tidyr` [Version 1.1.3; @R-tidyr] and `DescTools` [Version 0.99.45; @R-desctools] were used for data analysis and result presentation.
+The simulation was implemented using the statistical programming language R [Version 4.2, @R-base].
+The R package `anticlust` [Version 0.6.1, @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
+The R package `faux` [Version 1.1.0, @R-faux] was used to generate the multivariate normal data sets that were processed during the simulation.
+The R packages `dplyr` [Version 1.0.7, @R-dplyr], `ggplot2` [Version 3.3.5, @R-ggplot2], `papaja` [Version 0.1.0.9997, @R-papaja], `tidyr` [Version 1.1.3, @R-tidyr] and `DescTools` [Version 0.99.45, @R-desctools] were used for data analysis and result presentation.
 The simulation study is fully reproducible via code and data that is accessible from the accompanying OSF repository (**link TODO**).
@@
 
-The same local maximum search [Method LCW in @weitz1998] was applied to optimize each of the five competing anticlustering objectives.
+The same local maximum search [Method LCW, @weitz1998] was applied to optimize each of the five competing anticlustering objectives.
 Building on an initial random assignment of elements to groups, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
-That is, for each element, all possible exchanges are simulated and the resulting objective values are stored.
-After each exchange has been tested, the one exchange is realized that improves the objective the most.
+That is, for each element, all possible exchanges with elements in other clusters are simulated and the resulting objective values are stored.
+After each exchange has been evaluated, the one exchange is realized that improves the objective the most.
 No exchange is realized if no improvement is possible.
@@
 To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
-The simulation employed five repetitions of the local maximum search.
+The simulation study always employed five repetitions of the local maximum search.
 
@@
 The same procedure was applied to compute the discrepancy in standard deviations ($\Delta \mathit{SD}$), skewness ($\Delta \mathit{Skew}$) and kurtosis ($\Delta \mathit{Kurtosis}$).
-Quantifying discrepancy in correlations ($\Delta \mathit{Cor}$) followed the same rule, but discrepancies in correlations between groups had to be averaged across pairs of features instead of single features.
+Quantifying discrepancy in correlations ($\Delta \mathit{Cor}$) followed the same rule, but discrepancies in correlations had to be averaged across pairs of features instead of single features.
 
 ## Results
+
+\begin{singlespace}
 
@@
 library(tidyr)
-source("./Simulation_Study/0-functions-generate-data.R")
 
@@
 
-apa_table(
+foo <- apa_table(
   tab,
@@
 ```
+
+\end{singlespace}
 
@@
 
-As Figure 1 shows, with increasing $N$, k-means and the k-plus methods converge on the same level with regard to minimizing $\Delta \mathit{M}$.
+Figure 1 shows that with increasing $N$, k-means and the k-plus methods converged on the same level with regard to minimizing $\Delta \mathit{M}$.
 This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time---especially when $N$ increases.
 However, while k-plus anticlustering is capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria apparently aggravates fulfilling each single criterion.
-K-plus skew/kurtosis and k-plus correlation performed worse at minimizing discrepancy with regard to standard deviations than standard k-plus, which itself performed slightly worse than k-means anticlustering.
+Hence, k-plus skew/kurtosis and k-plus correlation performed worse at minimizing $\Delta \mathit{SD}$ than standard k-plus.
 Still, every k-plus method did what it was supposed to do: k-plus skew/kurtosis was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation was best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$.
@@
 These results illustrate the cost of optimizing several criteria at the same time: for low $N$, k-plus seems skew/kurtosis to be preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
-Increasing $N$ then facilitates to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
+Increasing $N$ then facilitated to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
 
-Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to somewhat addressed all distribution characteristics.
+Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to address all distribution characteristics.
 For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering was outperformed.
@@
 
+The appendix includes additional figures that split the simulation results according to the other variables that varied between data sets (number of features; correlation between features; standard deviation of features; number of groups).
+Here, the only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased.
+Because the number of covariances increases quadratically when the number of features increases, it gets comparatively more difficult to fulfill the correlation objective with a higher number of features.
+
 # K-plus anticlustering using the R package anticlust
 
-This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020], using an interface function called `kplus_anticlustering()`, which is available from version 0.6.3 onward.
+This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] using an interface function called `kplus_anticlustering()`, which is available from version 0.6.3 onward.
 The package can be installed in the R environment using the `install.packages()` command:
@@
 It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations, and the skewness was also very evenly matched.
-K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations and especially skewness.
+K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations skewness in particular.
 Diversity anticlustering was well-suited to match means, standard deviations and skew, but was slightly outperformed by k-plus.
@@
 Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups.
-This is due the diversity criterion being a measure of within-group heterogeneity.
-Because the diversity criterion is a sum rather than an average, the correspondence between within-group heterogeneity and between-group similarity does not hold for unequal-sized groups.
 K-plus anticlustering strives for between-group similarity regardless of group size.
-
 Unequal group sizes can be requested in `anticlust` by passing the different group sizes to the argument `K`.
@@
 
-This function call relied on the default k-plus settings ensuring similarity with regard to means and variances equalized---but with regard to higher order moments or covariances.
+This function call relied on the default k-plus setting that ensures similarity with regard to means and variances, but does not consider higher order moments or covariances.
 
@@
 Anticlustering is a partitioning method that aims for high between-group similarity and within-group heterogeneity.
-In this paper, I presented the k-plus criterion, which focuses on maximizing between-group similarity.
+In this paper, I presented the k-plus criterion for anticlustering, which focuses on maximizing between-group similarity.
 K-plus represents between-group similarity as discrepancies in distribution moments between groups: mean, variances, skewness, kurtosis, higher order moments, and even covariances.
 Depending on the application and the needs of a user, a different subset of these objectives can be included in the anticlustering process.
-The k-plus criterion is an extension of the classical k-means criterion, which only represents how similar groups are with regard to their means.
-Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by merely augmenting the input data.
+The k-plus criterion is an extension of the classical k-means criterion, which only reflects how similar groups are with regard to their means.
+Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
 
 A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.
-Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be a default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be aggravated in small data sets.
+Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be considered as the default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be aggravated in small data sets.
 In practice, users can try out different objectives and see which specification fulfils their needs.
@@
 by specifying that multiple distribution moments should equalized.
-If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variances).
+If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variance).
 Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.
@@
 
-The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion: A simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
-More sophisticated multicriterion approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives (means, variances, higher order moments, and covariances).
-@brusco2019 showed how direct algorithms can be used to optimize multiple criteria at the same time in anticlustering applications.
+The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion: A simple unweighted sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
+More sophisticated multicriterion approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
+@brusco2019 showed how a direct algorithm can be used to optimize multiple criteria at the same time in anticlustering applications.
 Their study focused on approximating the pareto set of efficient solutions that optimize diversity and dispersion, which is essential in applications that require high within-group heterogeneity.
-Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.
+Future research should investigate if k-plus anticlustering may profit from similar advanced algorithmic treatment.
 
@@
 Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
-To obtain a fair assessment In the simulation study, it was therefore important that each anticlustering objective was optimized using the same algorithm.
-Differences in performance cannot be interpreted if the criterion as well as the algorithm differs between methods.
+To obtain a fair assessment in the simulation study, it was therefore important that each anticlustering objective was optimized using the same algorithm.
+Differences in performance cannot be interpreted if the objective function as well as the algorithm used to optimize it differs between methods.
 It was of particular interest to compare the k-plus objective with the popular diversity objective.
@@
 \endgroup
+
+\newpage
+
+# (APPENDIX) Appendix {-}
+
+```{r child = "appendix.Rmd"}
+```

### 2022-12-28 paper.Rmd
minor

@@
 That is, individual stimuli are numerically coded on relevant dimensions [@dry2009].
-For stimuli in psycholinguistic experiments, for example, attributes may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982].
+For stimuli in psycholinguistic experiments, attributes may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982].
 In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977].
@@
 # Plot the results, by N
+axis_ticks <- seq(24, 300, by = 12)
+axis_labels <- axis_ticks
+axis_labels[c(FALSE, TRUE)] <- ""
+
 ldf %>% 
@@
   ylab("Mean discrepancy") +
-  theme_bw(base_size = 16)
+  theme_bw(base_size = 16) + 
+  scale_x_continuous(
+    breaks = axis_ticks, 
+    labels = axis_labels
+  )
 

### 2022-12-29 paper.Rmd
editing and moved partition example back into paper

@@
 Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
-Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and separation between clusters [@rokach2005; @steinley2006].
+Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and separation between clusters [@rokach2005].
 Anticlustering has many applications in research psychology [@steinley2006; @brusco2019; @papenberg2020].
@@
 The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between elements within the same group [@brusco2019; @gallego2013].
-By summing all pairwise dissimilarities within the same group, the diversity reflects the total degree of within-group heterogeneity.
-For anticlustering, high within-group heterogeneity is desired, which simultaneously ensures high between-group similarity.[^equalsizedgroups] Another important anticlustering criterion is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
-Maximizing the dispersion increases the within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
+By considering pairwise dissimilarities from between elements in the same group, the diversity reflects the total degree of within-group heterogeneity.
+High within-group diversity simultaneously ensures high between-group similarity.[^equalsizedgroups] Another important anticlustering criterion is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
+Maximizing the dispersion increases within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
 @brusco2019 convincingly argued that anticlustering applications striving for within-group heterogeneity should incorporate both dispersion and diversity, and they presented a bicriterion algorithm to approximate the Pareto efficient set of solutions according to both criteria.
@@
 
-Oftentimes, partitioning applications in psychology are focused on ensuring between-group similarity rather than within-group heterogeneity (even though both goals oftentimes coincide).
+Oftentimes, partitioning applications in psychology are focused on between-group similarity rather than within-group heterogeneity (even though both goals oftentimes coincide).
 High between-group similarity is for example desirable when preparing experimental conditions that use different stimulus sets, that should however be as similar as possible with regard to response-relevant attributes [@lahl2006].
@@
 Subsequently, a distance-based anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the stimuli.
-A different anticlusting approach relies on clustering objectives that can directly be computed using the attribute values.
+A different anticlusting approach relies on clustering objectives that can directly be computed on the attribute values.
 K-means is probably the best-known objective of this kind.
@@
 However, other distribution characteristics---such as the variance---are not targeted.
-@papenberg2020 showed that k-means anticlustering tends to over-optimize similarity with regard to the mean; even an entirely random group assignment usually resulted in more similar variances.
-When requiring overall between-group similarity, the negligence of all distribution characteristics other than the mean is misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
+@papenberg2020 showed that k-means anticlustering tends to over-optimize between-group similarity with regard to the mean at the cost of similarity in spread; even an entirely random group assignment usually resulted in more similar variances.
+When requiring overall between-group similarity, the negligence of distribution characteristics other than the mean is misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
 
@@
 K-means anticlustering is used to partition $N$ elements each having $P$ attributes into $K$ groups $(C_1, \ldots, C_K)$, where $C_k$ is the set of $n_k$ objects in the $k$'th group.
-The partitioning is subject to the requirement that each element is assigned to exactly one group, formalized by the following two constraints [@papenberg2020]:
-
-$$
-\bigcup\limits_{k = 1}^{K} C_j = X
-$$
-
-$$
-C_j \cap C_k = \emptyset, \; \forall j, k \in \{1, \ldots, K\}, \; j \ne k
-$$
-
-In anticlustering applications, an additional constraint is usually imposed on the group sizes $n_k$, where the most common restriction requires that all groups have equal size:
-
-$$
-n_k = \frac{N}{K}, \forall k \in \{1, \ldots, K\}
-$$
+The partitioning is subject to the requirement that each element is assigned to exactly one group.
+In anticlustering applications, an additional constraint is usually imposed on the group sizes $n_k$, where the most common restriction requires that all groups have equal size [@papenberg2020].
 
@@
 In the following, I present a variation of the $\mathit{SSE}$ that quantifies how the variance of numeric data differs between groups.
-To motivate the new criterion, we first consider a set of one-dimensional observations $(x_1, \ldots, x_N)$.
+To motivate the new criterion, we first consider a set of one-dimensional observations.
 The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
@@
 However, the input data reflects the squared difference between each data point and the attribute mean instead of the raw data points.
+Maximizing $\mathit{SSE_{Var}}$ will lead to similar variances across the $K$ groups in each of the $P$ attributes.
 
@@
 
-It is unlikely that an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is of interest without also considering the $\mathit{SSE}$ (i.e., equalizing means).
+It is unlikely that an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is of interest without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
 A combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall between-group similarity.
@@
 via $z$-standardization.
+
+```{r}
+
+# First rendering is slow because the data needs to be generated
+path <- "partition_objectives.csv"
+
+if (!file.exists(path)) {
+  set.seed(2207) # 2207
+  # Generate data and all partitions
+  N <- 14
+  K <- 2
+  partitions <- generate_partitions(N, K)
+  features <- rnorm(N, 15, 3)
+
+  # Compute objectives for each partition
+  # For each partition: Compute SSE objective, and SSE(VAR)
+  # 1. SSE
+  all_objectives_kMeans <- sapply(
+    partitions,
+    FUN = var_objective,
+    features = features
+  )
+
+  squared_distances_from_mean <- squared_from_mean(as.matrix(features))
+
+  # 2. SSE(VAR)
+  all_objectives_kVar <- sapply(
+    partitions,
+    FUN = var_objective,
+    features = squared_distances_from_mean
+  )
+
+  # 3. SSE(KPLUS)
+  # Use standardization to compute k-plus criterion!
+  all_objectives_kPlus <- sapply(
+    partitions,
+    FUN = var_objective,
+    features = scale(cbind(features, squared_distances_from_mean))
+  )
+
+  df <- data.frame(
+    kmeans = all_objectives_kMeans,
+    kvar = all_objectives_kVar,
+    kplus = all_objectives_kPlus
+  )
+
+  write.table(
+    df, path, row.names = FALSE, sep = ","
+  )
+} else {
+  df <- read.csv(path)
+}
+
+```
+
+When optimizing a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are close to the global optimum value.
+Generally, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree [@ferligoj1992direct].
+Usually, a tradeoff is necessary.
+However, a motivating example illustrates that aiming to simultaneously optimize similarity of means and variances is not only desirable, but may actually be feasible.
+I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
+Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
+Note that for both criteria, a larger value is better, i.e., indicates higher similarity with regard to means and variances, respectively.
+Therefore, to obtain overall between-group similarity, a partition should be chosen that is depicted in the upper right corner of the scatterplot.
+However, optimal k-means anticlustering would create two groups that are not particularly similar with regard to their variance.
+The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$.
+Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (the triangular symbol in Figure 1) is not well-suited to optimize $\mathit{SSE}$ at the same time.
+The diamond symbol represents the partition that maximizes (the unweighted) $SSE_{kplus}$, i.e., the sum of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$.
+This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
+
+```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
+The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$ and represents a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+
+# Plot SSE vs SSE(VAR)
+plot(
+  df[, c("kmeans", "kvar")], 
+  pch = 4, col = "darkgrey", 
+  las = 1, cex = .5,
+  xlab = "SSE",
+  ylab = "SSE_Var"
+)
+
+# Illustrate best partition wrt difference in means (i.e., SSE)
+points(
+  df$kmeans[which.max(df$kmeans)],
+  df$kvar[which.max(df$kmeans)],
+  cex = 1.2, pch = 22, bg = "white", lwd = 1.7
+)
+
+# Illustrate best partition wrt difference in variance (i.e., SSE(VAR))
+points(
+  df$kmeans[which.max(df$kvar)],
+  df$kvar[which.max(df$kvar)],
+  cex = 1.2, pch = 24, bg = "white", lwd = 1.7
+)
+
+# Illustrate best partition wrt k-plus criterion
+points(
+  df$kmeans[which.max(df$kplus)],
+  df$kvar[which.max(df$kplus)],
+  cex = 1.2, pch = 23, bg = "white", lwd = 1.7
+)
+```
 
@@
 
-Figure 1 shows that with increasing $N$, k-means and the k-plus methods converged on the same level with regard to minimizing $\Delta \mathit{M}$.
+Figure 2 shows that with increasing $N$, k-means and the k-plus methods converged on the same level with regard to minimizing $\Delta \mathit{M}$.
 This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time---especially when $N$ increases.
@@
 It is maybe surprising that k-plus skew/kurtosis does not seem to have performed much better than diversity anticlustering with regard to skewness and---in particular---kurtosis, even though the method has been specifically tailored to these criteria.
-Figure 1 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis was even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
+Figure 2 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis was even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
 However, with increasing $N$, it clearly outperformed diversity anticlustering.

### 2022-12-29 paper.Rmd
editing

@@
 
-When optimizing a bicriterion objective such as $SSE_{kplus}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are close to the global optimum value.
-Generally, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree [@ferligoj1992direct].
-Usually, a tradeoff is necessary.
-However, a motivating example illustrates that aiming to simultaneously optimize similarity of means and variances is not only desirable, but may actually be feasible.
+When optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are close to the global optimum value.
+However, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree, making a tradeoff necessary [@ferligoj1992direct].
+A motivating example illustrates that in the case of anticlustering, aiming to simultaneously optimize similarity of means and variances may however be feasible.
 I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
 Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
-Note that for both criteria, a larger value is better, i.e., indicates higher similarity with regard to means and variances, respectively.
-Therefore, to obtain overall between-group similarity, a partition should be chosen that is depicted in the upper right corner of the scatterplot.
-However, optimal k-means anticlustering would create two groups that are not particularly similar with regard to their variance.
+To divide the 14 elements into two groups that are similar with regard to their mean and variance, a partition should be chosen that is depicted in the upper right corner of the scatterplot.
 The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$.
 Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (the triangular symbol in Figure 1) is not well-suited to optimize $\mathit{SSE}$ at the same time.
-The diamond symbol represents the partition that maximizes (the unweighted) $SSE_{kplus}$, i.e., the sum of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$.
+The diamond symbol represents the partition that maximizes $\mathit{SSE_{kplus}}$, i.e., the sum of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$.
 This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
+Thus, the optimal partition according to $\mathit{SSE_{kplus}}$ fulfills the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ to an astonishing degree.
 
 ```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
-The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$ and represents a good tradeoff, satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$, and satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
 
@@
   las = 1, cex = .5,
-  xlab = "SSE",
-  ylab = "SSE_Var"
+  xlab = "",
+  ylab = ""
 )
+mtext(side = 2, text = expression(italic(SSE[Var])), line = 3)
+mtext(side = 1, text = expression(italic(SSE)), line = 3)
 
@@
 
-The outlined logic to equalize the variance, which is the second moment of a distribution, can be generalized to higher order moments in a straight forward manner.
+The k-plus approach to equalizing the variance between groups can be generalized to higher order moments in a straight forward manner.
 The *j*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ can be computed as
@@
 The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion: A simple unweighted sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
-More sophisticated multicriterion approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
-@brusco2019 showed how a direct algorithm can be used to optimize multiple criteria at the same time in anticlustering applications.
-Their study focused on approximating the pareto set of efficient solutions that optimize diversity and dispersion, which is essential in applications that require high within-group heterogeneity.
-Future research should investigate if k-plus anticlustering may profit from similar advanced algorithmic treatment.
-
-This paper's focus on the k-plus objective only instead of optimization algorithms followed two reasonings.
-First, the applications showed that the local maximum search---maximizing the unweighted k-plus sum---yields satisfying results in practice that probably suffice in most applications.
+The sole focus on the objective function followed two reasonings.
+First, the applications showed that the local maximum search, maximizing the unweighted k-plus sum, yielded satisfying results that probably suffice in most applications.
+It seems that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria at the same time (see Figure 1).
+Still, more sophisticated approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
+@brusco2019 gave an example of how direct algorithms can be employed to optimize multiple criteria at the same time in anticlustering applications aiming for high within-group heterogeneity.
+Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.
 Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it.

### 2022-12-29 paper.Rmd
minor

@@
 
-When optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are close to the global optimum value.
+When optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups.
 However, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree, making a tradeoff necessary [@ferligoj1992direct].
@@
 ```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
-The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$, and satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+The partition represented by the triangular minimizes the difference in variances; the partition represented by the square minimizes the difference in means; the partition represented by the diamond maximizes $SSE_{kplus}$ and satisfys both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
 
@@
 
-The appendix includes additional figures that split the simulation results according to the other variables that varied between data sets (number of features; correlation between features; standard deviation of features; number of groups).
+The Appendix includes additional figures that split the simulation results according to the other variables that varied between simulation runs (number of features; correlation between features; standard deviation of features; number of groups).
 Here, the only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased.
@@
 
-This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] using an interface function called `kplus_anticlustering()`, which is available from version 0.6.3 onward.
+This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] using the function `kplus_anticlustering()`, which is available from version 0.6.3 onward.
 The package can be installed in the R environment using the `install.packages()` command:
@@
 
-This function call relied on the default k-plus setting that ensures similarity with regard to means and variances, but does not consider higher order moments or covariances.
-
 ## Example Application II: Small data set ($N = 96$)
@@
 Table 4 illustrates the results that are obtained when using standard k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
-Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are still similar enough for any practical purpose---especially when considering inevitable measurement error in norming studies.
+Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are similar enough for any practical purposes.
 
@@
 In this paper, I presented the k-plus criterion for anticlustering, which focuses on maximizing between-group similarity.
-K-plus represents between-group similarity as discrepancies in distribution moments between groups: mean, variances, skewness, kurtosis, higher order moments, and even covariances.
-Depending on the application and the needs of a user, a different subset of these objectives can be included in the anticlustering process.
+K-plus represents between-group similarity as discrepancies in distribution moments between groups: mean, variances, skewness, kurtosis, higher order moments, and covariances.
+Depending on the application, a different subset of these objectives can be included in the anticlustering process.
 The k-plus criterion is an extension of the classical k-means criterion, which only reflects how similar groups are with regard to their means.
-Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
+Interestingly, however, the k-plus extension does not change the original k-means criterion because additional criteria can be incorporated by augmenting the input data.
 

### 2023-01-03 paper.Rmd
editing

@@
 A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple objectives.
-In particular, optimizing between-group similarity with regard to variances as well as means comes with almost no cost to either criterion; the k-plus extension is therefore generally preferred over classical k-means anticlustering.
-Higher order moments and covariance structure can also succesfully be equalized.
+In particular, optimizing between-group similarity with regard to variances usually does not compromise similarity with regard to means; the k-plus extension is therefore generally preferred over classical k-means anticlustering.
 Example are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
@@
 
-Oftentimes, partitioning applications in psychology are focused on between-group similarity rather than within-group heterogeneity (even though both goals oftentimes coincide).
-High between-group similarity is for example desirable when preparing experimental conditions that use different stimulus sets, that should however be as similar as possible with regard to response-relevant attributes [@lahl2006].
-Rather than relying on pairwise (dis)similarity ratings, stimulus selection tasks are usually conducted on the basis of attribute data.
+Oftentimes, partitioning applications in psychology are focused on between-group similarity rather than within-group heterogeneity, even though both goals oftentimes coincide.
+High between-group similarity is for example desirable when designing experimental conditions using different stimulus sets that should be as similar as possible with regard to response-relevant attributes [@lahl2006].
+Stimulus selection tasks are usually conducted on the basis of attribute data and not on pairwise (dis)similarity ratings.
 That is, individual stimuli are numerically coded on relevant dimensions [@dry2009].
@@
 However, other distribution characteristics---such as the variance---are not targeted.
-@papenberg2020 showed that k-means anticlustering tends to over-optimize between-group similarity with regard to the mean at the cost of similarity in spread; even an entirely random group assignment usually resulted in more similar variances.
-When requiring overall between-group similarity, the negligence of distribution characteristics other than the mean is misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
+@papenberg2020 showed that k-means anticlustering tends to over-optimize between-group similarity with regard to the mean at the cost of similarity in variance.
+However, when aiming for overall between-group similarity, the negligence of distribution characteristics other than the mean is misguided; similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
 
@@
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
-After formally introducing the k-plus objective, a simulation study is reported showing that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
+After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
 Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
@@
 For the purpose of problem formalization, I adopt the notation @steinley2006 provided in his comprehensive synthesis of half a century of research on k-means.
-K-means anticlustering is used to partition $N$ elements each having $P$ attributes into $K$ groups $(C_1, \ldots, C_K)$, where $C_k$ is the set of $n_k$ objects in the $k$'th group.
+K-means anticlustering is used to partition $N$ elements each having $P$ attributes into $K$ groups $(C_1, \ldots, C_K)$.
 The partitioning is subject to the requirement that each element is assigned to exactly one group.
-In anticlustering applications, an additional constraint is usually imposed on the group sizes $n_k$, where the most common restriction requires that all groups have equal size [@papenberg2020].
+In anticlustering applications, an additional constraint is usually imposed on the group sizes, with the most common restriction being that all groups have equal size [@papenberg2020].
 
@@
 To motivate the new criterion, we first consider a set of one-dimensional observations.
-The sample variance of a set of observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
+The sample variance of a set of $N$ observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
 
@@
 
-This formula again has the same form as $\mathit{SSE}$ (and $\mathit{SSE_{Var}}$), but is computed on different data.
+This formula again has the same form as $\mathit{SSE}$ but is computed on different data.
 In this case, both the original data $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function.
@@
 Hence, using the unweighted sum approach, k-plus anticlustering reduces to an augmentation of the data matrix, leaving the original k-means criterion unchanged.
-
-Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective, i.e., when just appending the squared-distance-features to the data set.
-Since the computation of the data matrix $\mathbf{Y}_{N \times P}$ involves squaring, the squared distances $y_{ij}$ usually have much larger values, on average, than the raw values $x_{ij}$.
-Therefore, the $\mathit{SSE_{Var}}$ criterion naturally receives greater weight when computing $\mathit{SSE_{kplus}}$ than the standard $\mathit{SSE}$ criterion.
-This imbalance is most likely unwarranted because optimizing similarity with regard to variances is not more important than optimizing similarity with regard to means.
-A solution is to standardize all values $x_{ij}$ and $y_{ij}$ to the same scale before computing $SSE_{kplus}$, e.g.
+Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective.
+Since the computation of the $y_{ij}$ involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$.
+This imbalance can strongly and unpredictably influence the relative weight of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ when computing $\mathit{SSE_{kplus}}$.
+A solution is to standardize all values $x_{ij}$ and $y_{ij}$ to the same scale before computing $\mathit{SSE_{kplus}}$, e.g.
 via $z$-standardization.
@@
 
-When optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups.
-However, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree, making a tradeoff necessary [@ferligoj1992direct].
-A motivating example illustrates that in the case of anticlustering, aiming to simultaneously optimize similarity of means and variances may however be feasible.
+When optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$, the ideal outcome would be to find a partitioning where both constituting criteria (mean and variance) are similar between groups.
+Although research on multi-objective optimization has shown that solutions oftentimes fail to simultaneously satisfy several criteria [@ferligoj1992direct], a motivating example illustrates that in the case of anticlustering, aiming to simultaneously optimize similarity of means and variances may be feasible.
 I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
 Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
-To divide the 14 elements into two groups that are similar with regard to their mean and variance, a partition should be chosen that is depicted in the upper right corner of the scatterplot.
 The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$.
 Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (the triangular symbol in Figure 1) is not well-suited to optimize $\mathit{SSE}$ at the same time.
-The diamond symbol represents the partition that maximizes $\mathit{SSE_{kplus}}$, i.e., the sum of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$.
-This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
+Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances.
+The diamond symbol represents the partition that maximizes $\mathit{SSE_{kplus}}$, which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
 Thus, the optimal partition according to $\mathit{SSE_{kplus}}$ fulfills the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ to an astonishing degree.
@@
 
-The k-plus approach to equalizing the variance between groups can be generalized to higher order moments in a straight forward manner.
-The *j*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ can be computed as
+The k-plus approach to equalizing the variance can be generalized to higher order moments. The *j*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ can be computed as
 
@@
 
-Because the covariance is also defined as an expected value, the k-plus criterion can be extended to reflect differences in covariances between data sets.
-The covariance between two variables $x$ and $y$ is defined as
+Because the covariance is also defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariance structure.
+The covariance between two variables $x = (x_1, \ldots, x_N)$ and $y = (y_1, \ldots, y_N)$ is defined as
 
@@
 
-Thus, to maximize between-group similarity with regard to covariance structure, it is possible to add an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
-This requires computing $\binom{N}{2}$ new variables and thus turns out be be rather costly: The number of additional variables needed to minimize differences with regard to covariance structure grows quadratically with increasing number of variables; the number of additional variables needed to minimize differences with regard to distribution moments only grows linearly with increasing number of variables.
+Thus, to maximize between-group similarity with regard to covariance structure, it is possible to augment the data set with an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
 
@@
 To thoroughly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.
-To implement all k-plus objectives, the simple unweighted sum outlined above approach was employed.
-Thus, when optimizing a k-plus criterion, the original data matrix was extended to incorporate the additional criteria.
-Subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
+To implement all k-plus objectives, the unweighted sum approach was employed.
+Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria; subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
 
@@
 The same local maximum search [Method LCW, @weitz1998] was applied to optimize each of the five competing anticlustering objectives.
-Building on an initial random assignment of elements to groups, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
-That is, for each element, all possible exchanges with elements in other clusters are simulated and the resulting objective values are stored.
-After each exchange has been evaluated, the one exchange is realized that improves the objective the most.
+Building on an initial random assignment, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
+That is, it starts with the first element and simulates all exchanges with elements that are currently assigned to a different group.
+After each exchange has been processed, the one exchange is realized that improves the objective function the most.
 No exchange is realized if no improvement is possible.
-The exchange procedure is repeated for each element.
-Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements, repeating the exchange procedure.
-The procedure stops as soon as an iteration through the entire data set no longer yielded an improvement, i.e., as soon as a local maximum has been found.
+The exchange process is repeated for each element.
+After that, the procedure restarts at the beginning and is repeated until an iteration through the entire data set no longer yields an improvement.
 To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
@@
 
-foo <- apa_table(
+apa_table(
   tab,
@@
 This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time---especially when $N$ increases.
-However, while k-plus anticlustering is capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria apparently aggravates fulfilling each single criterion.
-Hence, k-plus skew/kurtosis and k-plus correlation performed worse at minimizing $\Delta \mathit{SD}$ than standard k-plus.
-Still, every k-plus method did what it was supposed to do: k-plus skew/kurtosis was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation was best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$.
+However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria aggravated fulfilling each single criterion: k-plus skew/kurtosis and k-plus correlation performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus.
+However, every k-plus method did what it was supposed to do: k-plus skew/kurtosis was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation was best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$.
 At the same time, these two k-plus objectives maintained a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
@@
 However, with increasing $N$, it clearly outperformed diversity anticlustering.
-These results illustrate the cost of optimizing several criteria at the same time: for low $N$, k-plus seems skew/kurtosis to be preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
+These results illustrate the cost of optimizing several criteria at the same time: for low $N$, k-plus skew/kurtosis was apparently preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
 Increasing $N$ then facilitated to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.

### 2023-01-03 paper.Rmd
minor

@@
 
-The Appendix includes additional figures that split the simulation results according to the other variables that varied between simulation runs (number of features; correlation between features; standard deviation of features; number of groups).
-Here, the only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased.
+The Appendix includes additional figures splitting the simulation results according to the other variables that varied between simulation runs (number of features; correlation between features; standard deviation of features; number of groups).
+The only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased.
 Because the number of covariances increases quadratically when the number of features increases, it gets comparatively more difficult to fulfill the correlation objective with a higher number of features.

### 2023-01-03 paper.Rmd
minor

@@
 
-When optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$, the ideal outcome would be to find a partitioning where both constituting criteria (mean and variance) are similar between groups.
-Although research on multi-objective optimization has shown that solutions oftentimes fail to simultaneously satisfy several criteria [@ferligoj1992direct], a motivating example illustrates that in the case of anticlustering, aiming to simultaneously optimize similarity of means and variances may be feasible.
+Research on multi-objective optimization indicates that solutions oftentimes fail to simultaneously satisfy several criteria [@ferligoj1992direct].
+It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ can yield a partitioning where both constituting criteria (mean and variance) are similar between groups.
+A motivating example illustrates that in the case of k-plus anticlustering, aiming to simultaneously optimize similarity of means and variances may be feasible.
 I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.

### 2023-01-03 paper.Rmd
added APA7 support

@@
 
-documentclass     : "apa6"
 output            : 
@@
   - \usepackage{setspace}
+  - \usepackage{float}
+  - \floatstyle{plaintop}
+  - \restylefloat{figure}
+  - |
+    \makeatletter
+    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-1em}%
+      {\normalfont\normalsize\bfseries\typesectitle}}
+
+    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-\z@\relax}%
+      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+    \makeatother
+
+csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
+documentclass     : "apa7"
 ---
@@
 \begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
 <div id="refs" custom-style="Bibliography"></div>

### 2023-01-03 paper.Rmd
minor editing

@@
 
-Because the sample variance is basically defined as an average, k-means anticlustering can be employed to equalize the variance across groups, which is the central insight motivating the new k-plus criterion.
-Therefore, to obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, we have to compute the squared deviation between all individual values and the mean of an attribute.
-We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$, where
-
-$$
-y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2.
-$$
-
-To reflect how strongly the variances of all attributes vary across the $K$ clusters, we now define the criterion $\mathit{SSE_{Var}}$:
+Because the sample variance is basically defined as an average, k-means anticlustering can be employed to equalize the variance across groups, which is the central insight motivating the new k-plus criterion: To obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable, representing the squared distances to the overall mean of the variable.
+We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ where $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
+On the basis of the data matrix $\mathbf{Y}_{N \times P}$ we can define the criterion $\mathit{SSE_{Var}}$ reflecting how strongly the variances of all attributes vary across the $K$ clusters:
 
@@
 However, the input data reflects the squared difference between each data point and the attribute mean instead of the raw data points.
-Maximizing $\mathit{SSE_{Var}}$ will lead to similar variances across the $K$ groups in each of the $P$ attributes.
+Maximizing $\mathit{SSE_{Var}}$ will lead to similar variances across the $K$ groups in the $P$ attributes.
 
@@
 
-Research on multi-objective optimization indicates that solutions oftentimes fail to simultaneously satisfy several criteria [@ferligoj1992direct].
+Research on multiobjective optimization indicates that solutions oftentimes fail to simultaneously satisfy several criteria [@ferligoj1992direct].
 It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ can yield a partitioning where both constituting criteria (mean and variance) are similar between groups.

### 2023-01-04 paper.Rmd
editing

@@
 knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE, dev = "cairo_pdf")
+
+# some functions for motivating example
+
+# Variance objective function with changed order of the arguments 
+# (can be called by sapply() etc.)
+var_objective <- function(clusters, features) {
+  anticlust::variance_objective(features, clusters)
+}
+
+# Compute features for k-means-extended criterion
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
 
@@
 
-Because the sample variance is basically defined as an average, k-means anticlustering can be employed to equalize the variance across groups, which is the central insight motivating the new k-plus criterion: To obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable, representing the squared distances to the overall mean of the variable.
+Because the sample variance is basically defined as an average---and k-means anticlustering minimizes differences with regard to averages---k-means anticlustering can be employed to equalize the variance across groups.
+This constitutes the central insight motivating the new k-plus criterion.
+
+To obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable; the new variable represents the squared distance of each of the original values to the mean of the variable.
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ where $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
@@
 
-Analogously to the standard $\mathit{SSE}$, the centroid values $\overline{y}_j ^{(k)}$ are given as
-
-$$
-\overline{y}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} y_{ij}.
-$$
-
 Note that $\mathit{SSE_{Var}}$ has the same form as the standard $\mathit{SSE}$.
@@
 It is unlikely that an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is of interest without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
-A combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both objectives, which is a reasonable idea when striving for overall between-group similarity.
+A combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both criteria, which is a reasonable idea when striving for overall between-group similarity.
 The most simple approach for such a bicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
@@
 In this case, both the original data $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function.
-That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending these to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data matrix.
-Hence, using the unweighted sum approach, k-plus anticlustering reduces to an augmentation of the data matrix, leaving the original k-means criterion unchanged.
+That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending them to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
+Hence, using the unweighted sum approach, k-plus anticlustering leaves the original k-means criterion unchanged.
 Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective.
-Since the computation of the $y_{ij}$ involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$.
+Because the computation of the new attribute values $y_{ij}$ involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$.
 This imbalance can strongly and unpredictably influence the relative weight of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ when computing $\mathit{SSE_{kplus}}$.
-A solution is to standardize all values $x_{ij}$ and $y_{ij}$ to the same scale before computing $\mathit{SSE_{kplus}}$, e.g.
+To ensure an equal weight of both criteria, all attributes ($x_{ij}$ and $y_{ij}$) can be standardize before computing $\mathit{SSE_{kplus}}$, e.g.
 via $z$-standardization.
@@
 
-Research on multiobjective optimization indicates that solutions oftentimes fail to simultaneously satisfy several criteria [@ferligoj1992direct].
+Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria at the same time [@ferligoj1992direct].
 It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ can yield a partitioning where both constituting criteria (mean and variance) are similar between groups.
@@
 The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$.
-Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (the triangular symbol in Figure 1) is not well-suited to optimize $\mathit{SSE}$ at the same time.
+Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (represented by the triangular symbol) is not well-suited to optimize $\mathit{SSE}$ at the same time.
 Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances.
@@
 ```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
-The partition represented by the triangular minimizes the difference in variances; the partition represented by the square minimizes the difference in means; the partition represented by the diamond maximizes $SSE_{kplus}$ and satisfys both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+The partition represented by the triangular minimizes the difference in variances; the partition represented by the square minimizes the difference in means; the partition represented by the diamond maximizes $SSE_{kplus}$ and satisfies both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
 

### 2023-01-05 paper.Rmd
editing

@@
     Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity.
-Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximzing instead of minimizing a clustering objective function.
+Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximizing instead of minimizing a clustering objective function.
 Introducing k-plus, this paper presents an extension of the k-means clustering objective to maximize between-group similarity in anticlustering applications.
@@
 
-keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis"
-wordcount         : "X"
+keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, stimulus selection"
+wordcount         : "4972"
 
@@
 
-Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity [@brusco2019; @spath1986; @valev1998].
+Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity [@brusco2019; @spath1986; @valev1998; @valev1983].
 Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and separation between clusters [@rokach2005].
@@
 High between-group similarity is for example desirable when designing experimental conditions using different stimulus sets that should be as similar as possible with regard to response-relevant attributes [@lahl2006].
-Stimulus selection tasks are usually conducted on the basis of attribute data and not on pairwise (dis)similarity ratings.
+Such stimulus selection tasks are usually conducted on the basis of attribute data and not on pairwise (dis)similarity ratings.
 That is, individual stimuli are numerically coded on relevant dimensions [@dry2009].
@@
 Subsequently, a distance-based anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the stimuli.
-A different anticlusting approach relies on clustering objectives that can directly be computed on the attribute values.
+A different anticlusting approach uses clustering objectives that can directly be computed on the attribute values.
 K-means is probably the best-known objective of this kind.
@@
 After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
-Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].
+Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].
 
@@
 Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
-This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining $z_i = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, i.e., the concatenation of the $x_{ij}$ and $y_{ij}$ values, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
+This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining $z_{i.} = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
 
@@
 Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective.
-Because the computation of the new attribute values $y_{ij}$ involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$.
+Because the computation of the new attribute values $y_{ij}$ involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$ (e.g., see Figure 1).
 This imbalance can strongly and unpredictably influence the relative weight of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ when computing $\mathit{SSE_{kplus}}$.
@@
 The diamond symbol represents the partition that maximizes $\mathit{SSE_{kplus}}$, which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
-Thus, the optimal partition according to $\mathit{SSE_{kplus}}$ fulfills the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ to an astonishing degree.
+Thus, using the unweighted sum approach, the optimal partition according to $\mathit{SSE_{kplus}}$ fulfills the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ to an astonishing degree.
 
@@
 To implement all k-plus objectives, the unweighted sum approach was employed.
-Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria; subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
+Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.
+Subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
 
@@
 All anticlustering methods used a $z$-standardization of the data sets as a preprocessing step.
-In the case of k-plus anticlustering, the augmented input data---i.e., the original data as well as the features that were added to incorporate the additional optimization criteria---were subjected to a standardization, which is recommended.
 
@@
 
-The same local maximum search [Method LCW, @weitz1998] was applied to optimize each of the five competing anticlustering objectives.
-Building on an initial random assignment, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
+The same local maximum search was applied to optimize each of the five competing anticlustering objectives [Method LCW, @weitz1998].
+Building on an initial random allocation, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
 That is, it starts with the first element and simulates all exchanges with elements that are currently assigned to a different group.
-After each exchange has been processed, the one exchange is realized that improves the objective function the most.
+After each exchange has been evaluated, the one exchange is realized that improves the objective function the most.
 No exchange is realized if no improvement is possible.
@@
 To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
-The simulation study always employed five repetitions of the local maximum search.
+In the simulation study five repetitions were used.
 
@@
 However, with increasing $N$, it clearly outperformed diversity anticlustering.
-These results illustrate the cost of optimizing several criteria at the same time: for low $N$, k-plus skew/kurtosis was apparently preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
+These results illustrate the cost of optimizing several criteria at the same time: For low $N$, k-plus skew/kurtosis was apparently preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
 Increasing $N$ then facilitated to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
@@
 The Appendix includes additional figures splitting the simulation results according to the other variables that varied between simulation runs (number of features; correlation between features; standard deviation of features; number of groups).
-The only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased.
+Here, the only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased.
 Because the number of covariances increases quadratically when the number of features increases, it gets comparatively more difficult to fulfill the correlation objective with a higher number of features.
 
-# K-plus anticlustering using the R package anticlust
+# Applying k-plus anticlustering using the R package anticlust
 
@@
 @brielmann2019intense collected an additional rating dimension by measuring how the same 900 images were rated with regard to their beauty.
-In my application, I use all three features; the corresponding data is available from Github.[^githubbeauty]
+In my application, I used all three features; the corresponding data is available from Github.[^githubbeauty]
 
@@
 This behaviour can be adjusted using the boolean argument `standardize`.
-The documentation of the `anticlust` package also provides detailed information on how to use the `kplus_anticlustering()` function.
+The documentation of the `anticlust` package also provides detailed information on how to use the function `kplus_anticlustering()`.
 
 Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features.
-Table 2 also includes results of k-means anticlustering and diversity anticlustering.
+Table 2 also contains the results after applying k-means anticlustering and diversity anticlustering.
 It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations, and the skewness was also very evenly matched.
@@
 K-plus anticlustering strives for between-group similarity regardless of group size.
-Unequal group sizes can be requested in `anticlust` by passing the different group sizes to the argument `K`.
+Unequal group sizes can be requested by passing the different group sizes to the argument `K`.
 The following code was used to equalize means and variances between four groups using the k-plus approach:
@@
 
-Anticlustering is a partitioning method that aims for high between-group similarity and within-group heterogeneity.
+Anticlustering is a partitioning method that aims for high between-group similarity and high within-group heterogeneity.
 In this paper, I presented the k-plus criterion for anticlustering, which focuses on maximizing between-group similarity.
@@
 
-The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion: A simple unweighted sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective.
+The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion.
+A simple unweighted sum approach was used to optimize the k-plus objective, while more sophisticated algorithmic procedures exist to tackle multicriterion optimization problems [e.g., @brusco2009cross].
 The sole focus on the objective function followed two reasonings.
 First, the applications showed that the local maximum search, maximizing the unweighted k-plus sum, yielded satisfying results that probably suffice in most applications.
-It seems that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria at the same time (see Figure 1).
-Still, more sophisticated approaches may yield improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
-@brusco2019 gave an example of how direct algorithms can be employed to optimize multiple criteria at the same time in anticlustering applications aiming for high within-group heterogeneity.
+It is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria to a satisfactory degree (see Figure 1).
+Still, more sophisticated approaches may yield theoretically improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
+@brusco2019 gave an example of how direct algorithms can be employed to optimize multiple criteria in anticlustering applications aiming for high within-group heterogeneity.
 Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.
-Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it.
-To obtain a fair assessment in the simulation study, it was therefore important that each anticlustering objective was optimized using the same algorithm.
-Differences in performance cannot be interpreted if the objective function as well as the algorithm used to optimize it differs between methods.
+A second reason for focusing on the k-plus objective per se---independently from the algorithm that is used to optimize it---was to obtain a fair comparison to existing anticlustering objectives, i.e., the classical k-means criterion and the diversity.
+In the simulation study, differences in performance could not have been interpreted if the objective function as well as the algorithm used to optimize it had differed between anticlustering methods.
 It was of particular interest to compare the k-plus objective with the popular diversity objective.

### 2023-01-05 paper.Rmd
minor

@@
 By considering pairwise dissimilarities from between elements in the same group, the diversity reflects the total degree of within-group heterogeneity.
-High within-group diversity simultaneously ensures high between-group similarity.[^equalsizedgroups] Another important anticlustering criterion is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
+High within-group diversity simultaneously ensures high between-group similarity.[^equalsizedgroups] Another important anticlustering criterion based on the information in a dissimilarity matrix is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
 Maximizing the dispersion increases within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.

### 2023-01-10 appendix.Rmd
including Tim's comments

@@
 
-```{r Figure2, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *M*."}
+```{r Figure2, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by the number of features, averaged across the remaining variables that varied in the simulation)."}
 ### For the appendix -- other splits for the main results as plot (N is most important)
-# Plot the results, by M
+# Plot the results, by M (number of features)
 ldf %>% 
@@
       levels = c("means", "sd", "skew", "kur", "cor"),
-      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
+      labels = paste0(
+        "Delta[italic(",
+        c("M", "SD", "Skew", "Kurtosis", "Cor"),
+        ")]"
+      )
+    ) 
   ) %>% 
@@
   geom_line(aes(linetype = method), size = .85) + 
-  facet_grid(rows = vars(Objective), scales = "free") + 
+  facet_grid(rows = vars(Objective), scales = "free", labeller = label_parsed) + 
   ylab("Mean discrepancy") +
+  xlab("Number of features") +
   theme_bw(base_size = 16) + 
@@
 
-```{r Figure3, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *SD*."}
+```{r Figure3, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by *SD*, averaged across the remaining variables that varied in the simulation)."}
 # Plot the results, by SD
@@
   geom_line(aes(linetype = method), size = .85) + 
-  facet_grid(rows = vars(Objective), scales = "free") + 
-  ylab("Mean discrepancy") +
+  facet_grid(rows = vars(Objective), scales = "free", labeller = label_parsed) + 
+  ylab("Mean discrepancy")+
+  xlab(expression(italic(SD))) +
   theme_bw(base_size = 16) + 
@@
 
-```{r Figure4, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *K*."}
+```{r Figure4, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by *K*, averaged across the remaining variables that varied in the simulation)."}
 # Plot the results, by K
@@
   geom_line(aes(linetype = method), size = .85) + 
-  facet_grid(rows = vars(Objective), scales = "free") + 
-  ylab("Mean discrepancy") +
+  facet_grid(rows = vars(Objective), scales = "free", labeller = label_parsed) + 
+  ylab("Mean discrepancy")+
+  xlab(expression(italic(K))) +
   theme_bw(base_size = 16) + 
@@
 
-```{r Figure5, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *r*."}
+```{r Figure5, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by *r*, averaged across the remaining variables that varied in the simulation)."}
 # Plot the results, by r
@@
   geom_line(aes(linetype = method), size = .85) + 
-  facet_grid(rows = vars(Objective), scales = "free") + 
-  ylab("Mean discrepancy") +
+  facet_grid(rows = vars(Objective), scales = "free", labeller = label_parsed) + 
+  ylab("Mean discrepancy")+
+  xlab(expression(italic(r))) +
   theme_bw(base_size = 16) + 

### 2023-01-10 paper.Rmd
including Tim's comments

@@
 The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, higher order moments, and covariances), while the classical k-means criterion only reflects the degree to which groups differ in their means.
-It is shown that the k-plus criterion can be reduced to an augmentation of the input data while the original k-means objective remains unchanged.
+While constituting a new criterion for anticlustering, it is surprisingly shown that k-plus anticlustering can be implemented by optimizing the classical k-means criterion after the input data has been augmented with additional variables.
 A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple objectives.
 In particular, optimizing between-group similarity with regard to variances usually does not compromise similarity with regard to means; the k-plus extension is therefore generally preferred over classical k-means anticlustering.
-Example are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
+Examples are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
 
@@
 
-# Compute features for k-means-extended criterion
+# Compute features for k-plus criterion
 squared_from_mean <- function(data) {
@@
 Several anticlustering objective functions use pairwise dissimilarity ratings such as the Euclidean distance as input.
-The most prominent criterion of this kind is the *diversity*, which is the total sum of dissimilarities between elements within the same group [@brusco2019; @gallego2013].
-By considering pairwise dissimilarities from between elements in the same group, the diversity reflects the total degree of within-group heterogeneity.
+The most prominent criterion of this kind is the *diversity*, which is the total sum of pairwise dissimilarities between elements within the same group [@brusco2019; @gallego2013].
+By considering pairwise dissimilarities between elements in the same group, the diversity criterion reflects the total degree of within-group heterogeneity.
 High within-group diversity simultaneously ensures high between-group similarity.[^equalsizedgroups] Another important anticlustering criterion based on the information in a dissimilarity matrix is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
-Maximizing the dispersion increases within-group heterogeneity by ensuring that any two elements in the same group as dissimilar from each other as possible.
+Maximizing the dispersion increases within-group heterogeneity by ensuring that any two elements in the same group are as dissimilar from each other as possible.
 @brusco2019 convincingly argued that anticlustering applications striving for within-group heterogeneity should incorporate both dispersion and diversity, and they presented a bicriterion algorithm to approximate the Pareto efficient set of solutions according to both criteria.
@@
 In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977].
-In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019].
-Subsequently, a distance-based anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the stimuli.
-A different anticlusting approach uses clustering objectives that can directly be computed on the attribute values.
-K-means is probably the best-known objective of this kind.
+When attribute values are available, anticlustering approaches that do not require a dissimilarity matrix can be used to partition the data.
+K-means is probably the best-known clustering objective that can directly be computed on attribute values.
 In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized, usually using the k-means heuristic [@jain2010; @steinley2006; @brusco2006branch].
@@
 @papenberg2020 showed that k-means anticlustering tends to over-optimize between-group similarity with regard to the mean at the cost of similarity in variance.
-However, when aiming for overall between-group similarity, the negligence of distribution characteristics other than the mean is misguided; similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
+However, when aiming for overall between-group similarity, neglecting all distribution characteristics other than the mean is misguided; similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
 
@@
 If the $\mathit{SSE}$ is maximal---which is the goal of k-means *anticlustering*---, the cluster centroids are as close as possible to the overall centroid $\overline{\mathbf{x}}$ and therefore to each other [@spath1986].
-Thus, k-means anticlustering directly optimizes the similarity of the mean attribute values across clusters.
+Thus, k-means anticlustering directly optimizes the similarity of the mean attribute values between clusters.
 
@@
 
-To obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable; the new variable represents the squared distance of each of the original values to the mean of the variable.
+To obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable; the new variable represents the squared distance of each of the original values to the overall mean of the variable.
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ where $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
@@
 Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
-This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining $z_{i.} = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, x_{iP})$, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
+This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining $z_{i.} = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, y_{iP})$, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
 
@@
 This imbalance can strongly and unpredictably influence the relative weight of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ when computing $\mathit{SSE_{kplus}}$.
-To ensure an equal weight of both criteria, all attributes ($x_{ij}$ and $y_{ij}$) can be standardize before computing $\mathit{SSE_{kplus}}$, e.g.
+To ensure an equal weight of both criteria, all attributes ($x_{ij}$ and $y_{ij}$) can be standardized before computing $\mathit{SSE_{kplus}}$, e.g.
 via $z$-standardization.
@@
 
-```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
-The partition represented by the triangular minimizes the difference in variances; the partition represented by the square minimizes the difference in means; the partition represented by the diamond maximizes $SSE_{kplus}$ and satisfies both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+```{r, fig.cap = "The relationship between the objectives $\\mathit{SSE}$ and $\\mathit{SSE_{Var}}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
+The partition represented by the square maximizes $\\mathit{SSE}$ and thus minimizes the difference in means between groups; the partition represented by the triangular maximizes $\\mathit{SSE_{Var}}$ and thus minimizes the difference in variances between groups; the partition represented by the diamond maximizes $\\mathit{SSE_{kplus}}$ and satisfies both constituting criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
 
@@
 
-The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means; (ii) diversity anticlustering, maximizing the sum of Euclidean distances within groups; (iii) a simple random allocation of elements to groups.
+The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means; (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups; (iii) a simple random allocation of elements to groups.
 All anticlustering methods used a $z$-standardization of the data sets as a preprocessing step.
@@
 ## Results
+
+Table 1 displays the global simulation results aggregated across all conditions.
+Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but did not outperform a random assignment with regard to all other distribution characteristics.
+Replicating @papenberg2020, k-means' performance with regard to similarity in standard deviations was even below that of a random assignment.
+K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); all three k-plus approaches outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
 
@@
   note = "The global results of the simulation study aggregated across all conditions.
-Cells contain information about the average between-group discrepancy with regard to means, standard deviations, skewness, kurtosis, and correlations.
+Cells contain information about the average global between-group discrepancy with regard to means, standard deviations, skewness, kurtosis, and correlations.
 Lower values indicate less discrepancy, i.e., higher between-group similarity.
@@
 
-```{r Figure1, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by $N$."}
+```{r Figure1, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by $N$, averaged across the remaining variables that varied in the simulation)."}
 
@@
       levels = c("means", "sd", "skew", "kur", "cor"),
-      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
+      labels = paste0(
+        "Delta[italic(",
+        c("M", "SD", "Skew", "Kurtosis", "Cor"),
+        ")]"
+      )
+    ) 
   ) %>%
@@
   geom_line(aes(linetype = method), size = .85) + 
-  facet_grid(rows = vars(Objective), scales = "free") + 
-  ylab("Mean discrepancy") +
+  facet_grid(rows = vars(Objective), scales = "free", labeller = label_parsed) + 
+  ylab("Mean discrepancy")+
+  xlab(expression(italic(N))) +
   theme_bw(base_size = 16) + 
@@
 ```
-
-Table 1 displays the global simulation results aggregated across all conditions.
-Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but did not outperform a random assignment with regard to all other distribution characteristics.
-Replicating @papenberg2020, k-means' performance with regard to similarity in standard deviations was even below that of a random assignment.
-K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); all three k-plus approaches outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
 
@@
 
-I used k-plus anticlustering to divide the data set of 900 images into 9 groups of 100 images each. The task was accomplished using the following R code:
+I used the function `kplus_anticlustering()` to divide the data set of 900 images into 9 groups of 100 images each.
+The first argument of the function call is a variable named `features`, which contains the image data as a table: 900 rows representing images and 3 columns representing the features beauty, arousal, and valence.
+The accompanying OSF repository contains the full code to reproduce the example (**TODO**).
 
@@
 kplus_anticlustering(
-  features,
+  features, # data table with 900 rows and 3 features
   K = 9,
@@
 
-The code runs in about 10 seconds on a contemporary personal computer; the run time may vary depending on how many iterations are needed to find the local maximum.
-The variable `features` is a data table containing the image data: 900 rows representing images and 3 columns representing the features beauty, arousal, and valence.
-The accompanying OSF repository contains code to reproduce the example (**TODO**).
-
-The `kplus_anticlustering()` function has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion.
+The function call takes about 10 seconds on a contemporary personal computer; the run time may vary depending on how many iterations are needed to find the local maximum.
+The function `kplus_anticlustering()` has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion.
 Only the argument `variance` is set to `TRUE` by default---corresponding to the understanding that k-plus is an extension to k-means that at least equalizes variances in addition to means.
@@
 
-Table 4 illustrates the results that are obtained when using standard k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
+Table 4 illustrates the results that are obtained when using standard[^standardkplus] k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
 Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are similar enough for any practical purposes.
+
+[^standardkplus]: I refer to *standard* k-plus anticlustering as the minimization of differences in means and variances, but not in higher order moments or covariances.
 
@@
 In this paper, I presented the k-plus criterion for anticlustering, which focuses on maximizing between-group similarity.
-K-plus represents between-group similarity as discrepancies in distribution moments between groups: mean, variances, skewness, kurtosis, higher order moments, and covariances.
+K-plus represents between-group similarity as discrepancies in distribution moments between groups: means, variances, skewness, kurtosis, higher order moments, and covariances.
 Depending on the application, a different subset of these objectives can be included in the anticlustering process.
@@
 If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variance).
-Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.
-Moreover, using the open source R package `anticlust`, the (repeated) application of k-plus anticlustering is easily accessible and free.
+Using the open source R package `anticlust`, the (repeated) application of k-plus anticlustering is easily accessible and free.
 

### 2023-01-11 paper.Rmd
word count
added function declarations to chunk with example
added info on devel version of papaja

@@
 keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, stimulus selection"
-wordcount         : "4972"
+wordcount         : "5114"
 
@@
 ```{r setup, include = FALSE}
-library("papaja")
+library("papaja") # use the development version: `remotes::install_github("crsh/papaja@devel")`; currently 0.1.1.9001
 library("anticlust")
@@
 knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE, dev = "cairo_pdf")
-
-# some functions for motivating example
-
-# Variance objective function with changed order of the arguments 
-# (can be called by sapply() etc.)
-var_objective <- function(clusters, features) {
-  anticlust::variance_objective(features, clusters)
-}
-
-# Compute features for k-plus criterion
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
 
@@
 
-```{r}
+```{r motivatingExample}
+
+# some functions for motivating example
+
+# Variance objective function with changed order of the arguments 
+# (can be called by sapply() etc.)
+var_objective <- function(clusters, features) {
+  anticlust::variance_objective(features, clusters)
+}
+
+# Compute features for k-plus criterion
+squared_from_mean <- function(data) {
+  apply(data, 2, function(x) (x - mean(x))^2)
+}
 
@@
 The R package `faux` [Version 1.1.0, @R-faux] was used to generate the multivariate normal data sets that were processed during the simulation.
-The R packages `dplyr` [Version 1.0.7, @R-dplyr], `ggplot2` [Version 3.3.5, @R-ggplot2], `papaja` [Version 0.1.0.9997, @R-papaja], `tidyr` [Version 1.1.3, @R-tidyr] and `DescTools` [Version 0.99.45, @R-desctools] were used for data analysis and result presentation.
+The R packages `dplyr` [Version 1.0.7, @R-dplyr], `ggplot2` [Version 3.3.5, @R-ggplot2], `papaja` [Version 0.1.1.9001, @R-papaja], `tidyr` [Version 1.1.3, @R-tidyr] and `DescTools` [Version 0.99.45, @R-desctools] were used for data analysis and result presentation.
 The simulation study is fully reproducible via code and data that is accessible from the accompanying OSF repository (**link TODO**).

### 2023-01-11 Simulation_Study/0-functions-analyze-data.R
updated year in simlation study directory

@@ -1,6 +1,6 @@
 
 # Author: Martin Papenberg
-# Year: 2022
+# Year: 2023
 
 #' Compute objective values for a given method and data set
 #' 

### 2023-01-11 Simulation_Study/0-functions-anticlustering-methods.R

@@ -1,6 +1,6 @@
 
 # Author: Martin Papenberg
-# Year: 2022
+# Year: 2023
 
 #' Apply several anticlustering methods to a data set
 #' 

### 2023-01-11 Simulation_Study/0-functions-generate-data.R

@@ -1,6 +1,6 @@
 
 # Author: Martin Papenberg
-# Year: 2022
+# Year: 2023
 
 #' Generate a data set for k-plus anticlustering simulation.
 #' Uses faux::rnorm_multi to generate the multivariate normal data.

### 2023-01-11 Simulation_Study/1-Generate-Data.R

@@ -1,6 +1,6 @@
 
 # Author: Martin Papenberg
-# Year: 2022
+# Year: 2023
 
 # First step in the simulation: Generate data
 source("0-functions-generate-data.R")

### 2023-01-11 Simulation_Study/2-Call-Methods.R

@@ -1,6 +1,6 @@
 
 # Author: Martin Papenberg
-# Year: 2022
+# Year: 2023
 
 # Load required packages
 library(anticlust)

### 2023-01-11 Simulation_Study/4-Aggregate-Results.R

@@ -1,6 +1,6 @@
 
 # Author: Martin Papenberg
-# Year: 2022
+# Year: 2023
 
 # Load required packages
 library(dplyr)

### 2023-01-11 paper.Rmd
minor

@@
 ```{r, fig.cap = "The relationship between the objectives $\\mathit{SSE}$ and $\\mathit{SSE_{Var}}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
-The partition represented by the square maximizes $\\mathit{SSE}$ and thus minimizes the difference in means between groups; the partition represented by the triangular maximizes $\\mathit{SSE_{Var}}$ and thus minimizes the difference in variances between groups; the partition represented by the diamond maximizes $\\mathit{SSE_{kplus}}$ and satisfies both constituting criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+The partition represented by the square maximizes $\\mathit{SSE}$ and thus minimizes the difference in means between groups; the partition represented by the triangular maximizes $\\mathit{SSE_{Var}}$ and thus minimizes the difference in variances between groups.
+The partition represented by the diamond maximizes $\\mathit{SSE_{kplus}}$ and satisfies both constituting criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
 
@@
 
-First, I generated 10,000 data sets following a normal distribution.
+For the simulation, I generated 10,000 data sets following a normal distribution.
 The data sets were subsequently processed by the competing methods.
@@
 Figure 2 shows that with increasing $N$, k-means and the k-plus methods converged on the same level with regard to minimizing $\Delta \mathit{M}$.
-This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time---especially when $N$ increases.
+This is an important observation because it shows that including the k-plus terms in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time---especially when $N$ increases.
 However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria aggravated fulfilling each single criterion: k-plus skew/kurtosis and k-plus correlation performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus.

### 2023-01-11 paper.Rmd
load all packages in first chunk

@@
 ```{r setup, include = FALSE}
-library("papaja") # use the development version: `remotes::install_github("crsh/papaja@devel")`; currently 0.1.1.9001
-library("anticlust")
-library("knitr")
-library("dplyr")
-library("ggplot2")
-library("tidyr")
+library(papaja) # use the development version: `remotes::install_github("crsh/papaja@devel")`; currently 0.1.1.9001
+library(anticlust)
+library(knitr)
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(DescTools)
 # for formatting numbers in Rmarkdown, use my package prmisc (https://github.com/m-Py/prmisc)
-library("prmisc") # remotes::install_github("m-Py/prmisc") 
+library(prmisc) # remotes::install_github("m-Py/prmisc") 
 options(digits = 2)

### 2023-01-25 paper.Rmd
included Juli's comments

@@
     \makeatother
+  - |
+    \makeatletter
+    \renewcommand*\env@matrix[1][*\c@MaxMatrixCols c]{%
+      \hskip -\arraycolsep
+      \let\@ifnextchar\new@ifnextchar
+      \array{#1}}
+    \makeatother
 
@@
 In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977].
-When attribute values are available, anticlustering approaches that do not require a dissimilarity matrix can be used to partition the data.
+When attribute values are available, anticlustering approaches that do not require a dissimilarity matrix[^dissimilarityfromfeatures] can be used to partition the data.
 K-means is probably the best-known clustering objective that can directly be computed on attribute values.
@@
 
+[^dissimilarityfromfeatures]: It is still possible to convert the attribute data into a dissimilarity matrix, e.g.
+by computing all pairwise Euclidean distances, and then optimize a distance-based anticlustering criterion such as the diversity to obtain a partitioning.
+
 To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering.
@@
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
-After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well suited to create groups having minimal differences according to multiple distribution characteristics.
+After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics.
 Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].
@@
 
-The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all observations belonging the $k$'th cluster $C_k$ ($k = 1, \ldots, K$):
+The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all observations belonging to the $k$'th cluster $C_k$ ($k = 1, \ldots, K$):
 
@@
 
-Note there is a direct connection between the $\mathit{SSE}$ and the location of the cluster centroids [@spath1986].
+Note that there is a direct connection between the $\mathit{SSE}$ and the location of the cluster centroids [@spath1986].
 If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the overall mean value on the $p$th attribute:
@@
 
-To obtain an adaption of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable; the new variable represents the squared distance of each of the original values to the overall mean of the variable.
+To obtain an adaptation of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable; the new variable represents the squared distance of each of the original values to the overall mean of the variable.
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ where $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
@@
 Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
-This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining $z_{i.} = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, y_{iP})$, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
+This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining a new data matrix $\mathbf{Z}_{N \times 2P} = \{z_{ij}\}_{N \times 2P}$ as the augmentation of $\mathbf{X}_{N \times P}$ with $\mathbf{Y}_{N \times P}$, i.e.
+
+$$
+  \mathbf{Z}_{N \times 2P} =
+  \left[ {\begin{array}{cccc|cccc}
+    x_{11} & x_{12} & \cdots & x_{1P} & y_{11} & y_{12} & \cdots & y_{1P}\\
+    x_{21} & x_{22} & \cdots & x_{2P} & y_{21} & y_{22} & \cdots & y_{2P}\\
+    \vdots & \vdots & \ddots & \vdots & \vdots & \vdots & \ddots & \vdots\\
+    x_{N1} & x_{N2} & \cdots & x_{NP} & y_{N1} & y_{N2} & \cdots & y_{NP}\\
+  \end{array} } \right]
+$$
+
+and $z_{i.} = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, y_{iP})$, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
 
@@
   \sum\limits_{i \in C_k}
-  (z_{ij} - \overline{z}_j ^{(k)})^2
-$$.
+  (z_{ij} - \overline{z}_j ^{(k)})^2.
+$$
 
 This formula again has the same form as $\mathit{SSE}$ but is computed on different data.
-In this case, both the original data $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function.
-That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending them to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
+In this case, both the original values $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function.
+That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
 Hence, using the unweighted sum approach, k-plus anticlustering leaves the original k-means criterion unchanged.
+
 Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective.
@@
 The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means; (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups; (iii) a simple random allocation of elements to groups.
-All anticlustering methods used a $z$-standardization of the data sets as a preprocessing step.
 
@@
 After the five anticlustering methods and the random allocation procedure had been applied to the 10,000 data sets for $K = 2$, $K = 3$, $K = 4$, I investigated how well the competing methods were able to create similar groups.
-To this end, I investigated the discrepancy in means, standard deviations, skewness, kurtosis, and correlation between groups---for each data set for each $K$.
+To this end, I analyzed the discrepancy in means, standard deviations, skewness, kurtosis, and correlation between groups---for each data set for each $K$.
 To quantify the discrepancy in group means, the following computation was used: for each of the (2, 3, 4 or 5) features, all (2, 3 or 4) group means were computed.
@@
 Figure 2 shows that with increasing $N$, k-means and the k-plus methods converged on the same level with regard to minimizing $\Delta \mathit{M}$.
-This is an important observation because it shows that including the k-plus terms in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time---especially when $N$ increases.
-However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria aggravated fulfilling each single criterion: k-plus skew/kurtosis and k-plus correlation performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus.
+This is an important observation because it shows that k-plus anticlustering is capable of addressing multiple objectives at the same time---especially when $N$ increases.
+However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria made it more difficult to fulfill each single criterion: k-plus skew/kurtosis and k-plus correlation performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus.
 However, every k-plus method did what it was supposed to do: k-plus skew/kurtosis was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation was best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$.
@@
 It is maybe surprising that k-plus skew/kurtosis does not seem to have performed much better than diversity anticlustering with regard to skewness and---in particular---kurtosis, even though the method has been specifically tailored to these criteria.
-Figure 2 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis was even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
+Figure 2 gives a more fine grained assessment of this observation: Surprisingly, for small $N$, k-plus skew/kurtosis was even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
 However, with increasing $N$, it clearly outperformed diversity anticlustering.
@@
 assembled 900 open-access color images and collected ratings on two affective dimensions: arousal and valence.
-@brielmann2019intense collected an additional rating dimension by measuring how the same 900 images were rated with regard to their beauty.
+@brielmann2019intense collected additional norming data by measuring how the same 900 images were rated with regard to their beauty.
 In my application, I used all three features; the corresponding data is available from Github.[^githubbeauty]
@@
 
-The function call takes about 10 seconds on a contemporary personal computer; the run time may vary depending on how many iterations are needed to find the local maximum.
+The function call takes about 10 seconds on a contemporary personal computer.
 The function `kplus_anticlustering()` has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion.
@@
 It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations, and the skewness was also very evenly matched.
-K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations skewness in particular.
+K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations, and skewness in particular.
 Diversity anticlustering was well-suited to match means, standard deviations and skew, but was slightly outperformed by k-plus.
@@
 
-A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.
-Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be considered as the default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be aggravated in small data sets.
-In practice, users can try out different objectives and see which specification fulfils their needs.
+A simulation study and examples on real norming data showed that the k-plus criterion is well-suited to maximize between-group similarity with regard to multiple criteria.
+Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be considered as the default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be challenging in small data sets.
+In practice, users can try out different objectives and see which specification fulfills their needs.
 A reasonable approach is to start "greedy", e.g.
@@
 First, the applications showed that the local maximum search, maximizing the unweighted k-plus sum, yielded satisfying results that probably suffice in most applications.
-It is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria to a satisfactory degree (see Figure 1).
+It is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria to a satisfactory degree (as suggested by Figure 1).
 Still, more sophisticated approaches may yield theoretically improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
@@
 [^clusterediting]: In their study, they used the term anticluster editing to refer to the maximization of the diversity criterion.
-This is because the diversity criterion is minimized as part of the cluster editing method [@zahn1964; @shamir2004vf].
+This is because the diversity criterion is minimized as part of the cluster editing method [@shamir2004vf].
 

### 2023-01-25 paper.Rmd
added URL for OSF repo

@@
 The R packages `dplyr` [Version 1.0.7, @R-dplyr], `ggplot2` [Version 3.3.5, @R-ggplot2], `papaja` [Version 0.1.1.9001, @R-papaja], `tidyr` [Version 1.1.3, @R-tidyr] and `DescTools` [Version 0.99.45, @R-desctools] were used for data analysis and result presentation.
-The simulation study is fully reproducible via code and data that is accessible from the accompanying OSF repository (**link TODO**).
+The simulation study is fully reproducible via code and data that has been made accessible via an accompanying OSF repository [@osf2023].
 
@@
 The first argument of the function call is a variable named `features`, which contains the image data as a table: 900 rows representing images and 3 columns representing the features beauty, arousal, and valence.
-The accompanying OSF repository contains the full code to reproduce the example (**TODO**).
+The accompanying OSF repository contains the full code to reproduce the example [@osf2023].
 

### 2023-01-27 paper.Rmd
minor

@@
 In particular, optimizing between-group similarity with regard to variances usually does not compromise similarity with regard to means; the k-plus extension is therefore generally preferred over classical k-means anticlustering.
-Examples are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
+Examples are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust).
 
@@
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ where $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
-On the basis of the data matrix $\mathbf{Y}_{N \times P}$ we can define the criterion $\mathit{SSE_{Var}}$ reflecting how strongly the variances of all attributes vary across the $K$ clusters:
+On the basis of the data matrix $\mathbf{Y}_{N \times P}$ we define the criterion $\mathit{SSE_{Var}}$, reflecting how strongly the variances of all attributes vary across the $K$ clusters:
 
@@
 Replicating @papenberg2020, k-means' performance with regard to similarity in standard deviations was even below that of a random assignment.
-K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); all three k-plus approaches outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
+K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); all three k-plus approaches strongly outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
 
@@
 The first argument of the function call is a variable named `features`, which contains the image data as a table: 900 rows representing images and 3 columns representing the features beauty, arousal, and valence.
-The accompanying OSF repository contains the full code to reproduce the example [@osf2023].
+An OSF repository accompanying this manuscript contains the full code to reproduce the example [@osf2023].
 
@@
 This behaviour can be adjusted using the boolean argument `standardize`.
-The documentation of the `anticlust` package also provides detailed information on how to use the function `kplus_anticlustering()`.
 
@@
 It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations, and the skewness was also very evenly matched.
-K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations, and skewness in particular.
+K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations and skewness.
 Diversity anticlustering was well-suited to match means, standard deviations and skew, but was slightly outperformed by k-plus.
@@
 I make use of a norming data set of 96 word stimuli that was contributed to the `anticlust` package by Marie L.
-Schaper [@schaper2019metacognitive; @schaper2019metamory; also see @papenberg2020 for a description of the norming data].
+Schaper [@schaper2019metacognitive; @schaper2019metamory].
 After loading the `anticlust` package, the data set can be accessed as follows:
@@
 Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.
-A second reason for focusing on the k-plus objective per se---independently from the algorithm that is used to optimize it---was to obtain a fair comparison to existing anticlustering objectives, i.e., the classical k-means criterion and the diversity.
+A second reason for focusing on the k-plus objective per se---independently from the algorithm that is used to optimize it---was to obtain a fair comparison to existing anticlustering objectives.
 In the simulation study, differences in performance could not have been interpreted if the objective function as well as the algorithm used to optimize it had differed between anticlustering methods.

### 2023-01-27 paper.Rmd
minor change abstract

@@
 Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximizing instead of minimizing a clustering objective function.
-Introducing k-plus, this paper presents an extension of the k-means clustering objective to maximize between-group similarity in anticlustering applications.
-The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, higher order moments, and covariances), while the classical k-means criterion only reflects the degree to which groups differ in their means.
-While constituting a new criterion for anticlustering, it is surprisingly shown that k-plus anticlustering can be implemented by optimizing the classical k-means criterion after the input data has been augmented with additional variables.
+Introducing k-plus, this paper presents an extension of the classical k-means objective to maximize between-group similarity in anticlustering applications.
+The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, higher order moments, and covariances), whereas the k-means criterion only reflects the degree to which groups differ in their means.
+While constituting a new criterion for anticlustering, it is surprisingly shown that k-plus anticlustering can be implemented by optimizing the original k-means criterion after the input data has been augmented with additional variables.
 A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple objectives.

### 2023-01-27 paper.Rmd
some 'last changes'

@@
 [^dissimilarityfromfeatures]: It is still possible to convert the attribute data into a dissimilarity matrix, e.g.
-by computing all pairwise Euclidean distances, and then optimize a distance-based anticlustering criterion such as the diversity to obtain a partitioning.
+by computing all pairwise Euclidean distances, and then optimize a distance-based anticlustering criterion to obtain a partitioning.
 
@@
 This constitutes the central insight motivating the new k-plus criterion.
-
-To obtain an adaptation of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable; the new variable represents the squared distance of each of the original values to the overall mean of the variable.
-We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ where $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
-On the basis of the data matrix $\mathbf{Y}_{N \times P}$ we define the criterion $\mathit{SSE_{Var}}$, reflecting how strongly the variances of all attributes vary across the $K$ clusters:
+Thus, to obtain an adaptation of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
+We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
+On the basis of $\mathbf{Y}_{N \times P}$, we define the criterion $\mathit{SSE_{Var}}$, reflecting how strongly the variances of all attributes vary across the $K$ clusters:
 
@@
 
-This formula again has the same form as $\mathit{SSE}$ but is computed on different data.
-In this case, both the original values $x_{ij}$ as well as the squared distances $y_{ij}$ are used as attributes that enter the objective function.
-That is, computing the k-plus criterion can be done by first computing the values $y_{ij}$, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
-Hence, using the unweighted sum approach, k-plus anticlustering leaves the original k-means criterion unchanged.
+This term again has the same form as $\mathit{SSE}$ but is computed on different data.
+Hence, computing the k-plus criterion can be done by first computing the squared distances $y_{ij}$, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
+Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
 
-Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when optimizing the k-plus objective.
+Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when computing $\mathit{SSE_{kplus}}$.
 Because the computation of the new attribute values $y_{ij}$ involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$ (e.g., see Figure 1).
@@
 It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ can yield a partitioning where both constituting criteria (mean and variance) are similar between groups.
-A motivating example illustrates that in the case of k-plus anticlustering, aiming to simultaneously optimize similarity of means and variances may be feasible.
+A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible.
 I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
@@
 
-To thoroughly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to the mean and variance; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to the mean, variance, skew and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to the mean, variance and covariance structure.
+To thoroughly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to group means and variances; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to group means, variances, skewness and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to group means, variances and covariance structure.
 To implement all k-plus objectives, the unweighted sum approach was employed.
@@
 Subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
-
-The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means; (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups; (iii) a simple random allocation of elements to groups.
+The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means, (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups, (iii) a random allocation of elements to groups.
 
@@
 The R packages `dplyr` [Version 1.0.7, @R-dplyr], `ggplot2` [Version 3.3.5, @R-ggplot2], `papaja` [Version 0.1.1.9001, @R-papaja], `tidyr` [Version 1.1.3, @R-tidyr] and `DescTools` [Version 0.99.45, @R-desctools] were used for data analysis and result presentation.
-The simulation study is fully reproducible via code and data that has been made accessible via an accompanying OSF repository [@osf2023].
+The simulation study is fully reproducible via code and data that has been made accessible in an accompanying OSF repository [@osf2023].
 
@@
 
-It is maybe surprising that k-plus skew/kurtosis does not seem to have performed much better than diversity anticlustering with regard to skewness and---in particular---kurtosis, even though the method has been specifically tailored to these criteria.
+It is maybe surprising that k-plus skew/kurtosis does not seem to have performed much better than diversity anticlustering with regard to skewness and kurtosis, even though the method has been specifically tailored to these criteria.
 Figure 2 gives a more fine grained assessment of this observation: Surprisingly, for small $N$, k-plus skew/kurtosis was even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
@@
 
-In the first example, I use norming data for the OASIS image data set that is freely available online [@kurdi2017introducing].
+In the first example, I used the OASIS norming data that is freely available online [@kurdi2017introducing].
 Kurdi et al.
 assembled 900 open-access color images and collected ratings on two affective dimensions: arousal and valence.
-@brielmann2019intense collected additional norming data by measuring how the same 900 images were rated with regard to their beauty.
-In my application, I used all three features; the corresponding data is available from Github.[^githubbeauty]
-
-[^githubbeauty]: https://github.com/aenneb/OASIS-beauty
-
-I used the function `kplus_anticlustering()` to divide the data set of 900 images into 9 groups of 100 images each.
+@brielmann2019intense measured how the same 900 images were rated with regard to beauty.
+I used the function `kplus_anticlustering()` to divide all images into 9 groups of 100 images each.
 The first argument of the function call is a variable named `features`, which contains the image data as a table: 900 rows representing images and 3 columns representing the features beauty, arousal, and valence.
@@
 kplus_anticlustering(
-  features, # data table with 900 rows and 3 features
+  features,
   K = 9,

### 2023-01-27 paper.Rmd
word count
typos

@@
     corresponding : yes
-    address       : "Heinrich Heine Universität Düsseldorf,  Institut für Experimentelle Psychologie, Universitätsstraße 1, 40225 Düsseldorf, Germany"
+    address       : "Heinrich-Heine-Universität Düsseldorf,  Institut für Experimentelle Psychologie, Universitätsstraße 1, 40225 Düsseldorf, Germany"
     email         : "martin.papenberg@hhu.de"
@@
 keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, stimulus selection"
-wordcount         : "5114"
+wordcount         : "4992"
 
@@
 
-Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to address all distribution characteristics.
+Generally, it should be noted that diversity anticlustering is a good all-rounder method that tends to address all distribution characteristics.
 For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering was outperformed.
@@
 Only the argument `variance` is set to `TRUE` by default---corresponding to the understanding that k-plus is an extension to k-means that at least equalizes variances in addition to means.
-The arguments corresponding to skewness, kurtosis and convariance have to be spefically "turned on", i.e., by setting them to `TRUE` in the function call of `kplus_anticlustering()`.
+The arguments corresponding to skewness, kurtosis and covariance have to be specifically "turned on", i.e., by setting them to `TRUE` in the function call of `kplus_anticlustering()`.
 If other higher order moments should be included as part of the optimization, the optional argument `moments` can be used to specify the desired moments as an integer vector.
@@
 by specifying that multiple distribution moments should equalized.
-If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variance).
+If the results of such a greedy setup are not satisfying, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variance).
 Using the open source R package `anticlust`, the (repeated) application of k-plus anticlustering is easily accessible and free.

### 2023-01-27 old_scripts/Example1.R
removed old before submission

@@ -1,88 +1,0 @@
-
-library(tidyr)
-library(dplyr)
-library(anticlust)
-
-oasis <- read.csv(
-  "https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv"
-)
-
-features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
-
-K <- 9
-
-# k-plus variance / skew
-start <- Sys.time()
-groups_kplus <- kplus_anticlustering(
-  features,
-  K = K,
-  variance = TRUE,
-  skew = TRUE,
-  method = "local-maximum",
-  standardize = TRUE
-)
-Sys.time() - start
-
-# k-means
-start <- Sys.time()
-groups_kmeans <- anticlustering(
-  features,
-  K = K,
-  objective = "variance",
-  method = "local-maximum",
-  standardize = TRUE
-)
-Sys.time() - start
-
-# diversity
-start <- Sys.time()
-groups_diversity <- anticlustering(
-  features,
-  K = K,
-  objective = "diversity",
-  method = "local-maximum",
-  standardize = TRUE
-)
-Sys.time() - start
-
-
-# Functions to compute means / SD / skew per group 
-descriptives_by_group <- function(features, anticlusters, FUN, name) {
-  df <- data.frame(
-    t(data.frame(lapply(by(features, anticlusters, FUN), c)))
-  )
-  df$group <- 1:nrow(df)
-  df$Descriptive = name
-  df
-}
-
-
-all_descriptives_by_group <- function(features, anticlusters) {
-  means <- descriptives_by_group(features, anticlusters, colMeans, "mean")
-  sds <- descriptives_by_group(
-    features, 
-    anticlusters, 
-    function(x) sapply(x, sd),
-    "sds"
-  )
-  skew <- descriptives_by_group(
-    features, 
-    anticlusters, 
-    function(x) sapply(x, DescTools::Skew),
-    "skew"
-  )
-  rbind(means, sds, skew)
-}
-
-
-kplus <- all_descriptives_by_group(features, groups_kplus)
-kplus <- data.frame(kplus, Objective = "kplus")
-
-kmeans <- all_descriptives_by_group(features, groups_kmeans)
-kmeans <- data.frame(kmeans, Objective = "kmeans")
-diversity <- all_descriptives_by_group(features, groups_diversity)
-diversity <- data.frame(diversity, Objective = "diversity")
-
[…]

### 2023-01-27 old_scripts/Examples/Example-1-OASIS.R

@@ -1,23 +1,0 @@
-
-library(anticlust)
-
-oasis <- read.csv("https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv")
-
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
-extended <- cbind(features, squared_from_mean(features))
-
-oasis$ac <- anticlustering(
-  scale(extended),
-  K = 10,
-  categories = oasis$Category, 
-  objective = "variance",
-  method = "local-maximum"
-)
-
-mean_sd_tab(features, oasis$ac)
-
-table(oasis$ac, oasis$Category)

### 2023-01-27 old_scripts/Examples/Example-2-Unequal-Sizes.R

@@ -1,24 +1,0 @@
-# Unequal sizes
-
-library(anticlust)
-
-features <- schaper2019[, 3:6]
-
-K <- 3
-init <- sample(rep(1:3, nrow(schaper2019) * c(1/4, 1/4, 1/2)))
-
-# Compute features for k-means-extended criterion
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-anticlusters <- anticlustering(
-  scale(cbind(features, squared_from_mean(features))),
-  K = init,
-  objective = "variance",
-  method = "local-maximum", 
-  repetitions = 30
-)
-
-mean_sd_tab(features, anticlusters)
-table(anticlusters)

### 2023-01-27 old_scripts/Examples/Example-3-Schaper-Covariances.R

@@ -1,82 +1,0 @@
-
-# Example 3
-
-library(anticlust)
-
-# Features for k-plus objective
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-# Features for k-covariances objective
-covariance_features <- function(data) {
-  M <- ncol(data)
-  if (M < 2) {
-    stop("k-covariances can only be used when at least two features are present.")
-  }
-  # number of features for equalizing covariances
-  feature_combinations <- t(combn(M, 2))
-  # store new features in matrix
-  cov_feature_matrix <- matrix(ncol = nrow(feature_combinations), nrow = nrow(data))
-  for (i in 1:nrow(feature_combinations)) {
-    f1 <- feature_combinations[i, 1]
-    f2 <- feature_combinations[i, 2]
-    cov_feature_matrix[, i] <- (data[, f1] - mean(data[, f1])) * (data[, f2] - mean(data[, f2]))
-  }
-  cov_feature_matrix
-}
-
-
-upper_tri <- function(x) {
-  dims <- dim(x)
-  mat <- rep(NA, prod(dims))
-  dim(mat) <- dims
-  mat[upper.tri(mat)] <- x[upper.tri(x)]
-  mat
-}
-
-lower_tri <- function(x) {
-  dims <- dim(x)
-  mat <- rep(NA, prod(dims))
-  dim(mat) <- dims
-  mat[lower.tri(mat)] <- x[lower.tri(x)]
-  mat
-}
-
-plus_na <- function(x, y) {
-  x[is.na(x)] <- 0
-  y[is.na(y)] <- 0
-  x + y
-}
-
-
-features <- schaper2019[, 3:6]
-
-extended <- cbind(features, squared_from_mean(features))
-extended_cov <- cbind(extended, covariance_features(features))
-
-ac_kplus <- anticlustering(
-  scale(extended),
-  K = 2,
-  objective = "variance",
-  method = "local-maximum",
-  repetitions = 10
-)
-
-ac_cov <- anticlustering(
-  scale(extended_cov),
-  K = 2,
-  objective = "variance",
-  method = "local-maximum",
-  repetitions = 10
-)
-
-cors_kplus <- lapply(by(features, ac_kplus, cor), function(x) round(x, 2))
[…]

### 2023-01-27 old_scripts/Examples/Example-4-OASIS-Skew_Kurtosis.R

@@ -1,44 +1,0 @@
-
-library(anticlust)
-library(DescTools) # for computing skew and kurtosis
-
-oasis <- read.csv("https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv")
-
-# function to compute features for variance
-# moment = 2 -> variance; 3 = skew; 4 = kurtosis
-moment_features <- function(data, moment = 2) {
-  apply(data, 2, function(x) (x - mean(x))^moment)
-}
-
-features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
-extended <- cbind(features, moment_features(features))
-extended_skew_kurtosis <- cbind(
-  extended, 
-  moment_features(features, 3),
-  moment_features(features, 4)
-)
-
-# standard k-plus anticlustering
-anticlusters_cov <- anticlustering(
-  scale(extended),
-  K = 3,
-  objective = "variance"
-)
-
-# k-plus anticlustering, also considering skew and curtosis
-anticlusters_skew_kurtosis <- anticlustering(
-  scale(extended_skew_kurtosis),
-  K = 3,
-  objective = "variance"
-)
-
-
-by(features, anticlusters_skew_kurtosis, function(x) round(apply(x, 2, Skew), 2))
-by(features, anticlusters_cov, function(x) round(apply(x, 2, Skew), 2))
-
-by(features, anticlusters_skew_kurtosis, function(x) round(apply(x, 2, Kurt), 2))
-by(features, anticlusters_cov, function(x) round(apply(x, 2, Kurt), 2))
-
-mean_sd_tab(features, anticlusters_skew_kurtosis)
-mean_sd_tab(features, anticlusters_cov)
-

### 2023-01-27 old_scripts/K-Means-Vs-Difference.R

@@ -1,128 +1,0 @@
-
-# Investigate why k-variance is worse than k-means-variance for 
-# optimizing between-group similarity in variance
-
-# Reason: I validated the results using difference between minimum and
-# maximum *standard deviation* across groups. For K > 2, this is not 
-# equivalent to minimum and maximum *variance* across groups (which
-# is however minimized using the k-variance criterion!)
-
-library(anticlust)
-source("exchange_method.R")
-source("../misc-functions.R")
-
-N <- 15
-M <- 3
-K <- 3
-features <- matrix(rnorm(N * M), ncol = M)
-all_partitions <- generate_partitions(N, K)
-
-all_objectives_var <- sapply(
-  all_partitions,
-  FUN = sum_group_diff_min_max,
-  features = features,
-  fun = var
-)
-
-all_objectives_mean <- sapply(
-  all_partitions,
-  FUN = sum_group_diff_min_max,
-  features = features,
-  fun = mean
-)
-
-all_objectives_kVariances <- sapply(
-  all_partitions,
-  FUN = var_objective,
-  features = squared_from_mean(features)
-)
-
-all_objectives_kMeans <- sapply(
-  all_partitions,
-  FUN = var_objective,
-  features = features
-)
-
-all_objectives_kExtended <- sapply(
-  all_partitions,
-  FUN = var_objective,
-  features = cbind(squared_from_mean(features), features)
-)
-
-df <- data.frame(
-  kVar = all_objectives_kVariances,
-  diff_var  = all_objectives_var,
-  kMeans = all_objectives_kMeans,
-  diff_means  = all_objectives_mean,
-  kExt = all_objectives_kExtended
-)
-
-#plot(df, col = "darkgrey", las = 1, pch = 4, cex = 0.8)
-
-
-# Initialize a partition
-init <- sample(rep_len(1:K, N))
-init <- anticlust:::order_cluster_vector(init)
-
-# Do local search for three objectives
-partitions_means <- local_maximum_exchange(
-  features, 
-  init,
-  anticlust::variance_objective
-)
-partitions_variance <- local_maximum_exchange(
-  squared_from_mean(features), 
-  init,
-  anticlust::variance_objective
-)
-partitions_means_variance <- local_maximum_exchange(
-  cbind(features, squared_from_mean(features)), 
-  init,
-  anticlust::variance_objective
-)
-
[…]

### 2023-01-27 old_scripts/diff-min-max-mean-variance.R

@@ -1,129 +1,0 @@
-
-# Investigate why k-variance is worse than k-means-variance for 
-# optimizing between-group similarity in variance
-
-# Reason: I validated the results using difference between minimum and
-# maximum *standard deviation* across groups. For K > 2, this is not 
-# equivalent to minimum and maximum *variance* across groups (which
-# is however minimized using the k-variance criterion!)
-
-library(anticlust)
-source("exchange_method.R")
-source("../misc-functions.R")
-
-N <- 200
-M <- 4
-K <- 4
-features <- matrix(rnorm(N * M), ncol = M)
-dist_features <- squared_from_mean(features)
-combined_features <- cbind(features, dist_features)
-
-# Initialize a partition
-init <- sample(rep_len(1:K, N))
-init <- order_cluster_vector(init)
-
-# Do local search for three objectives
-partitions_means <- local_maximum_exchange(
-  features, 
-  init
-)
-partitions_variance <- local_maximum_exchange(
-  dist_features, 
-  init
-)
-partitions_means_variance <- local_maximum_exchange(
-  combined_features, 
-  init
-)
-
-objectives <- list()
-objectives[["diff_mean"]] <- list()
-objectives[["diff_var"]] <- list()
-
-funs <- list(diff_mean = mean, diff_var = var)
-
-for (i in c("diff_mean", "diff_var")) {
-  objectives[[i]][["kMeans"]] <- sapply(
-    partitions_means,
-    sum_group_diff_min_max,
-    features = features,
-    fun = funs[[i]]
-  )
-  objectives[[i]][["kVar"]] <- sapply(
-    partitions_variance,
-    sum_group_diff_min_max,
-    features = features,
-    fun = funs[[i]]
-  )
-  objectives[[i]][["kExt"]] <- sapply(
-    partitions_means_variance,
-    sum_group_diff_min_max,
-    features = features,
-    fun = funs[[i]]
-  )
-}
-
-
-
-sample_point <- function(clusters, features) {
-  cl <- sample(clusters)
-  c(
-    mean = sum_group_diff_min_max(cl, features, fun = mean),
-    var = sum_group_diff_min_max(cl, features, fun = var)
-  )
-}
-
[…]

### 2023-01-27 old_scripts/equivalent-objectives.R

@@ -1,28 +1,0 @@
-
-# Test equivalence of objectives
-
-# Check that I can actually just append the squared difference data as 
-# additional columns to the original data frame:
-
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-
-N <- 500
-M <- 20
-data <- matrix(rnorm(N * M), ncol = M)
-var_data <- squared_from_mean(data)
-
-K <- 10
-clusters <- sample(rep_len(1:K, N))
-
-obj1 <- variance_objective(data, clusters) + variance_objective(var_data, clusters)
-obj2 <- variance_objective(cbind(data, var_data), clusters)
-
-tolerance <- 1e-10
-
-obj1 - obj2 < tolerance
-
-obj1
-obj2

### 2023-01-27 old_scripts/exchange_method.R

@@ -1,218 +1,0 @@
-
-# Call the exchange method below until a local maximum is reached
-local_maximum_exchange <- function(features, init) {
-  best_partition <- init
-  found_partitions <- list(best_partition)
-  new_obj <- variance_objective(features, best_partition)
-  old_obj <- -Inf
-  while (new_obj > old_obj) {
-    partitions <- exchange_method(
-      features,
-      clusters = best_partition
-    )
-    best_partition <- partitions[[length(partitions)]]
-    old_obj <- new_obj
-    new_obj <- anticlust::variance_objective(features, best_partition)
-    found_partitions <- c(found_partitions, partitions)
-  }
-  found_partitions[!duplicated(found_partitions)]
-} 
-
-
-#' Solve k-means anticlustering using the modified exchange method
-#'
-#' @param data the data -- an N x M table of item features
-#' @param clusters An initial cluster assignment
-#'
-#' @return The anticluster assignment
-#'
-#' @noRd
-#'
-
-exchange_method <- function(data, clusters) {
-  N <- nrow(data)
-  best_total <- variance_objective_(clusters, data)
-  centers <- cluster_centers(data, clusters)
-  distances <- dist_from_centers(data, centers, squared = TRUE)
-  all_partitions <- list(clusters)
-  u <- 2
-  
-  ## frequencies of each cluster are required for updating cluster centers:
-  tab <- c(table(clusters))
-  for (i in 1:N) {
-    # cluster of current item
-    cluster_i <- clusters[i]
-    # get exchange partners for item i
-    exchange_partners <- 1:N
-    # exchange partners are not in the same cluster:
-    exchange_partners <- exchange_partners[clusters != clusters[i]]
-    # Sometimes an exchange cannot take place
-    if (length(exchange_partners) == 0) {
-      next
-    }
-    # container to store objectives associated with each exchange of item i:
-    comparison_objectives <- rep(NA, length(exchange_partners))
-    for (j in seq_along(exchange_partners)) {
-      ## Swap item i with all legal exchange partners and check out objective
-      # (a) Determine clusters of to-be-swapped elements
-      tmp_clusters <- clusters
-      tmp_swap <- exchange_partners[j]
-      cluster_j <- tmp_clusters[tmp_swap]
-      # (b) Swap the elements
-      tmp_clusters[i] <- cluster_j
-      tmp_clusters[tmp_swap] <- cluster_i
-      # (c) Update cluster centers after swap
[…]

### 2023-01-27 old_scripts/improved_objective.R

@@ -1,65 +1,0 @@
-
-library(anticlust)
-
-N <- 150
-M <- 2
-K <- 3
-features <- matrix(rnorm(N * M), ncol = M)
-#features <- schaper2019[, 3:6]
-
-nrep <- 30
-
-ac_kmeans <- anticlustering(
-  features,
-  K = K,
-  objective = "variance",
-  method = "local-maximum",
-  repetitions = nrep
-)
-
-ac_distance <- anticlustering(
-  features,
-  K = K,
-  objective = "distance",
-  method = "local-maximum",
-  repetitions = nrep,
-)
-
-ac_kmeans_var_only <- anticlustering(
-  squared_from_mean(features),
-  K = K,
-  objective = "variance",
-  method = "local-maximum",
-  repetitions = nrep
-)
-
-ac_kmeans_var <- anticlustering(
-  cbind(features, squared_from_mean(features)),
-  K = K,
-  objective = "variance",
-  method = "local-maximum",
-  repetitions = nrep
-)
-
-cat("Standard K-Means \n")
-print(mean_sd_tab(features, ac_kmeans, return_diff = TRUE))
-cat("K-Means + Squared Diff\n")
-print(mean_sd_tab(features, ac_kmeans_var, return_diff = TRUE))
-#cat("Anticluster Editing \n")
-#print(mean_sd_tab(features, ac_distance, return_diff = TRUE))
-cat("Only Squared Diff\n")
-print(mean_sd_tab(features, ac_kmeans_var_only, return_diff = TRUE))
-
-variance_objective(squared_from_mean(features), ac_kmeans_var) / 
-  variance_objective(squared_from_mean(features), ac_kmeans_var_only)
-
-variance_objective(features, ac_kmeans_var) / variance_objective(features, ac_kmeans)
-
-
-# WHY IS K-MEANS EXTENDED BETTER AT MINIMIZING MAXIMUM DIFFERENCE IN 
-# VARIANCE THAN K-VARIANCE ALONE? NOT SURE YET. (and this is only the 
-# case for K > 2)
-sum(group_diff_min_max(ac_kmeans, features, var))
-sum(group_diff_min_max(ac_kmeans_var, features, var))
-sum(group_diff_min_max(ac_kmeans_var_only, features, var))
-

### 2023-01-27 old_scripts/k_covariances_first_test.R

@@ -1,105 +1,0 @@
-library(anticlust)
-
-# Compute features for k-means-extended criterion
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-# Difference between minimum and maximum in a vector
-diff_min_max <- function(x) {
-  diff(range(x))
-}
-
-
-# Compute difference between a correlations in two groups
-diff_group_cors <- function(features, clusters) {
-  diff_min_max(by(features, clusters, function(x) cor(x[, 1], x[, 2])))
-}
-
-source("./Simulation_Study/0-functions-analyze-data.R")
-
-paired_data <- function(n, mx = 0, my = 0, sdx = 1, sdy = 1, r, empirical = FALSE) {
-  cor_matrix <- matrix(c(1, r, r, 1), ncol = 2)
-  sds <- c(sdx, sdy)
-  vars <- sds %*% t(sds)
-  cov_matrix <- vars * cor_matrix
-  data <- MASS::mvrnorm(
-    n, 
-    mu = c(mx, my), 
-    Sigma = cov_matrix, 
-    empirical = empirical
-  ) 
-  data <- data.frame(data)
-  colnames(data) <- c("x", "y")
-  data
-}
-
-sim_kcov <- function(X, N, r, K) {
-
-    K <- K
-
-    features <- as.matrix(paired_data(n = N, r = r))
-
-    anticlusters1 <- anticlustering(
-        features,
-        objective = "variance",
-        K = K
-    )
-
-    anticlusters2 <- anticlustering(
-        features,
-        objective = "kplus",
-        K = K
-    )
-
-    anticlusters3 <- anticlustering(
-        features,
-        objective = "diversity",
-        K = K
-    )
-    
-    # optimize covariance similarity
-    cov_features <- (features[, 1] - mean(features[,1])) * (features[, 2] - mean(features[,2]))
-    var_features <- squared_from_mean(features)
-    anticlusters4 <- anticlustering(
-        scale(cbind(features, var_features, cov_features)),
-        objective = "variance",
-        K = K
-    )
-    
-    rnd_anticlusters <- sample(anticlusters1)
-    
-    results <- matrix(
-            c(c(
-                random = var_means(rnd_anticlusters, features),
-                kmeans = var_means(anticlusters1, features),
-                kplus = var_means(anticlusters2, features),
-                diversity = var_means(anticlusters3, features),
-                kcov = var_means(anticlusters4, features)
-            ),
-            c(
[…]

### 2023-01-27 old_scripts/misc-functions.R

@@ -1,81 +1,0 @@
-
-# Compute features for k-means-extended criterion
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-# Get the last element of a list / vector
-last <- function(x) {
-  x[[length(x)]]
-}
-
-# Sum of difference between minimum and maximum value per feature of a 
-# data frame, by a group variable (i.e., total deviation in variances,
-# measured by difference between min/max variance, across several features)
-sum_group_diff_min_max <- function(clusters, features, fun) {
-  sum(group_diff_min_max(clusters, features, fun))
-}
-
-# Difference between minimum and maximum value per feature of a data frame,
-# by a group variable
-group_diff_min_max <- function(clusters, features, fun) {
-  apply(fun_by_group(features, clusters, fun), 1, diff_min_max)
-}
-
-# Function that computes a function for each feature of a data frame,
-# by a grouping variable
-fun_by_group <- function(features, clusters, fun) {
-  data.frame(lapply(by(features, clusters, function(x) apply(x, 2, fun)), c))
-}
-
-# Difference between minimum and maximum in a vector
-diff_min_max <- function(x) {
-  diff(range(x))
-}
-
-# short-cut function to determine difference in minimum and maximum
-# variance, across groups, for one feature
-diff_vars <- function(clusters, features) {
-  abs(diff(range(tapply(features, clusters, var))))
-}
-
-# short-cut function to determine difference in minimum and maximum
-# means, across groups, for one feature
-diff_mean <- function(clusters, features) {
-  abs(diff(range(tapply(features, clusters, mean))))
-}
-
-# Variance objective function with changed order of the arguments 
-# (can be called by sapply() etc.)
-var_objective <- function(clusters, features) {
-  anticlust::variance_objective(features, clusters)
-}
-
-
-# For a data frame (rows = partitions, columns = objectives), determine which 
-# elements are part of the Pareto efficient set. Output is a logical vector,
-# illustrating for each row if it is part of the Pareto efficient set.
-# `sense` can be "max" or "min", depending on whether high objective values
-# are good or low objective values. 
-pareto_set <- function(df, sense = "max") {
-  if (sense == "min") {
-    df <- -df
-  }
[…]

### 2023-01-27 old_scripts/pareto_front.R

@@ -1,113 +1,0 @@
-
-# Investigate why k-variance is worse than k-means-variance for 
-# optimizing between-group similarity in variance
-
-library(anticlust)
-source("../misc-functions.R")
-
-N <- 12
-K <- 3
-M <- 2
-all_partitions <- generate_partitions(N, K)
-features <- matrix(rnorm(N * M), ncol = M)
-
-all_objectives_mean <- sapply(
-  all_partitions,
-  FUN = var_objective,
-  features = features
-)
-
-all_objectives_var <- sapply(
-  all_partitions,
-  FUN = var_objective,
-  features = squared_from_mean(features)
-)
-
-df <- data.frame(
-  kMeans = all_objectives_mean,
-  kVariance  = all_objectives_var
-)
-
-# Compute Pareto efficient set!
-
-# For a data frame (rows = partitions, columns = objectives), determine which 
-# elements are part of the Pareto efficient set. Output is a logical vector,
-# illustrating for each row if it is part of the Pareto efficient set.
-
-pareto_set <- function(df) {
-  is_dominated <- rep(FALSE, nrow(df)) 
-  for (i in 1:nrow(df)) {
-    for (j in 1:nrow(df)) {
-      j_dominates_i <- all(df[i, ] <= df[j, ]) && any(df[i, ] < df[j, ])
-      if (j_dominates_i) {
-        is_dominated[i] <- TRUE
-        break
-      }
-    }
-  }
-  !is_dominated
-}
-
-# Try some speed optimization
-pareto_set2 <- function(df) {
-  df <- data.frame(df)
-  ncrits <- ncol(df)
-  df$ID <- 1:nrow(df)
-  df <- df[order(rowSums(df[, 1:ncrits]), decreasing = TRUE), ]
-  
-  is_dominated <- rep(FALSE, nrow(df)) 
-  for (i in nrow(df):1) {
-    for (j in 1:nrow(df)) {
-      j_dominates_i <- all(df[i, 1:ncrits] <= df[j, 1:ncrits]) && 
-                       any(df[i, 1:ncrits] < df[j, 1:ncrits])
-      if (j_dominates_i) {
-        is_dominated[df$ID[i]] <- TRUE
-        break
-      }
-    }
-  }
-  !is_dominated
-}
-
-#start <- Sys.time()
-#efficient_set2 <- pareto_set2(df)
-#Sys.time() - start 
-
-#start <- Sys.time()
[…]

### 2023-01-27 old_scripts/solution-structure-large-data.R

@@ -1,83 +1,0 @@
-
-# Investigate why k-variance is worse than k-means-variance for 
-# optimizing between-group similarity in variance
-
-source("exchange_method.R")
-source("../misc-functions.R")
-
-
-library(anticlust)
-library(tidyverse)
-
-N <- 100
-M <- 4
-K <- 4
-features <- matrix(rnorm(N * M), ncol = M)
-dist_features <- squared_from_mean(features)
-combined_features <- cbind(features, dist_features)
-
-# Initialize a partition
-init <- sample(rep_len(1:K, N))
-init <- order_cluster_vector(init)
-
-# Do local search for three objectives
-partitions_means <- local_maximum_exchange(
-  features, 
-  init
-)
-partitions_variance <- local_maximum_exchange(
-  dist_features, 
-  init
-)
-partitions_means_variance <- local_maximum_exchange(
-  combined_features, 
-  init
-)
-
-partitions <- list(
-  kMeans = partitions_means, 
-  kVariance = partitions_variance, 
-  kExtended = partitions_means_variance
-)
-all_features <- list(
-  kMeans = features, 
-  kVariance = dist_features
-)
-
-results_list <- list()
-# iterate over data and objectives
-for (i in names(partitions)) {
-  iteration <- 1:length(partitions[[i]])
-  for (j in names(all_features)) {
-    objectives <- sapply(partitions[[i]], var_objective, features = all_features[[j]])
-    results_list[[paste0(i, "-", j)]] <- data.frame(
-      objectives, method = i, objective = j, iteration = iteration
-    )
-  }
-}
-results <- do.call(rbind, results_list)
-rownames(results) <- NULL
-
-results <- pivot_wider(results, names_from = objective, values_from = objectives)
-
-
-# Plot data
-ggplot(results, aes(x = kMeans, y = kVariance, group = method)) +
-  geom_path(aes(col = method)) + 
-  geom_point(aes(col = method)) + 
-  geom_point(aes(x = kMeans[1], y = kVariance[1]))
-
-options(pillar.sigfig = 6)
-
-# K-Variance better optimizes k-variance criterion ....
-results %>% 
-  group_by(method) %>% 
-  summarize(
-    kVarianceObj = max(kVariance),
-    kMeansObj = max(kMeans))
-
-# ... but is worse at minimizing the difference in range in variance,
-# compared to k-extended  
-sum(group_diff_min_max(last(partitions_means), features, var))
[…]

### 2023-01-27 old_scripts/solution-structure.R

@@ -1,96 +1,0 @@
-
-# Investigate why k-variance is worse than k-means-variance for 
-# optimizing between-group similarity in variance
-
-source("exchange_method.R")
-
-library(anticlust)
-
-N <- 16
-K <- 2
-all_partitions <- generate_partitions(N, K)
-features <- as.matrix(rnorm(N))
-
-## Create an objective function that takes the partition
-## as first argument (then, we can use sapply to compute
-## the objective for each partition)
-diff_mean <- function(clusters, features) {
-  abs(diff(range(tapply(features, clusters, mean))))
-}
-
-diff_vars <- function(clusters, features) {
-  abs(diff(range(tapply(features, clusters, var))))
-}
-
-all_objectives_mean <- sapply(
-  all_partitions,
-  FUN = diff_mean,
-  features = features
-)
-
-all_objectives_var <- sapply(
-  all_partitions,
-  FUN = diff_vars,
-  features = features
-)
-
-df <- data.frame(
-  diff_mean = all_objectives_mean,
-  diff_var  = all_objectives_var
-)
-
-# exchange method for finding all objectives
-
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-# Initialize a partition
-init <- sample(rep_len(1:K, N))
-init <- anticlust:::order_cluster_vector(init)
-
-# Do local search for three objectives
-partitions_means <- local_maximum_exchange(
-  features, 
-  init,
-  anticlust::variance_objective
-)
-partitions_variance <- local_maximum_exchange(
-  squared_from_mean(features), 
-  init,
-  anticlust::variance_objective
-)
-partitions_means_variance <- local_maximum_exchange(
-  cbind(features, squared_from_mean(features)), 
-  init,
-  anticlust::variance_objective
-)
-
-plot(df, col = "darkgrey", las = 1, pch = 4, cex = 0.8)
-points(
-  df[match(partitions_means, all_partitions), ], 
-  col = "#28E2E5",
-  pch = 19, 
-  cex = 2,
-  type = "b",
-  lwd = 2
-)
-
-points(
-  df[match(partitions_variance, all_partitions), ], 
-  col = "#2297E6",
-  pch = 18, 
-  cex = 2,
-  type = "b",
-  lwd = 2
-)
-
-points(
-  df[match(partitions_means_variance, all_partitions), ], 
[…]

### 2023-01-27 render.sh

@@ -1,3 +1,0 @@
-#!/bin/sh
-
-R -e 'rmarkdown::render("paper.Rmd")'

### 2023-04-17 paper.Rmd
typo
working on problem formalization for revision

@@
   - \usepackage{float}
+  - \usepackage{amsmath}
   - \floatstyle{plaintop}
@@
 
-## Quantifying differences in variance
+## Minimizing discrepancy in variance between groups
 
-In the following, I present a variation of the $\mathit{SSE}$ that quantifies how the variance of numeric data differs between groups.
+In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variance between groups instead of only mean values.
 To motivate the new criterion, we first consider a set of one-dimensional observations.
@@
 This constitutes the central insight motivating the new k-plus criterion.
-Thus, to obtain an adaptation of the $\mathit{SSE}$ that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
+
+Thus, to obtain an adaptation of the $\mathit{SSE}$ that can be maximized in order to minimize differences in variances between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
-On the basis of $\mathbf{Y}_{N \times P}$, we define the criterion $\mathit{SSE_{Var}}$, reflecting how strongly the variances of all attributes vary across the $K$ clusters:
+On the basis of $\mathbf{Y}_{N \times P}$, we define the criterion $\mathit{SSE_{Var}}$:
 
@@
 However, the input data reflects the squared difference between each data point and the attribute mean instead of the raw data points.
-Maximizing $\mathit{SSE_{Var}}$ will lead to similar variances across the $K$ groups in the $P$ attributes.
+Maximizing $\mathit{SSE_{Var}}$ will lead to similar variances across the $K$ groups in the $P$ attributes (**only if the group means are close to the overall mean, discuss this as suggested in reviewer 1's comment 4**; $\mathit{SSE_{Var}}$ is only valid if the group means are close to the overall means of the data.).
 
-## A bicriterion model: K-plus anticlustering {#kplus}
+## A multicriterion optimization model: K-plus anticlustering {#kplus}
 
-It is unlikely that an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is of interest without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
-A combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, however, can be employed to simultaneously optimize both criteria, which is a reasonable idea when striving for overall between-group similarity.
-The most simple approach for such a bicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
-That is, both criteria are computed independently from each other and are then combined into a single criterion through a weighted sum:
+As discussed (**TODO**) in the previous section, an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is not feasible without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
+Moreover, only equalizing variances between groups is most likely not even of interest in most applications.
+Hence, when striving for similarity in group variances by maximizing $\mathit{SSE_{Var}}$, it is at the same time necessary and worthwhile to strive for similarity in means, thus to maximize $\mathit{SSE}$ at the same time.
+Hence, a combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ should be employed to simultaneously optimize both criteria, which also represents a reasonable idea when striving for overall between-group similarity.
+The most simple approach for such a multicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
+That is, all criteria are computed independently from each other and are then combined into a single criterion through a weighted sum
 
@@
 $$
-
-The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
-Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
-This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining a new data matrix $\mathbf{Z}_{N \times 2P} = \{z_{ij}\}_{N \times 2P}$ as the augmentation of $\mathbf{X}_{N \times P}$ with $\mathbf{Y}_{N \times P}$, i.e.
-
-$$
-  \mathbf{Z}_{N \times 2P} =
-  \left[ {\begin{array}{cccc|cccc}
-    x_{11} & x_{12} & \cdots & x_{1P} & y_{11} & y_{12} & \cdots & y_{1P}\\
-    x_{21} & x_{22} & \cdots & x_{2P} & y_{21} & y_{22} & \cdots & y_{2P}\\
-    \vdots & \vdots & \ddots & \vdots & \vdots & \vdots & \ddots & \vdots\\
-    x_{N1} & x_{N2} & \cdots & x_{NP} & y_{N1} & y_{N2} & \cdots & y_{NP}\\
-  \end{array} } \right]
-$$
-
-and $z_{i.} = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, y_{iP})$, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
-
-$$
-\mathit{SSE_{kplus}} =
-  \sum\limits_{j=1}^{2 P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (z_{ij} - \overline{z}_j ^{(k)})^2.
-$$
-
-This term again has the same form as $\mathit{SSE}$ but is computed on different data.
-Hence, computing the k-plus criterion can be done by first computing the squared distances $y_{ij}$, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
-Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
-
-Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when computing $\mathit{SSE_{kplus}}$.
-Because the computation of the new attribute values $y_{ij}$ involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$ (e.g., see Figure 1).
-This imbalance can strongly and unpredictably influence the relative weight of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ when computing $\mathit{SSE_{kplus}}$.
-To ensure an equal weight of both criteria, all attributes ($x_{ij}$ and $y_{ij}$) can be standardized before computing $\mathit{SSE_{kplus}}$, e.g.
-via $z$-standardization.
+where the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
 
@@
     FUN = var_objective,
-    features = scale(cbind(features, squared_distances_from_mean))
+    features = cbind(features, squared_distances_from_mean)
   )
@@
 
-Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria at the same time [@ferligoj1992direct].
-It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ can yield a partitioning where both constituting criteria (mean and variance) are similar between groups.
+Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria at the same time if the underlying objectives are in conflict with each other [@ferligoj1992direct].
+It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ (**TODO undefined**) can yield a partitioning where both constituting criteria (mean and variance) are similar between groups, because it is unknown if these objectives tend to align with each other or are in conflict.
 A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible.
@@
 Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances.
-The diamond symbol represents the partition that maximizes $\mathit{SSE_{kplus}}$, which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
-Thus, using the unweighted sum approach, the optimal partition according to $\mathit{SSE_{kplus}}$ fulfills the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ to an astonishing degree.
+The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ and $w_2$ were simply set to 1), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
+Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree, indicating that these objectives tend to align and are not in conflict with each other.
 
@@
 
-## Skewness, kurtosis, and higher order moments
+## Higher order moments
 
-The k-plus approach to equalizing the variance can be generalized to higher order moments. The *j*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ can be computed as
+Note that the k-plus logic to equalizing group variances as well as group means can be seamlessly generalized to higher order moments.
+The *t*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ can be computed as
 
 $$
-\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^j.
+\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^t.
 $$
 
-where $j = 2$ gives the variance, $j = 3$ the skewness, and $j = 4$ the kurtosis [@joanes1998].
-Thus, by changing $j$---the power of the deviation between mean and data points---we can use the k-plus logic to equalize any desired sample moment between groups.
-For each moment $j$ and for each variable $x$, this is accomplished by adding a new variable $y^{(j)} = (y^{(j)}_1, \ldots, y^{(j)}_N)$ with $y^{(j)}_i = (x_i - \overline{\mathbf{x}})^j$ to the data set and subsequently applying standard k-means anticlustering.
-I expect this to be particularly interesting for the skewness and kurtosis when aiming to minimize differences with regard to distribution asymmetry (skewness) and the propensity to include outliers (kurtosis).
+where $t = 2$ gives the variance, $t = 3$ the skewness, and $t = 4$ the kurtosis [@joanes1998].
+Thus, by changing $t$---the power of the deviation between mean and data points---we can use the k-plus logic to equalize any desired sample moment between groups.
+
+Next, I provide some notation needed to compute the general $\mathit{SSE}$ for k-plus anticlustering, which may include arbitrary distribution moments as part of the computation.
+For each moment $t$ and for column of the original data matrix $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ ($j = 1, \ldots, P$) with mean $\overline{\mathbf{x}}_j$, a new variable $x_{t.j} = (x_{t1j}, \ldots, x_{tNj})$ is computed where
+
+$$
+    x_{tij} = 
+\begin{cases}
+    x_{ij} & \text{if } t = 1\\
+    (x_{ij} - \overline{\mathbf{x}}_j)^t, & \text{if } t > 1.
+\end{cases}
+$$
+
+This computation results in an augmented data matrix $\mathbf{\hat{X}}_{N \times (T \cdot P)} = \{x_{tij}\}_{N \times (T \cdot P)}$ where $T$ is the number of distribution moments that should be equalized between groups.
+Note that when defining $x_{tij}$, the condition $t = 1$ is included to keep the unchanged original data as the first $P$ columns in the data, which is used to equalize the the mean---the first moment---between groups.
+Using the thus defined data matrix $\mathbf{\hat{X}}_{N \times (T \cdot P)}$, the generalized multicriterion for k-plus anticlustering can be formalized by the formula
+
+$$
+\mathit{SSE_{kplus}} =
+  \sum\limits_{t=1}^{T} 
+  w_t \left(
+  \sum\limits_{j=1}^{P} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (x_{tij} - \overline{x}_{.j} ^{(k)})^2
+  \right)
+$$
+
+where $w_t$ is the weight associated with the $t'th$ moment and $\overline{x}_{.j} ^{(k)}$ is the mean of the $j'th$ variable in the $k'th$ group.
+Maximizing $\mathit{SSE_{kplus}}$ with a given number of moments $T$ can be used to equalize these distribution moments between $K$ groups.
+Note that according to this definition the inclusion of a distribution moment $t$ always implies that all moments $z < t$ are also included as part of the criterion $\mathit{SSE_{kplus}}$; it is not deemed worthwhile to exclude lower order moments in favor of higher order moments.
+In particular, the first moment, represented by the original data, should always be included because otherwise the criterion $\mathit{SSE_{kplus}}$ is not valid for higher order moments.
+
+**TODO: include discussion of weight selection, as suggested by reviewer 2**
+
+## The unweighted sum approach for k-plus anticlustering
+
+**argue why unweighted standardization k-plus anticlustering is good, as demanded by reviewer 2; also maybe include small scale simulation study**
+
+Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
+This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining a new data matrix $\mathbf{Z}_{N \times 2P} = \{z_{ij}\}_{N \times 2P}$ as the augmentation of $\mathbf{X}_{N \times P}$ with $\mathbf{Y}_{N \times P}$, i.e.
+
+$$
+  \mathbf{Z}_{N \times 2P} =
+  \left[ {\begin{array}{cccc|cccc}
+    x_{11} & x_{12} & \cdots & x_{1P} & y_{11} & y_{12} & \cdots & y_{1P}\\
+    x_{21} & x_{22} & \cdots & x_{2P} & y_{21} & y_{22} & \cdots & y_{2P}\\
+    \vdots & \vdots & \ddots & \vdots & \vdots & \vdots & \ddots & \vdots\\
+    x_{N1} & x_{N2} & \cdots & x_{NP} & y_{N1} & y_{N2} & \cdots & y_{NP}\\
+  \end{array} } \right]
+$$
+
+and $z_{i.} = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, y_{iP})$, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
+
+$$
+\mathit{SSE_{kplus}} =
+  \sum\limits_{j=1}^{2 P} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (z_{ij} - \overline{z}_j ^{(k)})^2.
+$$
+
+This term again has the same form as $\mathit{SSE}$ but is computed on different data.
+Hence, computing the k-plus criterion can be done by first computing the squared distances $y_{ij}$, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
+Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
+
+Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when computing $\mathit{SSE_{kplus}}$.
+Because the computation of the new attribute values $y_{ij}$ involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$ (e.g., see Figure 1).
+This imbalance can strongly and unpredictably influence the relative weight of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ when computing $\mathit{SSE_{kplus}}$.
+To ensure an equal weight of both criteria, all attributes ($x_{ij}$ and $y_{ij}$) can be standardized before computing $\mathit{SSE_{kplus}}$, e.g.
+via $z$-standardization.
 
@@
 For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering was outperformed.
-As as side note, apart from k-plus correlation, diversity anticlustering was the only method that equalized correlation structure among groups.
+As a side note, apart from k-plus correlation, diversity anticlustering was the only method that equalized correlation structure among groups.
 To the best of my knowledge, this observation has not been made previously, even though the maximum diversity problem has been widely studied.

### 2023-04-17 paper.Rmd
problem formalization

@@
 $$
+\mathit{SSE_{kplus}} = 
 w_1 \sum\limits_{j=1}^{P} 
@@
 where the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
+Maximizing $\mathit{SSE_{kplus}}$ for a given $K$ will lead to similar means as well as variances between groups.
 
@@
 Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria at the same time if the underlying objectives are in conflict with each other [@ferligoj1992direct].
-It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ (**TODO undefined**) can yield a partitioning where both constituting criteria (mean and variance) are similar between groups, because it is unknown if these objectives tend to align with each other or are in conflict.
+It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ can yield a partitioning where both constituting criteria (mean and variance) are similar between groups, because it is unknown if these objectives tend to align with each other or are in conflict.
 A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible.
@@
 
-## Higher order moments
+## Including higher order moments
 
@@
 $$
-\mathit{SSE_{kplus}} =
+\mathit{SSE_{kplusGeneralized}} =
   \sum\limits_{t=1}^{T} 
@@
 where $w_t$ is the weight associated with the $t'th$ moment and $\overline{x}_{.j} ^{(k)}$ is the mean of the $j'th$ variable in the $k'th$ group.
-Maximizing $\mathit{SSE_{kplus}}$ with a given number of moments $T$ can be used to equalize these distribution moments between $K$ groups.
-Note that according to this definition the inclusion of a distribution moment $t$ always implies that all moments $z < t$ are also included as part of the criterion $\mathit{SSE_{kplus}}$; it is not deemed worthwhile to exclude lower order moments in favor of higher order moments.
-In particular, the first moment, represented by the original data, should always be included because otherwise the criterion $\mathit{SSE_{kplus}}$ is not valid for higher order moments.
+Maximizing $\mathit{SSE_{kplusGeneralized}}$ with a given number of moments $T$ can be used to equalize these distribution moments between $K$ groups.
+Note that according to this definition the inclusion of a distribution moment $t$ always implies that all moments $z < t$ are also included as part of the criterion $\mathit{SSE_{kplusGeneralized}}$; it is not deemed worthwhile to exclude lower order moments in favor of higher order moments.
+In particular, the first moment, represented by the original data, should always be included because otherwise the criterion $\mathit{SSE_{kplusGeneralized}}$ is not valid for higher order moments.
 

### 2023-04-17 paper.Rmd
continued

@@
 Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances.
-The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ and $w_2$ were simply set to 1), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
+The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ and $w_2$ were set to 1 when computing $\mathit{SSE_{kplus}}$), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
 Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree, indicating that these objectives tend to align and are not in conflict with each other.

### 2023-04-18 paper.Rmd
more work on formalization

@@
 However, the input data reflects the squared difference between each data point and the attribute mean instead of the raw data points.
-Maximizing $\mathit{SSE_{Var}}$ will lead to similar variances across the $K$ groups in the $P$ attributes (**only if the group means are close to the overall mean, discuss this as suggested in reviewer 1's comment 4**; $\mathit{SSE_{Var}}$ is only valid if the group means are close to the overall means of the data.).
+Maximizing $\mathit{SSE_{Var}}$ can be used to create groups that have similar variances across the $P$ attributes.
+As a caveat, it is crucial to note that maximizing $\mathit{SSE_{Var}}$ only reliably yields similar group variances if the group centroids $\overline{\mathbf{x}}^{(k)}$, ($k = 1, \dots, K$) are close to the overall data centroid $\overline{\mathbf{x}}$.
+To reiterate, this condition has to be met for the original data $\mathbf{X}_{N \times P}$.
+Thus, only considering $\mathbf{Y}_{N \times P}$ in order to minimize discrepancies in variance may not yield satisfying results; instead, it has to be ensured that the group means of the original data are also similar between groups.
+Fortunately, minimizing discrepancy with regard to group means can be accomplished by standard k-means anticlustering, i.e., the maximization of $\mathit{SSE}$.
+Combining the maximization of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ at the same time leads to the k-plus approach for anticlustering, which is detailed in the next section.
 
@@
 
-As discussed (**TODO**) in the previous section, an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is not feasible without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
-Moreover, only equalizing variances between groups is most likely not even of interest in most applications.
-Hence, when striving for similarity in group variances by maximizing $\mathit{SSE_{Var}}$, it is at the same time necessary and worthwhile to strive for similarity in means, thus to maximize $\mathit{SSE}$ at the same time.
-Hence, a combination of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ should be employed to simultaneously optimize both criteria, which also represents a reasonable idea when striving for overall between-group similarity.
-The most simple approach for such a multicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
-That is, all criteria are computed independently from each other and are then combined into a single criterion through a weighted sum
+As discussed in the previous section, an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is not feasible without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
+Moreover, an exclusive maximizing of $\mathit{SSE_{Var}}$ is most likely not even interesting in any real application.
+Hence, when striving for similarity in group variances by maximizing $\mathit{SSE_{Var}}$, it is at the same time necessary and worthwhile to strive for similarity in means, and thus to maximize $\mathit{SSE}$ at the same time.
+The most simple approach for a multicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
+That is, all criteria are computed independently from each other and are then combined into a single criterion through a weighted sum.
+By combining $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ in a single criterion, we define the objective function for k-plus anticlustering:
 
@@
 $$
-where the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
-Maximizing $\mathit{SSE_{kplus}}$ for a given $K$ will lead to similar means as well as variances between groups.
+Maximizing $\mathit{SSE_{kplus}}$ will lead to similar means and variances between $K$ groups, while the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
 
@@
 
-Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria at the same time if the underlying objectives are in conflict with each other [@ferligoj1992direct].
-It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ can yield a partitioning where both constituting criteria (mean and variance) are similar between groups, because it is unknown if these objectives tend to align with each other or are in conflict.
+Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict with each other [@ferligoj1992direct].
+As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria (mean and variance) are similar between groups.
 A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible.
@@
 Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree, indicating that these objectives tend to align and are not in conflict with each other.
+Moreover, the scatterplot in Figure 1 shows that there is a high concentration of partitions in the upper-right quadrant where the two objectives are in high accordance.
 
@@
 
-## Including higher order moments
+## Generalizing k-plus anticlustering to higher order moments
 
-Note that the k-plus logic to equalizing group variances as well as group means can be seamlessly generalized to higher order moments.
+The k-plus logic to equalizing group variances as well as group means can be seamlessly generalized to higher order moments.
 The *t*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ can be computed as
@@
 Thus, by changing $t$---the power of the deviation between mean and data points---we can use the k-plus logic to equalize any desired sample moment between groups.
-
-Next, I provide some notation needed to compute the general $\mathit{SSE}$ for k-plus anticlustering, which may include arbitrary distribution moments as part of the computation.
-For each moment $t$ and for column of the original data matrix $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ ($j = 1, \ldots, P$) with mean $\overline{\mathbf{x}}_j$, a new variable $x_{t.j} = (x_{t1j}, \ldots, x_{tNj})$ is computed where
+That is, we compute new variables for each original variable and for each distribution moment that should be considered in the anticlustering process.
+To formalize this idea, I now provide some notation to compute a generalized $\mathit{SSE}$ for k-plus anticlustering, which can include arbitrary distribution moments as part of the criterion.
+First, we define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean is $\overline{\mathbf{x}}_j$.
+Let $T$ be the number of distribution moments that should be equalized between $K$ groups.
+For each moment $t$ ($t = 1, \ldots, T$) and for each variable $x_{.j}$, a new *t-moment variable* $\hat{x}_{t.j} = (\hat{x}_{t1j}, \ldots, \hat{x}_{tNj})$ is defined where
 
 $$
-    x_{tij} = 
+    \hat{x}_{tij} = 
 \begin{cases}
@@
 
-This computation results in an augmented data matrix $\mathbf{\hat{X}}_{N \times (T \cdot P)} = \{x_{tij}\}_{N \times (T \cdot P)}$ where $T$ is the number of distribution moments that should be equalized between groups.
-Note that when defining $x_{tij}$, the condition $t = 1$ is included to keep the unchanged original data as the first $P$ columns in the data, which is used to equalize the the mean---the first moment---between groups.
-Using the thus defined data matrix $\mathbf{\hat{X}}_{N \times (T \cdot P)}$, the generalized multicriterion for k-plus anticlustering can be formalized by the formula
+This computation results in an augmented data matrix $\mathbf{\hat{X}}_{N \times \mathit{TP}} = \{\hat{x}_{tij}\}_{N \times \mathit{TP}}$, i.e.
+
+$$
+  \mathbf{\hat{X}}_{N \times \mathit{TP}} =
+  \left[ {\begin{array}{ccc|ccc|c|ccc}
+    x_{11} & \cdots & x_{1P} & 
+        (x_{11} - \overline{\mathbf{x}}_1)^2 & \cdots & (x_{1P} - \overline{\mathbf{x}}_P)^2 & 
+        \cdots & 
+        (x_{11} - \overline{\mathbf{x}}_1)^T & \cdots & (x_{1P} - \overline{\mathbf{x}}_P)^T \\
+    x_{21} & \cdots & x_{2P} & 
+        (x_{21} - \overline{\mathbf{x}}_1)^2 & \cdots & (x_{2P} - \overline{\mathbf{x}}_P)^2 & 
+        \cdots & 
+        (x_{21} - \overline{\mathbf{x}}_1)^T & \cdots & (x_{2P} - \overline{\mathbf{x}}_P)^T \\
+    \vdots & \ddots & \vdots & \vdots & \ddots & \vdots & \cdots & \vdots & \ddots & \vdots\\
+    x_{N1} & \cdots & x_{NP} & 
+        (x_{N1} - \overline{\mathbf{x}}_1)^2 & \cdots & (x_{NP} - \overline{\mathbf{x}}_P)^2 & 
+        \cdots & 
+        (x_{N1} - \overline{\mathbf{x}}_1)^T & \cdots & (x_{NP} - \overline{\mathbf{x}}_P)^T \\
+  \end{array} } \right]
+$$
+
+Note that when defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as the first $P$ columns in the data, which is used to equalize the the mean---the first moment---between groups.
+
+Using the data matrix $\mathbf{\hat{X}}_{N \times \mathit{TP}}$ and a vector of weights $\mathbf{w} = (w_1, \dots, w_T$), the generalized multicriterion for k-plus anticlustering $\mathit{SSE_{kplusGeneralized}}$ can be formalized:
 
@@
   \sum\limits_{i \in C_k}
-  (x_{tij} - \overline{x}_{.j} ^{(k)})^2
+  (x_{tij} - \overline{x}_{tj} ^{(k)})^2
   \right)
@@
 
-where $w_t$ is the weight associated with the $t'th$ moment and $\overline{x}_{.j} ^{(k)}$ is the mean of the $j'th$ variable in the $k'th$ group.
-Maximizing $\mathit{SSE_{kplusGeneralized}}$ with a given number of moments $T$ can be used to equalize these distribution moments between $K$ groups.
-Note that according to this definition the inclusion of a distribution moment $t$ always implies that all moments $z < t$ are also included as part of the criterion $\mathit{SSE_{kplusGeneralized}}$; it is not deemed worthwhile to exclude lower order moments in favor of higher order moments.
-In particular, the first moment, represented by the original data, should always be included because otherwise the criterion $\mathit{SSE_{kplusGeneralized}}$ is not valid for higher order moments.
+Here, $w_t$ is the weight associated with the $t$'th moment, $\overline{x}_{tj} ^{(k)}$ is mean for the $k$'th group on the $j$'th $t$-moment variable.
+Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to equalize the first $T$ distribution moments between $K$ groups.
+
+## The unweighted sum approach for k-plus anticlustering
 
 **TODO: include discussion of weight selection, as suggested by reviewer 2**
-
-## The unweighted sum approach for k-plus anticlustering
 

### 2023-04-18 paper.Rmd
minor

@@
 Note that when defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as the first $P$ columns in the data, which is used to equalize the the mean---the first moment---between groups.
-
 Using the data matrix $\mathbf{\hat{X}}_{N \times \mathit{TP}}$ and a vector of weights $\mathbf{w} = (w_1, \dots, w_T$), the generalized multicriterion for k-plus anticlustering $\mathit{SSE_{kplusGeneralized}}$ can be formalized:
@@
 
-Here, $w_t$ is the weight associated with the $t$'th moment, $\overline{x}_{tj} ^{(k)}$ is mean for the $k$'th group on the $j$'th $t$-moment variable.
+Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{x}_{tj} ^{(k)}$ is mean for the $k$'th group on the $j$'th $t$-moment variable.
 Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to equalize the first $T$ distribution moments between $K$ groups.

### 2023-04-18 appendix.Rmd
removing covariances from the submission (only keep in discussion)

--- 2022-12-28 appendix.Rmd
@@
 
-```{r Figure2, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *M*."}
+```{r Figure2, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness and kurtosis (split by the number of features, averaged across the remaining variables that varied in the simulation)."}
 ### For the appendix -- other splits for the main results as plot (N is most important)
-# Plot the results, by M
+# Plot the results, by M (number of features)
 ldf %>% 
@@
       Objective, 
-      levels = c("means", "sd", "skew", "kur", "cor"),
-      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
+      levels = c("means", "sd", "skew", "kur"),
+      labels = paste0(
+        "Delta[italic(",
+        c("M", "SD", "Skew", "Kurtosis"),
+        ")]"
+      )
+    ) 
   ) %>% 
@@
 
-```{r Figure3, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *SD*."}
+```{r Figure3, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness and kurtosis (split by *SD*, averaged across the remaining variables that varied in the simulation)."}
 # Plot the results, by SD
@@
 
-```{r Figure4, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *K*."}
+```{r Figure4, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness and kurtosis (split by *K*, averaged across the remaining variables that varied in the simulation)."}
 # Plot the results, by K
@@
 
-```{r Figure5, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *r*."}
+```{r Figure5, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness and kurtosis (split by *r*, averaged across the remaining variables that varied in the simulation)."}
 # Plot the results, by r

### 2023-04-18 paper.Rmd
removing covariances from the submission (only keep in discussion)

@@
 Introducing k-plus, this paper presents an extension of the classical k-means objective to maximize between-group similarity in anticlustering applications.
-The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, higher order moments, and covariances), whereas the k-means criterion only reflects the degree to which groups differ in their means.
+The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, skewness, and higher order moments), whereas the k-means criterion only reflects the degree to which groups differ in their means.
 While constituting a new criterion for anticlustering, it is surprisingly shown that k-plus anticlustering can be implemented by optimizing the original k-means criterion after the input data has been augmented with additional variables.
@@
 
-keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, stimulus selection"
+keywords          : "Anticlustering, k-means, k-plus, variance, skewness, kurtosis, stimulus selection"
 wordcount         : "4992"
@@
 K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to several distribution characteristics instead of only the mean.
-Specifically, k-plus offers the possibility to minimize differences with regard to the variance, covariances, and higher order moments such as skewness and kurtosis.
+Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis.
 After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics.
@@
 
-## Covariances
-
-Because the covariance is also defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariance structure.
-The covariance between two variables $x = (x_1, \ldots, x_N)$ and $y = (y_1, \ldots, y_N)$ is defined as
-
-$$
-\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}}).
-$$
-
-Thus, to maximize between-group similarity with regard to covariance structure, it is possible to augment the data set with an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
-
 # Simulation study
 
-To thoroughly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to group means and variances; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to group means, variances, skewness and kurtosis; (iii) k-plus *correlation*, minimizing differences with regard to group means, variances and covariance structure.
-To implement all k-plus objectives, the unweighted sum approach was employed.
+To thoroughly evaluate the k-plus approach to anticlustering, two k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to group means and variances; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to group means, variances, skewness and kurtosise.
+To implement the k-plus objectives, the unweighted sum approach was employed.
 Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.
@@
   df,
-  cols = paste0(c("means", "sd", "skew", "kur", "cor"), "_obj"),
+  cols = paste0(c("means", "sd", "skew", "kur"), "_obj"),
   names_to = "Objective",
@@
 )
+
+ldf <- ldf[ldf$method != "k-plus-correlation", ] # removed because this deserves closer attention than what was given in the paper; check out the first version of the preprint for the previous results
 
@@
   pivot_wider(names_from = Objective, values_from = Mean) %>% 
-  select(c(means, sd, skew, kur, cor))
+  select(c(means, sd, skew, kur))
 
@@
   "$\\Delta \\mathit{Skew}$",
-  "$\\Delta \\mathit{Kurtosis}$",
-  "$\\Delta \\mathit{Cor}$"
+  "$\\Delta \\mathit{Kurtosis}$"
 )
@@
   note = "The global results of the simulation study aggregated across all conditions.
-Cells contain information about the average global between-group discrepancy with regard to means, standard deviations, skewness, kurtosis, and correlations.
+Cells contain information about the average global between-group discrepancy with regard to means, standard deviations, skewness and kurtosis.
 Lower values indicate less discrepancy, i.e., higher between-group similarity.
@@
 
-```{r Figure1, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by $N$, averaged across the remaining variables that varied in the simulation)."}
+```{r Figure1, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness and kurtosis (split by $N$, averaged across the remaining variables that varied in the simulation)."}
 
@@
       Objective, 
-      levels = c("means", "sd", "skew", "kur", "cor"),
+      levels = c("means", "sd", "skew", "kur"),
       labels = paste0(
         "Delta[italic(",
-        c("M", "SD", "Skew", "Kurtosis", "Cor"),
+        c("M", "SD", "Skew", "Kurtosis"),
         ")]"
@@
 This is an important observation because it shows that k-plus anticlustering is capable of addressing multiple objectives at the same time---especially when $N$ increases.
-However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria made it more difficult to fulfill each single criterion: k-plus skew/kurtosis and k-plus correlation performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus.
-However, every k-plus method did what it was supposed to do: k-plus skew/kurtosis was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation was best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$.
-At the same time, these two k-plus objectives maintained a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
+However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria made it more difficult to fulfill each single criterion: k-plus skew/kurtosis performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus that only considered the mean and variances.
+However, as intended, k-plus skew/kurtosis was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$ and at the same time, this k-plus objective maintained a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
 
@@
 For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering was outperformed.
-As a side note, apart from k-plus correlation, diversity anticlustering was the only method that equalized correlation structure among groups.
-To the best of my knowledge, this observation has not been made previously, even though the maximum diversity problem has been widely studied.
-
 The Appendix includes additional figures splitting the simulation results according to the other variables that varied between simulation runs (number of features; correlation between features; standard deviation of features; number of groups).
-Here, the only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased.
-Because the number of covariances increases quadratically when the number of features increases, it gets comparatively more difficult to fulfill the correlation objective with a higher number of features.
 
@@
   kurtosis = FALSE,
-  covariances = FALSE,
   moments = NULL,
@@
 The function call takes about 10 seconds on a contemporary personal computer.
-The function `kplus_anticlustering()` has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion.
+The function `kplus_anticlustering()` has three boolean arguments to specify whether variance, skewness and kurtosis should be included as part of the k-plus criterion.
 Only the argument `variance` is set to `TRUE` by default---corresponding to the understanding that k-plus is an extension to k-means that at least equalizes variances in addition to means.
-The arguments corresponding to skewness, kurtosis and covariance have to be specifically "turned on", i.e., by setting them to `TRUE` in the function call of `kplus_anticlustering()`.
+The arguments corresponding to skewness and kurtosis have to be specifically "turned on", i.e., by setting them to `TRUE` in the function call of `kplus_anticlustering()`.
 If other higher order moments should be included as part of the optimization, the optional argument `moments` can be used to specify the desired moments as an integer vector.
@@
 
-[^standardkplus]: I refer to *standard* k-plus anticlustering as the minimization of differences in means and variances, but not in higher order moments or covariances.
+[^standardkplus]: I refer to *standard* k-plus anticlustering as the minimization of differences in means and variances, but not in higher order moments.
 
@@
 In this paper, I presented the k-plus criterion for anticlustering, which focuses on maximizing between-group similarity.
-K-plus represents between-group similarity as discrepancies in distribution moments between groups: means, variances, skewness, kurtosis, higher order moments, and covariances.
+K-plus represents between-group similarity as discrepancies in distribution moments between groups: means, variances, skewness, kurtosis, and higher order moments.
 Depending on the application, a different subset of these objectives can be included in the anticlustering process.
@@
 
+It should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariance, a possibility that was not investigated in the current paper.
+The covariance between two variables $x = (x_1, \ldots, x_N)$ and $y = (y_1, \ldots, y_N)$ is defined as $\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
+Thus, to maximize between-group similarity with regard to covariance structure, it is possible to augment the data set with an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
+Future research should examine how the k-plus approach can be adapted to including covariances, including a discussion of how to select weights for the possibly high number of additional variables.
+The unweighted standardization approach presented here may serve as a starting point for investigation.
+
 [^clusterediting]: In their study, they used the term anticluster editing to refer to the maximization of the diversity criterion.

### 2023-04-18 paper.Rmd
make contribution in simulation study clearer, as per reviews

@@
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis.
-After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics.
+After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional anticlustering criteria.
 Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].

### 2023-04-18 paper.Rmd
minor

@@
 As discussed in the previous section, an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is not feasible without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
-Moreover, an exclusive maximizing of $\mathit{SSE_{Var}}$ is most likely not even interesting in any real application.
+Moreover, maximizing only $\mathit{SSE_{Var}}$ is most likely not interesting in any real application.
 Hence, when striving for similarity in group variances by maximizing $\mathit{SSE_{Var}}$, it is at the same time necessary and worthwhile to strive for similarity in means, and thus to maximize $\mathit{SSE}$ at the same time.
 The most simple approach for a multicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
-That is, all criteria are computed independently from each other and are then combined into a single criterion through a weighted sum.
+That is, all criteria are computed independently from each other and are then combined into a single criterion through a weighted sum; a suitable optimization algorithm is then used to maximize the combined criterion.
 By combining $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ in a single criterion, we define the objective function for k-plus anticlustering:

### 2023-04-18 paper.Rmd
figure caption

@@
 ```{r, fig.cap = "The relationship between the objectives $\\mathit{SSE}$ and $\\mathit{SSE_{Var}}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
-The partition represented by the square maximizes $\\mathit{SSE}$ and thus minimizes the difference in means between groups; the partition represented by the triangular maximizes $\\mathit{SSE_{Var}}$ and thus minimizes the difference in variances between groups.
-The partition represented by the diamond maximizes $\\mathit{SSE_{kplus}}$ and satisfies both constituting criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+The partition represented by the square maximizes $\\mathit{SSE}$ and thus minimizes the difference in means between groups; the partition represented by the triangular maximizes $\\mathit{SSE_{Var}}$.
+The partition represented by the diamond maximizes $\\mathit{SSE_{kplus}}$ for $w_1 = w_2 = 1$, and satisfies both constituting criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
 

### 2023-04-19 paper.Rmd
more work on problem formalization

@@
 That is, we compute new variables for each original variable and for each distribution moment that should be considered in the anticlustering process.
-To formalize this idea, I now provide some notation to compute a generalized $\mathit{SSE}$ for k-plus anticlustering, which can include arbitrary distribution moments as part of the criterion.
+
+To formalize the computation of the k-plus criterion for an arbitraty number of distribution moments, some new notation has to be provided.
 First, we define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean is $\overline{\mathbf{x}}_j$.
@@
 
-This computation results in an augmented data matrix $\mathbf{\hat{X}}_{N \times \mathit{TP}} = \{\hat{x}_{tij}\}_{N \times \mathit{TP}}$, i.e.
+This computation results in a three dimensional tensor $\mathbf{\hat{X}}_{T \times N \times P}$ = $\{\hat{x}_{tij}\}_{T \times N \times P}$ where---in comparison to the original data matrix $\mathbf{X}_{N \times P}$---the additional dimension is given by the $T$ distribution moments that are considered in the anticlustering process.
+When defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as first entry in the "moment dimension" in $\mathbf{\hat{X}}_{T \times N \times P}$, which is used to equalize the the mean---the first moment---between groups.
+
+Using the tensor  $\mathbf{\hat{X}}_{T \times N \times P}$ and a vector of weights $\mathbf{w} = (w_1, \dots, w_T$), the generalized multicriterion for k-plus anticlustering $\mathit{SSE_{kplusGeneralized}}$ can be formalized:
 
 $$
-  \mathbf{\hat{X}}_{N \times \mathit{TP}} =
+\mathit{SSE_{kplusGeneralized}} =
+  \sum\limits_{t=1}^{T} 
+  w_t \left(
+  \sum\limits_{j=1}^{P} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (x_{tij} - \overline{x}_{tj} ^{(k)})^2
+  \right)
+$$
+
+Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{x}_{tj} ^{(k)}$ is mean for the $k$'th group on the $j$'th $t$-moment variable.
+Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to equalize the first $T$ distribution moments between $K$ groups.
+
+## The unweighted sum approach for k-plus anticlustering
+
+Some simplifications in the computation of $\mathit{SSE_{kplusGeneralized}}$ can be arranged when ignoring the moment weights $\mathbf{w}$, e.g.
+by setting $w_t = 1$ ($\forall t = 1, \dots, T$).
+This simplification leads to the *unweighted sum approach* for k-plus anticlustering.
+In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$, representing distribution moments can be dropped; instead, the tensor is flattened into a two dimensional matrix, where the variables representing the additional distribution moments are appended to the original data matrix $\mathbf{X}_{N \times P}$, resulting in the augmented matrix $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$:
+
+$$
+  \mathbf{\hat{X}'}_{N \times \mathit{TP}} =
   \left[ {\begin{array}{ccc|ccc|c|ccc}
@@
 
-Note that when defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as the first $P$ columns in the data, which is used to equalize the the mean---the first moment---between groups.
-Using the data matrix $\mathbf{\hat{X}}_{N \times \mathit{TP}}$ and a vector of weights $\mathbf{w} = (w_1, \dots, w_T$), the generalized multicriterion for k-plus anticlustering $\mathit{SSE_{kplusGeneralized}}$ can be formalized:
+When defining the row vectors of $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ with the notation $\hat{x}'_{i.} = (\hat{x}'_{i1}, \ldots, \hat{x}'_{i(TP)})$ for each $i = 1, \ldots, N$, the generalized criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as
 
@@
 \mathit{SSE_{kplusGeneralized}} =
-  \sum\limits_{t=1}^{T} 
-  w_t \left(
-  \sum\limits_{j=1}^{P} 
+  \sum\limits_{j=1}^{\mathit{TP}}
   \sum\limits_{k=1}^{K} 
   \sum\limits_{i \in C_k}
-  (x_{tij} - \overline{x}_{tj} ^{(k)})^2
-  \right)
+  (\hat{x}'_{ij} - \overline{\hat{x}'}_j ^{(k)})^2.
 $$
 
-Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{x}_{tj} ^{(k)}$ is mean for the $k$'th group on the $j$'th $t$-moment variable.
-Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to equalize the first $T$ distribution moments between $K$ groups.
-
-## The unweighted sum approach for k-plus anticlustering
-
-**TODO: include discussion of weight selection, as suggested by reviewer 2**
-
-**argue why unweighted standardization k-plus anticlustering is good, as demanded by reviewer 2; also maybe include small scale simulation study**
-
-Some simplifications can be arranged when ignoring the weights $w_1$ and $w_2$, e.g., by setting them both to 1.
-This simplification leads to the *unweighted sum approach* for k-plus anticlustering: When defining a new data matrix $\mathbf{Z}_{N \times 2P} = \{z_{ij}\}_{N \times 2P}$ as the augmentation of $\mathbf{X}_{N \times P}$ with $\mathbf{Y}_{N \times P}$, i.e.
-
-$$
-  \mathbf{Z}_{N \times 2P} =
-  \left[ {\begin{array}{cccc|cccc}
-    x_{11} & x_{12} & \cdots & x_{1P} & y_{11} & y_{12} & \cdots & y_{1P}\\
-    x_{21} & x_{22} & \cdots & x_{2P} & y_{21} & y_{22} & \cdots & y_{2P}\\
-    \vdots & \vdots & \ddots & \vdots & \vdots & \vdots & \ddots & \vdots\\
-    x_{N1} & x_{N2} & \cdots & x_{NP} & y_{N1} & y_{N2} & \cdots & y_{NP}\\
-  \end{array} } \right]
-$$
-
-and $z_{i.} = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, y_{iP})$, the bicriterion $\mathit{SSE_{kplus}}$ can be computed as
-
-$$
-\mathit{SSE_{kplus}} =
-  \sum\limits_{j=1}^{2 P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (z_{ij} - \overline{z}_j ^{(k)})^2.
-$$
-
-This term again has the same form as $\mathit{SSE}$ but is computed on different data.
-Hence, computing the k-plus criterion can be done by first computing the squared distances $y_{ij}$, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
+Thus, when setting all the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ again has the same form as the standard k-means criterion $\mathit{SSE}$, but it is computed on different data.
+Here, computing the k-plus criterion is done by first computing the additional $t$-moment variables, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
 Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
 
-Note that a problem may arise if the weights $w_1$ and $w_2$ are ignored when computing $\mathit{SSE_{kplus}}$.
-Because the computation of the new attribute values $y_{ij}$ involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$ (e.g., see Figure 1).
+Note that a problem may arise if the weights $\mathbf{w}$ are ignored when computing.
+Because the computation of the new $t$-moment variables involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$ (e.g., see Figure 1).
 This imbalance can strongly and unpredictably influence the relative weight of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ when computing $\mathit{SSE_{kplus}}$.
-To ensure an equal weight of both criteria, all attributes ($x_{ij}$ and $y_{ij}$) can be standardized before computing $\mathit{SSE_{kplus}}$, e.g.
+To ensure an equal weight of both criteria, all attributes---i.e., the original data as well as the new $t$-moment varibles---can be standardized before computing $\mathit{SSE_{kplusGeneralized}}$, e.g.
 via $z$-standardization.

### 2023-04-19 paper.Rmd
including simulation for weight selection

@@
 
-# Simulation study
+# Evaluation
 
-To thoroughly evaluate the k-plus approach to anticlustering, two k-plus criteria were implemented in a simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to group means and variances; (ii) k-plus *skew/kurtosis*, minimizing differences with regard to group means, variances, skewness and kurtosise.
-To implement the k-plus objectives, the unweighted sum approach was employed.
-Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.
-Subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
-The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means, (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups, (iii) a random allocation of elements to groups.
+Two simulation studies were conducted to evaluate the new k-plus objective for anticlustering.
+A first small scale simulation compared different approaches for weight selection for the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$.
+In a second large scale simulation study, the k-plus objective was compared to other popular anticlustering objectives, i.e., the k-means and diversity objectives.
 
-The simulation was implemented using the statistical programming language R [Version 4.2, @R-base].
+The simulation studies were implemented using the statistical programming language R [Version 4.2, @R-base].
 The R package `anticlust` [Version 0.6.1, @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
@@
 
-The same local maximum search was applied to optimize each of the five competing anticlustering objectives [Method LCW, @weitz1998].
+In both simulation studies, the same local maximum search was applied to optimize the competing anticlustering objectives [Method LCW, @weitz1998].
 Building on an initial random allocation, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
@@
 To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
-In the simulation study five repetitions were used.
+In the simulation studies, five repetitions were used.
 
-## Conditions
+## Simulation 1: Selecting weights for k-plus anticlustering
 
-For the simulation, I generated 10,000 data sets following a normal distribution.
+```{r getresultsSim1}
+files <- list.files("./Simulation_Study/Weight_Simulation", pattern = ".Rdata")
+results <- list()
+for (file in files) {
+  results <- c(results, get(load(paste0("Simulation_Study/Weight_Simulation/", file))))
+}
+
+```
+
+In general, the weighting scheme for k-plus anticlustering can become highly complex as it might encompass an arbitrary number of distribution moments, and hence, an arbitrary number of weights.
+While an exhaustive evaluation of the influence of weights is beyond the scope of the current paper, a small scale simulation study was conducted to investigate the influence of weights for a selected number of moments.
+The target was the optimization of $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, i.e.
+minizimizing differences with regard to means, variances, skew and kurtosis.
+The two unweighted approaches were implemented with $\mathbf{w} = (1, 1, 1, 1)$.
+In one case, a standardization of all columns of $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ was implemented.
+That is, the original as well as the augmented variables were standardized before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
+In the second case, the unstandardized matrix $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ was used.
+The four weighted approaches were implemented by giving a 10 times increased weight to one of the respective moments.
+That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favoring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favoring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, repectively.
+The weighted approaches did not use standardization of the input data.
+
+To compare the different weighting schemes, `r length(results)` data sets from were generated from a normal distribution ($M = 0$, $\mathit{SD} = 1$).
+The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 10 and 100 with the restriction that $N$ had to be divisible by $K$.
+For each data set, $\mathit{SSE_{kplusGeneralized}}$ was optimized using the local maximum algorithm, once for each of the six competing weighting schemes.
+
+```{r}
+final_tab_sim1 <- (Reduce("+", results) / length(results) * 100) |> round(2)
+rownames(final_tab_sim1) <- c(
+  "$\\mathbf{w} = (1, 1, 1, 1)$ / unstandardized",
+  "$\\mathbf{w} = (1, 1, 1, 1)$ / standardized",
+  "$\\mathbf{w} = (10, 1, 1, 1)$",
+  "$\\mathbf{w} = (1, 10, 1, 1)$",
+  "$\\mathbf{w} = (1, 1, 10, 1)$",
+  "$\\mathbf{w} = (1, 1, 1, 10)$"
+)
+colnames(final_tab_sim1) <- c(
+  "$\\mathit{SSE}$",
+  "$\\mathit{SSE_{Var}}$", 
+  "$\\mathit{SSE_{Skew}}$",
+  "$\\mathit{SSE_{Kurtosis}}$",
+  "Average"
+)
+note <- paste0("The results of Simulation 1 averaged across all ", length(results), " data sets.
+Cells indicate how well the maximization of the weighted $\\mathit{SSE_{kplusGeneralized}}$ fulfilled each of the four constituting criteria $\\mathit{SSE}$, $\\mathit{SSE_{Var}}$, $\\mathit{SSE_{Skew}}$, and $\\mathit{SSE_{Kurtosis}}$.
+Cell values represent relative performance, i.e.
+they were computed as the $\\mathit{SSE}$ that was attained by the given weighting scheme, divided by the best $\\mathit{SSE}$ that was obtained by any weighting scheme in a given simulation run.")
+
+apa_table(
+  final_tab_sim1, 
+  escape = FALSE, 
+  caption = "Results of Simulation study 1",
+  note = note,
+  stub_indents = list("Unweighted approaches" = 1:2,
+                      "Weighted approaches" = 3:6)
+)
+
+unstd_best <- sapply(results, function(x) names(which.max(x[, 5]))) == "unweighted_std"
+
+```
+
+Table 1 illustrates how well each of the six weighting schemes fulfilled each of the four constititung criteria, averaged across the `r length(results)` simulation runs.
+It is shown that for the weighted approaches, the criterion that received the highest weight was fulfilled to the highest degree, as should be expected.
+Interestingly, the unweighted approach including standardization had the best average performance across the four criteria.
+While the advantage over the other weighting schemes does not seem to be remarkable, this result held across an astonishing `r sum(unstd_best)` simulation runs (`r mean(unstd_best) * 100`%).
+Among the competing methods, the unweighted standardization approach therefore seems preferable.
+It is also a strong approach because it yielded the best results with regard to maximizing the standard $\mathit{SSE}$.
+Hence, the unweighted standardization approach implicitly gives high precedence to minimizing differences in means.
+This is a pleasant result because, as discussed earlier, for the higher order moment $\mathit{SSE}$s to be valid, it has to be ensured that the group means are close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ receives high priority.
+
+## Simulation 2: Comparing k-plus anticlustering to other objectives
+
+To thoroughly evaluate the k-plus approach to anticlustering, two specific k-plus criteria were compared to traditional anticlustering criteria in a second simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to group means and variances (by maximizing $\mathit{SSE_{kplus}}$); (ii) "generalized" k-plus anticlustering, minimizing differences with regard to group means, variances, skewness and kurtosis (by maximizing $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$).
+To implement the k-plus objectives, the unweighted standardization approach was employed that performed best in Simulation 1.
+Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.
+Subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
+The k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means, (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups, (iii) a random allocation of elements to groups.
+
+### Conditions
+
+For Simulation 2, I generated 10,000 data sets following a normal distribution.
 The data sets were subsequently processed by the competing methods.
@@
 
-## Evaluation criteria
+### Evaluation criteria
 
@@
 
-## Results
+### Results
 
@@
 Replicating @papenberg2020, k-means' performance with regard to similarity in standard deviations was even below that of a random assignment.
-K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); all three k-plus approaches strongly outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
+K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); both k-plus approaches strongly outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
 
@@
   tab,
-  caption = "Results of the simulation study",
+  caption = "Results of Simulation study 2",
   note = "The global results of the simulation study aggregated across all conditions.

### 2023-04-19 Simulation_Study/Weight_Simulation/functions_weight_simulation.R

@@ -1,0 +1,107 @@
+library(anticlust)
+
+# Compute features for k-plus criterion
+moment_features <- function(data, moment = 2) {
+  apply(data, 2, function(x) (x - mean(x))^moment)
+}
+
+kplus_objective_weighted <- function(features, clusters, moments, weights) {
+  objectives <- rep(NA, length(moments))
+  for (i in moments) {
+    if (i == 1) {
+      objectives[i] <- variance_objective(features, clusters)
+    } else {
+      moment_variables <- moment_features(features, moments[i])
+      objectives[i] <- variance_objective(moment_variables, clusters)
+    }
+  }
+  sum(objectives * weights)
+}
+
+kplus_mean_preferred <- function(features, clusters) {
+  kplus_objective_weighted(features, clusters, 1:4, c(10, 1, 1, 1))
+}
+
+kplus_var_preferred <- function(features, clusters) {
+  kplus_objective_weighted(features, clusters, 1:4, c(1, 10, 1, 1))
+}
+
+kplus_skew_preferred <- function(features, clusters) {
+  kplus_objective_weighted(features, clusters, 1:4, c(1, 1, 10, 1))
+}
+
+kplus_kurtosis_preferred <- function(features, clusters) {
+  kplus_objective_weighted(features, clusters, 1:4, c(1, 1, 1, 10))
+}
+
+clusters_for_all_methods <- function(data, K) {
+  all_results <- list()
+  repetitions <- 1
+  method <- "local-maximum"
+  all_results[["unweighted_unstd"]] <- anticlustering(
+    cbind(data, moment_features(data, 2), moment_features(data, 3), moment_features(data, 4)),
+    K = K,
+    objective = "variance",
+    method = method,
+    repetitions = repetitions
+  )
+  all_results[["unweighted_std"]] <- anticlustering(
+    scale(cbind(data, moment_features(data, 2), moment_features(data, 3), moment_features(data, 4))),
+    K = K,
+    objective = "variance",
+    method = method,
+    repetitions = repetitions
+  )
+  
+  all_results[["mean_preferred"]] <- anticlustering(
+    data,
+    K = K,
+    objective = kplus_mean_preferred,
+    method = method,
+    repetitions = repetitions
+  )
+  all_results[["var_preferred"]] <- anticlustering(
+    data,
+    K = K,
+    objective = kplus_var_preferred,
+    method = method,
+    repetitions = repetitions
+  )
[…]

### 2023-04-19 Simulation_Study/Weight_Simulation/weight_simulation.R

@@ -1,0 +1,21 @@
+
+
+
+source("./Simulation_Study/Weight_Simulation/functions_weight_simulation.R")
+
+nsim <- 10
+results <- list()
+for (i in 1:nsim) {
+  K <- sample(2:4, size = 1)
+  N <- sample((10:100)[10:100 %% K == 0], size = 1)
+  M <- sample(1:5, size = 1)
+  message("Starting to work on N = ", N, ", K = ", K, ", M = ", M, ".")
+  dat <- matrix(rnorm(N * M), ncol = M)
+  start <- Sys.time()
+  results[[i]] <- weight_sim(dat, K = K)
+  print(Sys.time() - start)
+  results[[i]] <- cbind(results[[i]], rowMeans(results[[i]]))
+  colnames(results[[i]]) <- c("SSE", "SSE_Var", "SSE_Skew", "SSE_Kurtosis", "Average")
+}
+
+save(results, file = paste0("./Simulation_Study/Weight_Simulation/Results Sim 1 ", date(), ".Rdata"))

### 2023-04-19 paper.Rmd
update table numbers

@@
 
-Table 1 displays the global simulation results aggregated across all conditions.
+Table 2 displays the global simulation results aggregated across all conditions.
 Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but did not outperform a random assignment with regard to all other distribution characteristics.
@@
 
-This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] using the function `kplus_anticlustering()`, which is available from version 0.6.3 onward.
+This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] using the function `kplus_anticlustering()`, which implements the unweighted sum approach for k-plus anticlustering.
 The package can be installed in the R environment using the `install.packages()` command:
@@
 
-Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features.
-Table 2 also contains the results after applying k-means anticlustering and diversity anticlustering.
+Table 3 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features.
+Table 3 also contains the results after applying k-means anticlustering and diversity anticlustering.
 It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations, and the skewness was also very evenly matched.
@@
 
-Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups.
+Table 4 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups.
 K-plus anticlustering strives for between-group similarity regardless of group size.
@@
 
-Table 4 illustrates the results that are obtained when using standard[^standardkplus] k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
+Table 5 illustrates the results that are obtained when using standard[^standardkplus] k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
 Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are similar enough for any practical purposes.

### 2023-04-19 Simulation_Study/Weight_Simulation/weight_simulation.R
print time for all data sets

@@ -3,19 +3,19 @@
 
 source("./Simulation_Study/Weight_Simulation/functions_weight_simulation.R")
 
-nsim <- 10
+nsim <- 30
 results <- list()
+start <- Sys.time()
 for (i in 1:nsim) {
   K <- sample(2:4, size = 1)
   N <- sample((10:100)[10:100 %% K == 0], size = 1)
   M <- sample(1:5, size = 1)
   message("Starting to work on N = ", N, ", K = ", K, ", M = ", M, ".")
   dat <- matrix(rnorm(N * M), ncol = M)
-  start <- Sys.time()
   results[[i]] <- weight_sim(dat, K = K)
-  print(Sys.time() - start)
   results[[i]] <- cbind(results[[i]], rowMeans(results[[i]]))
   colnames(results[[i]]) <- c("SSE", "SSE_Var", "SSE_Skew", "SSE_Kurtosis", "Average")
 }
+print(Sys.time() - start)
 
 save(results, file = paste0("./Simulation_Study/Weight_Simulation/Results Sim 1 ", date(), ".Rdata"))

### 2023-04-19 paper.Rmd
was undefined before

@@
 
-The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all observations belonging to the $k$'th cluster $C_k$ ($k = 1, \ldots, K$):
+The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all $n_k$ observations belonging to the $k$'th cluster $C_k$ ($k = 1, \ldots, K$):
 

### 2023-04-19 Simulation_Study/Weight_Simulation/functions_weight_simulation.R
use t as index as in paper

@@ -7,12 +7,12 @@ moment_features <- function(data, moment = 2) {
 
 kplus_objective_weighted <- function(features, clusters, moments, weights) {
   objectives <- rep(NA, length(moments))
-  for (i in moments) {
-    if (i == 1) {
-      objectives[i] <- variance_objective(features, clusters)
+  for (t in moments) {
+    if (t == 1) {
+      objectives[t] <- variance_objective(features, clusters)
     } else {
-      moment_variables <- moment_features(features, moments[i])
-      objectives[i] <- variance_objective(moment_variables, clusters)
+      moment_variables <- moment_features(features, moments[t])
+      objectives[t] <- variance_objective(moment_variables, clusters)
     }
   }
   sum(objectives * weights)

### 2023-04-19 Simulation_Study/Weight_Simulation/functions_weight_simulation.R
comments

@@ -38,10 +38,11 @@ clusters_for_all_methods <- function(data, K) {
   all_results <- list()
   repetitions <- 1
   method <- "local-maximum"
+  # antiweighted k-plus anticlustering really is just k-means anticlusteriung
   all_results[["unweighted_unstd"]] <- anticlustering(
     cbind(data, moment_features(data, 2), moment_features(data, 3), moment_features(data, 4)),
     K = K,
-    objective = "variance",
+    objective = "variance", # k-means
     method = method,
     repetitions = repetitions
   )

### 2023-04-19 paper.Rmd
typo

@@
 This computation results in a three dimensional tensor $\mathbf{\hat{X}}_{T \times N \times P}$ = $\{\hat{x}_{tij}\}_{T \times N \times P}$ where---in comparison to the original data matrix $\mathbf{X}_{N \times P}$---the additional dimension is given by the $T$ distribution moments that are considered in the anticlustering process.
-When defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as first entry in the "moment dimension" in $\mathbf{\hat{X}}_{T \times N \times P}$, which is used to equalize the the mean---the first moment---between groups.
+When defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as first entry in the "moment dimension" in $\mathbf{\hat{X}}_{T \times N \times P}$, which is used to equalize the mean---the first moment---between groups.
 

### 2023-04-19 Simulation_Study/Weight_Simulation/weight_simulation.R
use N = 20 as minimum in Simulation 1

@@ -8,7 +8,7 @@ results <- list()
 start <- Sys.time()
 for (i in 1:nsim) {
   K <- sample(2:4, size = 1)
-  N <- sample((10:100)[10:100 %% K == 0], size = 1)
+  N <- sample((20:100)[20:100 %% K == 0], size = 1)
   M <- sample(1:5, size = 1)
   message("Starting to work on N = ", N, ", K = ", K, ", M = ", M, ".")
   dat <- matrix(rnorm(N * M), ncol = M)

### 2023-04-19 paper.Rmd
misc

@@
 Some simplifications in the computation of $\mathit{SSE_{kplusGeneralized}}$ can be arranged when ignoring the moment weights $\mathbf{w}$, e.g.
-by setting $w_t = 1$ ($\forall t = 1, \dots, T$).
+by setting $w_t = 1$ ($t = 1, \dots, T$).
 This simplification leads to the *unweighted sum approach* for k-plus anticlustering.
@@
 Thus, when setting all the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ again has the same form as the standard k-means criterion $\mathit{SSE}$, but it is computed on different data.
-Here, computing the k-plus criterion is done by first computing the additional $t$-moment variables, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
+Here, computing the k-plus criterion is done by first creating the additional $t$-moment variables, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
 Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
 
-Note that a problem may arise if the weights $\mathbf{w}$ are ignored when computing.
-Because the computation of the new $t$-moment variables involves squaring, their range of values is expected to be very different from the raw values $x_{ij}$ (e.g., see Figure 1).
-This imbalance can strongly and unpredictably influence the relative weight of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ when computing $\mathit{SSE_{kplus}}$.
-To ensure an equal weight of both criteria, all attributes---i.e., the original data as well as the new $t$-moment varibles---can be standardized before computing $\mathit{SSE_{kplusGeneralized}}$, e.g.
+Note that a problem may arise if the weights $\mathbf{w}$ are ignored when computing $\mathit{SSE_{kplusGeneralized}}$.
+Because the computation of the new $t$-moment variables involves exponentiation, their range of values is expected to be very different from the raw values $x_{ij}$ (e.g., see Figure 1).
+This imbalance can strongly and unpredictably influence the relative weight of the different moments when computing $\mathit{SSE_{kplusGeneralized}}$.
+To ensure an equal weight of all criteria, all attributes---i.e., the original data as well as the new $t$-moment variables---can be standardized before computing $\mathit{SSE_{kplusGeneralized}}$, e.g.
 via $z$-standardization.
+This approach will henceforth be referred to as the unweighted standardization approach for k-plus anticlustering.
 
@@
 The R package `anticlust` [Version 0.6.1, @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
-The R package `faux` [Version 1.1.0, @R-faux] was used to generate the multivariate normal data sets that were processed during the simulation.
+The R package `faux` [Version 1.1.0, @R-faux] was used to generate the multivariate normal data sets that were processed during Simulation 2.
 The R packages `dplyr` [Version 1.0.7, @R-dplyr], `ggplot2` [Version 3.3.5, @R-ggplot2], `papaja` [Version 0.1.1.9001, @R-papaja], `tidyr` [Version 1.1.3, @R-tidyr] and `DescTools` [Version 0.99.45, @R-desctools] were used for data analysis and result presentation.
@@
 minizimizing differences with regard to means, variances, skew and kurtosis.
-The two unweighted approaches were implemented with $\mathbf{w} = (1, 1, 1, 1)$.
-In one case, a standardization of all columns of $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ was implemented.
-That is, the original as well as the augmented variables were standardized before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
-In the second case, the unstandardized matrix $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ was used.
-The four weighted approaches were implemented by giving a 10 times increased weight to one of the respective moments.
+Two unweighted approaches (using standardization vs.
+not using standardization) were implemented with $\mathbf{w} = (1, 1, 1, 1)$.
+The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented variables---of $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
+In the other case, the unstandardized matrix $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ was used.
+The four weighted approaches were implemented by giving a tenfold increased weight to only one of the four moments.
 That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favoring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favoring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, repectively.
@@
 To compare the different weighting schemes, `r length(results)` data sets from were generated from a normal distribution ($M = 0$, $\mathit{SD} = 1$).
-The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 10 and 100 with the restriction that $N$ had to be divisible by $K$.
+The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$.
 For each data set, $\mathit{SSE_{kplusGeneralized}}$ was optimized using the local maximum algorithm, once for each of the six competing weighting schemes.
@@
 ```{r}
-final_tab_sim1 <- (Reduce("+", results) / length(results) * 100) |> round(2)
+final_tab_sim1 <- round(Reduce("+", results) / length(results) * 100, 2)
 rownames(final_tab_sim1) <- c(
@@
 Interestingly, the unweighted approach including standardization had the best average performance across the four criteria.
-While the advantage over the other weighting schemes does not seem to be remarkable, this result held across an astonishing `r sum(unstd_best)` simulation runs (`r mean(unstd_best) * 100`%).
-Among the competing methods, the unweighted standardization approach therefore seems preferable.
-It is also a strong approach because it yielded the best results with regard to maximizing the standard $\mathit{SSE}$.
-Hence, the unweighted standardization approach implicitly gives high precedence to minimizing differences in means.
-This is a pleasant result because, as discussed earlier, for the higher order moment $\mathit{SSE}$s to be valid, it has to be ensured that the group means are close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ receives high priority.
+While the advantage over the other weighting schemes does not seem to be remarkable, this pattern was observed in an astonishing `r sum(unstd_best)` of all `r length(results)` simulation runs (`r mean(unstd_best) * 100`%).
+Among the competing weighting schemes, the unweighted standardization approach therefore seems preferable.
+It also yielded the best results with regard to maximizing the standard $\mathit{SSE}$.
+The unweighted standardization approach seemed to implicitly give high priority to minimizing differences in means.
+This is a pleasant result because, as discussed earlier, for the higher order moment $\mathit{SSE}$s to be valid, the group means have to be close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ has high priority.
 
@@
 This is an important observation because it shows that k-plus anticlustering is capable of addressing multiple objectives at the same time---especially when $N$ increases.
-However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria made it more difficult to fulfill each single criterion: k-plus skew/kurtosis performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus that only considered the mean and variances.
-However, as intended, k-plus skew/kurtosis was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$ and at the same time, this k-plus objective maintained a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
-
-It is maybe surprising that k-plus skew/kurtosis does not seem to have performed much better than diversity anticlustering with regard to skewness and kurtosis, even though the method has been specifically tailored to these criteria.
-Figure 2 gives a more fine grained assessment of this observation: Surprisingly, for small $N$, k-plus skew/kurtosis was even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
+However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria made it more difficult to fulfill each single criterion: generalized k-plus performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus that only considered the mean and variances.
+However, as intended, generalized k-plus was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$ and at the same time, this k-plus objective maintained a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
+
+It is maybe surprising that generalized k-plus does not seem to have performed much better than diversity anticlustering with regard to skewness and kurtosis, even though the objective has been specifically tailored to these criteria.
+Figure 2 gives a more fine grained assessment of this observation: Surprisingly, for small $N$, generalized k-plus was even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
 However, with increasing $N$, it clearly outperformed diversity anticlustering.
-These results illustrate the cost of optimizing several criteria at the same time: For low $N$, k-plus skew/kurtosis was apparently preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
+These results illustrate the cost of optimizing several criteria at the same time: For low $N$, generalized k-plus was apparently preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
 Increasing $N$ then facilitated to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.

### 2023-04-19 paper.Rmd
misc

@@
 Note that there is a direct connection between the $\mathit{SSE}$ and the location of the cluster centroids [@spath1986].
-If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_p$ is the overall mean value on the $p$th attribute:
-
-$$
-\overline{\mathbf{x}}_p = \frac{1}{N} \sum\limits_{i=1}^{N} x_{ip}
+If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_j$ ($j = 1, \ldots, P$) is the overall mean value on the $j$th attribute:
+
+$$
+\overline{\mathbf{x}}_j = \frac{1}{N} \sum\limits_{i=1}^{N} x_{ij}
 $$
@@
 Maximizing $\mathit{SSE_{Var}}$ can be used to create groups that have similar variances across the $P$ attributes.
-As a caveat, it is crucial to note that maximizing $\mathit{SSE_{Var}}$ only reliably yields similar group variances if the group centroids $\overline{\mathbf{x}}^{(k)}$, ($k = 1, \dots, K$) are close to the overall data centroid $\overline{\mathbf{x}}$.
+As a caveat, it is crucial to note that maximizing $\mathit{SSE_{Var}}$ only reliably yields similar group variances if the group centroids $\overline{\mathbf{x}}^{(k)}$ ($k = 1, \dots, K$) are close to the overall data centroid $\overline{\mathbf{x}}$.
 To reiterate, this condition has to be met for the original data $\mathbf{X}_{N \times P}$.
@@
 
-As discussed in the previous section, an exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is not feasible without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
-Moreover, maximizing only $\mathit{SSE_{Var}}$ is most likely not interesting in any real application.
+An exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is not feasible---and most likely not interesting on its own---without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
 Hence, when striving for similarity in group variances by maximizing $\mathit{SSE_{Var}}$, it is at the same time necessary and worthwhile to strive for similarity in means, and thus to maximize $\mathit{SSE}$ at the same time.
-The most simple approach for a multicriterion optimization is the weighted sum approach [@naidu2014; @marler2010].
+The most simple approach for an optimization that considers multiple criteria at the same time is the weighted sum approach [@naidu2014; @marler2010].
 That is, all criteria are computed independently from each other and are then combined into a single criterion through a weighted sum; a suitable optimization algorithm is then used to maximize the combined criterion.
@@
 This computation results in a three dimensional tensor $\mathbf{\hat{X}}_{T \times N \times P}$ = $\{\hat{x}_{tij}\}_{T \times N \times P}$ where---in comparison to the original data matrix $\mathbf{X}_{N \times P}$---the additional dimension is given by the $T$ distribution moments that are considered in the anticlustering process.
-When defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as first entry in the "moment dimension" in $\mathbf{\hat{X}}_{T \times N \times P}$, which is used to equalize the mean---the first moment---between groups.
+When defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as first entry in the "moment dimension" in $\mathbf{\hat{X}}_{T \times N \times P}$.
+It is used to equalize the mean---the first moment---between groups.
 
@@
 This simplification leads to the *unweighted sum approach* for k-plus anticlustering.
-In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$, representing distribution moments can be dropped; instead, the tensor is flattened into a two dimensional matrix, where the variables representing the additional distribution moments are appended to the original data matrix $\mathbf{X}_{N \times P}$, resulting in the augmented matrix $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$:
-
-$$
-  \mathbf{\hat{X}'}_{N \times \mathit{TP}} =
+In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$, representing distribution moments can be dropped; instead, the tensor is flattened into a two dimensional matrix, where the variables representing the additional distribution moments are appended to the original data matrix $\mathbf{X}_{N \times P}$, resulting in the augmented matrix $\mathbf{X'}_{N \times \mathit{TP}}$:
+
+$$
+  \mathbf{X'}_{N \times \mathit{TP}} =
   \left[ {\begin{array}{ccc|ccc|c|ccc}
@@
 
-When defining the row vectors of $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ with the notation $\hat{x}'_{i.} = (\hat{x}'_{i1}, \ldots, \hat{x}'_{i(TP)})$ for each $i = 1, \ldots, N$, the generalized criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as
+When defining the row vectors of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i.} = (x'_{i1}, \ldots, x'_{i(TP)})$ for each $i = 1, \ldots, N$, the criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as
 
@@
   \sum\limits_{i \in C_k}
-  (\hat{x}'_{ij} - \overline{\hat{x}'}_j ^{(k)})^2.
-$$
-
-Thus, when setting all the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ again has the same form as the standard k-means criterion $\mathit{SSE}$, but it is computed on different data.
-Here, computing the k-plus criterion is done by first creating the additional $t$-moment variables, appending them as new columns to the original data matrix, and then computing $\mathit{SSE}$ on the augmented data.
+  (x'_{ij} - \overline{x'}_j ^{(k)})^2.
+$$
+
+Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ again has the same form as the standard k-means criterion $\mathit{SSE}$, but it is computed on different data.
+Computing the k-plus criterion is then done by first creating the additional $t$-moment variables, appending them as new columns to the original data matrix, and computing $\mathit{SSE}$ on the augmented data.
 Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
@@
 This imbalance can strongly and unpredictably influence the relative weight of the different moments when computing $\mathit{SSE_{kplusGeneralized}}$.
-To ensure an equal weight of all criteria, all attributes---i.e., the original data as well as the new $t$-moment variables---can be standardized before computing $\mathit{SSE_{kplusGeneralized}}$, e.g.
+To ensure a comparable weight of all criteria, all attributes---i.e., the original data as well as the new $t$-moment variables---can be standardized before computing $\mathit{SSE_{kplusGeneralized}}$, e.g.
 via $z$-standardization.
@@
 The target was the optimization of $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, i.e.
-minizimizing differences with regard to means, variances, skew and kurtosis.
-Two unweighted approaches (using standardization vs.
-not using standardization) were implemented with $\mathbf{w} = (1, 1, 1, 1)$.
-The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented variables---of $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
-In the other case, the unstandardized matrix $\mathbf{\hat{X}'}_{N \times \mathit{TP}}$ was used.
-The four weighted approaches were implemented by giving a tenfold increased weight to only one of the four moments.
+minimizing differences with regard to means, variances, skew and kurtosis.
+Two unweighted approaches were implemented by defining $\mathbf{w} = (1, 1, 1, 1)$.
+The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented variables---of the input data $\mathbf{X'}_{N \times \mathit{TP}}$ before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
+In the second case, the unstandardized matrix $\mathbf{X'}_{N \times \mathit{TP}}$ was used.
+Four weighted approaches were implemented by giving a tenfold increased weight to one of the four moments, respectively.
 That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favoring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favoring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, repectively.
-The weighted approaches did not use standardization of the input data.
+The weighted approaches did not use a standardization of the input data.
 
@@
 
+# significance test
+sig_test <- prop.test(x = sum(unstd_best), n = length(unstd_best), p = 1/6)
+
 ```
@@
 It is shown that for the weighted approaches, the criterion that received the highest weight was fulfilled to the highest degree, as should be expected.
-Interestingly, the unweighted approach including standardization had the best average performance across the four criteria.
-While the advantage over the other weighting schemes does not seem to be remarkable, this pattern was observed in an astonishing `r sum(unstd_best)` of all `r length(results)` simulation runs (`r mean(unstd_best) * 100`%).
-Among the competing weighting schemes, the unweighted standardization approach therefore seems preferable.
-It also yielded the best results with regard to maximizing the standard $\mathit{SSE}$.
-The unweighted standardization approach seemed to implicitly give high priority to minimizing differences in means.
+Interestingly, the unweighted standardization approach had the best average performance across the four criteria.
+While the degree of advantage over the other weighting schemes by itself does not seem to be remarkable, this pattern was observed in an astonishing `r mean(unstd_best) * 100`% of all simulation runs (the level of chance would have been `r 1/6 * 100`%, `r format_p(sig_test$p.value)`).
+Among the competing weighting schemes, the unweighted standardization approach therefore yielded the most promising results.
+It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, comparable to the weighting scheme explicitly favoring $\mathit{SSE}$ (**TODO**: CHECK IF THESE RESULTS HOLD WHEN ALL DATA IS IN!).
+Hence, the unweighted standardization approach seemed to implicitly give high priority to minimizing differences in means.
 This is a pleasant result because, as discussed earlier, for the higher order moment $\mathit{SSE}$s to be valid, the group means have to be close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ has high priority.
@@
 After the five anticlustering methods and the random allocation procedure had been applied to the 10,000 data sets for $K = 2$, $K = 3$, $K = 4$, I investigated how well the competing methods were able to create similar groups.
-To this end, I analyzed the discrepancy in means, standard deviations, skewness, kurtosis, and correlation between groups---for each data set for each $K$.
+To this end, I analyzed the discrepancy in means, standard deviations, skewness and kurtosis between groups---for each data set for each $K$.
 To quantify the discrepancy in group means, the following computation was used: for each of the (2, 3, 4 or 5) features, all (2, 3 or 4) group means were computed.
@@
 The same procedure was applied to compute the discrepancy in standard deviations ($\Delta \mathit{SD}$), skewness ($\Delta \mathit{Skew}$) and kurtosis ($\Delta \mathit{Kurtosis}$).
-Quantifying discrepancy in correlations ($\Delta \mathit{Cor}$) followed the same rule, but discrepancies in correlations had to be averaged across pairs of features instead of single features.
 
@@
 ldf <- ldf[ldf$method != "k-plus-correlation", ] # removed because this deserves closer attention than what was given in the paper; check out the first version of the preprint for the previous results
+
+ldf$method[ldf$method == "k-plus-skew-kurtosis"] <- "k-plus generalized"
 

### 2023-04-19 weighting_scheme.R
removed initial approach for weight simulation

@@ -1,106 +1,0 @@
-library(anticlust)
-
-var_objective <- function(clusters, features) {
-  variance_objective(features, clusters)
-}
-
-# Compute features for k-plus criterion
-moment_features <- function(data, moment = 2) {
-  apply(data, 2, function(x) (x - mean(x))^moment)
-}
-
-weight_selection_sim <- function(X, mean_distr = 0, sd_distr = 1) {
-  N <- sample(c(12, 14, 16), size = 1)
-  K <- 2
-  M <- sample(1:5, size = 1)
-  partitions <- generate_partitions(N, K)
-  features <- matrix(rnorm(N * M, mean_distr, sd_distr), ncol = M)
-  
-  # Compute objectives for each partition
-  # For each partition: Compute SSE objective, and SSE(VAR)
-  # 1. SSE
-  all_objectives_kMeans <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = features
-  )
-  
-  squared_distances_from_mean <- moment_features(as.matrix(features))
-  cubic_distances_from_mean <- moment_features(as.matrix(features), moment = 3)
-  
-  # 2. SSE(VAR)
-  all_objectives_kVar <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = squared_distances_from_mean
-  )
-  
-  # 3. SSE(skew)
-  all_objectives_skew <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = cubic_distances_from_mean
-  )
-  
-  # 4. SSE(KPLUS)
-  all_objectives_kPlus <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = cbind(features, squared_distances_from_mean, cubic_distances_from_mean)
-  )
-  
-  # 5. SSE(KPLUS) - standardized
-  # Use standardization to compute k-plus criterion!
-  all_objectives_kPlus_std <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = scale(cbind(features, squared_distances_from_mean, cubic_distances_from_mean))
-  )
-  
-  df <- data.frame(
-    kmeans = all_objectives_kMeans,
-    kvar = all_objectives_kVar,
-    kskew = all_objectives_skew,
-    kplus = all_objectives_kPlus,
-    kplus_std = all_objectives_kPlus_std,
[…]

### 2023-04-19 Simulation_Study/Weight_Simulation/functions_weight_simulation.R
rerun simulation 100 times with 5 repetitions in algorithm

@@ -36,7 +36,7 @@ kplus_kurtosis_preferred <- function(features, clusters) {
 
 clusters_for_all_methods <- function(data, K) {
   all_results <- list()
-  repetitions <- 1
+  repetitions <- 5
   method <- "local-maximum"
   # antiweighted k-plus anticlustering really is just k-means anticlusteriung
   all_results[["unweighted_unstd"]] <- anticlustering(
@@ -105,4 +105,3 @@ weight_sim <- function(dat, K) {
   results <- t(t(objective_results) / apply(objective_results, 2, max))
   results 
 }
-

### 2023-04-19 Simulation_Study/Weight_Simulation/weight_simulation.R

@@ -3,7 +3,9 @@
 
 source("./Simulation_Study/Weight_Simulation/functions_weight_simulation.R")
 
-nsim <- 30
+set.seed(6758) # random.org gave me this one 
+
+nsim <- 100
 results <- list()
 start <- Sys.time()
 for (i in 1:nsim) {

### 2023-04-20 paper.Rmd
editing

@@
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis.
-After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional anticlustering criteria.
+After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional anticlustering criteria.
 Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].
@@
 
-In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variance between groups instead of only mean values.
+In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variance between groups instead of only means.
 To motivate the new criterion, we first consider a set of one-dimensional observations.
@@
   # 3. SSE(KPLUS)
-  # Use standardization to compute k-plus criterion!
   all_objectives_kPlus <- sapply(
@@
 Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict with each other [@ferligoj1992direct].
-As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria (mean and variance) are similar between groups.
+As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria are similar between groups.
 A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible.
@@
 Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances.
-The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ and $w_2$ were set to 1 when computing $\mathit{SSE_{kplus}}$), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
-Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree, indicating that these objectives tend to align and are not in conflict with each other.
-Moreover, the scatterplot in Figure 1 shows that there is a high concentration of partitions in the upper-right quadrant where the two objectives are in high accordance.
+The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ = $w_2$ = 1 in the computation of $\mathit{SSE_{kplus}}$), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
+Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree.
+This result indicates that the objectives $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are not in conflict with each other.
+Moreover, in Figure 1 there is a high concentration of partitions in the upper-right corner where the two objectives are in high agreement.
 
@@
 
-To formalize the computation of the k-plus criterion for an arbitraty number of distribution moments, some new notation has to be provided.
+To formalize the computation of the k-plus criterion for an arbitrary number of distribution moments, some new notation has to be provided.
 First, we define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean is $\overline{\mathbf{x}}_j$.
@@
 Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{x}_{tj} ^{(k)}$ is mean for the $k$'th group on the $j$'th $t$-moment variable.
-Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to equalize the first $T$ distribution moments between $K$ groups.
+Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to minimize differences with regard to the first $T$ distribution moments between $K$ groups.
 
@@
 
-When defining the row vectors of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i.} = (x'_{i1}, \ldots, x'_{i(TP)})$ for each $i = 1, \ldots, N$, the criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as
+When defining the row vectors of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i.} = (x'_{i1}, \ldots, x'_{i(TP)})$ ($i = 1, \ldots, N$), the criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as
 
@@
 The target was the optimization of $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, i.e.
-minimizing differences with regard to means, variances, skew and kurtosis.
+minimizing differences with regard to group means, variances, skewness and kurtosis.
 Two unweighted approaches were implemented by defining $\mathbf{w} = (1, 1, 1, 1)$.
@@
 To compare the different weighting schemes, `r length(results)` data sets from were generated from a normal distribution ($M = 0$, $\mathit{SD} = 1$).
-The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$.
+The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$ so that equal-sized groups could be created.
 For each data set, $\mathit{SSE_{kplusGeneralized}}$ was optimized using the local maximum algorithm, once for each of the six competing weighting schemes.
@@
 )
-note <- paste0("The results of Simulation 1 averaged across all ", length(results), " data sets.
-Cells indicate how well the maximization of the weighted $\\mathit{SSE_{kplusGeneralized}}$ fulfilled each of the four constituting criteria $\\mathit{SSE}$, $\\mathit{SSE_{Var}}$, $\\mathit{SSE_{Skew}}$, and $\\mathit{SSE_{Kurtosis}}$.
-Cell values represent relative performance, i.e.
-they were computed as the $\\mathit{SSE}$ that was attained by the given weighting scheme, divided by the best $\\mathit{SSE}$ that was obtained by any weighting scheme in a given simulation run.")
+note <- paste0("Cells indicate how well the maximization $\\mathit{SSE_{kplusGeneralized}}$ fulfilled each of the four constituting criteria $\\mathit{SSE}$, $\\mathit{SSE_{Var}}$, $\\mathit{SSE_{Skew}}$, and $\\mathit{SSE_{Kurtosis}}$, depending on the weighting scheme.
+For each simulation run, the relative performance was computed as the $\\mathit{SSE}$ that was obtained by the given weighting scheme, divided by the best $\\mathit{SSE}$ that was obtained by any weighting scheme in this simulation run.
+Cell values represent the average relative performance across all ", length(results), " simulation runs.")
 
@@
   escape = FALSE, 
-  caption = "Results of Simulation study 1",
+  caption = "Relative performance of the competing weighting schemes in Simulation 1.",
   note = note,
@@
 
-Table 1 illustrates how well each of the six weighting schemes fulfilled each of the four constititung criteria, averaged across the `r length(results)` simulation runs.
-It is shown that for the weighted approaches, the criterion that received the highest weight was fulfilled to the highest degree, as should be expected.
+Table 1 illustrates how well each of the six weighting schemes fulfilled each of the four constitutung criteria, averaged across the `r length(results)` simulation runs.
+First, the results show that for all weighting schemes, optimizing the combined criterion $\mathit{SSE_{kplusGeneralized}}$ yielded high performance with regard to all four criteria.
+Thus, the various k-plus criteria seem to align well when optimizing a combined criterion, which is good news for the k-plus approach to anticlustering.
+For the weighted approaches, the criterion that had received the highest weight was fulfilled to the highest degree, as should be expected.
 Interestingly, the unweighted standardization approach had the best average performance across the four criteria.
-While the degree of advantage over the other weighting schemes by itself does not seem to be remarkable, this pattern was observed in an astonishing `r mean(unstd_best) * 100`% of all simulation runs (the level of chance would have been `r 1/6 * 100`%, `r format_p(sig_test$p.value)`).
+While the degree of advantage over the other weighting schemes by itself does not seem to be remarkable, this pattern was observed in an astonishing `r mean(unstd_best) * 100`% of all simulation runs, while the level of chance would have been `r 1/6 * 100`% (`r format_p(sig_test$p.value)`).
 Among the competing weighting schemes, the unweighted standardization approach therefore yielded the most promising results.
@@
 
-A simulation study and examples on real norming data showed that the k-plus criterion is well-suited to maximize between-group similarity with regard to multiple criteria.
+Two simulation studies and examples on real norming data showed that the k-plus criterion is well-suited to maximize between-group similarity with regard to multiple criteria.
 Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be considered as the default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be challenging in small data sets.
@@
 
-The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion.
+The present paper focused on the new k-plus criterion for anticlustering, but only briefly considered the multicriterion nature of this criterion.
 A simple unweighted sum approach was used to optimize the k-plus objective, while more sophisticated algorithmic procedures exist to tackle multicriterion optimization problems [e.g., @brusco2009cross].
 The sole focus on the objective function followed two reasonings.
-First, the applications showed that the local maximum search, maximizing the unweighted k-plus sum, yielded satisfying results that probably suffice in most applications.
-It is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria to a satisfactory degree (as suggested by Figure 1).
+First, the applications and the simulations showed that the unweighted standardization approach yielded satisfying results that probably suffice in most applications.
+Moreover, it is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria to a satisfactory degree (as suggested by Figure 1).
 Still, more sophisticated approaches may yield theoretically improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
@@
 
-It should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariance, a possibility that was not investigated in the current paper.
+As a last outlook, it should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariances, a possibility that was not investigated in the current paper.
 The covariance between two variables $x = (x_1, \ldots, x_N)$ and $y = (y_1, \ldots, y_N)$ is defined as $\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.

### 2023-04-20 paper.Rmd
they do!

@@
 Among the competing weighting schemes, the unweighted standardization approach therefore yielded the most promising results.
-It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, comparable to the weighting scheme explicitly favoring $\mathit{SSE}$ (**TODO**: CHECK IF THESE RESULTS HOLD WHEN ALL DATA IS IN!).
+It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, comparable to the weighting scheme explicitly favoring $\mathit{SSE}$.
 Hence, the unweighted standardization approach seemed to implicitly give high priority to minimizing differences in means.

### 2023-04-20 paper.Rmd
i guess this cannot be said based on this design

@@
 
-Table 1 illustrates how well each of the six weighting schemes fulfilled each of the four constitutung criteria, averaged across the `r length(results)` simulation runs.
-First, the results show that for all weighting schemes, optimizing the combined criterion $\mathit{SSE_{kplusGeneralized}}$ yielded high performance with regard to all four criteria.
-Thus, the various k-plus criteria seem to align well when optimizing a combined criterion, which is good news for the k-plus approach to anticlustering.
-For the weighted approaches, the criterion that had received the highest weight was fulfilled to the highest degree, as should be expected.
+Table 1 illustrates how well each of the six weighting schemes fulfilled each of the four constituting criteria, averaged across the `r length(results)` simulation runs.
+For the weighted approaches, the criterion that received the highest weight during the optimization was fulfilled to the highest degree, as should be expected.
 Interestingly, the unweighted standardization approach had the best average performance across the four criteria.

### 2023-04-20 paper.Rmd
editing sim 1

@@
   escape = FALSE, 
-  caption = "Relative performance of the competing weighting schemes in Simulation 1.",
+  caption = "Average relative performance of the competing weighting schemes in Simulation 1.",
   note = note,
@@
 Interestingly, the unweighted standardization approach had the best average performance across the four criteria.
-While the degree of advantage over the other weighting schemes by itself does not seem to be remarkable, this pattern was observed in an astonishing `r mean(unstd_best) * 100`% of all simulation runs, while the level of chance would have been `r 1/6 * 100`% (`r format_p(sig_test$p.value)`).
+While the degree of advantage by itself does not seem to be remarkable, this pattern was observed in an astonishing `r mean(unstd_best) * 100`% of all simulation runs, while the level of chance would have been `r 1/6 * 100`% (`r format_p(sig_test$p.value)`).
 Among the competing weighting schemes, the unweighted standardization approach therefore yielded the most promising results.
-It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, comparable to the weighting scheme explicitly favoring $\mathit{SSE}$.
+It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, even slightly outperforming the weighting scheme that explicitly favored $\mathit{SSE}$.
 Hence, the unweighted standardization approach seemed to implicitly give high priority to minimizing differences in means.

### 2023-04-20 paper.Rmd
editing

@@
 The partitioning is subject to the requirement that each element is assigned to exactly one group.
-In anticlustering applications, an additional constraint is usually imposed on the group sizes, with the most common restriction being that all groups have equal size [@papenberg2020].
+In anticlustering applications, an additional constraint is usually imposed on the group sizes $n_k$ ($k = 1, \ldots K$), with the most common restriction being that all groups have equal size, i.e.
+$n_k = \frac{N}{K}$ [@papenberg2020].
 
@@
 
-The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all $n_k$ observations belonging to the $k$'th cluster $C_k$ ($k = 1, \ldots, K$):
+The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all observations belonging to the $k$'th cluster $C_k$ ($k = 1, \ldots, K$):
 
@@
 
-In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variance between groups instead of only means.
+In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variances between groups instead of only means.
 To motivate the new criterion, we first consider a set of one-dimensional observations.
@@
 That is, we compute new variables for each original variable and for each distribution moment that should be considered in the anticlustering process.
-
-To formalize the computation of the k-plus criterion for an arbitrary number of distribution moments, some new notation has to be provided.
-First, we define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean is $\overline{\mathbf{x}}_j$.
+Analogously to the computation of, we could then define a $\mathit{SSE_{Skew}}$ for $t = 3$ or a $\mathit{SSE_{Kurtosis}}$ for $t = 4$.
+However, to formalize the computation of the k-plus criterion for an arbitrary number of distribution moments, some new notation has to be provided.
+
+We define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean is $\overline{\mathbf{x}}_j$.
 Let $T$ be the number of distribution moments that should be equalized between $K$ groups.
@@
 
-Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ again has the same form as the standard k-means criterion $\mathit{SSE}$, but it is computed on different data.
+Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ has the same form as the standard k-means criterion $\mathit{SSE}$, but it is computed on different data.
 Computing the k-plus criterion is then done by first creating the additional $t$-moment variables, appending them as new columns to the original data matrix, and computing $\mathit{SSE}$ on the augmented data.
@@
 Note that a problem may arise if the weights $\mathbf{w}$ are ignored when computing $\mathit{SSE_{kplusGeneralized}}$.
-Because the computation of the new $t$-moment variables involves exponentiation, their range of values is expected to be very different from the raw values $x_{ij}$ (e.g., see Figure 1).
-This imbalance can strongly and unpredictably influence the relative weight of the different moments when computing $\mathit{SSE_{kplusGeneralized}}$.
+Because the computation of the new $t$-moment variables involves exponentiation, their range of values is expected to be very different from the original variables.
+This imbalance can strongly and unpredictably influence the relative weight of the different moments when computing $\mathit{SSE_{kplusGeneralized}}$ (e.g., see Figure 1).
 To ensure a comparable weight of all criteria, all attributes---i.e., the original data as well as the new $t$-moment variables---can be standardized before computing $\mathit{SSE_{kplusGeneralized}}$, e.g.
@@
 Four weighted approaches were implemented by giving a tenfold increased weight to one of the four moments, respectively.
-That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favoring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favoring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, repectively.
+That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favoring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favoring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, respectively.
 The weighted approaches did not use a standardization of the input data.
@@
 The argument `K` describes the number of groups.
-By default, the function ensures that all features are standardized before the optimization starts; in particular, this option enforces that all k-plus criteria receive the same weight during the optimization process.
-This behaviour can be adjusted using the boolean argument `standardize`.
+By default, the function ensures that all features---the original and the augmented variables---are standardized before the optimization starts.
+Thus, by default, the unweighted standardization approach is used.
+Whether standardization is used can be adjusted using the boolean argument `standardize`.
 

### 2023-04-20 paper.Rmd
some editing

@@
 That is, we compute new variables for each original variable and for each distribution moment that should be considered in the anticlustering process.
-Analogously to the computation of, we could then define a $\mathit{SSE_{Skew}}$ for $t = 3$ or a $\mathit{SSE_{Kurtosis}}$ for $t = 4$.
+Analogously to the computation of $\mathit{SSE_{Var}}$, we could then define an $\mathit{SSE_{Skew}}$ using $t = 3$ or an $\mathit{SSE_{Kurtosis}}$ using $t = 4$.
 However, to formalize the computation of the k-plus criterion for an arbitrary number of distribution moments, some new notation has to be provided.
@@
 
-Using the tensor $\mathbf{\hat{X}}_{T \times N \times P}$ and a vector of weights $\mathbf{w} = (w_1, \dots, w_T$), the generalized multicriterion for k-plus anticlustering $\mathit{SSE_{kplusGeneralized}}$ can be formalized:
+Using $\mathbf{\hat{X}}_{T \times N \times P}$ and a vector of weights $\mathbf{w} = (w_1, \dots, w_T$), the generalized multicriterion for k-plus anticlustering $\mathit{SSE_{kplusGeneralized}}$ can be formalized:
 
@@
 
-Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{x}_{tj} ^{(k)}$ is mean for the $k$'th group on the $j$'th $t$-moment variable.
+Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{x}_{tj} ^{(k)}$ is the $k$'th group mean on the $j$'th $t$-moment variable.
 Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to minimize differences with regard to the first $T$ distribution moments between $K$ groups.
@@
 This simplification leads to the *unweighted sum approach* for k-plus anticlustering.
-In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$, representing distribution moments can be dropped; instead, the tensor is flattened into a two dimensional matrix, where the variables representing the additional distribution moments are appended to the original data matrix $\mathbf{X}_{N \times P}$, resulting in the augmented matrix $\mathbf{X'}_{N \times \mathit{TP}}$:
+In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$, representing distribution moments can be dropped; instead, the tensor is flattened into a two-dimensional matrix, where the variables representing the additional distribution moments are appended to the original data matrix $\mathbf{X}_{N \times P}$, resulting in the augmented matrix $\mathbf{X'}_{N \times \mathit{TP}}$:
 

### 2023-04-20 paper.Rmd
explain notation for mean values in t moment variables

@@
   \sum\limits_{i \in C_k}
-  (x_{tij} - \overline{x}_{tj} ^{(k)})^2
+  (\hat{x}_{tij} - \overline{x}_{tj} ^{(k)})^2
   \right)
@@
 
-Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{x}_{tj} ^{(k)}$ is the $k$'th group mean on the $j$'th $t$-moment variable.
+Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{\mathbf{X}}^{(k)}_{T \times P} = \{\overline{x}_{tj}^{(k)}\}_{T \times P}$ is a centroid matrix containing the mean values for all $t$-moment variables in the $k$'th group, i.e.
+
+$$
+\overline{x}_{tj}^{(k)} = \frac{1}{n_k} \sum_{i \in C_k} \hat{x}_{tij}
+$$
+
 Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to minimize differences with regard to the first $T$ distribution moments between $K$ groups.

### 2023-04-20 Simulation_Study/Weight_Simulation/functions_weight_simulation.R
typo

@@ -38,7 +38,7 @@ clusters_for_all_methods <- function(data, K) {
   all_results <- list()
   repetitions <- 5
   method <- "local-maximum"
-  # antiweighted k-plus anticlustering really is just k-means anticlusteriung
+  # unweighted k-plus anticlustering really is just k-means anticlustering
   all_results[["unweighted_unstd"]] <- anticlustering(
     cbind(data, moment_features(data, 2), moment_features(data, 3), moment_features(data, 4)),
     K = K,

### 2023-04-20 Simulation_1/1_weight_simulation.R
restructured

@@ -1,7 +1,7 @@
 
 
 
-source("./Simulation_Study/Weight_Simulation/functions_weight_simulation.R")
+source("0_functions_weight_simulation.R")
 
 set.seed(6758) # random.org gave me this one 
 
@@ -20,4 +20,4 @@ for (i in 1:nsim) {
 }
 print(Sys.time() - start)
 
-save(results, file = paste0("./Simulation_Study/Weight_Simulation/Results Sim 1 ", date(), ".Rdata"))
+save(results, file = paste0("Results Sim 1 ", date(), ".Rdata"))

### 2023-04-20 Simulation_1/2_analysis.R
added analysis of Simulatio 1 as script

@@ -1,0 +1,18 @@
+# Read simulation results
+# (in case there are several results files)
+files <- list.files(".", pattern = ".Rdata")
+results <- list()
+for (file in files) {
+  results <- c(results, get(load(file)))
+}
+
+# Table reported in the manuscript: 
+round(Reduce("+", results) / length(results) * 100, 2)
+
+# how often was unweighted standardization approach best:
+unstd_best <- sapply(results, function(x) names(which.max(x[, 5]))) == "unweighted_std"
+sum(unstd_best)
+mean(unstd_best)
+
+# Significance test
+prop.test(x = sum(unstd_best), n = length(unstd_best), p = 1/6)

### 2023-04-25 paper.Rmd
changed directories of simulation results in paper file

@@
 ```{r getresultsSim1}
-files <- list.files("./Simulation_Study/Weight_Simulation", pattern = ".Rdata")
+dir <- "./Simulation_1/"
+files <- list.files(dir, pattern = ".Rdata")
 results <- list()
 for (file in files) {
-  results <- c(results, get(load(paste0("Simulation_Study/Weight_Simulation/", file))))
+  results <- c(results, get(load(paste0(dir, file))))
 }
@@
 for (K in 2:4) {
-  filename <- paste0("./Simulation_Study/results-K", K, "-objectives-raw.csv")
+  filename <- paste0("./Simulation_2/results-K", K, "-objectives-raw.csv")
   df <- read.csv(filename, sep = ";", stringsAsFactors = FALSE)

### 2023-04-25 paper.Rmd
this is better

@@
 
-When defining a new variable $y = (y_1, \ldots, y_N)$ as the squared deviation of each observation from the mean $\overline{\mathbf{x}}$, the sample variance of $x$ can be reformulated as the mean of $y$:
-
-$$
-y_i = (x_i - \overline{\mathbf{x}})^2
-$$
+When defining a new variable $y = (y_1, \ldots, y_N)$ as the squared deviation of each observation from the mean $\overline{\mathbf{x}}$, i.e.
+$y_i = (x_i - \overline{\mathbf{x}})^2$, the sample variance of $x$ can be reformulated as the mean of $y$:
 

### 2023-04-25 paper.Rmd
editing

@@
 An exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is not feasible---and most likely not interesting on its own---without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
-Hence, when striving for similarity in group variances by maximizing $\mathit{SSE_{Var}}$, it is at the same time necessary and worthwhile to strive for similarity in means, and thus to maximize $\mathit{SSE}$ at the same time.
 The most simple approach for an optimization that considers multiple criteria at the same time is the weighted sum approach [@naidu2014; @marler2010].
@@
 )
-note <- paste0("Cells indicate how well the maximization $\\mathit{SSE_{kplusGeneralized}}$ fulfilled each of the four constituting criteria $\\mathit{SSE}$, $\\mathit{SSE_{Var}}$, $\\mathit{SSE_{Skew}}$, and $\\mathit{SSE_{Kurtosis}}$, depending on the weighting scheme.
+note <- paste0("Cells indicate how well the maximization of $\\mathit{SSE_{kplusGeneralized}}$ fulfilled each of the four constituting criteria $\\mathit{SSE}$, $\\mathit{SSE_{Var}}$, $\\mathit{SSE_{Skew}}$, and $\\mathit{SSE_{Kurtosis}}$, depending on the weighting scheme.
 For each simulation run, the relative performance was computed as the $\\mathit{SSE}$ that was obtained by the given weighting scheme, divided by the best $\\mathit{SSE}$ that was obtained by any weighting scheme in this simulation run.
@@
 
-This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] using the function `kplus_anticlustering()`, which implements the unweighted sum approach for k-plus anticlustering.
-The package can be installed in the R environment using the `install.packages()` command:
+This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020].
+The package can be installed in the R environment using the `install.packages()` command.
 
@@
 ```
+
+The function `kplus_anticlustering()` available in `anticlust` implements the unweighted sum approach for k-plus anticlustering.
 

### 2023-04-26 paper.Rmd
editing

@@
 by computing all pairwise Euclidean distances, and then optimize a distance-based anticlustering criterion to obtain a partitioning.
+For likert scale data or binary attributes, the Manhattan distance may be preferred over the Euclidean distance, and for mixed-type attributes, it is even possible to employ a combined distance metric [@rokach2005].
 
@@
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis.
-After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional anticlustering criteria.
+After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming the traditional anticlustering criteria diversity and k-means.
 Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].
@@
 As a caveat, it is crucial to note that maximizing $\mathit{SSE_{Var}}$ only reliably yields similar group variances if the group centroids $\overline{\mathbf{x}}^{(k)}$ ($k = 1, \dots, K$) are close to the overall data centroid $\overline{\mathbf{x}}$.
-To reiterate, this condition has to be met for the original data $\mathbf{X}_{N \times P}$.
-Thus, only considering $\mathbf{Y}_{N \times P}$ in order to minimize discrepancies in variance may not yield satisfying results; instead, it has to be ensured that the group means of the original data are also similar between groups.
+This is because the within-group variance is calculated on the basis of the group mean $\overline{\mathbf{x}}^{(k)}$, but the additional variables computed to minimize differences with regard to the variance across groups are calculated using the overall mean $\overline{\mathbf{x}}$.
+Thus, $\mathit{SSE_{Var}}$ is only valid if (approximately) $\overline{\mathbf{x}}$ = $\overline{\mathbf{x}}^{(k)}$; otherwise we may introduce non-negligible measurement error when optimizing $\mathit{SSE_{Var}}$.
+Thus, only considering the new variables $\mathbf{Y}_{N \times P}$ in order to minimize discrepancies in variance may not yield satisfying results; instead, it has to be ensured that the group means of the original variables are also similar between groups.
 Fortunately, minimizing discrepancy with regard to group means can be accomplished by standard k-means anticlustering, i.e., the maximization of $\mathit{SSE}$.
@@
 
-## A multicriterion optimization model: K-plus anticlustering {#kplus}
+## A bicriterion model: K-plus anticlustering {#kplus}
 
@@
 
-We define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean is $\overline{\mathbf{x}}_j$.
-Let $T$ be the number of distribution moments that should be equalized between $K$ groups.
-For each moment $t$ ($t = 1, \ldots, T$) and for each variable $x_{.j}$, a new *t-moment variable* $\hat{x}_{t.j} = (\hat{x}_{t1j}, \ldots, \hat{x}_{tNj})$ is defined where
+We define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{\cdot j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean---the overall mean of the $j$'th attribute---is $\overline{\mathbf{x}}_j$.
+Let $T$ be the number of distribution moments that should be similar across $K$ groups.
+For each moment $t$ ($t = 1, \ldots, T$) and for each variable $x_{\cdot j}$ ($j = 1, \ldots, P$), a new *t-moment variable* $\hat{x}_{t \cdot j} = (\hat{x}_{t1j}, \ldots, \hat{x}_{tNj})$ is defined where
 
@@
 
-When defining the row vectors of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i.} = (x'_{i1}, \ldots, x'_{i(TP)})$ ($i = 1, \ldots, N$), the criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as
+When defining the rows of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i \cdot} = (x'_{i1}, \ldots, x'_{i(TP)})$ ($i = 1, \ldots, N$), the criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as
 
@@
 Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ has the same form as the standard k-means criterion $\mathit{SSE}$, but it is computed on different data.
-Computing the k-plus criterion is then done by first creating the additional $t$-moment variables, appending them as new columns to the original data matrix, and computing $\mathit{SSE}$ on the augmented data.
+Computing the k-plus criterion is then done by first creating the additional $t$-moment variables, appending them as new columns to the original data matrix, and computing the $\mathit{SSE}$ on the augmented data.
 Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
@@
 
-To thoroughly evaluate the k-plus approach to anticlustering, two specific k-plus criteria were compared to traditional anticlustering criteria in a second simulation study: (i) "standard" k-plus anticlustering, minimizing differences with regard to group means and variances (by maximizing $\mathit{SSE_{kplus}}$); (ii) "generalized" k-plus anticlustering, minimizing differences with regard to group means, variances, skewness and kurtosis (by maximizing $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$).
+To thoroughly evaluate the k-plus approach to anticlustering, two specific k-plus criteria were compared to two traditional anticlustering criteria: (a) "standard" k-plus anticlustering, minimizing differences with regard to group means and variances (by maximizing $\mathit{SSE_{kplus}}$); (b) "generalized" k-plus anticlustering, minimizing differences with regard to group means, variances, skewness and kurtosis (by maximizing $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$).
 To implement the k-plus objectives, the unweighted standardization approach was employed that performed best in Simulation 1.

### 2023-04-27 paper.Rmd
minor editing

@@
 by computing all pairwise Euclidean distances, and then optimize a distance-based anticlustering criterion to obtain a partitioning.
-For likert scale data or binary attributes, the Manhattan distance may be preferred over the Euclidean distance, and for mixed-type attributes, it is even possible to employ a combined distance metric [@rokach2005].
+For likert scale data or binary attributes, the Manhattan distance may be preferred over the Euclidean distance; for mixed-type attributes, it is even possible to employ a combined distance metric [@rokach2005].
 
@@
 
-As a last outlook, it should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariances, a possibility that was not investigated in the current paper.
-The covariance between two variables $x = (x_1, \ldots, x_N)$ and $y = (y_1, \ldots, y_N)$ is defined as $\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
-Thus, to maximize between-group similarity with regard to covariance structure, it is possible to augment the data set with an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
-Future research should examine how the k-plus approach can be adapted to including covariances, including a discussion of how to select weights for the possibly high number of additional variables.
+As an additional outlook, it should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariances, a possibility that was not investigated in the current paper.
+In this case, an additional variable has to be computed for each pair of variables $x = (x_1, \dots, x_N)$ and $y = (y_1, \dots, y_N)$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
+Future research should examine how the k-plus approach can be adapted to including covariances.
 The unweighted standardization approach presented here may serve as a starting point for investigation.

### 2023-04-27 paper.Rmd
including discussion on fair clustering and ARI (sort of)

@@
   )
+
+  # investigate the best 1% of k-means anticlustering partitions, for discussion
+  best_1percent_anticlusterings <- df$kmeans /max(df$kmeans) > .99
+  best_partitions <- partitions[best_1percent_anticlusterings]
+
+  neighbor_matrix <- anticlust:::selection_matrix_from_clusters
+  neighbourhoods <- lapply(best_partitions, neighbor_matrix)
+  neighbourhoods <- lapply(neighbourhoods, function(x) as.integer(x[lower.tri(x)]))
+  pairs <- Reduce("+", neighbourhoods)
+  # length(pairs)
+  # length(neighbourhoods[[1]])
+  # choose(14, 2)
+  # min(pairs) # 131
+  # max(pairs) # 292
+  # length(best_partitions) # 453
+  # -> all items are grouped with one another, and no items are always grouped with one another
+
 } else {
@@
 
+## Closing remarks on the nature of anticlustering
+
+```{r}
+df <- read.csv("partition_objectives.csv")
+```
+
+Traditionally, anticlustering is implemented by reversing an objective function used in classical cluster analysis.
+Whereas cluster analysis seeks homogeneity within groups and separation between groups, this reversal leads to heterogeneity within groups and similarity between groups.
+While tightly linked in a mathematical sense, an important semantic difference is that cluster analysis is concerned with finding "true" structure in the data e.g., by defining a taxonomy on naturally occurring groups [@everitt1979].
+Anticlustering instead imposes structure on the data, while the anticlusters have no underlying "true" meaning.
+An interesting side effect is that we are usually not concerned with the question of whether pairs of elements are assigned to the same anticluster or not.
+In cluster analysis, similar elements tend to be grouped in the same cluster, and in general this result will be observed across a variety of clustering approaches.
+In contrast, different anticlustering partitions tend to look very different.
+In Figure 1, for example, an astonishing `r mean(df$kmeans / max(df$kmeans) >.99) * 100`% of all partitions are within 1% of the global maximum in $\mathit{SSE}$ (i.e., the best k-means anticlustering solution).
+Among these `r sum(df$kmeans / max(df$kmeans) >.99)` partitions, every possible pair of elements is sometimes in the same group and sometimes not; the minimum amount of co-occurrence is 131 and the maximum amount of co-occurrence is 292.
+Thus, a high anticlustering objective does not depend on assigning individual items to the same or a different group.
+In contrast, there are only `r sum(min(df$kmeans)/ df$kmeans > .90)` partitions (`r mean(min(df$kmeans)/ df$kmeans > .90)*100`%) that approximate the global minimum in $\mathit{SSE}$ (i.e., the best k-means clustering solution) by even 10%.
+Thus, while there are many different high quality anticlustering partitions, there are only very few good clustering partitions, constituting another relevant difference between cluster analysis and anticlustering, despite them being directly and reversely related.
+Interestingly, the k-plus objective further loosens the link between cluster analysis and anticlustering.
+Reversing the k-plus anticlustering objective no longer yields a useful clustering interpretation, because identifying groups that are different with regard to higher order moments does not constitute a clustering in the usual sense.
+Thus, even though k-plus extends the classical k-means clustering criterion, it is a pure anticlustering objective.
+
+An interesting mixture of cluster analysis and anticlustering is fair clustering [e.g.
+@chhabra2021; @delBarrio2022].
+Fair clustering performs a cluster analysis under the restriction that the distribution of a set of "protected" attributes---such as race, sex or age---should be similar between clusters.
+Thus, as compared to pure cluster analysis, fair clustering biases the identification of groups with regard to the protected attributes, striving for algorithmic fairness.
+However, a relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data [@delBarrio2022], while anticlustering is not concerned with identifying the underlying structure of a data set.
+However, these methods are similar because anticlustering logically corresponds to fair clustering when only the protected attributes are considered.
+Hence, a fruitful outlook for research might be to investigate if approaches stemming from the literature on fair clustering can be adopted for anticlustering, and vice versa.
+
 ## Conclusion

### 2023-04-27 paper.Rmd
editing

--- 2023-04-27 paper.Rmd
@@
   )
+
+  # investigate the best 0.5% of k-means anticlustering partitions, for discussion
+  # best_0.1percent_anticlusterings <- df$kmeans /max(df$kmeans) > .999
+  # best_partitions <- partitions[best_0.1percent_anticlusterings]
+  # 
+  # neighbor_matrix <- anticlust:::selection_matrix_from_clusters
+  # neighbourhoods <- lapply(best_partitions, neighbor_matrix)
+  # neighbourhoods <- lapply(neighbourhoods, function(x) as.integer(x[lower.tri(x)]))
+  # pairs <- Reduce("+", neighbourhoods)
+  # length(pairs)
+  # length(neighbourhoods[[1]])
+  # choose(14, 2)
+  # min(pairs) # 40
+  # max(pairs) # 95
+  # length(best_partitions) # 146
+  # -> all items are grouped with one another, and no items are always grouped with one another
+
 } else {
@@
 
+## Closing remarks on the nature of anticlustering
+
+```{r}
+df <- read.csv("partition_objectives.csv")
+```
+
+Traditionally, anticlustering is implemented by reversing an objective function used in classical cluster analysis.
+Whereas cluster analysis seeks homogeneity within groups and separation between groups, the reversal results in heterogeneity within groups and similarity between groups.
+While tightly linked in a mathematical sense, an important semantic difference is that cluster analysis is concerned with finding "true" structure in the data, e.g., by defining a taxonomy on naturally occurring groups [@everitt1979].
+Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern.
+Correspondingly, different similarly adequate anticlustering partitions tend to look very different, while in cluster analysis, similar elements always tend to be grouped in the same cluster.
+In Figure 1, for example, we observe a strong concentration of anticlustering partitions: An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution.
+Among these `r sum(df$kmeans / max(df$kmeans) >.999)` partitions, every possible pair of elements is sometimes assigned to the same group and sometimes not; the minimum amount of co-occurrence is 40 and the maximum amount of co-occurrence is 95.
+Thus, a high anticlustering objective does not depend on assigning individual items either to the same or a different group.
+In contrast, there are only `r sum(min(df$kmeans)/ df$kmeans > .90)` partitions (`r mean(min(df$kmeans)/ df$kmeans > .90)*100`%) that approximate the global minimum in $\mathit{SSE}$ (i.e., the best k-means clustering solution) by even 10%.
+The clustering quality crucially depends on assigning similar elements to the same group.
+Hence, while there are many different high quality anticlustering partitions, there are only very few adequate clustering partitions, constituting another relevant difference between cluster analysis and anticlustering.
+Interestingly, the k-plus objective further loosens the link between cluster analysis and anticlustering.
+Reversing the k-plus anticlustering objective no longer yields a useful clustering interpretation, because identifying groups that are different with regard to higher order moments does not constitute a clustering in the usual sense.
+Thus, even though k-plus extends the classical k-means clustering criterion, it is a pure anticlustering objective.
+
+An interesting mixture of cluster analysis and anticlustering is fair clustering [e.g.
+@chhabra2021; @delBarrio2022].
+Fair clustering performs a cluster analysis under the restriction that the distribution of a set of "protected" attributes---such as race, sex or age---should be similar between clusters.
+Thus, as compared to pure cluster analysis, fair clustering biases the identification of groups with regard to the protected attributes, striving for algorithmic fairness.
+However, a relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data [@delBarrio2022], while anticlustering is not concerned with identifying the underlying structure of a data set.
+However, these methods are similar because anticlustering logically corresponds to fair clustering when only the protected attributes are considered.
+Hence, a fruitful outlook for research might be to investigate if approaches stemming from the literature on fair clustering can be adopted for anticlustering, and vice versa.
+
 ## Conclusion

### 2023-04-27 script_kplus_clustering.R
interesting

@@ -1,0 +1,81 @@
+# Compute features for k-plus criterion
+
+library(anticlust)
+moment_features <- function(data, moment = 2) {
+  apply(data, 2, function(x) (x - mean(x))^moment)
+}
+
+kmeans_clustering_objective <- function(features, clusters) {
+  -variance_objective(features, clusters)
+}
+
+kplus_clustering_objective <- function(features, clusters) {
+  variance_vars <- moment_features(features, 2)
+  all_vars <- scale(cbind(variance_vars))
+  -variance_objective(all_vars, clusters)
+}
+
+# Load data
+data(iris)
+cluster_data <- iris[, 1:2]
+
+# Do k-means clustering, store the clustering vector
+clusters <- anticlustering(
+  cluster_data,
+  K = 3,
+  objective = kmeans_clustering_objective,
+  method = "local-maximum"
+)
+
+# define 3 colors, 3 cex values, and 3 pch values that differ between
+# clusters - for more clusters, define more values.
+colors <- c("#a9a9a9", "#df536b", "#61d04f")
+cex <-  c(0.7, 1.2, 1.5)
+pch <- c(19, 15, 17)
+
+# Plot the data while visualizing the different clusters
+plot(
+  cluster_data,
+  col = colors[clusters],
+  cex = cex[clusters],
+  pch = pch[clusters]
+)
+
+# Do k-means clustering, store the clustering vector
+kplus_clusters <- anticlustering(
+  cluster_data,
+  K = 3,
+  objective = kplus_clustering_objective,
+  method = "local-maximum"
+)
+
+# Plot the data while visualizing the different clusters
+plot(
+  cluster_data,
+  col = colors[kplus_clusters],
+  cex = cex[kplus_clusters],
+  pch = pch[kplus_clusters]
+)
+
+
+
+###################
+
+library(anticlust)
+library(fossil)
+set.seed(12345)
+
+g1 <- anticlustering(iris[, -5], K = 3, objective = "kplus")
+g2 <- anticlustering(iris[, -5], K = 3, objective = "kplus")
+
+# Compare the two anticlustering solutions:
+rand.index(g1, g2)
+# Compare to random grouping:
+rand.index(sample(g1), sample(g2))
+
+# Compare to two clustering solutions: 
+c1 <- kmeans(iris[, -5], 3)$cluster
+c2 <- balanced_clustering(iris[, -5], 3)
+
+# High agreement, even though c2 is based on the restriction of equal group sizes:
+rand.index(c1, c2)

### 2023-04-28 paper.Rmd
editing

@@
 
-Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict with each other [@ferligoj1992direct].
-As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria are similar between groups.
+Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict with each other (Ferligoj & Batagelj, 1992).
+As $\mathit{SSE_{kplus}}$  is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria are similar between groups, i.e., whether these objectives are in alignment or conflicting.
 A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible.

### 2023-04-28 paper.Rmd
more editing discussion

@@
 Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern.
-Correspondingly, different similarly adequate anticlustering partitions tend to look very different, while in cluster analysis, similar elements always tend to be grouped in the same cluster.
-In Figure 1, for example, we observe a strong concentration of anticlustering partitions: An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution.
-Among these `r sum(df$kmeans / max(df$kmeans) >.999)` partitions, every possible pair of elements is sometimes assigned to the same group and sometimes not; the minimum amount of co-occurrence is 40 and the maximum amount of co-occurrence is 95.
-Thus, a high anticlustering objective does not depend on assigning individual items either to the same or a different group.
+Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, while good clustering solutions tend to look similarly.
+A good clustering solution depends on assigning similar elements into the same cluster and dissimilar elements into different clusters.
+A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.
+This difference between cluster analysis and anticlustering can also be clarified using Figure 1.
+Here, we observe a high concentration of suitable anticlustering partitions: An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution.
+Among these `r sum(df$kmeans / max(df$kmeans) >.999)` partitions, every possible pair of elements is sometimes assigned to the same group and sometimes not.
 In contrast, there are only `r sum(min(df$kmeans)/ df$kmeans > .90)` partitions (`r mean(min(df$kmeans)/ df$kmeans > .90)*100`%) that approximate the global minimum in $\mathit{SSE}$ (i.e., the best k-means clustering solution) by even 10%.
-The clustering quality crucially depends on assigning similar elements to the same group.
-Hence, while there are many different high quality anticlustering partitions, there are only very few adequate clustering partitions, constituting another relevant difference between cluster analysis and anticlustering.
+Thus, even though the problems are based on the same objective function, cluster analysis and anticlustering exhibit a stark discrepancy with regard to the solution space of partitions: There are many suitable anticlusterings but only few suitable clusterings.
 Interestingly, the k-plus objective further loosens the link between cluster analysis and anticlustering.
@@
 Thus, as compared to pure cluster analysis, fair clustering biases the identification of groups with regard to the protected attributes, striving for algorithmic fairness.
-However, a relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data [@delBarrio2022], while anticlustering is not concerned with identifying the underlying structure of a data set.
-However, these methods are similar because anticlustering logically corresponds to fair clustering when only the protected attributes are considered.
+A relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data [@delBarrio2022], while anticlustering is not concerned with identifying structure in the data; instead, anticlustering divides data sets into artificial subgroups in such a way that between-group similarity / within-group heterogeneity is maximized.
+However, these methods are logically similar because anticlustering corresponds to fair clustering when only considering the protected attributes.
 Hence, a fruitful outlook for research might be to investigate if approaches stemming from the literature on fair clustering can be adopted for anticlustering, and vice versa.

### 2023-04-28 paper.Rmd
still editing discussion

@@
 
-Traditionally, anticlustering is implemented by reversing an objective function used in classical cluster analysis.
+Anticlustering is usually implemented by reversing an objective function used in classical cluster analysis.
 Whereas cluster analysis seeks homogeneity within groups and separation between groups, the reversal results in heterogeneity within groups and similarity between groups.
@@
 Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern.
-Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, while good clustering solutions tend to look similarly.
-A good clustering solution depends on assigning similar elements into the same cluster and dissimilar elements into different clusters.
-A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.
-This difference between cluster analysis and anticlustering can also be clarified using Figure 1.
-Here, we observe a high concentration of suitable anticlustering partitions: An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution.
+Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, while good clustering solutions tend to look similar.
+A good clustering solution assigns similar elements into the same cluster and dissimilar elements into different clusters.
+A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This distinction can be clarified using Figure 1 where we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.
+An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution.
 Among these `r sum(df$kmeans / max(df$kmeans) >.999)` partitions, every possible pair of elements is sometimes assigned to the same group and sometimes not.
 In contrast, there are only `r sum(min(df$kmeans)/ df$kmeans > .90)` partitions (`r mean(min(df$kmeans)/ df$kmeans > .90)*100`%) that approximate the global minimum in $\mathit{SSE}$ (i.e., the best k-means clustering solution) by even 10%.
-Thus, even though the problems are based on the same objective function, cluster analysis and anticlustering exhibit a stark discrepancy with regard to the solution space of partitions: There are many suitable anticlusterings but only few suitable clusterings.
-Interestingly, the k-plus objective further loosens the link between cluster analysis and anticlustering.
-Reversing the k-plus anticlustering objective no longer yields a useful clustering interpretation, because identifying groups that are different with regard to higher order moments does not constitute a clustering in the usual sense.
+Thus, even though the problems are based on the same objective function, cluster analysis and anticlustering exhibit a stark discrepancy with regard to the solution space of partitions.
+Interestingly, the k-plus objective introduces an additional distinction between cluster analysis and anticlustering.
+Reversing the k-plus anticlustering objective no longer yields a useful clustering interpretation, because identifying groups that are different with regard to higher order moments does not constitute a clustering.
 Thus, even though k-plus extends the classical k-means clustering criterion, it is a pure anticlustering objective.
+
+[^dispersion]: An exception is given when using the dispersion objective, which ensures that the most similar elements are assigned to different clusters.
+However, anticlustering objectives striving for overall between-group similarity are mostly unaffected by the grouping of individual item pairs.
 
@@
 Thus, as compared to pure cluster analysis, fair clustering biases the identification of groups with regard to the protected attributes, striving for algorithmic fairness.
-A relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data [@delBarrio2022], while anticlustering is not concerned with identifying structure in the data; instead, anticlustering divides data sets into artificial subgroups in such a way that between-group similarity / within-group heterogeneity is maximized.
+A relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data [@delBarrio2022], while anticlustering divides data sets into artificial subgroups in such a way that between-group similarity / within-group heterogeneity is maximized.
 However, these methods are logically similar because anticlustering corresponds to fair clustering when only considering the protected attributes.

### 2023-05-02 paper.Rmd
editing

@@
 Let $T$ be the number of distribution moments that should be similar across $K$ groups.
-For each moment $t$ ($t = 1, \ldots, T$) and for each variable $x_{\cdot j}$ ($j = 1, \ldots, P$), a new *t-moment variable* $\hat{x}_{t \cdot j} = (\hat{x}_{t1j}, \ldots, \hat{x}_{tNj})$ is defined where
+For each moment $t$ ($t = 1, \ldots, T$) and for each variable $x_{\cdot j}$ ($j = 1, \ldots, P$), a new *k-plus variable* $\hat{x}_{t \cdot j} = (\hat{x}_{t1j}, \ldots, \hat{x}_{tNj})$ is defined where
 
@@
 
-Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{\mathbf{X}}^{(k)}_{T \times P} = \{\overline{x}_{tj}^{(k)}\}_{T \times P}$ is a centroid matrix containing the mean values for all $t$-moment variables in the $k$'th group, i.e.
+Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{\mathbf{X}}^{(k)}_{T \times P} = \{\overline{x}_{tj}^{(k)}\}_{T \times P}$ is a centroid matrix containing the mean values for all k-plus variables in the $k$'th group, i.e.
 
@@
 Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ has the same form as the standard k-means criterion $\mathit{SSE}$, but it is computed on different data.
-Computing the k-plus criterion is then done by first creating the additional $t$-moment variables, appending them as new columns to the original data matrix, and computing the $\mathit{SSE}$ on the augmented data.
+Computing the k-plus criterion is then done by first creating the additional k-plus variables, appending them as new columns to the original data matrix, and computing the $\mathit{SSE}$ on the augmented data.
 Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
@@
 Note that a problem may arise if the weights $\mathbf{w}$ are ignored when computing $\mathit{SSE_{kplusGeneralized}}$.
-Because the computation of the new $t$-moment variables involves exponentiation, their range of values is expected to be very different from the original variables.
+Because the computation of the new k-plus variables involves exponentiation, their range of values is expected to be very different from the original variables.
 This imbalance can strongly and unpredictably influence the relative weight of the different moments when computing $\mathit{SSE_{kplusGeneralized}}$ (e.g., see Figure 1).
-To ensure a comparable weight of all criteria, all attributes---i.e., the original data as well as the new $t$-moment variables---can be standardized before computing $\mathit{SSE_{kplusGeneralized}}$, e.g.
+To ensure a comparable weight of all criteria, all attributes---i.e., the original data as well as the new k-plus variables---can be standardized before computing $\mathit{SSE_{kplusGeneralized}}$, e.g.
 via $z$-standardization.
@@
 
-To thoroughly evaluate the k-plus approach to anticlustering, two specific k-plus criteria were compared to two traditional anticlustering criteria: (a) "standard" k-plus anticlustering, minimizing differences with regard to group means and variances (by maximizing $\mathit{SSE_{kplus}}$); (b) "generalized" k-plus anticlustering, minimizing differences with regard to group means, variances, skewness and kurtosis (by maximizing $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$).
+In Simulation 2, the performance of two k-plus criteria was compared to the performance of two traditional anticlustering criteria.
+"Standard" k-plus anticlustering was implemented by by maximizing $\mathit{SSE_{kplus}}$, i.e., minimizing differences with regard to group means and variances; "generalized" k-plus anticlustering was implemented by by maximizing $\mathit{SSE_{kplusGeneralized}}$ using $T = 4$, i.e., minimizing differences with regard to group means, variances, skewness and kurtosis.
 To implement the k-plus objectives, the unweighted standardization approach was employed that performed best in Simulation 1.
-Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.
+Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented k-plus variables---were $z$-standardized.
 Subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
@@
 
-[^dispersion]: An exception is given when using the dispersion objective, which ensures that the most similar elements are assigned to different clusters.
-However, anticlustering objectives striving for overall between-group similarity are mostly unaffected by the grouping of individual item pairs.
+[^dispersion]: An exception is given when employing the dispersion objective, which ensures that the most similar elements are assigned to different clusters.
+However, anticlustering objectives striving for overall between-group similarity are largely unaffected by the grouping of individual item pairs.
 
@@
 Fair clustering performs a cluster analysis under the restriction that the distribution of a set of "protected" attributes---such as race, sex or age---should be similar between clusters.
-Thus, as compared to pure cluster analysis, fair clustering biases the identification of groups with regard to the protected attributes, striving for algorithmic fairness.
+Thus, as compared to pure cluster analysis, fair clustering biases the group identification with regard to the protected attributes, striving for algorithmic fairness.
 A relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data [@delBarrio2022], while anticlustering divides data sets into artificial subgroups in such a way that between-group similarity / within-group heterogeneity is maximized.
-However, these methods are logically similar because anticlustering corresponds to fair clustering when only considering the protected attributes.
-Hence, a fruitful outlook for research might be to investigate if approaches stemming from the literature on fair clustering can be adopted for anticlustering, and vice versa.
+However, these methods are logically similar because fair  clustering is akin to anticlustering when only considering the protected attributes.
+Hence, a fruitful outlook for future research might be to investigate if approaches stemming from the literature on fair clustering can be adopted for anticlustering, and vice versa.
 

### 2023-05-03 paper.Rmd
make contribution of simulation more clear

@@
 
-To thoroughly evaluate the k-plus approach to anticlustering, two specific k-plus criteria were compared to two traditional anticlustering criteria: (a) "standard" k-plus anticlustering, minimizing differences with regard to group means and variances (by maximizing $\mathit{SSE_{kplus}}$); (b) "generalized" k-plus anticlustering, minimizing differences with regard to group means, variances, skewness and kurtosis (by maximizing $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$).
+In Simulation 2, the performance of two k-plus criteria was compared to the performance of two traditional anticlustering criteria.
+"Standard" k-plus anticlustering was implemented by maximizing $\mathit{SSE_{kplus}}$, thus minimizing differences with regard to group means and variances.
+"Generalized" k-plus anticlustering was implemented by maximizing $\mathit{SSE_{kplusGeneralized}}$ using $T = 4$, thus minimizing differences with regard to group means, variances, skewness and kurtosis.
 To implement the k-plus objectives, the unweighted standardization approach was employed that performed best in Simulation 1.
@@
 Subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
-The k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means, (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups, (iii) a random allocation of elements to groups.
+The k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means, (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups, and also (iii) a random allocation of elements to groups.
 

### 2023-05-03 paper.Rmd
after the revision its four anticlustering methods

@@
 
-After the five anticlustering methods and the random allocation procedure had been applied to the 10,000 data sets for $K = 2$, $K = 3$, $K = 4$, I investigated how well the competing methods were able to create similar groups.
+After the four anticlustering methods and the random allocation procedure had been applied to the 10,000 data sets for $K = 2$, $K = 3$, $K = 4$, I investigated how well the competing methods were able to create similar groups.
 To this end, I analyzed the discrepancy in means, standard deviations, skewness and kurtosis between groups---for each data set for each $K$.

### 2023-05-03 paper.Rmd
minor

@@
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis.
-After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming the traditional anticlustering criteria diversity and k-means.
+After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional anticlustering criteria such as the diversity and the classical k-means objective.
 Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].
@@
 A good clustering solution assigns similar elements into the same cluster and dissimilar elements into different clusters.
-A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This distinction can be clarified using Figure 1 where we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.
+A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This pattern was observed in SImulation 2, where it was shown that partitions obtaining high between-group similarity according to different criteria were not at all similar; the ARI was usually close to 0.
+For different clustering methods, there is usually a considerable amount of overlap between partitions.
+This  distinction between cluster analysis and anticlustering can also be clarified using Figure 1, where we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.
 An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution.

### 2023-05-03 paper.Rmd
include ARI in the results

@@
 The R package `faux` [Version 1.1.0, @R-faux] was used to generate the multivariate normal data sets that were processed during Simulation 2.
-The R packages `dplyr` [Version 1.0.7, @R-dplyr], `ggplot2` [Version 3.3.5, @R-ggplot2], `papaja` [Version 0.1.1.9001, @R-papaja], `tidyr` [Version 1.1.3, @R-tidyr] and `DescTools` [Version 0.99.45, @R-desctools] were used for data analysis and result presentation.
+The R packages `dplyr` [Version 1.0.7, @R-dplyr], `ggplot2` [Version 3.3.5, @R-ggplot2], `papaja` [Version 0.1.1.9001, @R-papaja], `tidyr` [Version 1.1.3, @R-tidyr], `fossil` [Version 0.4.0, @R-fossil], and `DescTools` [Version 0.99.45, @R-desctools] were used for data analysis and result presentation.
 The simulation study is fully reproducible via code and data that has been made accessible in an accompanying OSF repository [@osf2023].
@@
 
+As a secondary purpose of Simulation 2, I investigated the degree to which partitions identified by the different methods showed similarity.
+Specifically, I investigated whether different anticlustering solutions showed any overlap regarding the placement of individual elements in groups.
+In cluster analysis, high quality partitions according to different objectives tend to overlap to a large degree, but it is an open question whether this result also holds for different anticlustering partitions.
+To tackle this issue, I computed the adjusted Rand index (ARI) for each pair of partitions that were returned by the competing methods [@rand1971; @hubert1985].
+The ARI is an index of the similarity between two partitions, where an index of 0 indicates no similarity and a value of 1 indicates the highest possible similarity (i.e., when two partitions are the same).
+The pairwise ARIs were computed for each data set and for each $K$, i.e., 30,000 times.
+
 ### Results
@@
   tab,
-  caption = "Results of Simulation study 2",
+  caption = "Results of Simulation 2.",
   note = "The global results of the simulation study aggregated across all conditions.
@@
 
+```{r}
+
+ARIs <- as.matrix(read.table("Simulation_2/global_results_ari.csv", sep = ","))
+colnames(ARIs) <- c(
+  "diversity", "k-means", "k-plus", "k-plus generalized", "random"
+)
+rownames(ARIs) <- colnames(ARIs)
+ARIs <- round(ARIs, 3)
+ARIs[upper.tri(ARIs)] <- ""
+
+# removing leading 0
+ARIs <- gsub("0.", ".", ARIs, fixed = TRUE)
+
+apa_table(
+  ARIs,
+  caption = "Average pairwise ARI returned by the competing methods in Simulation 2.",
+  note = "Each cell value is the result of averaging across 30,000 data points (10,000 data sets $\\times$ $K = 2$, $K = 3$ or $K = 4$).",
+  escape = FALSE,
+  align = c("l")
+)
+
+```
+
+Table 3 illustrates the degree of overlap between anticlustering partitions that were identified in Simulation 2.
+Overall, it is shown that there was hardly any similarity between anticlustering partitions according to the different criteria.
+The average ARI was almost 0 in all cases---indicating no similarity of partitions whatsoever.
+This constitutes an interesting and novel result, illustrating that high between-group similarity can be achieved via very different partitionings.
+
 # Applying k-plus anticlustering using the R package anticlust
@@
 
-Table 3 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features.
-Table 3 also contains the results after applying k-means anticlustering and diversity anticlustering.
+Table 4 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features.
+Table 4 also contains the results after applying k-means anticlustering and diversity anticlustering.
 It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations, and the skewness was also very evenly matched.
@@
 
-Table 4 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups.
+Table 5 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups.
 K-plus anticlustering strives for between-group similarity regardless of group size.
@@
 
-Table 5 illustrates the results that are obtained when using standard[^standardkplus] k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
+Table 6 illustrates the results that are obtained when using standard[^standardkplus] k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
 Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are similar enough for any practical purposes.
@@
 Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern.
-Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, while good clustering solutions tend to look similar.
-A good clustering solution assigns similar elements into the same cluster and dissimilar elements into different clusters.
-A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This pattern was observed in SImulation 2, where it was shown that partitions obtaining high between-group similarity according to different criteria were not at all similar; the ARI was usually close to 0.
-For different clustering methods, there is usually a considerable amount of overlap between partitions.
-This  distinction between cluster analysis and anticlustering can also be clarified using Figure 1, where we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.
+Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, as shown in Simulation 2.
+In contrast, good clustering solutions tend to look similar because a good clustering is characterized by assigning similar elements into the same cluster and dissimilar elements into different clusters.
+A good anticlustering solution however does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This distinction can also be clarified using Figure 1, where we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.
 An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution.
-Among these `r sum(df$kmeans / max(df$kmeans) >.999)` partitions, every possible pair of elements is sometimes assigned to the same group and sometimes not.
 In contrast, there are only `r sum(min(df$kmeans)/ df$kmeans > .90)` partitions (`r mean(min(df$kmeans)/ df$kmeans > .90)*100`%) that approximate the global minimum in $\mathit{SSE}$ (i.e., the best k-means clustering solution) by even 10%.

### 2023-05-03 Simulation_2/ARI.R

@@ -1,0 +1,46 @@
+
+library(fossil)
+source("0-functions-analyze-data.R")
+
+# Compute all pairwise ARIs for a set of partitions
+# argument l is a list of partitions
+all_pairwise_aris <- function(l, ID) {
+  method_names <- gsub(paste0("-", ID), "", names(l))
+  ARIs <- matrix(NA, nrow = length(l), ncol = length(l))
+  colnames(ARIs) <- method_names
+  rownames(ARIs) <- method_names
+  for (i in seq_along(l)) {
+    for (j in seq_along(l)) {
+      ARIs[i, j] <- adj.rand.index(l[[i]], l[[j]])
+    }
+  }
+  ARIs
+}
+
+Ks <- 2:4
+results_by_k <- vector(mode = "list", length = 3)
+names(results_by_k) <- Ks
+for (K in Ks) {
+  data <- read.csv2(paste0("results-K", K, "-solutions.csv"), stringsAsFactors = FALSE)
+  
+  # ensure that the order of methods is good
+  n_methods <- length(unique(data$method))
+  apply(matrix(data$method, ncol = n_methods, byrow = TRUE), 2, function(x) all(x == x[1]))
+  all_solutions <- lapply(data$result, anticlusters_from_string)
+  names(all_solutions) <- paste0(rep_len(data$method[1:n_methods], length(all_solutions)), "-", data$ID)
+  
+  # exclude correlation method
+  all_solutions <- all_solutions[!grepl("correlation", names(all_solutions))]
+  
+  
+  all_aris <- vector(mode = "list", length = length(unique(data$ID)))
+  names(all_aris) <- unique(data$ID)
+  for (id in unique(data$ID)) {
+    current_solutions <- all_solutions[grepl(id, names(all_solutions))]
+    all_aris[[id]] <- all_pairwise_aris(current_solutions, id)
+  }
+  
+  results_by_k[[as.character(K)]] <- Reduce("+", all_aris) / length(all_aris)
+}
+
+write.table(Reduce("+", results_by_k) / length(results_by_k), "global_results_ari.csv", sep = ",")

### 2023-05-03 paper.Rmd
editing

@@
 
-As a secondary purpose of Simulation 2, I investigated the degree to which partitions identified by the different methods showed similarity.
+As a secondary purpose of Simulation 2, I investigated the degree to which partitions identified by the different anticlustering methods showed similarity.
 Specifically, I investigated whether different anticlustering solutions showed any overlap regarding the placement of individual elements in groups.
@@
 The average ARI was almost 0 in all cases---indicating no similarity of partitions whatsoever.
+Moreover, the ARI between two anticlustering partitions hardly differed from the ARI between an anticlustering partition and a random partition.
 This constitutes an interesting and novel result, illustrating that high between-group similarity can be achieved via very different partitionings.
@@
 Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern.
-Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, as shown in Simulation 2.
-In contrast, good clustering solutions tend to look similar because a good clustering is characterized by assigning similar elements into the same cluster and dissimilar elements into different clusters.
+Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, as was also shown in Simulation 2.
+In contrast, different clustering solutions tend to look similar because a good clustering is characterized by assigning similar elements into the same cluster and dissimilar elements into different clusters.
 A good anticlustering solution however does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This distinction can also be clarified using Figure 1, where we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.

### 2023-05-03 Simulation_2/ARI.R
this was not needed

@@ -24,14 +24,11 @@ for (K in Ks) {
   data <- read.csv2(paste0("results-K", K, "-solutions.csv"), stringsAsFactors = FALSE)
   
   # ensure that the order of methods is good
-  n_methods <- length(unique(data$method))
-  apply(matrix(data$method, ncol = n_methods, byrow = TRUE), 2, function(x) all(x == x[1]))
   all_solutions <- lapply(data$result, anticlusters_from_string)
-  names(all_solutions) <- paste0(rep_len(data$method[1:n_methods], length(all_solutions)), "-", data$ID)
+  names(all_solutions) <- paste0(data$method, "-", data$ID)
   
   # exclude correlation method
   all_solutions <- all_solutions[!grepl("correlation", names(all_solutions))]
-  
   
   all_aris <- vector(mode = "list", length = length(unique(data$ID)))
   names(all_aris) <- unique(data$ID)

### 2023-05-04 paper_submission.tex
added submitted paper tex for latexdiff

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, stimulus selection\newline\indent Word count: 4992}
\usepackage{csquotes}
\usepackage{setspace}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{figure}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\makeatletter
\renewcommand*\env@matrix[1][*\c@MaxMatrixCols c]{%
  \hskip -\arraycolsep
  \let\@ifnextchar\new@ifnextchar
  \array{#1}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={k-plus Anticlustering: An Improved k-means Criterion for Maximizing Between-Group Similarity},
  pdfauthor={Martin Papenberg1},
  pdflang={en-EN},
  pdfkeywords={Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, stimulus selection},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{k-plus Anticlustering: An Improved k-means Criterion for Maximizing Between-Group Similarity}
\author{Martin Papenberg\textsuperscript{1}}
\date{}

\shorttitle{k-plus anticlustering}

\authornote{

Martin Papenberg, Department of Experimental Psychology, Heinrich Heine University Düsseldorf.

Correspondence concerning this article should be addressed to Martin Papenberg, Heinrich-Heine-Universität Düsseldorf, Institut für Experimentelle Psychologie, Universitätsstraße 1, 40225 Düsseldorf, Germany. E-mail: \href{mailto:martin.papenberg@hhu.de}{\nolinkurl{martin.papenberg@hhu.de}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Heinrich Heine University Düsseldorf}

\abstract{%
Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity. Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximizing instead of minimizing a clustering objective function. Introducing k-plus, this paper presents an extension of the classical k-means objective to maximize between-group similarity in anticlustering applications. The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, higher order moments, and covariances), whereas the k-means criterion only reflects the degree to which groups differ in their means. While constituting a new criterion for anticlustering, it is surprisingly shown that k-plus anticlustering can be implemented by optimizing the original k-means criterion after the input data has been augmented with additional variables. A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple objectives. In particular, optimizing between-group similarity with regard to variances usually does not compromise similarity with regard to means; the k-plus extension is therefore generally preferred over classical k-means anticlustering. Examples are given on how k-plus anticlustering can be applied to real norming data using the open source R package \texttt{anticlust}, which is freely available via CRAN (\url{https://cran.r-project.org/package=anticlust}).
}

\begin{document}
\maketitle

Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity (Brusco et al., 2020; Späth, 1986; Valev, 1983, 1998). Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and separation between clusters (Rokach \& Maimon, 2005). Anticlustering has many applications in research psychology (Brusco et al., 2020; Papenberg \& Klau, 2021; Steinley, 2006). Examples include splitting tests into parts of equal difficulty (Gierl et al., 2017), assigning students to work groups (Baker \& Powell, 2002), and assigning stimuli to different, but parallel experimental conditions (Lahl \& Pietrowsky, 2006). Solving anticlustering problems ``by hand'' is a tedious and time-consuming task, and the quality of manual partitioning is usually subpar. Fortunately, anticlustering problems can be formalized as mathematical optimization problems (e.g., Baker \& Powell, 2002; Brusco et al., 2020; Fernández et al., 2013; Späth, 1986) and accessible open source software solutions to tackling these problems exist (Papenberg \& Klau, 2021).

Anticlustering methods are characterized by (a) an objective function that quantifies between-group similarity and/or within-group heterogeneity, and (b) an algorithm that determines how elements are assigned to groups to maximize the objective function. Several anticlustering objective functions use pairwise dissimilarity ratings such as the Euclidean distance as input. The most prominent criterion of this kind is the \emph{diversity}, which is the total sum of pairwise dissimilarities between elements within the same group (Brusco et al., 2020; Gallego et al., 2013). By considering pairwise dissimilarities between elements in the same group, the diversity criterion reflects the total degree of within-group heterogeneity. High within-group diversity simultaneously ensures high between-group similarity.\footnote{As will be shown later, the diversity objective only strictly equalizes within-group heterogeneity and between-group similarity when groups are equal-sized.} Another important anticlustering criterion based on the information in a dissimilarity matrix is the \emph{dispersion}, which is the minimum dissimilarity between any two elements within the same group (Fernández et al., 2013). Maximizing the dispersion increases within-group heterogeneity by ensuring that any two elements in the same group are as dissimilar from each other as possible. Brusco et al. (2020) convincingly argued that anticlustering applications striving for within-group heterogeneity should incorporate both dispersion and diversity, and they presented a bicriterion algorithm to approximate the Pareto efficient set of solutions according to both criteria.

Oftentimes, partitioning applications in psychology are focused on between-group similarity rather than within-group heterogeneity, even though both goals oftentimes coincide. High between-group similarity is for example desirable when designing experimental conditions using different stimulus sets that should be as similar as possible with regard to response-relevant attributes (Lahl \& Pietrowsky, 2006). Such stimulus selection tasks are usually conducted on the basis of attribute data and not on pairwise (dis)similarity ratings. That is, individual stimuli are numerically coded on relevant dimensions (Dry \& Storms, 2009). For stimuli in psycholinguistic experiments, attributes may consist of ratings for imagery and concreteness, as well as orthographic variables (Friendly et al., 1982). In other cases, attributes are binary and may represent the presence or absence of a feature (Tversky, 1977). When attribute values are available, anticlustering approaches that do not require a dissimilarity matrix\footnote{It is still possible to convert the attribute data into a dissimilarity matrix, e.g.~by computing all pairwise Euclidean distances, and then optimize a distance-based anticlustering criterion to obtain a partitioning.} can be used to partition the data. K-means is probably the best-known clustering objective that can directly be computed on attribute values. In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized, usually using the k-means heuristic (Brusco, 2006; Jain, 2010; Steinley, 2006). Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity (Aloise et al., 2009). Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful approach when aiming for high between-group similarity (Späth, 1986; Valev, 1983, 1998). Specifically, as Späth (1986) noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters. However, other distribution characteristics---such as the variance---are not targeted. Papenberg and Klau (2021) showed that k-means anticlustering tends to over-optimize between-group similarity with regard to the mean at the cost of similarity in variance. However, when aiming for overall between-group similarity, neglecting all distribution characteristics other than the mean is misguided; similar means can be obtained even when the underlying distributions are clearly different (e.g., Anscombe, 1973).

To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering. K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to several distribution characteristics instead of only the mean. Specifically, k-plus offers the possibility to minimize differences with regard to the variance, covariances, and higher order moments such as skewness and kurtosis. After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics. Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package \texttt{anticlust} (Papenberg, 2019; Papenberg \& Klau, 2021), an extension to the popular statistical programming language R (R Core Team, 2022).

\hypertarget{problem-formalization}{%
\section{Problem formalization}\label{problem-formalization}}

For the purpose of problem formalization, I adopt the notation Steinley (2006) provided in his comprehensive synthesis of half a century of research on k-means. K-means anticlustering is used to partition \(N\) elements each having \(P\) attributes into \(K\) groups \((C_1, \ldots, C_K)\). The partitioning is subject to the requirement that each element is assigned to exactly one group. In anticlustering applications, an additional constraint is usually imposed on the group sizes, with the most common restriction being that all groups have equal size (Papenberg \& Klau, 2021).

For a data matrix \(\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}\), k-means anticlustering aims to maximize the error sum of squares (\(\mathit{SSE}\)), which is the sum of the squared Euclidean distances between each data point and its cluster centroid:

\[
\mathit{SSE} = 
  \sum\limits_{j=1}^{P} 
  \sum\limits_{k=1}^{K} 
  \sum\limits_{i \in C_k}
  (x_{ij} - \overline{x}_j ^{(k)})^2
\]

The \emph{k}'th centroid \(\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})\) is a vector of length \(P\), where each entry is the mean value of one of the \(P\) attributes, computed across all observations belonging to the \(k\)'th cluster \(C_k\) (\(k = 1, \ldots, K\)):

\[
\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} x_{ij}
\]

Note that there is a direct connection between the \(\mathit{SSE}\) and the location of the cluster centroids (Späth, 1986). If the \(\mathit{SSE}\) is minimal---which is the goal of k-means \emph{clustering}---, the centroid values \(\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)\) are as far away as possible from the overall data centroid \(\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P\)), where \(\overline{\mathbf{x}}_p\) is the overall mean value on the \(p\)th attribute:

\[
\overline{\mathbf{x}}_p = \frac{1}{N} \sum\limits_{i=1}^{N} x_{ip}
\]

If the \(\mathit{SSE}\) is maximal---which is the goal of k-means \emph{anticlustering}---, the cluster centroids are as close as possible to the overall centroid \(\overline{\mathbf{x}}\) and therefore to each other (Späth, 1986). Thus, k-means anticlustering directly optimizes the similarity of the mean attribute values between clusters.

\hypertarget{quantifying-differences-in-variance}{%
\subsection{Quantifying differences in variance}\label{quantifying-differences-in-variance}}

In the following, I present a variation of the \(\mathit{SSE}\) that quantifies how the variance of numeric data differs between groups. To motivate the new criterion, we first consider a set of one-dimensional observations. The sample variance of a set of \(N\) observations \(x = (x_1, \ldots, x_N)\) is computed as the mean of the squared distances between observations and their mean \(\overline{\mathbf{x}}\):

\[
\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^2
\]

When defining a new variable \(y = (y_1, \ldots, y_N)\) as the squared deviation of each observation from the mean \(\overline{\mathbf{x}}\), the sample variance of \(x\) can be reformulated as the mean of \(y\):

\[
y_i = (x_i - \overline{\mathbf{x}})^2
\]

\[
\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} y_i
\]

Because the sample variance is basically defined as an average---and k-means anticlustering minimizes differences with regard to averages---k-means anticlustering can be employed to equalize the variance across groups. This constitutes the central insight motivating the new k-plus criterion. Thus, to obtain an adaptation of the \(\mathit{SSE}\) that quantifies the degree to which the variance differs between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable. We obtain a new data matrix \(\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}\) with \(y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2\). On the basis of \(\mathbf{Y}_{N \times P}\), we define the criterion \(\mathit{SSE_{Var}}\), reflecting how strongly the variances of all attributes vary across the \(K\) clusters:

\[
\mathit{SSE_{Var}} = 
  \sum\limits_{j=1}^{P} 
  \sum\limits_{k=1}^{K} 
  \sum\limits_{i \in C_k}
  (y_{ij} - \overline{y}_j ^{(k)})^2
\]

Note that \(\mathit{SSE_{Var}}\) has the same form as the standard \(\mathit{SSE}\). However, the input data reflects the squared difference between each data point and the attribute mean instead of the raw data points. Maximizing \(\mathit{SSE_{Var}}\) will lead to similar variances across the \(K\) groups in the \(P\) attributes.

\hypertarget{kplus}{%
\subsection{A bicriterion model: K-plus anticlustering}\label{kplus}}

It is unlikely that an exclusive maximization of \(\mathit{SSE_{Var}}\) (i.e., equalizing variances) is of interest without also considering the standard \(\mathit{SSE}\) (i.e., equalizing means). A combination of \(\mathit{SSE_{Var}}\) and \(\mathit{SSE}\), however, can be employed to simultaneously optimize both criteria, which is a reasonable idea when striving for overall between-group similarity. The most simple approach for such a bicriterion optimization is the weighted sum approach (Marler \& Arora, 2010; Naidu et al., 2014). That is, both criteria are computed independently from each other and are then combined into a single criterion through a weighted sum:

\[
w_1 \sum\limits_{j=1}^{P} 
  \sum\limits_{k=1}^{K} 
  \sum\limits_{i \in C_k}
  (x_{ij} - \overline{x}_j ^{(k)})^2 + 
w_2 \sum\limits_{j=1}^{P} 
  \sum\limits_{k=1}^{K} 
  \sum\limits_{i \in C_k}
  (y_{ij} - \overline{y}_j ^{(k)})^2
\]

The weights \(w_1\) and \(w_2\) can be chosen to adjust the relative importance of equalizing means and variances, respectively. Some simplifications can be arranged when ignoring the weights \(w_1\) and \(w_2\), e.g., by setting them both to 1. This simplification leads to the \emph{unweighted sum approach} for k-plus anticlustering: When defining a new data matrix \(\mathbf{Z}_{N \times 2P} = \{z_{ij}\}_{N \times 2P}\) as the augmentation of \(\mathbf{X}_{N \times P}\) with \(\mathbf{Y}_{N \times P}\), i.e.

\[
  \mathbf{Z}_{N \times 2P} =
  \left[ {\begin{array}{cccc|cccc}
    x_{11} & x_{12} & \cdots & x_{1P} & y_{11} & y_{12} & \cdots & y_{1P}\\
    x_{21} & x_{22} & \cdots & x_{2P} & y_{21} & y_{22} & \cdots & y_{2P}\\
    \vdots & \vdots & \ddots & \vdots & \vdots & \vdots & \ddots & \vdots\\
    x_{N1} & x_{N2} & \cdots & x_{NP} & y_{N1} & y_{N2} & \cdots & y_{NP}\\
  \end{array} } \right]
\]

and \(z_{i.} = (x_{i1}, \ldots, x_{iP}, y_{i1}, \ldots, y_{iP})\), the bicriterion \(\mathit{SSE_{kplus}}\) can be computed as

\[
\mathit{SSE_{kplus}} =
  \sum\limits_{j=1}^{2 P} 
  \sum\limits_{k=1}^{K} 
  \sum\limits_{i \in C_k}
  (z_{ij} - \overline{z}_j ^{(k)})^2.
\]

This term again has the same form as \(\mathit{SSE}\) but is computed on different data. Hence, computing the k-plus criterion can be done by first computing the squared distances \(y_{ij}\), appending them as new columns to the original data matrix, and then computing \(\mathit{SSE}\) on the augmented data. Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.

Note that a problem may arise if the weights \(w_1\) and \(w_2\) are ignored when computing \(\mathit{SSE_{kplus}}\). Because the computation of the new attribute values \(y_{ij}\) involves squaring, their range of values is expected to be very different from the raw values \(x_{ij}\) (e.g., see Figure 1). This imbalance can strongly and unpredictably influence the relative weight of \(\mathit{SSE}\) and \(\mathit{SSE_{Var}}\) when computing \(\mathit{SSE_{kplus}}\). To ensure an equal weight of both criteria, all attributes (\(x_{ij}\) and \(y_{ij}\)) can be standardized before computing \(\mathit{SSE_{kplus}}\), e.g.~via \(z\)-standardization.

Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria at the same time (Ferligoj \& Batagelj, 1992). It is therefore unclear whether optimizing a bicriterion objective such as \(\mathit{SSE_{kplus}}\) can yield a partitioning where both constituting criteria (mean and variance) are similar between groups. A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible. I created 14 data points following a univariate normal distribution (\(M = 15\), \(SD = 3\)) and then generated all 1716 possible ways to partition the 14 data points into \(K = 2\) equal-sized groups. Figure 1 shows the scatterplot of the criteria \(\mathit{SSE}\) and \(\mathit{SSE_{Var}}\) across partitions. The partition that is represented by the squared symbol has the overall best highest \(\mathit{SSE}\), but is outperformed by many other partitions with regard to \(\mathit{SSE_{Var}}\). Similarly, the partition that maximizes \(\mathit{SSE_{Var}}\) (represented by the triangular symbol) is not well-suited to optimize \(\mathit{SSE}\) at the same time. Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances. The diamond symbol represents the partition that maximizes \(\mathit{SSE_{kplus}}\), which yields 99.99\% of the global maximum \(\mathit{SSE}\), and 99.93\% of the global maximum value of \(\mathit{SSE_{Var}}\). Thus, using the unweighted sum approach, the optimal partition according to \(\mathit{SSE_{kplus}}\) fulfills the two constituting criteria \(\mathit{SSE}\) and \(\mathit{SSE_{Var}}\) to an astonishing degree.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}The relationship between the objectives \(\mathit{SSE}\) and \(\mathit{SSE_{Var}}\) across all ways to partition \(N = 14\) normally distributed random values into \(K = 2\) groups. The partition represented by the square maximizes \(\mathit{SSE}\) and thus minimizes the difference in means between groups; the partition represented by the triangular maximizes \(\mathit{SSE_{Var}}\) and thus minimizes the difference in variances between groups. The partition represented by the diamond maximizes \(\mathit{SSE_{kplus}}\) and satisfies both constituting criteria to an almost optimal degree.}
\end{figure}

\hypertarget{skewness-kurtosis-and-higher-order-moments}{%
\subsection{Skewness, kurtosis, and higher order moments}\label{skewness-kurtosis-and-higher-order-moments}}

The k-plus approach to equalizing the variance can be generalized to higher order moments. The \emph{j}'th sample moment of a variable \(x = (x_1, \ldots, x_N)\) can be computed as

\[
\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^j.
\]

where \(j = 2\) gives the variance, \(j = 3\) the skewness, and \(j = 4\) the kurtosis (Joanes \& Gill, 1998). Thus, by changing \(j\)---the power of the deviation between mean and data points---we can use the k-plus logic to equalize any desired sample moment between groups. For each moment \(j\) and for each variable \(x\), this is accomplished by adding a new variable \(y^{(j)} = (y^{(j)}_1, \ldots, y^{(j)}_N)\) with \(y^{(j)}_i = (x_i - \overline{\mathbf{x}})^j\) to the data set and subsequently applying standard k-means anticlustering. I expect this to be particularly interesting for the skewness and kurtosis when aiming to minimize differences with regard to distribution asymmetry (skewness) and the propensity to include outliers (kurtosis).

\hypertarget{covariances}{%
\subsection{Covariances}\label{covariances}}

Because the covariance is also defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariance structure. The covariance between two variables \(x = (x_1, \ldots, x_N)\) and \(y = (y_1, \ldots, y_N)\) is defined as

\[
\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}}).
\]

Thus, to maximize between-group similarity with regard to covariance structure, it is possible to augment the data set with an additional variable for each pair of variables \(x\) and \(y\), computed as \((x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})\).

\hypertarget{simulation-study}{%
\section{Simulation study}\label{simulation-study}}

To thoroughly evaluate the k-plus approach to anticlustering, three specific k-plus criteria were implemented in a simulation study: (i) ``standard'' k-plus anticlustering, minimizing differences with regard to group means and variances; (ii) k-plus \emph{skew/kurtosis}, minimizing differences with regard to group means, variances, skewness and kurtosis; (iii) k-plus \emph{correlation}, minimizing differences with regard to group means, variances and covariance structure. To implement all k-plus objectives, the unweighted sum approach was employed. Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were \(z\)-standardized. Subsequently, the original k-means criterion \(\mathit{SSE}\) was maximized on the basis of the augmented data set. The three k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means, (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups, (iii) a random allocation of elements to groups.

The simulation was implemented using the statistical programming language R (Version 4.2, R Core Team, 2022). The R package \texttt{anticlust} (Version 0.6.1, Papenberg, 2019; Papenberg \& Klau, 2021) was used to optimize the anticlustering objectives. The R package \texttt{faux} (Version 1.1.0, DeBruine, 2021) was used to generate the multivariate normal data sets that were processed during the simulation. The R packages \texttt{dplyr} (Version 1.0.7, Wickham et al., 2019), \texttt{ggplot2} (Version 3.3.5, Wickham, 2016), \texttt{papaja} (Version 0.1.1.9001, Aust \& Barth, 2018), \texttt{tidyr} (Version 1.1.3, Wickham, 2021) and \texttt{DescTools} (Version 0.99.45, Signorell, 2022) were used for data analysis and result presentation. The simulation study is fully reproducible via code and data that has been made accessible in an accompanying OSF repository (Papenberg, 2023).

\hypertarget{optimization-algorithm}{%
\subsection{Optimization algorithm}\label{optimization-algorithm}}

The same local maximum search was applied to optimize each of the five competing anticlustering objectives (Method LCW, Weitz \& Lakshminarayanan, 1998). Building on an initial random allocation, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin. That is, it starts with the first element and simulates all exchanges with elements that are currently assigned to a different group. After each exchange has been evaluated, the one exchange is realized that improves the objective function the most. No exchange is realized if no improvement is possible. The exchange process is repeated for each element. After that, the procedure restarts at the beginning and is repeated until an iteration through the entire data set no longer yields an improvement. To obtain better results, this local maximum search may be initialized multiple times (Späth, 1986). In the simulation study five repetitions were used.

\hypertarget{conditions}{%
\subsection{Conditions}\label{conditions}}

For the simulation, I generated 10,000 data sets following a normal distribution. The data sets were subsequently processed by the competing methods. The following properties were determined randomly for each data set: (a) the sample size \(N\) varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into \(K = 2\), \(3\), and \(4\) groups; (b) the number of features varied between 2 and 5; (c) the standard deviation of all features was set to 1, 2 or 3 (the mean was always zero); (d) the correlation between all features was \(r = 0\), \(r = .1\), \(r = .2\), \(r = .3\), \(r = .4\) or \(r = .5\). All anticlustering methods were applied to each of the 10,000 data sets using \(K = 2\), \(K = 3\) and \(K = 4\) (i.e., each method was applied 30,000 times). Groups sizes were always equal.

\hypertarget{evaluation-criteria}{%
\subsection{Evaluation criteria}\label{evaluation-criteria}}

After the five anticlustering methods and the random allocation procedure had been applied to the 10,000 data sets for \(K = 2\), \(K = 3\), \(K = 4\), I investigated how well the competing methods were able to create similar groups. To this end, I analyzed the discrepancy in means, standard deviations, skewness, kurtosis, and correlation between groups---for each data set for each \(K\). To quantify the discrepancy in group means, the following computation was used: for each of the (2, 3, 4 or 5) features, all (2, 3 or 4) group means were computed. For each feature, the difference between the minimum and maximum group mean was used as a measure of discrepancy; the global discrepancy in means (\(\Delta M\)) was then determined as the average discrepancy across features. The same procedure was applied to compute the discrepancy in standard deviations (\(\Delta \mathit{SD}\)), skewness (\(\Delta \mathit{Skew}\)) and kurtosis (\(\Delta \mathit{Kurtosis}\)). Quantifying discrepancy in correlations (\(\Delta \mathit{Cor}\)) followed the same rule, but discrepancies in correlations had to be averaged across pairs of features instead of single features.

\hypertarget{results}{%
\subsection{Results}\label{results}}

Table 1 displays the global simulation results aggregated across all conditions. Standard k-means performed best at minimizing discrepancy with regard to means (\(\Delta \mathit{M}\)), but did not outperform a random assignment with regard to all other distribution characteristics. Replicating Papenberg and Klau (2021), k-means' performance with regard to similarity in standard deviations was even below that of a random assignment. K-plus anticlustering addressed \(\Delta \mathit{M}\) almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations (\(\Delta \mathit{SD}\)); all three k-plus approaches strongly outperformed all other methods with regard to minimizing \(\Delta \mathit{SD}\).

\begin{singlespace}

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:Table1}Results of the simulation study}

\begin{tabular}{rrrrrr}
\toprule
 & \multicolumn{1}{c}{$\Delta M$} & \multicolumn{1}{c}{$\Delta \mathit{SD}$} & \multicolumn{1}{c}{$\Delta \mathit{Skew}$} & \multicolumn{1}{c}{$\Delta \mathit{Kurtosis}$} & \multicolumn{1}{c}{$\Delta \mathit{Cor}$}\\
\midrule
diversity & 0.06 & 0.15 & 0.34 & 0.68 & 0.11\\
k-means & 0.02 & 0.39 & 0.57 & 0.86 & 0.26\\
k-plus & 0.03 & 0.03 & 0.62 & 1.03 & 0.24\\
k-plus-correlation & 0.05 & 0.04 & 0.61 & 1.02 & 0.03\\
k-plus-skew-kurtosis & 0.04 & 0.04 & 0.26 & 0.65 & 0.24\\
random & 0.53 & 0.38 & 0.54 & 0.85 & 0.25\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} The global results of the simulation study aggregated across all conditions. Cells contain information about the average global between-group discrepancy with regard to means, standard deviations, skewness, kurtosis, and correlations. Lower values indicate less discrepancy, i.e., higher between-group similarity. Each cell value is the result of averaging across 30,000 data points (10,000 data sets $\times$ $K = 2$, $K = 3$ or $K = 4$).}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\end{singlespace}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/Figure1-1.pdf}
\caption{\label{fig:Figure1}Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by \(N\), averaged across the remaining variables that varied in the simulation).}
\end{figure}

Figure 2 shows that with increasing \(N\), k-means and the k-plus methods converged on the same level with regard to minimizing \(\Delta \mathit{M}\). This is an important observation because it shows that k-plus anticlustering is capable of addressing multiple objectives at the same time---especially when \(N\) increases. However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria made it more difficult to fulfill each single criterion: k-plus skew/kurtosis and k-plus correlation performed worse at minimizing \(\Delta \mathit{M}\) and \(\Delta \mathit{SD}\) than standard k-plus. However, every k-plus method did what it was supposed to do: k-plus skew/kurtosis was best at minimizing \(\Delta \mathit{Skew}\) and \(\Delta \mathit{Kurtosis}\), k-plus correlation was best at minimizing discrepancy with regard to \(\Delta \mathit{Cor}\). At the same time, these two k-plus objectives maintained a comparably good level of addressing \(\Delta \mathit{M}\) and \(\Delta \mathit{SD}\), outperforming diversity anticlustering.

It is maybe surprising that k-plus skew/kurtosis does not seem to have performed much better than diversity anticlustering with regard to skewness and kurtosis, even though the method has been specifically tailored to these criteria. Figure 2 gives a more fine grained assessment of this observation: Surprisingly, for small \(N\), k-plus skew/kurtosis was even worse than diversity anticlustering at minimizing \(\Delta \mathit{Skew}\) and \(\Delta \mathit{Kurtosis}\). However, with increasing \(N\), it clearly outperformed diversity anticlustering. These results illustrate the cost of optimizing several criteria at the same time: For low \(N\), k-plus skew/kurtosis was apparently preoccupied with fulfilling the objectives \(\Delta \mathit{M}\) and \(\Delta \mathit{SD}\). Increasing \(N\) then facilitated to also address \(\Delta \mathit{Skew}\) and \(\Delta \mathit{Kurtosis}\).

Generally, it should be noted that diversity anticlustering is a good all-rounder method that tends to address all distribution characteristics. For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering was outperformed. As as side note, apart from k-plus correlation, diversity anticlustering was the only method that equalized correlation structure among groups. To the best of my knowledge, this observation has not been made previously, even though the maximum diversity problem has been widely studied.

The Appendix includes additional figures splitting the simulation results according to the other variables that varied between simulation runs (number of features; correlation between features; standard deviation of features; number of groups). Here, the only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased. Because the number of covariances increases quadratically when the number of features increases, it gets comparatively more difficult to fulfill the correlation objective with a higher number of features.

\hypertarget{applying-k-plus-anticlustering-using-the-r-package-anticlust}{%
\section{Applying k-plus anticlustering using the R package anticlust}\label{applying-k-plus-anticlustering-using-the-r-package-anticlust}}

This section demonstrates how readers can easily employ k-plus anticlustering using the R package \texttt{anticlust} (Papenberg \& Klau, 2021) using the function \texttt{kplus\_anticlustering()}, which is available from version 0.6.3 onward. The package can be installed in the R environment using the \texttt{install.packages()} command:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{install.packages}\NormalTok{(}\StringTok{"anticlust"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{example-application-i}{%
\subsection{Example Application I}\label{example-application-i}}

In the first example, I used the OASIS norming data that is freely available online (Kurdi et al., 2017). Kurdi et al.~assembled 900 open-access color images and collected ratings on two affective dimensions: arousal and valence. Brielmann and Pelli (2019) measured how the same 900 images were rated with regard to beauty. I used the function \texttt{kplus\_anticlustering()} to divide all images into 9 groups of 100 images each. The first argument of the function call is a variable named \texttt{features}, which contains the image data as a table: 900 rows representing images and 3 columns representing the features beauty, arousal, and valence. An OSF repository accompanying this manuscript contains the full code to reproduce the example (Papenberg, 2023).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(anticlust) }\CommentTok{\# load the package anticlust}

\FunctionTok{kplus\_anticlustering}\NormalTok{(}
\NormalTok{  features,}
  \AttributeTok{K =} \DecValTok{9}\NormalTok{,}
  \AttributeTok{variance =} \ConstantTok{TRUE}\NormalTok{,}
  \AttributeTok{skew =} \ConstantTok{TRUE}\NormalTok{,}
  \AttributeTok{kurtosis =} \ConstantTok{FALSE}\NormalTok{,}
  \AttributeTok{covariances =} \ConstantTok{FALSE}\NormalTok{,}
  \AttributeTok{moments =} \ConstantTok{NULL}\NormalTok{,}
  \AttributeTok{method =} \StringTok{"local{-}maximum"}\NormalTok{,}
  \AttributeTok{standardize =} \ConstantTok{TRUE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The function call takes about 10 seconds on a contemporary personal computer. The function \texttt{kplus\_anticlustering()} has four boolean arguments to specify whether variance, skewness, kurtosis and covariances should be included as part of the k-plus criterion. Only the argument \texttt{variance} is set to \texttt{TRUE} by default---corresponding to the understanding that k-plus is an extension to k-means that at least equalizes variances in addition to means. The arguments corresponding to skewness, kurtosis and covariance have to be specifically ``turned on'', i.e., by setting them to \texttt{TRUE} in the function call of \texttt{kplus\_anticlustering()}. If other higher order moments should be included as part of the optimization, the optional argument \texttt{moments} can be used to specify the desired moments as an integer vector. The argument \texttt{method\ =\ "local-maximum"} ensures that the k-plus criterion is optimized using the local maximum search method described in the previous section. The argument \texttt{K} describes the number of groups. By default, the function ensures that all features are standardized before the optimization starts; in particular, this option enforces that all k-plus criteria receive the same weight during the optimization process. This behaviour can be adjusted using the boolean argument \texttt{standardize}.

Table 2 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features. Table 2 also contains the results after applying k-means anticlustering and diversity anticlustering. It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations, and the skewness was also very evenly matched. K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations and skewness. Diversity anticlustering was well-suited to match means, standard deviations and skew, but was slightly outperformed by k-plus. The latter result is not surprising because---unlike k-plus anticlustering---diversity anticlustering does not directly maximize similarity with regard to these criteria, but instead maximizes within-group heterogeneity.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-2}Descriptive statistics for OASIS features by group and anticlustering method.}

\begin{tabular}{crrrrrrrrr}
\toprule
 & \multicolumn{3}{c}{Beauty} & \multicolumn{3}{c}{Valence} & \multicolumn{3}{c}{Arousal} \\
\cmidrule(r){2-4} \cmidrule(r){5-7} \cmidrule(r){8-10}
Group & \multicolumn{1}{c}{$M$} & \multicolumn{1}{c}{$\mathit{SD}$} & \multicolumn{1}{c}{Skew} & \multicolumn{1}{c}{$M$} & \multicolumn{1}{c}{$\mathit{SD}$} & \multicolumn{1}{c}{Skew} & \multicolumn{1}{c}{$M$} & \multicolumn{1}{c}{$\mathit{SD}$} & \multicolumn{1}{c}{Skew}\\
\midrule
\textbf{K-plus anticlustering} &  &  &  &  &  &  &  &  & \\
\ \ \ 1 & 4.34 & 1.13 & -0.28 & 4.33 & 1.23 & -0.49 & 3.67 & 0.84 & -0.32\\
\ \ \ 2 & 4.34 & 1.13 & -0.29 & 4.33 & 1.23 & -0.49 & 3.67 & 0.84 & -0.33\\
\ \ \ 3 & 4.34 & 1.13 & -0.29 & 4.33 & 1.23 & -0.48 & 3.67 & 0.84 & -0.34\\
\ \ \ 4 & 4.34 & 1.13 & -0.28 & 4.33 & 1.23 & -0.49 & 3.67 & 0.84 & -0.33\\
\ \ \ 5 & 4.34 & 1.13 & -0.28 & 4.33 & 1.23 & -0.49 & 3.67 & 0.84 & -0.33\\
\ \ \ 6 & 4.34 & 1.13 & -0.29 & 4.33 & 1.23 & -0.49 & 3.67 & 0.84 & -0.33\\
\ \ \ 7 & 4.34 & 1.13 & -0.29 & 4.33 & 1.23 & -0.49 & 3.67 & 0.84 & -0.33\\
\ \ \ 8 & 4.34 & 1.13 & -0.28 & 4.33 & 1.23 & -0.49 & 3.67 & 0.84 & -0.33\\
\ \ \ 9 & 4.34 & 1.13 & -0.28 & 4.33 & 1.23 & -0.49 & 3.67 & 0.84 & -0.33\\
\textbf{K-means anticlustering} &  &  &  &  &  &  &  &  & \\
\ \ \ 1 & 4.34 & 1.20 & -0.41 & 4.33 & 1.17 & -0.46 & 3.67 & 0.79 & -0.26\\
\ \ \ 2 & 4.34 & 1.18 & -0.27 & 4.33 & 1.32 & -0.50 & 3.67 & 0.79 & -0.48\\
\ \ \ 3 & 4.34 & 1.14 & -0.29 & 4.33 & 1.28 & -0.48 & 3.67 & 0.84 & -0.40\\
\ \ \ 4 & 4.34 & 1.19 & -0.12 & 4.33 & 1.20 & -0.32 & 3.67 & 0.83 & -0.51\\
\ \ \ 5 & 4.34 & 1.00 & 0.02 & 4.33 & 1.14 & -0.34 & 3.67 & 0.92 & -0.28\\
\ \ \ 6 & 4.34 & 1.02 & -0.40 & 4.33 & 1.12 & -0.43 & 3.67 & 0.91 & -0.33\\
\ \ \ 7 & 4.34 & 1.11 & -0.51 & 4.33 & 1.21 & -0.78 & 3.67 & 0.89 & -0.30\\
\ \ \ 8 & 4.34 & 1.18 & -0.32 & 4.33 & 1.30 & -0.50 & 3.67 & 0.78 & -0.31\\
\ \ \ 9 & 4.34 & 1.11 & -0.22 & 4.33 & 1.34 & -0.52 & 3.67 & 0.80 & -0.07\\
\textbf{Diversity anticlustering} &  &  &  &  &  &  &  &  & \\
\ \ \ 1 & 4.34 & 1.14 & -0.30 & 4.34 & 1.23 & -0.47 & 3.67 & 0.84 & -0.33\\
\ \ \ 2 & 4.34 & 1.12 & -0.27 & 4.33 & 1.23 & -0.49 & 3.67 & 0.85 & -0.25\\
\ \ \ 3 & 4.34 & 1.12 & -0.27 & 4.33 & 1.24 & -0.50 & 3.67 & 0.84 & -0.36\\
\ \ \ 4 & 4.34 & 1.13 & -0.29 & 4.33 & 1.23 & -0.49 & 3.67 & 0.84 & -0.32\\
\ \ \ 5 & 4.34 & 1.13 & -0.30 & 4.33 & 1.23 & -0.47 & 3.67 & 0.84 & -0.33\\
\ \ \ 6 & 4.34 & 1.14 & -0.34 & 4.33 & 1.23 & -0.47 & 3.67 & 0.84 & -0.39\\
\ \ \ 7 & 4.34 & 1.12 & -0.26 & 4.33 & 1.24 & -0.51 & 3.67 & 0.84 & -0.34\\
\ \ \ 8 & 4.35 & 1.13 & -0.29 & 4.33 & 1.24 & -0.50 & 3.67 & 0.84 & -0.32\\
\ \ \ 9 & 4.34 & 1.12 & -0.25 & 4.33 & 1.23 & -0.50 & 3.67 & 0.84 & -0.33\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} The k-plus criterion that was optimized included a term to minimize differences with regard to means, variances, and skewness.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups. K-plus anticlustering strives for between-group similarity regardless of group size. Unequal group sizes can be requested by passing the different group sizes to the argument \texttt{K}. The following code was used to equalize means and variances between four groups using the k-plus approach:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{kplus\_anticlustering}\NormalTok{(}
\NormalTok{  features,}
  \AttributeTok{K =} \FunctionTok{c}\NormalTok{(}\DecValTok{100}\NormalTok{, }\DecValTok{100}\NormalTok{, }\DecValTok{100}\NormalTok{, }\DecValTok{600}\NormalTok{),}
  \AttributeTok{method =} \StringTok{"local{-}maximum"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-3}Performance of the diversity and k-plus objectives for unequal group sizes.}

\begin{tabular}{rrrrr}
\toprule
Group & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Beauty} & \multicolumn{1}{c}{Valence} & \multicolumn{1}{c}{Arousal}\\
\midrule
\textbf{K-plus anticlustering} &  &  &  & \\
\ \ \ 1 & 100 & 4.34 (1.13) & 4.33 (1.23) & 3.67 (0.84)\\
\ \ \ 2 & 100 & 4.34 (1.13) & 4.33 (1.23) & 3.67 (0.84)\\
\ \ \ 3 & 100 & 4.34 (1.13) & 4.33 (1.23) & 3.67 (0.84)\\
\ \ \ 4 & 600 & 4.34 (1.12) & 4.33 (1.23) & 3.67 (0.84)\\
\textbf{K-means anticlustering} &  &  &  & \\
\ \ \ 1 & 100 & 4.34 (1.11) & 4.33 (1.20) & 3.67 (0.94)\\
\ \ \ 2 & 100 & 4.34 (1.12) & 4.33 (1.14) & 3.67 (0.81)\\
\ \ \ 3 & 100 & 4.34 (1.12) & 4.33 (1.20) & 3.67 (0.76)\\
\ \ \ 4 & 600 & 4.34 (1.13) & 4.33 (1.25) & 3.67 (0.84)\\
\textbf{Diversity anticlustering} &  &  &  & \\
\ \ \ 1 & 100 & 4.43 (0.62) & 4.53 (0.76) & 3.64 (0.45)\\
\ \ \ 2 & 100 & 4.43 (0.61) & 4.54 (0.76) & 3.64 (0.45)\\
\ \ \ 3 & 100 & 4.43 (0.62) & 4.53 (0.76) & 3.64 (0.45)\\
\ \ \ 4 & 600 & 4.30 (1.30) & 4.23 (1.39) & 3.68 (0.98)\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} The cells contain mean values by cluster and the standard deviation (in parentheses).}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{example-application-ii-small-data-set-n-96}{%
\subsection{\texorpdfstring{Example Application II: Small data set (\(N = 96\))}{Example Application II: Small data set (N = 96)}}\label{example-application-ii-small-data-set-n-96}}

The simulation study indicated that k-plus may show decreased performance for smaller group sizes. Therefore, a second application illustrates that k-plus anticlustering yields satisfactory results in practice even when group sizes are smaller. I make use of a norming data set of 96 word stimuli that was contributed to the \texttt{anticlust} package by Marie L. Schaper (Schaper et al., 2019a, 2019b). After loading the \texttt{anticlust} package, the data set can be accessed as follows:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{data}\NormalTok{(schaper2019)}
\end{Highlighting}
\end{Shaded}

Table 4 illustrates the results that are obtained when using standard\footnote{I refer to \emph{standard} k-plus anticlustering as the minimization of differences in means and variances, but not in higher order moments or covariances.} k-plus anticlustering to divide the stimulus set into 6 groups of size 16 each. Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are similar enough for any practical purposes.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-6}Descriptive statistics by group for the Schaper et al. (2019a; 2019b) data set, after applying k-plus anticlustering.}

\begin{tabular}{rrrrrr}
\toprule
Group & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Typicality} & \multicolumn{1}{c}{Atypicality} & \multicolumn{1}{c}{Syllables} & \multicolumn{1}{c}{Frequency}\\
\midrule
1 & 16 & 4.49 (0.25) & 1.10 (0.07) & 3.44 (0.96) & 18.38 (2.45)\\
2 & 16 & 4.49 (0.26) & 1.10 (0.07) & 3.38 (0.96) & 18.38 (2.45)\\
3 & 16 & 4.49 (0.26) & 1.11 (0.07) & 3.44 (0.96) & 18.25 (2.44)\\
4 & 16 & 4.49 (0.26) & 1.10 (0.07) & 3.44 (0.89) & 18.25 (2.46)\\
5 & 16 & 4.50 (0.25) & 1.10 (0.07) & 3.38 (0.96) & 18.38 (2.42)\\
6 & 16 & 4.49 (0.25) & 1.10 (0.07) & 3.44 (0.96) & 18.25 (2.49)\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} The cells contain mean values by cluster and the standard deviation (in parentheses).}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Anticlustering is a partitioning method that aims for high between-group similarity and high within-group heterogeneity. In this paper, I presented the k-plus criterion for anticlustering, which focuses on maximizing between-group similarity. K-plus represents between-group similarity as discrepancies in distribution moments between groups: means, variances, skewness, kurtosis, higher order moments, and covariances. Depending on the application, a different subset of these objectives can be included in the anticlustering process. The k-plus criterion is an extension of the classical k-means criterion, which only reflects how similar groups are with regard to their means. Interestingly, however, the k-plus extension does not change the original k-means criterion because additional criteria can be incorporated by augmenting the input data.

A simulation study and examples on real norming data showed that the k-plus criterion is well-suited to maximize between-group similarity with regard to multiple criteria. Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be considered as the default replacement of classical k-means when striving for between-group similarity.\footnote{A reasonable exception is given when the input data consists of binary variables because for binary data, the mean and the variance are directly dependent.} The results did however show that fulfilling multiple criteria at the same time can be challenging in small data sets. In practice, users can try out different objectives and see which specification fulfills their needs. A reasonable approach is to start ``greedy'', e.g.~by specifying that multiple distribution moments should equalized. If the results of such a greedy setup are not satisfying, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variance). Using the open source R package \texttt{anticlust}, the (repeated) application of k-plus anticlustering is easily accessible and free.

\hypertarget{limitations-and-outlook}{%
\subsection{Limitations and outlook}\label{limitations-and-outlook}}

The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion. A simple unweighted sum approach was used to optimize the k-plus objective, while more sophisticated algorithmic procedures exist to tackle multicriterion optimization problems (e.g., Brusco \& Steinley, 2009). The sole focus on the objective function followed two reasonings. First, the applications showed that the local maximum search, maximizing the unweighted k-plus sum, yielded satisfying results that probably suffice in most applications. It is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria to a satisfactory degree (as suggested by Figure 1). Still, more sophisticated approaches may yield theoretically improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives. Brusco et al. (2020) gave an example of how direct algorithms can be employed to optimize multiple criteria in anticlustering applications aiming for high within-group heterogeneity. Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment. A second reason for focusing on the k-plus objective per se---independently from the algorithm that is used to optimize it---was to obtain a fair comparison to existing anticlustering objectives. In the simulation study, differences in performance could not have been interpreted if the objective function as well as the algorithm used to optimize it had differed between anticlustering methods. It was of particular interest to compare the k-plus objective with the popular diversity objective. Papenberg and Klau (2021) compared k-means anticlustering and diversity anticlustering\footnote{In their study, they used the term anticluster editing to refer to the maximization of the diversity criterion. This is because the diversity criterion is minimized as part of the cluster editing method (Shamir et al., 2004).} in a simulation study and concluded that the diversity criterion should be preferred because k-means does not equalize the spread of the distribution between groups. In contrast, maximizing the diversity leads to an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion has regained attractivity: When opting for similarity in means and variances---and for increasing \(N\), skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

This paper introduced k-plus anticlustering for maximizing between-group similarity, improving the classical k-means approach. K-plus anticlustering can be adopted whenever a group of elements has to be partitioned into equivalent parts. The method is accessible and easily usable via the R package \texttt{anticlust}.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-aloise2009np}{}}%
Aloise, D., Deshpande, A., Hansen, P., \& Popat, P. (2009). NP-hardness of euclidean sum-of-squares clustering. \emph{Machine Learning}, \emph{75}(2), 245--248. \url{https://doi.org/10.1007/s10994-009-5103-0}

\leavevmode\vadjust pre{\hypertarget{ref-anscombe1973}{}}%
Anscombe, F. J. (1973). Graphs in statistical analysis. \emph{The American Statistician}, \emph{27}(1), 17--21. \url{https://doi.org/10.1080/00031305.1973.10478966}

\leavevmode\vadjust pre{\hypertarget{ref-R-papaja}{}}%
Aust, F., \& Barth, M. (2018). \emph{{papaja}: {Create} {APA} manuscripts with {R Markdown}}. \url{https://github.com/crsh/papaja}

\leavevmode\vadjust pre{\hypertarget{ref-baker2002}{}}%
Baker, K. R., \& Powell, S. G. (2002). Methods for assigning students to groups: A study of alternative objective functions. \emph{Journal of the Operational Research Society}, \emph{53}(4), 397--404. \url{https://doi.org/10.1057/palgrave.jors.2601307}

\leavevmode\vadjust pre{\hypertarget{ref-brielmann2019intense}{}}%
Brielmann, A. A., \& Pelli, D. G. (2019). Intense beauty requires intense pleasure. \emph{Frontiers in Psychology}, \emph{10}, Article 2420. \url{https://doi.org/10.3389/fpsyg.2019.02420}

\leavevmode\vadjust pre{\hypertarget{ref-brusco2006branch}{}}%
Brusco, M. J. (2006). A repetitive branch-and-bound procedure for minimum within-cluster sums of squares partitioning. \emph{Psychometrika}, \emph{71}(2), 347--363. \url{https://doi.org/10.1007/s11336-004-1218-1}

\leavevmode\vadjust pre{\hypertarget{ref-brusco2019}{}}%
Brusco, M. J., Cradit, J. D., \& Steinley, D. (2020). Combining diversity and dispersion criteria for anticlustering: A bicriterion approach. \emph{British Journal of Mathematical and Statistical Psychology}, \emph{73}(3), 375--396. \url{https://doi.org/10.1111/bmsp.12186}

\leavevmode\vadjust pre{\hypertarget{ref-brusco2009cross}{}}%
Brusco, M. J., \& Steinley, D. (2009). Cross validation issues in multiobjective clustering. \emph{British Journal of Mathematical and Statistical Psychology}, \emph{62}(2), 349--368. \url{https://doi.org/10.1348/000711008X304385}

\leavevmode\vadjust pre{\hypertarget{ref-R-faux}{}}%
DeBruine, L. (2021). \emph{Faux: Simulation for factorial designs}. Zenodo. \url{https://doi.org/10.5281/zenodo.2669586}

\leavevmode\vadjust pre{\hypertarget{ref-dry2009}{}}%
Dry, M. J., \& Storms, G. (2009). Similar but not the same: A comparison of the utility of directly rated and feature-based similarity measures for generating spatial models of conceptual data. \emph{Behavior Research Methods}, \emph{41}(3), 889--900. \url{https://doi.org/10.3758/BRM.41.3.889}

\leavevmode\vadjust pre{\hypertarget{ref-ferligoj1992direct}{}}%
Ferligoj, A., \& Batagelj, V. (1992). Direct multicriteria clustering algorithms. \emph{Journal of Classification}, \emph{9}(1), 43--61. \url{https://doi.org/10.1007/BF02618467}

\leavevmode\vadjust pre{\hypertarget{ref-fernandez2013}{}}%
Fernández, E., Kalcsics, J., \& Nickel, S. (2013). The maximum dispersion problem. \emph{Omega}, \emph{41}(4), 721--730. \url{https://doi.org/10.1016/j.omega.2012.09.005}

\leavevmode\vadjust pre{\hypertarget{ref-friendly1982}{}}%
Friendly, M., Franklin, P. E., Hoffman, D., \& Rubin, D. C. (1982). The {Toronto} word pool: Norms for imagery, concreteness, orthographic variables, and grammatical usage for 1,080 words. \emph{Behavior Research Methods \& Instrumentation}, \emph{14}(4), 375--399. \url{https://doi.org/10.3758/BF03203275}

\leavevmode\vadjust pre{\hypertarget{ref-gallego2013}{}}%
Gallego, M., Laguna, M., Marti, R., \& Duarte, A. (2013). Tabu search with strategic oscillation for the maximally diverse grouping problem. \emph{Journal of the Operational Research Society}, \emph{64}(5), 724--734. \url{https://doi.org/10.1057/jors.2012.66}

\leavevmode\vadjust pre{\hypertarget{ref-gierl2017}{}}%
Gierl, M., Daniels, L., \& Zhang, X. (2017). Creating parallel forms to support on-demand testing for undergraduate students in psychology. \emph{Journal of Measurement and Evaluation in Education and Psychology}, \emph{8}(3), 288--302. \url{https://doi.org/10.21031/epod.305350}

\leavevmode\vadjust pre{\hypertarget{ref-jain2010}{}}%
Jain, A. K. (2010). Data clustering: 50 years beyond k-means. \emph{Pattern Recognition Letters}, \emph{31}(8), 651--666. \url{https://doi.org/10.1016/j.patrec.2009.09.011}

\leavevmode\vadjust pre{\hypertarget{ref-joanes1998}{}}%
Joanes, D. N., \& Gill, C. A. (1998). Comparing measures of sample skewness and kurtosis. \emph{Journal of the Royal Statistical Society: Series D (The Statistician)}, \emph{47}(1), 183--189. \url{https://doi.org/10.1111/1467-9884.00122}

\leavevmode\vadjust pre{\hypertarget{ref-kurdi2017introducing}{}}%
Kurdi, B., Lozano, S., \& Banaji, M. R. (2017). Introducing the open affective standardized image set (OASIS). \emph{Behavior Research Methods}, \emph{49}(2), 457--470. \url{https://doi.org/10.3758/s13428-016-0715-3}

\leavevmode\vadjust pre{\hypertarget{ref-lahl2006}{}}%
Lahl, O., \& Pietrowsky, R. (2006). EQUIWORD: A software application for the automatic creation of truly equivalent word lists. \emph{Behavior Research Methods}, \emph{38}(1), 146--152. \url{https://doi.org/10.3758/BF03192760}

\leavevmode\vadjust pre{\hypertarget{ref-marler2010}{}}%
Marler, R. T., \& Arora, J. S. (2010). The weighted sum method for multi-objective optimization: New insights. \emph{Structural and Multidisciplinary Optimization}, \emph{41}(6), 853--862. \url{https://doi.org/10.1007/s00158-009-0460-7}

\leavevmode\vadjust pre{\hypertarget{ref-naidu2014}{}}%
Naidu, K., Mokhlis, H., \& Bakar, A. A. (2014). Multiobjective optimization using weighted sum artificial bee colony algorithm for load frequency control. \emph{International Journal of Electrical Power \& Energy Systems}, \emph{55}, 657--667. \url{https://doi.org/10.1016/j.ijepes.2013.10.022}

\leavevmode\vadjust pre{\hypertarget{ref-R-anticlust}{}}%
Papenberg, M. (2019). \emph{Anticlust: Subset partitioning via anticlustering}. \url{https://cran.r-project.org/package=anticlust}

\leavevmode\vadjust pre{\hypertarget{ref-osf2023}{}}%
Papenberg, M. (2023). \emph{K-plus anticlustering {[}{Open Science Framework Repository}{]}}. \url{https://doi.org/10.17605/OSF.IO/7G3K2}

\leavevmode\vadjust pre{\hypertarget{ref-papenberg2020}{}}%
Papenberg, M., \& Klau, G. W. (2021). Using anticlustering to partition data sets into equivalent parts. \emph{Psychological Methods}, \emph{26}(2), 161--174. \url{https://doi.org/10.1037/met0000301}

\leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
R Core Team. (2022). \emph{R: A language and environment for statistical computing}. \url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-rokach2005}{}}%
Rokach, L., \& Maimon, O. (2005). Clustering methods. In O. Maimon \& L. Rokach (Eds.), \emph{Data mining and knowledge discovery handbook} (pp. 321--352). Springer. \url{https://doi.org/10.1007/0-387-25465-X_15}

\leavevmode\vadjust pre{\hypertarget{ref-schaper2019metacognitive}{}}%
Schaper, M. L., Kuhlmann, B. G., \& Bayen, U. J. (2019a). Metacognitive expectancy effects in source monitoring: Beliefs, in-the-moment experiences, or both? \emph{Journal of Memory and Language}, \emph{107}, 95--110. \url{https://doi.org/10.1016/j.jml.2019.03.009}

\leavevmode\vadjust pre{\hypertarget{ref-schaper2019metamory}{}}%
Schaper, M. L., Kuhlmann, B. G., \& Bayen, U. J. (2019b). Metamemory expectancy illusion and schema-consistent guessing in source monitoring. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{45}(3), 470--496. \url{https://doi.org/10.1037/xlm0000602}

\leavevmode\vadjust pre{\hypertarget{ref-shamir2004vf}{}}%
Shamir, R., Sharan, R., \& Tsur, D. (2004). {Cluster graph modification problems}. \emph{Discrete Applied Mathematics}, \emph{144}, 173--182. \url{https://doi.org/10.1016/j.dam.2004.01.007}

\leavevmode\vadjust pre{\hypertarget{ref-R-desctools}{}}%
Signorell, A. (2022). \emph{{DescTools}: Tools for descriptive statistics}. \url{https://cran.r-project.org/package=DescTools}

\leavevmode\vadjust pre{\hypertarget{ref-spath1986}{}}%
Späth, H. (1986). Anticlustering: Maximizing the variance criterion. \emph{Control and Cybernetics}, \emph{15}(2), 213--218.

\leavevmode\vadjust pre{\hypertarget{ref-steinley2006}{}}%
Steinley, D. (2006). K-means clustering: A half-century synthesis. \emph{British Journal of Mathematical and Statistical Psychology}, \emph{59}(1), 1--34. \url{https://doi.org/10.1348/000711005X48266}

\leavevmode\vadjust pre{\hypertarget{ref-tversky1977}{}}%
Tversky, A. (1977). Features of similarity. \emph{Psychological Review}, \emph{84}(4), 327--352. \url{https://doi.org/10.1037/0033-295X.84.4.327}

\leavevmode\vadjust pre{\hypertarget{ref-valev1983}{}}%
Valev, V. (1983). Set partition principles. In J. Kozesnik (Ed.), \emph{Transactions of the ninth {Prague} conference on information theory, statistical decision functions, and random processes ({Prague}, 1982)} (pp. 251--256). Springer Netherlands.

\leavevmode\vadjust pre{\hypertarget{ref-valev1998}{}}%
Valev, V. (1998). Set partition principles revisited. In A. Amin, D. Dori, P. Pudil, \& H. Freeman (Eds.), \emph{Advances in pattern recognition. {SSPR /SPR 1998}. Lecture notes in computer science} (pp. 875--881). Springer.

\leavevmode\vadjust pre{\hypertarget{ref-weitz1998}{}}%
Weitz, R., \& Lakshminarayanan, S. (1998). An empirical comparison of heuristic methods for creating maximally diverse groups. \emph{Journal of the Operational Research Society}, \emph{49}(6), 635--646. \url{https://doi.org/10.1057/palgrave.jors.2600510}

\leavevmode\vadjust pre{\hypertarget{ref-R-ggplot2}{}}%
Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}

\leavevmode\vadjust pre{\hypertarget{ref-R-tidyr}{}}%
Wickham, H. (2021). \emph{Tidyr: Tidy messy data}. \url{https://CRAN.R-project.org/package=tidyr}

\leavevmode\vadjust pre{\hypertarget{ref-R-dplyr}{}}%
Wickham, H., François, R., Henry, L., \& Müller, K. (2019). \emph{Dplyr: A grammar of data manipulation}. \url{https://CRAN.R-project.org/package=dplyr}

\end{CSLReferences}

\endgroup

\newpage

\hypertarget{appendix-appendix}{%
\appendix}

\hypertarget{section}{%
\section{}\label{section}}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/Figure2-1.pdf}
\caption{\label{fig:Figure2}Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by the number of features, averaged across the remaining variables that varied in the simulation).}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/Figure3-1.pdf}
\caption{\label{fig:Figure3}Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by \emph{SD}, averaged across the remaining variables that varied in the simulation).}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/Figure4-1.pdf}
\caption{\label{fig:Figure4}Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by \emph{K}, averaged across the remaining variables that varied in the simulation).}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/Figure5-1.pdf}
\caption{\label{fig:Figure5}Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness, kurtosis and correlations (split by \emph{r}, averaged across the remaining variables that varied in the simulation).}
\end{figure}

\end{document}

### 2023-05-04 paper.Rmd
this is consistent with the other ones

@@
 Note that there is a direct connection between the $\mathit{SSE}$ and the location of the cluster centroids [@spath1986].
-If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_j$ ($j = 1, \ldots, P$) is the overall mean value on the $j$th attribute:
+If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_j$ ($j = 1, \ldots, P$) is the overall mean value on the $j$'th attribute:
 

### 2023-05-04 paper.Rmd
minor

@@
 When defining a new variable $y = (y_1, \ldots, y_N)$ as the squared deviation of each observation from the mean $\overline{\mathbf{x}}$, i.e.
-$y_i = (x_i - \overline{\mathbf{x}})^2$, the sample variance of $x$ can be reformulated as the mean of $y$:
+$y_i = (x_i - \overline{\mathbf{x}})^2$ ($i = 1, \dots, N$), the sample variance of $x$ can be reformulated as the mean of $y$:
 
@@
 Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict with each other (Ferligoj & Batagelj, 1992).
-As $\mathit{SSE_{kplus}}$  is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria are similar between groups, i.e., whether these objectives are in alignment or conflicting.
+As $\mathit{SSE_{kplus}}$  is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria are similar between groups, i.e., whether these objectives are aligned or in conflict.
 A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible.
@@
 In the simulation studies, five repetitions were used.
+Note that the local maximum search can be performed on the basis of any arbitrary initial grouping.
+Specifically, the initial grouping can also be chosen in such a way that the groups are different sized; for example, if the data set cannot evenly be split, it is possible to ensure that the group sizes differ by 1 at most.
+However, the data sets in the simulation studies were always partitioned into equal-sized groups.
 
@@
 Two unweighted approaches were implemented by defining $\mathbf{w} = (1, 1, 1, 1)$.
-The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented variables---of the input data $\mathbf{X'}_{N \times \mathit{TP}}$ before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
+The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented k-plus variables---of the input data $\mathbf{X'}_{N \times \mathit{TP}}$ before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
 In the second case, the unstandardized matrix $\mathbf{X'}_{N \times \mathit{TP}}$ was used.
@@
 
-## Simulation 2: Comparing k-plus anticlustering to other objectives
+## Simulation 2: Comparing k-plus to other anticlustering objectives
 
@@
 "Generalized" k-plus anticlustering was implemented by maximizing $\mathit{SSE_{kplusGeneralized}}$ using $T = 4$, thus minimizing differences with regard to group means, variances, skewness and kurtosis.
-To implement the k-plus objectives, the unweighted standardization approach was employed that performed best in Simulation 1.
+The unweighted standardization approach, that performed best in Simulation 1, was used as the benchmark implementation for k-plus anticlustering that was compared to the competing objectives.
 Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.

### 2023-05-04 paper.Rmd
more editing

@@
 When defining a new variable $y = (y_1, \ldots, y_N)$ as the squared deviation of each observation from the mean $\overline{\mathbf{x}}$, i.e.
-$y_i = (x_i - \overline{\mathbf{x}})^2$ ($i = 1, \dots, N$), the sample variance of $x$ can be reformulated as the mean of $y$:
+$y_i = (x_i - \overline{\mathbf{x}})^2$, the sample variance of $x$ can be reformulated as the mean of $y$:
 
@@
 
-Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict with each other (Ferligoj & Batagelj, 1992).
+Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict (Ferligoj & Batagelj, 1992).
 As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria are similar between groups, i.e., whether these objectives are aligned or in conflict.
@@
 Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree.
-This result indicates that the objectives $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are not in conflict with each other.
-Moreover, in Figure 1 there is a high concentration of partitions in the upper-right corner where the two objectives are in high agreement.
 
@@
 
-Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{\mathbf{X}}^{(k)}_{T \times P} = \{\overline{x}_{tj}^{(k)}\}_{T \times P}$ is a centroid matrix containing the mean values for all k-plus variables in the $k$'th group, i.e.
+Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{\mathbf{X}}^{(k)}_{T \times P} = \{\overline{x}_{tj}^{(k)}\}_{T \times P}$ is a matrix containing the mean values for all k-plus variables in the $k$'th group, i.e.
 
@@
 
-Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ has the same form as the standard k-means criterion $\mathit{SSE}$, but it is computed on different data.
-Computing the k-plus criterion is then done by first creating the additional k-plus variables, appending them as new columns to the original data matrix, and computing the $\mathit{SSE}$ on the augmented data.
+Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ has the same form as the standard k-means criterion $\mathit{SSE}$, but is based on different data.
+Here, computing the k-plus criterion is done by first creating the additional k-plus variables, appending them as new columns to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data.
 Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
@@
 Two unweighted approaches were implemented by defining $\mathbf{w} = (1, 1, 1, 1)$.
-The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented k-plus variables---of the input data $\mathbf{X'}_{N \times \mathit{TP}}$ before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
-In the second case, the unstandardized matrix $\mathbf{X'}_{N \times \mathit{TP}}$ was used.
+The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented k-plus variables---of the input data before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
+In the second case, the unstandardized data matrix was used.
 Four weighted approaches were implemented by giving a tenfold increased weight to one of the four moments, respectively.
@@
 To compare the different weighting schemes, `r length(results)` data sets from were generated from a normal distribution ($M = 0$, $\mathit{SD} = 1$).
-The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$ so that equal-sized groups could be created.
+The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$.
 For each data set, $\mathit{SSE_{kplusGeneralized}}$ was optimized using the local maximum algorithm, once for each of the six competing weighting schemes.
@@
 "Generalized" k-plus anticlustering was implemented by maximizing $\mathit{SSE_{kplusGeneralized}}$ using $T = 4$, thus minimizing differences with regard to group means, variances, skewness and kurtosis.
-The unweighted standardization approach, that performed best in Simulation 1, was used as the benchmark implementation for k-plus anticlustering that was compared to the competing objectives.
+The unweighted standardization approach that performed best in Simulation 1 was used as the benchmark implementation for k-plus anticlustering that was compared to the competing objectives.
 Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.
@@
 The ARI is an index of the similarity between two partitions, where an index of 0 indicates no similarity and a value of 1 indicates the highest possible similarity (i.e., when two partitions are the same).
-The pairwise ARIs were computed for each data set and for each $K$, i.e., 30,000 times.
+The pairwise ARIs were computed for each data set and for each $K$.
 
@@
 Moreover, the ARI between two anticlustering partitions hardly differed from the ARI between an anticlustering partition and a random partition.
-This constitutes an interesting and novel result, illustrating that high between-group similarity can be achieved via very different partitionings.
+This constitutes an interesting result, illustrating that high between-group similarity can be achieved via very different partitionings.
 
@@
 The k-plus criterion is an extension of the classical k-means criterion, which only reflects how similar groups are with regard to their means.
-Interestingly, however, the k-plus extension does not change the original k-means criterion because additional criteria can be incorporated by augmenting the input data.
+Interestingly, when using the unweighted sum approach presented here, the k-plus extension does not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
 
@@
 Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.
+Specifically, it should be investigated if the bicriterion approach by Brusco et al.
+can be transformed into a multicriterion algorithm for the k-plus objective.
 A second reason for focusing on the k-plus objective per se---independently from the algorithm that is used to optimize it---was to obtain a fair comparison to existing anticlustering objectives.
@@
 Using the k-plus extension, however, the k-means criterion has regained attractivity: When opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.
+Future research should investigate whether this advantage also holds for other data distributions than normal.
+For example, it may be hypothesized that a generalized k-plus criterion is even more useful when applied to data that is not symmetric or contains outliers, because it is able to specifically incorporate the skewness and kurtosis as criteria in the anticlustering process.
 
 As an additional outlook, it should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariances, a possibility that was not investigated in the current paper.
-In this case, an additional variable has to be computed for each pair of variables $x = (x_1, \dots, x_N)$ and $y = (y_1, \dots, y_N)$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$.
-Future research should examine how the k-plus approach can be adapted to including covariances.
-The unweighted standardization approach presented here may serve as a starting point for investigation.
+In this case, we have to compute an additional variable for each of pair of variables whose covariance should be similar between groups.
+In future research, the unweighted standardization approach presented here may serve as a starting point for investigating covariance similarity in anticlustering applications.
 
@@
 Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, as was also shown in Simulation 2.
-In contrast, different clustering solutions tend to look similar because a good clustering is characterized by assigning similar elements into the same cluster and dissimilar elements into different clusters.
-A good anticlustering solution however does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This distinction can also be clarified using Figure 1, where we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.
+In contrast, different clustering solutions tend to look similar because a good clustering is always characterized by assigning similar elements into the same cluster and dissimilar elements into different clusters.
+A good anticlustering solution however does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This result was observed in Simulation 2 and can also be clarified using Figure 1.
+Here, we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.
 An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution.

### 2023-05-05 paper.Rmd
another round of editing

@@
 Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis.
-After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional anticlustering criteria such as the diversity and the classical k-means objective.
+After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional criteria such as the diversity and the k-means objective.
 Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].
@@
 
-In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variances between groups instead of only means.
+In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variances between groups.
 To motivate the new criterion, we first consider a set of one-dimensional observations.
@@
 
-Thus, to obtain an adaptation of the $\mathit{SSE}$ that can be maximized in order to minimize differences in variances between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
+Thus, to obtain an adaptation of the $\mathit{SSE}$ whose maximization can result in minimal differences in variances between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
@@
 However, the input data reflects the squared difference between each data point and the attribute mean instead of the raw data points.
-Maximizing $\mathit{SSE_{Var}}$ can be used to create groups that have similar variances across the $P$ attributes.
-As a caveat, it is crucial to note that maximizing $\mathit{SSE_{Var}}$ only reliably yields similar group variances if the group centroids $\overline{\mathbf{x}}^{(k)}$ ($k = 1, \dots, K$) are close to the overall data centroid $\overline{\mathbf{x}}$.
-This is because the within-group variance is calculated on the basis of the group mean $\overline{\mathbf{x}}^{(k)}$, but the additional variables computed to minimize differences with regard to the variance across groups are calculated using the overall mean $\overline{\mathbf{x}}$.
-Thus, $\mathit{SSE_{Var}}$ is only valid if (approximately) $\overline{\mathbf{x}}$ = $\overline{\mathbf{x}}^{(k)}$; otherwise we may introduce non-negligible measurement error when optimizing $\mathit{SSE_{Var}}$.
-Thus, only considering the new variables $\mathbf{Y}_{N \times P}$ in order to minimize discrepancies in variance may not yield satisfying results; instead, it has to be ensured that the group means of the original variables are also similar between groups.
-Fortunately, minimizing discrepancy with regard to group means can be accomplished by standard k-means anticlustering, i.e., the maximization of $\mathit{SSE}$.
+Maximizing $\mathit{SSE_{Var}}$ can be used to obtain groups that have similar variances across the $P$ attributes.
+As a caveat, however, it is crucial to note that maximizing $\mathit{SSE_{Var}}$ only reliably yields similar group variances if the group centroids $\overline{\mathbf{x}}^{(k)}$ ($k = 1, \dots, K$) are close to the overall data centroid $\overline{\mathbf{x}}$.
+This is because the within-group variance is calculated on the basis of the group centroid $\overline{\mathbf{x}}^{(k)}$, but the new squared distance variables $y_{ij}$ are calculated using the overall centroid $\overline{\mathbf{x}}$ as reference point.
+To ensure that $\mathit{SSE_{Var}}$ is a valid criterion for between-group similarity in variance, we must therefore simultaneously ensure that group centroids of the original variables are close to the overall data centroid (and therefore to each other).
+Ensuring that the group means are close to the overall centroid is accomplished when maximizing the classical $\mathit{SSE}$, i.e., via standard k-means anticlustering [@spath1986].
 Combining the maximization of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ at the same time leads to the k-plus approach for anticlustering, which is detailed in the next section.
@@
 The most simple approach for an optimization that considers multiple criteria at the same time is the weighted sum approach [@naidu2014; @marler2010].
-That is, all criteria are computed independently from each other and are then combined into a single criterion through a weighted sum; a suitable optimization algorithm is then used to maximize the combined criterion.
+That is, all criteria are computed independently from each other and are then combined into a single criterion through a weighted sum.
 By combining $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ in a single criterion, we define the objective function for k-plus anticlustering:
@@
 $$
+
 Maximizing $\mathit{SSE_{kplus}}$ will lead to similar means and variances between $K$ groups, while the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
@@
 Some simplifications in the computation of $\mathit{SSE_{kplusGeneralized}}$ can be arranged when ignoring the moment weights $\mathbf{w}$, e.g.
-by setting $w_t = 1$ ($t = 1, \dots, T$).
-This simplification leads to the *unweighted sum approach* for k-plus anticlustering.
+by setting $w_t = 1$ ($t = 1, \dots, T$), leading to the *unweighted sum approach* for k-plus anticlustering.
 In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$, representing distribution moments can be dropped; instead, the tensor is flattened into a two-dimensional matrix, where the variables representing the additional distribution moments are appended to the original data matrix $\mathbf{X}_{N \times P}$, resulting in the augmented matrix $\mathbf{X'}_{N \times \mathit{TP}}$:
@@
 A first small scale simulation compared different approaches for weight selection for the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$.
-In a second large scale simulation study, the k-plus objective was compared to other popular anticlustering objectives, i.e., the k-means and diversity objectives.
+In a second large scale simulation study, the k-plus objective was compared to other popular anticlustering objectives, specifically, the diversity and the k-means objectives.
 
@@
 Note that the local maximum search can be performed on the basis of any arbitrary initial grouping.
-Specifically, the initial grouping can also be chosen in such a way that the groups are different sized; for example, if the data set cannot evenly be split, it is possible to ensure that the group sizes differ by 1 at most.
-However, the data sets in the simulation studies were always partitioned into equal-sized groups.
+In particular, the initial grouping directly determines the resulting group sizes, because the following exchange procedure never changes the sizes of the groups.
+Hence, while the data sets in the simulation studies were always partitioned into equal-sized groups, it is possible to specify different sized groups through the initial assignment, or if the data set cannot evenly be split, it is possible to ensure that the group sizes differ by 1 at most.
 
@@
 While an exhaustive evaluation of the influence of weights is beyond the scope of the current paper, a small scale simulation study was conducted to investigate the influence of weights for a selected number of moments.
-The target was the optimization of $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, i.e.
+The results of this initial test determined the benchmark implementation of k-plus anticlustering in the follow-up large scale comparison in Simulation 2.
+
+In Simulation 1, the target was the optimization of $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, i.e.
 minimizing differences with regard to group means, variances, skewness and kurtosis.
+Six weighting schemes were compared.
 Two unweighted approaches were implemented by defining $\mathbf{w} = (1, 1, 1, 1)$.
 The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented k-plus variables---of the input data before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
-In the second case, the unstandardized data matrix was used.
+In the second case, the unstandardised data matrix was used.
 Four weighted approaches were implemented by giving a tenfold increased weight to one of the four moments, respectively.
-That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favoring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favoring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, respectively.
+That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favouring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favouring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, respectively.
 The weighted approaches did not use a standardization of the input data.
@@
 Table 1 illustrates how well each of the six weighting schemes fulfilled each of the four constituting criteria, averaged across the `r length(results)` simulation runs.
-For the weighted approaches, the criterion that received the highest weight during the optimization was fulfilled to the highest degree, as should be expected.
+For the weighted approaches, the criterion that had received the highest weight during the optimization was fulfilled to the highest degree, as should be expected.
 Interestingly, the unweighted standardization approach had the best average performance across the four criteria.
@@
 Among the competing weighting schemes, the unweighted standardization approach therefore yielded the most promising results.
-It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, even slightly outperforming the weighting scheme that explicitly favored $\mathit{SSE}$.
-Hence, the unweighted standardization approach seemed to implicitly give high priority to minimizing differences in means.
+It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, even slightly outperforming the weighting scheme that explicitly favoured $\mathit{SSE}$.
+Hence, the unweighted standardization approach implicitly gave high priority to minimizing differences in means.
 This is a pleasant result because, as discussed earlier, for the higher order moment $\mathit{SSE}$s to be valid, the group means have to be close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ has high priority.
@@
 "Generalized" k-plus anticlustering was implemented by maximizing $\mathit{SSE_{kplusGeneralized}}$ using $T = 4$, thus minimizing differences with regard to group means, variances, skewness and kurtosis.
-The unweighted standardization approach that performed best in Simulation 1 was used as the benchmark implementation for k-plus anticlustering that was compared to the competing objectives.
+The unweighted standardization approach that performed best in Simulation 1 was employed.
 Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.
-Subsequently, the original k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
+Subsequently, the standard k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
 The k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means, (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups, and also (iii) a random allocation of elements to groups.
@@
 
-As a secondary purpose of Simulation 2, I investigated the degree to which partitions identified by the different anticlustering methods showed similarity.
-Specifically, I investigated whether different anticlustering solutions showed any overlap regarding the placement of individual elements in groups.
-In cluster analysis, high quality partitions according to different objectives tend to overlap to a large degree, but it is an open question whether this result also holds for different anticlustering partitions.
+As a secondary purpose of Simulation 2, I investigated the similarity of partitions obtained when optimizing the different anticlustering objectives.
+In cluster analysis, partitions identified through different approaches tend to overlap considerably, but it is an open question whether this result also holds for different anticlustering partitions.
 To tackle this issue, I computed the adjusted Rand index (ARI) for each pair of partitions that were returned by the competing methods [@rand1971; @hubert1985].
 The ARI is an index of the similarity between two partitions, where an index of 0 indicates no similarity and a value of 1 indicates the highest possible similarity (i.e., when two partitions are the same).
-The pairwise ARIs were computed for each data set and for each $K$.
+The pairwise ARIs were computed for each simulation run, i.e., for data set and for each $K$.
 
@@
 Increasing $N$ then facilitated to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
-
 Generally, it should be noted that diversity anticlustering is a good all-rounder method that tends to address all distribution characteristics.
@@
 Table 3 illustrates the degree of overlap between anticlustering partitions that were identified in Simulation 2.
-Overall, it is shown that there was hardly any similarity between anticlustering partitions according to the different criteria.
+Overall, it is shown that there was hardly any similarity between anticlustering partitions that were obtained via optimization of the different criteria.
 The average ARI was almost 0 in all cases---indicating no similarity of partitions whatsoever.
 Moreover, the ARI between two anticlustering partitions hardly differed from the ARI between an anticlustering partition and a random partition.
-This constitutes an interesting result, illustrating that high between-group similarity can be achieved via very different partitionings.
+This constitutes an interesting result, illustrating that high between-group similarity can be achieved via very different partitions.
 
@@
 Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern.
-Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, as was also shown in Simulation 2.
-In contrast, different clustering solutions tend to look similar because a good clustering is always characterized by assigning similar elements into the same cluster and dissimilar elements into different clusters.
+Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different.
+In contrast, different clustering solutions tend to look similar because good clusterings are characterized by assigning similar elements into the same cluster and dissimilar elements into different clusters.
 A good anticlustering solution however does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This result was observed in Simulation 2 and can also be clarified using Figure 1.
@@
 Thus, as compared to pure cluster analysis, fair clustering biases the identification of groups with regard to the protected attributes, striving for algorithmic fairness.
-A relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data [@delBarrio2022], while anticlustering divides data sets into artificial subgroups in such a way that between-group similarity / within-group heterogeneity is maximized.
+A relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data, while anticlustering divides data sets into artificial subgroups in such a way that between-group similarity / within-group heterogeneity is maximized.
 However, these methods are logically similar because anticlustering corresponds to fair clustering when only considering the protected attributes.

### 2023-05-05 paper.Rmd
another round of editing

@@
 
-Maximizing $\mathit{SSE_{kplus}}$ will lead to similar means and variances between $K$ groups, while the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
-
 ```{r motivatingExample}
@@
 
-Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict (Ferligoj & Batagelj, 1992).
-As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria are similar between groups, i.e., whether these objectives are aligned or in conflict.
+By maximizing the bicriterion objective $\mathit{SSE_{kplus}}$, we hope to obtain a partitioning where means as well as variances are similar between groups.
+The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively, during the optimization.
+Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict rather than in alignment (Ferligoj & Batagelj, 1992).
+As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown if the two constituting criteria are aligned.
+That is, it has yet to be demonstrated that maximizing $\mathit{SSE_{kplus}}$ can yield a partition where means as well as variances are similar between groups.
 A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible.
@@
 
-Some simplifications in the computation of $\mathit{SSE_{kplusGeneralized}}$ can be arranged when ignoring the moment weights $\mathbf{w}$, e.g.
-by setting $w_t = 1$ ($t = 1, \dots, T$), leading to the *unweighted sum approach* for k-plus anticlustering.
+Some simplifications in the computation of $\mathit{SSE_{kplusGeneralized}}$ can be arranged when ignoring the moment weights $\mathbf{w}$ by setting $w_t = 1$ ($t = 1, \dots, T$); this simplification leads to the *unweighted sum approach* for k-plus anticlustering.
 In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$, representing distribution moments can be dropped; instead, the tensor is flattened into a two-dimensional matrix, where the variables representing the additional distribution moments are appended to the original data matrix $\mathbf{X}_{N \times P}$, resulting in the augmented matrix $\mathbf{X'}_{N \times \mathit{TP}}$:
@@
 
-Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ has the same form as the standard k-means criterion $\mathit{SSE}$, but is based on different data.
+Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ has the same form as the standard k-means criterion $\mathit{SSE}$.
 Here, computing the k-plus criterion is done by first creating the additional k-plus variables, appending them as new columns to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data.
@@
 A first small scale simulation compared different approaches for weight selection for the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$.
-In a second large scale simulation study, the k-plus objective was compared to other popular anticlustering objectives, specifically, the diversity and the k-means objectives.
+In a second large scale simulation study, the k-plus objective was compared to other anticlustering objectives striving for between-group similarity, specifically, the diversity and the k-means objectives.
 
@@
 Note that the local maximum search can be performed on the basis of any arbitrary initial grouping.
-In particular, the initial grouping directly determines the resulting group sizes, because the following exchange procedure never changes the sizes of the groups.
-Hence, while the data sets in the simulation studies were always partitioned into equal-sized groups, it is possible to specify different sized groups through the initial assignment, or if the data set cannot evenly be split, it is possible to ensure that the group sizes differ by 1 at most.
+In particular, the initial grouping directly determines the size of the anticlusters, because the exchange procedure following the set-up never changes the sizes of the groups.
+While the data sets in the simulation studies were always partitioned into equal-sized groups, it is also possible to request different-sized groups or, if the data set cannot be split evenly, it is possible to ensure that the group sizes differ by 1 at most.
 
@@
 The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented k-plus variables---of the input data before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
-In the second case, the unstandardised data matrix was used.
+In the second case, the unstandardised augmented data matrix was used.
 Four weighted approaches were implemented by giving a tenfold increased weight to one of the four moments, respectively.
@@
 Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but did not outperform a random assignment with regard to all other distribution characteristics.
-Replicating @papenberg2020, k-means' performance with regard to similarity in standard deviations was even below that of a random assignment.
 K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); both k-plus approaches strongly outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
@@
 
+# remove diagonal
+diag(ARIs) <- ""
+ARIs <- ARIs[, -ncol(ARIs)]
+
 apa_table(
@@
   escape = FALSE,
-  align = c("l")
+  align = c("r")
 )
@@
 Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.
-Specifically, it should be investigated if the bicriterion approach by Brusco et al.
+Specifically, it is of high interest to investigate if the bicriterion approach by Brusco et al.
 can be transformed into a multicriterion algorithm for the k-plus objective.
@@
 Future research should investigate whether this advantage also holds for other data distributions than normal.
-For example, it may be hypothesized that a generalized k-plus criterion is even more useful when applied to data that is not symmetric or contains outliers, because it is able to specifically incorporate the skewness and kurtosis as criteria in the anticlustering process.
+For example, it may be hypothesized that the generalized k-plus criterion is even more useful when applied to data that is not symmetric or contains outliers, because it is able to explicitly incorporate the skewness and kurtosis as criteria in the anticlustering process.
 
 As an additional outlook, it should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariances, a possibility that was not investigated in the current paper.
-In this case, we have to compute an additional variable for each of pair of variables whose covariance should be similar between groups.
-In future research, the unweighted standardization approach presented here may serve as a starting point for investigating covariance similarity in anticlustering applications.
+To take into account covariances, we have to compute an additional variable for each of pair of variables whose covariance should be similar between groups.
+The unweighted standardization approach presented here may serve as a starting point for investigating covariance similarity in anticlustering applications in the future.
 

### 2023-05-06 paper.Rmd
another small round of editing

@@
 
-Thus, to obtain an adaptation of the $\mathit{SSE}$ whose maximization can result in minimal differences in variances between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
+Thus, to obtain an adaptation of the $\mathit{SSE}$ whose maximization can minimize differences in variances, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
@@
 As a caveat, however, it is crucial to note that maximizing $\mathit{SSE_{Var}}$ only reliably yields similar group variances if the group centroids $\overline{\mathbf{x}}^{(k)}$ ($k = 1, \dots, K$) are close to the overall data centroid $\overline{\mathbf{x}}$.
-This is because the within-group variance is calculated on the basis of the group centroid $\overline{\mathbf{x}}^{(k)}$, but the new squared distance variables $y_{ij}$ are calculated using the overall centroid $\overline{\mathbf{x}}$ as reference point.
-To ensure that $\mathit{SSE_{Var}}$ is a valid criterion for between-group similarity in variance, we must therefore simultaneously ensure that group centroids of the original variables are close to the overall data centroid (and therefore to each other).
-Ensuring that the group means are close to the overall centroid is accomplished when maximizing the classical $\mathit{SSE}$, i.e., via standard k-means anticlustering [@spath1986].
+This is because the within-group variance is calculated on the basis of the group centroid $\overline{\mathbf{x}}^{(k)}$, but the new variables $y_{ij}$ are calculated using the overall centroid $\overline{\mathbf{x}}$ as reference point.
+To ensure that $\mathit{SSE_{Var}}$ is a valid criterion, we must therefore ensure that group centroids of the original variables are close to the overall data centroid (and therefore to each other).
+Ensuring that the group means are close to the overall mean is accomplished by maximizing the classical $\mathit{SSE}$, i.e., via standard k-means anticlustering [@spath1986].
 Combining the maximization of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ at the same time leads to the k-plus approach for anticlustering, which is detailed in the next section.
@@
 
-By maximizing the bicriterion objective $\mathit{SSE_{kplus}}$, we hope to obtain a partitioning where means as well as variances are similar between groups.
-The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively, during the optimization.
-Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict rather than in alignment (Ferligoj & Batagelj, 1992).
+The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
+By maximizing $\mathit{SSE_{kplus}}$ we seek to obtain a partitioning where means as well as variances are similar between groups.
+However, research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict rather than in alignment (Ferligoj & Batagelj, 1992).
 As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown if the two constituting criteria are aligned.
@@
 Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
+First, it is shown that the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ seem to be rather well aligned: There is a high concentration of partitions where both criteria have high values.
+At the same time, there are many partitions that fulfil only one of the criteria very well, but fail to also fulfil the other criterion.
 The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$.
@@
 Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances.
-The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ = $w_2$ = 1 in the computation of $\mathit{SSE_{kplus}}$), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
+The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
 Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree.
@@
 A first small scale simulation compared different approaches for weight selection for the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$.
-In a second large scale simulation study, the k-plus objective was compared to other anticlustering objectives striving for between-group similarity, specifically, the diversity and the k-means objectives.
+In general, the weighting scheme for k-plus anticlustering can become highly complex as it might encompass an arbitrary number of distribution moments, and hence, an arbitrary number of weights.
+While an exhaustive evaluation of the influence of weights was beyond the scope of the current paper, Simulation 1 investigated the relative performance of a selected number weighting schemes.
+The results of this initial test determined the benchmark implementation of k-plus anticlustering that was used in Simulation 2.
+Simulation 2 was a large scale comparison of the k-plus objective and other anticlustering objectives striving for between-group similarity: the diversity and the k-means objective.
 
@@
 In particular, the initial grouping directly determines the size of the anticlusters, because the exchange procedure following the set-up never changes the sizes of the groups.
-While the data sets in the simulation studies were always partitioned into equal-sized groups, it is also possible to request different-sized groups or, if the data set cannot be split evenly, it is possible to ensure that the group sizes differ by 1 at most.
+While the data sets in the simulation studies were always partitioned into equal-sized groups, it would also be possible to ensure that the group sizes differ by 1 at most if an even split is not possible; even different-sized groups can be requested if required by the application.
 
@@
 ```
-
-In general, the weighting scheme for k-plus anticlustering can become highly complex as it might encompass an arbitrary number of distribution moments, and hence, an arbitrary number of weights.
-While an exhaustive evaluation of the influence of weights is beyond the scope of the current paper, a small scale simulation study was conducted to investigate the influence of weights for a selected number of moments.
-The results of this initial test determined the benchmark implementation of k-plus anticlustering in the follow-up large scale comparison in Simulation 2.
 

### 2023-05-07 paper.Rmd
typos

@@
 by computing all pairwise Euclidean distances, and then optimize a distance-based anticlustering criterion to obtain a partitioning.
-For likert scale data or binary attributes, the Manhattan distance may be preferred over the Euclidean distance; for mixed-type attributes, it is even possible to employ a combined distance metric [@rokach2005].
+For Likert scale data or binary attributes, the Manhattan distance may be preferred over the Euclidean distance; for mixed-type attributes, it is even possible to employ a combined distance metric [@rokach2005].
 
@@
 
-The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of equalizing means and variances, respectively.
+The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of means and variances in the anticlustering process.
 By maximizing $\mathit{SSE_{kplus}}$ we seek to obtain a partitioning where means as well as variances are similar between groups.
-However, research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict rather than in alignment (Ferligoj & Batagelj, 1992).
+However, research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict rather than in alignment (@ferligoj1992direct).
 As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown if the two constituting criteria are aligned.
@@
 I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
-Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
+Figure 1 shows the scatter plot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
 First, it is shown that the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ seem to be rather well aligned: There is a high concentration of partitions where both criteria have high values.
@@
 After the four anticlustering methods and the random allocation procedure had been applied to the 10,000 data sets for $K = 2$, $K = 3$, $K = 4$, I investigated how well the competing methods were able to create similar groups.
-To this end, I analyzed the discrepancy in means, standard deviations, skewness and kurtosis between groups---for each data set for each $K$.
+To this end, I analysed the discrepancy in means, standard deviations, skewness and kurtosis between groups---for each data set for each $K$.
 To quantify the discrepancy in group means, the following computation was used: for each of the (2, 3, 4 or 5) features, all (2, 3 or 4) group means were computed.
@@
 This is an important observation because it shows that k-plus anticlustering is capable of addressing multiple objectives at the same time---especially when $N$ increases.
-However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria made it more difficult to fulfill each single criterion: generalized k-plus performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus that only considered the mean and variances.
+However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria made it more difficult to fulfil each single criterion: generalized k-plus performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus that only considered the mean and variances.
 However, as intended, generalized k-plus was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$ and at the same time, this k-plus objective maintained a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
@@
 Kurdi et al.
-assembled 900 open-access color images and collected ratings on two affective dimensions: arousal and valence.
+assembled 900 open-access colour images and collected ratings on two affective dimensions: arousal and valence.
 @brielmann2019intense measured how the same 900 images were rated with regard to beauty.
@@
 Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be considered as the default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be challenging in small data sets.
-In practice, users can try out different objectives and see which specification fulfills their needs.
+In practice, users can try out different objectives and see which specification fulfils their needs.
 A reasonable approach is to start "greedy", e.g.
@@
 First, the applications and the simulations showed that the unweighted standardization approach yielded satisfying results that probably suffice in most applications.
-Moreover, it is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria to a satisfactory degree (as suggested by Figure 1).
-Still, more sophisticated approaches may yield theoretically improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives.
+Moreover, it is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfil multiple criteria to a satisfactory degree (as suggested by Figure 1).
+Still, more sophisticated approaches may yield theoretically improved results, as well as the opportunity to deliberately select among partitions that may be preferred according to different k-plus objectives.
 @brusco2019 gave an example of how direct algorithms can be employed to optimize multiple criteria in anticlustering applications aiming for high within-group heterogeneity.

### 2023-05-07 paper.Rmd
next round of editing

@@
 
-Oftentimes, partitioning applications in psychology are focused on between-group similarity rather than within-group heterogeneity, even though both goals oftentimes coincide.
+Oftentimes, partitioning applications in psychology are focused on between-group similarity rather than within-group heterogeneity, even though both goals usually coincide.
 High between-group similarity is for example desirable when designing experimental conditions using different stimulus sets that should be as similar as possible with regard to response-relevant attributes [@lahl2006].
@@
 
-Thus, to obtain an adaptation of the $\mathit{SSE}$ whose maximization can minimize differences in variances, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
+Thus, to obtain an adaptation of the $\mathit{SSE}$ whose maximization can minimize differences in variances between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
 We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
@@
 This is because the within-group variance is calculated on the basis of the group centroid $\overline{\mathbf{x}}^{(k)}$, but the new variables $y_{ij}$ are calculated using the overall centroid $\overline{\mathbf{x}}$ as reference point.
-To ensure that $\mathit{SSE_{Var}}$ is a valid criterion, we must therefore ensure that group centroids of the original variables are close to the overall data centroid (and therefore to each other).
+To ensure that $\mathit{SSE_{Var}}$ is a valid criterion, we must therefore ensure that the group centroids of the original variables are close to the overall data centroid (and therefore to each other).
 Ensuring that the group means are close to the overall mean is accomplished by maximizing the classical $\mathit{SSE}$, i.e., via standard k-means anticlustering [@spath1986].
@@
 
-The weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of means and variances in the anticlustering process.
-By maximizing $\mathit{SSE_{kplus}}$ we seek to obtain a partitioning where means as well as variances are similar between groups.
-However, research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict rather than in alignment (@ferligoj1992direct).
+By maximizing $\mathit{SSE_{kplus}}$ we seek to obtain a partitioning where means as well as variances are similar between groups, while the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of means and variances.
+However, research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria, if the underlying objectives are in conflict rather than in alignment [@ferligoj1992direct].
 As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown if the two constituting criteria are aligned.
@@
 Figure 1 shows the scatter plot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
-First, it is shown that the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ seem to be rather well aligned: There is a high concentration of partitions where both criteria have high values.
+The criteria seem to be rather well aligned: There is a high concentration of partitions where both the $\mathit{SSE}$ and the $\mathit{SSE_{Var}}$ have high values.
 At the same time, there are many partitions that fulfil only one of the criteria very well, but fail to also fulfil the other criterion.
@@
 The partition represented by the square maximizes $\\mathit{SSE}$ and thus minimizes the difference in means between groups; the partition represented by the triangular maximizes $\\mathit{SSE_{Var}}$.
-The partition represented by the diamond maximizes $\\mathit{SSE_{kplus}}$ for $w_1 = w_2 = 1$, and satisfies both constituting criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}
+The partition represented by the diamond maximizes $\\mathit{SSE_{kplus}}$ using $w_1 = w_2 = 1$.", fig.height = 5, fig.width = 5}
 
@@
 Thus, by changing $t$---the power of the deviation between mean and data points---we can use the k-plus logic to equalize any desired sample moment between groups.
-That is, we compute new variables for each original variable and for each distribution moment that should be considered in the anticlustering process.
+That is, to extend the k-plus criterion to higher order moments, we compute a new variable for each original variable and for each distribution moment that should be considered in the anticlustering process.
 Analogously to the computation of $\mathit{SSE_{Var}}$, we could then define an $\mathit{SSE_{Skew}}$ using $t = 3$ or an $\mathit{SSE_{Kurtosis}}$ using $t = 4$.
@@
 
-This computation results in a three dimensional tensor $\mathbf{\hat{X}}_{T \times N \times P}$ = $\{\hat{x}_{tij}\}_{T \times N \times P}$ where---in comparison to the original data matrix $\mathbf{X}_{N \times P}$---the additional dimension is given by the $T$ distribution moments that are considered in the anticlustering process.
-When defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as first entry in the "moment dimension" in $\mathbf{\hat{X}}_{T \times N \times P}$.
+This computation results in a three dimensional tensor $\mathbf{\hat{X}}_{T \times N \times P}$ = $\{\hat{x}_{tij}\}_{T \times N \times P}$ where---in comparison to the original data matrix $\mathbf{X}_{N \times P}$---the additional dimension is given by the additional distribution moments that are considered in the anticlustering process.
+When defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as first entry of the "moment dimension" in $\mathbf{\hat{X}}_{T \times N \times P}$.
 It is used to equalize the mean---the first moment---between groups.
-
 Using $\mathbf{\hat{X}}_{T \times N \times P}$ and a vector of weights $\mathbf{w} = (w_1, \dots, w_T$), the generalized multicriterion for k-plus anticlustering $\mathit{SSE_{kplusGeneralized}}$ can be formalized:
@@
 
-Some simplifications in the computation of $\mathit{SSE_{kplusGeneralized}}$ can be arranged when ignoring the moment weights $\mathbf{w}$ by setting $w_t = 1$ ($t = 1, \dots, T$); this simplification leads to the *unweighted sum approach* for k-plus anticlustering.
-In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$, representing distribution moments can be dropped; instead, the tensor is flattened into a two-dimensional matrix, where the variables representing the additional distribution moments are appended to the original data matrix $\mathbf{X}_{N \times P}$, resulting in the augmented matrix $\mathbf{X'}_{N \times \mathit{TP}}$:
+Some simplifications in the computation of $\mathit{SSE_{kplusGeneralized}}$ can be arranged when ignoring the moment weights $\mathbf{w}$ by setting $w_t = 1$ ($t = 1, \dots, T$).
+This simplification leads to the *unweighted sum approach* for k-plus anticlustering.
+In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$ can be dropped.
+Instead, the tensor is flattened into a two-dimensional matrix by appending the newly computed k-plus variables $\hat{x}_{tij}$ ($t > 1$) to the original data matrix $\mathbf{X}_{N \times P}$.
+We obtain the augmented matrix $\mathbf{X'}_{N \times \mathit{TP}}$:
 
@@
 
-When defining the rows of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i \cdot} = (x'_{i1}, \ldots, x'_{i(TP)})$ ($i = 1, \ldots, N$), the criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as
+When defining the rows of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i \cdot} = (x'_{i1}, \ldots, x'_{i(TP)})$ ($i = 1, \ldots, N$), the criterion $\mathit{SSE_{kplusGeneralized}}$ can then be computed as
 
@@
 
-Thus, when setting the weights for all moments to 1, the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ has the same form as the standard k-means criterion $\mathit{SSE}$.
-Here, computing the k-plus criterion is done by first creating the additional k-plus variables, appending them as new columns to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data.
+Thus, when setting the weights $w_t = 1$ ($t = 1, \dots, T$), the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ can be expressed in the same form as the standard k-means criterion $\mathit{SSE}$.
+Thus, it is possible to define the k-plus criterion by first creating the additional k-plus variables, appending them as new columns to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data.
 Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
@@
 Two simulation studies were conducted to evaluate the new k-plus objective for anticlustering.
-A first small scale simulation compared different approaches for weight selection for the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$.
-In general, the weighting scheme for k-plus anticlustering can become highly complex as it might encompass an arbitrary number of distribution moments, and hence, an arbitrary number of weights.
-While an exhaustive evaluation of the influence of weights was beyond the scope of the current paper, Simulation 1 investigated the relative performance of a selected number weighting schemes.
-The results of this initial test determined the benchmark implementation of k-plus anticlustering that was used in Simulation 2.
+Simulation 1 compared a selected number weighting schemes for the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$.
+The results of this initial test determined the benchmark implementation of k-plus anticlustering that was used as contestant in the follow-up simulation.
 Simulation 2 was a large scale comparison of the k-plus objective and other anticlustering objectives striving for between-group similarity: the diversity and the k-means objective.
@@
 Note that the local maximum search can be performed on the basis of any arbitrary initial grouping.
-In particular, the initial grouping directly determines the size of the anticlusters, because the exchange procedure following the set-up never changes the sizes of the groups.
+In particular, the initial grouping determines the size of the anticlusters, because the exchange procedure following the set-up never changes the sizes of the groups.
 While the data sets in the simulation studies were always partitioned into equal-sized groups, it would also be possible to ensure that the group sizes differ by 1 at most if an even split is not possible; even different-sized groups can be requested if required by the application.
@@
 
-In Simulation 1, the target was the optimization of $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, i.e.
-minimizing differences with regard to group means, variances, skewness and kurtosis.
-Six weighting schemes were compared.
+Six weighting schemes were used to optimize $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, thus minimizing differences with regard to differences with regard to group means, variances, skewness and kurtosis.
 Two unweighted approaches were implemented by defining $\mathbf{w} = (1, 1, 1, 1)$.
-The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented k-plus variables---of the input data before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
-In the second case, the unstandardised augmented data matrix was used.
+The first unweighted approach implemented a $z$-standardization of all columns---the original as well as the augmented k-plus variables---of the input data before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
+The second unweighted approach used the unstandardised augmented data matrix.
 Four weighted approaches were implemented by giving a tenfold increased weight to one of the four moments, respectively.
@@
 
-To compare the different weighting schemes, `r length(results)` data sets from were generated from a normal distribution ($M = 0$, $\mathit{SD} = 1$).
-The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$.
-For each data set, $\mathit{SSE_{kplusGeneralized}}$ was optimized using the local maximum algorithm, once for each of the six competing weighting schemes.
+In the simulation, `r length(results)` data sets from were generated from a normal distribution ($M = 0$, $\mathit{SD} = 1$).
+Each data set was divided into groups via optimization of $\mathit{SSE_{kplusGeneralized}}$, once using each of the six weighting schemes.
+The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of groups $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$.
 
@@
 For the weighted approaches, the criterion that had received the highest weight during the optimization was fulfilled to the highest degree, as should be expected.
-Interestingly, the unweighted standardization approach had the best average performance across the four criteria.
+The unweighted standardization approach yielded the best average performance across the four criteria.
 While the degree of advantage by itself does not seem to be remarkable, this pattern was observed in an astonishing `r mean(unstd_best) * 100`% of all simulation runs, while the level of chance would have been `r 1/6 * 100`% (`r format_p(sig_test$p.value)`).
 Among the competing weighting schemes, the unweighted standardization approach therefore yielded the most promising results.
-It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, even slightly outperforming the weighting scheme that explicitly favoured $\mathit{SSE}$.
+It performed also remarkably strong with regard to the standard $\mathit{SSE}$, even slightly outperforming the weighting scheme that explicitly favoured the $\mathit{SSE}$.
 Hence, the unweighted standardization approach implicitly gave high priority to minimizing differences in means.
-This is a pleasant result because, as discussed earlier, for the higher order moment $\mathit{SSE}$s to be valid, the group means have to be close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ has high priority.
+This is a pleasant result because, as discussed earlier, for the higher order $\mathit{SSE}$s to be valid, the group means have to be close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ has high priority.
 
@@
 As a secondary purpose of Simulation 2, I investigated the similarity of partitions obtained when optimizing the different anticlustering objectives.
-In cluster analysis, partitions identified through different approaches tend to overlap considerably, but it is an open question whether this result also holds for different anticlustering partitions.
-To tackle this issue, I computed the adjusted Rand index (ARI) for each pair of partitions that were returned by the competing methods [@rand1971; @hubert1985].
+In cluster analysis, partitions identified through different approaches tend to overlap considerably, because similar elements tend to be placed in the same cluster while dissimilar elements tend to be placed in different clusters; this principle holds regardless of the specific clustering method.
+It is an open question whether the optimization of different anticlustering objectives leads to similar groupings of the elements in anticlusters.
+To address this question, I computed the adjusted Rand index (ARI) for each pair of partitions that were returned by the competing methods [@rand1971; @hubert1985].
 The ARI is an index of the similarity between two partitions, where an index of 0 indicates no similarity and a value of 1 indicates the highest possible similarity (i.e., when two partitions are the same).
@@
 
-Table 3 illustrates the degree of overlap between anticlustering partitions that were identified in Simulation 2.
-Overall, it is shown that there was hardly any similarity between anticlustering partitions that were obtained via optimization of the different criteria.
+To answer the question if different anticlustering partitions are similar or dissimilar, Table 3 illustrates the degree of overlap between anticlustering partitions that were identified in Simulation 2.
+Overall, there was hardly any similarity between anticlustering partitions that were obtained via optimization of the different criteria.
 The average ARI was almost 0 in all cases---indicating no similarity of partitions whatsoever.
-Moreover, the ARI between two anticlustering partitions hardly differed from the ARI between an anticlustering partition and a random partition.
+Even the pairwise ARI of the two k-plus criteria was fractional, although the criteria overlapped considerably in their optimization goal.
 This constitutes an interesting result, illustrating that high between-group similarity can be achieved via very different partitions.
@@
 
-This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020].
+This section demonstrates how readers can easily apply k-plus anticlustering using the R package `anticlust` [@papenberg2020].
 The package can be installed in the R environment using the `install.packages()` command.
@@
   kurtosis = FALSE,
-  moments = NULL,
+  T = NULL,
   method = "local-maximum",
@@
 The arguments corresponding to skewness and kurtosis have to be specifically "turned on", i.e., by setting them to `TRUE` in the function call of `kplus_anticlustering()`.
-If other higher order moments should be included as part of the optimization, the optional argument `moments` can be used to specify the desired moments as an integer vector.
+If other higher order moments should be included as part of the optimization, the optional argument `T` can be used to specify how many moments should be used.
 The argument `method = "local-maximum"` ensures that the k-plus criterion is optimized using the local maximum search method described in the previous section.
@@
 The k-plus criterion is an extension of the classical k-means criterion, which only reflects how similar groups are with regard to their means.
-Interestingly, when using the unweighted sum approach presented here, the k-plus extension does not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
+Interestingly, when using the unweighted sum approach presented here, the k-plus extension does not even change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
 
@@
 Using the k-plus extension, however, the k-means criterion has regained attractivity: When opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.
-Future research should investigate whether this advantage also holds for other data distributions than normal.
+Future research should investigate whether this advantage also holds for other data distributions than considered here.
 For example, it may be hypothesized that the generalized k-plus criterion is even more useful when applied to data that is not symmetric or contains outliers, because it is able to explicitly incorporate the skewness and kurtosis as criteria in the anticlustering process.
 
-As an additional outlook, it should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariances, a possibility that was not investigated in the current paper.
-To take into account covariances, we have to compute an additional variable for each of pair of variables whose covariance should be similar between groups.
-The unweighted standardization approach presented here may serve as a starting point for investigating covariance similarity in anticlustering applications in the future.
+As an additional outlook, it should be noted that because the covariance is defined as an expected value (i.e., an average), the k-plus logic can be extended to minimize differences with regard to covariances, a possibility that was not investigated in the current paper.
+Similar to addressing distribution moments, covariances could be incorporated by appending additional variables to a data set.
+In this case, it would be necessary to compute an additional variable for each of pair of variables, because the covariance is the product of the differences between two variables and their respective means.
 
@@
 In contrast, different clustering solutions tend to look similar because good clusterings are characterized by assigning similar elements into the same cluster and dissimilar elements into different clusters.
-A good anticlustering solution however does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This result was observed in Simulation 2 and can also be clarified using Figure 1.
+A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This result was observed in Simulation 2 and can also be clarified using Figure 1.
 Here, we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.
@@
 Thus, as compared to pure cluster analysis, fair clustering biases the identification of groups with regard to the protected attributes, striving for algorithmic fairness.
-A relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data, while anticlustering divides data sets into artificial subgroups in such a way that between-group similarity / within-group heterogeneity is maximized.
+A relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data, while anticlustering divides data sets into purely artificial subgroups.
 However, these methods are logically similar because anticlustering corresponds to fair clustering when only considering the protected attributes.

### 2023-05-08 paper.Rmd
more editing

@@
 To motivate the new criterion, we first consider a set of one-dimensional observations.
-The sample variance of a set of $N$ observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:
-
-$$
-\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^2
-$$
-
-When defining a new variable $y = (y_1, \ldots, y_N)$ as the squared deviation of each observation from the mean $\overline{\mathbf{x}}$, i.e.
-$y_i = (x_i - \overline{\mathbf{x}})^2$, the sample variance of $x$ can be reformulated as the mean of $y$:
+The sample variance of a set of $N$ observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{x}$:
+
+$$
+\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{x})^2
+$$
+
+When defining a new variable $y = (y_1, \ldots, y_N)$ as the squared deviation of each observation from the mean $\overline{x}$, i.e.
+$y_i = (x_i - \overline{x})^2$, the sample variance of $x$ can be reformulated as the mean of $y$:
 
@@
 The criteria seem to be rather well aligned: There is a high concentration of partitions where both the $\mathit{SSE}$ and the $\mathit{SSE_{Var}}$ have high values.
-At the same time, there are many partitions that fulfil only one of the criteria very well, but fail to also fulfil the other criterion.
+At the same time, there are many partitions that fulfil only one of the criteria, but fail to simultaneously fulfil the other criterion.
 The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$.
@@
 In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$ can be dropped.
-Instead, the tensor is flattened into a two-dimensional matrix by appending the newly computed k-plus variables $\hat{x}_{tij}$ ($t > 1$) to the original data matrix $\mathbf{X}_{N \times P}$.
+Instead, the tensor is flattened into a two-dimensional matrix by appending the newly computed k-plus variables $\hat{x}_{tij}$ ($t \ge 2$) to the original data matrix $\mathbf{X}_{N \times P}$.
 We obtain the augmented matrix $\mathbf{X'}_{N \times \mathit{TP}}$:
@@
 Thus, when setting the weights $w_t = 1$ ($t = 1, \dots, T$), the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ can be expressed in the same form as the standard k-means criterion $\mathit{SSE}$.
-Thus, it is possible to define the k-plus criterion by first creating the additional k-plus variables, appending them as new columns to the original data matrix, and then computing the $\mathit{SSE}$ on the augmented data.
+According to this principle, the k-plus criterion is computed by first creating the additional k-plus variables, appending them as new columns to the original data matrix, and then computing the standard $\mathit{SSE}$ on the augmented data.
 Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
@@
 Two simulation studies were conducted to evaluate the new k-plus objective for anticlustering.
-Simulation 1 compared a selected number weighting schemes for the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$.
+Simulation 1 compared a selected number of weighting schemes for the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$.
 The results of this initial test determined the benchmark implementation of k-plus anticlustering that was used as contestant in the follow-up simulation.
-Simulation 2 was a large scale comparison of the k-plus objective and other anticlustering objectives striving for between-group similarity: the diversity and the k-means objective.
+Simulation 2 was a large scale comparison of the k-plus objective and the most important traditional anticlustering objectives striving for between-group similarity: the diversity and the k-means objective.
 
@@
 Note that the local maximum search can be performed on the basis of any arbitrary initial grouping.
-In particular, the initial grouping determines the size of the anticlusters, because the exchange procedure following the set-up never changes the sizes of the groups.
+In particular, the initial grouping determines the sizes of the anticlusters, because the exchange procedure following the set-up never changes the sizes of the groups.
 While the data sets in the simulation studies were always partitioned into equal-sized groups, it would also be possible to ensure that the group sizes differ by 1 at most if an even split is not possible; even different-sized groups can be requested if required by the application.
@@
 note <- paste0("Cells indicate how well the maximization of $\\mathit{SSE_{kplusGeneralized}}$ fulfilled each of the four constituting criteria $\\mathit{SSE}$, $\\mathit{SSE_{Var}}$, $\\mathit{SSE_{Skew}}$, and $\\mathit{SSE_{Kurtosis}}$, depending on the weighting scheme.
-For each simulation run, the relative performance was computed as the $\\mathit{SSE}$ that was obtained by the given weighting scheme, divided by the best $\\mathit{SSE}$ that was obtained by any weighting scheme in this simulation run.
+For each simulation run, the relative performance was computed as the result that was obtained by the given weighting scheme, divided by the best result that was obtained by any weighting scheme in this simulation run.
 Cell values represent the average relative performance across all ", length(results), " simulation runs.")
@@
 Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.
-Subsequently, the standard k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
+Subsequently, the k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
 The k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means, (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups, and also (iii) a random allocation of elements to groups.
@@
 In cluster analysis, partitions identified through different approaches tend to overlap considerably, because similar elements tend to be placed in the same cluster while dissimilar elements tend to be placed in different clusters; this principle holds regardless of the specific clustering method.
-It is an open question whether the optimization of different anticlustering objectives leads to similar groupings of the elements in anticlusters.
+However, it is an open question whether the optimization of different anticlustering objectives leads to similar groupings of the elements in anticlusters.
 To address this question, I computed the adjusted Rand index (ARI) for each pair of partitions that were returned by the competing methods [@rand1971; @hubert1985].
@@
 [^dispersion]: An exception is given when using the dispersion objective, which ensures that the most similar elements are assigned to different clusters.
-However, anticlustering objectives striving for overall between-group similarity are mostly unaffected by the grouping of individual item pairs.
+However, anticlustering objectives striving for overall between-group similarity are largely unaffected by the grouping of individual item pairs.
 

### 2023-05-08 paper.Rmd
still editing for revision 1

@@
 by computing all pairwise Euclidean distances, and then optimize a distance-based anticlustering criterion to obtain a partitioning.
-For Likert scale data or binary attributes, the Manhattan distance may be preferred over the Euclidean distance; for mixed-type attributes, it is even possible to employ a combined distance metric [@rokach2005].
+For Likert scale data or binary attributes, the Manhattan distance may be preferred over the Euclidean distance; for mixed-type attributes, it is even possible to compute a combined distance metric [@rokach2005].
 
@@
 This is because the within-group variance is calculated on the basis of the group centroid $\overline{\mathbf{x}}^{(k)}$, but the new variables $y_{ij}$ are calculated using the overall centroid $\overline{\mathbf{x}}$ as reference point.
+Using $\overline{\mathbf{x}}$ as reference point is however necessary because the group means $\overline{\mathbf{x}}^{(k)}$ are determined by the composition of the groups $(C_1, \ldots, C_K)$, which is the end result of the optimization and not known in advance.
 To ensure that $\mathit{SSE_{Var}}$ is a valid criterion, we must therefore ensure that the group centroids of the original variables are close to the overall data centroid (and therefore to each other).
-Ensuring that the group means are close to the overall mean is accomplished by maximizing the classical $\mathit{SSE}$, i.e., via standard k-means anticlustering [@spath1986].
+Ensuring that the group means are close to the overall mean is accomplished by maximizing the classical $\mathit{SSE}$, i.e., via standard k-means anticlustering (Späth, 1986).
 Combining the maximization of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ at the same time leads to the k-plus approach for anticlustering, which is detailed in the next section.
@@
 
-By maximizing $\mathit{SSE_{kplus}}$ we seek to obtain a partitioning where means as well as variances are similar between groups, while the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of means and variances.
-However, research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria, if the underlying objectives are in conflict rather than in alignment [@ferligoj1992direct].
+By maximizing $\mathit{SSE_{kplus}}$, we seek to obtain a partitioning where means as well as variances are similar between groups, while the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of means and variances.
+However, research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict rather than in alignment [@ferligoj1992direct].
 As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown if the two constituting criteria are aligned.
@@
 
-To answer the question if different anticlustering partitions are similar or dissimilar, Table 3 illustrates the degree of overlap between anticlustering partitions that were identified in Simulation 2.
+To answer the question if different anticlustering partitions are similar or dissimilar, Table 3 illustrates the degree of similarity between anticlustering partitions that were identified in Simulation 2.
 Overall, there was hardly any similarity between anticlustering partitions that were obtained via optimization of the different criteria.
@@
 
-As an additional outlook, it should be noted that because the covariance is defined as an expected value (i.e., an average), the k-plus logic can be extended to minimize differences with regard to covariances, a possibility that was not investigated in the current paper.
+As an additional outlook, it should be noted that because the covariance is defined as an expected value (i.e., an average), the k-plus logic can be extended to minimize differences with regard to covariances.
 Similar to addressing distribution moments, covariances could be incorporated by appending additional variables to a data set.
 In this case, it would be necessary to compute an additional variable for each of pair of variables, because the covariance is the product of the differences between two variables and their respective means.
+Future research should investigate how covariances can properly be included with k-plus anticlustering.
 

### 2023-05-08 paper.Rmd
word count for revision
last minute changes before submission of revision

@@
     Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity.
-Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximizing instead of minimizing a clustering objective function.
-Introducing k-plus, this paper presents an extension of the classical k-means objective to maximize between-group similarity in anticlustering applications.
-The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, skewness, and higher order moments), whereas the k-means criterion only reflects the degree to which groups differ in their means.
-While constituting a new criterion for anticlustering, it is surprisingly shown that k-plus anticlustering can be implemented by optimizing the original k-means criterion after the input data has been augmented with additional variables.
+Anticlustering thereby reverses the logic of its better known twin—cluster analysis—and is usually approached by maximizing instead of minimizing a clustering objective function.
+This paper presents k-plus, an extension of the classical k-means objective to maximize between-group similarity in anticlustering applications.
+K-plus represents between-group similarity as discrepancy in distribution moments (means, variance, and higher order moments), whereas the k-means criterion only reflects groups differences with regard to means.
+While constituting a new criterion for anticlustering, it is shown that k-plus anticlustering can be implemented by optimizing the original k-means criterion after the input data has been augmented with additional variables.
 A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple objectives.
 In particular, optimizing between-group similarity with regard to variances usually does not compromise similarity with regard to means; the k-plus extension is therefore generally preferred over classical k-means anticlustering.
-Examples are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust).
+Examples are given on how k-plus anticlustering can be applied to real norming data using the open source R package anticlust, which is freely available via CRAN.
 
@@
 keywords          : "Anticlustering, k-means, k-plus, variance, skewness, kurtosis, stimulus selection"
-wordcount         : "4992"
+wordcount         : "6825"
 
@@
 ARIs <- ARIs[, -ncol(ARIs)]
+ARIs <- ARIs[-1, ]
 

### 2023-06-12 paper.Rmd
updated information on acceptance at BMSP

@@
     Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity.
-Anticlustering thereby reverses the logic of its better known twin—cluster analysis—and is usually approached by maximizing instead of minimizing a clustering objective function.
+Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximizing instead of minimizing a clustering objective function.
 This paper presents k-plus, an extension of the classical k-means objective to maximize between-group similarity in anticlustering applications.
@@
 documentclass     : "apa7"
+note              : "\\begin{flushleft} This is the preprint of a paper that is now in press at the British Journal of Mathematical and Statistical Psychology.
+The DOI will be made available as soon as it is exists.
+\\newline \\newline Papenberg, M.
+(in press).
+k-plus Anticlustering: An Improved k-means Criterion for Maximizing Between-Group Similarity.
+\\textit{British Journal of Mathematical and Statistical Psychology}.
+\\end{flushleft}"
+
 ---

### 2023-07-04 paper.Rmd
do not use mathbold for scalars

@@
 Note that there is a direct connection between the $\mathit{SSE}$ and the location of the cluster centroids [@spath1986].
-If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{\mathbf{x}}_1, \overline{\mathbf{x}}_2, \ldots, \overline{\mathbf{x}}_P$), where $\overline{\mathbf{x}}_j$ ($j = 1, \ldots, P$) is the overall mean value on the $j$'th attribute:
-
-$$
-\overline{\mathbf{x}}_j = \frac{1}{N} \sum\limits_{i=1}^{N} x_{ij}
+If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{x}_1, \overline{x}_2, \ldots, \overline{x}_P)$, where $\overline{x}_j$ ($j = 1, \ldots, P$) is the overall mean value on the $j$'th attribute:
+
+$$
+\overline{x}_j = \frac{1}{N} \sum\limits_{i=1}^{N} x_{ij}
 $$
@@
 Thus, to obtain an adaptation of the $\mathit{SSE}$ whose maximization can minimize differences in variances between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
-We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{\mathbf{x}}_j)^2$.
+We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$.
 On the basis of $\mathbf{Y}_{N \times P}$, we define the criterion $\mathit{SSE_{Var}}$:
@@
 $$
-\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}})^t.
+\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{x})^t.
 $$
@@
 
-We define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{\cdot j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean---the overall mean of the $j$'th attribute---is $\overline{\mathbf{x}}_j$.
+We define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{\cdot j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean---the overall mean of the $j$'th attribute---is $\overline{x}_j$.
 Let $T$ be the number of distribution moments that should be similar across $K$ groups.
@@
     x_{ij} & \text{if } t = 1\\
-    (x_{ij} - \overline{\mathbf{x}}_j)^t, & \text{if } t > 1.
+    (x_{ij} - \overline{x}_j)^t, & \text{if } t > 1.
 \end{cases}
@@
     x_{11} & \cdots & x_{1P} & 
-        (x_{11} - \overline{\mathbf{x}}_1)^2 & \cdots & (x_{1P} - \overline{\mathbf{x}}_P)^2 & 
+        (x_{11} - \overline{x}_1)^2 & \cdots & (x_{1P} - \overline{x}_P)^2 & 
         \cdots & 
-        (x_{11} - \overline{\mathbf{x}}_1)^T & \cdots & (x_{1P} - \overline{\mathbf{x}}_P)^T \\
+        (x_{11} - \overline{x}_1)^T & \cdots & (x_{1P} - \overline{x}_P)^T \\
     x_{21} & \cdots & x_{2P} & 
-        (x_{21} - \overline{\mathbf{x}}_1)^2 & \cdots & (x_{2P} - \overline{\mathbf{x}}_P)^2 & 
+        (x_{21} - \overline{x}_1)^2 & \cdots & (x_{2P} - \overline{x}_P)^2 & 
         \cdots & 
-        (x_{21} - \overline{\mathbf{x}}_1)^T & \cdots & (x_{2P} - \overline{\mathbf{x}}_P)^T \\
+        (x_{21} - \overline{x}_1)^T & \cdots & (x_{2P} - \overline{x}_P)^T \\
     \vdots & \ddots & \vdots & \vdots & \ddots & \vdots & \cdots & \vdots & \ddots & \vdots\\
     x_{N1} & \cdots & x_{NP} & 
-        (x_{N1} - \overline{\mathbf{x}}_1)^2 & \cdots & (x_{NP} - \overline{\mathbf{x}}_P)^2 & 
+        (x_{N1} - \overline{x}_1)^2 & \cdots & (x_{NP} - \overline{x}_P)^2 & 
         \cdots & 
-        (x_{N1} - \overline{\mathbf{x}}_1)^T & \cdots & (x_{NP} - \overline{\mathbf{x}}_P)^T \\
+        (x_{N1} - \overline{x}_1)^T & \cdots & (x_{NP} - \overline{x}_P)^T \\
   \end{array} } \right]

### 2023-07-04 paper.Rmd
added DOI

@@
   - \usepackage{amsmath}
+  - \usepackage{hyperref}
   - \floatstyle{plaintop}
@@
 documentclass     : "apa7"
-note              : "\\begin{flushleft} This is the preprint of a paper that is now in press at the British Journal of Mathematical and Statistical Psychology.
-The DOI will be made available as soon as it is exists.
-\\newline \\newline Papenberg, M.
-(in press).
-k-plus Anticlustering: An Improved k-means Criterion for Maximizing Between-Group Similarity.
-\\textit{British Journal of Mathematical and Statistical Psychology}.
-\\end{flushleft}"
-
+note              : "\\begin{flushleft} This paper is not the copy of record and may not exactly replicate the final, authoritative version of the article.
+The final article is available from the \\emph{British Journal of Mathematical and Statistical Psychology} via its DOI: \\href{https://doi.org/10.1111/bmsp.12315}{10.1111/bmsp.12315} \\end{flushleft}"
 ---

### 2024-01-31 Simulation_2/Analysis_BILS.R
sig test

@@ -32,6 +32,8 @@ mean(df$VANILLA_FOUND_OPTIMUM)
 tapply(df$VANILLA_FOUND_OPTIMUM, df$K, mean) |> round(2)
 tapply(df$VANILLA_FOUND_OPTIMUM, list(df$K, df$separate_dispersion_distances), mean) |> round(2)
 
+chisq.test(table(df$VANILLA_FOUND_OPTIMUM, df$separate_dispersion_distances))
+
 # not sure if this analysis makes sense:
 colMeans(df[df$VANILLA_FOUND_OPTIMUM & df$separate_dispersion_distances == 0, grepl("DIV", colnames(df))])
 

### 2024-01-31 paper.Rmd
starting new paper

@@
 ---
-title             : "k-plus Anticlustering: An Improved k-means Criterion for Maximizing Between-Group Similarity"
-shorttitle        : "k-plus anticlustering"
+title             : "An improved anticlustering method to maximize stimulus dispersion in experiments"
+shorttitle        : "Maximum stimulus dispersion"
 
@@
     email         : "martin.papenberg@hhu.de"
+  - name          : "Martin Breuer"
+    affiliation   : "1"
+    corresponding : no
+  - name          : "Max Diekhoff"
+    affiliation   : "1"
+    corresponding : no
+  - name          : "Gunnar W. Klau"
+    affiliation   : "1"
+    corresponding : no
 
@@
 
-  Martin Papenberg, Department of Experimental Psychology, Heinrich Heine University Düsseldorf.
+  Martin Papenberg, Department of Experimental Psychology, Heinrich 
+  Heine University Düsseldorf. Martin Breuer, Gunnar W. Klau, Max Diekhoff, 
+  Department of Computer Science, Heinrich Heine University Düsseldorf.
 
@@
 
-    Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity.
-Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximizing instead of minimizing a clustering objective function.
-This paper presents k-plus, an extension of the classical k-means objective to maximize between-group similarity in anticlustering applications.
-K-plus represents between-group similarity as discrepancy in distribution moments (means, variance, and higher order moments), whereas the k-means criterion only reflects groups differences with regard to means.
-While constituting a new criterion for anticlustering, it is shown that k-plus anticlustering can be implemented by optimizing the original k-means criterion after the input data has been augmented with additional variables.
-A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple objectives.
-In particular, optimizing between-group similarity with regard to variances usually does not compromise similarity with regard to means; the k-plus extension is therefore generally preferred over classical k-means anticlustering.
-Examples are given on how k-plus anticlustering can be applied to real norming data using the open source R package anticlust, which is freely available via CRAN.
+    When assigning stimuli to sets in experimental psychology, researchers strive for high similarity between stimulus sets and high pairwise dissimilarity of stimuli within sets.
+Because stimulus assignment is usually substandard when done manually by researchers, automated algorithmic approaches are preferred to accomplish this goal.
+Brusco et al.
+(2020) proposed a bicriterion anticlustering algorithm (BILS) that simultaneously maximizes the diversity (corresponding to between-set similarity) and the dispersion (corresponding to pairwise stimulus dissimilarity within-sets).
+We provide a first independent assessment of BILS and present BILS-E, which extends BILS in three ways.
+First, a globally optimal value of dispersion is always returned.
+Second, the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices, for example to include cannot-link constraints in the anticlustering process.
+Third, BILS-E allows that other anticlustering objectives of between-group similarity, such as k-means or k-plus, can be optimized instead of only the diversity.
+Using our optimal approach as benchmark in a simulation study, we show that the original BILS identifies optimal dispersion values for much larger data sets than previously analyzed, but also highlight conditions where it fails to do so.
+A practical example shows how BILS-E can be used for stimulus assignment in experimental psychology.
+The original BILS as well as our extension are both freely available via the open source R package anticlust.
 
@@
 
-keywords          : "Anticlustering, k-means, k-plus, variance, skewness, kurtosis, stimulus selection"
-wordcount         : "6825"
+keywords          : "Anticlustering, maximum dispersion, optimal algorithm, bicriterion optimization"
+wordcount         : "12345"
 
@@
   papaja::apa6_pdf
-appendix: "appendix.Rmd"
 
@@
 documentclass     : "apa7"
-note              : "\\begin{flushleft} This paper is not the copy of record and may not exactly replicate the final, authoritative version of the article.
-The final article is available from the \\emph{British Journal of Mathematical and Statistical Psychology} via its DOI: \\href{https://doi.org/10.1111/bmsp.12315}{10.1111/bmsp.12315} \\end{flushleft}"
+
 ---
@@
 library(knitr)
-library(dplyr)
-library(ggplot2)
-library(tidyr)
-library(DescTools)
+# library(dplyr)
+# library(ggplot2)
+# library(tidyr)
+# library(DescTools)
 # for formatting numbers in Rmarkdown, use my package prmisc (https://github.com/m-Py/prmisc)
@@
 
-# Problem formalization
-
-For the purpose of problem formalization, I adopt the notation @steinley2006 provided in his comprehensive synthesis of half a century of research on k-means.
-K-means anticlustering is used to partition $N$ elements each having $P$ attributes into $K$ groups $(C_1, \ldots, C_K)$.
-The partitioning is subject to the requirement that each element is assigned to exactly one group.
-In anticlustering applications, an additional constraint is usually imposed on the group sizes $n_k$ ($k = 1, \ldots K$), with the most common restriction being that all groups have equal size, i.e.
-$n_k = \frac{N}{K}$ [@papenberg2020].
-
-For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means anticlustering aims to maximize the error sum of squares ($\mathit{SSE}$), which is the sum of the squared Euclidean distances between each data point and its cluster centroid:
-
-$$
-\mathit{SSE} = 
-  \sum\limits_{j=1}^{P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (x_{ij} - \overline{x}_j ^{(k)})^2
-$$
-
-The *k*'th centroid $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ is a vector of length $P$, where each entry is the mean value of one of the $P$ attributes, computed across all observations belonging to the $k$'th cluster $C_k$ ($k = 1, \ldots, K$):
-
-$$
-\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} x_{ij}
-$$
-
-Note that there is a direct connection between the $\mathit{SSE}$ and the location of the cluster centroids [@spath1986].
-If the $\mathit{SSE}$ is minimal---which is the goal of k-means *clustering*---, the centroid values $\overline{\mathbf{x}}^{(k)} (k = 1, \ldots, K)$ are as far away as possible from the overall data centroid $\overline{\mathbf{x}} = (\overline{x}_1, \overline{x}_2, \ldots, \overline{x}_P)$, where $\overline{x}_j$ ($j = 1, \ldots, P$) is the overall mean value on the $j$'th attribute:
-
-$$
-\overline{x}_j = \frac{1}{N} \sum\limits_{i=1}^{N} x_{ij}
-$$
-
-If the $\mathit{SSE}$ is maximal---which is the goal of k-means *anticlustering*---, the cluster centroids are as close as possible to the overall centroid $\overline{\mathbf{x}}$ and therefore to each other [@spath1986].
-Thus, k-means anticlustering directly optimizes the similarity of the mean attribute values between clusters.
-
-## Minimizing discrepancy in variance between groups
-
-In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variances between groups.
-To motivate the new criterion, we first consider a set of one-dimensional observations.
-The sample variance of a set of $N$ observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{x}$:
-
-$$
-\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{x})^2
-$$
-
-When defining a new variable $y = (y_1, \ldots, y_N)$ as the squared deviation of each observation from the mean $\overline{x}$, i.e.
-$y_i = (x_i - \overline{x})^2$, the sample variance of $x$ can be reformulated as the mean of $y$:
-
-$$
-\mathit{Var(x)} = \frac{1}{N} \sum\limits_{i=1}^{N} y_i
-$$
-
-Because the sample variance is basically defined as an average---and k-means anticlustering minimizes differences with regard to averages---k-means anticlustering can be employed to equalize the variance across groups.
-This constitutes the central insight motivating the new k-plus criterion.
-
-Thus, to obtain an adaptation of the $\mathit{SSE}$ whose maximization can minimize differences in variances between groups, a new variable is computed for each original variable: The new variable represents the squared distance of each of the original values to the overall mean of the variable.
-We obtain a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$.
-On the basis of $\mathbf{Y}_{N \times P}$, we define the criterion $\mathit{SSE_{Var}}$:
-
-$$
-\mathit{SSE_{Var}} = 
-  \sum\limits_{j=1}^{P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (y_{ij} - \overline{y}_j ^{(k)})^2
-$$
-
-Note that $\mathit{SSE_{Var}}$ has the same form as the standard $\mathit{SSE}$.
-However, the input data reflects the squared difference between each data point and the attribute mean instead of the raw data points.
-Maximizing $\mathit{SSE_{Var}}$ can be used to obtain groups that have similar variances across the $P$ attributes.
-As a caveat, however, it is crucial to note that maximizing $\mathit{SSE_{Var}}$ only reliably yields similar group variances if the group centroids $\overline{\mathbf{x}}^{(k)}$ ($k = 1, \dots, K$) are close to the overall data centroid $\overline{\mathbf{x}}$.
-This is because the within-group variance is calculated on the basis of the group centroid $\overline{\mathbf{x}}^{(k)}$, but the new variables $y_{ij}$ are calculated using the overall centroid $\overline{\mathbf{x}}$ as reference point.
-Using $\overline{\mathbf{x}}$ as reference point is however necessary because the group means $\overline{\mathbf{x}}^{(k)}$ are determined by the composition of the groups $(C_1, \ldots, C_K)$, which is the end result of the optimization and not known in advance.
-To ensure that $\mathit{SSE_{Var}}$ is a valid criterion, we must therefore ensure that the group centroids of the original variables are close to the overall data centroid (and therefore to each other).
-Ensuring that the group means are close to the overall mean is accomplished by maximizing the classical $\mathit{SSE}$, i.e., via standard k-means anticlustering (Späth, 1986).
-Combining the maximization of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ at the same time leads to the k-plus approach for anticlustering, which is detailed in the next section.
-
-## A bicriterion model: K-plus anticlustering {#kplus}
-
-An exclusive maximization of $\mathit{SSE_{Var}}$ (i.e., equalizing variances) is not feasible---and most likely not interesting on its own---without also considering the standard $\mathit{SSE}$ (i.e., equalizing means).
-The most simple approach for an optimization that considers multiple criteria at the same time is the weighted sum approach [@naidu2014; @marler2010].
-That is, all criteria are computed independently from each other and are then combined into a single criterion through a weighted sum.
-By combining $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ in a single criterion, we define the objective function for k-plus anticlustering:
-
-$$
-\mathit{SSE_{kplus}} = 
-w_1 \sum\limits_{j=1}^{P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (x_{ij} - \overline{x}_j ^{(k)})^2 + 
-w_2 \sum\limits_{j=1}^{P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (y_{ij} - \overline{y}_j ^{(k)})^2
-$$
-
-```{r motivatingExample}
-
-# some functions for motivating example
-
-# Variance objective function with changed order of the arguments 
-# (can be called by sapply() etc.)
-var_objective <- function(clusters, features) {
-  anticlust::variance_objective(features, clusters)
-}
-
-# Compute features for k-plus criterion
-squared_from_mean <- function(data) {
-  apply(data, 2, function(x) (x - mean(x))^2)
-}
-
-# First rendering is slow because the data needs to be generated
-path <- "partition_objectives.csv"
-
-if (!file.exists(path)) {
-  set.seed(2207) # 2207
-  # Generate data and all partitions
-  N <- 14
-  K <- 2
-  partitions <- generate_partitions(N, K)
-  features <- rnorm(N, 15, 3)
-
-  # Compute objectives for each partition
-  # For each partition: Compute SSE objective, and SSE(VAR)
-  # 1. SSE
-  all_objectives_kMeans <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = features
-  )
-
-  squared_distances_from_mean <- squared_from_mean(as.matrix(features))
-
-  # 2. SSE(VAR)
-  all_objectives_kVar <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = squared_distances_from_mean
-  )
-
-  # 3. SSE(KPLUS)
-  all_objectives_kPlus <- sapply(
-    partitions,
-    FUN = var_objective,
-    features = cbind(features, squared_distances_from_mean)
-  )
-
-  df <- data.frame(
-    kmeans = all_objectives_kMeans,
-    kvar = all_objectives_kVar,
-    kplus = all_objectives_kPlus
-  )
-
-  write.table(
-    df, path, row.names = FALSE, sep = ","
-  )
-
-  # investigate the best 0.5% of k-means anticlustering partitions, for discussion
-  # best_0.1percent_anticlusterings <- df$kmeans /max(df$kmeans) > .999
-  # best_partitions <- partitions[best_0.1percent_anticlusterings]
-  # 
-  # neighbor_matrix <- anticlust:::selection_matrix_from_clusters
-  # neighbourhoods <- lapply(best_partitions, neighbor_matrix)
-  # neighbourhoods <- lapply(neighbourhoods, function(x) as.integer(x[lower.tri(x)]))
-  # pairs <- Reduce("+", neighbourhoods)
-  # length(pairs)
-  # length(neighbourhoods[[1]])
-  # choose(14, 2)
-  # min(pairs) # 40
-  # max(pairs) # 95
-  # length(best_partitions) # 146
-  # -> all items are grouped with one another, and no items are always grouped with one another
-
-} else {
-  df <- read.csv(path)
-}
-
-```
-
-By maximizing $\mathit{SSE_{kplus}}$, we seek to obtain a partitioning where means as well as variances are similar between groups, while the weights $w_1$ and $w_2$ can be chosen to adjust the relative importance of means and variances.
-However, research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict rather than in alignment [@ferligoj1992direct].
-As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown if the two constituting criteria are aligned.
-That is, it has yet to be demonstrated that maximizing $\mathit{SSE_{kplus}}$ can yield a partition where means as well as variances are similar between groups.
-A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible.
-I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups.
-Figure 1 shows the scatter plot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions.
-The criteria seem to be rather well aligned: There is a high concentration of partitions where both the $\mathit{SSE}$ and the $\mathit{SSE_{Var}}$ have high values.
-At the same time, there are many partitions that fulfil only one of the criteria, but fail to simultaneously fulfil the other criterion.
-The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$.
-Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (represented by the triangular symbol) is not well-suited to optimize $\mathit{SSE}$ at the same time.
-Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances.
-The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$, which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$.
-Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree.
-
-```{r, fig.cap = "The relationship between the objectives $\\mathit{SSE}$ and $\\mathit{SSE_{Var}}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups.
-The partition represented by the square maximizes $\\mathit{SSE}$ and thus minimizes the difference in means between groups; the partition represented by the triangular maximizes $\\mathit{SSE_{Var}}$.
-The partition represented by the diamond maximizes $\\mathit{SSE_{kplus}}$ using $w_1 = w_2 = 1$.", fig.height = 5, fig.width = 5}
-
-# Plot SSE vs SSE(VAR)
-plot(
-  df[, c("kmeans", "kvar")], 
-  pch = 4, col = "darkgrey", 
-  las = 1, cex = .5,
-  xlab = "",
-  ylab = ""
-)
-mtext(side = 2, text = expression(italic(SSE[Var])), line = 3)
-mtext(side = 1, text = expression(italic(SSE)), line = 3)
-
-# Illustrate best partition wrt difference in means (i.e., SSE)
-points(
-  df$kmeans[which.max(df$kmeans)],
-  df$kvar[which.max(df$kmeans)],
-  cex = 1.2, pch = 22, bg = "white", lwd = 1.7
-)
-
-# Illustrate best partition wrt difference in variance (i.e., SSE(VAR))
-points(
-  df$kmeans[which.max(df$kvar)],
-  df$kvar[which.max(df$kvar)],
-  cex = 1.2, pch = 24, bg = "white", lwd = 1.7
-)
-
-# Illustrate best partition wrt k-plus criterion
-points(
-  df$kmeans[which.max(df$kplus)],
-  df$kvar[which.max(df$kplus)],
-  cex = 1.2, pch = 23, bg = "white", lwd = 1.7
-)
-```
-
-## Generalizing k-plus anticlustering to higher order moments
-
-The k-plus logic to equalizing group variances as well as group means can be seamlessly generalized to higher order moments.
-The *t*'th sample moment of a variable $x = (x_1, \ldots, x_N)$ can be computed as
-
-$$
-\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{x})^t.
-$$
-
-where $t = 2$ gives the variance, $t = 3$ the skewness, and $t = 4$ the kurtosis [@joanes1998].
-Thus, by changing $t$---the power of the deviation between mean and data points---we can use the k-plus logic to equalize any desired sample moment between groups.
-That is, to extend the k-plus criterion to higher order moments, we compute a new variable for each original variable and for each distribution moment that should be considered in the anticlustering process.
-Analogously to the computation of $\mathit{SSE_{Var}}$, we could then define an $\mathit{SSE_{Skew}}$ using $t = 3$ or an $\mathit{SSE_{Kurtosis}}$ using $t = 4$.
-However, to formalize the computation of the k-plus criterion for an arbitrary number of distribution moments, some new notation has to be provided.
-
-We define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{\cdot j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean---the overall mean of the $j$'th attribute---is $\overline{x}_j$.
-Let $T$ be the number of distribution moments that should be similar across $K$ groups.
-For each moment $t$ ($t = 1, \ldots, T$) and for each variable $x_{\cdot j}$ ($j = 1, \ldots, P$), a new *k-plus variable* $\hat{x}_{t \cdot j} = (\hat{x}_{t1j}, \ldots, \hat{x}_{tNj})$ is defined where
-
-$$
-\hat{x}_{tij} = 
-\begin{cases}
-    x_{ij} & \text{if } t = 1\\
-    (x_{ij} - \overline{x}_j)^t, & \text{if } t > 1.
-\end{cases}
-$$
-
-This computation results in a three dimensional tensor $\mathbf{\hat{X}}_{T \times N \times P}$ = $\{\hat{x}_{tij}\}_{T \times N \times P}$ where---in comparison to the original data matrix $\mathbf{X}_{N \times P}$---the additional dimension is given by the additional distribution moments that are considered in the anticlustering process.
-When defining $\hat{x}_{tij}$, the condition $t = 1$ is separately included to keep the unchanged original data as first entry of the "moment dimension" in $\mathbf{\hat{X}}_{T \times N \times P}$.
-It is used to equalize the mean---the first moment---between groups.
-Using $\mathbf{\hat{X}}_{T \times N \times P}$ and a vector of weights $\mathbf{w} = (w_1, \dots, w_T$), the generalized multicriterion for k-plus anticlustering $\mathit{SSE_{kplusGeneralized}}$ can be formalized:
-
-$$
-\mathit{SSE_{kplusGeneralized}} =
-  \sum\limits_{t=1}^{T} 
-  w_t \left(
-  \sum\limits_{j=1}^{P} 
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (\hat{x}_{tij} - \overline{x}_{tj} ^{(k)})^2
-  \right)
-$$
-
-Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{\mathbf{X}}^{(k)}_{T \times P} = \{\overline{x}_{tj}^{(k)}\}_{T \times P}$ is a matrix containing the mean values for all k-plus variables in the $k$'th group, i.e.
-
-$$
-\overline{x}_{tj}^{(k)} = \frac{1}{n_k} \sum_{i \in C_k} \hat{x}_{tij}
-$$
-
-Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to minimize differences with regard to the first $T$ distribution moments between $K$ groups.
-
-## The unweighted sum approach for k-plus anticlustering
-
-Some simplifications in the computation of $\mathit{SSE_{kplusGeneralized}}$ can be arranged when ignoring the moment weights $\mathbf{w}$ by setting $w_t = 1$ ($t = 1, \dots, T$).
-This simplification leads to the *unweighted sum approach* for k-plus anticlustering.
-In this case, the additional dimension in $\mathbf{\hat{X}}_{T \times N \times P}$ can be dropped.
-Instead, the tensor is flattened into a two-dimensional matrix by appending the newly computed k-plus variables $\hat{x}_{tij}$ ($t \ge 2$) to the original data matrix $\mathbf{X}_{N \times P}$.
-We obtain the augmented matrix $\mathbf{X'}_{N \times \mathit{TP}}$:
-
-$$
-  \mathbf{X'}_{N \times \mathit{TP}} =
-  \left[ {\begin{array}{ccc|ccc|c|ccc}
-    x_{11} & \cdots & x_{1P} & 
-        (x_{11} - \overline{x}_1)^2 & \cdots & (x_{1P} - \overline{x}_P)^2 & 
-        \cdots & 
-        (x_{11} - \overline{x}_1)^T & \cdots & (x_{1P} - \overline{x}_P)^T \\
-    x_{21} & \cdots & x_{2P} & 
-        (x_{21} - \overline{x}_1)^2 & \cdots & (x_{2P} - \overline{x}_P)^2 & 
-        \cdots & 
-        (x_{21} - \overline{x}_1)^T & \cdots & (x_{2P} - \overline{x}_P)^T \\
-    \vdots & \ddots & \vdots & \vdots & \ddots & \vdots & \cdots & \vdots & \ddots & \vdots\\
-    x_{N1} & \cdots & x_{NP} & 
-        (x_{N1} - \overline{x}_1)^2 & \cdots & (x_{NP} - \overline{x}_P)^2 & 
-        \cdots & 
-        (x_{N1} - \overline{x}_1)^T & \cdots & (x_{NP} - \overline{x}_P)^T \\
-  \end{array} } \right]
-$$
-
-When defining the rows of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i \cdot} = (x'_{i1}, \ldots, x'_{i(TP)})$ ($i = 1, \ldots, N$), the criterion $\mathit{SSE_{kplusGeneralized}}$ can then be computed as
-
-$$
-\mathit{SSE_{kplusGeneralized}} =
-  \sum\limits_{j=1}^{\mathit{TP}}
-  \sum\limits_{k=1}^{K} 
-  \sum\limits_{i \in C_k}
-  (x'_{ij} - \overline{x'}_j ^{(k)})^2.
-$$
-
-Thus, when setting the weights $w_t = 1$ ($t = 1, \dots, T$), the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$ can be expressed in the same form as the standard k-means criterion $\mathit{SSE}$.
-According to this principle, the k-plus criterion is computed by first creating the additional k-plus variables, appending them as new columns to the original data matrix, and then computing the standard $\mathit{SSE}$ on the augmented data.
-Using the unweighted sum approach, k-plus anticlustering therefore leaves the original k-means criterion unchanged.
-
-Note that a problem may arise if the weights $\mathbf{w}$ are ignored when computing $\mathit{SSE_{kplusGeneralized}}$.
-Because the computation of the new k-plus variables involves exponentiation, their range of values is expected to be very different from the original variables.
-This imbalance can strongly and unpredictably influence the relative weight of the different moments when computing $\mathit{SSE_{kplusGeneralized}}$ (e.g., see Figure 1).
-To ensure a comparable weight of all criteria, all attributes---i.e., the original data as well as the new k-plus variables---can be standardized before computing $\mathit{SSE_{kplusGeneralized}}$, e.g.
-via $z$-standardization.
-This approach will henceforth be referred to as the unweighted standardization approach for k-plus anticlustering.
-
 # Evaluation
@@
 
-## Simulation 1: Selecting weights for k-plus anticlustering
-
-```{r getresultsSim1}
-dir <- "./Simulation_1/"
-files <- list.files(dir, pattern = ".Rdata")
-results <- list()
-for (file in files) {
-  results <- c(results, get(load(paste0(dir, file))))
-}
-
-```
-
-Six weighting schemes were used to optimize $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, thus minimizing differences with regard to differences with regard to group means, variances, skewness and kurtosis.
-Two unweighted approaches were implemented by defining $\mathbf{w} = (1, 1, 1, 1)$.
-The first unweighted approach implemented a $z$-standardization of all columns---the original as well as the augmented k-plus variables---of the input data before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized.
-The second unweighted approach used the unstandardised augmented data matrix.
-Four weighted approaches were implemented by giving a tenfold increased weight to one of the four moments, respectively.
-That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favouring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favouring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, respectively.
-The weighted approaches did not use a standardization of the input data.
-
-In the simulation, `r length(results)` data sets from were generated from a normal distribution ($M = 0$, $\mathit{SD} = 1$).
-Each data set was divided into groups via optimization of $\mathit{SSE_{kplusGeneralized}}$, once using each of the six weighting schemes.
-The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of groups $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$.
-
-```{r}
-final_tab_sim1 <- round(Reduce("+", results) / length(results) * 100, 2)
-rownames(final_tab_sim1) <- c(
-  "$\\mathbf{w} = (1, 1, 1, 1)$ / unstandardized",
-  "$\\mathbf{w} = (1, 1, 1, 1)$ / standardized",
-  "$\\mathbf{w} = (10, 1, 1, 1)$",
-  "$\\mathbf{w} = (1, 10, 1, 1)$",
-  "$\\mathbf{w} = (1, 1, 10, 1)$",
-  "$\\mathbf{w} = (1, 1, 1, 10)$"
-)
-colnames(final_tab_sim1) <- c(
-  "$\\mathit{SSE}$",
-  "$\\mathit{SSE_{Var}}$", 
-  "$\\mathit{SSE_{Skew}}$",
-  "$\\mathit{SSE_{Kurtosis}}$",
-  "Average"
-)
-note <- paste0("Cells indicate how well the maximization of $\\mathit{SSE_{kplusGeneralized}}$ fulfilled each of the four constituting criteria $\\mathit{SSE}$, $\\mathit{SSE_{Var}}$, $\\mathit{SSE_{Skew}}$, and $\\mathit{SSE_{Kurtosis}}$, depending on the weighting scheme.
-For each simulation run, the relative performance was computed as the result that was obtained by the given weighting scheme, divided by the best result that was obtained by any weighting scheme in this simulation run.
-Cell values represent the average relative performance across all ", length(results), " simulation runs.")
-
-apa_table(
-  final_tab_sim1, 
-  escape = FALSE, 
-  caption = "Average relative performance of the competing weighting schemes in Simulation 1.",
-  note = note,
-  stub_indents = list("Unweighted approaches" = 1:2,
-                      "Weighted approaches" = 3:6)
-)
-
-unstd_best <- sapply(results, function(x) names(which.max(x[, 5]))) == "unweighted_std"
-
-# significance test
-sig_test <- prop.test(x = sum(unstd_best), n = length(unstd_best), p = 1/6)
-
-```
-
-Table 1 illustrates how well each of the six weighting schemes fulfilled each of the four constituting criteria, averaged across the `r length(results)` simulation runs.
-For the weighted approaches, the criterion that had received the highest weight during the optimization was fulfilled to the highest degree, as should be expected.
-The unweighted standardization approach yielded the best average performance across the four criteria.
-While the degree of advantage by itself does not seem to be remarkable, this pattern was observed in an astonishing `r mean(unstd_best) * 100`% of all simulation runs, while the level of chance would have been `r 1/6 * 100`% (`r format_p(sig_test$p.value)`).
-Among the competing weighting schemes, the unweighted standardization approach therefore yielded the most promising results.
-It performed also remarkably strong with regard to the standard $\mathit{SSE}$, even slightly outperforming the weighting scheme that explicitly favoured the $\mathit{SSE}$.
-Hence, the unweighted standardization approach implicitly gave high priority to minimizing differences in means.
-This is a pleasant result because, as discussed earlier, for the higher order $\mathit{SSE}$s to be valid, the group means have to be close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ has high priority.
-
-## Simulation 2: Comparing k-plus to other anticlustering objectives
-
-In Simulation 2, the performance of two k-plus criteria was compared to the performance of two traditional anticlustering criteria.
-"Standard" k-plus anticlustering was implemented by maximizing $\mathit{SSE_{kplus}}$, thus minimizing differences with regard to group means and variances.
-"Generalized" k-plus anticlustering was implemented by maximizing $\mathit{SSE_{kplusGeneralized}}$ using $T = 4$, thus minimizing differences with regard to group means, variances, skewness and kurtosis.
-The unweighted standardization approach that performed best in Simulation 1 was employed.
-Thus, to optimize a k-plus criterion, the original data set was augmented to incorporate the additional criteria and all variables---the original and the augmented variables---were $z$-standardized.
-Subsequently, the k-means criterion $\mathit{SSE}$ was maximized on the basis of the augmented data set.
-The k-plus methods were compared to (i) k-means anticlustering, minimizing differences with regard to group means, (ii) diversity anticlustering, maximizing the sum of pairwise Euclidean distances within groups, and also (iii) a random allocation of elements to groups.
-
-### Conditions
+## Conditions
 
@@
 
-### Evaluation criteria
+## Evaluation criteria
 
-After the four anticlustering methods and the random allocation procedure had been applied to the 10,000 data sets for $K = 2$, $K = 3$, $K = 4$, I investigated how well the competing methods were able to create similar groups.
-To this end, I analysed the discrepancy in means, standard deviations, skewness and kurtosis between groups---for each data set for each $K$.
-To quantify the discrepancy in group means, the following computation was used: for each of the (2, 3, 4 or 5) features, all (2, 3 or 4) group means were computed.
-For each feature, the difference between the minimum and maximum group mean was used as a measure of discrepancy; the global discrepancy in means ($\Delta M$) was then determined as the average discrepancy across features.
-The same procedure was applied to compute the discrepancy in standard deviations ($\Delta \mathit{SD}$), skewness ($\Delta \mathit{Skew}$) and kurtosis ($\Delta \mathit{Kurtosis}$).
-
-As a secondary purpose of Simulation 2, I investigated the similarity of partitions obtained when optimizing the different anticlustering objectives.
-In cluster analysis, partitions identified through different approaches tend to overlap considerably, because similar elements tend to be placed in the same cluster while dissimilar elements tend to be placed in different clusters; this principle holds regardless of the specific clustering method.
-However, it is an open question whether the optimization of different anticlustering objectives leads to similar groupings of the elements in anticlusters.
-To address this question, I computed the adjusted Rand index (ARI) for each pair of partitions that were returned by the competing methods [@rand1971; @hubert1985].
-The ARI is an index of the similarity between two partitions, where an index of 0 indicates no similarity and a value of 1 indicates the highest possible similarity (i.e., when two partitions are the same).
-The pairwise ARIs were computed for each simulation run, i.e., for data set and for each $K$.
-
-### Results
-
-Table 2 displays the global simulation results aggregated across all conditions.
-Standard k-means performed best at minimizing discrepancy with regard to means ($\Delta \mathit{M}$), but did not outperform a random assignment with regard to all other distribution characteristics.
-K-plus anticlustering addressed $\Delta \mathit{M}$ almost as well as k-means, and at the same time minimized discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$); both k-plus approaches strongly outperformed all other methods with regard to minimizing $\Delta \mathit{SD}$.
-
-\begin{singlespace}
-
-```{r Table1}
-
-# Load required packages
-library(dplyr)
-library(ggplot2)
-library(tidyr)
-
-## Analyze data for K = 2 and K = 3 and K = 4
-simulation_results <- list()
-for (K in 2:4) {
-  filename <- paste0("./Simulation_2/results-K", K, "-objectives-raw.csv")
-  df <- read.csv(filename, sep = ";", stringsAsFactors = FALSE)
-  df$K <- K
-  simulation_results[[paste0("K-", K)]] <- df
-}
-
-df <- do.call(rbind, simulation_results)
-rownames(df) <- NULL
-
-# Make long format
-ldf <- pivot_longer(
-  df,
-  cols = paste0(c("means", "sd", "skew", "kur"), "_obj"),
-  names_to = "Objective",
-  names_pattern = "(.*)_obj"
-)
-
-ldf <- ldf[ldf$method != "k-plus-correlation", ] # removed because this deserves closer attention than what was given in the paper; check out the first version of the preprint for the previous results
-
-ldf$method[ldf$method == "k-plus-skew-kurtosis"] <- "k-plus generalized"
-
-## Global results, aggregated across all simulation variables
-tab <- ldf %>% 
-  group_by(method, Objective) %>% 
-  summarise(Mean = round(mean(value), 3)) %>% 
-  pivot_wider(names_from = Objective, values_from = Mean) %>% 
-  select(c(means, sd, skew, kur))
-
-colnames(tab) <- c("",
-  "$\\Delta M$", 
-  "$\\Delta \\mathit{SD}$",
-  "$\\Delta \\mathit{Skew}$",
-  "$\\Delta \\mathit{Kurtosis}$"
-)
-
-apa_table(
-  tab,
-  caption = "Results of Simulation 2.",
-  note = "The global results of the simulation study aggregated across all conditions.
-Cells contain information about the average global between-group discrepancy with regard to means, standard deviations, skewness and kurtosis.
-Lower values indicate less discrepancy, i.e., higher between-group similarity.
-Each cell value is the result of averaging across 30,000 data points (10,000 data sets $\\times$ $K = 2$, $K = 3$ or $K = 4$).",
-  escape = FALSE,
-  align = c("r")
-)
-
-```
-
-\end{singlespace}
-
-```{r Figure1, fig.width = 8, fig.height = 10, fig.cap = "Results of the simulation study: Depicts the performance of the anticlustering methods with regard to minimizing discrepancy in means, variances, skewness and kurtosis (split by $N$, averaged across the remaining variables that varied in the simulation)."}
-
-# Plot the results, by N
-axis_ticks <- seq(24, 300, by = 12)
-axis_labels <- axis_ticks
-axis_labels[c(FALSE, TRUE)] <- ""
-
-ldf %>% 
-  group_by(method, Objective, N) %>% 
-  summarise(Mean = mean(value)) %>% 
-  filter(method != "random") %>% 
-  mutate(
-    Objective = ordered(
-      Objective, 
-      levels = c("means", "sd", "skew", "kur"),
-      labels = paste0(
-        "Delta[italic(",
-        c("M", "SD", "Skew", "Kurtosis"),
-        ")]"
-      )
-    ) 
-  ) %>%
-  ggplot(aes(x = N, y = Mean, colour = method)) + 
-  geom_line(aes(linetype = method), size = .85) + 
-  facet_grid(rows = vars(Objective), scales = "free", labeller = label_parsed) + 
-  ylab("Mean discrepancy")+
-  xlab(expression(italic(N))) +
-  theme_bw(base_size = 16) + 
-  scale_x_continuous(
-    breaks = axis_ticks, 
-    labels = axis_labels
-  )
-
-```
-
-Figure 2 shows that with increasing $N$, k-means and the k-plus methods converged on the same level with regard to minimizing $\Delta \mathit{M}$.
-This is an important observation because it shows that k-plus anticlustering is capable of addressing multiple objectives at the same time---especially when $N$ increases.
-However, while k-plus anticlustering was capable of addressing multiple objectives simultaneously, increasing the number of optimization criteria made it more difficult to fulfil each single criterion: generalized k-plus performed worse at minimizing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$ than standard k-plus that only considered the mean and variances.
-However, as intended, generalized k-plus was best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$ and at the same time, this k-plus objective maintained a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.
-
-It is maybe surprising that generalized k-plus does not seem to have performed much better than diversity anticlustering with regard to skewness and kurtosis, even though the objective has been specifically tailored to these criteria.
-Figure 2 gives a more fine grained assessment of this observation: Surprisingly, for small $N$, generalized k-plus was even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
-However, with increasing $N$, it clearly outperformed diversity anticlustering.
-These results illustrate the cost of optimizing several criteria at the same time: For low $N$, generalized k-plus was apparently preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$.
-Increasing $N$ then facilitated to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.
-Generally, it should be noted that diversity anticlustering is a good all-rounder method that tends to address all distribution characteristics.
-For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering was outperformed.
-The Appendix includes additional figures splitting the simulation results according to the other variables that varied between simulation runs (number of features; correlation between features; standard deviation of features; number of groups).
-
-```{r}
-
-ARIs <- as.matrix(read.table("Simulation_2/global_results_ari.csv", sep = ","))
-colnames(ARIs) <- c(
-  "diversity", "k-means", "k-plus", "k-plus generalized", "random"
-)
-rownames(ARIs) <- colnames(ARIs)
-ARIs <- round(ARIs, 3)
-ARIs[upper.tri(ARIs)] <- ""
-
-# removing leading 0
-ARIs <- gsub("0.", ".", ARIs, fixed = TRUE)
-
-# remove diagonal
-diag(ARIs) <- ""
-ARIs <- ARIs[, -ncol(ARIs)]
-ARIs <- ARIs[-1, ]
-
-apa_table(
-  ARIs,
-  caption = "Average pairwise ARI returned by the competing methods in Simulation 2.",
-  note = "Each cell value is the result of averaging across 30,000 data points (10,000 data sets $\\times$ $K = 2$, $K = 3$ or $K = 4$).",
-  escape = FALSE,
-  align = c("r")
-)
-
-```
-
-To answer the question if different anticlustering partitions are similar or dissimilar, Table 3 illustrates the degree of similarity between anticlustering partitions that were identified in Simulation 2.
-Overall, there was hardly any similarity between anticlustering partitions that were obtained via optimization of the different criteria.
-The average ARI was almost 0 in all cases---indicating no similarity of partitions whatsoever.
-Even the pairwise ARI of the two k-plus criteria was fractional, although the criteria overlapped considerably in their optimization goal.
-This constitutes an interesting result, illustrating that high between-group similarity can be achieved via very different partitions.
-
-# Applying k-plus anticlustering using the R package anticlust
-
-This section demonstrates how readers can easily apply k-plus anticlustering using the R package `anticlust` [@papenberg2020].
-The package can be installed in the R environment using the `install.packages()` command.
-
-```R
-install.packages("anticlust")
-```
-
-The function `kplus_anticlustering()` available in `anticlust` implements the unweighted sum approach for k-plus anticlustering.
-
-## Example Application I
-
-In the first example, I used the OASIS norming data that is freely available online [@kurdi2017introducing].
-Kurdi et al.
-assembled 900 open-access colour images and collected ratings on two affective dimensions: arousal and valence.
-@brielmann2019intense measured how the same 900 images were rated with regard to beauty.
-I used the function `kplus_anticlustering()` to divide all images into 9 groups of 100 images each.
-The first argument of the function call is a variable named `features`, which contains the image data as a table: 900 rows representing images and 3 columns representing the features beauty, arousal, and valence.
-An OSF repository accompanying this manuscript contains the full code to reproduce the example [@osf2023].
-
-```R
-library(anticlust) # load the package anticlust
-
-kplus_anticlustering(
-  features,
-  K = 9,
-  variance = TRUE,
-  skew = TRUE,
-  kurtosis = FALSE,
-  T = NULL,
-  method = "local-maximum",
-  standardize = TRUE
-)
-```
-
-The function call takes about 10 seconds on a contemporary personal computer.
-The function `kplus_anticlustering()` has three boolean arguments to specify whether variance, skewness and kurtosis should be included as part of the k-plus criterion.
-Only the argument `variance` is set to `TRUE` by default---corresponding to the understanding that k-plus is an extension to k-means that at least equalizes variances in addition to means.
-The arguments corresponding to skewness and kurtosis have to be specifically "turned on", i.e., by setting them to `TRUE` in the function call of `kplus_anticlustering()`.
-If other higher order moments should be included as part of the optimization, the optional argument `T` can be used to specify how many moments should be used.
-The argument `method = "local-maximum"` ensures that the k-plus criterion is optimized using the local maximum search method described in the previous section.
-The argument `K` describes the number of groups.
-By default, the function ensures that all features---the original and the augmented variables---are standardized before the optimization starts.
-Thus, by default, the unweighted standardization approach is used.
-Whether standardization is used can be adjusted using the boolean argument `standardize`.
-
-Table 4 shows the results of the anticlustering application by listing the descriptive statistics (means, standard deviations and skewness) for each of the 9 groups and for each of the 3 features.
-Table 4 also contains the results after applying k-means anticlustering and diversity anticlustering.
-It is shown that---up to two decimals---k-plus anticlustering perfectly matched all groups with regard to the features' means and standard deviations, and the skewness was also very evenly matched.
-K-means anticlustering also perfectly matched the mean values, but showed decreased performance with regard to similarity in standard deviations and skewness.
-Diversity anticlustering was well-suited to match means, standard deviations and skew, but was slightly outperformed by k-plus.
-The latter result is not surprising because---unlike k-plus anticlustering---diversity anticlustering does not directly maximize similarity with regard to these criteria, but instead maximizes within-group heterogeneity.
-
-```{r}
-
-example1_file <- "results_oasis_same_size.csv"
-if (!file.exists(example1_file)) {
-
-    oasis <- read.csv(
-    "https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv"
-    )
-
-    features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
-
-    K <- 9
-
-    # k-plus variance / skew
-    groups_kplus <- kplus_anticlustering(
-        features,
-        K = K,
-        variance = TRUE,
-        skew = TRUE,
-        method = "local-maximum",
-        standardize = TRUE
-    )
-
-    # k-means
-    groups_kmeans <- anticlustering(
-        features,
-        K = K,
-        objective = "variance",
-        method = "local-maximum",
-        standardize = TRUE
-    )
-
-    # diversity
-    groups_diversity <- anticlustering(
-        features,
-        K = K,
-        objective = "diversity",
-        method = "local-maximum",
-        standardize = TRUE
-    )
-
-    # Functions to compute means / SD / skew per group 
-    descriptives_by_group <- function(features, anticlusters, FUN, name) {
-        df <- data.frame(
-            t(data.frame(lapply(by(features, anticlusters, FUN), c)))
-        )
-        df$group <- 1:nrow(df)
-        df$Descriptive = name
-        df
-    }
-
-    all_descriptives_by_group <- function(features, anticlusters) {
-    means <- descriptives_by_group(features, anticlusters, colMeans, "mean")
-    sds <- descriptives_by_group(
-        features, 
-        anticlusters, 
-        function(x) sapply(x, sd),
-        "sds"
-    )
-    skew <- descriptives_by_group(
-        features, 
-        anticlusters, 
-        function(x) sapply(x, DescTools::Skew),
-        "skew"
-    )
-    rbind(means, sds, skew)
-    }
-
-    kplus <- all_descriptives_by_group(features, groups_kplus)
-    kplus <- data.frame(kplus, Objective = "kplus")
-
-    kmeans <- all_descriptives_by_group(features, groups_kmeans)
-    kmeans <- data.frame(kmeans, Objective = "kmeans")
-    diversity <- all_descriptives_by_group(features, groups_diversity)
-    diversity <- data.frame(diversity, Objective = "diversity")
-
-    all_objs <- rbind(kplus, kmeans, diversity) %>% 
-    pivot_wider(names_from = Descriptive, values_from = c(beauty_mean, Valence_mean, Arousal_mean))
-
-    write.table(
-        all_objs, example1_file, row.names = FALSE, sep = ";"
-    )
-
-} else {
-  all_objs <- read.csv(example1_file, sep = ";")
-}
-
-table_all_objs <- subset(all_objs, select = -Objective)
-table_all_objs <- as.matrix(table_all_objs)
-colnames(table_all_objs) <- c("Group", rep(c("$M$", "$\\mathit{SD}$", "Skew"), 3))
-table_all_objs <- apply(table_all_objs, 2, force_or_cut)
-#mode(table_all_objs) <- "character"
-
-apa_table(
-  table_all_objs,
-  caption = "Descriptive statistics for OASIS features by group and anticlustering method.",
-  note = "The k-plus criterion that was optimized included a term to minimize differences with regard to means, variances, and skewness.",
-  escape = FALSE,
-  align = c("c", rep("r", 9)),
-  col_spanners = list(Beauty = c(2, 4), Valence = c(5, 7), Arousal = c(8, 10)), 
-  stub_indents = list("\\textbf{K-plus anticlustering}" = 1:9,
-                      "\\textbf{K-means anticlustering}" = 10:18,
-                      "\\textbf{Diversity anticlustering}" = 19:27)
-)
-```
-
-Table 5 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups.
-K-plus anticlustering strives for between-group similarity regardless of group size.
-Unequal group sizes can be requested by passing the different group sizes to the argument `K`.
-The following code was used to equalize means and variances between four groups using the k-plus approach:
-
-```R
-kplus_anticlustering(
-  features,
-  K = c(100, 100, 100, 600),
-  method = "local-maximum"
-)
-```
-
-```{r}
-
-path <- "results_oasis_unequal_size.csv"
-
-if (!file.exists(path)) {
-  oasis <- read.csv(
-    "https://raw.githubusercontent.com/aenneb/OASIS-beauty/master/means_per_image.csv"
-  )
-
-  features <- subset(oasis, select = c(beauty_mean, Valence_mean, Arousal_mean))
-
-  K <- c(100, 100, 100, 600)
-  mean_sd_tabs <- list()
-  for (criterion in c("kplus", "variance", "diversity")) {
-    if (criterion == "kplus") {
-        oasis[[paste(criterion, "group", sep = "_")]] <-  kplus_anticlustering(
-            features,
-            K = c(100, 100, 100, 600),
-            method = "local-maximum"
-        )
-    } else {
-        oasis[[paste(criterion, "group", sep = "_")]] <- anticlustering(
-        features,
-        standardize = TRUE,
-        K = K,
-        objective = criterion,
-        method = "local-maximum"
-        )
-    }
-    mean_sd_tabs[[criterion]] <- mean_sd_tab(
-      features,
-      oasis[[paste(criterion, "group", sep = "_")]]
-    )
-    mean_sd_tabs[[criterion]] <- cbind(1:length(K), K, mean_sd_tabs[[criterion]])
-  }
-
-  # append all tables to the same table:
-  oasis_tab <- do.call(rbind, mean_sd_tabs)
-  rownames(oasis_tab) <- NULL
-  colnames(oasis_tab) <- c("Group", "N", "Beauty", "Valence", "Arousal")
-
-  write.table(
-    oasis_tab, path, row.names = FALSE, sep = ";"
-  )
-
-} else {
-  oasis_tab <- read.csv(path, sep = ";")
-}
-
-apa_table(
-  oasis_tab,
-  caption = "Performance of the diversity and k-plus objectives for unequal group sizes.",
-  note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
-  escape = FALSE,
-  align = c("r"),
-  stub_indents = list("\\textbf{K-plus anticlustering}" = 1:4,
-                      "\\textbf{K-means anticlustering}" = 5:8,
-                      "\\textbf{Diversity anticlustering}" = 9:12)
-)
-```
+## Results
 
@@
 
-Table 6 illustrates the results that are obtained when using standard[^standardkplus] k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each.
-Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are similar enough for any practical purposes.
-
-[^standardkplus]: I refer to *standard* k-plus anticlustering as the minimization of differences in means and variances, but not in higher order moments.
-
-```{r, echo = FALSE}
-features <- schaper2019[, 3:6]
-anticlusters <- anticlustering(
-  features,
-  K = 6,
-  objective = "kplus",
-  method = "local-maximum",
-  repetitions = 5,
-  standardize = TRUE
-)
-```
-
-```{r}
-
-tab <- mean_sd_tab(features, anticlusters)
-tab <- cbind(1:6, 96 / 6, tab)
-colnames(tab) <- c("Group", "N", "Typicality", "Atypicality", "Syllables", "Frequency")
-
-apa_table(
-  tab,
-  caption = "Descriptive statistics by group for the Schaper et al. (2019a; 2019b) data set, after applying k-plus anticlustering.",
-  note = "The cells contain mean values by cluster and the standard deviation (in parentheses).",
-  escape = FALSE,
-  align = c("r")
-)
-
-```
-
 # Discussion
-
-Anticlustering is a partitioning method that aims for high between-group similarity and high within-group heterogeneity.
-In this paper, I presented the k-plus criterion for anticlustering, which focuses on maximizing between-group similarity.
-K-plus represents between-group similarity as discrepancies in distribution moments between groups: means, variances, skewness, kurtosis, and higher order moments.
-Depending on the application, a different subset of these objectives can be included in the anticlustering process.
-The k-plus criterion is an extension of the classical k-means criterion, which only reflects how similar groups are with regard to their means.
-Interestingly, when using the unweighted sum approach presented here, the k-plus extension does not even change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.
-
-Two simulation studies and examples on real norming data showed that the k-plus criterion is well-suited to maximize between-group similarity with regard to multiple criteria.
-Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be considered as the default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be challenging in small data sets.
-In practice, users can try out different objectives and see which specification fulfils their needs.
-A reasonable approach is to start "greedy", e.g.
-by specifying that multiple distribution moments should equalized.
-If the results of such a greedy setup are not satisfying, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variance).
-Using the open source R package `anticlust`, the (repeated) application of k-plus anticlustering is easily accessible and free.
-
-[^binarydata]: A reasonable exception is given when the input data consists of binary variables because for binary data, the mean and the variance are directly dependent.
-
-## Limitations and outlook
-
-The present paper focused on the new k-plus criterion for anticlustering, but only briefly considered the multicriterion nature of this criterion.
-A simple unweighted sum approach was used to optimize the k-plus objective, while more sophisticated algorithmic procedures exist to tackle multicriterion optimization problems [e.g., @brusco2009cross].
-The sole focus on the objective function followed two reasonings.
-First, the applications and the simulations showed that the unweighted standardization approach yielded satisfying results that probably suffice in most applications.
-Moreover, it is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfil multiple criteria to a satisfactory degree (as suggested by Figure 1).
-Still, more sophisticated approaches may yield theoretically improved results, as well as the opportunity to deliberately select among partitions that may be preferred according to different k-plus objectives.
-@brusco2019 gave an example of how direct algorithms can be employed to optimize multiple criteria in anticlustering applications aiming for high within-group heterogeneity.
-Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment.
-Specifically, it is of high interest to investigate if the bicriterion approach by Brusco et al.
-can be transformed into a multicriterion algorithm for the k-plus objective.
-A second reason for focusing on the k-plus objective per se---independently from the algorithm that is used to optimize it---was to obtain a fair comparison to existing anticlustering objectives.
-In the simulation study, differences in performance could not have been interpreted if the objective function as well as the algorithm used to optimize it had differed between anticlustering methods.
-It was of particular interest to compare the k-plus objective with the popular diversity objective.
-@papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should be preferred because k-means does not equalize the spread of the distribution between groups.
-In contrast, maximizing the diversity leads to an appropriate balance between location and spread.
-Using the k-plus extension, however, the k-means criterion has regained attractivity: When opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.
-Future research should investigate whether this advantage also holds for other data distributions than considered here.
-For example, it may be hypothesized that the generalized k-plus criterion is even more useful when applied to data that is not symmetric or contains outliers, because it is able to explicitly incorporate the skewness and kurtosis as criteria in the anticlustering process.
-
-As an additional outlook, it should be noted that because the covariance is defined as an expected value (i.e., an average), the k-plus logic can be extended to minimize differences with regard to covariances.
-Similar to addressing distribution moments, covariances could be incorporated by appending additional variables to a data set.
-In this case, it would be necessary to compute an additional variable for each of pair of variables, because the covariance is the product of the differences between two variables and their respective means.
-Future research should investigate how covariances can properly be included with k-plus anticlustering.
-
-[^clusterediting]: In their study, they used the term anticluster editing to refer to the maximization of the diversity criterion.
-This is because the diversity criterion is minimized as part of the cluster editing method [@shamir2004vf].
-
-## Closing remarks on the nature of anticlustering
-
-```{r}
-df <- read.csv("partition_objectives.csv")
-```
-
-Anticlustering is usually implemented by reversing an objective function used in classical cluster analysis.
-Whereas cluster analysis seeks homogeneity within groups and separation between groups, the reversal results in heterogeneity within groups and similarity between groups.
-While tightly linked in a mathematical sense, an important semantic difference is that cluster analysis is concerned with finding "true" structure in the data, e.g., by defining a taxonomy on naturally occurring groups [@everitt1979].
-Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern.
-Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different.
-In contrast, different clustering solutions tend to look similar because good clusterings are characterized by assigning similar elements into the same cluster and dissimilar elements into different clusters.
-A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This result was observed in Simulation 2 and can also be clarified using Figure 1.
-Here, we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions.
-An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution.
-In contrast, there are only `r sum(min(df$kmeans)/ df$kmeans > .90)` partitions (`r mean(min(df$kmeans)/ df$kmeans > .90)*100`%) that approximate the global minimum in $\mathit{SSE}$ (i.e., the best k-means clustering solution) by even 10%.
-Thus, even though the problems are based on the same objective function, cluster analysis and anticlustering exhibit a stark discrepancy with regard to the solution space of partitions.
-Interestingly, the k-plus objective introduces an additional distinction between cluster analysis and anticlustering.
-Reversing the k-plus anticlustering objective no longer yields a useful clustering interpretation, because identifying groups that are different with regard to higher order moments does not constitute a clustering.
-Thus, even though k-plus extends the classical k-means clustering criterion, it is a pure anticlustering objective.
-
-[^dispersion]: An exception is given when using the dispersion objective, which ensures that the most similar elements are assigned to different clusters.
-However, anticlustering objectives striving for overall between-group similarity are largely unaffected by the grouping of individual item pairs.
-
-An interesting mixture of cluster analysis and anticlustering is fair clustering [e.g.
-@chhabra2021; @delBarrio2022].
-Fair clustering performs a cluster analysis under the restriction that the distribution of a set of "protected" attributes---such as race, sex or age---should be similar between clusters.
-Thus, as compared to pure cluster analysis, fair clustering biases the identification of groups with regard to the protected attributes, striving for algorithmic fairness.
-A relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data, while anticlustering divides data sets into purely artificial subgroups.
-However, these methods are logically similar because anticlustering corresponds to fair clustering when only considering the protected attributes.
-Hence, a fruitful outlook for research might be to investigate if approaches stemming from the literature on fair clustering can be adopted for anticlustering, and vice versa.
-
-## Conclusion
-
-This paper introduced k-plus anticlustering for maximizing between-group similarity, improving the classical k-means approach.
-K-plus anticlustering can be adopted whenever a group of elements has to be partitioned into equivalent parts.
-The method is accessible and easily usable via the R package `anticlust`.
 
@@
 \endgroup
-
-\newpage
-
-# (APPENDIX) Appendix {-}
-
-```{r child = "appendix.Rmd"}
-```

### 2024-01-31 paper.Rmd
abtract and outline

@@
     When assigning stimuli to sets in experimental psychology, researchers strive for high similarity between stimulus sets and high pairwise dissimilarity of stimuli within sets.
-Because stimulus assignment is usually substandard when done manually by researchers, automated algorithmic approaches are preferred to accomplish this goal.
+Because stimulus assignment is usually inadequate when done manually, automated algorithmic approaches are preferred.
 Brusco et al.
-(2020) proposed a bicriterion anticlustering algorithm (BILS) that simultaneously maximizes the diversity (corresponding to between-set similarity) and the dispersion (corresponding to pairwise stimulus dissimilarity within-sets).
-We provide a first independent assessment of BILS and present BILS-E, which extends BILS in three ways.
-First, a globally optimal value of dispersion is always returned.
-Second, the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices, for example to include cannot-link constraints in the anticlustering process.
-Third, BILS-E allows that other anticlustering objectives of between-group similarity, such as k-means or k-plus, can be optimized instead of only the diversity.
-Using our optimal approach as benchmark in a simulation study, we show that the original BILS identifies optimal dispersion values for much larger data sets than previously analyzed, but also highlight conditions where it fails to do so.
-A practical example shows how BILS-E can be used for stimulus assignment in experimental psychology.
+(2020) proposed a bicriterion anticlustering algorithm (BILS) that simultaneously maximizes the diversity (corresponding to between-set similarity) and the dispersion (corresponding to pairwise stimulus dissimilarity within sets).
+We provide a first independent assessment of BILS and extend it in three ways.
+First, by prefixing an integer linear programming algorithm, our extension ensures that a globally optimal value of dispersion is always returned; without it, it is unknown if the BILS's result is optimal.
+Second, the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices, for example to prevent specific pairs of stimuli from appearing in the same condition.
+Third, our extension allows that other anticlustering objectives of between-group similarity -- such as the k-means or k-plus objectives -- can be optimized instead of only the diversity.
+Two simulation studies provide an extensive evaluation of the original BILS and our extension.
+Using the optimal approach as benchmark, we show that the original BILS identifies optimal dispersion values for much larger data sets than previously analyzed, but we also highlight conditions where it fails to do so reliably.
+A practical example shows how the BILS can be used for stimulus assignment in experimental psychology.
 The original BILS as well as our extension are both freely available via the open source R package anticlust.
@@
 ```
+
+## Outline
+
+- Start with use case: Stimulus assignment in within-subjects experiments (and stick with it throughout the paper, only briefly mention other use cases)
+  * Needed: high similarity between stimulus set to reduce variance in responses attributed to other factors than the experimental manipulation.
+This is possible using `anticlust`, which has been applied many times since its introduction in @papenberg2020 (cite references)
+  * Also needed (point of this paper!): High pairwise dissimilarity of stimuli in the same set. 
+- Simultaneous optimization of between-set similarity and pairwise within-set dissimilarity was introduced by Brusco et al.'s alogrithm, which is also available in `anticlust` since 2021 (version 0.6.0)
+- Here, we provide researchers with a practical introduction to their method and also present several extensions, which allow it to cover more use cases
+  (1) Always return optimal solution
+  (2) Use different distance matrix for dispersion and diversity (if different information is needed for both criteria, or to induce custom cannot-link constraints)
+  (3) Extend the BILS's capacities towards other measures of between-group similarity (k-means / k-plus)
+- Technical introduction to "vanilla" BILS
+- Technical introduction to our extensions
+  * Optimal dispersion method: Basic algorithm + graph coloring ILP
+    * How to maintain optimal dispersion while the BILS runs? Outline three approaches!
+  * Different distance matrices: does not change algorithm, just computation / implementation
+  * Extend towards k-means: Use squared Euclidean distance (not really extension, just different input); compute "average" diversity if group sizes are unequal (k-plus is basically the same, but do not use the additional k-plus variables for the dispersion distance matrix)
+- Simulation Study 1: Two research questions
+  (1) How often does vanilla BILS find optimal solutions
+  (2) Which of the three extension is best for optimizing between-group similarity while maintaining optimal dispersion?
+- Simulation Study 2: More difficult conditions (do not continuous data; K > 4 groups)
+- Application (presenting R code)
+  * Optimize k-plus criterion using norming data; dispersion is based on pairwise orthographic dissimilarity between words
+- Discussion
+
+## Introduction
+
+In within-subjects experiments, participants are presented with several stimulus sets that are affiliated with the different experimental conditions.
+Due to carry-over effects, showing the same stimuli in different conditions is oftentimes prohibited.
+This is evident in studies on memory, where studying a list of words in one conditions necessarily makes it impossible to repeat their study in a later phase; memory cannot be erased after an experimental block has been completed [@lahl2006].
+While the experimental logic dictates that different lists must be used, the lists should be as parallel as possible with regard to variables that affect the response variable.
+@lintz2021 gave two reasons why it is important that study participants should be confronted with balanced lists: (a) if lists are not sufficiently parallel, the variance of within-subjects effect sizes due to the experimental manipulation will be inflated, reducing statistical power---if there is no proper statistical control of the stimulus materials, which is often the case [@judd2012]; (b) highly unbalanced sets may introduce secondary effects beyond the influence of the intrinsic stimulus features.
+That is, highly unequal lists may induce a strategic behavior shift between experimental blocks.
+Such a shift may for example result from suspicion, boredom or frustration, invalidating any observed differences between conditions.
+Importantly, this bias cannot be accounted for using statistical analysis.
+
+**Using anticlustering to equate stimulus sets**
+
+**BUT DISPERSION IS ALSO IMPORTANT!**
+
+For example, in tasks of recognition memory, participants commonly have to decide if a target stimuli (e.g., a word) has been presented in an earlier phase of the experiment.
+In this case, words that are orthographically similar to a target word should serve as distractors.
+
+The BILS algorithm has been available in the anticlust package since 2021, but we assume that many researchers do not yet know about its potential, also given that the introduction by @brusco2019 was rather technical in nature.
+For now, users seem to rely on the unicriterion methods focusing on between-set similarity only.
+Therefore, we provide readers with an application-focused introduction and extend BILS with parameters that may be useful when designing experiments.
+
+---
+
+**OLD**
 

### 2024-02-01 Simulation_2/Analysis_BILS.R
better aggregate analysis

@@ -36,10 +36,30 @@ tapply(df$VANILLA_FOUND_OPTIMUM, list(df$K, df$separate_dispersion_distances), m
 
 chisq.test(table(df$VANILLA_FOUND_OPTIMUM, df$separate_dispersion_distances))
 
-# not sure if this analysis makes sense:
-colMeans(df[df$VANILLA_FOUND_OPTIMUM & df$separate_dispersion_distances == 0, grepl("DIV", colnames(df))])
+# Descriptives:
+df |>
+  filter(VANILLA_FOUND_OPTIMUM == 1) |> # this actually introduces a bias
+  group_by(separate_dispersion_distances, K) |>
+  summarize(
+    E_1 = mean(DIV_E_1),
+    E_ALL = mean(DIV_E_ALL),
+    E_ALL_RESTRICTED = mean(DIV_E_ALL_RESTRICTED),
+    VANILLA = mean(DIV_VANILLA)
+  )
 
-colMeans(df[df$VANILLA_FOUND_OPTIMUM & df$separate_dispersion_distances == 1, grepl("DIV", colnames(df))])
+
+# Descriptives without bias, only Extended versions
+df |>
+  group_by(separate_dispersion_distances, K) |>
+  summarize(
+    E_1 = mean(DIV_E_1),
+    E_ALL = mean(DIV_E_ALL),
+    E_ALL_RESTRICTED = mean(DIV_E_ALL_RESTRICTED)
+  )
+
+
+
+
 
 tt <- df[df$VANILLA_FOUND_OPTIMUM, ]
 
@@ -53,11 +73,11 @@ t.test(tt$DIV_E_ALL, tt$DIV_E_ALL_RESTRICTED, paired = TRUE)
 
 # Important take aways (preliminary)
 
-# Diversity is lower if the dispersion is optimized on the basis of different data (that makes sense I guess)
-# E_ALL is better than E_1 (it seems...)
-# VANILLA BILS has more difficulties finding the optimal solution if the dispersion is optimized on the basis of a different data set
-# The max dispersion problem can be solved in reasonable time using an open source solver (!) for rather large data sets (N = 300, K = 4)
-# Restricted method may be best (?). This would indicate that using LCW is just as effective as BILS
+# - Diversity is lower if the dispersion is optimized on the basis of different data (that makes sense I guess)
+# - E_ALL is better than E_1
+# - VANILLA BILS has more difficulties finding the optimal diversity if the dispersion is optimized on the basis of a different data set
[…]

### 2024-02-01 paper.Rmd
minor

@@
 Brusco et al.
-(2020) proposed a bicriterion anticlustering algorithm (BILS) that simultaneously maximizes the diversity (corresponding to between-set similarity) and the dispersion (corresponding to pairwise stimulus dissimilarity within sets).
+(2020) proposed a bicriterion anticlustering algorithm (BILS) that simultaneously maximizes the diversity (corresponding to between-set similarity) and the dispersion (corresponding to pairwise within-set dissimilarity ).
 We provide a first independent assessment of BILS and extend it in three ways.
@@
 Two simulation studies provide an extensive evaluation of the original BILS and our extension.
-Using the optimal approach as benchmark, we show that the original BILS identifies optimal dispersion values for much larger data sets than previously analyzed, but we also highlight conditions where it fails to do so reliably.
-A practical example shows how the BILS can be used for stimulus assignment in experimental psychology.
+Using the optimal approach as benchmark, we show that the original BILS identifies optimal dispersion values for much larger data sets than previously analyzed, but we also highlight conditions where it fails to do so.
+A practical example shows how the extended BILS can be used for stimulus assignment in experimental psychology.
 The original BILS as well as our extension are both freely available via the open source R package anticlust.
@@
 Due to carry-over effects, showing the same stimuli in different conditions is oftentimes prohibited.
-This is evident in studies on memory, where studying a list of words in one conditions necessarily makes it impossible to repeat their study in a later phase; memory cannot be erased after an experimental block has been completed [@lahl2006].
+This is evident in studies on memory, where studying a list of words in one conditions necessarily makes it impossible to repeat their study without being influenced by the earlier study phase; memory cannot be erased after an experimental block has been completed [@lahl2006].
 While the experimental logic dictates that different lists must be used, the lists should be as parallel as possible with regard to variables that affect the response variable.
-@lintz2021 gave two reasons why it is important that study participants should be confronted with balanced lists: (a) if lists are not sufficiently parallel, the variance of within-subjects effect sizes due to the experimental manipulation will be inflated, reducing statistical power---if there is no proper statistical control of the stimulus materials, which is often the case [@judd2012]; (b) highly unbalanced sets may introduce secondary effects beyond the influence of the intrinsic stimulus features.
-That is, highly unequal lists may induce a strategic behavior shift between experimental blocks.
-Such a shift may for example result from suspicion, boredom or frustration, invalidating any observed differences between conditions.
-Importantly, this bias cannot be accounted for using statistical analysis.
+@lintz2021 gave two reasons as to why it is important that study participants should be confronted with balanced lists: (a) if lists are not sufficiently parallel, the variance of within-subjects effect sizes due to the experimental manipulation will be inflated, reducing statistical power---if there is no proper statistical control of the stimulus materials, which is often the case [@judd2012]; (b) highly unbalanced sets may introduce secondary effects beyond the influence of the intrinsic stimulus features.
+For example, highly biased lists may induce a strategic behavior shift between experimental blocks, possibly due to suspicion, boredom or frustration.
+These effects invalidate any observed differences between conditions and importantly, this bias cannot be accounted for via statistical techniques.
+
+Traditionally, assigning stimuli to experimental conditions was done manually [e.g., @schaper2019metamory] or randomly [e.g., @kroneisen2018], both of which methods do not guarantee that the stimulus sets will be similar [@papenberg2020; @cutler1981; @vancasteren2007; @armstrong2012].
+Generally, automated approaches are preferred that quantify similarity on the basis of the stimulus features and use an algorithm to assign stimuli in such a way that this similarity is optimized [@brusco2019; @papenberg2020].
 

### 2024-02-01 paper.Rmd
minor

@@
 
-Traditionally, assigning stimuli to experimental conditions was done manually [e.g., @schaper2019metamory] or randomly [e.g., @kroneisen2018], both of which methods do not guarantee that the stimulus sets will be similar [@papenberg2020; @cutler1981; @vancasteren2007; @armstrong2012].
+Traditionally, assigning stimuli to experimental conditions was done manually [e.g., @schaper2019metamory] or randomly [e.g., @kroneisen2018], both of which methods however do not ensure that the stimulus sets will be similar [@papenberg2020; @cutler1981; @vancasteren2007; @armstrong2012].
 Generally, automated approaches are preferred that quantify similarity on the basis of the stimulus features and use an algorithm to assign stimuli in such a way that this similarity is optimized [@brusco2019; @papenberg2020].
+One approach that has recently gained traction
 

### 2024-02-01 Application/Application.R

@@ -1,0 +1,164 @@
+# https://osf.io/w8jzn/ # OFS project page for paper
+# https://osf.io/y7efm
+
+file <- ifelse(file.exists("Materials Experiment 3.csv"), "Materials Experiment 3.csv", "https://osf.io/download/y7efm/")
+
+df <- read.csv(file)
+
+# Matrix of normalized Levenshtein distances
+Levenshtein_Dist <- function(x) {
+  x <- as.character(x)
+  item_combinations <- t(combn(length(x), 2))
+  dists <- matrix(NA, nrow = length(x), ncol = length(x))
+  for (combs in 1:nrow(item_combinations)) {
+    j <- item_combinations[combs, ][1]
+    i <- item_combinations[combs, ][2]
+    normalizer <- sum(c(nchar(x[i]), nchar(x[j])))
+    dists[i, j] <- adist(x[i], x[j]) / normalizer
+  }
+  as.dist(dists)
+}
+
+positions_min_dist <- function(x) {
+  positions_by_value(x, min)
+}
+positions_max_dist <- function(x) {
+  positions_by_value(x, max)
+}
+
+positions_by_value <- function(x, fun = min, value = NULL) {
+  tt <- as.matrix(x)
+  tt[upper.tri(tt)] <- NA
+  diag(tt) <- NA
+  if (!is.null(value)) {
+    return(which(tt == value, arr.ind = TRUE))
+  }
+  which(tt == fun(tt, na.rm = TRUE), arr.ind = TRUE)
+}
+
+lt <- Levenshtein_Dist(df$Item)
+hist(lt)
+min(lt)
+
+# observe pairs of similar orthography
+df[t(positions_min_dist(lt)),]
+df[t(positions_by_value(lt, value = sort(lt)[2])), ]
+df[t(positions_by_value(lt, value = sort(lt)[4])), ]
+df[t(positions_by_value(lt, value = sort(lt)[5])), ]
+df[t(positions_by_value(lt, value = sort(lt)[7])), ]
+# Definition of Levenshtein normalization as done above seems to be reasonable
+
+bathroom <- subset(df, Scene == "bathroom")
+bathroom_dists <- Levenshtein_Dist(bathroom$Item)
+runs <- 5000
+opt <- optimal_dispersion( # E ALL NOT GOOD HERE?!
+  bathroom_dists,
+  K = 3
+)
+
+features <- bathroom[, c("ConsistencyRating", "InconsistencyRating_Bedroom",
[…]

### 2024-02-21 Simulation_2/Analysis_BILS.R
analyze results in dependence of duplicates

@@ -29,7 +29,7 @@ tapply(df$time_vanilla_s, list(df$K), max) |> round(2)
 table(df$RUNS_BILS_VANILLA)
 table(df$RUNS_BILS_VANILLA, df$K)
 
-(table(df$RUNS_BILS_VANILLA, df$K) / sum(table(df$RUNS_BILS_VANILLA))) |> round(3) * 100
+prop.table(table(df$RUNS_BILS_VANILLA, df$K), margin = 2) |> round(3) * 100
 
 # this is highly interesting: time to solve max dispersion problem optimally is related to the number of runs the heuristic requires to find optimal solution (not surprising, but nice)
 tapply(df$time_optimal_s, list(df$RUNS_BILS_VANILLA), median) |> round(2)
@@ -37,6 +37,7 @@ tapply(df$time_optimal_s, list(df$RUNS_BILS_VANILLA), median) |> round(2)
 df$VANILLA_FOUND_OPTIMUM <- df$RUNS_BILS_VANILLA != -1
 tapply(df$VANILLA_FOUND_OPTIMUM, df$K, mean)
 
+sum(df$N_DUPLICATE_PARTITIONS == (df$RUNS_BILS_VANILLA - 1)) # !!!!
 
 df$VANILLA_FOUND_OPTIMUM_AFTER_5 <- df$RUNS_BILS_VANILLA == 5
 mean(df$VANILLA_FOUND_OPTIMUM_AFTER_5)
@@ -58,9 +59,11 @@ df |>
 
 table(df$N_DUPLICATE_PARTITIONS > 0, df$K) # also check out results in dependence of duplicate partitions!
 
+df$MAXIMUM_RESTRICTION <- df$N_DUPLICATE_PARTITIONS == (df$RUNS_BILS_VANILLA - 1)
+
 df |>
   filter(VANILLA_FOUND_OPTIMUM == 1) |> # this actually introduces a bias
-  group_by(K, N_DUPLICATE_PARTITIONS > 0) |>
+  group_by(K, MAXIMUM_RESTRICTION) |>
   summarize(
     E_1 = mean(DIV_E_1),
     E_ALL = mean(DIV_E_ALL),
@@ -68,6 +71,9 @@ df |>
     VANILLA = mean(DIV_VANILLA)
   )
 
+# if there is only one unique input partition, VANILLA is better than RESTRICTED,
+# but otherwise not!
+  
 # Descriptives without bias, only Extended versions
 df |>
   group_by(K) |>

### 2024-02-21 Simulation_2/Analysis_BILS.R
some cleanup

@@ -1,7 +1,6 @@
 library(anticlust)
 library(tidyr)
 library(dplyr)
-library(BayesFactor)
 
 df <- read.csv("results_bils.csv", sep = ";")
 sum(!complete.cases(df)) # =)
@@ -55,9 +54,10 @@ df |>
     E_ALL = mean(DIV_E_ALL),
     E_ALL_RESTRICTED = mean(DIV_E_ALL_RESTRICTED),
     VANILLA = mean(DIV_VANILLA)
-  )
+  ) |>
+  as.data.frame()
 
-table(df$N_DUPLICATE_PARTITIONS > 0, df$K) # also check out results in dependence of duplicate partitions!
+table(df$N_DUPLICATE_PARTITIONS > 0, df$K) # also checkCP Conference Index in DBLP out results in dependence of duplicate partitions!
 
 df$MAXIMUM_RESTRICTION <- df$N_DUPLICATE_PARTITIONS == (df$RUNS_BILS_VANILLA - 1)
 
@@ -83,11 +83,7 @@ df |>
     E_ALL_RESTRICTED = mean(DIV_E_ALL_RESTRICTED)
   )
 
-
-
-
-
-tt <- df[df$VANILLA_FOUND_OPTIMUM, ]
+tt <- df[df$VANILLA_FOUND_OPTIMUM, ] # slight bias in favor of VANILLA, in particular for K = 6, K = 7
 
 t.test(tt$DIV_VANILLA, tt$DIV_E_1, paired = TRUE)
 t.test(tt$DIV_VANILLA, tt$DIV_E_ALL, paired = TRUE)
@@ -95,44 +91,13 @@ t.test(tt$DIV_VANILLA, tt$DIV_E_ALL_RESTRICTED, paired = TRUE)
 t.test(tt$DIV_E_1, tt$DIV_E_ALL, paired = TRUE)
 t.test(tt$DIV_E_1, tt$DIV_E_ALL_RESTRICTED, paired = TRUE)
 t.test(tt$DIV_E_ALL, tt$DIV_E_ALL_RESTRICTED, paired = TRUE)
-# VANILLA seems to best optimize the diversity, but does not always find the optimal dispersion. so this analysis is actually flawes because of systematic missing values... what is the most appropriate analysis? BILS-E with restrictions may even be better with regard to max diversity! (even in flawed analysis with dropouts)
 
 # Important take aways (preliminary)
 
[…]

### 2024-02-21 Simulation_2/Analysis_BILS.R
comment to include analysis

@@ -53,7 +53,7 @@ df |>
     E_1 = mean(DIV_E_1),
     E_ALL = mean(DIV_E_ALL),
     E_ALL_RESTRICTED = mean(DIV_E_ALL_RESTRICTED),
-    VANILLA = mean(DIV_VANILLA)
+    VANILLA = mean(DIV_VANILLA) # also add N per row to illustrate bias!
   ) |>
   as.data.frame()
 

### 2024-02-21 Simulation_2/Analysis_BILS.R
stuff

@@ -36,7 +36,7 @@ tapply(df$time_optimal_s, list(df$RUNS_BILS_VANILLA), median) |> round(2)
 df$VANILLA_FOUND_OPTIMUM <- df$RUNS_BILS_VANILLA != -1
 tapply(df$VANILLA_FOUND_OPTIMUM, df$K, mean)
 
-sum(df$N_DUPLICATE_PARTITIONS == (df$RUNS_BILS_VANILLA - 1)) # !!!!
+mean(df$N_DUPLICATE_PARTITIONS == (df$RUNS_BILS_VANILLA - 1)) # !!!!
 
 df$VANILLA_FOUND_OPTIMUM_AFTER_5 <- df$RUNS_BILS_VANILLA == 5
 mean(df$VANILLA_FOUND_OPTIMUM_AFTER_5)
@@ -53,13 +53,14 @@ df |>
     E_1 = mean(DIV_E_1),
     E_ALL = mean(DIV_E_ALL),
     E_ALL_RESTRICTED = mean(DIV_E_ALL_RESTRICTED),
-    VANILLA = mean(DIV_VANILLA) # also add N per row to illustrate bias!
+    VANILLA = mean(DIV_VANILLA), # also add N per row to illustrate bias!
+    N = n()
   ) |>
   as.data.frame()
 
-table(df$N_DUPLICATE_PARTITIONS > 0, df$K) # also checkCP Conference Index in DBLP out results in dependence of duplicate partitions!
+table(df$N_DUPLICATE_PARTITIONS > 0, df$K) # also check out results in dependence of duplicate partitions!
 
-df$MAXIMUM_RESTRICTION <- df$N_DUPLICATE_PARTITIONS == (df$RUNS_BILS_VANILLA - 1)
+df$MAXIMUM_RESTRICTION <- as.numeric(df$N_DUPLICATE_PARTITIONS == (df$RUNS_BILS_VANILLA - 1))
 
 df |>
   filter(VANILLA_FOUND_OPTIMUM == 1) |> # this actually introduces a bias
@@ -68,9 +69,14 @@ df |>
     E_1 = mean(DIV_E_1),
     E_ALL = mean(DIV_E_ALL),
     E_ALL_RESTRICTED = mean(DIV_E_ALL_RESTRICTED),
-    VANILLA = mean(DIV_VANILLA)
-  )
+    VANILLA = mean(DIV_VANILLA),
+    N = n()
+  ) |>
+  as.data.frame() |>
+  round(2)
 
[…]

### 2024-02-21 paper.Rmd
minor stuff abstract

@@
 Brusco et al.
-(2020) proposed a bicriterion anticlustering algorithm (BILS) that simultaneously maximizes the diversity (corresponding to between-set similarity) and the dispersion (corresponding to pairwise within-set dissimilarity ).
+(2020) proposed a bicriterion anticlustering algorithm (BILS) that simultaneously maximizes the diversity (corresponding to between-set similarity) and the dispersion (corresponding to pairwise within-set dissimilarity).
 We provide a first independent assessment of BILS and extend it in three ways.
-First, by prefixing an integer linear programming algorithm, our extension ensures that a globally optimal value of dispersion is always returned; without it, it is unknown if the BILS's result is optimal.
+First, by prefixing an exact algorithm, our extension ensures that a globally optimal value of dispersion is always returned; without it, it is unknown if the BILS's result is optimal.
 Second, the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices, for example to prevent specific pairs of stimuli from appearing in the same condition.
 Third, our extension allows that other anticlustering objectives of between-group similarity -- such as the k-means or k-plus objectives -- can be optimized instead of only the diversity.
-Two simulation studies provide an extensive evaluation of the original BILS and our extension.
+A simulation study provides an extensive evaluation of the original BILS and our extension.
 Using the optimal approach as benchmark, we show that the original BILS identifies optimal dispersion values for much larger data sets than previously analyzed, but we also highlight conditions where it fails to do so.

### 2024-02-21 paper.Rmd
minor

@@
   * Extend towards k-means: Use squared Euclidean distance (not really extension, just different input); compute "average" diversity if group sizes are unequal (k-plus is basically the same, but do not use the additional k-plus variables for the dispersion distance matrix)
-- Simulation Study 1: Two research questions
+- Simulation Study
   (1) How often does vanilla BILS find optimal solutions
   (2) Which of the three extension is best for optimizing between-group similarity while maintaining optimal dispersion?
-- Simulation Study 2: More difficult conditions (do not continuous data; K > 4 groups)
 - Application (presenting R code)

### 2024-02-29 paper.Rmd
abtract

@@
 ---
-title             : "An improved anticlustering method to maximize stimulus dispersion in experiments"
+title             : "An optimal anticlustering method to maximize stimulus dispersion in experiments"
 shorttitle        : "Maximum stimulus dispersion"
@@
 
-    When assigning stimuli to sets in experimental psychology, researchers strive for high similarity between stimulus sets and high pairwise dissimilarity of stimuli within sets.
+    When assigning stimuli to experimental conditions, researchers strive for high similarity between stimulus sets and high pairwise dissimilarity of stimuli within sets.
 Because stimulus assignment is usually inadequate when done manually, automated algorithmic approaches are preferred.
 Brusco et al.
-(2020) proposed a bicriterion anticlustering algorithm (BILS) that simultaneously maximizes the diversity (corresponding to between-set similarity) and the dispersion (corresponding to pairwise within-set dissimilarity).
-We provide a first independent assessment of BILS and extend it in three ways.
-First, by prefixing an exact algorithm, our extension ensures that a globally optimal value of dispersion is always returned; without it, it is unknown if the BILS's result is optimal.
-Second, the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices, for example to prevent specific pairs of stimuli from appearing in the same condition.
-Third, our extension allows that other anticlustering objectives of between-group similarity -- such as the k-means or k-plus objectives -- can be optimized instead of only the diversity.
-A simulation study provides an extensive evaluation of the original BILS and our extension.
-Using the optimal approach as benchmark, we show that the original BILS identifies optimal dispersion values for much larger data sets than previously analyzed, but we also highlight conditions where it fails to do so.
-A practical example shows how the extended BILS can be used for stimulus assignment in experimental psychology.
-The original BILS as well as our extension are both freely available via the open source R package anticlust.
+(2020) proposed a bicriterion anticlustering algorithm (BILS) to obtain a stimulus partitioning that simultaneously achieves high values according to two criteria: the diversity, corresponding to between-set similarity, and the dispersion, corresponding to pairwise within-set dissimilarity.
+As a heuristic, however, the BILS does not necessarily identify globally optimal criterion values.
+We present a novel exact algorithm that returns (multiple) globally optimal partitions with regard to the dispersion.
+Despite its theoretical computational hardness, it scales to rather large data sets and can be applied to solve most real-world use cases.
+In a simulation study, we evaluate several alternatives to combine our optimal approach with the BILS heuristic, which -- due to its bicriterion approach -- naturally lends itself to providing high diversity on top of an optimal dispersion.
+We also provide two further extensions to the BILS: First, the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices.
+Second, we allow that other anticlustering objectives of between-group similarity -- such as the k-means or k-plus objectives -- can be optimized instead of only the diversity.
+The original BILS as well as our extensions are freely available via the free and open source R package `anticlust`.
+Practical examples are included to illustrate their application.
 

### 2024-02-29 Simulation_2/BILS.R
adding some logs to simulation

@@ -6,7 +6,7 @@ library(anticlust)
 source("BILS_METHODS.R")
 
 RUNS_MBPI <- 10
-BATCH_SIZE_SIMULATION <- 5
+BATCH_SIZE_SIMULATION <- 300
 
 # Do not do the entire simulation in a single R session and adjust BATCH_SIZE_SIMULATION accordingly
 
@@ -32,7 +32,7 @@ for (K in 2:7) {
       } else {
         dispersion_distances <- dist(data)
       }
-      
+      cat("   Running VANILLA with", RUNS_MBPI, "repetitions\n")
       start_vanilla <- Sys.time()
       GROUPS_BILS_VANILLA <- BILS_VANILLA(
         data, 
@@ -52,14 +52,14 @@ for (K in 2:7) {
         min_dispersion_considered = dispersion_objective(dispersion_distances, GROUPS_BILS_VANILLA)
       )
       end <- Sys.time()
-      cat("   Found optimal solution!\n")
+      cat("   Exact method found optimal solution!\n")
 
       # Rerun BILS_VANILLA if the dispersion it returns is not optimal
       # do 100, 1000, 10000
       
       if (dispersion_objective(dispersion_distances, GROUPS_BILS_VANILLA) != opt$dispersion) {
         for (RUNS in c(100, 1000, 10000)) {
-          cat("Trying", RUNS, "repetitions for each heuristic\n")
+          cat("   VANILLA did not find optimal dispersion. Trying", RUNS, "repetitions for VANILLA.\n")
           start_vanilla <- Sys.time()
           GROUPS_BILS_VANILLA <- BILS_VANILLA(
             data, 
@@ -70,15 +70,18 @@ for (K in 2:7) {
           end_vanilla <- Sys.time()
           # test if optimum was found
           if (dispersion_objective(dispersion_distances, GROUPS_BILS_VANILLA) == opt$dispersion) {
+            cat("   VANILLA FOUND OPTIMAL SOLUTION\n")
             RUNS_TILL_OPTIMUM <- RUNS
             break
           } else {
             if (RUNS == 10000) {
+              cat("   Vanilla found no optimal solution\n")
               RUNS_TILL_OPTIMUM <- -1 # encode that no optimum was found even with 10000 repetitions
             }
           }
         }
       } else {
[…]

### 2024-02-29 Simulation_2/BILS_METHODS.R
this used to many repetitions

@@ -55,14 +55,14 @@ BILS_E_ALL_RESTRICTED_ALT <- function(data, init_partitions, cannot_link, disper
   distances <- as.matrix(dist(data))
   
   copy_distances <- distances
-  distances[rbind(cannot_link, t(apply(cannot_link, 1, rev)))] <- -(sum(distances) + 1) * 999999999
+  distances[rbind(cannot_link, t(apply(cannot_link, 1, rev)))] <- -(sum(distances) + 1)
   
   half_runs <- nrow(init_partitions)/2
   
   PARTITIONS <- bicriterion_anticlustering(
     distances, 
     K = init_partitions[1, ], # mostly irrelevant if `init_partitions` is passed
-    R = rep(half_runs, 2),
+    R = c(half_runs, 0),
     init_partitions = init_partitions[1:half_runs, ],
     dispersion_distances = dispersion_distances
   )
@@ -71,7 +71,7 @@ BILS_E_ALL_RESTRICTED_ALT <- function(data, init_partitions, cannot_link, disper
   PARTITIONS2 <- bicriterion_anticlustering(
     copy_distances, 
     K = PARTITIONS[1,], # 
-    R = c(nrow(PARTITIONS), nrow(init_partitions)/2),
+    R = c(nrow(PARTITIONS), half_runs),
     init_partitions = PARTITIONS,
     dispersion_distances = dispersion_distances
   )

### 2024-02-29 paper.Rmd
abstract

@@
 Brusco et al.
-(2020) proposed a bicriterion anticlustering algorithm (BILS) to obtain a stimulus partitioning that simultaneously achieves high values according to two criteria: the diversity, corresponding to between-set similarity, and the dispersion, corresponding to pairwise within-set dissimilarity.
+(2020) proposed a bicriterion algorithm (BILS) to obtain stimulus partitionings that simultaneously satisfy two anticlustering criteria: the diversity, corresponding to between-set similarity, and the dispersion, corresponding to pairwise within-set dissimilarity.
 As a heuristic, however, the BILS does not necessarily identify globally optimal criterion values.
 We present a novel exact algorithm that returns (multiple) globally optimal partitions with regard to the dispersion.
-Despite its theoretical computational hardness, it scales to rather large data sets and can be applied to solve most real-world use cases.
-In a simulation study, we evaluate several alternatives to combine our optimal approach with the BILS heuristic, which -- due to its bicriterion approach -- naturally lends itself to providing high diversity on top of an optimal dispersion.
+Despite its theoretical computational hardness, our method scales to rather large data sets and can be applied to solve most real-world use cases.
+In a simulation study, we evaluate several options to combine our optimal approach with the BILS heuristic, which -- due to its bicriterion approach -- naturally lends itself to providing high diversity on top of an optimal dispersion.
 We also provide two further extensions to the BILS: First, the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices.

### 2024-03-04 paper.Rmd
some abstract / outline

@@
 In a simulation study, we evaluate several options to combine our optimal approach with the BILS heuristic, which -- due to its bicriterion approach -- naturally lends itself to providing high diversity on top of an optimal dispersion.
-We also provide two further extensions to the BILS: First, the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices.
+We also provide two further extensions to the BILS.
+First, we allow that the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices.
 Second, we allow that other anticlustering objectives of between-group similarity -- such as the k-means or k-plus objectives -- can be optimized instead of only the diversity.
@@
   * Also needed (point of this paper!): High pairwise dissimilarity of stimuli in the same set. 
-- Simultaneous optimization of between-set similarity and pairwise within-set dissimilarity was introduced by Brusco et al.'s alogrithm, which is also available in `anticlust` since 2021 (version 0.6.0)
-- Here, we provide researchers with a practical introduction to their method and also present several extensions, which allow it to cover more use cases
-  (1) Always return optimal solution
+- Simultaneous optimization of between-set similarity and pairwise within-set dissimilarity was introduced by Brusco et al.'s algorithm, which is also available in `anticlust` since 2021 (version 0.6.0)
+- Here, we provide researchers with a practical introduction to their method and also present several extensions
+  (1) Always return optimal solution and combine optimal dispersion with maximum diversity (BILS, LCW, optimal ILP)
   (2) Use different distance matrix for dispersion and diversity (if different information is needed for both criteria, or to induce custom cannot-link constraints)
@@
   * Different distance matrices: does not change algorithm, just computation / implementation
-  * Extend towards k-means: Use squared Euclidean distance (not really extension, just different input); compute "average" diversity if group sizes are unequal (k-plus is basically the same, but do not use the additional k-plus variables for the dispersion distance matrix)
+  * Extend towards k-means: Use squared Euclidean distance + "average diversity" (k-plus is basically the same, but do not use the additional k-plus variables for the dispersion distance matrix)
 - Simulation Study
   (1) How often does vanilla BILS find optimal solutions
-  (2) Which of the three extension is best for optimizing between-group similarity while maintaining optimal dispersion?
+  (2) Which of the alternatives is best for optimizing between-group similarity while maintaining optimal dispersion?
 - Application (presenting R code)
@@
 - Discussion
+  * Future extensions: Graph coloring returns several solutions, LCW as first phase in BILS
 

### 2024-03-04 Simulation_2/Analysis_BILS.R
additional descriptives + interpretation

@@ -33,8 +33,8 @@ df$N_Category <- santoku::chop(df$N, breaks = c(20, 60, 100))
 tapply(df$time_optimal_s, list(df$K, df$N_Category), median) |> round(2)
 
 
-plot(tapply(df$time_optimal_s, santoku::chop(df$N, breaks = c(20, 40, 60, 80, 100)), median) |> round(2), xlab = "N", ylab = "Time (s)")
-plot(tapply(df$time_vanilla_s, santoku::chop(df$N, breaks = c(20, 40, 60, 80, 100)), median) |> round(2), xlab = "N", ylab = "Time (s)")
+# plot(tapply(df$time_optimal_s, santoku::chop(df$N, breaks = c(20, 40, 60, 80, 100)), median) |> round(2), xlab = "N", ylab = "Time (s)")
+# plot(tapply(df$time_vanilla_s, santoku::chop(df$N, breaks = c(20, 40, 60, 80, 100)), median) |> round(2), xlab = "N", ylab = "Time (s)")
 
 
 df$VANILLA_FOUND_OPTIMUM <- df$DISP_VANILLA == df$DISP_E_1
@@ -73,13 +73,17 @@ df |>
 # What happens in maximally restricted data sets?
 
 # What happens in maximally restricted data sets? (with VANILLA, biased)
-df$MAXIMUM_RESTRICTION <- as.numeric(df$N_DUPLICATE_PARTITIONS == 99)
+df$RESTRICTION <- "none" 
+df$RESTRICTION[df$N_DUPLICATE_PARTITIONS == 99] <- "maximal"
+df$RESTRICTION[df$N_DUPLICATE_PARTITIONS > 0 & df$N_DUPLICATE_PARTITIONS < 99] <- "some"
+df$RESTRICTION <- ordered(df$RESTRICTION, levels = c("none", "some", "maximal"))
 
-table(df$MAXIMUM_RESTRICTION)
-table(df$MAXIMUM_RESTRICTION, df$K)
+
+table(df$RESTRICTION)
+table(df$RESTRICTION, df$K)
 
 df |>
-  group_by(K, MAXIMUM_RESTRICTION) |>
+  group_by(K, RESTRICTION) |>
   summarize(
     E_1 = mean(DIV_E_1),
     E_1_ILS = mean(DIV_E_1_ILS),
@@ -88,15 +92,42 @@ df |>
[…]

### 2024-03-04 Simulation_2/Analysis_BILS.R
store best performer by degree of restriction

@@ -98,7 +98,7 @@ df |>
 
 # Just group by RESTRICTION
 
-df |>
+(results_by_restriction <- df |>
   group_by(RESTRICTION) |>
   summarize(
     E_1 = mean(DIV_E_1),
@@ -111,7 +111,10 @@ df |>
     LCW_ILS = mean(DIV_LCW_ILS),
     N = n()
   ) |> 
-  as.data.frame()  # Tibble wtf you doing with decimals
+  as.data.frame())  # Tibble wtf you doing with decimals
+
+results_only <- subset(results_by_restriction, select = -c(RESTRICTION, N))
+(best_by_restriction <- colnames(results_only)[apply(results_only, 1, which.max)])
 
 ## E_ALL_ILS best for maximally restricted data sets!!
 
@@ -125,8 +128,9 @@ df |>
 #    (i.e., E_ALL_RESTRICTED and LCW_RESTRICTED are best).
 # 2. If the problem is somewhat restricted, restricting the search space slighly
 #    is best (LCW_ILS + E_ALL_RESTRICTED_ILS are best).
-# 3. If the problem is maximally restricted, E_ALL_ILS is best. 
-# -> This case differentiation can actually be implemented in the `anticlust` interface! 
+# 3. If the problem is maximally restricted, E_ALL_ILS is best (LCW_ILS + E_ALL_RESTRICTED_ILS still good)
+# -> This case differentiation can actually be implemented in the `anticlust` interface because
+#    the degree of restriction is known!
 # (depending on the input method; give a warning when using LCW with a maximally restricted data set!) 
 
 df |>

### 2024-03-04 Simulation_2/BILS.R
I prefer K = 7 -> K = 2 in the simulation

@@ -6,11 +6,11 @@ library(anticlust)
 source("BILS_METHODS.R")
 
 TOTAL_REPETITIONS_HEURISTICS <- 100
-BATCH_SIZE_SIMULATION <- 3
+BATCH_SIZE_SIMULATION <- 50
 
 # Do not do the entire simulation in a single R session and adjust BATCH_SIZE_SIMULATION accordingly
 
-for (K in 2:7) {
+for (K in 7:2) {
   files <- list.files(paste0("./datasets/K", K, "/"), full.names = FALSE)
   # Do not replicate previous files:
   if (file.exists("results_bils.csv")) {

### 2024-03-04 Simulation_2/Analysis_BILS.R
no more need for old take aways, results much clearer now

@@ -191,13 +191,4 @@ t.test(df$DIV_E_ALL, df$DIV_E_ALL_RESTRICTED, paired = TRUE)
 t.test(df$DIV_E_ALL_RESTRICTED, df$DIV_E_ALL_RESTRICTED_ALT, paired = TRUE)
 t.test(df$DIV_E_ALL_RESTRICTED_ALT, df$DIV_LCW, paired = TRUE)
 
-# Important take aways (preliminary)
-
-# - E_ALL is better than E_1; E_ALL_RESTRICTED is usually best extension 
-# - VANILLA BILS has more difficulties finding the optimal diversity if the dispersion is optimized on the basis of a different data set
-# - Diversity in general is lower if the dispersion is optimized on the basis of different data (that makes sense I guess)
-# - The max dispersion problem can be solved in reasonable time using an open source solver (!) for rather large data sets and K
-# - Restricted method is the best extension that ensures the maximum dispersion (This would indicate that using unicriterion LCW is just as effective as BILS)
-# - Interpretation: Best for optimizing diversity while maintaining the global optimum in dispersion: restricted version if there are many init partitions that have the optimal value in dispersion. Worst: Only pass 1 init partition. Vanilla is usually good, but does not necessarily find global optimum (this can be checked using the optimal_dispersion() function). If the global optimum is needed, use restricted version; if there are no duplicates: stick with it.
-# - RESTRICTED METHOD IS NOT GOOD IF THE NUMBER OF UNIQUE PARTITIONS IS SMALL 
-# - Duplicate partitions more likely arise for smaller group sizes (i.e., larger K, and smaller N)
+# TODO: Better inference statistics (regression / ANOVA)

### 2024-03-05 Simulation_2/Analysis_BILS.R
minor stuff results

@@ -82,6 +82,30 @@ df$RESTRICTION <- ordered(df$RESTRICTION, levels = c("none", "some", "maximal"))
 table(df$RESTRICTION)
 table(df$RESTRICTION, df$K)
 
+# Just group by RESTRICTION
+
+(results_by_restriction <- df |>
+    group_by(RESTRICTION) |>
+    summarize(
+      N = n(),
+      E_1 = mean(DIV_E_1),
+      E_1_ILS = mean(DIV_E_1_ILS),
+      E_ALL = mean(DIV_E_ALL),
+      E_ALL_ILS = mean(DIV_E_ALL_ILS),
+      E_ALL_RESTRICTED = mean(DIV_E_ALL_RESTRICTED),
+      E_ALL_RESTRICTED_ILS = mean(DIV_E_ALL_RESTRICTED_ILS),
+      LCW = mean(DIV_LCW),
+      LCW_ILS = mean(DIV_LCW_ILS)
+    ) |> 
+    as.data.frame())  # Tibble wtf you doing with decimals
+
+results_only <- subset(results_by_restriction, select = -c(RESTRICTION, N))
+results_by_restriction$Best_Performer <- colnames(results_only)[apply(results_only, 1, which.max)]
+results_by_restriction[, 3:10] <- round(results_by_restriction[, 3:10], 2)
+results_by_restriction
+
+# Group by restriction and K
+
 df |>
   group_by(K, RESTRICTION) |>
   summarize(
@@ -96,37 +120,20 @@ df |>
     N = n()
   )
 
-# Just group by RESTRICTION
-
-(results_by_restriction <- df |>
-  group_by(RESTRICTION) |>
-  summarize(
-    E_1 = mean(DIV_E_1),
-    E_1_ILS = mean(DIV_E_1_ILS),
-    E_ALL = mean(DIV_E_ALL),
-    E_ALL_ILS = mean(DIV_E_ALL_ILS),
-    E_ALL_RESTRICTED = mean(DIV_E_ALL_RESTRICTED),
-    E_ALL_RESTRICTED_ILS = mean(DIV_E_ALL_RESTRICTED_ILS),
-    LCW = mean(DIV_LCW),
-    LCW_ILS = mean(DIV_LCW_ILS),
-    N = n()
-  ) |> 
-  as.data.frame())  # Tibble wtf you doing with decimals
-
-results_only <- subset(results_by_restriction, select = -c(RESTRICTION, N))
[…]

### 2024-03-05 paper.Rmd
it is now more than three approaches
different title?

@@
 ---
-title             : "An optimal anticlustering method to maximize stimulus dispersion in experiments"
-shorttitle        : "Maximum stimulus dispersion"
+title             : "Extending the Bicriterion Approach for Anticlustering: Optimal and Heuristic Approaches"
+shorttitle        : "Extending the Bicriterion Approach for Anticlustering"
 
@@
   * Optimal dispersion method: Basic algorithm + graph coloring ILP
-    * How to maintain optimal dispersion while the BILS runs? Outline three approaches!
+    * How to maintain optimal dispersion while the BILS runs?
   * Different distance matrices: does not change algorithm, just computation / implementation

### 2024-03-05 paper.Rmd
outline

@@
 - Simultaneous optimization of between-set similarity and pairwise within-set dissimilarity was introduced by Brusco et al.'s algorithm, which is also available in `anticlust` since 2021 (version 0.6.0)
-- Here, we provide researchers with a practical introduction to their method and also present several extensions
+- Outline: 
+  * Here, we present several alternatives of a bicriterion optimization that are based on optimally solving the max dispersion problem before maximizing the diversity, and we compare them to "vanilla" BILS
     (1) Always return optimal solution and combine optimal dispersion with maximum diversity (BILS, LCW, optimal ILP)
@@
     (3) Extend the BILS's capacities towards other measures of between-group similarity (k-means / k-plus)
+  * Simulation study as evaluation of the different bicriterion approaches
 - Technical introduction to "vanilla" BILS
@@
   * Optimal dispersion method: Basic algorithm + graph coloring ILP
-    * How to maintain optimal dispersion while the BILS runs?
+    * Computational Feasibility
+    * How to maintain optimal dispersion while the BILS runs? (Optimal ILP, BILS, LCW)
+    * Obtaining an (optimal) Pareto set
   * Different distance matrices: does not change algorithm, just computation / implementation

### 2024-03-05 paper.Rmd
abstract

@@
 
-    When assigning stimuli to experimental conditions, researchers strive for high similarity between stimulus sets and high pairwise dissimilarity of stimuli within sets.
-Because stimulus assignment is usually inadequate when done manually, automated algorithmic approaches are preferred.
+    Numerous applications in psychological research require that a data set is partitioned via the maximization---instead of the more common minimization---of a clustering criterion to obtain high similarity between groups and heterogeneity within groups.
+Such anticlustering may consider criteria of overall between-group similarity and/or pairwise within-group dissimilarity.
 Brusco et al.
-(2020) proposed a bicriterion algorithm (BILS) to obtain stimulus partitionings that simultaneously satisfy two anticlustering criteria: the diversity, corresponding to between-set similarity, and the dispersion, corresponding to pairwise within-set dissimilarity.
+(2020) proposed a bicriterion algorithm (BILS) to obtain partitionings that simultaneously maximize the diversity, corresponding to between-group similarity, and the dispersion, corresponding to pairwise within-group dissimilarity.
 As a heuristic, however, the BILS does not necessarily identify globally optimal criterion values.
-We present a novel exact algorithm that returns (multiple) globally optimal partitions with regard to the dispersion.
-Despite its theoretical computational hardness, our method scales to rather large data sets and can be applied to solve most real-world use cases.
-In a simulation study, we evaluate several options to combine our optimal approach with the BILS heuristic, which -- due to its bicriterion approach -- naturally lends itself to providing high diversity on top of an optimal dispersion.
-We also provide two further extensions to the BILS.
-First, we allow that the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices.
-Second, we allow that other anticlustering objectives of between-group similarity -- such as the k-means or k-plus objectives -- can be optimized instead of only the diversity.
-The original BILS as well as our extensions are freely available via the free and open source R package `anticlust`.
+We provide four extensions to the bicriterion approach for anticlustering: First, we present a novel exact algorithm that returns multiple globally optimal partitions with regard to the dispersion.
+Despite its theoretical computational hardness, our method scales to rather large data sets (e.g., $N > 5000$ for $K = 3$ groups).
+Second, by extending existing optimal and heuristic approaches (including BILS), we evaluate how to best obtain a high diversity on top of an optimal dispersion provided by our algorithm.
+Third, we allow that the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices.
+Forth, by reformulating the k-means clustering criterion as an average diversity criterion, we allow that other anticlustering objectives of between-group similarity can be optimized alongside the dispersion instead of only the diversity.
+The original BILS as well as our extensions are available via the free and open source R package `anticlust`.
 Practical examples are included to illustrate their application.

### 2024-03-06 Simulation_2/Analysis_BILS.R
this file is quite the mess, the ANOVA is not useful

@@ -44,8 +44,14 @@ tapply(df$VANILLA_FOUND_OPTIMUM, df$K, mean)
 
 table(df$N_DUPLICATE_PARTITIONS)
 
+colMedian <- function(df) {
+  apply(df, 2, median)
+}
+
+FUN <- colMedian
+
 # Display global results across all conditions
-GLOBAL_RESULTS <- data.frame(Diversity = colMeans(df[, grepl("DIV_E|DIV_LCW", colnames(df))]))
+GLOBAL_RESULTS <- data.frame(Diversity = FUN(df[, grepl("DIV_E|DIV_LCW", colnames(df))]))
 GLOBAL_RESULTS[order(GLOBAL_RESULTS$Diversity), , drop = FALSE] |> round(2)
 
 df$RESTRICTION <- "none" 
@@ -56,8 +62,6 @@ df$RESTRICTION <- ordered(df$RESTRICTION, levels = c("none", "some", "maximal"))
 
 table(df$RESTRICTION)
 table(df$RESTRICTION, df$K)
-
-# (TODO: I really need the data in long format for better descriptives...)
 
 # There seems to be a clear order in the results: 3 groups of performers: 
 # E_1 and E_1_ILS are worst (E_1 is particularly bad for some reason)
@@ -92,7 +96,7 @@ m1 <- aov_ez(
 m1
 
 aov_tab <- m1$anova_table
-aov_tab <- aov_tab[aov_tab$pes > .004 & aov_tab$`Pr(>F)` < .001, ]
+aov_tab <- aov_tab[aov_tab$pes > .0035 & aov_tab$`Pr(>F)` < .001, ]
 nice(aov_tab[order(aov_tab$pes, decreasing = TRUE), ])
 
 pairs(emmeans(m1, specs = "Method"))
@@ -113,8 +117,30 @@ pairs(emmeans(m1, ~ Method | N_centered + K_centered, cov.reduce = function(x) q
 # Interactions with N, M and K are all driven by E_1... it gets increasingly bad when 
 # the problem size increases, and this seems to multiply with the different factors (N, M, K).
 
+# Do not use E_1 in ANOVA, it seems to produce several interactions ...
+# Some inference statistics ...
+no_e1 <- dfl |> filter(!grepl("E_1", Method))
+m2 <- aov_ez(
+  id = "file",
+  dv = "Log_div",
[…]

### 2024-03-06 Simulation_2/Analysis_BILS_new.R
try a new cleaner analysis

@@ -1,0 +1,128 @@
+
+library(tidyr)
+library(dplyr)
+library(santoku)
+
+df <- read.csv("results_bils.csv", sep = ";")
+
+# Two functions for printing descriptive results
+
+# Function to add the difference between best and worst method
+# primarily shows in which conditions E_1 is really bad
+add_discrepancy <- function(results, colpattern = "DIV") {
+  results_only <- results[, grepl(colpattern, colnames(results))]
+  MIN <- apply(results_only, 1, min)
+  MAX <- apply(results_only, 1, max)
+  results$Discrepancy <- MAX - MIN
+  results
+}
+
+add_best_method <- function(results, colpattern = "DIV") {
+  results_only <- results[, grepl(colpattern, colnames(results))]
+  results$Best_Performer <- colnames(results_only)[apply(results_only, 1, which.max)]
+  results
+}
+
+df$VANILLA_FOUND_OPTIMUM <- df$DISP_VANILLA == df$DISP_E_1
+df$RESTRICTION <- "none" 
+df$RESTRICTION[df$N_DUPLICATE_PARTITIONS == 99] <- "maximal"
+df$RESTRICTION[df$N_DUPLICATE_PARTITIONS > 0 & df$N_DUPLICATE_PARTITIONS < 99] <- "some"
+df$RESTRICTION <- ordered(df$RESTRICTION, levels = c("none", "some", "maximal"))
+
+sort(colMeans(df[df$VANILLA_FOUND_OPTIMUM == 1, grepl("DIV", colnames(df))])) # global results including VANILLA
+
+# Did Vanilla find optimum? Depending on Restriction
+prop.table(table(df$RESTRICTION, c("NO OPTIMUM", "FOUND OPTIMUM")[df$VANILLA_FOUND_OPTIMUM+1]), margin = 2) |> round(2)
+
+dfw <- select(df, file, N, M, K, RESTRICTION, starts_with("DIV"), -DIV_VANILLA)
+
+dfw$DIV_MAX <- apply(dfw[, grepl("DIV", colnames(dfw))], 1, max)
+
+almost_has_maximum_div <- ((dfw[, grepl("DIV", colnames(dfw))] / dfw$DIV_MAX) > .999) * 1 # maybe good quantity?
+almost_has_maximum_div <- almost_has_maximum_div[, colnames(almost_has_maximum_div) != "DIV_MAX"]
+colnames(almost_has_maximum_div) <- paste0("MAX_", colnames(almost_has_maximum_div))
[…]

### 2024-03-06 paper.Rmd
word

@@
 As a heuristic, however, the BILS does not necessarily identify globally optimal criterion values.
-We provide four extensions to the bicriterion approach for anticlustering: First, we present a novel exact algorithm that returns multiple globally optimal partitions with regard to the dispersion.
+Here we provide four extensions to the bicriterion approach for anticlustering: First, we present a novel exact algorithm that returns multiple globally optimal partitions with regard to the dispersion.
 Despite its theoretical computational hardness, our method scales to rather large data sets (e.g., $N > 5000$ for $K = 3$ groups).

### 2024-03-06 paper.Rmd
abstract

--- 2024-03-05 paper.Rmd
@@
 
-    Numerous applications in psychological research require that a data set is partitioned via the maximization---instead of the more common minimization---of a clustering criterion to obtain high similarity between groups and heterogeneity within groups.
-Such anticlustering may consider criteria of overall between-group similarity and/or pairwise within-group dissimilarity.
+    Numerous applications in psychological research require that a data set is partitioned via the maximization---instead of the more common minimization---of a clustering criterion.
+Such anticlustering seeks for high similarity between groups (i.e., maximum diversity) and/or high pairwise dissimilarity within groups (i.e., maximum dispersion).
 Brusco et al.
-(2020) proposed a bicriterion algorithm (BILS) to obtain partitionings that simultaneously maximize the diversity, corresponding to between-group similarity, and the dispersion, corresponding to pairwise within-group dissimilarity.
-As a heuristic, however, the BILS does not necessarily identify globally optimal criterion values.
-We provide four extensions to the bicriterion approach for anticlustering: First, we present a novel exact algorithm that returns multiple globally optimal partitions with regard to the dispersion.
+(2020) proposed a bicriterion algorithm (BILS) that simultaneously seeks for maximum diversity and dispersion, introducing the bicriterion approach for anticlustering.
+Here, we advance the bicriterion approach in several ways.
+As two minor extensions, we allow that the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices, and that other anticlustering objectives of between-group similarity can be optimized instead of only the diversity.
+As our main contribution, we present an exact algorithm that returns multiple globally optimal partitions with regard to the dispersion.
 Despite its theoretical computational hardness, our method scales to rather large data sets (e.g., $N > 5000$ for $K = 3$ groups).
-Second, by extending existing optimal and heuristic approaches (including BILS), we evaluate how to best obtain a high diversity on top of an optimal dispersion provided by our algorithm.
-Third, we allow that the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices.
-Forth, by reformulating the k-means clustering criterion as an average diversity criterion, we allow that other anticlustering objectives of between-group similarity can be optimized alongside the dispersion instead of only the diversity.
+By extending existing optimal and heuristic approaches (including BILS), we evaluate how to best obtain high diversity on top of an optimal dispersion.
+Our analysis reveals an important factor determining the difficulty of the bicriterion anticlustering problem: the level of restriction induced by the maximum dispersion constraint.
+Depending on the level of restriction, different algorithmic choices are recommended.
 The original BILS as well as our extensions are available via the free and open source R package `anticlust`.

### 2024-03-06 paper.Rmd
some on abstract, probably does not make it better

@@
 Here, we advance the bicriterion approach in several ways.
-As two minor extensions, we allow that the dispersion and the diversity can be optimized on the basis of different dissimilarity matrices, and that other anticlustering objectives of between-group similarity can be optimized instead of only the diversity.
-As our main contribution, we present an exact algorithm that returns multiple globally optimal partitions with regard to the dispersion.
-Despite its theoretical computational hardness, our method scales to rather large data sets (e.g., $N > 5000$ for $K = 3$ groups).
-By extending existing optimal and heuristic approaches (including BILS), we evaluate how to best obtain high diversity on top of an optimal dispersion.
+As a heuristic, the BILS does not necessarily identify globally optimal criterion values.
+We combine a novel exact algorithm for the dispersion with the BILS heuristic to obtain high diversity on top of an optimal dispersion.
+We present several adaptations of the BILS and evaluate which is best suited to accomplish this goal.
+Despite its theoretical computational hardness, our optimal method scales to large data sets (e.g., $N > 5000$ for $K = 3$ groups).
 Our analysis reveals an important factor determining the difficulty of the bicriterion anticlustering problem: the level of restriction induced by the maximum dispersion constraint.
 Depending on the level of restriction, different algorithmic choices are recommended.
+In particular, the more constrained the problem becomes, the more important is the iterated local search (ILS), the second phase of the BILS.
 The original BILS as well as our extensions are available via the free and open source R package `anticlust`.

### 2024-03-07 paper.Rmd
abstract

@@
 Here, we advance the bicriterion approach in several ways.
-As a heuristic, the BILS does not necessarily identify globally optimal criterion values.
-We combine a novel exact algorithm for the dispersion with the BILS heuristic to obtain high diversity on top of an optimal dispersion.
-We present several adaptations of the BILS and evaluate which is best suited to accomplish this goal.
-Despite its theoretical computational hardness, our optimal method scales to large data sets (e.g., $N > 5000$ for $K = 3$ groups).
-Our analysis reveals an important factor determining the difficulty of the bicriterion anticlustering problem: the level of restriction induced by the maximum dispersion constraint.
-Depending on the level of restriction, different algorithmic choices are recommended.
-In particular, the more constrained the problem becomes, the more important is the iterated local search (ILS), the second phase of the BILS.
-The original BILS as well as our extensions are available via the free and open source R package `anticlust`.
+As our main contribution, we combine a novel exact algorithm for maximum dispersion with the BILS heuristic to obtain high diversity on top of an optimal dispersion.
+Despite its theoretical computational hardness, our optimal method scales to rather large data sets.
+It generates multiple partitions that the BILS can use as initializations during its local search, thus guaranteeing that the BILS also returns a partition having optimal dispersion.
+In a simulation study, we compared several adaptations of the BILS.
+To obtain high diversity, initializing the BILS with multiple optimal partitions is preferable to using only one.
+Moreover, in data sets where the constraint of optimal dispersion severely restricts the search for feasible partitions, the iterated local search phase of the BILS is crucial.
+While the original BILS oftentimes finds optimal criterion values for the dispersion, our simulation illustrates conditions when it does not.
+The optimal dispersion algorithm and the BILS including our extensions are available via the free and open source R package `anticlust`.
 Practical examples are included to illustrate their application.

### 2024-03-07 Simulation_2/ANALYSIS_BILS_VANILLA.R
analysis

@@ -1,5 +1,17 @@
 # ANALYSIS_BILS_VANILLA.R
 
+df <- read.csv("results_bils.csv", sep = ";")
+df$VANILLA_FOUND_OPTIMUM <- df$DISP_VANILLA == df$DISP_E_1
+df$RESTRICTION <- "none" 
+df$RESTRICTION[df$N_DUPLICATE_PARTITIONS == 99] <- "maximal"
+df$RESTRICTION[df$N_DUPLICATE_PARTITIONS > 0 & df$N_DUPLICATE_PARTITIONS < 99] <- "some"
+df$RESTRICTION <- factor(df$RESTRICTION, levels = c("none", "some", "maximal"))
+
+m1 <- glm(VANILLA_FOUND_OPTIMUM ~ N + K + M + RESTRICTION, data = df, family = binomial(link = "logit"))
+summary(m1)
+m2 <- glm(VANILLA_FOUND_OPTIMUM ~ N + K + M + relevel(RESTRICTION, ref = 2), data = df, family = binomial(link = "logit"))
+summary(m2)
+
 
 table(df$RUNS_BILS_VANILLA)
 table(df$RUNS_BILS_VANILLA, df$K)

### 2024-03-07 Simulation_2/Analysis_BILS_new.R

@@ -30,6 +30,22 @@ df$RESTRICTION[df$N_DUPLICATE_PARTITIONS > 0 & df$N_DUPLICATE_PARTITIONS < 99] <
 df$RESTRICTION <- ordered(df$RESTRICTION, levels = c("none", "some", "maximal"))
 
 sort(colMeans(df[df$VANILLA_FOUND_OPTIMUM == 1, grepl("DIV", colnames(df))])) # global results including VANILLA
+# THE ALL_ILS adaptations perform better than VANILLA even when VANILLA found optimum (which is a biased analysis!):
+t.test(
+  df$DIV_VANILLA[df$VANILLA_FOUND_OPTIMUM == 1], 
+  df$DIV_E_ALL_ILS[df$VANILLA_FOUND_OPTIMUM == 1], 
+  paired = TRUE
+)
+t.test(
+  df$DIV_VANILLA[df$VANILLA_FOUND_OPTIMUM == 1], 
+  df$DIV_E_ALL_RESTRICTED_ILS[df$VANILLA_FOUND_OPTIMUM == 1], 
+  paired = TRUE
+)
+t.test(
+  df$DIV_VANILLA[df$VANILLA_FOUND_OPTIMUM == 1], 
+  df$DIV_LCW_ILS[df$VANILLA_FOUND_OPTIMUM == 1], 
+  paired = TRUE
+)
 
 # Did Vanilla find optimum? Depending on Restriction
 prop.table(table(df$RESTRICTION, c("NO OPTIMUM", "FOUND OPTIMUM")[df$VANILLA_FOUND_OPTIMUM+1]), margin = 2) |> round(2)

### 2024-03-07 Simulation_2/Analysis_BILS_new.R
interpretation

@@ -130,6 +130,10 @@ dflD |>
   add_discrepancy() |> 
   add_best_method()
 
+# For no restriction: Small difference between E_ALL variants and LCW variants, E_1 bad. Slight advantage for non-ILS variants.
+# When restriction increases: ILS becomes more important.
+# For maximum restriction, ILS is crucial. Interestingly: E_1 is good for small data sets with maximum restriction.
+
 # E_1 may actually be good for small data sets + maximum restriction
 # (but I would still prefer a different method due to the danger of "serious fails")
 

### 2024-03-07 paper.Rmd
minor abstract

@@
     Numerous applications in psychological research require that a data set is partitioned via the maximization---instead of the more common minimization---of a clustering criterion.
-Such anticlustering seeks for high similarity between groups (i.e., maximum diversity) and/or high pairwise dissimilarity within groups (i.e., maximum dispersion).
+Such anticlustering seeks for high similarity between groups (i.e., maximum diversity) or high pairwise dissimilarity within groups (i.e., maximum dispersion).
 Brusco et al.
@@
 To obtain high diversity, initializing the BILS with multiple optimal partitions is preferable to using only one.
-Moreover, in data sets where the constraint of optimal dispersion severely restricts the search for feasible partitions, the iterated local search phase of the BILS is crucial.
-While the original BILS oftentimes finds optimal criterion values for the dispersion, our simulation illustrates conditions when it does not.
+Moreover, in data sets where the constraint of optimal dispersion severely restricts the search for feasible partitions, the iterated local search phase of the BILS is crucial to obtain high diversity.
+While the original BILS oftentimes finds optimal dispersion values, our simulation illustrates conditions when it does not.
 The optimal dispersion algorithm and the BILS including our extensions are available via the free and open source R package `anticlust`.

### 2024-03-07 paper.Rmd
why not

@@
 Brusco et al.
-(2020) proposed a bicriterion algorithm (BILS) that simultaneously seeks for maximum diversity and dispersion, introducing the bicriterion approach for anticlustering.
+(2020) proposed a powerful bicriterion algorithm (BILS) that simultaneously seeks for maximum diversity and dispersion, introducing the bicriterion approach for anticlustering.
 Here, we advance the bicriterion approach in several ways.
@@
 Despite its theoretical computational hardness, our optimal method scales to rather large data sets.
-It generates multiple partitions that the BILS can use as initializations during its local search, thus guaranteeing that the BILS also returns a partition having optimal dispersion.
+It generates multiple partitions that we use as initializations of the BILS's local search, thus guaranteeing that the BILS also returns a partition having optimal dispersion.
 In a simulation study, we compared several adaptations of the BILS.

### 2024-03-11 paper.Rmd
some text

@@
 Despite its theoretical computational hardness, our optimal method scales to rather large data sets.
-It generates multiple partitions that we use as initializations of the BILS's local search, thus guaranteeing that the BILS also returns a partition having optimal dispersion.
+It generates multiple partitions that we use as initializations of the BILS's local search, thus guaranteeing that the BILS also returns a partition with optimal dispersion.
 In a simulation study, we compared several adaptations of the BILS.
 To obtain high diversity, initializing the BILS with multiple optimal partitions is preferable to using only one.
-Moreover, in data sets where the constraint of optimal dispersion severely restricts the search for feasible partitions, the iterated local search phase of the BILS is crucial to obtain high diversity.
+Moreover, in data sets where maintaining optimal dispersion severely restricts the search for feasible partitions, the iterated local search phase of the BILS is crucial to obtain high diversity.
 While the original BILS oftentimes finds optimal dispersion values, our simulation illustrates conditions when it does not.
@@
 
-## Outline
+Anticlustering problems arise in many research fields including psychology [@brusco2019; @schaper2023], test assessment [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network transmission [@mohebi2022], and operations research [e.g.
+@gallego2013; @gliesch2021].
+Anticlustering seeks for similarity between groups and heterogeinity within groups by maximizing a clustering objective criterion.
+Thus, anticlustering is the logical and mathematical opposite of classical clustering, which conducts partitioning via minimization of the same criteria [@steinley2006; @brusco2019; @spath1986]
 
-- Start with use case: Stimulus assignment in within-subjects experiments (and stick with it throughout the paper, only briefly mention other use cases)
-  * Needed: high similarity between stimulus set to reduce variance in responses attributed to other factors than the experimental manipulation.
-This is possible using `anticlust`, which has been applied many times since its introduction in @papenberg2020 (cite references)
-  * Also needed (point of this paper!): High pairwise dissimilarity of stimuli in the same set. 
-- Simultaneous optimization of between-set similarity and pairwise within-set dissimilarity was introduced by Brusco et al.'s algorithm, which is also available in `anticlust` since 2021 (version 0.6.0)
-- Outline: 
-  * Here, we present several alternatives of a bicriterion optimization that are based on optimally solving the max dispersion problem before maximizing the diversity, and we compare them to "vanilla" BILS
-    (1) Always return optimal solution and combine optimal dispersion with maximum diversity (BILS, LCW, optimal ILP)
-    (2) Use different distance matrix for dispersion and diversity (if different information is needed for both criteria, or to induce custom cannot-link constraints)
-    (3) Extend the BILS's capacities towards other measures of between-group similarity (k-means / k-plus)
-  * Simulation study as evaluation of the different bicriterion approaches
-- Technical introduction to "vanilla" BILS
-- Technical introduction to our extensions
-  * Optimal dispersion method: Basic algorithm + graph coloring ILP
-    * Computational Feasibility
-    * How to maintain optimal dispersion while the BILS runs? (Optimal ILP, BILS, LCW)
-    * Obtaining an (optimal) Pareto set
-  * Different distance matrices: does not change algorithm, just computation / implementation
-  * Extend towards k-means: Use squared Euclidean distance + "average diversity" (k-plus is basically the same, but do not use the additional k-plus variables for the dispersion distance matrix)
-- Simulation Study
-  (1) How often does vanilla BILS find optimal solutions
-  (2) Which of the alternatives is best for optimizing between-group similarity while maintaining optimal dispersion?
-- Application (presenting R code)
-  * Optimize k-plus criterion using norming data; dispersion is based on pairwise orthographic dissimilarity between words
-- Discussion
-  * Future extensions: Graph coloring returns several solutions, LCW as first phase in BILS
+More formally, anticlustering is used to partition $N$ objects into $K$ groups $C = (C_1, \ldots, C_K)$, where $C_k$ is the set of $n_k$ objects in the $k^{th}$ group, and $K$ and $n_k$ are predetermined by the application.
+In many anticlustering applications, equal-sized groups are required, i.e., $n_k = \frac{N}{K}$ [@schulz2022; @gallego2013].
+The partitioning is characterized by assignment to exhaustive and mutually exclusive subgroups:
 
-## Introduction
+\begin{align}
+\bigcup\limits_{k = 1}^{K} C_k = X \label{formalization:1} \\
+C_j \cap C_k = \emptyset, \; \forall j, k \in \{1, \ldots, K\}, \; j \ne k\label{formalization:2} \\
+\vert C_k \vert = n_k, \; \forall k \in \{1, \ldots, K\} \label{formalization:3}
+\end{align}
+
+We assume that a data matrix $\mathbf{X}_{N \times M} = \{x_{ij}\}_{N \times M}$ is available where the $M$ columns contain numeric measurements on the $N$ objects.
+The partitioning is obtained by maximizing a clustering criterion $f_i$.
+For example, *k*-means anticlustering aims to maximize between-group similarity by maximizing the error sum of squares ($\mathit{SSE}$), i.e., the sum the squared distances between each object and its cluster center [@spath1986; @papenberg2024]:
+
+$$
+f_1(C) = 
+  \sum\limits_{j=1}^{M} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (x_{ij} - \overline{x}_j ^{(k)})^2
+$$
+
+The *diversity* is another important anticlustering criterion representing between-group similarity.
+It is computed as the within-group sum of pairwise dissimilarities [@brusco2019; @papenberg2020; @feo1990]:
+
+$$
+f_2(C) = \sum_{i < j) \in C_k} d_{ij}
+$$
+
+The diversity can for example be computed by first converting the data matrix $\mathbf{X}_{N \times M}$ into a matrix of pairwise Euclidean distances $\mathbf{D}_{N \times N} = \{d_{ij}\}_{N \times N}$, where
+
+$$
+d_{ij} = \sqrt{\sum\limits_{m = 1}^{M}(x_{im} - x_{jm})^2}
+$$
+
+However, note that most anticlustering algorithms that use a dissimilarity matrix as input are agnostic as to how it was was generated [@brusco2019; cf. @schulz2022].
+
+Unlike $f_1$ and $f_2$, the *dispersion* criterion is not a sum, but it is defined as the minimum dissimilarity across any two objects that are part of the same group [@brusco2019; @fernandez2013]:
+
+$$
+f_3(C) = \textrm{min}_{(i < j) \in C_k} d_{ij}
+$$
+
+The dispersion does not represent between-group similarity, but instead it is measure of the worst-case pairwise dissimilarity.
+Maximizing the dispersion ensures that any two objects in the same group are as dissimilar from each other as possible.
+
+@brusco2019 outlined the importance of simultaneously considering multiple criteria when tackling anticlustering problems and introduced the bicriterion optimization model for anticlustering.
+In particular, they argued that anticlustering algorithms should incorporate an objective of between-group similarity---such as the diversity---and pairwise within-group dissimilarity---i.e., the dispersion---at the same time.
+
+## Algorithms for anticlustering
+
+Heuristics dominate (local maximum search @spath1986; @weitz1998, multiple repetitions robust for k-means @steinley2007, ILS @brusco2017, VARIABLE NEIGHBORHOOD SEARCH @urovsevic2014; tabu search @gallego2013; multiple improvement phases based on initial local maximum search @yang2022).
+Exact algorithms for diversity [@papenberg2020; @schulz2022] and for the dispersion [@fernandez2013; @gliesch2021; @brucker1978].
+Exact algorithms for diversity perform much worse than for the reversed clustering problems, where many hundred elements can often be processed.
+Maximum diversity can be achieved with much better success.
+Our novel contribution is to combine optimal maximum dispersion with high diversity.
+We present a novel algorithm for maximum diversity that is similar to the approaches by @fernandez2013 and @gliesch2021 and is based on the logic that was already introduced by @brucker1978.
+It scales to hundreds or even thousands of data points and therefore can be used to solve most real-world applications we have in mind.
+We then proceed to discuss how on top of an optimal dispersion, we can implement high diversity.
+A logical next step is to obtain optimal maximum diversity as well, which is however only limited to rather small data sets, as we show.
+Due to its bicriterion nature, the BILS by @brusco2019 is naturally suited to maintain optimum dispersion while optimizing diversity.
+We discuss several adaptations of the BILS, which are compared in a simulation study.
 

### 2024-03-11 paper.Rmd
intro

@@
 We assume that a data matrix $\mathbf{X}_{N \times M} = \{x_{ij}\}_{N \times M}$ is available where the $M$ columns contain numeric measurements on the $N$ objects.
-The partitioning is obtained by maximizing a clustering criterion $f_i$.
-For example, *k*-means anticlustering aims to maximize between-group similarity by maximizing the error sum of squares ($\mathit{SSE}$), i.e., the sum the squared distances between each object and its cluster center [@spath1986; @papenberg2024]:
+The partitioning is obtained by maximizing a clustering criterion $f_i(C)$.
+For example, *k*-means anticlustering aims to maximize between-group similarity by maximizing the error sum of squares, i.e., the sum the squared distances between each object and its cluster center [@spath1986; @papenberg2024]:
 
@@
 
-The *diversity* is another important anticlustering criterion representing between-group similarity.
+The $k^{th}$ cluster center $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ contains the mean values of the $M$ variables computed across all observations belonging to the  $k^{th}$ group:
+
+$$
+\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} x_{ij}
+$$
+
+Maximizing $f_1$ ensures that the cluster centers, i.e., the mean values of the $M$ measurements, are as close to each other as possible.
+The *k*-plus criterion is an extension of the $k$-means criterion for anticlustering that can be implemented by adding fictitious to the data matrix $\mathbf{X}_{N \times M}$ [@papenberg2024].
+*K*-plus anticlustering is used to equate not only means between clusters, but higher order moments such as the variance or skewness as well.
+Because the approach can be reduced to an augmentation of the data set, *k*-plus anticlustering leaves the original *k*-means criterion $f_1$ unchanged.
+
+Other anticlustering criteria usually use as input a matrix $\mathbf{D}_{N \times N} = \{d_{ij}\}_{N \times N}$ of pairwise dissimilarity measurements.
+The *diversity* criterion is the most important criterion of this kind.
 It is computed as the within-group sum of pairwise dissimilarities [@brusco2019; @papenberg2020; @feo1990]:
@@
 $$
-f_2(C) = \sum_{i < j) \in C_k} d_{ij}
+f_2(C) = \sum_{(i < j) \in C_k} d_{ij}
 $$
 
-The diversity can for example be computed by first converting the data matrix $\mathbf{X}_{N \times M}$ into a matrix of pairwise Euclidean distances $\mathbf{D}_{N \times N} = \{d_{ij}\}_{N \times N}$, where
+Maximizing the diversity simultaneously minimizes the sum of distances between objects not in the same group.
+Therefore, the diversity represents within-group heterogeneity as well as between-group similarity [@feo1990].
+Being a sum instead of an average, however, it no longer adequately captures between-group similarity if group sizes are not equal.
+The diversity is oftentimes obtained by first converting the data matrix $\mathbf{X}_{N \times M}$ into a matrix of pairwise Euclidean distances.
+However, most anticlustering algorithms that use a dissimilarity matrix as input are agnostic as to how it was was generated [@brusco2019; cf.
+@schulz2022].
 
-$$
-d_{ij} = \sqrt{\sum\limits_{m = 1}^{M}(x_{im} - x_{jm})^2}
-$$
-
-However, note that most anticlustering algorithms that use a dissimilarity matrix as input are agnostic as to how it was was generated [@brusco2019; cf. @schulz2022].
-
-Unlike $f_1$ and $f_2$, the *dispersion* criterion is not a sum, but it is defined as the minimum dissimilarity across any two objects that are part of the same group [@brusco2019; @fernandez2013]:
+Unlike $f_1$ and $f_2$, the *dispersion* criterion does not represent between-group similarity, but is a pure measure of within-group heterogeneity.
+It is defined as the minimum dissimilarity across any two objects that are part of the same group [@brusco2019; @fernandez2013]:
 
@@
 
-The dispersion does not represent between-group similarity, but instead it is measure of the worst-case pairwise dissimilarity.
-Maximizing the dispersion ensures that any two objects in the same group are as dissimilar from each other as possible.
+The dispersion is a measure of the worst-case pairwise dissimilarity. Maximizing the dispersion ensures that any two objects in the same group are as dissimilar from each other as possible.
 
-@brusco2019 outlined the importance of simultaneously considering multiple criteria when tackling anticlustering problems and introduced the bicriterion optimization model for anticlustering.
+@brusco2019 outlined the importance of simultaneously considering multiple criteria when tackling anticlustering problems.
 In particular, they argued that anticlustering algorithms should incorporate an objective of between-group similarity---such as the diversity---and pairwise within-group dissimilarity---i.e., the dispersion---at the same time.

### 2024-03-11 paper.Rmd
k-plus

@@
 
-Anticlustering problems arise in many research fields including psychology [@brusco2019; @schaper2023], test assessment [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network transmission [@mohebi2022], and operations research [e.g.
+Partitioning data sets via anticlustering has uses in many research fields including psychology [@brusco2019; @schaper2023], test assessment [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network transmission [@mohebi2022], and operations research [e.g.
 @gallego2013; @gliesch2021].
 Anticlustering seeks for similarity between groups and heterogeinity within groups by maximizing a clustering objective criterion.
-Thus, anticlustering is the logical and mathematical opposite of classical clustering, which conducts partitioning via minimization of the same criteria [@steinley2006; @brusco2019; @spath1986]
+Thus, anticlustering is the logical and mathematical opposite of classical clustering, which partitions data sets via minimization of the same criteria [@steinley2006; @brusco2019; @spath1986]
 
@@
 We assume that a data matrix $\mathbf{X}_{N \times M} = \{x_{ij}\}_{N \times M}$ is available where the $M$ columns contain numeric measurements on the $N$ objects.
-The partitioning is obtained by maximizing a clustering criterion $f_i(C)$.
+The partitioning is obtained by maximizing a clustering criterion $f_i(C, \mathbf{X})$.
 For example, *k*-means anticlustering aims to maximize between-group similarity by maximizing the error sum of squares, i.e., the sum the squared distances between each object and its cluster center [@spath1986; @papenberg2024]:
@@
 $$
-f_1(C) = 
+f_1(C, \mathbf{X}_{N \times M}) = 
   \sum\limits_{j=1}^{M} 
@@
 Because the approach can be reduced to an augmentation of the data set, *k*-plus anticlustering leaves the original *k*-means criterion $f_1$ unchanged.
+For example, to equate means and variances, we define a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$.
+Then, the *k*-plus criterion is computed as
+
+$$
+f_1^*(C, \mathbf{(X | Y)}_{N \times 2M}) = 
+  \sum\limits_{j=1}^{M} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (x_{ij} - \overline{x}_j ^{(k)})^2 + (y_{ij} - \overline{y}_j ^{(k)})^2
+$$
 

### 2024-03-11 paper.Rmd
text

--- 2024-03-11 paper.Rmd
@@
 
-Anticlustering problems arise in many research fields including psychology [@brusco2019; @schaper2023], test assessment [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network transmission [@mohebi2022], and operations research [e.g.
+Partitioning data sets via anticlustering has uses in many research fields including psychology [@brusco2019; @schaper2023], test assessment [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network systems [@mohebi2022], and operations research [e.g.
 @gallego2013; @gliesch2021].
-Anticlustering seeks for similarity between groups and heterogeinity within groups by maximizing a clustering objective criterion.
-Thus, anticlustering is the logical and mathematical opposite of classical clustering, which conducts partitioning via minimization of the same criteria [@steinley2006; @brusco2019; @spath1986]
+Anticlustering seeks for similarity between groups and heterogeneity within groups by maximizing a clustering objective criterion.
+Thus, anticlustering is the logical and mathematical opposite of classical clustering, which partitions data sets via minimization of the same criteria [@steinley2006; @brusco2019; @spath1986].
 
-More formally, anticlustering is used to partition $N$ objects into $K$ groups $C = (C_1, \ldots, C_K)$, where $C_k$ is the set of $n_k$ objects in the $k^{th}$ group, and $K$ and $n_k$ are predetermined by the application.
+Formally, anticlustering is used to partition $N$ objects into $K$ groups $C = (C_1, \ldots, C_K)$, where $C_k$ is the set of $n_k$ objects in the $k^{th}$ group, and $K$ and $n_k$ are predetermined by the application.
 In many anticlustering applications, equal-sized groups are required, i.e., $n_k = \frac{N}{K}$ [@schulz2022; @gallego2013].
@@
 
-We assume that a data matrix $\mathbf{X}_{N \times M} = \{x_{ij}\}_{N \times M}$ is available where the $M$ columns contain numeric measurements on the $N$ objects.
-The partitioning is obtained by maximizing a clustering criterion $f_i(C)$.
+We assume that a data matrix $\mathbf{Z}$ is used as input for computing the anticlustering criterion.
+$\mathbf{Z}$ is either a feature matrix $\mathbf{X}_{N \times M} = \{x_{ij}\}_{N \times M}$ where the $M$ columns contain numeric measurements on the $N$ objects, or a symmetric matrix $\mathbf{D}_{N \times N} = \{d_{ij}\}_{N \times N}$ of pairwise dissimilarity measurements.
+The partitioning is obtained by maximizing a clustering criterion $f_i(C, \mathbf{Z})$.
 For example, *k*-means anticlustering aims to maximize between-group similarity by maximizing the error sum of squares, i.e., the sum the squared distances between each object and its cluster center [@spath1986; @papenberg2024]:
@@
 $$
-f_1(C) = 
+f_1(C, \mathbf{X}_{N \times M}) =
   \sum\limits_{j=1}^{M} 
@@
 
-Maximizing $f_1$ ensures that the cluster centers, i.e., the mean values of the $M$ measurements, are as close to each other as possible.
+Maximizing $f_1$ ensures that the cluster centers, i.e., the mean values of the $M$ measurements, are as close to each other as possible [@spath1986].
 The *k*-plus criterion is an extension of the $k$-means criterion for anticlustering that can be implemented by adding fictitious to the data matrix $\mathbf{X}_{N \times M}$ [@papenberg2024].
@@
 Because the approach can be reduced to an augmentation of the data set, *k*-plus anticlustering leaves the original *k*-means criterion $f_1$ unchanged.
-
-Other anticlustering criteria usually use as input a matrix $\mathbf{D}_{N \times N} = \{d_{ij}\}_{N \times N}$ of pairwise dissimilarity measurements.
-The *diversity* criterion is the most important criterion of this kind.
-It is computed as the within-group sum of pairwise dissimilarities [@brusco2019; @papenberg2020; @feo1990]:
+For example, to equate means and variances, we define a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$ and augment the original matrix with it.
+Then, the *k*-plus criterion is computed as
 
 $$
-f_2(C) = \sum_{(i < j) \in C_k} d_{ij}
+f_2(C, \mathbf{(X | Y)_{N \times 2M}}) =
+  \sum\limits_{j=1}^{M} 
+  \sum\limits_{k=1}^{K} 
+  \sum\limits_{i \in C_k}
+  (x_{ij} - \overline{x}_j ^{(k)})^2 + (y_{ij} - \overline{y}_j ^{(k)})^2
+$$
+
+Another important anticlustering criterion is the *diversity*, which uses a dissimilarity matrix $\mathbf{D}_{N \times N}$ as input.
+It is computed as the within-group sum of pairwise dissimilarities across all clusters [@brusco2019; @papenberg2020; @feo1990]:
+
+$$
+f_3(C, \mathbf{D}_{N \times N}) = \sum_{(i < j) \in C_k} d_{ij}
 $$
@@
 @schulz2022].
+If $\mathbf{D}_{N \times N}$ is computed via the squared Euclidean distances, we note an important equivalence between the k-means criterion and the diversity.
+The k-means criterion can also be expressed using the pairwise squared Euclidean distances between data points (Brusco, 2006; Späth, 1980).
+Based on $\mathbf{D}_{N \times N}$, the k-means criterion is then computed as
 
-Unlike $f_1$ and $f_2$, the *dispersion* criterion does not represent between-group similarity, but is a pure measure of within-group heterogeneity.
+$$
+f_1(C, \mathbf{X}_{N \times M}) =
+f_4(C, \mathbf{D}_{N \times N}) = \sum\limits_{k = 1}^{K}\left( \frac{1}{n_k} \sum_{(i<j)\in C_k} d_{ij} \right)
+$$
+
+That is, "the sum of squared distances of a collection of points from the centroid associated with those points is equal to the sum of the pairwise squared distances between those points divided by the size (number of objects) of the collection" (Brusco, 2006, p.
+350).
+
+An important special case of anticlustering applications is that all groups are equal-sized, i.e., $n_k = \frac{N}{K}$ ($k = 1, \dots K$). Here, we obtain
+
+$$
+\mathit{SSE} = \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
+$$
+
+In this case, the normalizing factor $\frac{N}{K}$ can be ignored when maximizing $f_1$. Thus, the criterion reduces to the sum of the (squared) distances between elements in the same cluster:
+
+$$
+\mathit{SSE_{Diversity}} = \sum\limits_{k = 1}^{K}\sum_{(i<j)\in C_k} d_{ij}
+$$
+
+This term corresponds to the standard diversity $f_2$, with the distances $d_{ij}$ representing squared Euclidean distances.
+We denote $f_4$ as the *average diversity criterion*, which maintain unconditional equivalence between k-means and the diversity.
+
+Unlike $f_1$, $f_2$, $f_3$, and $f_4$, the *dispersion* criterion does not represent between-group similarity, but is a pure measure of within-group heterogeneity.
 It is defined as the minimum dissimilarity across any two objects that are part of the same group [@brusco2019; @fernandez2013]:
@@
 
----
-
-**OLD**
-
-Anticlustering refers to the process of partitioning elements into disjoint groups with the goal of obtaining high between-group similarity and high within-group heterogeneity [@brusco2019; @spath1986; @valev1998; @valev1983].
-Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and separation between clusters [@rokach2005].
-Anticlustering has many applications in research psychology [@steinley2006; @brusco2019; @papenberg2020].
-Examples include splitting tests into parts of equal difficulty [@gierl2017], assigning students to work groups [@baker2002], and assigning stimuli to different, but parallel experimental conditions [@lahl2006].
-Solving anticlustering problems "by hand" is a tedious and time-consuming task, and the quality of manual partitioning is usually subpar.
-Fortunately, anticlustering problems can be formalized as mathematical optimization problems [e.g., @baker2002; @brusco2019; @spath1986; @fernandez2013] and accessible open source software solutions to tackling these problems exist [@papenberg2020].
-
-Anticlustering methods are characterized by (a) an objective function that quantifies between-group similarity and/or within-group heterogeneity, and (b) an algorithm that determines how elements are assigned to groups to maximize the objective function.
-Several anticlustering objective functions use pairwise dissimilarity ratings such as the Euclidean distance as input.
-The most prominent criterion of this kind is the *diversity*, which is the total sum of pairwise dissimilarities between elements within the same group [@brusco2019; @gallego2013].
-By considering pairwise dissimilarities between elements in the same group, the diversity criterion reflects the total degree of within-group heterogeneity.
-High within-group diversity simultaneously ensures high between-group similarity.[^equalsizedgroups] Another important anticlustering criterion based on the information in a dissimilarity matrix is the *dispersion*, which is the minimum dissimilarity between any two elements within the same group [@fernandez2013].
-Maximizing the dispersion increases within-group heterogeneity by ensuring that any two elements in the same group are as dissimilar from each other as possible.
-@brusco2019 convincingly argued that anticlustering applications striving for within-group heterogeneity should incorporate both dispersion and diversity, and they presented a bicriterion algorithm to approximate the Pareto efficient set of solutions according to both criteria.
-
-[^equalsizedgroups]: As will be shown later, the diversity objective only strictly equalizes within-group heterogeneity and between-group similarity when groups are equal-sized.
-
-Oftentimes, partitioning applications in psychology are focused on between-group similarity rather than within-group heterogeneity, even though both goals usually coincide.
-High between-group similarity is for example desirable when designing experimental conditions using different stimulus sets that should be as similar as possible with regard to response-relevant attributes [@lahl2006].
-Such stimulus selection tasks are usually conducted on the basis of attribute data and not on pairwise (dis)similarity ratings.
-That is, individual stimuli are numerically coded on relevant dimensions [@dry2009].
-For stimuli in psycholinguistic experiments, attributes may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982].
-In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977].
-When attribute values are available, anticlustering approaches that do not require a dissimilarity matrix[^dissimilarityfromfeatures] can be used to partition the data.
-K-means is probably the best-known clustering objective that can directly be computed on attribute values.
-In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized, usually using the k-means heuristic [@jain2010; @steinley2006; @brusco2006branch].
-Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np].
-Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful approach when aiming for high between-group similarity [@spath1986; @valev1998; @valev1983].
-Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters.
-However, other distribution characteristics---such as the variance---are not targeted.
-@papenberg2020 showed that k-means anticlustering tends to over-optimize between-group similarity with regard to the mean at the cost of similarity in variance.
-However, when aiming for overall between-group similarity, neglecting all distribution characteristics other than the mean is misguided; similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].
-
-[^dissimilarityfromfeatures]: It is still possible to convert the attribute data into a dissimilarity matrix, e.g.
-by computing all pairwise Euclidean distances, and then optimize a distance-based anticlustering criterion to obtain a partitioning.
-For Likert scale data or binary attributes, the Manhattan distance may be preferred over the Euclidean distance; for mixed-type attributes, it is even possible to compute a combined distance metric [@rokach2005].
-
-To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering.
-K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to several distribution characteristics instead of only the mean.
-Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis.
-After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional criteria such as the diversity and the k-means objective.
-Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].
-
-# Evaluation
-
-Two simulation studies were conducted to evaluate the new k-plus objective for anticlustering.
-Simulation 1 compared a selected number of weighting schemes for the k-plus criterion $\mathit{SSE_{kplusGeneralized}}$.
-The results of this initial test determined the benchmark implementation of k-plus anticlustering that was used as contestant in the follow-up simulation.
-Simulation 2 was a large scale comparison of the k-plus objective and the most important traditional anticlustering objectives striving for between-group similarity: the diversity and the k-means objective.
-
-The simulation studies were implemented using the statistical programming language R [Version 4.2, @R-base].
-The R package `anticlust` [Version 0.6.1, @R-anticlust; @papenberg2020] was used to optimize the anticlustering objectives.
-The R package `faux` [Version 1.1.0, @R-faux] was used to generate the multivariate normal data sets that were processed during Simulation 2.
-The R packages `dplyr` [Version 1.0.7, @R-dplyr], `ggplot2` [Version 3.3.5, @R-ggplot2], `papaja` [Version 0.1.1.9001, @R-papaja], `tidyr` [Version 1.1.3, @R-tidyr], `fossil` [Version 0.4.0, @R-fossil], and `DescTools` [Version 0.99.45, @R-desctools] were used for data analysis and result presentation.
-The simulation study is fully reproducible via code and data that has been made accessible in an accompanying OSF repository [@osf2023].
-
-## Optimization algorithm
-
-In both simulation studies, the same local maximum search was applied to optimize the competing anticlustering objectives [Method LCW, @weitz1998].
-Building on an initial random allocation, the algorithm proceeds by swapping elements between groups in such a way that each swap improves the objective criterion by the largest possible margin.
-That is, it starts with the first element and simulates all exchanges with elements that are currently assigned to a different group.
-After each exchange has been evaluated, the one exchange is realized that improves the objective function the most.
-No exchange is realized if no improvement is possible.
-The exchange process is repeated for each element.
-After that, the procedure restarts at the beginning and is repeated until an iteration through the entire data set no longer yields an improvement.
-To obtain better results, this local maximum search may be initialized multiple times [@spath1986].
-In the simulation studies, five repetitions were used.
-Note that the local maximum search can be performed on the basis of any arbitrary initial grouping.
-In particular, the initial grouping determines the sizes of the anticlusters, because the exchange procedure following the set-up never changes the sizes of the groups.
-While the data sets in the simulation studies were always partitioned into equal-sized groups, it would also be possible to ensure that the group sizes differ by 1 at most if an even split is not possible; even different-sized groups can be requested if required by the application.
-
-## Conditions
-
-For Simulation 2, I generated 10,000 data sets following a normal distribution.
-The data sets were subsequently processed by the competing methods.
-The following properties were determined randomly for each data set: (a) the sample size $N$ varied between 24 and 300 with intermediate steps of 12 so that each data set could be split evenly into $K = 2$, $3$, and $4$ groups; (b) the number of features varied between 2 and 5; (c) the standard deviation of all features was set to 1, 2 or 3 (the mean was always zero); (d) the correlation between all features was $r = 0$, $r = .1$, $r = .2$, $r = .3$, $r = .4$ or $r = .5$.
-All anticlustering methods were applied to each of the 10,000 data sets using $K = 2$, $K = 3$ and $K = 4$ (i.e., each method was applied 30,000 times).
-Groups sizes were always equal.
-
-## Evaluation criteria
-
-## Results
-
-## Example Application II: Small data set ($N = 96$)
-
-The simulation study indicated that k-plus may show decreased performance for smaller group sizes.
-Therefore, a second application illustrates that k-plus anticlustering yields satisfactory results in practice even when group sizes are smaller.
-I make use of a norming data set of 96 word stimuli that was contributed to the `anticlust` package by Marie L.
-Schaper [@schaper2019metacognitive; @schaper2019metamory].
-After loading the `anticlust` package, the data set can be accessed as follows:
-
-```{r, echo = TRUE}
-data(schaper2019)
-```
-
 # Discussion

### 2024-03-11 paper.Rmd
intro

@@
 *K*-plus anticlustering is used to equate not only means between clusters, but higher order moments such as the variance or skewness as well.
-Because the approach can be reduced to an augmentation of the data set, *k*-plus anticlustering leaves the original *k*-means criterion $f_1$ unchanged.
 For example, to equate means and variances, we define a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$ and augment the original matrix with it.
@@
 $$
-f_3(C, \mathbf{D}_{N \times N}) = \sum_{(i < j) \in C_k} d_{ij}
+f_3(C, \mathbf{D}_{N \times N}) = \sum\limits_{k = 1}^{K} \sum_{(i < j) \in C_k} d_{ij}
 $$
@@
 @schulz2022].
+
 If $\mathbf{D}_{N \times N}$ is computed via the squared Euclidean distances, we note an important equivalence between the k-means criterion and the diversity.
-The k-means criterion can also be expressed using the pairwise squared Euclidean distances between data points (Brusco, 2006; Späth, 1980).
+The k-means criterion can be expressed using the pairwise squared Euclidean distances between data points (Brusco, 2006; Späth, 1980).
 Based on $\mathbf{D}_{N \times N}$, the k-means criterion is then computed as
@@
 $$
-f_1(C, \mathbf{X}_{N \times M}) =
-f_4(C, \mathbf{D}_{N \times N}) = \sum\limits_{k = 1}^{K}\left( \frac{1}{n_k} \sum_{(i<j)\in C_k} d_{ij} \right)
+f_4(C, \mathbf{D}_{N \times N}) = \sum\limits_{k = 1}^{K}\left( \frac{1}{n_k} \sum_{(i<j)\in C_k} d_{ij} \right) = f_1(C, \mathbf{X}_{N \times M})
 $$
@@
 350).
-
-An important special case of anticlustering applications is that all groups are equal-sized, i.e., $n_k = \frac{N}{K}$ ($k = 1, \dots K$). Here, we obtain
+For arbitrary measures of dissimilarity (and not just the squared Euclidean distance), we denote $f_4$ as the *average diversity criterion*, which  represents between-group similarity regardless of group sizes.
+However, an important special case of anticlustering applications is that all groups are equal-sized, i.e., $n_k = \frac{N}{K}$ ($k = 1, \dots K$).
+Here, we obtain
 
 $$
-\mathit{SSE} = \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
+f_4(C, \mathbf{D}_{N \times N}) = \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
 $$
 
-In this case, the normalizing factor $\frac{N}{K}$ can be ignored when maximizing $f_1$. Thus, the criterion reduces to the sum of the (squared) distances between elements in the same cluster:
-
-$$
-\mathit{SSE_{Diversity}} = \sum\limits_{k = 1}^{K}\sum_{(i<j)\in C_k} d_{ij}
-$$
-
-This term corresponds to the standard diversity $f_2$, with the distances $d_{ij}$ representing squared Euclidean distances.
-We denote $f_4$ as the *average diversity criterion*, which maintain unconditional equivalence between k-means and the diversity.
+In this case, the normalizing factor $\frac{N}{K}$ can be ignored when maximizing $f_4$.
+Thus, the criterion reduces to the sum of the (squared) distances between elements in the same cluster.
+Thus, for equal-sized groups, $f_4$ is reduced to the standard diversity $f_3$---at least in of the partitionings that are obtained via maximization.
+Moreover, if the $d_{ij}$ are squared Euclidean distances, $f_4$ is equivalent to the k-means objective $f_1$.
+Hence, for equal-sized groups, the diversity anticlustering criterion $f_3$ is equivalent to the k-means criterion $f_1$.
 
@@
 $$
-f_3(C) = \textrm{min}_{(i < j) \in C_k} d_{ij}
+f_5(C) = \textrm{min}_{(i < j) \in C_k} d_{ij}
 $$
@@
 In particular, they argued that anticlustering algorithms should incorporate an objective of between-group similarity---such as the diversity---and pairwise within-group dissimilarity---i.e., the dispersion---at the same time.
+To this end, @brusco2019 presented a bicriterion algorithm that simultaneously maximizes the diversity $f_3$ and the dispersion $f_5$.
+(**say some general stuff about the BILS before explaining it in detail later...**)
+
+We endorse Brusco et al.'s bicriterion approach for anticlustering, and advance it in several ways in the current paper.
+Specifically, we offer several contributions: First, we allow that not only the standard diversity $f_3$ can be optimized alongside the dispersion, but instead we allow that the k-means, k-plus or the average diversity can be used as measures of between-group similarity (i.e., the criteria $f_1$, $f_2$ and $f_4$).
+Second, we allow that the dispersion and $f_1$ -- $f_4$ can be optimized using different dissimilarity matrices.
+This is useful if both criteria should be based on different information, or to induce custom cannot-link constraints, as we will show.
+As out most important contribution, we combine a novel exact algorithm for maximum dispersion with the BILS heuristic to obtain high diversity on top of an optimal dispersion.
+Despite its theoretical computational hardness, our optimal method scales to rather large data sets.
+It generates multiple partitions that we use as initializations of the BILS's local search, thus guaranteeing that the BILS also returns a partition with optimal dispersion.
+In a simulation study, we compared several adaptations of the BILS.
+
+## The BILS
 

### 2024-03-11 paper.Rmd
remove old stuff

@@
 
-In within-subjects experiments, participants are presented with several stimulus sets that are affiliated with the different experimental conditions.
-Due to carry-over effects, showing the same stimuli in different conditions is oftentimes prohibited.
-This is evident in studies on memory, where studying a list of words in one conditions necessarily makes it impossible to repeat their study without being influenced by the earlier study phase; memory cannot be erased after an experimental block has been completed [@lahl2006].
-While the experimental logic dictates that different lists must be used, the lists should be as parallel as possible with regard to variables that affect the response variable.
-@lintz2021 gave two reasons as to why it is important that study participants should be confronted with balanced lists: (a) if lists are not sufficiently parallel, the variance of within-subjects effect sizes due to the experimental manipulation will be inflated, reducing statistical power---if there is no proper statistical control of the stimulus materials, which is often the case [@judd2012]; (b) highly unbalanced sets may introduce secondary effects beyond the influence of the intrinsic stimulus features.
-For example, highly biased lists may induce a strategic behavior shift between experimental blocks, possibly due to suspicion, boredom or frustration.
-These effects invalidate any observed differences between conditions and importantly, this bias cannot be accounted for via statistical techniques.
-
-Traditionally, assigning stimuli to experimental conditions was done manually [e.g., @schaper2019metamory] or randomly [e.g., @kroneisen2018], both of which methods however do not ensure that the stimulus sets will be similar [@papenberg2020; @cutler1981; @vancasteren2007; @armstrong2012].
-Generally, automated approaches are preferred that quantify similarity on the basis of the stimulus features and use an algorithm to assign stimuli in such a way that this similarity is optimized [@brusco2019; @papenberg2020].
-One approach that has recently gained traction
-
-**Using anticlustering to equate stimulus sets**
-
-**BUT DISPERSION IS ALSO IMPORTANT!**
-
-For example, in tasks of recognition memory, participants commonly have to decide if a target stimuli (e.g., a word) has been presented in an earlier phase of the experiment.
-In this case, words that are orthographically similar to a target word should serve as distractors.
-
-The BILS algorithm has been available in the anticlust package since 2021, but we assume that many researchers do not yet know about its potential, also given that the introduction by @brusco2019 was rather technical in nature.
-For now, users seem to rely on the unicriterion methods focusing on between-set similarity only.
-Therefore, we provide readers with an application-focused introduction and extend BILS with parameters that may be useful when designing experiments.
-
 # Discussion

### 2024-03-12 paper.Rmd
absatz
intro

@@
 
-Partitioning data sets via anticlustering has uses in many research fields including psychology [@brusco2019; @schaper2023], test assessment [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network systems [@mohebi2022], and operations research [e.g.
+Partitioning data sets via anticlustering has uses in many research fields including psychology [@brusco2019; @schaper2023], test assembly [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network systems [@mohebi2022], and operations research [e.g.
 @gallego2013; @gliesch2021].
@@
 
-Formally, anticlustering is used to partition $N$ objects into $K$ groups $C = (C_1, \ldots, C_K)$, where $C_k$ is the set of $n_k$ objects in the $k^{th}$ group, and $K$ and $n_k$ are predetermined by the application.
-In many anticlustering applications, equal-sized groups are required, i.e., $n_k = \frac{N}{K}$ [@schulz2022; @gallego2013].
-The partitioning is characterized by assignment to exhaustive and mutually exclusive subgroups:
+Formally, anticlustering is used to partition $N$ objects into $K$ exhaustive and mutually exclusive subgroups $C = (C_1, \ldots, C_K)$. The partitioning restrictions are formalized as follows:
 
 \begin{align}
-\bigcup\limits_{k = 1}^{K} C_k = X \label{formalization:1} \\
-C_j \cap C_k = \emptyset, \; \forall j, k \in \{1, \ldots, K\}, \; j \ne k\label{formalization:2} \\
-\vert C_k \vert = n_k, \; \forall k \in \{1, \ldots, K\} \label{formalization:3}
+\vert C_k \vert = n_k, \; (k = 1, \ldots, K) \label{formalization:1} \\
+\sum_{k = 1}^{K} n_k = N \label{formalization:2} \\
+C_j \cap C_k = \emptyset (j, k = 1, \ldots, K), \; j \ne k\label{formalization:3} \\
 \end{align}
 
-We assume that a data matrix $\mathbf{Z}$ is used as input for computing the anticlustering criterion.
+$C_k$ is the set of $n_k$ objects in the $k^{th}$ group, and $K$ and $n_k$ are predetermined by the application.
+We will refer to the subsets $C_k$ ($k = 1, \ldots, K$), as anticlusters, sets, or groups interchangeably, and to the collection $C$ as the anticlustering partitioning.
+In many anticlustering applications, equal-sized groups are required, i.e., $n_k = \frac{N}{K}$ ($k = 1, \ldots, K$) [@schulz2022; @gallego2013].
+
+We assume that a data matrix $\mathbf{Z}$ is used as input for computing the anticlustering criterion on the basis of a partitioning $C$.
 $\mathbf{Z}$ is either a feature matrix $\mathbf{X}_{N \times M} = \{x_{ij}\}_{N \times M}$ where the $M$ columns contain numeric measurements on the $N$ objects, or a symmetric matrix $\mathbf{D}_{N \times N} = \{d_{ij}\}_{N \times N}$ of pairwise dissimilarity measurements.
@@
 
-The $k^{th}$ cluster center $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ contains the mean values of the $M$ variables computed across all observations belonging to the  $k^{th}$ group:
-
 $$
@@
 
-Maximizing $f_1$ ensures that the cluster centers, i.e., the mean values of the $M$ measurements, are as close to each other as possible [@spath1986].
+The $k^{th}$ cluster center $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ contains the mean values of the $M$ variables computed across all observations belonging to the $k^{th}$ group.
+Maximizing $f_1$ ensures that the cluster centers are as close to each other as possible [@spath1986].
+
 The *k*-plus criterion is an extension of the $k$-means criterion for anticlustering that can be implemented by adding fictitious to the data matrix $\mathbf{X}_{N \times M}$ [@papenberg2024].
@@
 
-If $\mathbf{D}_{N \times N}$ is computed via the squared Euclidean distances, we note an important equivalence between the k-means criterion and the diversity.
+If $\mathbf{D}_{N \times N}$ is computed via the squared Euclidean distances, we note an important equivalence between the k-means criterion $f_1$ and the diversity $f_3$.
 The k-means criterion can be expressed using the pairwise squared Euclidean distances between data points (Brusco, 2006; Späth, 1980).
-Based on $\mathbf{D}_{N \times N}$, the k-means criterion is then computed as
+If we let $\mathbf{D'}_{N \times N}$ = \{$d'_{ij}$\} representing the squared Euclidean distance, i.e., $d'_{ij} = \sum\limits_{m = 1}^{M}(x_{im} - x_{jm})^2$, the k-means criterion is given as
 
 $$
-f_4(C, \mathbf{D}_{N \times N}) = \sum\limits_{k = 1}^{K}\left( \frac{1}{n_k} \sum_{(i<j)\in C_k} d_{ij} \right) = f_1(C, \mathbf{X}_{N \times M})
+f_4(C, \mathbf{D'}_{N \times N}) = \sum\limits_{k = 1}^{K}\left( \frac{1}{n_k} \sum_{(i<j)\in C_k} d'_{ij} \right) = f_1(C, \mathbf{X}_{N \times M}).
 $$
@@
 350).
-For arbitrary measures of dissimilarity (and not just the squared Euclidean distance), we denote $f_4$ as the *average diversity criterion*, which  represents between-group similarity regardless of group sizes.
-However, an important special case of anticlustering applications is that all groups are equal-sized, i.e., $n_k = \frac{N}{K}$ ($k = 1, \dots K$).
-Here, we obtain
+For arbitrary measures of dissimilarity (and not just the squared Euclidean distance), we denote $f_4$ as the *average diversity criterion*.
+An important special case of anticlustering applications is that all groups are equal-sized, i.e., $n_k = \frac{N}{K}$ ($k = 1, \dots K$).
+In this context, $f_4$ can be simplified to
 
 $$
-f_4(C, \mathbf{D}_{N \times N}) = \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
+f_4^*(C, \mathbf{D'}_{N \times N}) = \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
 $$
 
-In this case, the normalizing factor $\frac{N}{K}$ can be ignored when maximizing $f_4$.
-Thus, the criterion reduces to the sum of the (squared) distances between elements in the same cluster.
-Thus, for equal-sized groups, $f_4$ is reduced to the standard diversity $f_3$---at least in of the partitionings that are obtained via maximization.
-Moreover, if the $d_{ij}$ are squared Euclidean distances, $f_4$ is equivalent to the k-means objective $f_1$.
-Hence, for equal-sized groups, the diversity anticlustering criterion $f_3$ is equivalent to the k-means criterion $f_1$.
+In the context of maximization, the normalizing factor $\frac{N}{K}$ can be ignored because it does not depend on the partitioning $C$.
+Thus, the criterion reduces to the sum of the (squared) distances between elements in the same cluster, i.e., the standard diversity $f_3$.
+Therefore, for equal-sized groups, the diversity anticlustering criterion $f_3$ is equivalent to the k-means criterion $f_1$, if the diversity is computed via the squared Euclidean distance.
+We will use this insight later when we apply algorithms for maximum diversity to $k$-means and $k$-plus anticlustering.
 

### 2024-03-12 paper.Rmd
intro

@@
 
-Partitioning data sets via anticlustering has uses in many research fields including psychology [@brusco2019; @schaper2023], test assembly [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network systems [@mohebi2022], and operations research [e.g.
+Partitioning data sets via anticlustering has uses in many research fields including psychology [@brusco2019; @schaper2023; @papenberg2020], test assembly [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network systems [@mohebi2022], and operations research [e.g.
 @gallego2013; @gliesch2021].
@@
 $C_k$ is the set of $n_k$ objects in the $k^{th}$ group, and $K$ and $n_k$ are predetermined by the application.
-We will refer to the subsets $C_k$ ($k = 1, \ldots, K$), as anticlusters, sets, or groups interchangeably, and to the collection $C$ as the anticlustering partitioning.
-In many anticlustering applications, equal-sized groups are required, i.e., $n_k = \frac{N}{K}$ ($k = 1, \ldots, K$) [@schulz2022; @gallego2013].
+We will refer to the subsets $C_k$ ($k = 1, \ldots, K$), as anticlusters, clusters, or groups interchangeably, and to the collection $C$ as the anticlustering partitioning.
 
@@
 
-The *k*-plus criterion is an extension of the $k$-means criterion for anticlustering that can be implemented by adding fictitious to the data matrix $\mathbf{X}_{N \times M}$ [@papenberg2024].
+The *k*-plus criterion is an extension of the $k$-means criterion for anticlustering that can be implemented by adding fictitious variables to the data matrix $\mathbf{X}_{N \times M}$ [@papenberg2024].
 *K*-plus anticlustering is used to equate not only means between clusters, but higher order moments such as the variance or skewness as well.
-For example, to equate means and variances, we define a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$ and augment the original matrix with it.
+For example, to equate means and variances, we define a new data matrix $\mathbf{Y}_{N \times M} = \{y_{ij}\}_{N \times M}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$, where $\overline{x}_j$ is the overall mean of the $j^{th}$ attribute, and augment the original matrix with it.
 Then, the *k*-plus criterion is computed as
@@
 Being a sum instead of an average, however, it no longer adequately captures between-group similarity if group sizes are not equal.
-The diversity is oftentimes obtained by first converting the data matrix $\mathbf{X}_{N \times M}$ into a matrix of pairwise Euclidean distances.
-However, most anticlustering algorithms that use a dissimilarity matrix as input are agnostic as to how it was was generated [@brusco2019; cf.
-@schulz2022].
 
-If $\mathbf{D}_{N \times N}$ is computed via the squared Euclidean distances, we note an important equivalence between the k-means criterion $f_1$ and the diversity $f_3$.
-The k-means criterion can be expressed using the pairwise squared Euclidean distances between data points (Brusco, 2006; Späth, 1980).
+The diversity is oftentimes obtained by first converting the data matrix $\mathbf{X}_{N \times M}$ into a matrix of pairwise Euclidean distances [@brusco2019; @papenberg2020; @papenberg2024; @gallego2013].
+If $\mathbf{D}_{N \times N}$ is obtained by using the squared pairwise Euclidean distances instead, we note an important equivalence between the k-means criterion $f_1$ and the diversity $f_3$.
+Traditionally, the k-means criterion is denoted via a sum of squared distances between data points and cluster centroids.
+However, it can also be expressed using the pairwise squared Euclidean distances between data points (Brusco, 2006; Späth, 1980).
 If we let $\mathbf{D'}_{N \times N}$ = \{$d'_{ij}$\} representing the squared Euclidean distance, i.e., $d'_{ij} = \sum\limits_{m = 1}^{M}(x_{im} - x_{jm})^2$, the k-means criterion is given as
@@
 $$
-f_4(C, \mathbf{D'}_{N \times N}) = \sum\limits_{k = 1}^{K}\left( \frac{1}{n_k} \sum_{(i<j)\in C_k} d'_{ij} \right) = f_1(C, \mathbf{X}_{N \times M}).
+f_4(C, \mathbf{D'}_{N \times N}) = 
+  \sum\limits_{k = 1}^{K}\left( \frac{1}{n_k} \sum_{(i<j)\in C_k} d'_{ij} \right) = 
+  f_1(C, \mathbf{X}_{N \times M}).
 $$
@@
 350).
-For arbitrary measures of dissimilarity (and not just the squared Euclidean distance), we denote $f_4$ as the *average diversity criterion*.
 An important special case of anticlustering applications is that all groups are equal-sized, i.e., $n_k = \frac{N}{K}$ ($k = 1, \dots K$).
-In this context, $f_4$ can be simplified to
+In this context, $f_4$ can be simplified to $f^*$ via
 
 $$
-f_4^*(C, \mathbf{D'}_{N \times N}) = \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
+f_4^*(C, \mathbf{D'}_{N \times N}) = 
+  \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij} \\
+  \propto \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij} = f_3(C, \mathbf{D'}_{N \times N})
 $$
@@
 In the context of maximization, the normalizing factor $\frac{N}{K}$ can be ignored because it does not depend on the partitioning $C$.
-Thus, the criterion reduces to the sum of the (squared) distances between elements in the same cluster, i.e., the standard diversity $f_3$.
-Therefore, for equal-sized groups, the diversity anticlustering criterion $f_3$ is equivalent to the k-means criterion $f_1$, if the diversity is computed via the squared Euclidean distance.
-We will use this insight later when we apply algorithms for maximum diversity to $k$-means and $k$-plus anticlustering.
+Thus, the criterion $f_4^*$ reduces to the sum of the (squared) distances between elements in the same cluster, i.e., the standard diversity $f_3$.
+Therefore, for equal-sized groups, the diversity anticlustering criterion $f_3$ is equivalent to the k-means criterion $f_1$---if the diversity is computed via the squared Euclidean distance.
+We will use this insight later when we discuss applying algorithms for maximum diversity to the $k$-means and $k$-plus objectives.
+For arbitrary measures of dissimilarity (and not just the squared Euclidean distance), we denote $f_4$ as the *average diversity criterion*.
 

### 2024-03-14 paper.Rmd
minor

@@
 *K*-plus anticlustering is used to equate not only means between clusters, but higher order moments such as the variance or skewness as well.
-For example, to equate means and variances, we define a new data matrix $\mathbf{Y}_{N \times M} = \{y_{ij}\}_{N \times M}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$, where $\overline{x}_j$ is the overall mean of the $j^{th}$ attribute, and augment the original matrix with it.
+For example, to equate means and variances, we define a new data matrix $\mathbf{Y}_{N \times M} = \{y_{ij}\}_{N \times M}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$, where $\overline{x}_j$ is the overall mean of the $j^{th}$ attribute.
 Then, the *k*-plus criterion is computed as
@@
 However, it can also be expressed using the pairwise squared Euclidean distances between data points (Brusco, 2006; Späth, 1980).
-If we let $\mathbf{D'}_{N \times N}$ = \{$d'_{ij}$\} representing the squared Euclidean distance, i.e., $d'_{ij} = \sum\limits_{m = 1}^{M}(x_{im} - x_{jm})^2$, the k-means criterion is given as
+If we let $\mathbf{D'}_{N \times N}$ = \{$d'_{ij}$\} represent the squared Euclidean distance, i.e., $d'_{ij} = \sum\limits_{m = 1}^{M}(x_{im} - x_{jm})^2$, the k-means criterion can be expressed as as
 
@@
 An important special case of anticlustering applications is that all groups are equal-sized, i.e., $n_k = \frac{N}{K}$ ($k = 1, \dots K$).
-In this context, $f_4$ can be simplified to $f^*$ via
+In this context, $f_4$ can be simplified to $f_4^*$ via
 
@@
 f_4^*(C, \mathbf{D'}_{N \times N}) = 
-  \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij} \\
+  \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
+$$
+
+In the context of maximization, the normalizing factor $\frac{N}{K}$ can be ignored because it does not depend on the partitioning $C$. Hence,
+
+$$
+f_4^*(C, \mathbf{D'}_{N \times N})
   \propto \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij} = f_3(C, \mathbf{D'}_{N \times N})
@@
 
-In the context of maximization, the normalizing factor $\frac{N}{K}$ can be ignored because it does not depend on the partitioning $C$.
-Thus, the criterion $f_4^*$ reduces to the sum of the (squared) distances between elements in the same cluster, i.e., the standard diversity $f_3$.
+For anticlustering, the criterion $f_4^*$ (i.e., the k-means criterion for equal-sized groups) reduces to the sum of the (squared) distances between elements in the same cluster, i.e., the standard diversity $f_3$.
 Therefore, for equal-sized groups, the diversity anticlustering criterion $f_3$ is equivalent to the k-means criterion $f_1$---if the diversity is computed via the squared Euclidean distance.
@@
 
-@brusco2019 outlined the importance of simultaneously considering multiple criteria when tackling anticlustering problems.
+@brusco2019 outlined the importance of simultaneously considering multiple criteria when addressing anticlustering problems.
 In particular, they argued that anticlustering algorithms should incorporate an objective of between-group similarity---such as the diversity---and pairwise within-group dissimilarity---i.e., the dispersion---at the same time.
 To this end, @brusco2019 presented a bicriterion algorithm that simultaneously maximizes the diversity $f_3$ and the dispersion $f_5$.
-(**say some general stuff about the BILS before explaining it in detail later...**)
-
+While in the next section, we will discuss their approach in more detail.
 We endorse Brusco et al.'s bicriterion approach for anticlustering, and advance it in several ways in the current paper.

### 2024-03-14 paper.Rmd
outline

@@
 
-Formally, anticlustering is used to partition $N$ objects into $K$ exhaustive and mutually exclusive subgroups $C = (C_1, \ldots, C_K)$. The partitioning restrictions are formalized as follows:
+Formally, anticlustering is used to partition $N$ objects into $K$ exhaustive and mutually exclusive subsets $C = (C_1, \ldots, C_K)$. The partitioning restrictions are formalized as follows:
 
@@
 To this end, @brusco2019 presented a bicriterion algorithm that simultaneously maximizes the diversity $f_3$ and the dispersion $f_5$.
-While in the next section, we will discuss their approach in more detail.
 We endorse Brusco et al.'s bicriterion approach for anticlustering, and advance it in several ways in the current paper.
-Specifically, we offer several contributions: First, we allow that not only the standard diversity $f_3$ can be optimized alongside the dispersion, but instead we allow that the k-means, k-plus or the average diversity can be used as measures of between-group similarity (i.e., the criteria $f_1$, $f_2$ and $f_4$).
-Second, we allow that the dispersion and $f_1$ -- $f_4$ can be optimized using different dissimilarity matrices.
-This is useful if both criteria should be based on different information, or to induce custom cannot-link constraints, as we will show.
-As out most important contribution, we combine a novel exact algorithm for maximum dispersion with the BILS heuristic to obtain high diversity on top of an optimal dispersion.
-Despite its theoretical computational hardness, our optimal method scales to rather large data sets.
-It generates multiple partitions that we use as initializations of the BILS's local search, thus guaranteeing that the BILS also returns a partition with optimal dispersion.
-In a simulation study, we compared several adaptations of the BILS.
+First, we allow that not only the diversity $f_3$ can be optimized alongside the dispersion, but instead we allow that the k-means, k-plus or the average diversity can be used as measures of between-group similarity (i.e., the criteria $f_1$, $f_2$ and $f_4$).
+Second, we allow that the dispersion can be optimized on the basis of a different dissimilarity matrix than the diversity (or a different objective of between-group similarity).
+This is useful if the criteria should be based on different information, or to induce custom cannot-link constraints, as we will show.
+As our most important contribution, we combine a novel exact algorithm for maximum dispersion with the BILS heuristic to obtain high diversity on top of an optimal dispersion.
+Our paper is organized as follows.
+In the next section, we review the bicriterion method by @brusco2019 in detail.
+After that, we provide the technical background for our extensions.
+We then provide several evaluations of our contributions: (a) we probe the practical feasibility of our optimal algorithm for maximum dispersion, (b) we compare different adaptations of the BILS with regard to their ability to providing high diversity on top of an optimal dispersion, and (c) we provide practical examples on using our new methods.
 

### 2024-03-14 paper.Rmd
shown here
introduction

@@
 
-Partitioning data sets via anticlustering has uses in many research fields including psychology [@brusco2019; @schaper2023; @papenberg2020], test assembly [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network systems [@mohebi2022], and operations research [e.g.
+Partitioning data sets via anticlustering has uses in many research fields including psychology [@brusco2019; @schaper2023], test assembly [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network systems [@mohebi2022], and operations research [e.g.
 @gallego2013; @gliesch2021].
-Anticlustering seeks for similarity between groups and heterogeneity within groups by maximizing a clustering objective criterion.
-Thus, anticlustering is the logical and mathematical opposite of classical clustering, which partitions data sets via minimization of the same criteria [@steinley2006; @brusco2019; @spath1986].
+In psychological research, an important application is the assignment of stimuli to different experimental sets that are presented in within-subjects designs: The different sets should be as similar as possible with regard to variables that affect the responses, and presenting the same set repeatedly is usually prohibited due to carry-over effects.
+Anticlustering is a powerful tool for automatically solving such tasks, which are time-consuming and usually inadequately handled manually [@lahl2006; @lintz2021; @cutler1981].
+The first introduction to anticlustering in a Psychological journal was provided by @brusco2019.
+@papenberg2020 presented the free and open source software package `anticlust` that implements many algorithms for anticlustering, thus providing researchers with a tool for solving a problem that frequently occurs in psychological research, but previously lacked an appropriate software solution.
 
@@
 \sum_{k = 1}^{K} n_k = N \label{formalization:2} \\
-C_j \cap C_k = \emptyset (j, k = 1, \ldots, K), \; j \ne k\label{formalization:3} \\
+C_j \cap C_k = \emptyset (j, k = 1, \ldots, K), \; j \ne k\label{formalization:3}
 \end{align}
@@
 
+Anticlustering seeks for similarity between groups and heterogeneity within groups by maximizing a clustering objective criterion.
+Thus, anticlustering is the logical and mathematical opposite of classical clustering, which partitions data sets via minimization of the same criteria [@steinley2006; @brusco2019; @spath1986].
 We assume that a data matrix $\mathbf{Z}$ is used as input for computing the anticlustering criterion on the basis of a partitioning $C$.
@@
 The partitioning is obtained by maximizing a clustering criterion $f_i(C, \mathbf{Z})$.
-For example, *k*-means anticlustering aims to maximize between-group similarity by maximizing the error sum of squares, i.e., the sum the squared distances between each object and its cluster center [@spath1986; @papenberg2024]:
+For example, *k*-means anticlustering aims to maximize between-group similarity by maximizing the error sum of squares, i.e., the sum the squared distances between each object and its cluster center [@spath1986; @steinley2006]:
 
@@
 
-Maximizing the diversity simultaneously minimizes the sum of distances between objects not in the same group.
+Maximizing the diversity $f_3$ simultaneously minimizes the sum of distances between objects not in the same group.
 Therefore, the diversity represents within-group heterogeneity as well as between-group similarity [@feo1990].
-Being a sum instead of an average, however, it no longer adequately captures between-group similarity if group sizes are not equal.
+Being a sum instead of an average, however, it no longer adequately captures between-group similarity if group sizes are not equal [@papenberg2024].
 
-The diversity is oftentimes obtained by first converting the data matrix $\mathbf{X}_{N \times M}$ into a matrix of pairwise Euclidean distances [@brusco2019; @papenberg2020; @papenberg2024; @gallego2013].
+The diversity is oftentimes computed by first converting a feature matrix $\mathbf{X}_{N \times M}$ into a matrix of pairwise Euclidean distances [@brusco2019; @papenberg2020; @papenberg2024; @gallego2013].
 If $\mathbf{D}_{N \times N}$ is obtained by using the squared pairwise Euclidean distances instead, we note an important equivalence between the k-means criterion $f_1$ and the diversity $f_3$.
@@
 However, it can also be expressed using the pairwise squared Euclidean distances between data points (Brusco, 2006; Späth, 1980).
-If we let $\mathbf{D'}_{N \times N}$ = \{$d'_{ij}$\} represent the squared Euclidean distance, i.e., $d'_{ij} = \sum\limits_{m = 1}^{M}(x_{im} - x_{jm})^2$, the k-means criterion can be expressed as as
+If we let $\mathbf{D'}_{N \times N}$ = \{$d'_{ij}$\} represent the squared Euclidean distance, i.e., $d'_{ij} = \sum\limits_{m = 1}^{M}(x_{im} - x_{jm})^2$, the k-means criterion can be expressed as $f_4$:
 
@@
 f_4^*(C, \mathbf{D'}_{N \times N}) = 
-  \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
+  \frac{K}{N} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
 $$
 
-In the context of maximization, the normalizing factor $\frac{N}{K}$ can be ignored because it does not depend on the partitioning $C$. Hence,
+In the context of maximization, the factor $\frac{K}{N}$ can be ignored because it does not depend on the partitioning $C$.
 
@@
 
-For anticlustering, the criterion $f_4^*$ (i.e., the k-means criterion for equal-sized groups) reduces to the sum of the (squared) distances between elements in the same cluster, i.e., the standard diversity $f_3$.
-Therefore, for equal-sized groups, the diversity anticlustering criterion $f_3$ is equivalent to the k-means criterion $f_1$---if the diversity is computed via the squared Euclidean distance.
+Therefore, the criterion $f_4^*$ (i.e., the k-means criterion for equal-sized groups) reduces to the standard diversity $f_3$---if the diversity is computed via the squared Euclidean distance.
 We will use this insight later when we discuss applying algorithms for maximum diversity to the $k$-means and $k$-plus objectives.
-For arbitrary measures of dissimilarity (and not just the squared Euclidean distance), we denote $f_4$ as the *average diversity criterion*.
+For arbitrary measures of dissimilarity (and not just the squared Euclidean distance), we denote $f_4$ as the *average diversity* criterion.
 
 Unlike $f_1$, $f_2$, $f_3$, and $f_4$, the *dispersion* criterion does not represent between-group similarity, but is a pure measure of within-group heterogeneity.
-It is defined as the minimum dissimilarity across any two objects that are part of the same group [@brusco2019; @fernandez2013]:
+It is defined as the minimum dissimilarity across any two objects that are part of the same group [@fernandez2013]:
 
@@
 
-The dispersion is a measure of the worst-case pairwise dissimilarity. Maximizing the dispersion ensures that any two objects in the same group are as dissimilar from each other as possible.
+The dispersion is a measure of the worst-case pairwise dissimilarity [@brusco2019].
+Maximizing the dispersion ensures that any two objects in the same group are as dissimilar from each other as possible, which is desireable when striving for high within-group heterogeneity.
 
 @brusco2019 outlined the importance of simultaneously considering multiple criteria when addressing anticlustering problems.
-In particular, they argued that anticlustering algorithms should incorporate an objective of between-group similarity---such as the diversity---and pairwise within-group dissimilarity---i.e., the dispersion---at the same time.
-To this end, @brusco2019 presented a bicriterion algorithm that simultaneously maximizes the diversity $f_3$ and the dispersion $f_5$.
-We endorse Brusco et al.'s bicriterion approach for anticlustering, and advance it in several ways in the current paper.
-First, we allow that not only the diversity $f_3$ can be optimized alongside the dispersion, but instead we allow that the k-means, k-plus or the average diversity can be used as measures of between-group similarity (i.e., the criteria $f_1$, $f_2$ and $f_4$).
+In particular, they argued that anticlustering algorithms should incorporate an objective of between-group similarity such as the diversity, and pairwise within-group dissimilarity, i.e., the dispersion.
+To this end, @brusco2019 presented a bicriterion algorithm (*BILS*) that simultaneously maximizes the diversity $f_3$ and the dispersion $f_5$ by approximating the Pareto efficient set of partitions according to both criteria.
+By investigating the Pareto set for many problem instances, they found that a considerable improvement in one of these criteria is often possible with little or even no sacrifice in the other.
+We endorse Brusco et al.'s bicriterion approach for anticlustering and extend it in several ways.
+First, we allow that not only the diversity $f_3$ can be optimized alongside the dispersion, but we also allow that the k-means, k-plus or the average diversity can be employed as measures of between-group similarity during the bicriterion optimization (i.e., the criteria $f_1$, $f_2$ and $f_4$).
 Second, we allow that the dispersion can be optimized on the basis of a different dissimilarity matrix than the diversity (or a different objective of between-group similarity).
-This is useful if the criteria should be based on different information, or to induce custom cannot-link constraints, as we will show.
-As our most important contribution, we combine a novel exact algorithm for maximum dispersion with the BILS heuristic to obtain high diversity on top of an optimal dispersion.
+This is useful if the objectives should be based on different information, or, as we discuss below, to induce custom pairwise cannot-link constraints.
+As our most important contribution, we present a novel exact algorithm for maximum dispersion and investigate different algorithmic techniques for providing high diversity on top of it.
+All of the methods newly developed here as well as the original BILS algorithm are available via the free and open source `R` package `anticlust` [@papenberg2020; @R-anticlust].
+
 Our paper is organized as follows.
-In the next section, we review the bicriterion method by @brusco2019 in detail.
-After that, we provide the technical background for our extensions.
-We then provide several evaluations of our contributions: (a) we probe the practical feasibility of our optimal algorithm for maximum dispersion, (b) we compare different adaptations of the BILS with regard to their ability to providing high diversity on top of an optimal dispersion, and (c) we provide practical examples on using our new methods.
+In the next section, we discuss algorithms for maximizing anticlustering criteria.
+In particular, we review the bicriterion algorithm by @brusco2019 in detail.
+After that, we provide the technical background for our own extensions of the bicriterion approach.
+We continue by evaluating these contributions: (a) we probe the practical feasibility of our optimal algorithm for maximum dispersion for varying $N$ and $K$; (b) we compare different algorithmic approaches in their ability to provide high diversity on top of an optimal dispersion; (c) we apply our new methods to real norming data and present functioning `R` code to reproduce the examples [@R-base].
 

### 2024-03-14 paper.Rmd
abtract

@@
 Here, we advance the bicriterion approach in several ways.
-As our main contribution, we combine a novel exact algorithm for maximum dispersion with the BILS heuristic to obtain high diversity on top of an optimal dispersion.
+As our most important contribution, we present a novel exact algorithm for maximum dispersion and investigate different algorithmic techniques for providing high diversity on top of it.
 Despite its theoretical computational hardness, our optimal method scales to rather large data sets.
-It generates multiple partitions that we use as initializations of the BILS's local search, thus guaranteeing that the BILS also returns a partition with optimal dispersion.
-In a simulation study, we compared several adaptations of the BILS.
-To obtain high diversity, initializing the BILS with multiple optimal partitions is preferable to using only one.
-Moreover, in data sets where maintaining optimal dispersion severely restricts the search for feasible partitions, the iterated local search phase of the BILS is crucial to obtain high diversity.
-While the original BILS oftentimes finds optimal dispersion values, our simulation illustrates conditions when it does not.
-The optimal dispersion algorithm and the BILS including our extensions are available via the free and open source R package `anticlust`.
+It generates partitions that we use as initializations of the BILS's local search, thus guaranteeing that the BILS also returns a partition with optimal dispersion.
+In a simulation study, we find that initializing the BILS with multiple optimal partitions is preferable to using only one.
+While a constrained single criterion local search performs comparably to the BILS, the iterated local search phase of the BILS is crucial to obtain diversity when the constraint of optimal dispersion severely restricts the search for feasible partitions.
+The original BILS oftentimes finds optimal dispersion values, but our simulation illustrates conditions when it does not.
+All of the methods developed here as well as the original BILS algorithm are available via the free and open source `R` package `anticlust`.
 Practical examples are included to illustrate their application.
